### 2015-10-04 writeup/draft.tex
ready

\documentclass[twoside]{article}

% ------
% Fonts and typesetting settings
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\linespread{1.05} % Palatino needs more space between lines
\usepackage{microtype}

% ------
% Page layout
\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
\usepackage[font=it]{caption}
\usepackage{paralist}
\usepackage{multicol, hyperref, graphicx}

% ------
% Lettrines
\usepackage{lettrine}
\usepackage{color}

% ------
% Abstract
\usepackage{abstract}
	\renewcommand{\abstractnamefont}{\normalfont\bfseries}
	\renewcommand{\abstracttextfont}{\normalfont\small\itshape}

% ------
% Titling (section/subsection)
\usepackage{titlesec}
\renewcommand\thesection{\Roman{section}}
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{}

% ------
% Header/footer
\usepackage{fancyhdr}
	\pagestyle{fancy}
	\fancyhead{}
	\fancyfoot{}
	\fancyhead[C]{ $\bullet$ Draft $\bullet$}
	\fancyfoot[RO,LE]{\thepage}

% My Shortcuts

%useful shortcuts
\def\R{\ensuremath{\mathbb{R}}} %\ensuremath adds math mode, if forgotten
\def\Q{\ensuremath{\mathbb{Q}}}
\def\N{\ensuremath{\mathbb{N}}}
\def\Z{\ensuremath{\mathbb{Z}}}
\def\C{\ensuremath{\mathbb{C}}}

%shorcuts with arguments
\newcommand{\abs}[1]{\left\vert#1\right\vert} %nice absolute values
\newcommand{\bt}[1]{\textbf{#1}} %bold
\newcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
\newcommand{\cb}[1]{\centerline{\fbox{#1}}} %centered box
\newcommand{\bp}[1]{\fbox{\parbox{0.8\textwidth}{#1}}} %box paragraph
\newcommand{\norm}[1]{\left\lVert#1\right\rVert} %vector norm
\newcommand{\notimplies}{% does not imply
  \mathrel{{\ooalign{\hidewidth$\not\phantom{=}$\hidewidth\cr$\implies$}}}}
\renewcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations

%colors
\definecolor{javagreen}{rgb}{0.25,0.5,0.35} %dark green color
\definecolor{lightblue}{rgb}{0.149,0.545,0.824} %solarized blue
\definecolor{sred}{rgb}{0.863, 0.196, 0.184} %solarized red

\newcommand{\blue}[1]{{\leavevmode\color{lightblue}{#1}}} %solarized blue 
\newcommand{\green}[1]{{\leavevmode\color{javagreen}{#1}}} %command for green
\newcommand{\red}[1]{{\leavevmode\color{sred}{#1}}} %solarized red
\newcommand{\gray}[1]{{\leavevmode\color[gray]{0.5}{#1}}} %gray text

%environment
\newcommand{\tab}{\phantom{ssss}}
%----------------

% ------
% Clickable URLs (optional)
\usepackage{hyperref}

% ------
% Maketitle metadata
\title{\vspace{-5mm}%
	\fontsize{24pt}{12pt}\selectfont
	\textbf{What's So Special About Philosophy?} 
	}	
\author{%
\fontsize{14pt}{14pt}\selectfont
	Unraveling Wikipedia's First Link Network \vspace{-2mm}\\
	}
\date{}

%figures
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\maketitle
\thispagestyle{fancy}
%========================ABSTRACT====================================
\begin{abstract}
\fontsize{12pt}{12pt}
\selectfont
\noindent Apples, oranges, and the most obscure Dylan song too---is any article a few clicks from Philosophy? The surprising answer is yes, $95\%$ of the time (according to a popular blog post).

Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. Yet, we have never closely examined the curious web tying one entry to another. By following the first link in an article, we connect entries to form a directed network within Wikipedia: Wikpedia's First Link Network. Our aim is to study Wikipedia's First Link Network for insight into the topics, ideas, people, objects, and events we link. ((improve this sentence))

We first map the First Link Network by algorithmically parsing the first link. Then, apply a river network model to identify the basins of attraction, loop structure, and characterize the many paths from one page's first link to the next. ((wording is unclear: as if there are many paths from one page))

\red{rewrite abstract}

\end{abstract}

%==================================BASE====================================================

\begin{multicols}{2}

\lettrine[nindent=0em,lines=2]{A}t no point in history has a larger or more meticulously indexed collection of human knowledge existed \cite{1}. In amassing such an awe-inspiring collection, we naturally linked topics, inventions, people, objects, places, and events 
\footnote{while anyone with internet access is free to contribute "neutral, cited information," it's worth noting the population of contributors is not necessarily representative of the overall english speaking population. Notably, the contributors are overwhelmingly male ((cite, expand)).} across space and time. The web we weaved, and continue to weave, is a wealth of information about those notable inventions, places, figures, and ideas, but also about us: the ones weaving the web. Here we explore this web as a directed network formed by the first link connecting one article to another.

We build Wikipedia's First link network by following the first link, not in parenthesis, inside the main body of each article of the English version of Wikipedia. For the directed network to meaningfully reflect how we associate one article to another, we exclude links in parenthesis containing pronunciation keys or disambiguations. We also exclude any links in the side-bar elements, as well as any links to external pages, files, or other WikiMedia projects outstide of Wikipedia such as Wiktionary ((lookup others))
\footnote{
These constraints are the same which led to the popular blog post/s claiming nearly every article is within fewer than 20 first links away from philosophy. 
}
thus, forming a true network of Wikipedia.

\includegraphics[scale=0.18]{graphics/wiki_train.png}
\ \\

\section{Making a Map: Finding the First Links}
To map Wikipedia's First Link Network, we use the freely-available xml dump of all 4.7 million English articles \cite{2}. We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time. Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new article per day on average \cite{3}, our aim is to characterize the dynamics of the First Link Network, not to record a particular link between one page to another.

The xml dump provided demarkates article elements in in Media Wiki markup, a markup language with special syntax for links. 
Media Wiki markup also includes templates for audio files, images, side-bar information.
While a human can accurately idenfity the first link, to map the entire First Link Network of 4.7 million pages, we programmatically untangled the body text from side-bar, header box, and bad link elements.

While some libraries exist for Media Wiki Markup, we opted to develop our own algorithm for parsing the first link in the xml version of each article. 
\footnote{Approaches using existing libraries led to several bugs including trouble with nested links, nested parenthesis, unclosed tags, escape characters as well as compatibility with other libraries used to parse the xml.}
Our parsing algorithm aimed to: 1. squash the initial bugs 2. eliminate the need for several passes. To process an article with one pass, we developed a hierachical system of flags:
\begin{itemize}
    \item Flag 1: inside template?
    \begin{itemize}
        \item Flag 2: inside <ref>, <div>?
        \begin{itemize}
            \item Flag 3: inside ()?
        \end{itemize}
    \end{itemize}
\end{itemize}

The algorithm loops in three characters chunks, shifting by one character steps through the article markup. Proceeding in three character chuncks eliminates any nesting issues. If any flag triggers are detected the Flag is raised. Once a flag is raised, we stop processing and proceed to the next character. A first link is correctly identified only if Flags 1, 2, and 3 are all off. Then the entire link is retrieved. We then confirm whether the link is legitimate, meaning it does not link to an image, audio file, external page or other Wikimedia project via keyword triggers provided by Wiki Media \cite{4}. The first link of an article is then the earliest link in the text to pass all three flags and the false link check.

Instead of relying on a sample of articles from which to generalize, we opted to process all 11.blah articles in the xml dump containing the 4.7 million articles along with redirected pages and disambugations, which we filter. To process the entirety of Wikipedia, we divided the xml dump into collections of articles, distributing the processing across 112 cores of the UVM sueprcomputer cluster. We then obtained a complete map of the First Link Network.

\subsection*{testing the algorithm}
\section{Using the Map: Navigating the First Links}

\textcolor{red}{reword the shit below}
\gray{
quote analysis from other areas/dodds paper: Droplets accumulating\\
first paper by Dodds cites Horton-Strahler (leaf pruning) not drop accumulation
}

To understand the structure of Wikipedia's First Link Network, we need to understand how the links flow from one page to another. The procedure, inspired by work in networks/enviromental something ((ask dodds)) is to treat each page as a starting point.

We follow the path of the first links until we reach either a dead link--a page linking to an invalid link (a file, another wikimedia project, or an external webpage)-- or a loop, where one page in the path is revisted.

What we can then measure are various metrics including the number of pages whose path leads to a particular page ((confusing wording)), the path length, even loops of a particular size. Combining these measures, we can then examine are particular properties of how pages related: for example the most popular two-loops.

Thus, from the directed network of First Links, we associate to each node several properties. Fortunately, we need not worry about issues of sampling error or statistiacal accuracy, since we do not rely on any sampling. Instead, thanks to the VACC we are able to follow every single possible path, tallying the properties of each node in the network.

Whether philosophy or another page is at the center of Wikipedia's First Link Network is a question about the page with the most traverses. More importantly, instead of a mere page, we can precisely identify the exact loop structure (or basin) where most pages feed.

Not only this, we can answer whether the top loop or page is in fact special--whether the number of pages feeding into the page/loop is measurably different from the other top pages/basins.

((add diagram to explain flow of procedure))

\section{The Stories of our Journey: our results}

\subsection{Clicks (beans...)}
Distribution is zipf's law:
((plot))
8.6 million of the 11 million pages in our raw dataset (11 million includes redirects to the subset of 4.2 million actual pages) have no clicks, with less than 1\% of pages accruing more than 100 clicks (precisely 8.56 million of the 11 have no visits).

((maybe sample and graph log-log to see if it's linear to justify zipf's law))

average number of visits is 21.

Even examining only the 1000 pages with the most page visist, we find a majority have no visit:

\includegraphics[scale=0.5]{graphics/top_1k_pages.png}

Of the top 100 pages, we see fewer than 30, contain an overwhelming majority of the traversals 6, 7 million:

\includegraphics[scale=0.5]{graphics/top_100_pages.png}

Maximum number of traversals is 7.4 million, but the mean is a mere 20 visits.

The top 7 pages, the ones with teh most visits comprise a loop around philosophy:

\includegraphics[scale=0.5]{graphics/top_pages_list.png}

\subsection{Longest Path}

The longest first link path span around 365 pages. These paths traverse calendar related pages including religious calendars such as Orthodox Liturgics and history pages. The Eastern Orthodox Liturgics pages for example are dated. Visit any page's Orthodox Liturgics and the first link will send to tomorrow's liturgics. The enxt days links to the following day's liturgics. Curiously, these are complete loops. The last calendar day of the year, simply links back to the first restarting the loop. Other notable pages with a path length of 80 include the legislative assembly of Ontario, 1868 in the UK,

\subsection{Loops - basics}
Although the loop of 7 pages around philosophy does attract the greatest number of visits, several other loops around natural fundamental concepts also exist. 
The top non-philosophy basins include a basin around community, another around landmass.

Identifying basins explains the structure of the network, but it doesn't reveal which page/s are a funnel for the basin. By following a similar approach and measuring where the drop accumulation is funneled we find:

\includegraphics[scale=0.5]{graphics/top_funnels.png}

confirming philosophy is indeed the largest funnel for the philosophy loop, implying it is in fact at the center of wikipedia. Nonetheless, curious foundational funnels also surface such as the word presentation, Tree of life (biology), Geographical regions such as Southern Europe, the US Constituion, health care and BBC. 
This amalgamation mixes both foundational topics, confirming teh theory whereby the first link generalizes a specific page by hooking it to a broader concept. The list also highlights broader topics that are perhaps less foundational, but seem to be at the forefront of people's minds. Health Care while not a foundational discipline such as biology or philosophy ranks at the top perhaps due to the importance we as a society attach to it. Similarly BBC, Hip Hop Music surface perhpas out of a cultural ttraction.

\red{how many two loops are there compared to the rest of wikipedia?}

In addition to measuring pages that funnel to accumulation, we can also measure true loops instead of just basins. These are pages which sit in a closed loop as opposed ones leading to a dead link.

Furthermore, since we can rank pages, we can rank loops based on the measure associated with each page. We first examine two loops:

\includegraphics[scale=0.4]{graphics/top_2loops.png}

Two loops reveal ideas that are closely linked to each other or nealry synomous such a secondary school the place wher esecondary education takes place. A lecture which is a form of a presentation, or fedualism and its examples.

Investigating the list beyond simply the top pages, but looking at a typical 2-loop we find a more natural link between (where typical means between 5 and 15 visits) linking a person to what they are most known for: say a book and author: Anatomy of Britain links to Anthony Sampson (and the reverse; an inventor and their invention: Voere and VEC-91; and an event links to the creator of the event: Peotry Bus Tour links to Wave Books.

While often a less intimate connection, 3-loops reveal a similar pattern. The top ranking 3-loops are:

\includegraphics[scale=0.4]{graphics/top_3loops.png}

There are 74k true 3loops.

\subsection{Later: Visualize loops}

\href{http://projects.flowingdata.com/tut/interactive_network_demo/ }{D3 Example 1}

\href{http://christophergandrud.github.io/d3Network/}{simple D3 Network}

\section{Conclusion}

\gray{Next: word stem to see how broader ideas are related (taxonomy of ideas).
    * Broader: what if we consider all links (or first par)? other languages? 
        * popular page by links to it. }

\gray{* Is the network stable over time? 
        * analyze a few newer datasets. }

The analysis remarkably demonstrates the links in Wikipedia truly exihibit a power-law like distribution where most pages link up to a handful of pages. Furthermore, Philosophy does occupy not only a special place inside the most popular basin of attraction. Not only that, the philosophy page itself is the page through which the feeds flow. This implies there is something special even from a mathematical standpoint. Philosophy isn't just a popular link, it lies in an exceptionally high percentage, unique to its loop.

Aside from the philosophy loop, other foundational concepts also emerge as the centers for basins of attraction. Fundamental concepts such as community, government, or philosophy. The network also has a prelevance of 2-loops ((find exact proportion or number)) which link two closely associated concepts such as an inventor or the invention as well as tie synonyms together. Other artifacts such as tremendously long loops based around calendar days demonstrate how we group ideas such as holidays, hisotiric events, or religious, not by their ties to other ideas, but first by where they fall on calendar days.

Based on the flow from specific pages to more general concepts, the results also agree with the heuristic provided here. Whereby, we link a very specific page to a broader concept, then to an ever broader concept. For example, the Wikipedia page for Banana links to fruit, then botanical then, ((confirm, fill-in)) then biology, then science and eventually philosophy. It is curious that Wikipedia, the largest encyclopedia, which by its very meaning aims to capture information about all branches of human knowledge, has naturally evolved through the works of ((x million editors/edits)) so that Knowledge, philsophy are the center links, occupying such a special place.

\end{multicols}

\subsection{Terms (aka style guide)}

use article instead of page

traversal = node weight
path length = degree of separation from a cluster
path is the flow from one page until a dead link or a repeated article

cycle (permutation from abstract algebra) instead of loop

basin of attraction or cluster? as opposed to cycle

invalid link (instead of dead or bad link)

To Do:
\textcolor{red}{
\begin{enumerate}
    \item Add limitations of work: snap shot in time, parsing is imperfect
    \item Add better graphics
    \item post code and results to github (add link to paper)
\end{enumerate}}

\newpage

\begin{thebibliography}{10}
\bibitem{1}
\href{ http://www.technologyreview.com/featuredstory/520446/the-decline-of-wikipedia/}{"The Decline of Wikipedia", MIT Tech Review}

\bibitem{2}
\href{https://en.wikipedia.org/wiki/Wikipedia:Database_download}{Wikipedia Database Download}

\bibitem{3}
\href{https://en.wikipedia.org/wiki/Wikipedia:Statistics}{Wikipedia Statistics}

\bibitem{4}
\href{http://en.wikipedia.org/wiki/Help:Special_page}{Media Markup Special Pages}
\end{thebibliography}

\end{document}

### 2015-10-04 writeup/results_section.tex
ready

\documentclass[twoside]{article}

% ------
% Fonts and typesetting settings
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\linespread{1.05} % Palatino needs more space between lines
\usepackage{microtype}
\usepackage{textcomp}

% ------
% Page layout
\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
\usepackage[font=it]{caption}
\usepackage{paralist}
\usepackage{multicol, hyperref, graphicx}

% ------
% Lettrines
\usepackage{lettrine}
\usepackage{color}

% ------
% Abstract
\usepackage{abstract}
	\renewcommand{\abstractnamefont}{\normalfont\bfseries}
	\renewcommand{\abstracttextfont}{\normalfont\small\itshape}

% ------
% Titling (section/subsection)
\usepackage{titlesec}
\renewcommand\thesection{\Roman{section}}
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{}

% ------
% Header/footer
\usepackage{fancyhdr}
	\pagestyle{fancy}
	\fancyhead{}
	\fancyfoot{}
	\fancyhead[C]{ $\bullet$ Draft $\bullet$}
	\fancyfoot[RO,LE]{\thepage}

% My Shortcuts

%useful shortcuts
\def\R{\ensuremath{\mathbb{R}}} %\ensuremath adds math mode, if forgotten
\def\Q{\ensuremath{\mathbb{Q}}}
\def\N{\ensuremath{\mathbb{N}}}
\def\Z{\ensuremath{\mathbb{Z}}}
\def\C{\ensuremath{\mathbb{C}}}

%shorcuts with arguments
\newcommand{\abs}[1]{\left\vert#1\right\vert} %nice absolute values
\newcommand{\bt}[1]{\textbf{#1}} %bold
\newcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
\newcommand{\cb}[1]{\centerline{\fbox{#1}}} %centered box
\newcommand{\bp}[1]{\fbox{\parbox{0.8\textwidth}{#1}}} %box paragraph
\newcommand{\norm}[1]{\left\lVert#1\right\rVert} %vector norm
\newcommand{\notimplies}{% does not imply
  \mathrel{{\ooalign{\hidewidth$\not\phantom{=}$\hidewidth\cr$\implies$}}}}
\renewcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations

%colors
\definecolor{javagreen}{rgb}{0.25,0.5,0.35} %dark green color
\definecolor{lightblue}{rgb}{0.149,0.545,0.824} %solarized blue
\definecolor{sred}{rgb}{0.863, 0.196, 0.184} %solarized red

\newcommand{\blue}[1]{{\leavevmode\color{lightblue}{#1}}} %solarized blue 
\newcommand{\green}[1]{{\leavevmode\color{javagreen}{#1}}} %command for green
\newcommand{\red}[1]{{\leavevmode\color{sred}{#1}}} %solarized red
\newcommand{\gray}[1]{{\leavevmode\color[gray]{0.5}{#1}}} %gray text

%environment
\newcommand{\tab}{\phantom{ssss}}
%----------------

% ------
% Clickable URLs (optional)
\usepackage{hyperref}

% ------
% Maketitle metadata
\title{\vspace{-5mm}%
	\fontsize{24pt}{12pt}\selectfont
	\textbf{What's So Special About Philosophy?} 
	}	
\author{%
\fontsize{14pt}{14pt}\selectfont
	Unraveling Wikipedia's First Link Network \vspace{-2mm}\\
	}
\date{}

%figures
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\maketitle
\thispagestyle{fancy}
%========================ABSTRACT====================================

\begin{abstract}
\fontsize{12pt}{12pt}
\selectfont
%\noindent Apples, oranges, and the most obscure Dylan song too---is any article a few clicks from Philosophy? The surprising answer is yes, $95\%$ of the time (according to a popular blog post).
%
%Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. Yet, we have never closely examined the curious web tying one entry to another. By following the first link in an article, we connect entries to form a directed network within Wikipedia: Wikpedia's First Link Network. Our aim is to study Wikipedia's First Link Network for insight into the topics, ideas, people, objects, and events we link. ((improve this sentence))
%
%We first map the First Link Network by algorithmically parsing the first link. Then, apply a river network model to identify the basins of attraction, loop structure, and characterize the many paths from one page's first link to the next. ((wording is unclear: as if there are many paths from one page))
%
\red{rewrite abstract}

\end{abstract}

%========================RESUTLS====================================
\fontsize{11pt}{11pt}
\selectfont
\section{Results}

\subsection{Traversals}

For each article, we the first link from one page to another up to an invalid link or cycle. By tracking the articles along each path, we can assign each article a rank measuring the number of paths containing a particular article. Through our traversal algorithm, we measure a total of 232.4 million 
traversals through the network of first links.

Remarkably, the distribution of traversals per article exihibits a power-law distribution \red{((citation + justification)) for justification look at definition of power-law; consider log-log with line}. Of the 11 million articles, $78\%$ have 0 traversals and $99\%$ have fewer than 100 traversals. Of the highest ranking 100 articles, fewer than $30$ articles have an overwhelming majority, \red{((~7 million, confirm))} of the total number of traversals.

\begin{figure}[H]
\centering
\caption{distribution of traversals}
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{graphics/top_1k_pages.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{graphics/top_100_pages.png}
    \end{subfigure}
\end{figure}

\begin{figure}[H]
\centering
    \caption{highest ranking articles by number of traversals}
    \begin{subfigure}[b]{0.8\textwidth}
        \includegraphics[width=\textwidth, angle=-90]{graphics/top_articles_list.png}
    \end{subfigure}
\end{figure}

The ranking by traversals reveals a cyclic structure marked by identical traversal numbers for the highest ranking 7 articles. By analyzing the traversal feeds, we can identify which articles in particular are the funnels or highest ranking within a particular cluster.

\red{\bt{TODO}}\\
\red{((look at other descriptives about a distribution or even descriptors of power-law))}\\

\subsection{Traversal Path Length}

By tracking the number of traversals to a repeated article or an invalid link, we can measure
the path length starting at any given article. 
We discovered the longest path length is 365, following 365 calendar days of Orthodox Liturgics, where the first link of today's Liturgic article links to the next day's and so on. Curiously, 
at on the last calendar day the article links back to the first day.
Similarly lengthy traversal paths included articles on the history of a nation on a particular year, for example 1953 in Scotland or 1560s in architecture.

Of the 11 million articles in our sample, 5.5 million had a path lenght of zero, meaning the first link is an invalid link or the article links to itslef (see section of cycles).
The most frequent path length is 29, with an interquartile range of (26, 30). The distribution of 
path lengths is similarly power-law like with few pages at the extreme of 355 traversals in their path length, while the majority is between 26-30.

\begin{figure}[H]
\centering
    \caption{Longest Path Lengths}
    \begin{subfigure}[b]{0.5\textwidth}
        \includegraphics[width=\textwidth]{graphics/top_1k_path_length.png}
    \end{subfigure}
\end{figure}

\subsection{Basins of Attraction and Traversal Feeds}

A basin of attraction in our network is collection of path-connected articles. These articles do not necessarily form a loop, but may end at an invalid link. These basins are one way of measuring the center the of a network. If we rank the basins of attraction based on the accumulated number of ttraversals we find: many basins around philosophy:

\begin{figure}[H]
\centering
    \caption{Longest Path Lengths}
    \begin{subfigure}[b]{0.5\textwidth}
        \includegraphics[width=\textwidth]{graphics/top_basins.png}
    \end{subfigure}
\end{figure}

Excluding any basins containing Philosophy, we find basins of attractions around Community, 
Landmass, Federal Government, Belief System, Presentation, and similarly foundational human concepts. This remarkable results confirms the foundational quality these ideas have in our english-based society. 
Whereby these concepts naturally arose out fo the links structure formed by teh owrk of many individuals working independtly on many topics often far aflied from one another. 
Curiously, gas and India both appear right after the foundational ideas, perhaps indicating the importance of these perhaps geographically or by population (second most populous country in the world) and gas which could refer to one fothe four fundamental states of matter or one of the most prized resources in the world.

\red{((maybe add graphic))}

Although identifying basins of attraction identifies a group of fundamental ideas, 
using the accumulated traversals, we can't identify whether one or more of the pages is a funnel. 
Among Philosophy, Existence, Knowledge, one page may be a funnel while the remaining simply happened to be connected to an article with many traversals. By measuring, where the traversal feeds, that is the number of traversals an article feeds into a loop, we can identify special funnel pages, 
which are part of a larger basin of attraction. Remarkable, within the Philosophy basin, 
Philosophy seems to be the singular funnel, accumulating an overwhelming majority of the traversals in the philosophy basin, with reality much farther behind (contributing a mere $.2\%$ of the 
traversals in the loop):

\begin{figure}[H]
\centering
    \caption{Top Funnels}
    \begin{subfigure}[b]{0.5\textwidth}
        \includegraphics[width=\textwidth]{graphics/top_funnels.png}
    \end{subfigure}
\end{figure}

By measuring the number of traversals an article feeds into a loop, rather than measuring the accumulation, we measure how remarkable the traversals are around Philosophy. The second ranking page on teh list, Presentation, is dwarfed at a mere $.4\%$ of the traversal feeds Philosophy holds. 
Nonetheless, the list reveals a curious list of top pages including biology as the foundation of life, presentation a fundamental human act, as well as ideas of geographical and political import. Naturally, the page for Census-Designated Place makes an appearance on the top 10, as an accumulation place for the geographical regions in the US. Other topics of cultural and political importance today are Health Care (around which much political debate, science, and economics has been discussed), engaging a large number of our population ((cite something)). As well as Hip Hop or BBC, which are of cultural import. Other high ranking pages include historical events such as the Cold War, Fossil Fuel.

\subsection{Network Cycles}

By tracking traversals we can also measure true network cycles (loops) rather than a simple chain
of connected articles with a dead end. While many of the lengthy paths we discussed in an earlier section terminate at an invalid link, the Eastern Orthodox Liturgics do form a true cycle in the network with a lenght of 365, one article per calendar day looping back to the January first. Other, less lengthy cycles also exist with lengths between 60-75 traversals. These cycles include Historical pages, Judicial Pages such as the Legislative Assembly of Ontario and various Japanese eras dating back to the 8th or 9th centuries.

Since articles iwth a path-length of two yielded curious results, we also measured and ranked the top 2-cycles in the network by accumulated traversals:

\begin{figure}[H]
\centering
\caption{highest ranking 2-Cycles}
    \begin{subfigure}[b]{0.8\textwidth}
        \includegraphics[width=\textwidth]{graphics/top_2loops.png}
    \end{subfigure}
\end{figure}

Of the 11 million artilces, only 83.6k are true 2-cycles. The average cycle traversals is 8, fewer than the typical article. However, Presentation does hold 31k traversals.

The top articles tended to be mere synoms rather than meaningful connections: Lecture and Presentation; Secondary eduaction adn seconday school; biological network and biological system; photogay and photograph. Outside the higest ranking articles, we does see a more meaningful connection among articles in 2-cycles: inventor to product: Voere -> VEC-91; book to author: Anatomy of Britain -> Anthony Sampson; event to organizer: Poetry Bus Tour -> Wave Books.

We also measured 3-cycles, cycles with length fewer than 100 etc. 3-Cycles were still expressing similar ideas or capturing synomous articles for similar ideas: Tree of life (biology) -> Tree of life (disambiguation) and Tree of life. Similarly with Cinema of India, Indian Cinemay and Telugu cinema  There were 74K 3-cycles. Once we looked beyond cycles of lenght 10 however, philosophy seemed to dominate along with the remaining list of higest ranking articles--this confirms the typical path length for high ranking pages is typically less than 10.

\begin{figure}[H]
\centering
\caption{highest ranking 3-Cycles}
    \begin{subfigure}[b]{0.8\textwidth}
        \includegraphics[width=\textwidth]{graphics/top_3loops.png}
    \end{subfigure}
\end{figure}
\newpage
\section*{Style Guide}

\begin{itemize}
\item Web Terms
    \begin{itemize}
        \item \bt{article} instead of page
        \item \bt{invalid link} (instead of dead or bad link)
    \end{itemize}
\item Measurements
    \begin{itemize}
        \item \bt{traversals} = node weight
        \item \bt{path length} = number of articles traversed until an invalid link or a repeated article
        \item \bt{traversal feeds} = traversals before a cycle is reached.
    \end{itemize}
\item Network Properties
    \begin{itemize}
        \item \bt{cycle} (permutation from abstract algebra) instead of loop
        \item \bt{path-connected articles} for both cycles and articles along the same path leading to an 
    invalid link
        \item \bt{basin of attraction} a path-connected group of articles (not necessarily at loop) with a high ranking number of traversals.
    \end{itemize}
\end{itemize}

\end{document}

### 2015-10-04 code/combine_clicks.py

@@ -1,0 +1,34 @@
+
+import json, gc
+
+clicks_path = "/users/m/s/msibrahi/v4/results/clicks/"
+fln_path = "/users/m/s/msibrahi/v4/results/flnetwork/fln.json"
+
+#load list of pages
+with open(fln_path) as f:
+    fln = json.load(f)
+
+clicks = {}
+
+clicks_files = []
+
+#add clicks_files to list
+for i in range(0, 112):
+    with open(clicks_path +str(i)+"click.json") as f:
+        clicks_chunk = json.load(f)
+        clicks_files.append(clicks_chunk)
+
+p = 0 
+for page in fln.iterkeys():
+    p += 1
+    if p % 10000 == 0: gc.collect()
+    page_clicks = 0 
+    for i in range(0, 112):
+        if page in clicks_files[i]:
+            page_clicks += clicks_files[i][page]
+    clicks[page] = page_clicks
+        
+
+
+with open(clicks_path + "clicks.json", "w") as f:
+    json.dump(clicks, f)

### 2015-10-04 code/combine_feeder.py

@@ -1,0 +1,33 @@
+import json, gc
+
+clicks_path = "/users/m/s/msibrahi/v4/results/feed_count/"
+fln_path = "/users/m/s/msibrahi/v4/results/flnetwork/fln.json"
+
+#load list of pages
+with open(fln_path) as f:
+    fln = json.load(f)
+
+clicks = {}
+
+clicks_files = []
+
+#add clicks_files to list
+for i in range(0, 112):
+    with open(clicks_path +str(i)+"feed_count.json") as f:
+        clicks_chunk = json.load(f)
+        clicks_files.append(clicks_chunk)
+
+p = 0 
+for page in fln.iterkeys():
+    p += 1
+    if p % 10000 == 0: gc.collect()
+    page_clicks = 0 
+    for i in range(0, 112):
+        if page in clicks_files[i]:
+            page_clicks += clicks_files[i][page]
+    clicks[page] = page_clicks
+        
+
+
+with open(clicks_path + "feed_count.json", "w") as f:
+    json.dump(clicks, f)

### 2015-10-04 code/combine_fln.py

@@ -1,0 +1,15 @@
+import json
+
+fln_path = "/users/m/s/msibrahi/v4/results/flnetwork/"
+
+fln = {}
+for i in range(0, 112):
+    with open(fln_path+str(i)+".json") as f:
+        fl = json.load(f)
+        fln.update(fl)
+
+#write fln.json 
+
+with open(fln_path+"fln.json", "w") as f:
+    json.dump(fln, f)
+

### 2015-10-04 code/combine_length.py

@@ -1,0 +1,35 @@
+
+import json, gc
+
+length_path = "/users/m/s/msibrahi/v4/results/path_length/"
+fln_path = "/users/m/s/msibrahi/v4/results/flnetwork/fln.json"
+
+#load list of pages
+with open(fln_path) as f:
+    fln = json.load(f)
+
+lengths = {}
+
+length_files = []
+
+#add length_files to list
+for i in range(0, 112):
+    with open(length_path +str(i)+"length.json") as f:
+        lengths_chunk = json.load(f)
+        length_files.append(lengths_chunk)
+
+p = 0 
+for page in fln.iterkeys():
+    p += 1
+    if p % 10000 == 0: gc.collect()
+    page_length = 0 
+    for i in range(0, 112):
+        if page in length_files[i]:
+            page_length += length_files[i][page]
+    lengths[page] = page_length
+        
+
+
+with open(length_path + "lengths.json", "w") as f:
+    json.dump(lengths, f)
+

### 2015-10-04 code/combine_nloops.py

@@ -1,0 +1,19 @@
+
+import json, gc
+
+length_path = "/users/m/s/msibrahi/v4/results/nloops/"
+
+nloops = {}
+
+#list of files 
+files = [str(i)+"nloops.json" for i in range(112)]
+ 
+for file in files:
+    with open(length_path + file) as f:
+        nloop = json.load(f)
+    nloops.update(nloop)
+
+
+with open(length_path + "nloops.json", "w") as f:
+    json.dump(nloops, f)
+

### 2015-10-04 code/combine_paths.py

@@ -1,0 +1,19 @@
+
+import json, gc, sys
+
+paths_path = "/users/m/s/msibrahi/v4/results/paths/"
+
+paths = {}
+s = int(sys.argv[1])
+n = int(sys.argv[2])
+
+for i in range(s, n):
+    with open(paths_path+str(i)+"path.json") as f:
+        path_chunk = json.load(f)
+        paths.update(path_chunk)
+
+#write fln.json 
+
+with open(paths_path + "cpaths" + str(s) + "_" + str(n) + ".json", "w") as f:
+    json.dump(paths, f)
+

### 2015-10-04 code/create_fln.py

@@ -1,0 +1,115 @@
+"""creates nth part of first link network using nth piece of xml dump"""
+#takes a file number, stores nth first link network as json
+import xml.etree.ElementTree as ET
+import first_link_txt, gc, json, string, sys
+#depends on first link extractor script "first_link_txt"
+
+def create_tree(data_path):
+    """xml tree object for parsing"""
+    tree = ET.iterparse(data_path)
+    tree = iter(tree) #iterable
+    return tree
+
+def get_title(elem):
+    """return title after checking"""
+    #validate title
+    if first_link_txt.check_link(elem.text):
+        try: 
+            title = elem.text
+        except:
+            title = ""
+    else: 
+        title = ""
+    return title
+
+def get_fl(elem):
+    """validate, return first link"""
+    try:
+        fl = first_link_txt.run_parser(elem.text)
+    except:
+        fl = ""
+    return fl
+    
+def process_xml(data_path):
+    """goes through each page in xml
+    returns dictionary: title -> first link"""
+    tree = create_tree(data_path)
+    
+    #initialize vars
+    fldict = {}
+    p = 0
+    title = ""
+    fl = ""
+
+    #root
+    event, root = tree.next()
+    
+    for action, elem in tree:
+        #free memory
+        if p % 10000 == 0: gc.collect() 
+        
+        if elem.tag[-4:] == "page":
+            #count
+            p += 1
+            #if p > 1000: break
+            
+            #write
+            if title and fl:
+                fldict[title] = fl
+            
+            #reset
+            title = ""
+            fl = ""
+            
+        if elem.tag[-5:] == "title":
+            title = get_title(elem)
+        
+        if elem.tag[-4:] == "text" and title:
+            fl = get_fl(elem)
+
+        elem.clear()
+        root.clear()
+    
+    return fldict
+
+def store_json(fldict, out_path):
+    """stores fldict as json file"""
+    with open(out_path, 'w') as outfile:
+        json.dump(fldict, outfile)
+
+def run_process(data_path, out_path):
+    """process xml and stores result in json"""
+    fldict = process_xml(data_path)
+    print fldict
+    store_json(fldict, out_path)
+    
+#to load json
+#with open(out_path, 'r') as jdict:
+#    fldict = json.load(jdict)
+
+def return_file_list():
+    """returns list of smallxx.xml files"""
[…]

### 2015-10-04 code/explore/get_loop.py

@@ -1,0 +1,54 @@
+"""create a dictionary of links traversed for nth fl chuck"""
+#takes a file num and creates xpath.json
+
+import json, gc, sys
+
+
+#goal path[page] = [link1, link2, link3]
+
+#load fln dictionary
+net_path = "/Users/mark/Desktop/wiki_v4/"
+
+with open(net_path + "fln.json") as f:
+    fln = json.load(f)
+def correct_cap(title):
+    """returns properly capitalized key, blank otherwise"""
+    try:
+        fl = fln[title]
+        return title
+    except:
+    #capitalize first letter only
+        try:
+            fl = fln[title[0].upper() + title[1:]]
+            return title[0].upper() + title[1:]
+        except:
+    #try title case
+            try:
+                fl = fln[title.title()]
+                return title.title()
+            except KeyError:
+                return ""
+
+def traverse(start):
+    """returns list of links traversed"""
+    #stop when link is repeated or dead link
+    links_traversed = []
+    try:
+        fl = correct_cap(fln[start])
+    except KeyError:
+        return "broken link" 
+    i = 0
+    while fl not in links_traversed and fl !="":
+        #maybe add clicks here?
+        i += 1
+        #break if too large
+        if i > 10000: 
+            fl = "exceeded 10k click limit"
+            break
+        #append fl
+        links_traversed.append(fl)
+        #jump to next link
+        fl = correct_cap(fln[fl])
+    return links_traversed
+
+            

### 2015-10-04 code/find_nloops.py

@@ -1,0 +1,108 @@
+"""
+create a dictionary of page: loop length
+includes only pages inside true loops
+"""
+#takes a file num and creates xpath.json
+
+import json, gc, sys
+
+
+#goal path[page] = [link1, link2, link3]
+
+def correct_cap(title):
+    """returns properly capitalized key, blank otherwise"""
+    try:
+        fl = fln[title]
+        return title
+    except:
+    #capitalize first letter only
+        try:
+            fl = fln[title[0].upper() + title[1:]]
+            return title[0].upper() + title[1:]
+        except:
+    #try title case
+            try:
+                fl = fln[title.title()]
+                return title.title()
+            except KeyError:
+                return ""
+
+def traverse_loops(fln, start):
+    """
+    if true loop,
+    returns list of links traversed
+    """
+    #stop when link is repeated or dead link
+    links_traversed = []
+    try:
+        fl = correct_cap(fln[start])
+    except KeyError:
+        return False
+    i = 0
+    while fl not in links_traversed and fl !="":
+        #maybe add clicks here?
+        i += 1
+        #break if too large
+        if i > 10000: 
+            fl = "exceeded 10k click limit"
+            break
+        #append fl
+        links_traversed.append(fl)
+        #jump to next link
+        fl = correct_cap(fln[fl])
+    if fl in links_traversed:
+        return links_traversed
+    else:
+        return False
+
+def run_traverse(fln, fln_chunk):       
+    """
+    loops through pages in fln_chunk and 
+    1. store true loops in dict 
+        - length of loop is key
+    """
+    nloops_dict = {}
+    #key is length -> list of loops (stored as sets)
+    i = 0
+    for page in fln_chunk.iterkeys():
+        i +=1
+        if i % 10000 == 0: 
+            print i
+            gc.collect()
+        #traverse path
+        page_path = traverse_loops(fln, page)
+        if page_path:
+            length = len(page_path)
+            nloops_dict[page] = length
+    return nloops_dict
+
+def write_result(dict, out_path):
+    """writes paths to json file"""
+    with open(out_path, 'w') as f:
+        json.dump(dict, f)
+
+            
+if __name__ == "__main__":
+    net_path = "/users/m/s/msibrahi/v4/results/flnetwork/"
+    #net_path = "/Users/mark/Desktop/wiki_v4/"
+
[…]

### 2015-10-04 code/find_paths.py

@@ -1,0 +1,125 @@
+"""create a dictionary of links traversed for nth fl chuck"""
+#takes a file num and creates xpath.json
+
+import json, gc, sys
+
+
+#goal path[page] = [link1, link2, link3]
+
+def correct_cap(title):
+    """returns properly capitalized key, blank otherwise"""
+    try:
+        fl = fln[title]
+        return title
+    except:
+    #capitalize first letter only
+        try:
+            fl = fln[title[0].upper() + title[1:]]
+            return title[0].upper() + title[1:]
+        except:
+    #try title case
+            try:
+                fl = fln[title.title()]
+                return title.title()
+            except KeyError:
+                return ""
+
+def traverse(fln, start):
+    """returns list of links traversed"""
+    #stop when link is repeated or dead link
+    links_traversed = []
+    try:
+        fl = correct_cap(fln[start])
+    except KeyError:
+        return "broken link" 
+    i = 0
+    while fl not in links_traversed and fl !="":
+        #maybe add clicks here?
+        i += 1
+        #break if too large
+        if i > 10000: 
+            fl = "exceeded 10k click limit"
+            break
+        #append fl
+        links_traversed.append(fl)
+        #jump to next link
+        fl = correct_cap(fln[fl])
+    return links_traversed
+
+def return_feeder_page(fln, page, page_path):
+    """returns feeder page if any"""
+    if len(page_path) > 1:
+        fl = correct_cap(fln[page_path[-1]])
+        if fl in page_path and fl != "":
+            return fl
+    return ""
+
+
+def run_traverse(fln, fln_chunk):       
+    """loops through pages in fln_chunk and 
+    1. stores paths traversed
+    2. stores clicks
+    3. stores len of path
+    4. stores clicks of page feeding into loop """
+    paths_chunk = {}
+    clicks_chunk = {} #page: clicks
+    path_length_chunk = {}  #page : path length
+    feed_count_chunk = {} #page : feed clicks
+    i = 0
+    for page in fln_chunk.iterkeys():
+        i +=1
+        if i % 10000 == 0: 
+            print i
+            gc.collect()
+        #traverse path
+        page_path = traverse(fln, page)
+        #add to dictionary
+        paths_chunk[page] = page_path
+        #update clicks 
[…]

### 2015-10-04 code/first_link_txt.py

@@ -1,0 +1,120 @@
+#parse page text for first link
+
+#for a list of pages: https://en.wikipedia.org/wiki/Special:AllPages
+#entire dump is https://dumps.wikimedia.org/enwiki/20141008/
+"""
+1. article body via xml tag
+2. clean tags: <ref>, <div>
+3. clean Media wiki templates {{ }}
+5. disregard parenthesis, not embedeed in links
+6. find first link 
+    * elimanting outermost false links:
+        *Image, wiktionary etc.
+"""
+import xml.etree.ElementTree as ET
+
+#first level of hierachy is check whether {{}}
+    #proceed to ()
+    #then test link
+    #link
+
+def inside_char(char, marker, tracker, i):
+    #checks whether inside char such as parentheses or wiki_template
+        #handles nested 
+    if char == marker[0]:
+        tracker.append(i)
+    elif char == marker[1]:
+        try:
+            tracker.pop()
+        except IndexError:
+            pass
+    return tracker
+        
+def grab_links(body):
+    #returns list of outer-most links not in parentheses or template or tags
+    links = []
+    link_char = []
+    
+    w_marker = ["{{", "}}"]
+    w_temp = [] #in template?
+    
+    par_marker = ["(", ")"]
+    par = [] #in parentheses?
+    
+    rtag_marker = ["<ref", "</re"]
+    rtag = [] #in <ref> tag?
+    
+    dtag_marker = ["<div", "</di"]
+    dtag = []
+    
+    skip_char = []
+    
+    for i, c in enumerate(body):
+        if i in skip_char: continue #eliminates double counting
+        char = body[i:i+2]
+        tag = body[i:i+4]
+        
+        #wiki template
+        w_temp = inside_char(char, w_marker, w_temp, i)
+        if char in w_marker: skip_char.append(i+1)
+        if w_temp:
+            continue #doesn't process if inside wiki template
+        
+        #parentheses
+        par = inside_char(c, par_marker, par, i)
+        if par:
+            continue
+        
+        #<ref> or <div>
+        rtag = inside_char(tag, rtag_marker, rtag, i)
+        dtag = inside_char(tag, dtag_marker, dtag, i)
+        if rtag or dtag:
+            continue
+        
+        #clear to add outer-most link
+        if char == '[[':
+            link_char.append(i)
+        elif char == ']]' and len(link_char) == 1:
+            links.append(body[link_char[0]:i+2])
+            link_char.pop()
+        elif char == ']]' and len(link_char) > 1:
[…]

### 2015-10-04 code/old_code/chopped_traverse.py

@@ -1,0 +1,105 @@
+
+#loop through entries in dictionary
+#traverse adding +1
+#stop once already traversed or dead
+
+
+import json, gc, sys
+
+net_path = "/users/m/s/msibrahi/v3/results/flnetwork/fldict.json"
+
+with open(net_path) as f:
+    fldict = json.load(f)
+
+
+#tdict = {page: clicks, cycle length, end page}
+
+def get_fl(fldict, title):
+    """adjusts case until title is a valid key, returns fl"""
+    #try as is
+    try:
+        fl = fldict[title]
+    except:
+    #capitalize first letter only
+        try:
+            fl = fldict[title[0].upper() + title[1:]]
+        except:
+    #try title case
+            try:
+                fl = fldict[title.title()]
+            except:
+                fl = ""
+    return fl
+            
+def test_traverse(fldict, start, n):
+    fl = get_fl(fldict, start)
+    for i in range(n):
+        print fl
+        if fl == "": break
+        fl = get_fl(fldict, fl)
+        
+def correct_cap(fldict, title):
+    """returns properly capitalized key, blank otherwise"""
+    try:
+        fl = fldict[title]
+        return title
+    except:
+    #capitalize first letter only
+        try:
+            fl = fldict[title[0].upper() + title[1:]]
+            return title[0].upper() + title[1:]
+        except:
+    #try title case
+            try:
+                fl = fldict[title.title()]
+                return title.title()
+            except KeyError:
+                return ""
+                
+
+    
+def traverse(fldict, start):
+    """traverses first link network until repeated link (or dead)
+    returns tdict: {page: clicks, cycle length, end page}"""
+    gc.collect() #memory
+    links_traversed = [] #change to set?
+    fl = get_fl(fldict, start)
+    i = 0 
+    while fl not in links_traversed and fl != "":
+        i +=1
+        if i > 10000: #limit to 10000 traverses
+            fl = "exceeded 10000 click limit"
+            break
+        fl = correct_cap(fldict, fl)
+        links_traversed.append(fl)
+        #add +1 click
+        clicks = tdict.get(fl, [0,0,""])[0] + 1
+        tdict[fl] = [clicks, tdict.get(fl, [0,0,""])[1], tdict.get(fl, [0,0,""])[2]]
+        #next click    
+        fl = get_fl(fldict, fl)
[…]

### 2015-10-04 code/old_code/combine_tdict.py

@@ -1,0 +1,67 @@
+# coding: utf-8
+
+#load all dictionaries into memory
+#load fldict into memory
+
+#loop through fldict, building master tdict
+
+import json, gc
+
+
+#load
+
+#fldict
+net_path = "/users/m/s/msibrahi/v3/results/flnetwork/fldict.json"
+with open(net_path) as f:
+    fldict = json.load(f)
+    
+
+#load ntdict
+ntdict_path = "/users/m/s/msibrahi/v3/results/tdict/"
+ntdict_names = ["tdict"+str(i)+".json" for i in range(0, 112)]
+ntdict_list = []
+
+for ntdict_name in ntdict_names:
+    with open(ntdict_path + ntdict_name) as f:
+        ntdict_list.append(json.load(f))
+
+
+
+
+#tdict = {page: clicks, cycle length, end page}
+tdict = {}
+
+
+def update_tdict(page, n):
+    """updates tdict with entry from nth tdict"""
+    ntdict = ntdict_list[n]
+    clicks, cycle_length, end_page = ntdict.get(page,[0,0,""])
+    
+    #update 
+    clicks = clicks + tdict.get(page,[0])[0]
+    cycle_length = cycle_length + tdict.get(page,[0,0])[1]
+    end_page = end_page + tdict.get(page, [0,0,""])[2]
+    
+    #assign
+    tdict[page] = [clicks, cycle_length, end_page]
+
+
+#loop through fldict
+i = 0
+for page in fldict.iterkeys():
+    i += 1
+    if i > 100: break
+    if i % 1000 == 0: gc.collect() #memory
+    
+    for n in range(0,112):
+        update_tdict(page,n)
+
+
+
+#write tdict
+with open(ntdict_path + "tdict.json", 'w') as outfile:
+    json.dump(tdict, outfile)
+
+
+
+

### 2015-10-04 code/old_code/old_vacc/chopped_traverse.py

@@ -1,0 +1,105 @@
+
+#loop through entries in dictionary
+#traverse adding +1
+#stop once already traversed or dead
+
+
+import json, gc, sys
+
+net_path = "/users/m/s/msibrahi/v3/results/flnetwork/fldict.json"
+
+with open(net_path) as f:
+    fldict = json.load(f)
+
+
+#tdict = {page: clicks, cycle length, end page}
+
+def get_fl(fldict, title):
+    """adjusts case until title is a valid key, returns fl"""
+    #try as is
+    try:
+        fl = fldict[title]
+    except:
+    #capitalize first letter only
+        try:
+            fl = fldict[title[0].upper() + title[1:]]
+        except:
+    #try title case
+            try:
+                fl = fldict[title.title()]
+            except:
+                fl = ""
+    return fl
+            
+def test_traverse(fldict, start, n):
+    fl = get_fl(fldict, start)
+    for i in range(n):
+        print fl
+        if fl == "": break
+        fl = get_fl(fldict, fl)
+        
+def correct_cap(fldict, title):
+    """returns properly capitalized key, blank otherwise"""
+    try:
+        fl = fldict[title]
+        return title
+    except:
+    #capitalize first letter only
+        try:
+            fl = fldict[title[0].upper() + title[1:]]
+            return title[0].upper() + title[1:]
+        except:
+    #try title case
+            try:
+                fl = fldict[title.title()]
+                return title.title()
+            except KeyError:
+                return ""
+                
+
+    
+def traverse(fldict, start):
+    """traverses first link network until repeated link (or dead)
+    returns tdict: {page: clicks, cycle length, end page}"""
+    gc.collect() #memory
+    links_traversed = [] #change to set?
+    fl = get_fl(fldict, start)
+    i = 0 
+    while fl not in links_traversed and fl != "":
+        i +=1
+        if i > 10000: #limit to 10000 traverses
+            fl = "exceeded 10000 click limit"
+            break
+        links_traversed.append(fl)
+        #add +1 click
+        clicks = tdict.get(fl, [0,0,""])[0] + 1
+        tdict[fl] = [clicks, tdict.get(fl, [0,0,""])[1], tdict.get(fl, [0,0,""])[2]]
+        #next click    
+        fl = correct_cap(fldict, get_fl(fldict, fl))
[…]

### 2015-10-04 code/old_code/old_vacc/combine_json.py

@@ -1,0 +1,13 @@
+
+import json
+
+fldict = {}
+for i in range(112):
+    fpath = "/users/m/s/msibrahi/v3/results/flnetwork/" + str(i) + ".json"
+    with open(fpath, 'r') as f:
+        data = json.load(f)
+        fldict.update(data)
+
+out_path = "/users/m/s/msibrahi/v3/results/flnetwork/fldict.json"
+with open(out_path, 'w') as outfile:
+    json.dump(fldict, outfile)

### 2015-10-04 code/old_code/old_vacc/combine_tdict.py

@@ -1,0 +1,66 @@
+# coding: utf-8
+
+#load all dictionaries into memory
+#load fldict into memory
+
+#loop through fldict, building master tdict
+
+import json, gc
+
+
+#load
+
+#fldict
+net_path = "/users/m/s/msibrahi/v3/results/flnetwork/fldict.json"
+with open(net_path) as f:
+    fldict = json.load(f)
+    
+
+#load ntdict
+ntdict_path = "/users/m/s/msibrahi/v3/results/tdict/"
+ntdict_names = ["tdict"+str(i)+".json" for i in range(0, 112)]
+ntdict_list = []
+
+for ntdict_name in ntdict_names:
+    with open(ntdict_path + ntdict_name) as f:
+        ntdict_list.append(json.load(f))
+
+
+
+
+#tdict = {page: clicks, cycle length, end page}
+tdict = {}
+
+
+def update_tdict(page, n):
+    """updates tdict with entry from nth tdict"""
+    ntdict = ntdict_list[n]
+    clicks, cycle_length, end_page = ntdict.get(page,[0,0,""])
+    
+    #update 
+    clicks = clicks + tdict.get(page,[0])[0]
+    cycle_length = cycle_length + tdict.get(page,[0,0])[1]
+    end_page = end_page + tdict.get(page, [0,0,""])[2]
+    
+    #assign
+    tdict[page] = [clicks, cycle_length, end_page]
+
+
+#loop through fldict
+i = 0
+for page in fldict.iterkeys():
+    i += 1
+    #if i > 100: break
+    if i % 10000 == 0: gc.collect() #memory
+    
+    for n in range(0,112):
+        update_tdict(page,n)
+
+
+#write tdict
+with open(ntdict_path + "tdict.json", 'w') as outfile:
+    json.dump(tdict, outfile)
+
+
+
+

### 2015-10-04 code/old_code/old_vacc/first_link_txt.py

@@ -1,0 +1,133 @@
+# -*- coding: utf-8 -*-
+# <nbformat>3.0</nbformat>
+
+#parse page xml (fed through dump)
+
+#for a list of pages: https://en.wikipedia.org/wiki/Special:AllPages
+#entire dump is https://dumps.wikimedia.org/enwiki/20141008/
+"""
+1. article body via xml tag
+2. clean tags: <ref>, <div>
+3. clean Media wiki templates {{ }}
+5. disregard parenthesis, not embedeed in links
+6. find first link 
+    * elimanting outermost false links:
+        *Image, wiktionary etc.
+"""
+import urllib2
+import xml.etree.ElementTree as ET
+
+# <codecell>
+
+def get_body(page_xml):
+    return page_xml
+
+# <codecell>
+
+#first level of hierachy is check whether {{}}
+    #proceed to ()
+    #then test link
+    #link
+
+def inside_char(char, marker, tracker, i):
+    #checks whether inside char such as parentheses or wiki_template
+        #handles nested 
+    if char == marker[0]:
+        tracker.append(i)
+    elif char == marker[1]:
+        try:
+            tracker.pop()
+        except IndexError:
+            pass
+        
+    return tracker
+        
+def grab_links(body):
+    #returns list of outer-most links not in parentheses or template or tags
+    links = []
+    link_char = []
+    
+    w_marker = ["{{", "}}"]
+    w_temp = [] #in template?
+    
+    par_marker = ["(", ")"]
+    par = [] #in parentheses?
+    
+    rtag_marker = ["<ref", "</re"]
+    rtag = [] #in <ref> tag?
+    
+    dtag_marker = ["<div", "</di"]
+    dtag = []
+    
+    skip_char = []
+    
+    for i, c in enumerate(body):
+        if i in skip_char: continue #eliminates double counting
+        char = body[i:i+2]
+        tag = body[i:i+4]
+        
+        #wiki template
+        w_temp = inside_char(char, w_marker, w_temp, i)
+        if char in w_marker: skip_char.append(i+1)
+        if w_temp:
+            continue #doesn't process if inside wiki template
+        
+        #parentheses
+        par = inside_char(c, par_marker, par, i)
+        if par:
+            continue
+        
+        #<ref> or <div>
+        rtag = inside_char(tag, rtag_marker, rtag, i)
+        dtag = inside_char(tag, dtag_marker, dtag, i)
+        if rtag or dtag:
+            continue
+        
[…]

### 2015-10-04 code/old_code/old_vacc/process_xml_v3.py

@@ -1,0 +1,111 @@
+
+
+import xml.etree.ElementTree as ET
+import first_link_txt, gc, json, string, sys
+
+#takes file number, stores first link network as json
+
+def create_tree(data_path):
+    """xml tree object for parsing"""
+    tree = ET.iterparse(data_path)
+    tree = iter(tree) #iterable
+    return tree
+
+def get_title(elem):
+    """return title after checking"""
+    #validate title
+    if first_link_txt.check_link(elem.text):
+        try: 
+            title = elem.text
+        except:
+            title = ""
+    else: 
+        title = ""
+    return title
+
+def get_fl(elem):
+    """validate, return first link"""
+    try:
+        fl = first_link_txt.run_parser(elem.text)
+    except:
+        fl = ""
+    return fl
+    
+def process_xml(data_path):
+    """goes through each page in xml
+    returns dictionary: title -> first link"""
+    tree = create_tree(data_path)
+    
+    #initialize vars
+    fldict = {}
+    p = 0
+    title = ""
+    fl = ""
+
+    #root
+    event, root = tree.next()
+    
+    for action, elem in tree:
+        #free memory
+        if p % 10000 == 0: gc.collect() 
+        
+        if elem.tag[-4:] == "page":
+            #count
+            p += 1
+            if p > 10: break
+            
+            #write
+            if title and fl:
+                fldict[title] = fl
+            
+            #reset
+            title = ""
+            fl = ""
+            
+        if elem.tag[-5:] == "title":
+            title = get_title(elem)
+        
+        if elem.tag[-4:] == "text" and title:
+            fl = get_fl(elem)
+
+        elem.clear()
+        root.clear()
+    
+    return fldict
+
+def store_json(fldict, out_path):
+    """stores fldict as json file"""
+    with open(out_path, 'w') as outfile:
+        json.dump(fldict, outfile)
+
+def run_process(data_path, out_path):
+    """process xml and stores result in json"""
+    fldict = process_xml(data_path)
+    print fldict
+    store_json(fldict, out_path)
+    
+#to load json
+#with open(out_path, 'r') as jdict:
+#    fldict = json.load(jdict)
+
+
+if __name__ == "__main__":
+    #can be called with argument for file number 
+    #list of 112 files
+    letter_range = ["a", "b", "c", "d"]
[…]

### 2015-10-04 code/old_code/old_vacc/run_vacc.py

@@ -1,0 +1,30 @@
+
+contents = """# This job needs 1 compute node with 1 processor per node.
+#PBS -l nodes=1:ppn=1
+# It should be allowed to run for up to 1 hour.
+#PBS -l walltime=06:00:00
+# Name of job.
+#PBS -N myjob
+# Join STDERR TO STDOUT.  (omit this if you want separate STDOUT AND STDERR)
+#PBS -j oe   
+# Send me mail on job start, job end and if job aborts
+#PBS -M mark.s.ibrahim@uvm.edu
+#PBS -m bea
+
+cd /users/m/s/msibrahi/v3/code/
+echo "This is myjob running on " `hostname`
+python process_xml_v3.py """
+
+import subprocess, os
+
+
+log_path = "/users/m/s/msibrahi/v3/code/vacc_logs/"
+os.chdir(log_path)
+
+for i in range(0, 112):
+    with open('wikijob'+str(i)+".script", "w") as f:
+        f.write(contents + str(i))
+
+for i in range(0,112):
+    command = "qsub wikijob" + str(i) + ".script" 
+    subprocess.call(command, shell=True)

### 2015-10-04 code/old_code/old_vacc/run_vacc_combine_tdict.py

@@ -1,0 +1,32 @@
+
+contents = """
+# This job needs 1 compute node with 1 processor per node.
+#PBS -l nodes=1:ppn=1,pmem=6gb,pvmem=7gb
+# It should be allowed to run for up to 1 hour.
+#PBS -l walltime=06:00:00
+# Name of job.
+#PBS -N combine_tdict
+# Join STDERR TO STDOUT.  (omit this if you want separate STDOUT AND STDERR)
+#PBS -j oe   
+# Send me mail on job start, job end and if job aborts
+#PBS -M mark.s.ibrahim@uvm.edu
+#PBS -m bea
+
+cd /users/m/s/msibrahi/v3/code/
+echo "This is myjob running on " `hostname`
+python step_combine_tdict.py """
+
+import subprocess, os
+
+
+log_path = "/users/m/s/msibrahi/v3/code/vacc_logs/"
+os.chdir(log_path)
+n=11
+
+for i in range(0,n):
+    with open('combine_tdict'+str(i)+".script", "w") as f:
+        f.write(contents + str(i))
+
+for i in range(0,n):
+    command = "qsub combine_tdict" + str(i) + ".script" 
+    subprocess.call(command, shell=True)

### 2015-10-04 code/old_code/old_vacc/run_vacc_traverse.py

@@ -1,0 +1,31 @@
+
+contents1 = """# This job needs 1 compute node with 1 processor per node.
+#PBS -l nodes=1:ppn=1,pmem=4gb,pvmem=5gb
+# It should be allowed to run for up to 1 hour.
+#PBS -l walltime=08:00:00
+# Name of job.
+#PBS -N traverse"""
+
+contents2="""# Join STDERR TO STDOUT.  (omit this if you want separate STDOUT AND STDERR)
+#PBS -j oe   
+# Send me mail on job start, job end and if job aborts
+#PBS -M mark.s.ibrahim@uvm.edu
+#PBS -m bea
+
+cd /users/m/s/msibrahi/v3/code/
+echo "This is myjob running on " `hostname`
+python chopped_traverse.py """
+
+import subprocess, os
+
+
+log_path = "/users/m/s/msibrahi/v3/code/vacc_logs/"
+os.chdir(log_path)
+
+for i in [112,113, 114]:
+    with open('traverse_job'+str(i)+".script", "w") as f:
+        f.write(contents1 + str(i) +"\n" + contents2 + str(i))
+
+for i in [112, 113, 114]:
+    command = "qsub traverse_job" + str(i) + ".script" 
+    subprocess.call(command, shell=True)

### 2015-10-04 code/old_code/old_vacc/step_combine_tdict.py

@@ -1,0 +1,110 @@
+
+# coding: utf-8
+
+#load all dictionaries into memory
+#load fldict into memory
+
+#loop through fldict, building 10 master tdict chucks
+    # need to add 100, 111 tdict manually
+#then run step_tdict 2 
+
+import json, gc, sys
+import datetime
+import shutil
+
+
+
+#load fldict
+def load_nfldict(n):
+    net_path = "/users/m/s/msibrahi/v3/results/flnetwork/"+str(n)+".json"
+    with open(net_path) as f:
+        fldict = json.load(f)
+    return fldict
+
+    
+
+#load ntdict
+ntdict_path = "/users/m/s/msibrahi/v3/results/tdict/"
+ntdict_names = ["tdict"+str(i)+".json" for i in range(0, 112)]
+ntdict_list = []
+
+def load_ntdict():
+    for ntdict_name in ntdict_names:
+        with open(ntdict_path + ntdict_name) as f:
+            ntdict_list.append(json.load(f))
+
+
+
+#tdict = {page: clicks, cycle length, end page}
+tdict = {}
+
+
+def update_tdict(page, n):
+    """updates tdict with entry from nth tdict"""
+    clicks, cycle_length, end_page = ntdict_list[n][page]
+    
+    #update 
+    if page in tdict:
+        clicks = clicks + tdict[page][0]
+        cycle_length = cycle_length + tdict[page][1]
+        if tdict[page][2] not in end_page:
+            end_page = end_page + tdict[page][2]
+        
+    #assign
+    tdict[page] = [clicks, cycle_length, end_page.strip()]
+
+
+#loop through pages in json (nfldict)  between (file_sec, +1)
+def loop_combine(n):
+    n = int(n)
+    for ln in range(n*10,(n+1)*10):
+        nfldict = load_nfldict(n)
+        p = 0
+        for page, v in nfldict.iteritems():
+            p += 1
+            #if p > 100: break
+            if p % 10000 == 0: gc.collect() #memory
+            
+            for i in range(0,112):
+                if page in ntdict_list[i]:
+                    update_tdict(page,i)
+
+
+#write tdict
+def write_tdict(n):
+    write_path = ntdict_path + "intermediate/tdict" + str(n) + ".json"
+    with open(write_path, 'w') as outfile:
+        json.dump(tdict, outfile)
+
+def run_show(n):
[…]

### 2015-10-04 code/old_code/old_vacc/step_combine_tdict2.py

@@ -1,0 +1,96 @@
+
+# coding: utf-8
+
+#load all dictionaries into memory
+#load fldict into memory
+
+#loop through fldict, building master tdict
+
+import json, gc, sys
+import datetime
+
+
+
+
+#load ntdict
+ntdict_path = "/users/m/s/msibrahi/v3/results/tdict/intermediate/"
+ntdict_names = ["tdict"+str(i)+".json" for i in range(0, 16)]
+ntdict_list = []
+
+def load_ntdict():
+    for ntdict_name in ntdict_names:
+        with open(ntdict_path + ntdict_name) as f:
+            ntdict_list.append(json.load(f))
+
+
+
+#tdict = {page: clicks, cycle length, end page}
+tdict = {}
+
+
+def update_tdict(page, n):
+    """updates tdict with entry from nth tdict"""
+    clicks, cycle_length, end_page = ntdict_list[n][page]
+    
+    #update 
+    if page in tdict:
+        clicks = clicks + tdict[page][0]
+        cycle_length = cycle_length + tdict[page][1]
+        if tdict[page][2] not in end_page:
+            end_page = end_page + tdict[page][2]
+        
+    #assign
+    tdict[page] = [clicks, cycle_length, end_page.strip()]
+
+
+#loop through tdict in intermediate folder
+def get_keys():
+    full_keys = []
+    for i in range(16):
+        nkeys = ntdict_list[i].keys()
+        full_keys += nkeys
+        full_keys = list(set(full_keys))
+        if i == 5: gc.collect()
+    return list(full_keys)
+
+def loop_combine(full_keys):
+    full_keys = get_keys()
+    p = 0 
+    for page in full_keys:
+        p += 1
+        #if p > 100: break
+        if p % 10000 == 0: gc.collect() #memory
+        
+        for i in range(0,16):
+            if page in ntdict_list[i]:
+                update_tdict(page,i)
+
+
+#write tdict
+def write_tdict():
+    write_path = ntdict_path + "/full_tdict.json"
+    with open(write_path, 'w') as outfile:
+        json.dump(tdict, outfile)
+
+def run_show():
+    #load ntdict
+    load_ntdict()
+    #get list of keys in intermediate
+    full_keys = get_keys()
+    #loop combine
+    loop_combine(full_keys)
+    #write
+    write_tdict()
+
+run_show()
+
+
+    
+#run time for 100 pages on 6th list (n=6)
+    # < 1 
[…]

### 2015-10-04 code/old_code/old_vacc/traverse.py

@@ -1,0 +1,97 @@
+
+#loop through entries in dictionary
+#traverse adding +1
+#stop once already traversed or dead
+
+
+import json, gc
+
+net_path = "/users/m/s/msibrahi/v3/results/flnetwork/fldict.json"
+
+with open(net_path) as f:
+    fldict = json.load(f)
+
+
+#tdict = {page: clicks, cycle length, end page}
+
+def get_fl(fldict, title):
+    """adjusts case until title is a valid key, returns fl"""
+    #try as is
+    try:
+        fl = fldict[title]
+    except:
+    #capitalize first letter only
+        try:
+            fl = fldict[title[0].upper() + title[1:]]
+        except:
+    #try title case
+            try:
+                fl = fldict[title.title()]
+            except:
+                fl = ""
+    return fl
+            
+def test_traverse(fldict, start, n):
+    fl = get_fl(fldict, start)
+    for i in range(n):
+        print fl
+        if fl == "": break
+        fl = get_fl(fldict, fl)
+        
+def correct_cap(fldict, title):
+    """returns properly capitalized key, blank otherwise"""
+    try:
+        fl = fldict[title]
+        return title
+    except:
+    #capitalize first letter only
+        try:
+            fl = fldict[title[0].upper() + title[1:]]
+            return title[0].upper() + title[1:]
+        except:
+    #try title case
+            try:
+                fl = fldict[title.title()]
+                return title.title()
+            except KeyError:
+                return ""
+                
+
+    
+def traverse(fldict, start):
+    """traverses first link network until repeated link (or dead)
+    returns tdict: {page: clicks, cycle length, end page}"""
+    gc.collect() #memory
+    links_traversed = [] #change to set?
+    fl = get_fl(fldict, start)
+    i = 0 
+    while fl not in links_traversed and fl != "":
+        i +=1
+        if i > 10000: #limit to 10000 traverses
+            fl = "exceeded 10000 click limit"
+            break
+        links_traversed.append(fl)
+        #add +1 click
+        clicks = tdict.get(fl, [0,0,""])[0] + 1
+        tdict[fl] = [clicks, tdict.get(fl, [0,0,""])[1], tdict.get(fl, [0,0,""])[2]]
+        #next click    
+        fl = correct_cap(fldict, get_fl(fldict, fl))
+    #store cycle length and end page
[…]

### 2015-10-04 code/old_code/process/chop_wiki.py

@@ -1,0 +1,68 @@
+
+# coding: utf-8
+
+# In[20]:
+
+#create smallaa.xml with proper pages tags
+
+import string, os
+
+
+# In[20]:
+
+
+
+
+# In[21]:
+
+#construct list of 100 files
+letter_range = ["a", "b", "c", "d"]
+file_list = [a + b for a in ["small" + a for a in letter_range] for b in list(string.ascii_lowercase)]
+extra_files = ["small" + a for a in ["ea", "eb", "ec", "ed", "ee", "ef", "eg", "eh"]]
+
+full_file_list = file_list + extra_files
+
+
+#data location
+data_path = "/Users/mark/Desktop/wiki_split/"
+data_path = "/Volumes/mark external/wikipedia/clean_small/"
+data_path = "/users/m/s/msibrahi/full_wiki_data/"
+
+for file_num in range(len(full_file_list)):
+    #get lines
+    xml_file = data_path + full_file_list[file_num+1]
+    cut_off = ""
+    with open(xml_file, 'r') as f:
+        for line in f:
+            cut_off = cut_off + line
+            if "</page>" in line:
+                break
+
+    #append lines
+    xml_write_file = data_path + full_file_list[file_num]
+    with open(xml_write_file, 'a') as f:
+        f.write(cut_off)
+    
+    #rename file
+    os.rename(xml_write_file, xml_write_file+".xml")
+
+
+# In[14]:
+
+
+
+
+# In[14]:
+
+
+
+
+# In[14]:
+
+
+
+
+# In[ ]:
+
+
+

### 2015-10-04 code/old_code/process/first_link.py

@@ -1,0 +1,151 @@
+# -*- coding: utf-8 -*-
+# <nbformat>3.0</nbformat>
+
+# <codecell>
+
+#parse page xml (fed through dump)
+
+#for a list of pages: https://en.wikipedia.org/wiki/Special:AllPages
+#entire dump is https://dumps.wikimedia.org/enwiki/20141008/
+"""
+1. article body via xml tag
+2. clean tags: <ref>, <div>
+3. clean Media wiki templates {{ }}
+5. disregard parenthesis, not embedeed in links
+6. find first link 
+    * elimanting outermost false links:
+        *Image, wiktionary etc.
+"""
+import urllib2
+import xml.etree.ElementTree as ET
+
+# <codecell>
+
+def get_body(page_xml):
+    #returns xml body text
+    #up to first 25k characters
+    
+    #extract article body
+    tree = ET.fromstring(page_xml)
+    try:
+        for child in tree:
+            if child.tag[-8:] == "revision":
+                for j in child:
+                    if j.tag[-4:] == "text":
+                        body = j.text
+                        
+
+    except:
+        print "can't parse article body"
+        return False
+    return body[:25000] #returns body 
+
+# <codecell>
+
+#first level of hierachy is check whether {{}}
+    #proceed to ()
+    #then test link
+    #link
+
+def inside_char(char, marker, tracker, i):
+    #checks whether inside char such as parentheses or wiki_template
+        #handles nested 
+    if char == marker[0]:
+        tracker.append(i)
+    elif char == marker[1]:
+        try:
+            tracker.pop()
+        except IndexError:
+            pass
+        
+    return tracker
+        
+def grab_links(body):
+    #returns list of outer-most links not in parentheses or template or tags
+    links = []
+    link_char = []
+    
+    w_marker = ["{{", "}}"]
+    w_temp = [] #in template?
+    
+    par_marker = ["(", ")"]
+    par = [] #in parentheses?
+    
+    rtag_marker = ["<ref", "</re"]
+    rtag = [] #in <ref> tag?
+    
+    dtag_marker = ["<div", "</di"]
+    dtag = []
+    
+    skip_char = []
+    
+    for i, c in enumerate(body):
+        if i in skip_char: continue #eliminates double counting
+        char = body[i:i+2]
+        tag = body[i:i+4]
+        
+        #wiki template
+        w_temp = inside_char(char, w_marker, w_temp, i)
+        if char in w_marker: skip_char.append(i+1)
[…]

### 2015-10-04 code/old_code/process/first_link_txt.py

@@ -1,0 +1,133 @@
+# -*- coding: utf-8 -*-
+# <nbformat>3.0</nbformat>
+
+#parse page xml (fed through dump)
+
+#for a list of pages: https://en.wikipedia.org/wiki/Special:AllPages
+#entire dump is https://dumps.wikimedia.org/enwiki/20141008/
+"""
+1. article body via xml tag
+2. clean tags: <ref>, <div>
+3. clean Media wiki templates {{ }}
+5. disregard parenthesis, not embedeed in links
+6. find first link 
+    * elimanting outermost false links:
+        *Image, wiktionary etc.
+"""
+import urllib2
+import xml.etree.ElementTree as ET
+
+# <codecell>
+
+def get_body(page_xml):
+    return page_xml
+
+# <codecell>
+
+#first level of hierachy is check whether {{}}
+    #proceed to ()
+    #then test link
+    #link
+
+def inside_char(char, marker, tracker, i):
+    #checks whether inside char such as parentheses or wiki_template
+        #handles nested 
+    if char == marker[0]:
+        tracker.append(i)
+    elif char == marker[1]:
+        try:
+            tracker.pop()
+        except IndexError:
+            pass
+        
+    return tracker
+        
+def grab_links(body):
+    #returns list of outer-most links not in parentheses or template or tags
+    links = []
+    link_char = []
+    
+    w_marker = ["{{", "}}"]
+    w_temp = [] #in template?
+    
+    par_marker = ["(", ")"]
+    par = [] #in parentheses?
+    
+    rtag_marker = ["<ref", "</re"]
+    rtag = [] #in <ref> tag?
+    
+    dtag_marker = ["<div", "</di"]
+    dtag = []
+    
+    skip_char = []
+    
+    for i, c in enumerate(body):
+        if i in skip_char: continue #eliminates double counting
+        char = body[i:i+2]
+        tag = body[i:i+4]
+        
+        #wiki template
+        w_temp = inside_char(char, w_marker, w_temp, i)
+        if char in w_marker: skip_char.append(i+1)
+        if w_temp:
+            continue #doesn't process if inside wiki template
+        
+        #parentheses
+        par = inside_char(c, par_marker, par, i)
+        if par:
+            continue
+        
+        #<ref> or <div>
+        rtag = inside_char(tag, rtag_marker, rtag, i)
+        dtag = inside_char(tag, dtag_marker, dtag, i)
+        if rtag or dtag:
+            continue
+        
[…]

### 2015-10-04 code/old_code/process/mini_chop.py

@@ -1,0 +1,56 @@
+
+# coding: utf-8
+
+# In[20]:
+
+#create smallaa.xml with proper pages tags
+
+import string, os
+
+
+# In[20]:
+
+
+
+
+# In[21]:
+
+data_path = "/users/m/s/msibrahi/full_wiki_data/"
+
+xml_write_file = data_path + "smallee"
+xml_file = data_path + "smallef.xml"
+
+cut_off = ""
+with open(xml_file, 'r') as f:
+    for line in f:
+        cut_off = cut_off + line
+        if "</page>" in line:
+            break
+
+#append lines
+with open(xml_write_file, 'a') as f:
+    f.write(cut_off)
+
+#rename file
+os.rename(xml_write_file, xml_write_file+".xml")
+
+
+# In[14]:
+
+
+
+
+# In[14]:
+
+
+
+
+# In[14]:
+
+
+
+
+# In[ ]:
+
+
+

### 2015-10-04 code/old_code/process/process_xml_v3.py

@@ -1,0 +1,112 @@
+
+
+import xml.etree.ElementTree as ET
+import first_link_txt, gc, json, string, sys
+
+
+#data_path = "/Volumes/mark external/wikipedia/enwiki_20141106.xml"
+#data_path = "/Volumes/mark external/wikipedia/clean_small/smallaf.xml"
+data_path = "/Users/mark/Desktop/wiki_v3/smallag.xml"
+out_path = "/Users/mark/Desktop/test.json"
+
+
+
+def create_tree(data_path):
+    """xml tree object for parsing"""
+    parser = ET.XMLParser()
+    tree = ET.iterparse(data_path)
+    return tree
+
+def get_title(elem):
+    """return title after checking"""
+    #validate title
+    if first_link_txt.check_link(elem.text):
+        try: 
+            title = elem.text
+        except:
+            title = ""
+    else: 
+        title = ""
+    return title
+
+def get_fl(elem):
+    """validate, return first link"""
+    try:
+        fl = first_link_txt.run_parser(elem.text)
+    except:
+        fl = ""
+    return fl
+    
+def process_xml(data_path):
+    """goes through each page in xml
+    returns dictionary: title -> first link"""
+    tree = create_tree(data_path)
+    
+    #initialize vars
+    fldict = {}
+    p = 0
+    title = ""
+    fl = ""
+    
+    for action, elem in tree:
+        #free memory
+        if p % 100 == 0: gc.collect() 
+        
+        if elem.tag[-4:] == "page":
+            #count
+            p += 1
+            if p > 1000: break
+            
+            #write
+            if title and fl:
+                fldict[title] = fl
+            
+            #reset
+            title = ""
+            fl = ""
+            
+        if elem.tag[-5:] == "title":
+            title = get_title(elem)
+        
+        if elem.tag[-4:] == "text" and title:
+            fl = get_fl(elem)
+    
+    return fldict
+
+def store_json(fldict, out_path):
+    """stores fldict as json file"""
+    with open(out_path, 'w') as outfile:
+        json.dump(fldict, outfile)
+
+def run_process(data_path, out_path):
+    """process xml and stores result in json"""
+    fldict = process_xml(data_path)
+    store_json(fldict, out_path)
+    
+#to load json
+#with open(out_path, 'r') as jdict:
+#    fldict = json.load(jdict)
+
+
[…]

### 2015-10-04 code/old_code/traverse.py

@@ -1,0 +1,98 @@
+
+#loop through entries in dictionary
+#traverse adding +1
+#stop once already traversed or dead
+
+
+import json, gc
+import pandas as pd
+
+net_path = "/users/m/s/msibrahi/v3/results/flnetwork/fldict.json"
+
+with open(net_path) as f:
+    fldict = json.load(f)
+
+
+#tdict = {page: clicks, cycle length, end page}
+
+def get_fl(fldict, title):
+    """adjusts case until title is a valid key, returns fl"""
+    #try as is
+    try:
+        fl = fldict[title]
+    except:
+    #capitalize first letter only
+        try:
+            fl = fldict[title[0].upper() + title[1:]]
+        except:
+    #try title case
+            try:
+                fl = fldict[title.title()]
+            except:
+                fl = ""
+    return fl
+            
+def test_traverse(fldict, start, n):
+    fl = get_fl(fldict, start)
+    for i in range(n):
+        print fl
+        if fl == "": break
+        fl = get_fl(fldict, fl)
+        
+def correct_cap(fldict, title):
+    """returns properly capitalized key, blank otherwise"""
+    try:
+        fl = fldict[title]
+        return title
+    except:
+    #capitalize first letter only
+        try:
+            fl = fldict[title[0].upper() + title[1:]]
+            return title[0].upper() + title[1:]
+        except:
+    #try title case
+            try:
+                fl = fldict[title.title()]
+                return title.title()
+            except KeyError:
+                return ""
+                
+
+    
+def traverse(fldict, start):
+    """traverses first link network until repeated link (or dead)
+    returns tdict: {page: clicks, cycle length, end page}"""
+    gc.collect() #memory
+    links_traversed = [] #change to set?
+    fl = get_fl(fldict, start)
+    i = 0 
+    while fl not in links_traversed and fl != "":
+        i +=1
+        if i > 10000: #limit to 10000 traverses
+            fl = "exceeded 10000 click limit"
+            break
+        fl = correct_cap(fldict, fl)
+        links_traversed.append(fl)
+        #add +1 click
+        clicks = tdict.get(fl, [0,0,""])[0] + 1
+        tdict[fl] = [clicks, tdict.get(fl, [0,0,""])[1], tdict.get(fl, [0,0,""])[2]]
+        #next click    
+        fl = get_fl(fldict, fl)
[…]

### 2015-10-04 code/run_combinepaths.py

@@ -1,0 +1,30 @@
+contents = """# This job needs 1 compute node with 1 processor per node.
+#PBS -l nodes=1:ppn=1,pmem=8gb,pvmem=9gb
+# It should be allowed to run for up to 6 hours.
+#PBS -l walltime=06:00:00
+# Name of job.
+#PBS -N combine_paths
+# Join STDERR TO STDOUT.  (omit this if you want separate STDOUT AND STDERR)
+#PBS -j oe   
+# Send me mail on job start, job end and if job aborts
+#PBS -M mark.s.ibrahim@uvm.edu
+#PBS -m bea
+
+cd /users/m/s/msibrahi/v4/code/
+echo "This is myjob running on " `hostname`
+python combine_paths.py """
+
+import subprocess, os
+
+
+log_path = "/users/m/s/msibrahi/v4/code/vacc_logs/"
+os.chdir(log_path)
+
+chunks = [(0,20), (20, 40), (40, 60), (60, 80), (80,100), (100,112)]
+
+for chunk in chunks:
+    with open("combinepathsjob"+ str(chunk[0])+ ".script", "w") as f:
+        f.write(contents + str(chunk[0]) + " " + str(chunk[1]))
+
+    command = "qsub " + "combinepathsjob" + str(chunk[0]) + ".script"
+    subprocess.call(command, shell=True)

### 2015-10-04 code/run_fln.py

@@ -1,0 +1,31 @@
+contents1 = """# This job needs 1 compute node with 1 processor per node.
+#PBS -l nodes=1:ppn=1
+# It should be allowed to run for up to 1 hour.
+#PBS -l walltime=06:00:00
+# Name of job.
+#PBS -N create_fln"""
+
+contents2 =""" # Join STDERR TO STDOUT.  (omit this if you want separate STDOUT AND STDERR)
+#PBS -j oe   
+# Send me mail on job start, job end and if job aborts
+#PBS -M mark.s.ibrahim@uvm.edu
+#PBS -m bea
+
+cd /users/m/s/msibrahi/v4/code/
+echo "This is myjob running on " `hostname`
+python create_fln.py """
+
+import subprocess, os
+
+
+log_path = "/users/m/s/msibrahi/v4/code/vacc_logs/"
+os.chdir(log_path)
+n = 112
+
+for i in range(0, n):
+    with open('wikijob'+str(i)+".script", "w") as f:
+        f.write(contents1 + str(i) + "\n" + contents2 + str(i))
+
+for i in range(0, n):
+    command = "qsub wikijob" + str(i) + ".script" 
+    subprocess.call(command, shell=True)

### 2015-10-04 code/run_nloops.py

@@ -1,0 +1,31 @@
+contents1 = """# This job needs 1 compute node with 1 processor per node.
+#PBS -l nodes=1:ppn=1,pmem=8gb,pvmem=9gb
+# It should be allowed to run for up to 1 hour.
+#PBS -l walltime=06:00:00
+# Name of job.
+#PBS -N find_nloops"""
+
+contents2 =""" # Join STDERR TO STDOUT.  (omit this if you want separate STDOUT AND STDERR)
+#PBS -j oe   
+# Send me mail on job start, job end and if job aborts
+#PBS -M mark.s.ibrahim@uvm.edu
+#PBS -m bea
+
+cd /users/m/s/msibrahi/v4/code/
+echo "This is myjob running on " `hostname`
+python find_nloops.py """
+
+import subprocess, os
+
+
+log_path = "/users/m/s/msibrahi/v4/code/vacc_logs/"
+os.chdir(log_path)
+n = 112
+
+for i in range(0, n):
+    with open('wikijob'+str(i)+".script", "w") as f:
+        f.write(contents1 + str(i) + "\n" + contents2 + str(i))
+
+for i in range(0, n):
+    command = "qsub wikijob" + str(i) + ".script" 
+    subprocess.call(command, shell=True)

### 2015-10-04 code/run_paths.py

@@ -1,0 +1,31 @@
+contents1 = """# This job needs 1 compute node with 1 processor per node.
+#PBS -l nodes=1:ppn=1,pmem=8gb,pvmem=9gb
+# It should be allowed to run for up to 6 hours.
+#PBS -l walltime=06:00:00
+# Name of job.
+#PBS -N find_paths"""
+
+contents2 =""" # Join STDERR TO STDOUT.  (omit this if you want separate STDOUT AND STDERR)
+#PBS -j oe   
+# Send me mail on job start, job end and if job aborts
+#PBS -M mark.s.ibrahim@uvm.edu
+#PBS -m bea
+
+cd /users/m/s/msibrahi/v4/code/
+echo "This is myjob running on " `hostname`
+python find_paths.py """
+
+import subprocess, os
+
+
+log_path = "/users/m/s/msibrahi/v4/code/vacc_logs/"
+os.chdir(log_path)
+n = 112
+
+for i in range(0, n):
+    with open('wikijob'+str(i)+".script", "w") as f:
+        f.write(contents1 + str(i) + "\n" + contents2 + str(i))
+
+for i in range(0, n):
+    command = "qsub wikijob" + str(i) + ".script" 
+    subprocess.call(command, shell=True)

### 2015-10-04 writeup/draft.tex
rough draft

@@
 \usepackage{microtype}
+\usepackage{textcomp}
 
@@
 \usepackage{multicol, hyperref, graphicx}
+\usepackage{parskip}
 
@@
 \usepackage{subcaption}
+\usepackage{float}
 
@@
 %========================ABSTRACT====================================
+
 \begin{abstract}
@@
 \selectfont
-\noindent Apples, oranges, and the most obscure Dylan song too---is any article a few clicks from Philosophy? The surprising answer is yes, $95\%$ of the time (according to a popular blog post).
+
+Apples, oranges, and the most obscure Dylan song too---is everything a few clicks from Philosophy? In Wikipedia, the surprising answer is yes.
 
 Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed.
-Yet, we have never closely examined the curious web tying one entry to another.
-By following the first link in an article, we connect entries to form a directed network within Wikipedia: Wikpedia's First Link Network.
-Our aim is to study Wikipedia's First Link Network for insight into the topics, ideas, people, objects, and events we link.
-((improve this sentence))
+Yet, we have never closely examined the naturally formed web tying one entry to another.
+By following the first link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network.
+Here we study Wikipedia's First Link Network for insight into how we link the largest collection of topics, ideas, people, objects, and events.
+Structurally, we discover Wikipedia's First Link Network is by many metrics a scale-free network with Philosophy at a salient center by many orders of magnitude.
+We also learn the network generalizes specific events, objects, or phenomena to broader basins around Community, State, and Science.
 
-We first map the First Link Network by algorithmically parsing the first link.
-Then, apply a river network model to identify the basins of attraction, loop structure, and characterize the many paths from one page's first link to the next.
-((wording is unclear: as if there are many paths from one page))
-
-\red{rewrite abstract}
-
+\green{
+conclusions to mention
+\begin{itemize}
+    \item finding many small-world phenomenan in the number of first links flowing through a page, the path lenghts, and ...\\
+    \item also cementing Philosophy as an extremely peculiar article, orders of magnitude away from other article in Wikipedia by our metrics.
+    \item links generalize specific event, object, or phenomemon to a broader concept. Mention synonyms
+\end{itemize}
+}
 \end{abstract}
 
-%==================================BASE====================================================
+\fontsize{11pt}{11pt}
+\selectfont
 
-\begin{multicols}{2}
+%========================INTRODUCTION====================================
+At no point in history has a larger or more meticulously indexed collection of human knowledge existed.
+((cite))
+In amassing such an awe-inspiring collection, we formed an equally impressive web. 
+Through the efforts of millions of individuals, working independently,
+((cite))
+we naturally linked topics, inventions, people, objects, places, and events across space and time.
+The web we weaved, and continue to weave, is a wealth of information not only about those notable inventions, 
+places, figures, and ideas, but also about \textit{relationships} among them.
+Here we study the relationships in this naturally arising web through the hyperlinks connecting one article to another.
 
-\lettrine[nindent=0em,lines=2]{A}t no point in history has a larger or more meticulously indexed collection of human knowledge existed \cite{1}.
-In amassing such an awe-inspiring collection, we naturally linked topics, inventions, people, objects, places, and events
-\footnote{while anyone with internet access is free to contribute "neutral, cited information," it's worth noting the population of contributors is not necessarily representative of the overall english speaking population.
-Notably, the contributors are overwhelmingly male ((cite, expand)).} across space and time.
-The web we weaved, and continue to weave, is a wealth of information about those notable inventions, places, figures, and ideas, but also about us: the ones weaving the web.
-Here we explore this web as a directed network formed by the first link connecting one article to another.
+We build Wikipedia's First Link Network by following the first link, not in parenthesis, inside the main body of each article in the English version of Wikipedia. 
+For the directed network to meaningfully reflect how we associate one article to another, we exclude links in parenthesis, 
+disregarding pronunciation keys or disambiguations.
+We also exclude any links in the side-bar elements, as well as any links to external pages, files, or WikiMedia 
+projects outside of Wikipedia (such as Wiktionary).
+((footnote: this also stays true to the original blog post))
+The result is a directed network placing every article in a broader web of ideas.
 
-We build Wikipedia's First link network by following the first link, not in parenthesis, inside the main body of each article of the English version of Wikipedia.
-For the directed network to meaningfully reflect how we associate one article to another, we exclude links in parenthesis containing pronunciation keys or disambiguations.
-We also exclude any links in the side-bar elements, as well as any links to external pages, files, or other WikiMedia projects outstide of Wikipedia such as Wiktionary ((lookup others))
-\footnote{
-These constraints are the same which led to the popular blog post/s claiming nearly every article is within fewer than 20 first links away from philosophy. 
-}
-thus, forming a true network of Wikipedia.
+%===========================Data_and_Parsing================================
+\section{Constructing the First Link Network}
 
-\includegraphics[scale=0.18]{graphics/wiki_train.png}
-\ \\
+To map Wikipedia's First Link Network, we use the freely-available XML dump of the English Wikipedia. 
+Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
+eliminating any statistical error due to sampling.
+We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
+The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
+and disambiguations.
+Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
+((cite)), our aim is to characterize the dynamics of the First Link Network, not record a particular link between one
+article and another.
 
-\section{Making a Map: Finding the First Links}
-To map Wikipedia's First Link Network, we use the freely-available xml dump of all 4.7 million English articles \cite{2}.
-We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
-Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new article per day on average \cite{3}, our aim is to characterize the dynamics of the First Link Network, not to record a particular link between one page to another.
+Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
+Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
+information.
+While a human can accurately identify the first link, to map the entire First Link Network of 11 million articles, we programmatically untangled the body text from side-bar, header box, and bad link elements.
 
-The xml dump provided demarkates article elements in in Media Wiki markup, a markup language with special syntax for links. 
-Media Wiki markup also includes templates for audio files, images, side-bar information.
-While a human can accurately idenfity the first link, to map the entire First Link Network of 4.7 million pages, we programmatically untangled the body text from side-bar, header box, and bad link elements.
-
-While some libraries exist for Media Wiki Markup, we opted to develop our own algorithm for parsing the first link in the xml version of each article. 
-\footnote{Approaches using existing libraries led to several bugs including trouble with nested links, nested parenthesis, unclosed tags, escape characters as well as compatibility with other libraries used to parse the xml.}
-Our parsing algorithm aimed to: 1. squash the initial bugs 2. eliminate the need for several passes. To process an article with one pass, we developed a hierachical system of flags:
+While some libraries exist for MediaWiki Markup, we opted to develop our own algorithm for parsing the first link in the XML version of each article. 
+\footnote{Approaches using existing libraries led to several bugs 
+including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
+as well as compatibility with other libraries used to parse the XML.}
+Our parsing algorithm aimed to: 
+1. squash the initial bugs 
+2. eliminate the need for several passes. 
+To process an article with one pass, we developed a hierarchical system of flags:
 \begin{itemize}
-    \item Flag 1: inside template?
+    \item[Flag 1:] inside template?
     \begin{itemize}
-        \item Flag 2: inside <ref>, <div>?
+        \item[Flag 2:] inside <ref>, <div>?
         \begin{itemize}
-            \item Flag 3: inside ()?
+            \item[Flag 3:] inside ()?
         \end{itemize}
@@
 \end{itemize}
+((better graphic to come))
 
-The algorithm loops in three characters chunks, shifting by one character steps through the article markup.
-Proceeding in three character chuncks eliminates any nesting issues.
-If any flag triggers are detected the Flag is raised.
-Once a flag is raised, we stop processing and proceed to the next character.
-A first link is correctly identified only if Flags 1, 2, and 3 are all off.
-Then the entire link is retrieved.
-We then confirm whether the link is legitimate, meaning it does not link to an image, audio file, external page or other Wikimedia project via keyword triggers provided by Wiki Media \cite{4}.
-The first link of an article is then the earliest link in the text to pass all three flags and the false link check.
+The algorithm loops in three-character chunks to account for potentially nested elements, 
+shifting by one character steps through the article markup.
+If any markup triggers for a flag are detected a flag is raised. 
+Once a flag is raised, we stop processing and proceed to the next character
+until the flag's closing markup is detected.
+A first link is identified only if Flags 1, 2, and 3 are all off.
+In this case, the entire link is retrieved. 
+We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
+as well as common file extensions for 
+((cite))
+images, audio files, and the like.
+The first link of an article is then the earliest valid link with unraised flags.
 
-Instead of relying on a sample of articles from which to generalize, we opted to process all 11.blah articles in the xml dump containing the 4.7 million articles along with redirected pages and disambugations, which we filter.
-To process the entirety of Wikipedia, we divided the xml dump into collections of articles, distributing the processing across 112 cores of the UVM sueprcomputer cluster.
-We then obtained a complete map of the First Link Network.
+To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
+across 112 cores of the UVM supercomputer cluster
+((cite))
+We then joined the results to form a hash table containing every Wikipedia article and its corresponding
+first link. The resulting network is the basis of our analysis.
 
-\subsection*{testing the algorithm}
-\section{Using the Map: Navigating the First Links}
+%========================Traversing the Network====================================
+\section{Traversing the Network}
 
-\textcolor{red}{reword the shit below}
-\gray{
-quote analysis from other areas/dodds paper: Droplets accumulating\\
-first paper by Dodds cites Horton-Strahler (leaf pruning) not drop accumulation
-}
+To understand the structure of Wikipedia's First Link Network, we
+aimed to characterize the dynamics of the flow from one article to another. 
+Do links tend towards a particular article, group of articles, or groups of articles? 
+What is the flow of links through a typical article?
+What types of cycle (loop) structures exist in the network?
+What are typical paths from one article to loop or a dead end (invalid link)? 
+((add other questions we end up answering))
+Are there exceptionally long or short paths? 
+Are there articles funneling the flow of links towards a particular path?
+In answering these questions about the directed network, we aimed to characterize 
+how so many independent articles relate to each other.
 
-To understand the structure of Wikipedia's First Link Network, we need to understand how the links flow from one page to another.
-The procedure, inspired by work in networks/enviromental something ((ask dodds)) is to treat each page as a starting point.
+The n-degree, or number of links directly pointing to a particular article,
+while a natural measure, fails to fully capture the dynamics needed to answer these questions.
+The n-degree measures only the particular set of links to a particular article, 
+rather than the richer dynamics of how the links flow through the network: 
+where articles tend, what the typical and atypical 
+paths through the network are, which articles funnel relatively more links and so on.
 
-We follow the path of the first links until we reach either a dead link--a page linking to an invalid link (a file, another wikimedia project, or an external webpage)-- or a loop, where one page in the path is revisted.
+Consequently, we developed three metrics to capture the dynamics of how the First Links flow. 
+First, we traverse the network to measure the accumulated number of visits each article receives as we 
+follow every possible link path: \\
 
-What we can then measure are various metrics including the number of pages whose path leads to a particular page ((confusing wording)), the path length, even loops of a particular size.
-Combining these measures, we can then examine are particular properties of how pages related: for example the most popular two-loops.
+\begin{figure}[H]
+\centering
+    \caption{traversing the network}
+\begin{subfigure}[b]{0.8\textwidth}
+    \includegraphics[width=\textwidth]{graphics/traverse_algo.png}
+\end{subfigure}
+\end{figure}
 
-Thus, from the directed network of First Links, we associate to each node several properties.
-Fortunately, we need not worry about issues of sampling error or statistiacal accuracy, since we do not rely on any sampling.
-Instead, thanks to the VACC we are able to follow every single possible path, tallying the properties of each node in the network.
+((diagram and explain))
 
-Whether philosophy or another page is at the center of Wikipedia's First Link Network is a question about the page with the most traverses.
-More importantly, instead of a mere page, we can precisely identify the exact loop structure (or basin) where most pages feed.
+Second, we traverse the network in the same manner, but end our paths once we enter a cycle (loop in the network).
+We call the accumulated number of visits, the number of traversal funnels, because we quantify
+which articles funnel more link paths in particular direction or cycle.
 
-Not only this, we can answer whether the top loop or page is in fact special--whether the number of pages feeding into the page/loop is measurably different from the other top pages/basins.
+Finally, we can also track the path length, or the number of links until we repeat an article or reach an invalid link.
+Additionally, by recording the history of articles we traverse to map cycles in the network.
+The three metrics: traversal visits, traversals funnels, and path length, along with our path history
+yield a powerful arsenal of information with which to study the network. 
+We can effectively answer questions about how the First Links Flow, identify both peculiar and popular
+articles (and groups of articles), as well as link paths.
 
-((add diagram to explain flow of procedure))
+%========================Results====================================
+\section{Discoveries}
 
-\section{The Stories of our Journey: our results}
+We traversed every possible path through the network, measuring the accumulated number of traversal visits.
+We traversed a total of 232.4 million first links.
 
-\subsection{Clicks (beans...)}
-Distribution is zipf's law:
-((plot))
-8.6 million of the 11 million pages in our raw dataset (11 million includes redirects to the subset of 4.2 million actual pages) have no clicks, with less than 1\% of pages accruing more than 100 clicks (precisely 8.56 million of the 11 have no visits).
+\subsection{Traversal Visits}
 
-((maybe sample and graph log-log to see if it's linear to justify zipf's law))
+As a distribution, the number traversal visits per article appears scale-free. The majority of articles have fewer than 30 traversal visits, while few 
+have 5 order of magnitude more traversal visits. 
+Specifically, $99.$((fill in 99.something)) $\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
+Meanwhile, the highest ranking 30 articles have ((blah of the traversal visits)).
 
-average number of visits is 21.
+((diagram))
 
-Even examining only the 1000 pages with the most page visist, we find a majority have no visit:
+The highest ranking articles include Philosophy alongside related articles such as "Existence", "Quality", and "Reality".
+Other high ranking articles include similarly foundational ideas or abstracted disciplines such as "Science", "Biology", 
+"Community", "State", "Earth", "Human Geography", "Information", and "Communication".
 
-\includegraphics[scale=0.5]{graphics/top_1k_pages.png}
+((diagram: highest ranking articles))
 
-Of the top 100 pages, we see fewer than 30, contain an overwhelming majority of the traversals 6, 7 million:
+The recurrence of an exact number of traversal visits suggests some articles are part of a cycle. 
+The "Philosophy" article for example sits in what seems like a cycle of seven other articles; "Hypothesis" appears to sit in a 
+cycle of 6 other articles including "Experiment", "Fact", and "Knowledge".
 
-\includegraphics[scale=0.5]{graphics/top_100_pages.png}
+((support characterization of high ranking articles as abstract foundational ideas ))\\
+((maybe add synonym analysis here))\\
 
-Maximum number of traversals is 7.4 million, but the mean is a mere 20 visits.
+To confirm the suggested cycle structure, we recorded the history of articles traversed along a path.
 
-The top 7 pages, the ones with teh most visits comprise a loop around philosophy:
+\subsection{Network Cycles}
+By tracking the articles traversed we were able to identify the cycle structures
+by the reappearance of an article in our path history.
+We first identified 2-cycles, meaning a pair of articles with first link pointing to one another.
+Of the 11 million articles, 84 thousand are 2-cycles. 
+The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than different, yet connected ideas:
+"Health Care" and "Medical Case Management", "Broadcasting House" and "BBC", "Secondary Education" and "Secondary School".
 
-\includegraphics[scale=0.5]{graphics/top_pages_list.png}
+((diagram, check why some don't have a matching number of visits))
 
-\subsection{Longest Path}
+Outside of the highest ranking 2-cycles, the typical 2-cycle signals a connection between different, yet very closely related ideas. 
+Link patterns such as inventor to product ("Voere" to "VEC-91"), event to organizer ("Poetry Bus Tour" to "Weave Books"), and book to author ("Anatomy of Britain" to "Anthony Sampson").
 
-The longest first link path span around 365 pages.
-These paths traverse calendar related pages including religious calendars such as Orthodox Liturgics and history pages.
-The Eastern Orthodox Liturgics pages for example are dated.
-Visit any page's Orthodox Liturgics and the first link will send to tomorrow's liturgics.
-The enxt days links to the following day's liturgics.
-Curiously, these are complete loops.
-The last calendar day of the year, simply links back to the first restarting the loop.
-Other notable pages with a path length of 80 include the legislative assembly of Ontario, 1868 in the UK,
+Similarly, 3-cycles captured a synonymous or close relation among 3 articles: "Tree of life (Biology)", "Tree of life (disambiguation)", 
+and "Tree of life"; "Cinema of India", "Indian Cinema", and "Telugu Cinema". Once we extend our cycle size beyond a length of 6 however, 
+"Philosophy" along with the remaining list of high ranking articles by traversal visits dominate.
 
-\subsection{Loops - basics}
-Although the loop of 7 pages around philosophy does attract the greatest number of visits, several other loops around natural fundamental concepts also exist. 
-The top non-philosophy basins include a basin around community, another around landmass.
+((diagram of 3-cycles))\\
+((diagram of cycles beyond length 10))
 
-Identifying basins explains the structure of the network, but it doesn't reveal which page/s are a funnel for the basin.
-By following a similar approach and measuring where the drop accumulation is funneled we find:
+The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day. 
+Curiously, on the last calendar day, the last article simply links back to January 1, forming a 365-cycle.
+Other lengthy cycles span 60-75 articles including collections of articles on national histories such as "Japanese Eras" 
+or judicial bodies such as the "Legislative Assembly of Ontario".
 
-\includegraphics[scale=0.5]{graphics/top_funnels.png}
+\subsection{Basins}
+We can group articles lying on the same path to identify basins of path-connected articles, not necessarily forming a perfectly closed cycle.
+Ranking basins by traversal visits, we find many of highest ranking basins are around "Philosophy" as we might expect. 
+Looking beyond "Philosophy" however, we find high ranking basins around similarly foundational ideas:
+"Community", "Landmass", "Federal Government", "Presentation", and "Belief System". 
+These concepts naturally emerge from the First Link Network potentially indicating pillars, which 
+anchor specific knowledge in a broader, simpler concept.
 
-confirming philosophy is indeed the largest funnel for the philosophy loop, implying it is in fact at the center of wikipedia.
-Nonetheless, curious foundational funnels also surface such as the word presentation, Tree of life (biology), Geographical regions such as Southern Europe, the US Constituion, health care and BBC.
-This amalgamation mixes both foundational topics, confirming teh theory whereby the first link generalizes a specific page by hooking it to a broader concept.
-The list also highlights broader topics that are perhaps less foundational, but seem to be at the forefront of people's minds.
-Health Care while not a foundational discipline such as biology or philosophy ranks at the top perhaps due to the importance we as a society attach to it.
-Similarly BBC, Hip Hop Music surface perhpas out of a cultural ttraction.
+\subsection{Path Length}
 
-\red{how many two loops are there compared to the rest of wikipedia?}
+In addition to identifying cycles and cycle lengths, we also measure the traversal path length, which includes articles outside
+of cycles. 
+Path length measures the number of links traversed until a repeated or invalid link. 
+We discovered the longest path length is also 365, matching the longest cycle of Orthodox Liturgics. 
+We also found similarly lengthy paths following the evolution of a place or topic through time: 
+"1953 in Scotland" or "1569s Architecture", with articles sequentially proceeding by year, decade or era.
 
-In addition to measuring pages that funnel to accumulation, we can also measure true loops instead of just basins. These are pages which sit in a closed loop as opposed ones leading to a dead link.
+Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
+The most common path length is 29, with an interquartile range (26, 30).
+The distribution of path lengths is similarly scale-free with few articles at the extreme of 365 path lengths, while the majority 
+is between 26-30:
 
-Furthermore, since we can rank pages, we can rank loops based on the measure associated with each page. We first examine two loops:
+((diagram of path length distribution))
 
-\includegraphics[scale=0.4]{graphics/top_2loops.png}
+\subsection{Traversal Funnels}
 
-Two loops reveal ideas that are closely linked to each other or nealry synomous such a secondary school the place wher esecondary education takes place.
-A lecture which is a form of a presentation, or fedualism and its examples.
+Measuring only traversal visits however is limiting as it does not distinguish whether a particular article in a cycle 
+is a funnel, directing many more paths inside the cycle than others. 
+To distinguish among articles in a cycle, we also measured traversal funnels, or the number of 
+paths an article directs towards a cycle. Here we count the number of traversal visits up to a cycle, 
+so that accumulation does not flow to all articles in a cycle.
+The importance here is to distinguish an article that happens to be connected to another article with many traversal visits 
+from an article directly funneling many paths.
 
-Investigating the list beyond simply the top pages, but looking at a typical 2-loop we find a more natural link between (where typical means between 5 and 15 visits) linking a person to what they are most known for: say a book and author: Anatomy of Britain links to Anthony Sampson (and the reverse; an inventor and their invention: Voere and VEC-91; and an event links to the creator of the event: Peotry Bus Tour links to Wave Books.
+Measuring traversal funnels reveals a dramatically different structure where "Philosophy" dramatically stands out:
 
-While often a less intimate connection, 3-loops reveal a similar pattern. The top ranking 3-loops are:
+((funnels diagram))
 
-\includegraphics[scale=0.4]{graphics/top_3loops.png}
+Philosophy is not only a stand out with its number of traversal visits, but also by the number articles "Philosophy" funnels into
+its cycle: ((99.something)) $\%$. 
+The second contributor to the cycle is "Reality" with a mere $.2\%$ of accumulated traversals visits in the cycle.
+Even next to the contribution of other articles funneling into other cycles, "Philosophy" is a singularity, dwarfing
+the traversal feeds of "Presentation" the second highest ranking article: 
+a mere $0.4\%$ of the traversal feeds "Philosophy" holds.
+Nevertheless, the other high ranking funnels are remarkably topical, culturally and politically important ideas.
+For example, "Health Care", a recently high-contested legislative topic ((add google results trend)) appears high on the list
+as it does in google search terms or ((media articles?)).
+Other high ranking articles include key historical events such as the "Cold War" or critical scarce resource with recent 
+media discussion such as "Fossil Fuel".
+This coincidence of recent relevance and traversal feed rank suggests the First Link Network measurably represents
+meaningful relationships not only among ideas, but also to society ((english speaking)).
 
-There are 74k true 3loops.
+%========================Conclusion====================================
+\section{Reflections}
 
-\subsection{Later: Visualize loops}
+The findings here should only be considered within the limitations of their context.
+We examined only the English version of Wikipedia at a particular moment in time.
+Furthermore, we only studied the first link in the main body of each article
+as a means to related one article to another. Finally, Wikipedia, while the largest 
+collection of human knowledge, is rife with the biases of the many contributing editors--
+male middle aged ((substantiate and confirm)). Nevertheless, the findings do reveal
+generalizable relationships, point to foundational notions, and uncover many curiosities.
 
-\href{http://projects.flowingdata.com/tut/interactive_network_demo/ }{D3 Example 1}
+Among the curiosities is the multiple appearance of scale-free distributions within the network. 
+The three metrics we developed: path length, traversal visits, and traversal funnels are all marked 
+by scale-free distributions. Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
+articles are responsible for funneling most paths. When measured against the traversal funnels, 
+"Philosophy" emerges as an exceptional article by orders of magnitude. 
+Nevertheless, many other foundational ideas emerged naturally within the First Link Network. 
+Basins around "Community", "State", and "Science" reveal a foundational structure within the network. 
+More curious is the emergence of recently prominent political and economics topics such as "Fossil Fuel" and "Health Care" 
+within the highest ranking funnels. 
+Wikipedia seems to reflect not only timeless foundations, but also the topical (at least within English speaking society).
 
-\href{http://christophergandrud.github.io/d3Network/}{simple D3 Network}
+These findings form the basis for future work towards the creation of a taxonomy where 
+every idea, event, or object sits within a hierarchy of connected notions.
+The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
+Applications could range from an enhanced thesaurus of ideas to psychological insights into how human form associations.
+Specifically, an ever-evolving reference of related hierarchical concepts can be applied to search engine algorithms 
+or natural language processing.
 
-\section{Conclusion}
+%========================Style Guide====================================
+\newpage
+\section*{Style Guide}
 
-\gray{Next: word stem to see how broader ideas are related (taxonomy of ideas).
-    * Broader: what if we consider all links (or first par)? other languages? 
-        * popular page by links to it. }
+\begin{itemize}
+\item Web Terms
+    \begin{itemize}
+        \item \bt{article} instead of page
+        \item \bt{invalid link} (instead of dead or bad link)
+    \end{itemize}
+\item Measurements
+    \begin{itemize}
+        \item \bt{traversals visits} = node weight based on number of paths crossing an article.
+        \item \bt{path length} = number of articles traversed until an invalid link or a repeated article
+        \item \bt{traversal funnels} = traversals before a cycle is reached.
+    \end{itemize}
+\item Network Properties
+    \begin{itemize}
+        \item \bt{cycle} = permutation or a true loop
+        \item \bt{path-connected articles} for both cycles and articles along the same path leading to an 
+    invalid link
+        \item \bt{basin} a path-connected group of articles (not necessarily forming a cycle)
+    \end{itemize}
+\end{itemize}
 
-\gray{* Is the network stable over time? 
-        * analyze a few newer datasets. }
+- add quotes around article names: Philosophy should be "Philosophy"
 
-The analysis remarkably demonstrates the links in Wikipedia truly exihibit a power-law like distribution where most pages link up to a handful of pages.
-Furthermore, Philosophy does occupy not only a special place inside the most popular basin of attraction.
-Not only that, the philosophy page itself is the page through which the feeds flow.
-This implies there is something special even from a mathematical standpoint.
-Philosophy isn't just a popular link, it lies in an exceptionally high percentage, unique to its loop.
-
-Aside from the philosophy loop, other foundational concepts also emerge as the centers for basins of attraction.
-Fundamental concepts such as community, government, or philosophy.
-The network also has a prelevance of 2-loops ((find exact proportion or number)) which link two closely associated concepts such as an inventor or the invention as well as tie synonyms together.
-Other artifacts such as tremendously long loops based around calendar days demonstrate how we group ideas such as holidays, hisotiric events, or religious, not by their ties to other ideas, but first by where they fall on calendar days.
-
-Based on the flow from specific pages to more general concepts, the results also agree with the heuristic provided here.
-Whereby, we link a very specific page to a broader concept, then to an ever broader concept.
-For example, the Wikipedia page for Banana links to fruit, then botanical then, ((confirm, fill-in)) then biology, then science and eventually philosophy.
-It is curious that Wikipedia, the largest encyclopedia, which by its very meaning aims to capture information about all branches of human knowledge, has naturally evolved through the works of ((x million editors/edits)) so that Knowledge, philsophy are the center links, occupying such a special place.
-
-\end{multicols}
-
-\subsection{Terms (aka style guide)}
-
-use article instead of page
-
-traversal = node weight
-path length = degree of separation from a cluster
-path is the flow from one page until a dead link or a repeated article
-
-cycle (permutation from abstract algebra) instead of loop
-
-basin of attraction or cluster? as opposed to cycle
-
-invalid link (instead of dead or bad link)
-
-To Do:
-\textcolor{red}{
-\begin{enumerate}
-    \item Add limitations of work: snap shot in time, parsing is imperfect
-    \item Add better graphics
-    \item post code and results to github (add link to paper)
-\end{enumerate}}
-
-\newpage
-
-\begin{thebibliography}{10}
-\bibitem{1}
-\href{ http://www.technologyreview.com/featuredstory/520446/the-decline-of-wikipedia/}{"The Decline of Wikipedia", MIT Tech Review}
-
-\bibitem{2}
-\href{https://en.wikipedia.org/wiki/Wikipedia:Database_download}{Wikipedia Database Download}
-
-\bibitem{3}
-\href{https://en.wikipedia.org/wiki/Wikipedia:Statistics}{Wikipedia Statistics}
-
-\bibitem{4}
-\href{http://en.wikipedia.org/wiki/Help:Special_page}{Media Markup Special Pages}
-\end{thebibliography}
-
+- add thank you to RJ for pointing out the Blog Post
 \end{document}

### 2015-10-04 writeup/rough.tex
rough draft

\documentclass[twoside]{article}

% ------
% Fonts and typesetting settings
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\linespread{1.05} % Palatino needs more space between lines
\usepackage{microtype}

% ------
% Page layout
\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
\usepackage[font=it]{caption}
\usepackage{paralist}
\usepackage{multicol, hyperref, graphicx}

% ------
% Lettrines
\usepackage{lettrine}
\usepackage{color}

% ------
% Abstract
\usepackage{abstract}
	\renewcommand{\abstractnamefont}{\normalfont\bfseries}
	\renewcommand{\abstracttextfont}{\normalfont\small\itshape}

% ------
% Titling (section/subsection)
\usepackage{titlesec}
\renewcommand\thesection{\Roman{section}}
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{}

% ------
% Header/footer
\usepackage{fancyhdr}
	\pagestyle{fancy}
	\fancyhead{}
	\fancyfoot{}
	\fancyhead[C]{ $\bullet$ Draft $\bullet$}
	\fancyfoot[RO,LE]{\thepage}

% My Shortcuts

%useful shortcuts
\def\R{\ensuremath{\mathbb{R}}} %\ensuremath adds math mode, if forgotten
\def\Q{\ensuremath{\mathbb{Q}}}
\def\N{\ensuremath{\mathbb{N}}}
\def\Z{\ensuremath{\mathbb{Z}}}
\def\C{\ensuremath{\mathbb{C}}}

%shorcuts with arguments
\newcommand{\abs}[1]{\left\vert#1\right\vert} %nice absolute values
\newcommand{\bt}[1]{\textbf{#1}} %bold
\newcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
\newcommand{\cb}[1]{\centerline{\fbox{#1}}} %centered box
\newcommand{\bp}[1]{\fbox{\parbox{0.8\textwidth}{#1}}} %box paragraph
\newcommand{\norm}[1]{\left\lVert#1\right\rVert} %vector norm
\newcommand{\notimplies}{% does not imply
  \mathrel{{\ooalign{\hidewidth$\not\phantom{=}$\hidewidth\cr$\implies$}}}}
\renewcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations

%colors
\definecolor{javagreen}{rgb}{0.25,0.5,0.35} %dark green color
\definecolor{lightblue}{rgb}{0.149,0.545,0.824} %solarized blue
\definecolor{sred}{rgb}{0.863, 0.196, 0.184} %solarized red

\newcommand{\blue}[1]{{\leavevmode\color{lightblue}{#1}}} %solarized blue 
\newcommand{\green}[1]{{\leavevmode\color{javagreen}{#1}}} %command for green
\newcommand{\red}[1]{{\leavevmode\color{sred}{#1}}} %solarized red
\newcommand{\gray}[1]{{\leavevmode\color[gray]{0.5}{#1}}} %gray text

%environment
\newcommand{\tab}{\phantom{ssss}}
%----------------

% ------
% Clickable URLs (optional)
\usepackage{hyperref}

% ------
% Maketitle metadata
\title{\vspace{-5mm}%
	\fontsize{24pt}{12pt}\selectfont
	\textbf{What's So Special About Philosophy?} 
	}	
\author{%
\fontsize{14pt}{14pt}\selectfont
	Unraveling Wikipedia's First Link Network \vspace{-2mm}\\
	}
\date{}

%figures
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\maketitle
\thispagestyle{fancy}
%========================ABSTRACT====================================
\begin{abstract}
\fontsize{12pt}{12pt}
\selectfont
\noindent Apples, oranges, and the most obscure Dylan song too---is any article a few clicks from Philosophy? The surprising answer is yes, $95\%$ of the time (according to a popular blog post).

Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. Yet, we have never closely examined the curious web tying one entry to another. By following the first link in an article, we connect entries to form a directed network within Wikipedia: Wikpedia's First Link Network. Our aim is to study Wikipedia's First Link Network for insight into the topics, ideas, people, objects, and events we link. ((improve this sentence))

We first map the First Link Network by algorithmically parsing the first link. Then, apply a river network model to identify the basins of attraction, loop structure, and characterize the many paths from one page's first link to the next. ((wording is unclear: as if there are many paths from one page))

\red{rewrite abstract}

\end{abstract}

%==================================BASE====================================================

\begin{multicols}{2}

\lettrine[nindent=0em,lines=2]{A}t no point in history has a larger or more meticulously indexed collection of human knowledge existed \cite{1}. In amassing such an awe-inspiring collection, we naturally linked topics, inventions, people, objects, places, and events 
\footnote{while anyone with internet access is free to contribute "neutral, cited information," it's worth noting the population of contributors is not necessarily representative of the overall english speaking population. Notably, the contributors are overwhelmingly male ((cite, expand)).} across space and time. The web we weaved, and continue to weave, is a wealth of information about those notable inventions, places, figures, and ideas, but also about us: the ones weaving the web. Here we explore this web as a directed network formed by the first link connecting one article to another.

We build Wikipedia's First link network by following the first link, not in parenthesis, inside the main body of each article of the English version of Wikipedia. For the directed network to meaningfully reflect how we associate one article to another, we exclude links in parenthesis containing pronunciation keys or disambiguations. We also exclude any links in the side-bar elements, as well as any links to external pages, files, or other WikiMedia projects outstide of Wikipedia such as Wiktionary ((lookup others))
\footnote{
These constraints are the same which led to the popular blog post/s claiming nearly every article is within fewer than 20 first links away from philosophy. 
}
thus, forming a true network of Wikipedia.

\includegraphics[scale=0.18]{graphics/wiki_train.png}
\ \\

\section{Making a Map: Finding the First Links}
To map Wikipedia's First Link Network, we use the freely-available xml dump of all 4.7 million English articles \cite{2}. We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time. Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new article per day on average \cite{3}, our aim is to characterize the dynamics of the First Link Network, not to record a particular link between one page to another.

The xml dump provided demarkates article elements in in Media Wiki markup, a markup language with special syntax for links. 
Media Wiki markup also includes templates for audio files, images, side-bar information.
While a human can accurately idenfity the first link, to map the entire First Link Network of 4.7 million pages, we programmatically untangled the body text from side-bar, header box, and bad link elements.

While some libraries exist for Media Wiki Markup, we opted to develop our own algorithm for parsing the first link in the xml version of each article. 
\footnote{Approaches using existing libraries led to several bugs including trouble with nested links, nested parenthesis, unclosed tags, escape characters as well as compatibility with other libraries used to parse the xml.}
Our parsing algorithm aimed to: 1. squash the initial bugs 2. eliminate the need for several passes. To process an article with one pass, we developed a hierachical system of flags:
\begin{itemize}
    \item Flag 1: inside template?
    \begin{itemize}
        \item Flag 2: inside <ref>, <div>?
        \begin{itemize}
            \item Flag 3: inside ()?
        \end{itemize}
    \end{itemize}
\end{itemize}

The algorithm loops in three characters chunks, shifting by one character steps through the article markup. Proceeding in three character chuncks eliminates any nesting issues. If any flag triggers are detected the Flag is raised. Once a flag is raised, we stop processing and proceed to the next character. A first link is correctly identified only if Flags 1, 2, and 3 are all off. Then the entire link is retrieved. We then confirm whether the link is legitimate, meaning it does not link to an image, audio file, external page or other Wikimedia project via keyword triggers provided by Wiki Media \cite{4}. The first link of an article is then the earliest link in the text to pass all three flags and the false link check.

Instead of relying on a sample of articles from which to generalize, we opted to process all 11.blah articles in the xml dump containing the 4.7 million articles along with redirected pages and disambugations, which we filter. To process the entirety of Wikipedia, we divided the xml dump into collections of articles, distributing the processing across 112 cores of the UVM sueprcomputer cluster. We then obtained a complete map of the First Link Network.

\subsection*{testing the algorithm}
\section{Using the Map: Navigating the First Links}

\textcolor{red}{reword the shit below}
\gray{
quote analysis from other areas/dodds paper: Droplets accumulating\\
first paper by Dodds cites Horton-Strahler (leaf pruning) not drop accumulation
}

To understand the structure of Wikipedia's First Link Network, we need to understand how the links flow from one page to another. The procedure, inspired by work in networks/enviromental something ((ask dodds)) is to treat each page as a starting point.

We follow the path of the first links until we reach either a dead link--a page linking to an invalid link (a file, another wikimedia project, or an external webpage)-- or a loop, where one page in the path is revisted.

What we can then measure are various metrics including the number of pages whose path leads to a particular page ((confusing wording)), the path length, even loops of a particular size. Combining these measures, we can then examine are particular properties of how pages related: for example the most popular two-loops.

Thus, from the directed network of First Links, we associate to each node several properties. Fortunately, we need not worry about issues of sampling error or statistiacal accuracy, since we do not rely on any sampling. Instead, thanks to the VACC we are able to follow every single possible path, tallying the properties of each node in the network.

Whether philosophy or another page is at the center of Wikipedia's First Link Network is a question about the page with the most traverses. More importantly, instead of a mere page, we can precisely identify the exact loop structure (or basin) where most pages feed.

Not only this, we can answer whether the top loop or page is in fact special--whether the number of pages feeding into the page/loop is measurably different from the other top pages/basins.

((add diagram to explain flow of procedure))

\section{The Stories of our Journey: our results}

\subsection{Clicks (beans...)}
Distribution is zipf's law:
((plot))
8.6 million of the 11 million pages in our raw dataset (11 million includes redirects to the subset of 4.2 million actual pages) have no clicks, with less than 1\% of pages accruing more than 100 clicks (precisely 8.56 million of the 11 have no visits).

((maybe sample and graph log-log to see if it's linear to justify zipf's law))

average number of visits is 21.

Even examining only the 1000 pages with the most page visist, we find a majority have no visit:

\includegraphics[scale=0.5]{graphics/top_1k_pages.png}

Of the top 100 pages, we see fewer than 30, contain an overwhelming majority of the traversals 6, 7 million:

\includegraphics[scale=0.5]{graphics/top_100_pages.png}

Maximum number of traversals is 7.4 million, but the mean is a mere 20 visits.

The top 7 pages, the ones with teh most visits comprise a loop around philosophy:

\includegraphics[scale=0.5]{graphics/top_pages_list.png}

\subsection{Longest Path}

The longest first link path span around 365 pages. These paths traverse calendar related pages including religious calendars such as Orthodox Liturgics and history pages. The Eastern Orthodox Liturgics pages for example are dated. Visit any page's Orthodox Liturgics and the first link will send to tomorrow's liturgics. The enxt days links to the following day's liturgics. Curiously, these are complete loops. The last calendar day of the year, simply links back to the first restarting the loop. Other notable pages with a path length of 80 include the legislative assembly of Ontario, 1868 in the UK,

\subsection{Loops - basics}
Although the loop of 7 pages around philosophy does attract the greatest number of visits, several other loops around natural fundamental concepts also exist. 
The top non-philosophy basins include a basin around community, another around landmass.

Identifying basins explains the structure of the network, but it doesn't reveal which page/s are a funnel for the basin. By following a similar approach and measuring where the drop accumulation is funneled we find:

\includegraphics[scale=0.5]{graphics/top_funnels.png}

confirming philosophy is indeed the largest funnel for the philosophy loop, implying it is in fact at the center of wikipedia. Nonetheless, curious foundational funnels also surface such as the word presentation, Tree of life (biology), Geographical regions such as Southern Europe, the US Constituion, health care and BBC. 
This amalgamation mixes both foundational topics, confirming teh theory whereby the first link generalizes a specific page by hooking it to a broader concept. The list also highlights broader topics that are perhaps less foundational, but seem to be at the forefront of people's minds. Health Care while not a foundational discipline such as biology or philosophy ranks at the top perhaps due to the importance we as a society attach to it. Similarly BBC, Hip Hop Music surface perhpas out of a cultural ttraction.

\red{how many two loops are there compared to the rest of wikipedia?}

In addition to measuring pages that funnel to accumulation, we can also measure true loops instead of just basins. These are pages which sit in a closed loop as opposed ones leading to a dead link.

Furthermore, since we can rank pages, we can rank loops based on the measure associated with each page. We first examine two loops:

\includegraphics[scale=0.4]{graphics/top_2loops.png}

Two loops reveal ideas that are closely linked to each other or nealry synomous such a secondary school the place wher esecondary education takes place. A lecture which is a form of a presentation, or fedualism and its examples.

Investigating the list beyond simply the top pages, but looking at a typical 2-loop we find a more natural link between (where typical means between 5 and 15 visits) linking a person to what they are most known for: say a book and author: Anatomy of Britain links to Anthony Sampson (and the reverse; an inventor and their invention: Voere and VEC-91; and an event links to the creator of the event: Peotry Bus Tour links to Wave Books.

While often a less intimate connection, 3-loops reveal a similar pattern. The top ranking 3-loops are:

\includegraphics[scale=0.4]{graphics/top_3loops.png}

There are 74k true 3loops.

\subsection{Later: Visualize loops}

\href{http://projects.flowingdata.com/tut/interactive_network_demo/ }{D3 Example 1}

\href{http://christophergandrud.github.io/d3Network/}{simple D3 Network}

\section{Conclusion}

\gray{Next: word stem to see how broader ideas are related (taxonomy of ideas).
    * Broader: what if we consider all links (or first par)? other languages? 
        * popular page by links to it. }

\gray{* Is the network stable over time? 
        * analyze a few newer datasets. }

The analysis remarkably demonstrates the links in Wikipedia truly exihibit a power-law like distribution where most pages link up to a handful of pages. Furthermore, Philosophy does occupy not only a special place inside the most popular basin of attraction. Not only that, the philosophy page itself is the page through which the feeds flow. This implies there is something special even from a mathematical standpoint. Philosophy isn't just a popular link, it lies in an exceptionally high percentage, unique to its loop.

Aside from the philosophy loop, other foundational concepts also emerge as the centers for basins of attraction. Fundamental concepts such as community, government, or philosophy. The network also has a prelevance of 2-loops ((find exact proportion or number)) which link two closely associated concepts such as an inventor or the invention as well as tie synonyms together. Other artifacts such as tremendously long loops based around calendar days demonstrate how we group ideas such as holidays, hisotiric events, or religious, not by their ties to other ideas, but first by where they fall on calendar days.

Based on the flow from specific pages to more general concepts, the results also agree with the heuristic provided here. Whereby, we link a very specific page to a broader concept, then to an ever broader concept. For example, the Wikipedia page for Banana links to fruit, then botanical then, ((confirm, fill-in)) then biology, then science and eventually philosophy. It is curious that Wikipedia, the largest encyclopedia, which by its very meaning aims to capture information about all branches of human knowledge, has naturally evolved through the works of ((x million editors/edits)) so that Knowledge, philsophy are the center links, occupying such a special place.

\end{multicols}

\subsection{Terms (aka style guide)}

use article instead of page

traversal = node weight
path length = degree of separation from a cluster
path is the flow from one page until a dead link or a repeated article

cycle (permutation from abstract algebra) instead of loop

basin of attraction or cluster? as opposed to cycle

invalid link (instead of dead or bad link)

To Do:
\textcolor{red}{
\begin{enumerate}
    \item Add limitations of work: snap shot in time, parsing is imperfect
    \item Add better graphics
    \item post code and results to github (add link to paper)
\end{enumerate}}

\newpage

\begin{thebibliography}{10}
\bibitem{1}
\href{ http://www.technologyreview.com/featuredstory/520446/the-decline-of-wikipedia/}{"The Decline of Wikipedia", MIT Tech Review}

\bibitem{2}
\href{https://en.wikipedia.org/wiki/Wikipedia:Database_download}{Wikipedia Database Download}

\bibitem{3}
\href{https://en.wikipedia.org/wiki/Wikipedia:Statistics}{Wikipedia Statistics}

\bibitem{4}
\href{http://en.wikipedia.org/wiki/Help:Special_page}{Media Markup Special Pages}
\end{thebibliography}

\end{document}

### 2015-10-06 writeup/draft.tex
first draft

@@
 \usepackage[font=it]{caption}
-\usepackage{paralist}
-\usepackage{multicol, hyperref, graphicx}
+\usepackage{multicol, hyperref}
 \usepackage{parskip}
@@
 
-Apples, oranges, and the most obscure Dylan song too---is everything a few clicks from Philosophy? In Wikipedia, the surprising answer is yes.
+Apples, oranges, and the most obscure Dylan song too---is everything a few clicks from Philosophy? 
+Within Wikipedia, the surprising answer is yes: nearly all 
+paths\footnote{
+nearly $99.6\%$ of first link paths, ending inside Wikipedia, lead to the Philosophy article (see traversal funnels discussion for details).}
+\footnote{
+An analysis performed in 2011 by Mat Kelcey posted on his blog (and a smiliar post by Ilmari Karonen) suggest $94.5\%$ of articles end up at Philosophy.
+}
+lead to Philosophy.
+Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
+\footnote{
+cite MIT Tech Review Article: http://www.technologyreview.com/featuredstory/520446/the-decline-of-wikipedia/
+}
+More than information about a topic though, Wikipedia is a marvelous web of naturally emerging relationships.  
+By following the first link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
+Here we study Wikipedia's First Link Network for insight into how we related topics, ideas, people, objects, and events.
 
-Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed.
-Yet, we have never closely examined the naturally formed web tying one entry to another.
-By following the first link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network.
-Here we study Wikipedia's First Link Network for insight into how we link the largest collection of topics, ideas, people, objects, and events.
-Structurally, we discover Wikipedia's First Link Network is by many metrics a scale-free network with Philosophy at a salient center by many orders of magnitude.
-We also learn the network generalizes specific events, objects, or phenomena to broader basins around Community, State, and Science.
+We algorithmically parse every article on the English Wikipedia to construct a map of the First Link Network. 
+We then traverse the network, following every possible path to uncover the network's structure.
+We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general with 
+basins around fundamental notions such as Community, State, and Science. 
+Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
+These findings enrich our view of how we connect and structure
+an ever growing load of information.
 
-\green{
-conclusions to mention
-\begin{itemize}
-    \item finding many small-world phenomenan in the number of first links flowing through a page, the path lenghts, and ...\\
-    \item also cementing Philosophy as an extremely peculiar article, orders of magnitude away from other article in Wikipedia by our metrics.
-    \item links generalize specific event, object, or phenomemon to a broader concept. Mention synonyms
-\end{itemize}
-}
 \end{abstract}
@@
 At no point in history has a larger or more meticulously indexed collection of human knowledge existed.
-((cite))
 In amassing such an awe-inspiring collection, we formed an equally impressive web. 
@@
 projects outside of Wikipedia (such as Wiktionary).
-((footnote: this also stays true to the original blog post))
+This procedure corresponds to the original claim in 2008 about the percentage of pages with first links to 
+"Philosophy"\footnote{
+https://en.wikipedia.org/wiki/Wikipedia:GettingtoPhilosophy
+}
+\footnote{
+http://xkcd.com/903/
+}
+\footnote{
+https://www.reddit.com/r/pics/comments/gpdhb/trythiswikipediamindfk/
+}
+
 The result is a directed network placing every article in a broader web of ideas.
 
+\begin{figure}[H]
+\centering
+\caption{Wikipedia Train}
+    \begin{subfigure}[b]{0.8\textwidth}
+        \includegraphics[width=\textwidth]{graphics/wiki_train.png}    
+    \end{subfigure}
+\end{figure}
 %===========================Data_and_Parsing================================
@@
 information.
-While a human can accurately identify the first link, to map the entire First Link Network of 11 million articles, we programmatically untangled the body text from side-bar, header box, and bad link elements.
+While a human can accurately identify the first link, to map the entire First Link Network of 11 million articles, we programmatically untangled the body text from side-bar, header box, and invalid link elements.
 
-While some libraries exist for MediaWiki Markup, we opted to develop our own algorithm for parsing the first link in the XML version of each article. 
-\footnote{Approaches using existing libraries led to several bugs 
+While some libraries exist for MediaWiki Markup, we opted to develop our own algorithm for parsing the first link in the XML version of each article.\footnote{
+Approaches using existing libraries led to several bugs 
 including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
@@
 \end{itemize}
-((better graphic to come))
+
+\red{[[better graphic to come]]}
 
@@
 shifting by one character steps through the article markup.
-If any markup triggers for a flag are detected a flag is raised. 
+If any markup triggers for a flag are detected, a flag is raised. 
 Once a flag is raised, we stop processing and proceed to the next character
-until the flag's closing markup is detected.
+until the flag's closing markup.
 A first link is identified only if Flags 1, 2, and 3 are all off.
@@
 We then joined the results to form a hash table containing every Wikipedia article and its corresponding
-first link. The resulting network is the basis of our analysis.
+first link. The resulting network map is the basis of our analysis.
 
@@
 aimed to characterize the dynamics of the flow from one article to another. 
-Do links tend towards a particular article, group of articles, or groups of articles? 
+Do links tend towards a particular article, group of articles, or various groups of articles? 
 What is the flow of links through a typical article?
@@
 What are typical paths from one article to loop or a dead end (invalid link)? 
-((add other questions we end up answering))
 Are there exceptionally long or short paths? 
@@
 Consequently, we developed three metrics to capture the dynamics of how the First Links flow. 
-First, we traverse the network to measure the accumulated number of visits each article receives as we 
-follow every possible link path: \\
+First, we traverse the network to measure the accumulated number of visits each article receives: \\
 
@@
 
-((diagram and explain))
+\red{[[explain with better diagram]]}
 
-Second, we traverse the network in the same manner, but end our paths once we enter a cycle (loop in the network).
-We call the accumulated number of visits, the number of traversal funnels, because we quantify
-which articles funnel more link paths in particular direction or cycle.
+\begin{figure}[H]
+\centering
+    \caption{traverse procedure}
+\begin{subfigure}[b]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{graphics/traverse_procedure.png}
+\end{subfigure}
+\end{figure}
 
-Finally, we can also track the path length, or the number of links until we repeat an article or reach an invalid link.
-Additionally, by recording the history of articles we traverse to map cycles in the network.
+We also stop if the link is invalid. 
+By repeating the procedure starting at each article, we traverse
+every possible first link path through the network.
+The method is order agnostic with respect to the order of articles selected.
+As long as each article is selected as a starting point, the resulting set of measured traversal visits is equivalent. 
+More important, these traversals visits measure the accumulated flow of first 
+links to an article, preserving the link relations through the network.
+We call this value the number of traversal visits for a given article.
+
+\red{[[expand, explain better]]}
+
+Precedence for the traversal approach comes from methods used in \red{[[river networks? leaf pruning approach in environmental studies? what to cite?]]}
+
+Second, we traverse the network in the same manner, but end our paths once we enter a cycle.
+We call the accumulated number of articles funneled, the number of traversal funnels, because we quantify
+which articles funnel more link paths in a particular direction or cycle.
+
+Finally, we also track the path length, or the number of links until a repeated article or invalid link.
+Additionally, by recording the history of articles we traversed, we were able to map cycles in the network.
 The three metrics: traversal visits, traversals funnels, and path length, along with our path history
@@
 
-We traversed every possible path through the network, measuring the accumulated number of traversal visits.
-We traversed a total of 232.4 million first links.
+We followed every possible path through the network, taking 232 million steps along the way to measure the accumulated number of traversal visits.
 
@@
 have 5 order of magnitude more traversal visits. 
-Specifically, $99.$((fill in 99.something)) $\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
-Meanwhile, the highest ranking 30 articles have ((blah of the traversal visits)).
+Specifically, $99.76\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
+Meanwhile, the highest ranking 30 articles have an extremely disproportionate number of traversal visits.
 
-((diagram))
+\begin{figure}[H]
+\centering
+    \caption{distribution of traversal visits}
+    \begin{subfigure}[b]{\textwidth}
+        \includegraphics[width=\textwidth]{graphics/top_1k_100_pages.png}
+    \end{subfigure}
+\end{figure}
+
+\red{[[change title from "pages" to "articles" for consistency]]} \\
+\red{[[justify scale-free with log-log or power measure]]} \\
 
@@
 
-((diagram: highest ranking articles))
+\begin{figure}[H]
+\centering
+    \caption{highest ranking articles by number of traversals}
+    \begin{subfigure}[b]{0.8\textwidth}
+        \includegraphics[width=\textwidth, angle=-90]{graphics/top_articles_list.png}
+    \end{subfigure}
+\end{figure}
 
@@
 cycle of 6 other articles including "Experiment", "Fact", and "Knowledge".
+To confirm the suggested cycle structure, we recorded the history of articles traversed along a path.
 
-((support characterization of high ranking articles as abstract foundational ideas ))\\
-((maybe add synonym analysis here))\\
-
-To confirm the suggested cycle structure, we recorded the history of articles traversed along a path.
+\subsubsection{Measure Number of Synonyms}
+\red{
+[[ add synonym analysis to support characterization of high ranking articles
+as abstract foundational notions. \\
+- idea is to correlate the number of synonyms a word has with its traversal visit rank.
+]]}
 
@@
 
-((diagram, check why some don't have a matching number of visits))
+\begin{figure}[H]
+\centering
+\caption{highest ranking 2-Cycles}
+    \begin{subfigure}[b]{0.8\textwidth}
+        \includegraphics[width=\textwidth]{graphics/top_2loops.png}
+    \end{subfigure}
+\end{figure}
+
+\red{[[better representation to come\\
+look into articles that don't seem to have a partner]]}
 
@@
 
-((diagram of 3-cycles))\\
+\begin{figure}[H]
+\centering
+\caption{highest ranking 3-Cycles}
+    \begin{subfigure}[b]{0.8\textwidth}
+        \includegraphics[width=\textwidth]{graphics/top_3loops.png}
+    \end{subfigure}
+\end{figure}
 ((diagram of cycles beyond length 10))
@@
 We also found similarly lengthy paths following the evolution of a place or topic through time: 
-"1953 in Scotland" or "1569s Architecture", with articles sequentially proceeding by year, decade or era.
+"1953 in Scotland" or "1560s Architecture", with articles sequentially proceeding by year, decade or era.
 
@@
 
-((diagram of path length distribution))
+\begin{figure}[H]
+\centering
+    \caption{Longest Path Lengths}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{graphics/top_1k_path_length.png}
+    \end{subfigure}
+\end{figure}
 
@@
 
-Measuring traversal funnels reveals a dramatically different structure where "Philosophy" dramatically stands out:
+Measuring traversal funnels reveals a dramatically different structure where "Philosophy" stands unmatched by orders of magnitude:
 
-((funnels diagram))
+\begin{figure}[H]
+\centering
+    \caption{Top Funnels}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{graphics/top_funnels.png}
+    \end{subfigure}
+\end{figure}
 
-Philosophy is not only a stand out with its number of traversal visits, but also by the number articles "Philosophy" funnels into
-its cycle: ((99.something)) $\%$. 
-The second contributor to the cycle is "Reality" with a mere $.2\%$ of accumulated traversals visits in the cycle.
-Even next to the contribution of other articles funneling into other cycles, "Philosophy" is a singularity, dwarfing
-the traversal feeds of "Presentation" the second highest ranking article: 
-a mere $0.4\%$ of the traversal feeds "Philosophy" holds.
+Philosophy is not only a stand out the number of traversal visits, but also by the number articles "Philosophy" funnels into
+its cycle: \red{[[99.something]]} $\%$. 
+The second contributor to the Philosophy cycle is "Reality", funneling in a mere $.2\%$ of traversals visits in the cycle.
+Even next to the contribution of the largest funnels, "Philosophy" is a singularity. 
+In proportion, "Presentation" holds only $0.4\%$ next to the number of traversal feeds for "Philosophy".
 Nevertheless, the other high ranking funnels are remarkably topical, culturally and politically important ideas.
-For example, "Health Care", a recently high-contested legislative topic ((add google results trend)) appears high on the list
+For example, "Health Care", a recently high-contested legislative topic \red{[[add google results trend]]} appears high on the list
 as it does in google search terms or ((media articles?)).
@@
 collection of human knowledge, is rife with the biases of the many contributing editors--
-male middle aged ((substantiate and confirm)). Nevertheless, the findings do reveal
+male middle aged \red{[[substantiate and confirm]]}. Nevertheless, the findings do reveal
 generalizable relationships, point to foundational notions, and uncover many curiosities.
@@
 
-These findings form the basis for future work towards the creation of a taxonomy where 
+Future work could analyze other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network beyond the first link to a subset or potentially all links.
+These findings also form the basis for the creation of a taxonomy where 
 every idea, event, or object sits within a hierarchy of connected notions.
 The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
-Applications could range from an enhanced thesaurus of ideas to psychological insights into how human form associations.
+Applications could range from an enhanced thesaurus of ideas to psychological insights into how humans form associations.
 Specifically, an ever-evolving reference of related hierarchical concepts can be applied to search engine algorithms 
@@
 
+- thank you to my friend RJ for pointing out the xkcdc comic about links to Philosophy
+
 %========================Style Guide====================================
-\newpage
-\section*{Style Guide}
 
-\begin{itemize}
-\item Web Terms
-    \begin{itemize}
-        \item \bt{article} instead of page
-        \item \bt{invalid link} (instead of dead or bad link)
-    \end{itemize}
-\item Measurements
-    \begin{itemize}
-        \item \bt{traversals visits} = node weight based on number of paths crossing an article.
-        \item \bt{path length} = number of articles traversed until an invalid link or a repeated article
-        \item \bt{traversal funnels} = traversals before a cycle is reached.
-    \end{itemize}
-\item Network Properties
-    \begin{itemize}
-        \item \bt{cycle} = permutation or a true loop
-        \item \bt{path-connected articles} for both cycles and articles along the same path leading to an 
-    invalid link
-        \item \bt{basin} a path-connected group of articles (not necessarily forming a cycle)
-    \end{itemize}
-\end{itemize}
+%\newpage
+%
+%\section*{Style Guide}
+%
+%\begin{itemize}
+%\item Web Terms
+%    \begin{itemize}
+%        \item \bt{article} instead of page
+%        \item \bt{invalid link} (instead of dead or bad link)
+%    \end{itemize}
+%\item Measurements
+%    \begin{itemize}
+%        \item \bt{traversals visits} = node weight based on number of paths crossing an article.
+%        \item \bt{path length} = number of articles traversed until an invalid link or a repeated article
+%        \item \bt{traversal funnels} = traversals before a cycle is reached.
+%    \end{itemize}
+%\item Network Properties
+%    \begin{itemize}
+%        \item \bt{cycle} = permutation or a true loop
+%        \item \bt{path-connected articles} for both cycles and articles along the same path leading to an 
+%    invalid link
+%        \item \bt{basin} a path-connected group of articles (not necessarily forming a cycle)
+%    \end{itemize}
+%\end{itemize}
+%
+%- add quotes around article names: Philosophy should be "Philosophy"
+%
+%- add thank you to RJ for pointing out the Blog Post
 
-- add quotes around article names: Philosophy should be "Philosophy"
-
-- add thank you to RJ for pointing out the Blog Post
 \end{document}

### 2015-10-17 writeup/draft.tex
created traversal figures

@@
 \begin{itemize}
-    \item[Flag 1:] inside template?
+    \item[Flag 1:] inside Wikimedia template?\\
+        * \bt{trigger}: \{\{
     \begin{itemize}
@@
     \caption{distribution of traversal visits}
-    \begin{subfigure}[b]{\textwidth}
-        \includegraphics[width=\textwidth]{graphics/top_1k_100_pages.png}
+        \includegraphics[width=\textwidth]{graphics/traversals_per_article.png}
+\end{figure}
+
+distribution of visits blah blah
+\begin{figure}[H]
+\centering
+\caption{traversal visit distribution of highest ranking articles}
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\linewidth]{graphics/top_1k_article_traversals.png}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\linewidth]{graphics/top_100_article_traversals.png}
 \end{subfigure}
 \end{figure}
-
-\red{[[change title from "pages" to "articles" for consistency]]} \\
-\red{[[justify scale-free with log-log or power measure]]} \\
 
@@
     \caption{highest ranking articles by number of traversals}
-    \begin{subfigure}[b]{0.8\textwidth}
-        \includegraphics[width=\textwidth, angle=-90]{graphics/top_articles_list.png}
-    \end{subfigure}
+        \includegraphics[width=0.6\textwidth]{graphics/articles_ranked.png}
 \end{figure}
@@
 \subsubsection{Measure Number of Synonyms}
+
+\green{
+number of synonyms
+top 100 articles, average:
+5.0 
+9.1 (excluding zero)
+median: 7.0
+
+remaining articles average:
+0.5
+5.8 (excluding zero)
+median: 3.0
+}
 \red{

### 2015-10-25 writeup/draft.tex
network d3 figures complete
parsing algo flags figure

@@
 	\fontsize{24pt}{12pt}\selectfont
-	\textbf{What's So Special About Philosophy?} 
+	\textbf{What's so Special About Philosophy?} 
 	}	
@@
 By following the first link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
-Here we study Wikipedia's First Link Network for insight into how we related topics, ideas, people, objects, and events.
+Here we study the English edition of Wikipedia's First Link Network for insight into how we relate topics, ideas, people, objects, and events.
 
-We algorithmically parse every article on the English Wikipedia to construct a map of the First Link Network. 
-We then traverse the network, following every possible path to uncover the network's structure.
+We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
+In a novel approach to uncover structure, we traverse every possible path through the network, 
+measuring the accumulation of first links, path lengths, basins, cycles, and even particular articles funneling links into the cycles.
 We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general with 
@@
 places, figures, and ideas, but also about \textit{relationships} among them.
-Here we study the relationships in this naturally arising web through the hyperlinks connecting one article to another.
+We study the relationships in this naturally arising web through the hyperlinks connecting one article to another.
 
@@
 This procedure corresponds to the original claim in 2008 about the percentage of pages with first links to 
-"Philosophy"\footnote{
+"Philosophy."\footnote{
 https://en.wikipedia.org/wiki/Wikipedia:GettingtoPhilosophy
@@
 
-To map Wikipedia's First Link Network, we use the freely-available XML dump of the English Wikipedia. 
+To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
 Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
@@
 Our parsing algorithm aimed to: 
-1. squash the initial bugs 
-2. eliminate the need for several passes. 
+1) accurately identify the first link among other page elements
+2) efficiently do so, without the need for several passes through the data.
 To process an article with one pass, we developed a hierarchical system of flags:
-\begin{itemize}
-    \item[Flag 1:] inside Wikimedia template?\\
-        * \bt{trigger}: \{\{
-    \begin{itemize}
-        \item[Flag 2:] inside <ref>, <div>?
-        \begin{itemize}
-            \item[Flag 3:] inside ()?
-        \end{itemize}
-    \end{itemize}
-\end{itemize}
 
-\red{[[better graphic to come]]}
+\begin{figure}[H]
+\centering
+\caption{parsing algorithm for XML}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{graphics/flags.pdf}    
+    \end{subfigure}
+\end{figure}
 
@@
     \begin{subfigure}[b]{0.8\textwidth}
-        \includegraphics[width=\textwidth]{graphics/top_2loops.png}
+        \includegraphics[width=\textwidth]{graphics/2_cycles.pdf}
     \end{subfigure}
@@
     \begin{subfigure}[b]{0.8\textwidth}
-        \includegraphics[width=\textwidth]{graphics/top_3loops.png}
+        \includegraphics[width=\textwidth]{graphics/3_cycles.pdf}
     \end{subfigure}
@@
 
+\begin{figure}[H]
+\centering
+\caption{highest ranking 3-Cycles}
+    \begin{subfigure}[b]{0.8\textwidth}
+        \includegraphics[width=\textwidth]{graphics/basins.pdf}
+    \end{subfigure}
+\end{figure}
+
 \subsection{Path Length}
@@
 
+\begin{figure}[H]
+\centering
+    \caption{Path Length Distribution}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{graphics/path_lengths_boxplot.png}
+    \end{subfigure}
+\end{figure}
 \begin{figure}[H]
@@
     \caption{Top Funnels}
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{graphics/top_funnels.png}
+    \begin{subfigure}[b]{0.9\textwidth}
+        \includegraphics[width=\textwidth]{graphics/funnels.pdf}
     \end{subfigure}
@@
 
+\section{Supplementary}
+
+\begin{figure}[H]
+\centering
+\caption{highest ranking 2-Cycles}
+    \begin{subfigure}[b]{0.8\textwidth}
+        \includegraphics[width=\textwidth]{graphics/top_2loops.png}
+    \end{subfigure}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\caption{highest ranking 2-Cycles}
+    \begin{subfigure}[b]{0.8\textwidth}
+        \includegraphics[width=\textwidth]{graphics/top_3loops.png}
+    \end{subfigure}
+\end{figure}
+\begin{figure}[H]
+\centering
+    \caption{Top Funnels}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{graphics/top_funnels.png}
+    \end{subfigure}
+\end{figure}
 %========================Style Guide====================================

### 2015-10-27 writeup/draft.tex
edits complete

@@
 paths\footnote{
-nearly $99.6\%$ of first link paths, ending inside Wikipedia, lead to the Philosophy article (see traversal funnels discussion for details).}
+nearly $99.6\%$ of First Link paths, ending inside Wikipedia, lead to the Philosophy article (see traversal funnels discussion for details).}
 \footnote{
@@
 More than information about a topic though, Wikipedia is a marvelous web of naturally emerging relationships.  
-By following the first link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
+By following the First Link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
 Here we study the English edition of Wikipedia's First Link Network for insight into how we relate topics, ideas, people, objects, and events.
@@
 In a novel approach to uncover structure, we traverse every possible path through the network, 
-measuring the accumulation of first links, path lengths, basins, cycles, and even particular articles funneling links into the cycles.
+measuring the accumulation of First Links, path lengths, basins, cycles, and even particular articles funneling links into the cycles.
 We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general with 
@@
 
-We build Wikipedia's First Link Network by following the first link, not in parenthesis, inside the main body of each article in the English version of Wikipedia. 
+We build Wikipedia's First Link Network by following the First Link, not in parenthesis, inside the main body of each article in the English edition of Wikipedia. 
 For the directed network to meaningfully reflect how we associate one article to another, we exclude links in parenthesis, 
@@
 projects outside of Wikipedia (such as Wiktionary).
-This procedure corresponds to the original claim in 2008 about the percentage of pages with first links to 
+This procedure corresponds to the original claim in 2008 about the percentage of pages with First Links to 
 "Philosophy."\footnote{
@@
 \end{figure}
+
 %===========================Data_and_Parsing================================
@@
 
-Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
+Wikipedia renders and stores articles in MediaWiki markup, a markup langauge with syntax and keywords to format and mark elements in a page.
 Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
 information.
-While a human can accurately identify the first link, to map the entire First Link Network of 11 million articles, we programmatically untangled the body text from side-bar, header box, and invalid link elements.
+While a human can accurately identify the First Link, to map the entire First Link Network of 11 million articles, we programmatically untangled the body text from side-bar, header box, and invalid link elements.
 
-While some libraries exist for MediaWiki Markup, we opted to develop our own algorithm for parsing the first link in the XML version of each article.\footnote{
+While some libraries exist for MediaWiki Markup, we opted to develop our own algorithm for parsing the First Link in the XML version of each article.\footnote{
 Approaches using existing libraries led to several bugs 
@@
 Our parsing algorithm aimed to: 
-1) accurately identify the first link among other page elements
+1) accurately identify the First Link among other page elements
 2) efficiently do so, without the need for several passes through the data.
@@
 \centering
-\caption{parsing algorithm for XML}
+\caption{parsing algorithm of Wikipedia's XML dump}
     \begin{subfigure}[b]{0.5\textwidth}
         \includegraphics[width=\textwidth]{graphics/flags.pdf}    
+        \caption{ The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
+Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
+Finally, we catch parenthesis to ensure we do not capture a link to a pronunciation key.}
     \end{subfigure}
+
 \end{figure}
@@
 until the flag's closing markup.
-A first link is identified only if Flags 1, 2, and 3 are all off.
+A First Link is identified only if Flags 1, 2, and 3 are all off.
 In this case, the entire link is retrieved. 
@@
 images, audio files, and the like.
-The first link of an article is then the earliest valid link with unraised flags.
+The First Link of an article is then the earliest valid link with unraised flags.
 
@@
 We then joined the results to form a hash table containing every Wikipedia article and its corresponding
-first link. The resulting network map is the basis of our analysis.
+First Link. The resulting network map is the basis of our analysis.
 
@@
 What types of cycle (loop) structures exist in the network?
-What are typical paths from one article to loop or a dead end (invalid link)? 
+What are typical paths from one article to loop or an invalid link? 
 Are there exceptionally long or short paths? 
@@
 
-Consequently, we developed three metrics to capture the dynamics of how the First Links flow. 
-First, we traverse the network to measure the accumulated number of visits each article receives: \\
+Consequently, to capture the dynamics of how the First Links flow, we actually traverse every possible path through the network. 
+We developed three metrics to uncover structure: traversal visits, traversal funnels, and path length.
+Each metric captures one essence of how the links flow.
+Traversal visits gauge the accumulation of links; traversal funnels gauge the influence on link path; path length measures the number of First Links to a repeated or invalid link.
+
+The algorithm for traversing the network begins by selecting any article.
+We then proceed to the next article by following the First Link---recalling a First Link is a link in the main body of the article leading to another Wikipedia article.
+We repeat until the First Link is invalid or repeated to form a path.
+The collection of articles is path-connected conveying a flow of concepts from one article to another.
+The method is order agnostic with respect to which articles are selected first.
+As long as each article is selected eventually, the resulting metrics are equivalent.
 
@@
 \centering
-    \caption{traversing the network}
-\begin{subfigure}[b]{0.8\textwidth}
-    \includegraphics[width=\textwidth]{graphics/traverse_algo.png}
+    \caption{traversal visit algorithm on a sample network}
+\begin{subfigure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{graphics/traversal_visit_algo_figure.pdf}
+    \caption{The traversal visit vectors are an adjacency matrix for the paths through the network: the first column is indicates the path formed starting with article A.
+The number of traversal visits for article A is then the number of paths containing A or the sum of the first row in our matrix.}
 \end{subfigure}
@@
 
-\red{[[explain with better diagram]]}
+The first metric, traversal visits, gauges where the flow of first accumulates by associating a count to each article.
+Every path through the article increments the associated count by 1.
+We can construct a vector to represent the articles incremented on a particular path, combining these vectors to form a matrix.
+The total accumulated counts for a given article is then simply the sum of the row in our matrix corresponding to the article of interest.
+Node A in our example has 7 traversal visits, since every path includes article A.
+Whereas, node E only has two traversal visits since only two paths contain article E for example.
+What we have then is a metric signaling which article contains a larger accumulation of first links.
+
+While traversal visits measure accumulation, we cannot gauge whether a particular article exerts a greater influence on flow of the first link paths.
+To distinguish between an article that simply happened to fall within a cycle from an article actually funneling the flow of first links, we developed a second metric called traversal funnels.
+We traverse the network in the same manner, but end our paths once we enter a cycle.
+We again increment the associated count of each article by one if the path up to a cycle contains the article.
+We call the accumulated count the number of traversal funnels.
+By measuring traversal funnels, we quantify
+which articles funnel more link paths in a particular direction or cycle.
 
@@
 \centering
-    \caption{traverse procedure}
-\begin{subfigure}[b]{0.3\textwidth}
-    \includegraphics[width=\textwidth]{graphics/traverse_procedure.png}
+    \caption{traversal funnel algorithm on a sample network}
+\begin{subfigure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{graphics/traversal_funnel_algo_figure.pdf}
+    \caption{The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
+We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.}
 \end{subfigure}
@@
 
-We also stop if the link is invalid. 
-By repeating the procedure starting at each article, we traverse
-every possible first link path through the network.
-The method is order agnostic with respect to the order of articles selected.
-As long as each article is selected as a starting point, the resulting set of measured traversal visits is equivalent. 
-More important, these traversals visits measure the accumulated flow of first 
-links to an article, preserving the link relations through the network.
-We call this value the number of traversal visits for a given article.
-
-\red{[[expand, explain better]]}
-
-Precedence for the traversal approach comes from methods used in \red{[[river networks? leaf pruning approach in environmental studies? what to cite?]]}
-
-Second, we traverse the network in the same manner, but end our paths once we enter a cycle.
-We call the accumulated number of articles funneled, the number of traversal funnels, because we quantify
-which articles funnel more link paths in a particular direction or cycle.
-
-Finally, we also track the path length, or the number of links until a repeated article or invalid link.
-Additionally, by recording the history of articles we traversed, we were able to map cycles in the network.
+Finally, we track the number of links until a repeated article or invalid link, to find the typical path length within the network.
+By recording the history of articles we traversed, we are also able to uncover several other network charactersitics including basins, defined as a collection of path-connected articles ranked by traversal visits, and cycles of various lengths.
 The three metrics: traversal visits, traversals funnels, and path length, along with our path history
 yield a powerful arsenal of information with which to study the network.
-We can effectively answer questions about how the First Links Flow, identify both peculiar and popular
-articles (and groups of articles), as well as link paths.
 
@@
 
-distribution of visits blah blah
+To more accurately gauge the distribution, we constructed a log-log plot of the entire dataset: log(traversal visits) against log(rank).
+We observe a fairly linear fit, as is characteristic of scale-free networks, with an r-value of -.93.
+A handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none.
+The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.
+
 \begin{figure}[H]
@@
 The highest ranking articles include Philosophy alongside related articles such as "Existence", "Quality", and "Reality".
-Other high ranking articles include similarly foundational ideas or abstracted disciplines such as "Science", "Biology", 
-"Community", "State", "Earth", "Human Geography", "Information", and "Communication".
 
@@
 \centering
-    \caption{highest ranking articles by number of traversals}
+    \caption{highest ranking articles by number of traversal visits}
         \includegraphics[width=0.6\textwidth]{graphics/articles_ranked.png}
@@
 
-\subsubsection{Measure Number of Synonyms}
+\subsubsection{A flow from general to specific}
 
-\green{
-number of synonyms
-top 100 articles, average:
-5.0 
-9.1 (excluding zero)
-median: 7.0
+The highest ranking articles by traversal visits are 
+broad, global topics: 
+many are academic disciplines such as "Science," "Math," 
+"Geography," "Biology," and "Physics"; others are abstract fundamental concepts such as 
+"Community", "State", "Earth", "Information", "Communication", and "Power."
+Since traversal visits measure the accumulation of First Link visits in Wikipedia, 
+the highest ranking articles suggest a flow from specific to general. 
+For example the flow of First Links for the "Banana" article begins at a very concrete
+and specific topic then flows into progressively broader and broader disciplines, eventually 
+culminating at "Philosophy: "Banana" links to the broader category of "Fruit," which then links to 
+"Botany," eventually "Biology," then "Science" and ultimately "Philosophy."
 
-remaining articles average:
-0.5
-5.8 (excluding zero)
-median: 3.0
-}
-\red{
-[[ add synonym analysis to support characterization of high ranking articles
-as abstract foundational notions. \\
-- idea is to correlate the number of synonyms a word has with its traversal visit rank.
-]]}
+One means to measure the specificity of an article is to identify the number of synonyms available for a word or topic.
+The reasoning here is that a  broader topic likely has many more synonyms relative to a specific, concrete topic.
+Banana has fewer synonyms than botanical for example.
+To quantify the observed flow, we measured the number of synonyms the article title contained in WordNet---the largest lexical database of the English language.
+We find the highest ranking 100 articles have on average 5 more synonyms than the typical article; a difference of 2.5 synonyms if we compare the median number of synonyms in each group.
+As suggested by the median, many articles have no synonyms as we might expect, because titles with more than one word are not likely to appear in a thesaurus.
+Since many articles have no synonyms, we also compared the number of synonyms in the highest-ranking versus typical article, this time excluding all articles without at least 1 synonym.
+We still find the highest ranking 100 articles with an average of 9.0 (median of 7.0) synonyms whereas the remaining articles on average 5.8 synonyms (median of 7.0), even with the exclusion of all articles without any synonyms.
+The quantifiable difference in synonyms corroborates the flow of links from concrete, specific articles to broader disciplines or fundamental notions.
 
@@
 by the reappearance of an article in our path history.
-We first identified 2-cycles, meaning a pair of articles with first link pointing to one another.
+We first identified 2-cycles, meaning a pair of articles with First Link pointing to one another.
 Of the 11 million articles, 84 thousand are 2-cycles. 
@@
 \end{figure}
-
-\red{[[better representation to come\\
-look into articles that don't seem to have a partner]]}
 
@@
 \end{figure}
-((diagram of cycles beyond length 10))
 
@@
 Philosophy is not only a stand out the number of traversal visits, but also by the number articles "Philosophy" funnels into
-its cycle: \red{[[99.something]]} $\%$. 
+its cycle. Of any article, the number of traversal funnels Philosophy holds exceeds 
+all others by at least two orders of magnitude.
+Next to the contribution of the largest funnels, "Philosophy" is a singularity. 
+In proportion, "Presentation" holds only $0.4\%$ next to the number of traversal feeds for "Philosophy".
 The second contributor to the Philosophy cycle is "Reality", funneling in a mere $.2\%$ of traversals visits in the cycle.
-Even next to the contribution of the largest funnels, "Philosophy" is a singularity. 
-In proportion, "Presentation" holds only $0.4\%$ next to the number of traversal feeds for "Philosophy".
 Nevertheless, the other high ranking funnels are remarkably topical, culturally and politically important ideas.
-For example, "Health Care", a recently high-contested legislative topic \red{[[add google results trend]]} appears high on the list
-as it does in google search terms or ((media articles?)).
+For example, "Health Care", a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August-2009 and February-2010.
 Other high ranking articles include key historical events such as the "Cold War" or critical scarce resource with recent 
@@
 This coincidence of recent relevance and traversal feed rank suggests the First Link Network measurably represents
-meaningful relationships not only among ideas, but also to society ((english speaking)).
+meaningful relationships not only among ideas, but also to English-speaking society.
 
@@
 We examined only the English version of Wikipedia at a particular moment in time.
-Furthermore, we only studied the first link in the main body of each article
+Furthermore, we only studied the First Link in the main body of each article
 as a means to related one article to another. Finally, Wikipedia, while the largest 
-collection of human knowledge, is rife with the biases of the many contributing editors--
-male middle aged \red{[[substantiate and confirm]]}. Nevertheless, the findings do reveal
+collection of human knowledge, is rife with the biases of the many contributing editors ((cite))
+. Nevertheless, the findings do reveal
 generalizable relationships, point to foundational notions, and uncover many curiosities.
@@
 
-Future work could analyze other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network beyond the first link to a subset or potentially all links.
+Future work could analyze other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network beyond the First Link to a subset or potentially all links.
 These findings also form the basis for the creation of a taxonomy where 
@@
 - thank you to my friend RJ for pointing out the xkcdc comic about links to Philosophy
+
+\newpage
 

### 2015-10-27 writeup/draft.tex
raggedbottom option for easier reading
footnote with no markers added
3cycles graph updated

@@
 	\fontsize{24pt}{12pt}\selectfont
-	\textbf{What's so Special About Philosophy?} 
+	\textbf{What's so Special about Philosophy?} 
 	}	
@@
 
+%text, figure spacing
+\raggedbottom
+
+%footnote withou marker
+\newcommand\blfootnote[1]{%
+  \begingroup
+    \renewcommand\thefootnote{}\footnote{#1}%
+      \addtocounter{footnote}{-1}%
+   \endgroup }
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 Within Wikipedia, the surprising answer is yes: nearly all 
-paths\footnote{
+paths lead to Philosophy.
+\footnote{
 nearly $99.6\%$ of First Link paths, ending inside Wikipedia, lead to the Philosophy article (see traversal funnels discussion for details).}
@@
 }
-lead to Philosophy.
 Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
@@
 \centering
-\caption{highest ranking 2-Cycles}
+\caption{highest ranking 2-cycles}
     \begin{subfigure}[b]{0.8\textwidth}
@@
 \centering
-\caption{highest ranking 3-Cycles}
+\caption{highest ranking 3-cycles}
     \begin{subfigure}[b]{0.8\textwidth}
@@
 
+\blfootnote{
 - thank you to my friend RJ for pointing out the xkcdc comic about links to Philosophy
+}
 

### 2015-10-27 writeup/draft.tex
final edits

@@
 
-The recurrence of an exact number of traversal visits suggests some articles are part of a cycle. 
-The "Philosophy" article for example sits in what seems like a cycle of seven other articles; "Hypothesis" appears to sit in a 
-cycle of 6 other articles including "Experiment", "Fact", and "Knowledge".
-To confirm the suggested cycle structure, we recorded the history of articles traversed along a path.
-
 \subsubsection{A flow from general to specific}
@@
 \subsection{Network Cycles}
-By tracking the articles traversed we were able to identify the cycle structures
-by the reappearance of an article in our path history.
+
+The recurrence of an exact number of traversal visits suggests some articles are part of a cycle. 
+The "Philosophy" article for example sits in what seems like a cycle of seven other articles; "Hypothesis" appears to sit in a 
+cycle of 6 other articles including "Experiment", "Fact", and "Knowledge".
+To confirm the suggested cycle structure, we recorded the history of articles traversed along a path.
+
 We first identified 2-cycles, meaning a pair of articles with First Link pointing to one another.
-Of the 11 million articles, 84 thousand are 2-cycles. 
+Of the 11 million articles, 84,000 are 2-cycles. 
 The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than different, yet connected ideas:
@@
 In proportion, "Presentation" holds only $0.4\%$ next to the number of traversal feeds for "Philosophy".
-The second contributor to the Philosophy cycle is "Reality", funneling in a mere $.2\%$ of traversals visits in the cycle.
+The second contributor to the Philosophy cycle is "Reality", funneling in a mere $0.2\%$ of traversals visits in the cycle.
 Nevertheless, the other high ranking funnels are remarkably topical, culturally and politically important ideas.
@@
 
-Future work could analyze other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network beyond the First Link to a subset or potentially all links.
+Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the First Link.
 These findings also form the basis for the creation of a taxonomy where 

### 2015-11-16 writeup/2015-11-wikipedia-network.tex
paper-esque version

--- 2015-10-27 writeup/draft.tex
@@
-\documentclass[twoside]{article}
+\documentclass[pre,twocolumn,twoside,superscriptaddress,floatfix, aps, 10pt]{revtex4-1}
 
-% ------
+\usepackage{currfile}
+\lefthyphenmin=3
+\righthyphenmin=2
+
+\usepackage{graphicx,epsfig,verbatim,enumerate}
+\usepackage{amssymb,amsmath}
+\usepackage{ifthen}
+
 % Fonts and typesetting settings
@@
 
-% ------
-% Page layout
-\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
-\usepackage[font=it]{caption}
-\usepackage{multicol, hyperref}
-\usepackage{parskip}
+\usepackage{longtable}
 
-% ------
+\usepackage{mathtools}
+
+\newboolean{twocolswitch}
+
+\newcommand{\www}[1]{\url{#1}}
+\newcommand{\req}[1]{(\ref{#1})}
+\newcommand{\Req}[1]{Eq.~(\ref{#1})}
+
 % Lettrines
 \usepackage{lettrine}
-\usepackage{color}
-
-% ------
-% Abstract
-\usepackage{abstract}
-	\renewcommand{\abstractnamefont}{\normalfont\bfseries}
-	\renewcommand{\abstracttextfont}{\normalfont\small\itshape}
-
-% ------
-% Titling (section/subsection)
-\usepackage{titlesec}
-\renewcommand\thesection{\Roman{section}}
-\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{}
-
-% ------
-% Header/footer
-\usepackage{fancyhdr}
-	\pagestyle{fancy}
-	\fancyhead{}
-	\fancyfoot{}
-	\fancyhead[C]{ $\bullet$ Draft $\bullet$}
-	\fancyfoot[RO,LE]{\thepage}
-
-% My Shortcuts
 
@@
 \newcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-\newcommand{\cb}[1]{\centerline{\fbox{#1}}} %centered box
-\newcommand{\bp}[1]{\fbox{\parbox{0.8\textwidth}{#1}}} %box paragraph
 \newcommand{\norm}[1]{\left\lVert#1\right\rVert} %vector norm
-\newcommand{\notimplies}{% does not imply
-  \mathrel{{\ooalign{\hidewidth$\not\phantom{=}$\hidewidth\cr$\implies$}}}}
 \renewcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
+
+%piecewise function
+
+%\begin{displaymath}
+%   f(x) = \left\{
+%     \begin{array}{lr}
+%       1 & : x \in \mathbb{Q}\\
+%       0 & : x \notin \mathbb{Q}
+%     \end{array}
+%   \right.
+%\end{displaymath}
+
+%environment
+\newcommand{\tab}{\phantom{ssss}}
+
+\usepackage{color}
+\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}
 
@@
 \newcommand{\gray}[1]{{\leavevmode\color[gray]{0.5}{#1}}} %gray text
+\newcommand{\Prob}[1]{{\rm Pr}\left(#1\right)}
 
-%environment
-\newcommand{\tab}{\phantom{ssss}}
-%----------------
+\setboolean{twocolswitch}{true}
 
-% ------
-% Clickable URLs (optional)
-\usepackage{hyperref}
+\begin{document}
 
-% ------
-% Maketitle metadata
-\title{\vspace{-5mm}%
-	\fontsize{24pt}{12pt}\selectfont
-	\textbf{What's so Special about Philosophy?} 
+\title{\protect
+A Web of Human Knowledge: 
+Unravelling Wikipedia's First Link Network
 }
-\author{%
-\fontsize{14pt}{14pt}\selectfont
-	Unraveling Wikipedia's First Link Network \vspace{-2mm}\\
+
+\author{
+\firstname{Mark}
+\surname{Ibrahim}
 }
-\date{}
+\email{mark.s.ibrahim@uvm.edu}
 
-%figures
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{float}
+\affiliation{Department of Mathematics \& Statistics, 
+    Computational Story Lab,  \\
+    The University of Vermont, Burlington, VT 05401.}
 
-%text, figure spacing
-\raggedbottom
+\author{
+\firstname{Christopher}
+\surname{M. Danforth}
+}
+\email{chris.danforth@uvm.edu}
 
-%footnote withou marker
-\newcommand\blfootnote[1]{%
-  \begingroup
-    \renewcommand\thefootnote{}\footnote{#1}%
-      \addtocounter{footnote}{-1}%
-   \endgroup }
-%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-\maketitle
-\thispagestyle{fancy}
-%========================ABSTRACT====================================
+\affiliation{Department of Mathematics \& Statistics, 
+    Computational Story Lab,  \\
+    The University of Vermont, Burlington, VT 05401.}
+
+\author{
+\firstname{Peter}
+\surname{Sheridan Dodds}
+}
+
+\email{peter.dodds@uvm.edu}
+
+\affiliation{Department of Mathematics \& Statistics, 
+    Computational Story Lab,  \\
+    The University of Vermont, Burlington, VT 05401.}
 
 \begin{abstract}
-\fontsize{12pt}{12pt}
-\selectfont
-
-Apples, oranges, and the most obscure Dylan song too---is everything a few clicks from Philosophy? 
-Within Wikipedia, the surprising answer is yes: nearly all 
-paths lead to Philosophy.
-\footnote{
-nearly $99.6\%$ of First Link paths, ending inside Wikipedia, lead to the Philosophy article (see traversal funnels discussion for details).}
-\footnote{
-An analysis performed in 2011 by Mat Kelcey posted on his blog (and a smiliar post by Ilmari Karonen) suggest $94.5\%$ of articles end up at Philosophy.
-}
-Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
-\footnote{
-cite MIT Tech Review Article: http://www.technologyreview.com/featuredstory/520446/the-decline-of-wikipedia/
-}
-More than information about a topic though, Wikipedia is a marvelous web of naturally emerging relationships.  
-By following the First Link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
-Here we study the English edition of Wikipedia's First Link Network for insight into how we relate topics, ideas, people, objects, and events.
-
-We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
-In a novel approach to uncover structure, we traverse every possible path through the network, 
-measuring the accumulation of First Links, path lengths, basins, cycles, and even particular articles funneling links into the cycles.
-We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general with 
-basins around fundamental notions such as Community, State, and Science. 
-Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
-These findings enrich our view of how we connect and structure
-an ever growing load of information.
+  \protect
+  this is the glorious abstract of abstractness summarizing methods,
+  results, and contribution of paper our understanding of the world
 
@@
 
-\fontsize{11pt}{11pt}
-\selectfont
+\maketitle
 
-%========================INTRODUCTION====================================
-At no point in history has a larger or more meticulously indexed collection of human knowledge existed.
-In amassing such an awe-inspiring collection, we formed an equally impressive web. 
-Through the efforts of millions of individuals, working independently,
-((cite))
-we naturally linked topics, inventions, people, objects, places, and events across space and time.
-The web we weaved, and continue to weave, is a wealth of information not only about those notable inventions, 
-places, figures, and ideas, but also about \textit{relationships} among them.
-We study the relationships in this naturally arising web through the hyperlinks connecting one article to another.
+\section*{Introduction}
 
-We build Wikipedia's First Link Network by following the First Link, not in parenthesis, inside the main body of each article in the English edition of Wikipedia. 
-For the directed network to meaningfully reflect how we associate one article to another, we exclude links in parenthesis, 
-disregarding pronunciation keys or disambiguations.
-We also exclude any links in the side-bar elements, as well as any links to external pages, files, or WikiMedia 
-projects outside of Wikipedia (such as Wiktionary).
-This procedure corresponds to the original claim in 2008 about the percentage of pages with First Links to 
-"Philosophy."\footnote{
-https://en.wikipedia.org/wiki/Wikipedia:GettingtoPhilosophy
+Wikipedia is a towering achievement of the modern era. 
+At no point in history has a larger or more meticulously indexed collection of human knowledge 
+existed.
+Wikipedia contains 37 million articles in 283 languages, 
+with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
+\cite{stats} \cite{drugs}.
+All this information does not lie dormant. 
+Wikipedia is the sixth most visited
+site in the world, surpassing 18 billion page views in a single month. 
+The most widely used general reference, Wikipedia has amassed an awe-inspiring collection of human knowledge.
+
+It's not surprising then to see Wikipedia as the object of many studies. 
+Researchers have examined the cultural dynamics among editors,
+\cite{editors}
+the accuracy of the content relative to traditional encyclopedias,
+\cite{accuracy1}
+\cite{accuracy2}
+the disciplines and topics covered 
+\cite{coverage}
+and bias against portions of the population
+\cite{bias_women}.
+Wikipedia's content has also proven a powerful tool. 
+Researchers have used Wikipedia to identify missing dictionary entries
+\cite{missing_entries},
+cluster short text
+\cite{clustering},
+compute semantic relatedness
+\cite{semantic_relatedness},
+and disambiguate meaning \cite{disambiguating}.
+
+While these many studies have dissected and fruitfully applied Wikipedia's content,
+few have examined the connections among the many articles.
+Beyond information, Wikipedia is a web 
+relating ideas. 
+Within the body text of Wikipedia articles hyperlinks reference
+other Wikipedia articles.
+A hyperlink from one Wikipedia article to another is a way to relate the ideas
+conveyed in the two articles 
+\cite{relevance}.
+The notion that hyperlinks convey information about the content of a
+page has proved successful in numerous domains from search engine algorithms 
+such as PageRank 
+\cite{pagerank} 
+to topic classification
+\cite{classifier}.
+We treat a hyperlink as a mechanism to connect two ideas.
+
+The authors of a Wikipedia article choose where and whether to include a 
+hyperlink reference to another Wikipedia article.
+Authors must include hyperlinks in the article's markup.
+In describing a particular topic, each set of authors decides which articles are the most
+relevant references.
+For example, the hyperlinks in the "Train" article depend only on what its authors 
+decide are pertinent articles to reference.
+The collection of hyperlinks emerges from the independent choice of each article's authors. 
+The first hyperlink to another Wikipedia article marks the earliest moment in the body text where the authors
+choose to connect the information discussed to that of another article.
+By following the first hyperlink to another article in the English edition of
+Wikipedia, we connect the ideas in one article to those in another, ultimately forming a directed network: 
+{\it Wikipedia's First Link Network} (FLN).
+
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
+  \caption{
+    \textbf{First Link Path For "Train"}
+    We follow the first link to another Wikipedia article
+    in the main body of the article---the area
+    inside the green rectangle, which excludes 
+    side bar elements, the navigation bar and title. 
+    In this example, the first link to another Wikipedia article is "Rail Transport." 
+    We can again select the first link on the "Rail Transport"
+    article, repeating the process to 
+    form a path of First Links.
   }
-\footnote{
-http://xkcd.com/903/
-}
-\footnote{
-https://www.reddit.com/r/pics/comments/gpdhb/trythiswikipediamindfk/
-}
+  \label{fig:Train First Links}
+\end{figure*}
 
-The result is a directed network placing every article in a broader web of ideas.
+What information could the first links possibly reveal?
+Inspired by the claim that the majority of first links lead to the 
+"Philosophy" article---popularized by an xkcd comic and subsequently
+discussed in blog posts ---we holistically study 
+Wikipedia's FLN as a map relating areas of human knowledge. 
+The map is a hierarchical taxonomy where "Train" links to a parent node, 
+"Rail Transport," while many child nodes, "Steel" and "Horsepower," link up to 
+"Train." Unlike previous taxonomies
+created by individuals
+\cite{locke}
+\cite{descartes}
+\cite{aristotle} 
+or a select group of individuals 
+\cite{hist_thesaurus}, 
+the organization of ideas in the FLN 
+does not emerge from a concerted centralized effort, 
+but rather from the independent choice of every article's authors.
+The FLN is a wealth of relations among all the inventions, places,
+figures, objects, and events described in Wikipedia, across space and time.
 
-\begin{figure}[H]
-\centering
-\caption{Wikipedia Train}
-    \begin{subfigure}[b]{0.8\textwidth}
-        \includegraphics[width=\textwidth]{graphics/wiki_train.png}    
-    \end{subfigure}
-\end{figure}
+Our goal is to gain insight into how the information on Wikipedia is organized and related
+by studying the structure of the FLN.
+In a novel approach, we develop metrics to capture 
+the dominant features of the FLN's structure.
+We measure dynamics of the FLN as a flow, quantifying 
+aspects of the FLN from the accumulation of first links around articles 
+to the influence an article exerts in shaping the FLN.
+Together with cycles, n-degree, depth, and the content of the articles, 
+we build our analysis of the relations among the ideas in Wikipedia.
 
-%===========================Data_and_Parsing================================
-\section{Constructing the First Link Network}
-
-To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
-Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
-eliminating any statistical error due to sampling.
-We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
-The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
-and disambiguations.
-Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
-((cite)), our aim is to characterize the dynamics of the First Link Network, not record a particular link between one
-article and another.
-
-Wikipedia renders and stores articles in MediaWiki markup, a markup langauge with syntax and keywords to format and mark elements in a page.
-Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
-information.
-While a human can accurately identify the First Link, to map the entire First Link Network of 11 million articles, we programmatically untangled the body text from side-bar, header box, and invalid link elements.
-
-While some libraries exist for MediaWiki Markup, we opted to develop our own algorithm for parsing the First Link in the XML version of each article.\footnote{
-Approaches using existing libraries led to several bugs 
-including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
-as well as compatibility with other libraries used to parse the XML.}
-Our parsing algorithm aimed to: 
-1) accurately identify the First Link among other page elements
-2) efficiently do so, without the need for several passes through the data.
-To process an article with one pass, we developed a hierarchical system of flags:
-
-\begin{figure}[H]
-\centering
-\caption{parsing algorithm of Wikipedia's XML dump}
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{graphics/flags.pdf}    
-        \caption{ The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
-Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
-Finally, we catch parenthesis to ensure we do not capture a link to a pronunciation key.}
-    \end{subfigure}
-
-\end{figure}
-
-The algorithm loops in three-character chunks to account for potentially nested elements, 
-shifting by one character steps through the article markup.
-If any markup triggers for a flag are detected, a flag is raised. 
-Once a flag is raised, we stop processing and proceed to the next character
-until the flag's closing markup.
-A First Link is identified only if Flags 1, 2, and 3 are all off.
-In this case, the entire link is retrieved. 
-We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
-as well as common file extensions for 
-((cite))
-images, audio files, and the like.
-The First Link of an article is then the earliest valid link with unraised flags.
-
-To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
-across 112 cores of the UVM supercomputer cluster
-((cite))
-We then joined the results to form a hash table containing every Wikipedia article and its corresponding
-First Link. The resulting network map is the basis of our analysis.
-
-%========================Traversing the Network====================================
-\section{Traversing the Network}
+\section*{Traversing the First Link Network}
 
@@
 aimed to characterize the dynamics of the flow from one article to another. 
-Do links tend towards a particular article, group of articles, or various groups of articles? 
+Do links tend towards a particular article, group of articles, or groups of articles? 
 What is the flow of links through a typical article?
@@
 The n-degree, or number of links directly pointing to a particular article,
-while a natural measure, fails to fully capture the dynamics needed to answer these questions.
+inadequately captures the dynamics needed to answer these questions.
 The n-degree measures only the particular set of links to a particular article, 
@@
 
-Consequently, to capture the dynamics of how the First Links flow, we actually traverse every possible path through the network. 
-We developed three metrics to uncover structure: traversal visits, traversal funnels, and path length.
-Each metric captures one essence of how the links flow.
-Traversal visits gauge the accumulation of links; traversal funnels gauge the influence on link path; path length measures the number of First Links to a repeated or invalid link.
-
+Consequently, to capture the dynamics of how the First Links flow, we actually traverse every possible path through the FLN. 
 The algorithm for traversing the network begins by selecting any article. 
@@
 
-\begin{figure}[H]
+We developed three metrics to uncover structure: 
+traversal visits, traversal funnels, and path length. 
+Each metric captures one essence of how the links flow. 
+Traversal visits gauge the accumulation of links; 
+traversal funnels gauge the influence on link path; 
+path length measures the number of First Links to a repeated or invalid link.
+
+\begin{figure*}[tp!]
   \centering	
-    \caption{traversal visit algorithm on a sample network}
-\begin{subfigure}[b]{\textwidth}
   \includegraphics[width=\textwidth]{graphics/traversal_visit_algo_figure.pdf}  
-    \caption{The traversal visit vectors are an adjacency matrix for the paths through the network: the first column is indicates the path formed starting with article A.
-The number of traversal visits for article A is then the number of paths containing A or the sum of the first row in our matrix.}
-\end{subfigure}
-\end{figure}
+  \caption{
+    \textbf{Traversal Visit Algorithm on a sample network}
+     The traversal visit vectors are an adjacency matrix for the paths through the network: the first column is indicates the path formed starting with article A.
+The number of traversal visits for article A is then the number of paths containing A or the sum of the first row in our matrix:
+     $\sum_{i=1}^7 A_{\text{visit, i}} = 7$.
+  }
+  \label{fig:Traversal Visits}
+\end{figure*}
 
-The first metric, traversal visits, gauges where the flow of first accumulates by associating a count to each article.
-Every path through the article increments the associated count by 1.
-We can construct a vector to represent the articles incremented on a particular path, combining these vectors to form a matrix.
+The first metric, {\it traversal} visits, gauges where the flow of first accumulates by associating a count to each article. 
+Every path through the article increments the associated count by one. We can construct a vector to represent the articles incremented on a particular path, combining these vectors to form a matrix. 
 The total accumulated counts for a given article is then simply the sum of the row in our matrix corresponding to the article of interest. 
-Node A in our example has 7 traversal visits, since every path includes article A.
+Node A in our example (see figure ~\ref{fig:Traversal Visits}) has 7 traversal visits, since every path includes article A. 
 Whereas, node E only has two traversal visits since only two paths contain article E for example.
@@
 We again increment the associated count of each article by one if the path up to a cycle contains the article. 
-We call the accumulated count the number of traversal funnels.
-By measuring traversal funnels, we quantify
-which articles funnel more link paths in a particular direction or cycle.
+We call the accumulated count the number of {\it traversal funnels}. By measuring traversal funnels, we quantify
+which articles funnel more link paths in a particular direction or cycle 
+(see figure ~\ref{fig:Traversal Funnels})
+.
 
-\begin{figure}[H]
+\begin{figure*}[tp!]
   \centering	
-    \caption{traversal funnel algorithm on a sample network}
-\begin{subfigure}[b]{\textwidth}
-    \includegraphics[width=\textwidth]{graphics/traversal_funnel_algo_figure.pdf}
-    \caption{The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
-We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.}
-\end{subfigure}
-\end{figure}
+  \includegraphics[width=\columnwidth]{graphics/traversal_funnel_algo_figure.pdf}
+  \caption{
+    \textbf{Traversal Funnel Algorithm on a sample network}
+  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
+We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.
+For example
+  the number of traversal funnels for article $E$ is 
+  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
+  \label{fig:Traversal Funnels}
+
+\end{figure*}
 
 Finally, we track the number of links until a repeated article or invalid link, to find the typical path length within the network.
-By recording the history of articles we traversed, we are also able to uncover several other network charactersitics including basins, defined as a collection of path-connected articles ranked by traversal visits, and cycles of various lengths.
+By recording the history of articles we traversed, we are also able to uncover several other network characteristics including basins, defined as a collection of path-connected articles ranked by traversal visits, and cycles of various lengths.
 The three metrics: traversal visits, traversals funnels, and path length, along with our path history
@@
 
-%========================Results====================================
 \section{Discoveries}
@@
 
-\begin{figure}[H]
+\begin{figure*}[tp!]
   \centering	
-    \caption{distribution of traversal visits}
-        \includegraphics[width=\textwidth]{graphics/traversals_per_article.png}
-\end{figure}
+  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
+  \caption{
+    \textbf{Distribution of Traversal Visits}
+  }
+  We fit the distribution to a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal visits. 
+  The model explains $86\%$ of the variation in the data, yielding an r-value of $-0.93$ 
+  and a power law exponent of $-0.636$. The horizontal flattening around the highest
+  ranking articles is a result of the cyclic structure (see discussion on cycles).
+  \label{fig:Distribution of Visits}
 
-To more accurately gauge the distribution, we constructed a log-log plot of the entire dataset: log(traversal visits) against log(rank).
+\end{figure*}
+
+To more accurately gauge the distribution, we construct a log-log plot of the entire dataset: log(traversal visits) against log(rank).
 We observe a fairly linear fit, as is characteristic of scale-free networks, with an r-value of -.93.
@@
 
-\begin{figure}[H]
-\centering
-\caption{traversal visit distribution of highest ranking articles}
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.9\linewidth]{graphics/top_1k_article_traversals.png}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.9\linewidth]{graphics/top_100_article_traversals.png}
-\end{subfigure}
-\end{figure}
-
 The highest ranking articles include Philosophy alongside related articles such as "Existence", "Quality", and "Reality".
 
-\begin{figure}[H]
+\begin{figure*}[tp!]
   \centering	
-    \caption{highest ranking articles by number of traversal visits}
-        \includegraphics[width=0.6\textwidth]{graphics/articles_ranked.png}
-\end{figure}
+  \includegraphics[width=\columnwidth]{graphics/articles_ranked.png}
+  \caption{
+    \textbf{highest ranking articles by number of traversal visits}
+  }
+  We compute the number of traversal visits for each article in the FLN (see 
+  Traversal Algorithm section for details). In doing so, we can rank each article
+  by the accumulation of first links. Articles with a greater number of traversal visits
+  mark greater points of first link accumulation. The highest ranking articles by traversal visits reveals where the greatest accumulation occurs.
+  \label{fig:highest visits}
+\end{figure*}
 
@@
 
-\begin{figure}[H]
+\begin{figure*}[tp!]
   \centering	
-\caption{highest ranking 2-cycles}
-    \begin{subfigure}[b]{0.8\textwidth}
   \includegraphics[width=\textwidth]{graphics/2_cycles.pdf}
-    \end{subfigure}
-\end{figure}
+  \caption{
+    \textbf{highest ranking 2-cycles}
+  }
+  \label{fig:2-cycles}
+
+\end{figure*}
 
@@
 
-\begin{figure}[H]
+\begin{figure*}[tp!]
   \centering	
-\caption{highest ranking 3-cycles}
-    \begin{subfigure}[b]{0.8\textwidth}
   \includegraphics[width=\textwidth]{graphics/3_cycles.pdf}
-    \end{subfigure}
-\end{figure}
+  \caption{
+    \textbf{highest ranking 3-cycles}
+  }
+  \label{fig:3-cycles}
+\end{figure*}
 
@@
 \subsection{Basins}
-We can group articles lying on the same path to identify basins of path-connected articles, not necessarily forming a perfectly closed cycle.
-Ranking basins by traversal visits, we find many of highest ranking basins are around "Philosophy" as we might expect. 
-Looking beyond "Philosophy" however, we find high ranking basins around similarly foundational ideas:
+
+We can group articles lying on the same path to identify {\it basins}: 
+a group of path-connected articles.
+Since cycles identify only groups of articles with a closed set of links, 
+we additionally measure and rank basins to capture groups of closely related
+articles branching outside of a cycle into the rest of the FLN.
+Akin to river networks, these basins are areas of FLN accumulation with a path 
+flowing outwards to the rest of the FLN.
+The highest ranking basins by the number of traversal visits are groups of articles
+around "Philosophy" given the high number of traversal visits "Philosophy" has. 
+However, restricting our attention to the 
+ single articles around which basins formed,
+ we find high ranking basins around similarly foundational ideas:
 "Community", "Landmass", "Federal Government", "Presentation", and "Belief System". 
+Around each article is a sets of related ideas. Around "community" are groups of articles expressing similar ideas: "Political Union," "State," and "Federation." 
 These concepts naturally emerge from the First Link Network potentially indicating pillars, which 
@@
 
-\begin{figure}[H]
+\begin{figure*}[tp!]
   \centering	
-\caption{highest ranking 3-Cycles}
-    \begin{subfigure}[b]{0.8\textwidth}
-        \includegraphics[width=\textwidth]{graphics/basins.pdf}
-    \end{subfigure}
-\end{figure}
+  \includegraphics[width=\columnwidth]{graphics/basins.pdf}
+  \caption{
+    \textbf{basins around a single article}
+  }
+  \label{fig:basins}
+
+\end{figure*}
 
@@
 
-\begin{figure}[H]
+\begin{figure}[tp!]
   \centering	
-    \caption{Path Length Distribution}
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{graphics/path_lengths_boxplot.png}
-    \end{subfigure}
+  \includegraphics[width=\columnwidth]{graphics/path_lengths_boxplot.png}
+  \caption{
+    \textbf{Path Length Distribution}
+  }
+  \label{fig:Path Length Distribution}
 \end{figure}
-\begin{figure}[H]
+
+\begin{figure*}[tp!]
   \centering	
-    \caption{Longest Path Lengths}
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{graphics/top_1k_path_length.png}
-    \end{subfigure}
-\end{figure}
+  \includegraphics[width=\columnwidth]{graphics/top_1k_path_length.png}  
+  \caption{
+    \textbf{Path Length Distribution for Top 1k articles}
+  }
+  \label{fig:Top 1k Path Length Distribution}
+\end{figure*}
 
@@
 Measuring traversal funnels reveals a dramatically different structure where "Philosophy" stands unmatched by orders of magnitude:
-
-\begin{figure}[H]
+\begin{figure*}[tp!]
   \centering	
-    \caption{Top Funnels}
-    \begin{subfigure}[b]{0.9\textwidth}
   \includegraphics[width=\textwidth]{graphics/funnels.pdf}
-    \end{subfigure}
-\end{figure}
+  \caption{
+    \textbf{Funnels}
+  }
+  \label{fig:Funnels}
+\end{figure*}
 
@@
 %========================Conclusion====================================
+
 \section{Reflections}
@@
 
-\blfootnote{
-- thank you to my friend RJ for pointing out the xkcdc comic about links to Philosophy
-}
+\acknowledgments
+Thank you to my friend RJ for pointing out the xkcd comic describing how the majority of links lead to "Philosophy," inspiring this research.
 
@@
 
-\section{Supplementary}
+\section*{Appendix}
 
-\begin{figure}[H]
+\subsection*{Constructing The First Link Network}
+
+To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
+Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
+eliminating any statistical error due to sampling.
+We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
+The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
+and disambiguations.
+Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
+((cite)), our aim is to characterize the dynamics of the First Link Network, not record a particular link between one
+article and another.
+
+Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
+Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
+information.
+While a human could accurately identify the First Link, to map the entire First Link Network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
+
+While some libraries exist for MediaWiki Markup,
+Approaches using existing libraries led to several bugs 
+including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
+as well as compatibility with other libraries used to parse the XML.
+Consequently,  we developed an algorithm for parsing the First Link in the XML version of each article.
+Our parsing algorithm aimed to: 
+1) accurately identify the First Link among other page elements, and 
+2) efficiently do so---that is without needing for several passes through the data.
+To process an article in a single pass, we developed a hierarchical system of flags:
+
+\begin{figure*}[tp!]
   \centering	
-\caption{highest ranking 2-Cycles}
-    \begin{subfigure}[b]{0.8\textwidth}
-        \includegraphics[width=\textwidth]{graphics/top_2loops.png}
-    \end{subfigure}
-\end{figure}
+  \includegraphics[width=\columnwidth]{graphics/flags.pdf}  
+  \caption{
+    \textbf{Parsing Algorithm of Wikipedia's XML dump}
+     The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
+Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
+Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.
+  }
+  \label{fig:parsing algorithm}
+\end{figure*}
 
-\begin{figure}[H]
+The algorithm loops in three-character chunks to account for potentially nested elements, 
+shifting by one character steps through the article markup.
+If any markup triggers for a flag are detected, a flag is raised. 
+Once a flag is raised, we stop processing and proceed to the next character
+until the flag's closing markup.
+A First Link is identified only if Flags 1, 2, and 3 are all off.
+In this case, the entire link is retrieved. 
+We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
+as well as common file extensions for 
+((cite))
+images, audio files, and the like.
+The First Link of an article is then the earliest valid link with unraised flags.
+
+To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
+across 112 cores of the UVM supercomputer cluster
+\cite{vacc}
+We then joined the results to form a hash table containing every Wikipedia article and its corresponding
+First Link. The resulting network map is the basis of our analysis.
+
+\subsection*{Supplementary Tables}
+
+\begin{figure*}[tp!]
   \centering	
-\caption{highest ranking 2-Cycles}
-    \begin{subfigure}[b]{0.8\textwidth}
-        \includegraphics[width=\textwidth]{graphics/top_3loops.png}
-    \end{subfigure}
-\end{figure}
-\begin{figure}[H]
+  \includegraphics[width=\columnwidth]{graphics/top_2loops.png}
+  \caption{
+    \textbf{highest ranking 2-Cycles}
+  }
+  \label{fig:Highest Ranking 2-Cyles}
+\end{figure*}
+
+\begin{figure*}[tp!]
   \centering	
-    \caption{Top Funnels}
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{graphics/top_funnels.png}
-    \end{subfigure}
-\end{figure}
-%========================Style Guide====================================
+  \includegraphics[width=\columnwidth]{graphics/top_3loops.png}
+  \caption{
+    \textbf{highest ranking 3-Cycles}
+  }
+  \label{fig:Highest Ranking 3-Cyles}
+\end{figure*}
 
-%\newpage
-%
-%\section*{Style Guide}
-%
-%\begin{itemize}
-%\item Web Terms
-%    \begin{itemize}
-%        \item \bt{article} instead of page
-%        \item \bt{invalid link} (instead of dead or bad link)
-%    \end{itemize}
-%\item Measurements
-%    \begin{itemize}
-%        \item \bt{traversals visits} = node weight based on number of paths crossing an article.
-%        \item \bt{path length} = number of articles traversed until an invalid link or a repeated article
-%        \item \bt{traversal funnels} = traversals before a cycle is reached.
-%    \end{itemize}
-%\item Network Properties
-%    \begin{itemize}
-%        \item \bt{cycle} = permutation or a true loop
-%        \item \bt{path-connected articles} for both cycles and articles along the same path leading to an 
-%    invalid link
-%        \item \bt{basin} a path-connected group of articles (not necessarily forming a cycle)
-%    \end{itemize}
-%\end{itemize}
-%
-%- add quotes around article names: Philosophy should be "Philosophy"
-%
-%- add thank you to RJ for pointing out the Blog Post
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/top_3loops.png}
+  \caption{
+    \textbf{highest ranking 3-Cycles}
+  }
+  \label{fig:Highest Ranking 3-Cyles}
+\end{figure*}
+
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/top_3loops.png}
+  \caption{
+    \textbf{highest ranking 3-Cycles}
+  }
+  \label{fig:Highest Ranking 3-Cyles}
+\end{figure*}
+
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
+  \caption{
+    \textbf{Highest Ranking Articles by Traversal Funnels}
+  }
+  \label{fig:Highest Ranking by Traversal Funnels}
+\end{figure*}
+
+\begin{thebibliography}{1}
+
+      \bibitem{editors} Iba, Takashi, et al.
+"Analyzing the creative editing behavior of Wikipedia editors: Through dynamic social network analysis." Procedia-Social and Behavioral Sciences 2.4 (2010): 6441-6456.
+
+      \bibitem{accuracy1} Holman Rector, Lucy. "Comparison of Wikipedia and other encyclopedias for accuracy, breadth, and depth in historical articles." Reference services review 36.1 (2008): 7-22.
+
+      \bibitem{accuracy2} Giles, Jim. "Internet encyclopaedias go head to head." Nature 438.7070 (2005): 900-901.
+
+       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. "An analysis of topical coverage of Wikipedia." Journal of Computer‐Mediated Communication 13.2 (2008): 429-440.
+
+        \bibitem{bias_women} Hill, Benjamin Mako, and Aaron Shaw.
+"The Wikipedia gender gap revisited: characterizing survey response bias with propensity score estimation." PloS one 8.6 (2013): e65782.
+
+      \bibitem{missing_entries} Williams, Jake Ryland, Clark, Eric M., Bagrow, James P., Danforth, Christopher M.
+\& Dodds, Peter Sheridan (2015).
+Identifying missing dictionary entries with frequency-conserving context models.
+Phys.
+Rev.
+E, 92, 042808.
+
+      \bibitem{clustering} Banerjee, Somnath, Krishnan Ramanathan, and Ajay Gupta.
+"Clustering short texts using wikipedia." Proceedings of the 30th annual international ACM SIGIR conference on Research and development in information retrieval.
+ACM, 2007.
+
+      \bibitem{semantic_relatedness} Gabrilovich, Evgeniy, and Shaul Markovitch. "Computing Semantic Relatedness Using Wikipedia-based Explicit Semantic Analysis." IJCAI. Vol. 7. 2007.
+
+      \bibitem{disambiguating} Cucerzan, Silviu. "Large-Scale Named Entity Disambiguation Based on Wikipedia Data." EMNLP-CoNLL. Vol. 7. 2007.
+
+      \bibitem{drugs} Clauson, Kevin A., et al. "Scope, completeness, and accuracy of drug information in Wikipedia." Annals of Pharmacotherapy 42.12 (2008): 1814-1821.
+
+      \bibitem{stats} Wikipedia contributors. "Wikipedia:Statistics." Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 9 Nov. 2015. Web. 13 Nov. 2015.
+
+      \bibitem{pagerank} Page, Lawrence, et al. "The PageRank citation ranking: bringing order to the Web." (1999).
+
+      \bibitem{classifier} Chakrabarti, Soumen, Mukul Joshi, and Vivek Tawde.
+"Enhanced topic distillation using text, markup tags, and hyperlinks." Proceedings of the 24th annual international ACM SIGIR conference on Research and development in information retrieval.
+ACM, 2001.
+
+      \bibitem{relevance} Kamps, Jaap, and Marijn Koolen. "Is Wikipedia link structure different?." Proceedings of the second ACM international conference on Web search and data mining. ACM, 2009.
+
+    \bibitem{locke} Bolton, Martha Brandt.
+"The Taxonomy of Ideas in Locke’s {\it Essay} ", The Cambridge Companion to Locke's 'Essay Concerning Human Understanding'.
+Ed.Lex Newman..
+1st ed.
+Cambridge: Cambridge University Press, 2007.
+67-100.
+Cambridge Companions Online.
+Web.
+13 November 2015.
+http://dx.doi.org/10.1017/CCOL0521834333.004.
+
+    \bibitem{aristotle} Studtmann, Paul, "Aristotle's Categories", The Stanford Encyclopedia of Philosophy (Summer 2014 Edition), Edward N.
+Zalta (ed.), \url{http://plato.stanford.edu/archives/sum2014/entries/aristotle-categories/}
+
+    \bibitem{descartes} Smith, Kurt, "Descartes' Theory of Ideas", The Stanford Encyclopedia of Philosophy (Spring 2014 Edition), Edward N.
+Zalta (ed.), \url{http://plato.stanford.edu/archives/spr2014/entries/descartes-ideas/}
+
+    \bibitem{hist_thesaurus} What is the Historical Thesaurus of the OED - Oxford English Dictionary (Oxford English Dictionary)
+    http://public.oed.com/historical-thesaurus-of-the-oed/what-is-the-historical-thesaurus-of-the-oed/
+
+    \bibitem{vacc} The authors acknowledge the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
+    \url{http://www.uvm.edu/vacc}
+
+\end{thebibliography}
 

### 2015-11-16 writeup/2015-11-wikipedia-network.tex
paper-form

@@
   \protect
-  this is the glorious abstract of abstractness summarizing methods,
-  results, and contribution of paper our understanding of the world
+Apples, oranges, and the most obscure Dylan song too---is everything a few clicks from Philosophy? 
+Within Wikipedia, the surprising answer is yes: nearly all 
+paths lead to Philosophy.
+Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
+More than information about a topic though, Wikipedia is a marvelous web of naturally emerging relationships.  
+By following the First Link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
+Here we study the English edition of Wikipedia's First Link Network for insight into how we relate topics, ideas, people, objects, and events.
+
+We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
+In a novel approach to uncover structure, we traverse every possible path through the network, 
+measuring the accumulation of First Links, path lengths, basins, cycles, and even particular articles funneling links into the cycles.
+We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general with 
+basins around fundamental notions such as Community, State, and Science. 
+Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
+These findings enrich our view of how we connect and structure
+an ever growing load of information.
 

### 2015-11-23 writeup/2015-11-wikipedia-network.tex
ndegree section updated

@@
 
-\section*{Introduction}
+\section{Introduction}
 
@@
 Wikipedia is the sixth most visited
-site in the world, surpassing 18 billion page views in a single month. 
+site in the world, surpassing 18 billion page views in a single month
+\cite{wiki_views}.
 The most widely used general reference, Wikipedia has amassed an awe-inspiring collection of human knowledge.
@@
 
-\section*{Traversing the First Link Network}
+\section{Traversing the First Link Network}
 
@@
 To more accurately gauge the distribution, we construct a log-log plot of the entire dataset: log(traversal visits) against log(rank).
-We observe a fairly linear fit, as is characteristic of scale-free networks, with an r-value of -.93.
+We observe a fairly linear fit, as is characteristic of scale-free networks, with an r-value of -.93 and
+a power-law exponent of $-0.636$.
 A handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none.
@@
 
+\subsection{Degree Distribution}
+
+We identified the set of first links directly referencing each article. 
+Similar to the number of visits, we can examine the overall distribution and 
+first links to understand the direct structure of taking one step through the 
+FLN.
+
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
+  \caption{
+    \textbf{Highest Ranking Articles by n-degree}
+  }
+  We rank each article by the number of direct first links to the article (n-degree). 
+  \label{fig:ndegree list}
+
+\end{figure*}
+The highest ranking article by n-degree, or the number of direct first links, is 
+"United States" with 80,249 direct first links. Other high-ranking articles
+include foundational abstract concepts such as "village," "species,"; 
+sports associations such as "American Football," "Association Football"; 
+and developed nations such as "France," "Japan," "Russia," "Australia," and 
+the "Netherlands."
+
+"Philosophy" and other Philosophical concepts which hold many traversal visits,
+do not appear near the top when ranked by n-degree. 
+"Philosophy" has an n-degree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W.
+Morris."
+While many articles accumulate at "Philosophy," the accumulation is not the 
+result of many articles directly referencing "Philosophy." 
+Instead, the accumulation of first links, as we argued in our 
+discussion of traversal visits, flows towards Philosophy as the 
+ultimate anchor when genralizing from specific articles towards broader notions.
+
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
+  \caption{
+    \textbf{FLN Degree Distribution}
+  }
+  We construct a log-log plot fit our results to a linear model. The result is a 
+  an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
+  The distribution appears scale free with most articles having fewer than 9 
+  direct first links, while few hold most.
+  \label{fig:degree distribution}
+
+\end{figure*}
+As a distribution, we examine the degree of the articles in the FLN with at least
+one direct first link. The average n-degree 
+is 3.6 direct links with a standard deviation of 89.5 links.
+Only 4826 articles have more than 100 direct first links, $75\%$ have fewer than 
+9. Overall the n-degree appears scale-free with an r value of -0.98 as a linear log-log fit and a power law exponent of -0.79.
+
 \subsection{Traversal Funnels}
 
-Measuring only traversal visits however is limiting as it does not distinguish whether a particular article in a cycle 
+Measuring only traversal visits and n-degree is limiting as it does not distinguish whether a particular article in a cycle 
 is a funnel, directing many more paths inside the cycle than others. 
@@
 meaningful relationships not only among ideas, but also to English-speaking society.
+
+\red{TODO: add two regimes for power-law distribution}
+
+\subsection{Article Popularity}
+
+Using a recently released api, we compare the popularity of a page to our 
+findings about the FLN. We measure popularity as the total number 
+of page views in the english edition of Wikipedia in the month of 
+October 2015 (the earliest full month for which the data is available). 
+We find \dots
 
@@
 
-\section*{Appendix}
+\section{Appendix}
 
-\subsection*{Constructing The First Link Network}
+\subsection{Constructing The First Link Network}
 
@@
 
-\subsection*{Supplementary Tables}
+\subsection{Supplementary Tables}
 
@@
 
+    \bibitem{wiki_views} Page Views for Wikipedia, Non-mobile site, Normalized (Page Views for Wikipedia, Non-mobile site, Normalized)
+    \url{https://stats.wikimedia.org/EN/TablesPageViewsMonthly.htm}
 \end{thebibliography}

### 2015-11-23 writeup/2015-11-wikipedia-network.tex
article popularity and traversal visits

@@
 
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/top_1k_path_length.png}  
-  \caption{
-    \textbf{Path Length Distribution for Top 1k articles}
-  }
-  \label{fig:Top 1k Path Length Distribution}
-\end{figure*}
-
 \subsection{Degree Distribution}
@@
 \end{figure*}
-The highest ranking article by n-degree, or the number of direct first links, is 
+The highest ranking article by n-degree is 
 "United States" with 80,249 direct first links. Other high-ranking articles
@@
 
-"Philosophy" and other Philosophical concepts which hold many traversal visits,
-do not appear near the top when ranked by n-degree. 
+"Philosophy" and other philosophical concepts with many traversal visits
+are not high-ranking by n-degree.
 "Philosophy" has an n-degree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W.
@@
 meaningful relationships not only among ideas, but also to English-speaking society. 
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\textwidth]{graphics/funnels_distribution.png}
+  \caption{
+    \textbf{Distribution of Traversal Funnels}
+  }
+  \label{fig:Funnels Distribution}
+\end{figure*}
 
-\red{TODO: add two regimes for power-law distribution}
+\red{TODO: add discussion of distribution}
 
@@
 October 2015 (the earliest full month for which the data is available). 
-We find \dots
+We find the highest ranking 1000 articles by traversal visits have an average of
+$70$ thousand page views in October with high variation: the standard deviation 
+is $1.1$ million page views. 
+The number of page views has a skewed distrubtion with $75\%$ of articles
+reaching fewer than $73$ thousand page views in October.
+The article for "United States" has the most page 
+views of the highest ranking $1000$ articles by traversal visits with 
+$1.4$ million views in October. The next most popular articles are 
+"India," "World War II," and the "United Kingdom" each with $0.7-1.03$ million page views.
 
+The "Philosophy" article despite outranking every article in traversal visits,
+has only $240$ thousand page views in October compared to the $1.4$ million
+page views "United States" has.
+While "Philosophy" is not itself the most popular article, "Philosophy" serves as 
+an accumulation points for the ideas. Wikipedia users do not login specificially
+to read about "Philosophy," but the authors relate ideas to "Philosophy" 
+by ultimately anchoring the flow of first links to "Philosophy."
+
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/views_visits.png}
+  \caption{
+    \textbf{Article Popularity (by page views) compared to Traversal Visits 
+    for Highest ranking 1000 articles}
+    We use the total number of page views provided by Wikipedia for the month
+    of October 2015 to compare each article's popularity to traversal visits.
+    While the most popular articles do not necessarily correspond to the articles
+    with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
+  }
+  \label{fig:Views and Visits}
+
+\end{figure*}
+
+\red{REDO analysis}
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/funnels_visits.png}
+  \caption{
+    \textbf{Article Popularity (by page views) compared to Traversal Funnels
+    for Highest ranking 1000 articles}
+  }
+    We use the total number of page views provided by Wikipedia for the month
+    of October 2015 to compare each article's popularity to traversal funnels.
+  \label{fig:Views and Funnels}
+\end{figure*}
 %========================Conclusion====================================

### 2015-11-27 writeup/2015-11-wikipedia-network.tex
traversal algo expanded explanation

@@
 The authors of a Wikipedia article choose where and whether to include a 
-hyperlink reference to another Wikipedia article.
-Authors must include hyperlinks in the article's markup.
+hyperlink reference to another Wikipedia article
+by including hyperlinks in the article's markup.
 In describing a particular topic, each set of authors decides which articles are the most
@@
 to the influence an article exerts in shaping the FLN.
-Together with cycles, n-degree, depth, and the content of the articles, 
+Together with cycles, ndegree, depth, and the content of the articles, 
 we build our analysis of the relations among the ideas in Wikipedia.
@@
 
-To understand the structure of Wikipedia's First Link Network, we
-aimed to characterize the dynamics of the flow from one article to another. 
-Do links tend towards a particular article, group of articles, or groups of articles? 
-What is the flow of links through a typical article?
-What types of cycle (loop) structures exist in the network?
-What are typical paths from one article to loop or an invalid link? 
-Are there exceptionally long or short paths? 
-Are there articles funneling the flow of links towards a particular path?
-In answering these questions about the directed network, we aimed to characterize 
-how so many independent articles relate to each other.
+One essential measure of a directed network structure is the degree distribution 
+\cite{newman}. 
+Degree distribution has been used to study many phenomena from disease outbreak 
+\cite{disease} 
+to the dynamics of social networks 
+\cite{social_nets}.
+The degree distribution in the FLN describes how many first links point to a particular article. 
+We measure the indegree in the FLN to understand which articles authors choose to reference in 
+the first link.
+A higher indegree indicates more authors selected the particular article
+as a relevant reference in describing their topic. On the other hand, articles with zero 
+indegree have no references---they are outer leaves in the network. 
+Similar to PageRank, the indegree is also a way to rank the articles in Wikipedia.
+We can rank articles by indegree to see which articles are most and least referenced in the FLN. 
+Additionally, we can assess the degree distribution to see whether authors tend to reference all articles
+equally or only a few, as observed in many social networks \cite{social_nets}.
 
-The n-degree, or number of links directly pointing to a particular article,
-inadequately captures the dynamics needed to answer these questions.
-The n-degree measures only the particular set of links to a particular article, 
-rather than the richer dynamics of how the links flow through the network: 
-where articles tend, what the typical and atypical 
-paths through the network are, which articles funnel relatively more links and so on.
+The degree distribution captures one step in the FLN, measuring   
+the destination of a given link. 
+Indegree measures only the particular set of first links between two articles, 
+rather than the richer dynamics of how many articles fit together---how links 
+create a flow through the FLN.
+One way to expand the analysis beyond a single step is to map the full path of 
+an article's first link. 
+Each article is both a destination and a link to another article, similar to an element 
+in a linked list data structure. 
+By following the chain of links we can construct a path through the network with each article 
+leading to the next via the first link. 
+For example "Train" has a first link to "Conveyance of Passengers and Goods," which is itself
+an article with a first link to "Goods" and so on. The path starting at "Train" 
+is the collection of articles obtained by following the first links starting at "Train."
+The path contains "Goods," "Economics," "Social Science," and other articles related to 
+"Train," but not directly referenced by a first link.
 
-Consequently, to capture the dynamics of how the First Links flow, we actually traverse every possible path through the FLN. 
-The algorithm for traversing the network begins by selecting any article. 
-We then proceed to the next article by following the First Link---recalling a First Link is a link in the main body of the article leading to another Wikipedia article. 
-We repeat until the First Link is invalid or repeated to form a path. 
+Applying a similar concept with a 
+different starting article, we form 
+a path through the FLN moving from one article to the next.
+A path in the FLN connects a group of articles allowing us to measure how two articles relate,
+even if the articles are not directly linked. 
 The collection of articles is path-connected conveying a flow of concepts from one article to another. 
-The method is order agnostic with respect to which articles are selected first.
-As long as each article is selected eventually, the resulting metrics are equivalent.
+The notion of paths allows us to map not only which articles link directly to "Philosophy" 
+for example, but also which ideas are eventually anchored in "Philosophy" even if the articles
+are not directly linked.
+The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
+By constructing paths through the FLN, we 
+map how the first links flow through the FLN.
 
-We developed three metrics to uncover structure: 
-traversal visits, traversal funnels, and path length. 
-Each metric captures one essence of how the links flow. 
-Traversal visits gauge the accumulation of links; 
-traversal funnels gauge the influence on link path; 
-path length measures the number of First Links to a repeated or invalid link.
+In a novel approach, we use the notion of a path through the FLN to describe 
+the structure of the FLN as a flow relating ideas. 
+Grounded in the notion of flow through river networks,
+\cite{geo_basins}
+we characterize features of the FLN as a flow via first links.
+Previous studies have used flow to characterize the structure of river networks
+\cite{dodds} and describe the organization of food systems as transportation networks
+\cite{food_webs}.
+With paths to mark flow through the FLN, we develop new metrics to 
+guage the accumulation of references, 
+the length of the path relating articles, and the influence a particular
+article exerts in shaping the flow of links through the FLN.
 
@@
 
-The first metric, {\it traversal} visits, gauges where the flow of first accumulates by associating a count to each article. 
-Every path through the article increments the associated count by one. We can construct a vector to represent the articles incremented on a particular path, combining these vectors to form a matrix. 
-The total accumulated counts for a given article is then simply the sum of the row in our matrix corresponding to the article of interest. 
-Node A in our example (see figure ~\ref{fig:Traversal Visits}) has 7 traversal visits, since every path includes article A. 
-Whereas, node E only has two traversal visits since only two paths contain article E for example.
-What we have then is a metric signaling which article contains a larger accumulation of first links.
+Our goal is to map the flow of first links by following the paths of first links 
+starting at each article. 
+Since a path reveals how the ideas in one article may eventually be 
+anchored in another, the first metric we develop quantifies the accumulation of first links 
+within the FLN. 
+The algorithm begins by selecting an article then traversing the path formed
+by following the first links. Each time a first link references an article, we increment a count
+associated with the article. 
+We repeat until the first link is invalid (outside the FLN) or repeated. The collection of articles is path-connected conveying a flow of concepts from one article to another. 
+We select a second article and repeat the process until we have 
+constructed a path for each article in the network. We call the resulting count associated with each article the number of {\it traversal visits}. The number of traversal visits of an article 
+measures the number of references flowing to the ideas in the article.
 
-While traversal visits measure accumulation, we cannot gauge whether a particular article exerts a greater influence on flow of the first link paths. 
-To distinguish between an article that simply happened to fall within a cycle from an article actually funneling the flow of first links, we developed a second metric called traversal funnels. 
-We traverse the network in the same manner, but end our paths once we enter a cycle. 
-We again increment the associated count of each article by one if the path up to a cycle contains the article. 
-We call the accumulated count the number of {\it traversal funnels}. By measuring traversal funnels, we quantify
-which articles funnel more link paths in a particular direction or cycle 
-(see figure ~\ref{fig:Traversal Funnels})
-.
+We can map the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
+(see figure~\ref{fig:Traversal Visits}), the path starting at article A is the 
+first column in the traversal vists matrix with $1$ if the path contains a given article and 
+$0$ if the path doesn't. The second column indicates the path starting at article B and so on. 
+To compute the number traversal visits for article A we sum the corresponding row in our matrix,
+which totals the number of paths containing article A. The total is a analogous to the flow 
+in a river network with traversal visists measuring the accumulation of water near point A. 
+By constructing a traversal visits matrix, we can similarly compute the number of traversal 
+visits by summing along the corresponding row for our node. The traversal visits matrix for 
+Wikipedia consists 121 million entries to account for each path through the FLN. 
+We measure accumulation of first link references by summing the corresponding row for the 
+article of interest.
+
+In addition to accumulation, the traversal visits algorithm creates paths 
+which uncover depth in the FLN. By measuring the number of first links between
+two articles, we get an additional piece of information we call 
+{\it path length.} We can compute the length a path by summing 
+along a column in our traversal visits matrix.
+The path length describes how closely related ideas are, guaging the FLN's depth.
+The greater the number of first links separating two articles, the greater the number of ideas needed to relate the ideas.
+Although,
+"Train" is related to "Economics," there are several more specific articles
+bridging the connection:
+"Train" is more specifically related to transportation, whose object
+is often goods. Goods are one of the fundamental objects of study in Economics.
+Described in links, this relationship is captured by a relatedness of $4$ 
+first links ultimately connecting "Train" to "Economics." 
+We can examine path length as a relative measure of how closely connected ideas
+may be, but also as an aggregate measure describing the overall separation or 
+connectedness in the FLN.
+We can characterize whether ideas a closely connected by a handful of links, 
+many, or whether a more nuianced organization exists.
+
+One possible path through the FLN is a {\it cycle} or a group of articles 
+linking to one another inside a loop. In our sample network 
+(see figure~\ref{fig:Traversal Visits}) 
+a cycle exists among nodes A, B, and C. Since node C links back to the starting node A, the path starting at node A forms a {\it 3-cycle}. 
+By recording the history of the articles along a path in the FLN, we can identify
+the types of cycle structures within the FLN. Furthermore, since each article 
+in the a cycle has an associated traversal visit count, we can rank cycles by 
+the number of references directed towards a cycle. 
+We can also form and rank {\it basins} in the network by identifying groups of 
+path-connected articles, not necessarily forming a perfect cycle.
+A basin connects a group of articles which ultimately reference other articles
+in the FLN. By identifying basins and cycles, we pinpoint group of closely related articles.
 
@@
 
-Finally, we track the number of links until a repeated article or invalid link, to find the typical path length within the network.
-By recording the history of articles we traversed, we are also able to uncover several other network characteristics including basins, defined as a collection of path-connected articles ranked by traversal visits, and cycles of various lengths.
-The three metrics: traversal visits, traversals funnels, and path length, along with our path history
-yield a powerful arsenal of information with which to study the network.
+While traversal visits measure accumulation, each article's first link also 
+influences the shape of the FLN. 
+At a large point of accumulation, a single article's first link 
+can exert great influence over the shape of the FLN by directing many
+references towards a particular path. To distinguish between an article 
+that simply happened to fall within a cycle from an article funneling 
+many first links, we develop a second metric called {\it traversal funnels}.
+
+To measure traversal funnels, we traverse the FLN in the same manner as we 
+did for traversal visits, but end a path once we enter a cycle.
+We are then able to distinguish between an article related to many other ideas
+only by virtue of its place in a cycle from an article exerting influence over where the first links flow. 
+An article with a large number of traversal funnels directs many references
+towards a particular path. In our sample network 
+(see figure~\ref{fig:Traversal Funnels}) article C 
+directs the flow of links towards the 3-cycle, while articles A and B are 
+receipients of the flow---without exerting any influence themselves. 
+While articles A, B, and C all have the same number of traversal visits, 
+article C has 4 traversal funnels (A and B have none). By incrementing the count of 
+first link references on paths only up to a cycle, we can identifying which articles
+exert the greatest influence over the network's structure.
+
+By studying the FLN not only as collection of directly linked pairs of articles, but
+as a flow of paths, 
+we build a powerful arsenal of information with which to study the FLN. 
+From accumulated references, cycles, basins to influence, we can measure how the many articles in Wikipedia are organized and related.
 
@@
 
+%\green{TODO:
+%rewrite \\
+%introduce metric\\
+%find two or three salient points to discuss in each section \\
+%tie into bigger picture
+%\begin{itemize}
+%    \item reorder so paths and flow are described first then funnels and visits
+%    \item Basins expand explanation and break up into 2 paragraphs
+%    \item add more captions to figures
+%\end{itemize}
+%}
 We followed every possible path through the network, taking 232 million steps along the way to measure the accumulated number of traversal visits.
@@
 Banana has fewer synonyms than botanical for example.
-To quantify the observed flow, we measured the number of synonyms the article title contained in WordNet---the largest lexical database of the English language.
+To quantify the observed flow, we measured the number of synonyms the article title contained in WordNet---the largest lexical database of the English language
+\cite{wordnet}. 
 We find the highest ranking 100 articles have on average 5 more synonyms than the typical article; a difference of 2.5 synonyms if we compare the median number of synonyms in each group. 
@@
 cycle of 6 other articles including "Experiment", "Fact", and "Knowledge".
-To confirm the suggested cycle structure, we recorded the history of articles traversed along a path.
+To confirm the suggested cyclic structure, we record the history of articles traversed along a path. For example starting on "Train" we construct an path of articles: 
+"Rain Transport,"
+"Conveyance of Passengers and Goods," 
+"Goods," 
+"Economics,"
+"Social Science,"
+"Academic Disciplines,"
+"Knowledge,"
+"Awareness,"
+"Consciousness,"
+"Quality,"
+"Philosophy."
 
@@
 
-\subsection{Path Length}
+\subsection{Depth of the FLN}
 
@@
 of cycles. 
-Path length measures the number of links traversed until a repeated or invalid link. 
+Path length measures the number of links traversed until a repeated or invalid link as a means to guage the network's depth.
 We discovered the longest path length is also 365, matching the longest cycle of Orthodox Liturgics. 
@@
 The most common path length is 29, with an interquartile range (26, 30).
+The median path length far exceeds the popular 6-degrees of separation (see for 
+example the average number of academic publications separating scholars 
+\cite{six_degrees}), indicating a greater network depth. The network depth suggests a large network of spanning a variety of ideas, which Wikipedia certainly is.
 The distribution of path lengths is similarly scale-free with few articles at the extreme of 365 path lengths, while the majority 
-is between 26-30:
+is between 26-30.
 
@@
   }
+  The median network depth measured by the number of first links in a path
+  is 29. The 365 cycle for "Orthodox" liturgics is the outlier to the right, while
+  other historical articles about "Scotland," "UK" and so on are slightly outside
+  the third quartile. More than $75\%$ of articles have path lenghts between 
+  $0$ and $50$ links.
+
   \label{fig:Path Length Distribution}
@@
   \caption{
-    \textbf{Highest Ranking Articles by n-degree}
+    \textbf{Highest Ranking Articles by ndegree}
   }
-  We rank each article by the number of direct first links to the article (n-degree). 
+  We rank each article by the number of direct first links to the article (ndegree). 
   \label{fig:ndegree list}
@@
 \end{figure*}
-The highest ranking article by n-degree is 
+The highest ranking article by ndegree is 
 "United States" with 80,249 direct first links. Other high-ranking articles
@@
 "Philosophy" and other philosophical concepts with many traversal visits
-are not high-ranking by n-degree.
-"Philosophy" has an n-degree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W.
+are not high-ranking by ndegree.
+"Philosophy" has an ndegree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W.
 Morris."
@@
 As a distribution, we examine the degree of the articles in the FLN with at least
-one direct first link. The average n-degree 
+one direct first link. The average ndegree 
 is 3.6 direct links with a standard deviation of 89.5 links.
 Only 4826 articles have more than 100 direct first links, $75\%$ have fewer than 
-9. Overall the n-degree appears scale-free with an r value of -0.98 as a linear log-log fit and a power law exponent of -0.79.
+9. Overall the ndegree appears scale-free with an r value of -0.98 as a linear log-log fit and a power law exponent of -0.79.
 
@@
 
-Measuring only traversal visits and n-degree is limiting as it does not distinguish whether a particular article in a cycle 
+Measuring only traversal visits and ndegree is limiting as it does not distinguish whether a particular article in a cycle 
 is a funnel, directing many more paths inside the cycle than others. 
@@
 Measuring traversal funnels reveals a dramatically different structure where "Philosophy" stands unmatched by orders of magnitude:
+
 \begin{figure*}[tp!]
@@
   \centering	
-  \includegraphics[width=\textwidth]{graphics/funnels_distribution.png}
+  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
   \caption{
@@
 
-\red{TODO: add discussion of distribution}
-
 \subsection{Article Popularity}
 
-Using a recently released api, we compare the popularity of a page to our 
-findings about the FLN. We measure popularity as the total number 
-of page views in the english edition of Wikipedia in the month of 
-October 2015 (the earliest full month for which the data is available). 
+Wikipedia released an API to measure article popularity by page views
+starting November 2015. The page views adds another dimension to our
+findings about structure by revealing 
+how users view the information on Wikipedia. 
+We measure popularity as the total number 
+of page views in the English edition of Wikipedia in the month of 
+October 2015---the earliest full month for which the data is available. 
 We find the highest ranking 1000 articles by traversal visits have an average of
@@
 is $1.1$ million page views. 
-The number of page views has a skewed distrubtion with $75\%$ of articles
+The number of page views has a skewed distribution with $75\%$ of articles
 reaching fewer than $73$ thousand page views in October.
@@
 has only $240$ thousand page views in October compared to the $1.4$ million
-page views "United States" has.
+page views for "United States".
 While "Philosophy" is not itself the most popular article, "Philosophy" serves as 
-an accumulation points for the ideas. Wikipedia users do not login specificially
+an accumulation points for the ideas. Wikipedia users do not login specifically
 to read about "Philosophy," but the authors relate ideas to "Philosophy" 
@@
 
-\red{REDO analysis}
+We also analyze article popularity for the highest ranking 1000 articles by 
+traversal funnels. In October, the average page views per article is 
+$22$ thousand with a standard deviation of $95$ thousand views. The distribution
+is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
+"Halloween" is the most popular article with $2.8$ million views in October,
+likely a result of Halloween falling in the month of October. 
+Other popular articles include "Scientology" ($463$ thousand views), "Clint Eastwood" 
+($341$ thousand views),
+and the "Cold War" ($298$ thousand views) although each has significantly fewer views compared to the views for "Halloween."
+Other standout popular October articles include "24-hour" and "12-hour" clocks likely due
+to Daylight savings in October and "Marriage" likely due to the drop in the number of
+weddings in the months following October
+\cite{weddings}.
+"Philosophy" is ranks is the seventh most popular of the highest ranking articles by traversal funnels, but appears no where in the top 20 among the highest ranking
+articles by traversal visits. "Philosophy" is a relatively popular article 
+among articles influencing the shape of the network, but less popular among 
+articles with a large accumulation among the flow of first links.
+
 \begin{figure*}[tp!]
@@
     of October 2015 to compare each article's popularity to traversal funnels.
+    In the top left is "Halloween" with a significantly greater 
+    number of page views in the month of October.
   \label{fig:Views and Funnels}
 \end{figure*}
+
 %========================Conclusion====================================
@@
   \caption{
-    \textbf{highest ranking 3-Cycles}
+    \textbf{highest ranking 3-cycles}
   }
-  \label{fig:Highest Ranking 3-Cyles}
-\end{figure*}
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/top_3loops.png}
-  \caption{
-    \textbf{highest ranking 3-Cycles}
-  }
-  \label{fig:Highest Ranking 3-Cyles}
-\end{figure*}
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/top_3loops.png}
-  \caption{
-    \textbf{highest ranking 3-Cycles}
-  }
-  \label{fig:Highest Ranking 3-Cyles}
+  \label{fig:Highest Ranking 3-cyles}
 \end{figure*}
@@
     \url{https://stats.wikimedia.org/EN/TablesPageViewsMonthly.htm}
+
+    \bibitem{weddings} Wedding statistics in the United States (Wedding statistics in the United States)
+    \url{http://www.soundvision.com/article/wedding-statistics-in-the-united-states}
+
+    \bibitem{six_degrees} Elmacioglu, Ergin, and Dongwon Lee. "On six degrees of separation in DBLP-DB and more." ACM SIGMOD Record 34.2 (2005): 33-40.
+
+    \bibitem{wordnet} Princeton University "About WordNet." WordNet. Princeton University. 2010. 
+    \url{http://wordnet.princeton.edu}
+
+    \bibitem{newman} Newman, Mark EJ. "The structure and function of complex networks." SIAM review 45.2 (2003): 167-256.
+
+    \bibitem{disease} Eubank, Stephen, et al. "Modelling disease outbreaks in realistic urban social networks." Nature 429.6988 (2004): 180-184.
+
+    \bibitem{social_nets} Newman, Mark EJ, Duncan J.
+Watts, and Steven H.
+Strogatz.
+"Random graph models of social networks." Proceedings of the National Academy of Sciences 99.suppl 1 (2002): 2566-2572.
+
+    \bibitem{geo_basins} Horton, Robert E.
+"Erosional development of streams and their drainage basins; hydrophysical approach to quantitative morphology." Geological society of America bulletin 56.3 (1945): 275-370.
+    APA
+
+    \bibitem{dodds} Dodds, Peter Sheridan, and Daniel H. Rothman. "Unified view of scaling laws for river networks." Physical Review E 59.5 (1999): 4865.
+
+    \bibitem{food_webs} Garlaschelli, Diego, Guido Caldarelli, and Luciano Pietronero. "Universal scaling relations in food webs." Nature 423.6936 (2003): 165-168.
+
 \end{thebibliography}

### 2015-11-27 writeup/2015-11-wikipedia-network.tex
fixed ndegree

@@
 to the influence an article exerts in shaping the FLN.
-Together with cycles, ndegree, depth, and the content of the articles, 
+Together with cycles, indegree, depth, and the content of the articles, 
 we build our analysis of the relations among the ideas in Wikipedia.
@@
 
-%\green{TODO:
-%rewrite \\
-%introduce metric\\
-%find two or three salient points to discuss in each section \\
-%tie into bigger picture
-%\begin{itemize}
-%    \item reorder so paths and flow are described first then funnels and visits
-%    \item Basins expand explanation and break up into 2 paragraphs
-%    \item add more captions to figures
-%\end{itemize}
-%}
+\green{TODO:
+rewrite \\
+introduce metric\\
+find two or three salient points to discuss in each section \\
+tie into bigger picture
+\begin{itemize}
+    \item reorder so paths and flow are described first then funnels and visits
+    \item Basins expand explanation and break up into 2 paragraphs
+    \item add more captions to figures
+\end{itemize}
+}
+
+\subsection{Degree Distribution}
+
+We identified the set of first links directly referencing each article. 
+Similar to the number of visits, we can examine the overall distribution and 
+first links to understand the direct structure of taking one step through the 
+FLN.
+
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
+  \caption{
+    \textbf{Highest Ranking Articles by indegree}
+  }
+  We rank each article by the number of direct first links to the article (indegree). 
+  \label{fig:ndegree list}
+
+\end{figure*}
+The highest ranking article by indegree is 
+"United States" with 80,249 direct first links. Other high-ranking articles
+include foundational abstract concepts such as "village," "species,"; 
+sports associations such as "American Football," "Association Football"; 
+and developed nations such as "France," "Japan," "Russia," "Australia," and 
+the "Netherlands."
+
+"Philosophy" and other philosophical concepts with many traversal visits
+are not high-ranking by indegree.
+"Philosophy" has an indegree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W.
+Morris."
+While many articles accumulate at "Philosophy" (see traversal visits discussion), 
+the accumulation is not the 
+result of many articles directly referencing "Philosophy." 
+Instead, the accumulation of first links, as we argue in our 
+discussion of traversal visits, flows towards Philosophy as the 
+ultimate anchor when genralizing from specific articles towards broader notions.
+
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
+  \caption{
+    \textbf{FLN Degree Distribution}
+  }
+  We construct a log-log plot fit our results to a linear model. The result is a 
+  an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
+  The distribution appears scale free with most articles having fewer than 9 
+  direct first links, while few hold most.
+  \label{fig:degree distribution}
+\end{figure*}
+
+As a distribution, we examine the degree of the articles in the FLN with at least
+one direct first link. The average indegree 
+is 3.6 direct links with a standard deviation of 89.5 links.
+Only 4826 articles have more than 100 direct first links, $75\%$ have fewer than 
+9. Overall the indegree appears scale-free with an r value of -0.98 as a linear log-log fit and a power law exponent of -0.79.
+
+\subsection{Traversal Visits}
+
 We followed every possible path through the network, taking 232 million steps along the way to measure the accumulated number of traversal visits.
-
-\subsection{Traversal Visits}
 
@@
 The quantifiable difference in synonyms corroborates the flow of links from concrete, specific articles to broader disciplines or fundamental notions.
+
+\subsection{Depth of the FLN}
+
+In addition to identifying cycles and cycle lengths, we also measure the traversal path length, which includes articles outside
+of cycles. 
+Path length measures the number of links traversed until a repeated or invalid link as a means to guage the network's depth.
+We discovered the longest path length is also 365, matching the longest cycle of Orthodox Liturgics. 
+We also found similarly lengthy paths following the evolution of a place or topic through time: 
+"1953 in Scotland" or "1560s Architecture", with articles sequentially proceeding by year, decade or era.
+
+Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
+The most common path length is 29, with an interquartile range (26, 30).
+The median path length far exceeds the popular 6-degrees of separation (see for 
+example the average number of academic publications separating scholars 
+\cite{six_degrees}), indicating a greater network depth. The network depth suggests a large network of spanning a variety of ideas, which Wikipedia certainly is.
+The distribution of path lengths is similarly scale-free with few articles at the extreme of 365 path lengths, while the majority 
+is between 26-30.
+
+\begin{figure}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/path_lengths_boxplot.png}
+  \caption{
+    \textbf{Path Length Distribution}
+  }
+  The median network depth measured by the number of first links in a path
+  is 29. The 365 cycle for "Orthodox" liturgics is the outlier to the right, while
+  other historical articles about "Scotland," "UK" and so on are slightly outside
+  the third quartile. More than $75\%$ of articles have path lenghts between 
+  $0$ and $50$ links.
+
+  \label{fig:Path Length Distribution}
+\end{figure}
 
@@
 
-\subsection{Depth of the FLN}
-
-In addition to identifying cycles and cycle lengths, we also measure the traversal path length, which includes articles outside
-of cycles. 
-Path length measures the number of links traversed until a repeated or invalid link as a means to guage the network's depth.
-We discovered the longest path length is also 365, matching the longest cycle of Orthodox Liturgics. 
-We also found similarly lengthy paths following the evolution of a place or topic through time: 
-"1953 in Scotland" or "1560s Architecture", with articles sequentially proceeding by year, decade or era.
-
-Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
-The most common path length is 29, with an interquartile range (26, 30).
-The median path length far exceeds the popular 6-degrees of separation (see for 
-example the average number of academic publications separating scholars 
-\cite{six_degrees}), indicating a greater network depth. The network depth suggests a large network of spanning a variety of ideas, which Wikipedia certainly is.
-The distribution of path lengths is similarly scale-free with few articles at the extreme of 365 path lengths, while the majority 
-is between 26-30.
-
-\begin{figure}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/path_lengths_boxplot.png}
-  \caption{
-    \textbf{Path Length Distribution}
-  }
-  The median network depth measured by the number of first links in a path
-  is 29. The 365 cycle for "Orthodox" liturgics is the outlier to the right, while
-  other historical articles about "Scotland," "UK" and so on are slightly outside
-  the third quartile. More than $75\%$ of articles have path lenghts between 
-  $0$ and $50$ links.
-
-  \label{fig:Path Length Distribution}
-\end{figure}
-
-\subsection{Degree Distribution}
-
-We identified the set of first links directly referencing each article. 
-Similar to the number of visits, we can examine the overall distribution and 
-first links to understand the direct structure of taking one step through the 
-FLN.
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
-  \caption{
-    \textbf{Highest Ranking Articles by ndegree}
-  }
-  We rank each article by the number of direct first links to the article (ndegree). 
-  \label{fig:ndegree list}
-
-\end{figure*}
-The highest ranking article by ndegree is 
-"United States" with 80,249 direct first links. Other high-ranking articles
-include foundational abstract concepts such as "village," "species,"; 
-sports associations such as "American Football," "Association Football"; 
-and developed nations such as "France," "Japan," "Russia," "Australia," and 
-the "Netherlands."
-
-"Philosophy" and other philosophical concepts with many traversal visits
-are not high-ranking by ndegree.
-"Philosophy" has an ndegree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W.
-Morris."
-While many articles accumulate at "Philosophy," the accumulation is not the 
-result of many articles directly referencing "Philosophy." 
-Instead, the accumulation of first links, as we argued in our 
-discussion of traversal visits, flows towards Philosophy as the 
-ultimate anchor when genralizing from specific articles towards broader notions.
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
-  \caption{
-    \textbf{FLN Degree Distribution}
-  }
-  We construct a log-log plot fit our results to a linear model. The result is a 
-  an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
-  The distribution appears scale free with most articles having fewer than 9 
-  direct first links, while few hold most.
-  \label{fig:degree distribution}
-
-\end{figure*}
-As a distribution, we examine the degree of the articles in the FLN with at least
-one direct first link. The average ndegree 
-is 3.6 direct links with a standard deviation of 89.5 links.
-Only 4826 articles have more than 100 direct first links, $75\%$ have fewer than 
-9. Overall the ndegree appears scale-free with an r value of -0.98 as a linear log-log fit and a power law exponent of -0.79.
-
 \subsection{Traversal Funnels}
 
-Measuring only traversal visits and ndegree is limiting as it does not distinguish whether a particular article in a cycle 
+Measuring only traversal visits and indegree is limiting as it does not distinguish whether a particular article in a cycle 
 is a funnel, directing many more paths inside the cycle than others. 

### 2015-11-29 writeup/2015-11-wikipedia-network.tex
backup

@@
 
-We identified the set of first links directly referencing each article. 
-Similar to the number of visits, we can examine the overall distribution and 
-first links to understand the direct structure of taking one step through the 
-FLN.
-
+We first identify the set of first links directly referencing a particular article 
+by studying the indegree of the FLN. 
+A higher indegree suggests more authors found the ideas in the article 
+a pertinent direct reference. 
 \begin{figure*}[tp!]
@@
   }
-  We rank each article by the number of direct first links to the article (indegree). 
+  We rank each article by the number of direct first links to the article (indegree). The highest-ranking articles tend to represent abstraction for concepts
+  comprised of various facets and parts.
   \label{fig:ndegree list}
-
-\end{figure*}
-The highest ranking article by indegree is 
-"United States" with 80,249 direct first links. Other high-ranking articles
+\end{figure*}
+We rank all 11 million articles by indegree to find 
+the "United States" with 80,249 direct first links as the most referenced
+Wikipedia article. 
+Other high-ranking articles
 include foundational abstract concepts such as "village," "species,"; 
@@
 and developed nations such as "France," "Japan," "Russia," "Australia," and 
-the "Netherlands."
+the "Netherlands." These high ranking articles are useful abstractions: nations
+describe a collection of individuals, a common culture, language, or 
+geographical proximity; sports teams describe an ever changing collection of 
+sports players often associated with a cultural identity and a geographical 
+region. Since abstractions such as nations and teams are inherently comprised
+of many parts, authors reference the larger abstraction when describing 
+a part. For example, in describing the geographical region of north America, 
+the United States, Canada, or Mexico are natural references anchoring the 
+geographical location to a larger national identity.
 
 "Philosophy" and other philosophical concepts with many traversal visits
-are not high-ranking by indegree.
+are not among the highest-ranking articles by indegree.
 "Philosophy" has an indegree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W.
@@
 ultimate anchor when genralizing from specific articles towards broader notions.
-
 \begin{figure*}[tp!]
@@
 \end{figure*}
-
-As a distribution, we examine the degree of the articles in the FLN with at least
-one direct first link. The average indegree 
-is 3.6 direct links with a standard deviation of 89.5 links.
-Only 4826 articles have more than 100 direct first links, $75\%$ have fewer than 
-9. Overall the indegree appears scale-free with an r value of -0.98 as a linear log-log fit and a power law exponent of -0.79.
+The FLN's indegree exihibits a scale-free distribution where a few articles 
+receive a most direct first references, while most articles receive few or none.
+The average indegree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5 links.
+Only 4826 articles have more than 100 direct first links and $75\%$ of articles
+have fewer than 9. 
+When fit to a linear model on a log-log scale (log(rank) versus log(indegree)), 
+the model's r value is $-0.98$ suggesting a strong log-log linear fit 
+with a power law exponent of $-0.79$.
+
+\centerline{***}
+
+Beyond indegree, our work expands the analysis of relations to understand 
+how many more than two articles are organized and related. 
+We develop the notion of a path to describe the structure of the FLN as a 
+flow relating ideas. We traverse every possible path through the FLN, taking 
+$232$ million steps along the way.
+
+\subsection{Depth of the FLN}
+
+We first seek to describe the depth of the FLN: how many links does a 
+connection of ideas span? We gauge the FLN's depth via path length or the number of links traversed until a repeated or invalid link (one outside the FLN). 
+We discover the longest path length is 365
+corresponding to the yearly calendar of Orthodox Liturgics.
+Each day's Liturgics link to the next day's. Curiously, on the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
+(see discussion of cycles).
+We also found similarly lengthy paths following the evolution of a place or topic through time: 
+"1953 in Scotland" or "1560s Architecture", with articles sequentially proceeding by year, decade or era.
+The longest paths connecting ideas are organized temporally.
+
+\begin{figure}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/path_lengths_boxplot.png}
+  \caption{
+    \textbf{Path Length Distribution}
+  }
+  The median network depth measured by the number of first links in a path
+  is 29. The 365 cycle for "Orthodox" liturgics is the outlier to the right, while
+  other historical articles about "Scotland," "UK" and so on are slightly outside
+  the third quartile. More than $75\%$ of articles have path lengths between 
+  $0$ and $50$ links.
+
+  \label{fig:Path Length Distribution}
+\end{figure}
+
+Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
+This roughly corresponds to the official number of articles on Wikipedia 
+$~4.7$ million as of November 2014---approximately half of the 11 million 
+articles in the XML dump are redirects or disambiguations, not full articles.
+The most common path length is 29, with an interquartile range (26, 30).
+The median path length far exceeds the popular 6-degrees of separation (see for 
+example the average number of academic publications separating scholars 
+\cite{six_degrees}), indicating a greater network depth. The network depth suggests a large network of spanning a variety of ideas.
+As a distribution, more than $75\%$ of articles have a path length below 
+$50$ first links 
+while a few temporally organized paths exceed 50 links 
+(see figure~\ref{fig:Path Length Distribution}). 
+How can we characterize the flow of ideas along a path? Next, we develop 
+a metric to gauge the accumulation first link references
+and characterize a global organization 
+of ideas among articles.
 
@@
 
-We followed every possible path through the network, taking 232 million steps along the way to measure the accumulated number of traversal visits.
+We followed every possible path through the network, incrementing
+the number of traversal visits for every first link reference along a path 
+to generate a total of 232 million traversal visits.
 
@@
 The quantifiable difference in synonyms corroborates the flow of links from concrete, specific articles to broader disciplines or fundamental notions.
-
-\subsection{Depth of the FLN}
-
-In addition to identifying cycles and cycle lengths, we also measure the traversal path length, which includes articles outside
-of cycles. 
-Path length measures the number of links traversed until a repeated or invalid link as a means to guage the network's depth.
-We discovered the longest path length is also 365, matching the longest cycle of Orthodox Liturgics. 
-We also found similarly lengthy paths following the evolution of a place or topic through time: 
-"1953 in Scotland" or "1560s Architecture", with articles sequentially proceeding by year, decade or era.
-
-Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
-The most common path length is 29, with an interquartile range (26, 30).
-The median path length far exceeds the popular 6-degrees of separation (see for 
-example the average number of academic publications separating scholars 
-\cite{six_degrees}), indicating a greater network depth. The network depth suggests a large network of spanning a variety of ideas, which Wikipedia certainly is.
-The distribution of path lengths is similarly scale-free with few articles at the extreme of 365 path lengths, while the majority 
-is between 26-30.
-
-\begin{figure}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/path_lengths_boxplot.png}
-  \caption{
-    \textbf{Path Length Distribution}
-  }
-  The median network depth measured by the number of first links in a path
-  is 29. The 365 cycle for "Orthodox" liturgics is the outlier to the right, while
-  other historical articles about "Scotland," "UK" and so on are slightly outside
-  the third quartile. More than $75\%$ of articles have path lenghts between 
-  $0$ and $50$ links.
-
-  \label{fig:Path Length Distribution}
-\end{figure}
 
@@
 articles branching outside of a cycle into the rest of the FLN.
-Akin to river networks, these basins are areas of FLN accumulation with a path 
+We rank basins by the total number of traversal visits in each of the articles
+along the path forming the path.
+Akin to river networks, these basins are areas of accumulation with a path 
 flowing outwards to the rest of the FLN.
+
 The highest ranking basins by the number of traversal visits are groups of articles
-around "Philosophy" given the high number of traversal visits "Philosophy" has. 
-However, restricting our attention to the 
- single articles around which basins formed,
- we find high ranking basins around similarly foundational ideas:
-"Community", "Landmass", "Federal Government", "Presentation", and "Belief System". 
-Around each article is a sets of related ideas. Around "community" are groups of articles expressing similar ideas: "Political Union," "State," and "Federation." 
-These concepts naturally emerge from the First Link Network potentially indicating pillars, which 
-anchor specific knowledge in a broader, simpler concept.
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/basins.pdf}
-  \caption{
-    \textbf{basins around a single article}
-  }
-  \label{fig:basins}
-
-\end{figure*}
+around "Philosophy." 
+Since the accumulation of first links to "Philosophy" is so high, 
+the paths leading to "Philosophy" carry many references.
+The highest ranking paths include branches of philosophy flowing through 
+"Awareness,"Existence," and "Consciousness" to "Philosophy." Other paths
+include concepts around "Mathematics," scientific "Experiments," 
+"Biology," and "Fact."
+These paths link many specific articles to "Philosophy," each funneled through a particular domain.
+
+Excluding basins around "Philosophy" we find other basins around 
+foundational ideas such as "Community," "Landmass," "Federal Government," 
+"Presentation," and "Belief System." 
+The basins around each of these foundational notions are 
+various paths containing related articles. For example around 
+"Community" we find basins
+flowing from "United States" to "Federal Republic" to "Political Union" to "State" culminating at "Community"; we also find basins flowing from 
+"Public Policy" to "Executive (government)" to "Government" to State" and then 
+to "Community"; we also find paths flowing through a similar chain beginning
+with "Democracy," another beginning with "Constitution," another at 
+"Dictatorship," and so on. The ideas build from specific means of organizing
+a community (or society) then build up to "Community." 
+Other basins around landmass for example begin at specific geographical regions
+such as "Eastern Europe" building up to "Continent" and finally "Landmass."
+Each basin accumulates first link references through various 
+natural paths linking general themes to the foundational concept.
+
+Which specific articles direct the path towards a particular concepts? 
+Next we study the FLN using traversal funnels to understand the influence
+a particular article exerts in directing the flow of first link references.
 
@@
 
-Measuring only traversal visits and indegree is limiting as it does not distinguish whether a particular article in a cycle 
-is a funnel, directing many more paths inside the cycle than others. 
+To distinguish articles exerting great influence 
 To distinguish among articles in a cycle, we also measured traversal funnels, or the number of 
@@
 so that accumulation does not flow to all articles in a cycle.
-The importance here is to distinguish an article that happens to be connected to another article with many traversal visits 
-from an article directly funneling many paths.
+
+Measuring traversal funnels reveals a dramatically different structure where "Philosophy" stands unmatched by orders of magnitude:
+
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\textwidth]{graphics/funnels.pdf}
+  \caption{
+    \textbf{Funnels}
+  }
+  \label{fig:Funnels}
+\end{figure*}
+
+Philosophy is not only a stand out the number of traversal visits, but also by the number articles "Philosophy" funnels into
+its cycle. Of any article, the number of traversal funnels Philosophy holds exceeds 
+all others by at least two orders of magnitude.
+Next to the contribution of the largest funnels, "Philosophy" is a singularity. 
+In proportion, "Presentation" holds only $0.4\%$ next to the number of traversal feeds for "Philosophy".
+The second contributor to the Philosophy cycle is "Reality", funneling in a mere $0.2\%$ of traversals visits in the cycle.
+Nevertheless, the other high ranking funnels are remarkably topical, culturally and politically important ideas.
+For example, "Health Care", a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August-2009 and February-2010.
+Other high ranking articles include key historical events such as the "Cold War" or critical scarce resource with recent 
+media discussion such as "Fossil Fuel".
+This coincidence of recent relevance and traversal feed rank suggests the First Link Network measurably represents
+meaningful relationships not only among ideas, but also to English-speaking society. 
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
+  \caption{
+    \textbf{Distribution of Traversal Funnels}
+  }
+  \label{fig:Funnels Distribution}
+\end{figure*}
+
+\subsection{Article Popularity}
+
+Wikipedia released an API to measure article popularity by page views
+starting November 2015. The page views adds another dimension to our
+findings about structure by revealing 
+how users view the information on Wikipedia. 
+We measure popularity as the total number 
+of page views in the English edition of Wikipedia in the month of 
+October 2015---the earliest full month for which the data is available. 
+We find the highest ranking 1000 articles by traversal visits have an average of
+$70$ thousand page views in October with high variation: the standard deviation 
+is $1.1$ million page views. 
+The number of page views has a skewed distribution with $75\%$ of articles
+reaching fewer than $73$ thousand page views in October.
+The article for "United States" has the most page 
+views of the highest ranking $1000$ articles by traversal visits with 
+$1.4$ million views in October. The next most popular articles are 
+"India," "World War II," and the "United Kingdom" each with $0.7-1.03$ million page views.
+
+The "Philosophy" article despite outranking every article in traversal visits,
+has only $240$ thousand page views in October compared to the $1.4$ million
+page views for "United States".
+While "Philosophy" is not itself the most popular article, "Philosophy" serves as 
+an accumulation points for the ideas. Wikipedia users do not login specifically
+to read about "Philosophy," but the authors relate ideas to "Philosophy" 
+by ultimately anchoring the flow of first links to "Philosophy."
+
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/views_visits.png}
+  \caption{
+    \textbf{Article Popularity (by page views) compared to Traversal Visits 
+    for Highest ranking 1000 articles}
+    We use the total number of page views provided by Wikipedia for the month
+    of October 2015 to compare each article's popularity to traversal visits.
+    While the most popular articles do not necessarily correspond to the articles
+    with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
+  }
+  \label{fig:Views and Visits}
+
+\end{figure*}
+
+We also analyze article popularity for the highest ranking 1000 articles by 
+traversal funnels. In October, the average page views per article is 
+$22$ thousand with a standard deviation of $95$ thousand views. The distribution
+is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
+"Halloween" is the most popular article with $2.8$ million views in October,
+likely a result of Halloween falling in the month of October. 
+Other popular articles include "Scientology" ($463$ thousand views), "Clint Eastwood" 
+($341$ thousand views),
+and the "Cold War" ($298$ thousand views) although each has significantly fewer views compared to the views for "Halloween."
+Other standout popular October articles include "24-hour" and "12-hour" clocks likely due
+to Daylight savings in October and "Marriage" likely due to the drop in the number of
+weddings in the months following October
+\cite{weddings}.
+"Philosophy" is ranks is the seventh most popular of the highest ranking articles by traversal funnels, but appears no where in the top 20 among the highest ranking
+articles by traversal visits. "Philosophy" is a relatively popular article 
+among articles influencing the shape of the network, but less popular among 
+To distinguish among articles in a cycle, we also measured traversal funnels, or the number of 
+paths an article directs towards a cycle. Here we count the number of traversal visits up to a cycle, 
+so that accumulation does not flow to all articles in a cycle.
 

### 2015-11-29 writeup/test_com.tex
backup

--- 2015-11-29 writeup/2015-11-wikipedia-network.tex
@@
-\documentclass[pre,twocolumn,twoside,superscriptaddress,floatfix, aps, 10pt]{revtex4-1}
-
-\usepackage{currfile}
-\lefthyphenmin=3
-\righthyphenmin=2
-
-\usepackage{graphicx,epsfig,verbatim,enumerate}
-\usepackage{amssymb,amsmath}
-\usepackage{ifthen}
-
-% Fonts and typesetting settings
-\usepackage[sc]{mathpazo}
-\usepackage[T1]{fontenc}
-\linespread{1.05} % Palatino needs more space between lines
-\usepackage{microtype}
-\usepackage{textcomp}
-
-\usepackage{longtable}
-
-\usepackage{mathtools}
-
-\newboolean{twocolswitch}
-
-\newcommand{\www}[1]{\url{#1}}
-\newcommand{\req}[1]{(\ref{#1})}
-\newcommand{\Req}[1]{Eq.~(\ref{#1})}
-
-% Lettrines
-\usepackage{lettrine}
-
-%useful shortcuts
-\def\R{\ensuremath{\mathbb{R}}} %\ensuremath adds math mode, if forgotten
-\def\Q{\ensuremath{\mathbb{Q}}}
-\def\N{\ensuremath{\mathbb{N}}}
-\def\Z{\ensuremath{\mathbb{Z}}}
-\def\C{\ensuremath{\mathbb{C}}}
-
-%shorcuts with arguments
-\newcommand{\abs}[1]{\left\vert#1\right\vert} %nice absolute values
-\newcommand{\bt}[1]{\textbf{#1}} %bold
-\newcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert} %vector norm
-\renewcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-
-%piecewise function
-
-%\begin{displaymath}
-%   f(x) = \left\{
-%     \begin{array}{lr}
-%       1 & : x \in \mathbb{Q}\\
-%       0 & : x \notin \mathbb{Q}
-%     \end{array}
-%   \right.
-%\end{displaymath}
-
-%environment
-\newcommand{\tab}{\phantom{ssss}}
-
-\usepackage{color}
-\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}
-
-%colors
-\definecolor{javagreen}{rgb}{0.25,0.5,0.35} %dark green color
-\definecolor{lightblue}{rgb}{0.149,0.545,0.824} %solarized blue
-\definecolor{sred}{rgb}{0.863, 0.196, 0.184} %solarized red
-
-\newcommand{\blue}[1]{{\leavevmode\color{lightblue}{#1}}} %solarized blue 
-\newcommand{\green}[1]{{\leavevmode\color{javagreen}{#1}}} %command for green
-\newcommand{\red}[1]{{\leavevmode\color{sred}{#1}}} %solarized red
-\newcommand{\gray}[1]{{\leavevmode\color[gray]{0.5}{#1}}} %gray text
-\newcommand{\Prob}[1]{{\rm Pr}\left(#1\right)}
-
-\setboolean{twocolswitch}{true}
-
-\begin{document}
-
-\title{\protect
-A Web of Human Knowledge: 
-Unravelling Wikipedia's First Link Network
-}
-
-\author{
-\firstname{Mark}
-\surname{Ibrahim}
-}
-\email{mark.s.ibrahim@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Christopher}
-\surname{M. Danforth}
-}
-\email{chris.danforth@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Peter}
-\surname{Sheridan Dodds}
-}
-
-\email{peter.dodds@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    The University of Vermont, Burlington, VT 05401.}
-
-\begin{abstract}
-  \protect
-Apples, oranges, and the most obscure Dylan song too---is everything a few clicks from Philosophy? 
-Within Wikipedia, the surprising answer is yes: nearly all 
-paths lead to Philosophy.
-Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
-More than information about a topic though, Wikipedia is a marvelous web of naturally emerging relationships.  
-By following the First Link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
-Here we study the English edition of Wikipedia's First Link Network for insight into how we relate topics, ideas, people, objects, and events.
-
-We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
-In a novel approach to uncover structure, we traverse every possible path through the network, 
-measuring the accumulation of First Links, path lengths, basins, cycles, and even particular articles funneling links into the cycles.
-We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general with 
-basins around fundamental notions such as Community, State, and Science. 
-Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
-These findings enrich our view of how we connect and structure
-an ever growing load of information.
-
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-
-Wikipedia is a towering achievement of the modern era. 
-At no point in history has a larger or more meticulously indexed collection of human knowledge 
-existed.
-Wikipedia contains 37 million articles in 283 languages, 
-with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
-\cite{stats} \cite{drugs}.
-All this information does not lie dormant. 
-Wikipedia is the sixth most visited
-site in the world, surpassing 18 billion page views in a single month
-\cite{wiki_views}.
-The most widely used general reference, Wikipedia has amassed an awe-inspiring collection of human knowledge.
-
-It's not surprising then to see Wikipedia as the object of many studies. 
-Researchers have examined the cultural dynamics among editors,
-\cite{editors}
-the accuracy of the content relative to traditional encyclopedias,
-\cite{accuracy1}
-\cite{accuracy2}
-the disciplines and topics covered 
-\cite{coverage}
-and bias against portions of the population
-\cite{bias_women}.
-Wikipedia's content has also proven a powerful tool. 
-Researchers have used Wikipedia to identify missing dictionary entries
-\cite{missing_entries},
-cluster short text
-\cite{clustering},
-compute semantic relatedness
-\cite{semantic_relatedness},
-and disambiguate meaning \cite{disambiguating}.
-
-While these many studies have dissected and fruitfully applied Wikipedia's content,
-few have examined the connections among the many articles.
-Beyond information, Wikipedia is a web 
-relating ideas. 
-Within the body text of Wikipedia articles hyperlinks reference
-other Wikipedia articles.
-A hyperlink from one Wikipedia article to another is a way to relate the ideas
-conveyed in the two articles 
-\cite{relevance}.
-The notion that hyperlinks convey information about the content of a
-page has proved successful in numerous domains from search engine algorithms 
-such as PageRank 
-\cite{pagerank} 
-to topic classification
-\cite{classifier}.
-We treat a hyperlink as a mechanism to connect two ideas.
-
-The authors of a Wikipedia article choose where and whether to include a 
-hyperlink reference to another Wikipedia article
-by including hyperlinks in the article's markup.
-In describing a particular topic, each set of authors decides which articles are the most
-relevant references.
-For example, the hyperlinks in the "Train" article depend only on what its authors 
-decide are pertinent articles to reference.
-The collection of hyperlinks emerges from the independent choice of each article's authors. 
-The first hyperlink to another Wikipedia article marks the earliest moment in the body text where the authors
-choose to connect the information discussed to that of another article.
-By following the first hyperlink to another article in the English edition of
-Wikipedia, we connect the ideas in one article to those in another, ultimately forming a directed network: 
-{\it Wikipedia's First Link Network} (FLN).
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
-  \caption{
-    \textbf{First Link Path For "Train"}
-    We follow the first link to another Wikipedia article
-    in the main body of the article---the area
-    inside the green rectangle, which excludes 
-    side bar elements, the navigation bar and title. 
-    In this example, the first link to another Wikipedia article is "Rail Transport." 
-    We can again select the first link on the "Rail Transport"
-    article, repeating the process to 
-    form a path of First Links.
-  }
-  \label{fig:Train First Links}
-\end{figure*}
-
-What information could the first links possibly reveal?
-Inspired by the claim that the majority of first links lead to the 
-"Philosophy" article---popularized by an xkcd comic and subsequently
-discussed in blog posts ---we holistically study 
-Wikipedia's FLN as a map relating areas of human knowledge. 
-The map is a hierarchical taxonomy where "Train" links to a parent node, 
-"Rail Transport," while many child nodes, "Steel" and "Horsepower," link up to 
-"Train." Unlike previous taxonomies
-created by individuals
-\cite{locke}
-\cite{descartes}
-\cite{aristotle} 
-or a select group of individuals 
-\cite{hist_thesaurus}, 
-the organization of ideas in the FLN 
-does not emerge from a concerted centralized effort, 
-but rather from the independent choice of every article's authors.
-The FLN is a wealth of relations among all the inventions, places,
-figures, objects, and events described in Wikipedia, across space and time.
-
-Our goal is to gain insight into how the information on Wikipedia is organized and related
-by studying the structure of the FLN.
-In a novel approach, we develop metrics to capture 
-the dominant features of the FLN's structure.
-We measure dynamics of the FLN as a flow, quantifying 
-aspects of the FLN from the accumulation of first links around articles 
-to the influence an article exerts in shaping the FLN.
-Together with cycles, indegree, depth, and the content of the articles, 
-we build our analysis of the relations among the ideas in Wikipedia.
-
-\section{Traversing the First Link Network}
-
-One essential measure of a directed network structure is the degree distribution 
-\cite{newman}. 
-Degree distribution has been used to study many phenomena from disease outbreak 
-\cite{disease} 
-to the dynamics of social networks 
-\cite{social_nets}.
-The degree distribution in the FLN describes how many first links point to a particular article. 
-We measure the indegree in the FLN to understand which articles authors choose to reference in 
-the first link.
-A higher indegree indicates more authors selected the particular article
-as a relevant reference in describing their topic. On the other hand, articles with zero 
-indegree have no references---they are outer leaves in the network. 
-Similar to PageRank, the indegree is also a way to rank the articles in Wikipedia.
-We can rank articles by indegree to see which articles are most and least referenced in the FLN. 
-Additionally, we can assess the degree distribution to see whether authors tend to reference all articles
-equally or only a few, as observed in many social networks \cite{social_nets}.
-
-The degree distribution captures one step in the FLN, measuring   
-the destination of a given link. 
-Indegree measures only the particular set of first links between two articles, 
-rather than the richer dynamics of how many articles fit together---how links 
-create a flow through the FLN.
-One way to expand the analysis beyond a single step is to map the full path of 
-an article's first link. 
-Each article is both a destination and a link to another article, similar to an element 
-in a linked list data structure. 
-By following the chain of links we can construct a path through the network with each article 
-leading to the next via the first link. 
-For example "Train" has a first link to "Conveyance of Passengers and Goods," which is itself
-an article with a first link to "Goods" and so on. The path starting at "Train" 
-is the collection of articles obtained by following the first links starting at "Train."
-The path contains "Goods," "Economics," "Social Science," and other articles related to 
-"Train," but not directly referenced by a first link.
-
-Applying a similar concept with a 
-different starting article, we form 
-a path through the FLN moving from one article to the next.
-A path in the FLN connects a group of articles allowing us to measure how two articles relate,
-even if the articles are not directly linked. 
-The collection of articles is path-connected conveying a flow of concepts from one article to another. 
-The notion of paths allows us to map not only which articles link directly to "Philosophy" 
-for example, but also which ideas are eventually anchored in "Philosophy" even if the articles
-are not directly linked.
-The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
-By constructing paths through the FLN, we 
-map how the first links flow through the FLN.
-
-In a novel approach, we use the notion of a path through the FLN to describe 
-the structure of the FLN as a flow relating ideas. 
-Grounded in the notion of flow through river networks,
-\cite{geo_basins}
-we characterize features of the FLN as a flow via first links.
-Previous studies have used flow to characterize the structure of river networks
-\cite{dodds} and describe the organization of food systems as transportation networks
-\cite{food_webs}.
-With paths to mark flow through the FLN, we develop new metrics to 
-guage the accumulation of references, 
-the length of the path relating articles, and the influence a particular
-article exerts in shaping the flow of links through the FLN.
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\textwidth]{graphics/traversal_visit_algo_figure.pdf}  
-  \caption{
-    \textbf{Traversal Visit Algorithm on a sample network}
-     The traversal visit vectors are an adjacency matrix for the paths through the network: the first column is indicates the path formed starting with article A.
-The number of traversal visits for article A is then the number of paths containing A or the sum of the first row in our matrix:
-     $\sum_{i=1}^7 A_{\text{visit, i}} = 7$.
-  }
-  \label{fig:Traversal Visits}
-\end{figure*}
-
-Our goal is to map the flow of first links by following the paths of first links 
-starting at each article. 
-Since a path reveals how the ideas in one article may eventually be 
-anchored in another, the first metric we develop quantifies the accumulation of first links 
-within the FLN. 
-The algorithm begins by selecting an article then traversing the path formed
-by following the first links. Each time a first link references an article, we increment a count
-associated with the article. 
-We repeat until the first link is invalid (outside the FLN) or repeated. The collection of articles is path-connected conveying a flow of concepts from one article to another. 
-We select a second article and repeat the process until we have 
-constructed a path for each article in the network. We call the resulting count associated with each article the number of {\it traversal visits}. The number of traversal visits of an article 
-measures the number of references flowing to the ideas in the article.
-
-We can map the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
-(see figure~\ref{fig:Traversal Visits}), the path starting at article A is the 
-first column in the traversal vists matrix with $1$ if the path contains a given article and 
-$0$ if the path doesn't. The second column indicates the path starting at article B and so on. 
-To compute the number traversal visits for article A we sum the corresponding row in our matrix,
-which totals the number of paths containing article A. The total is a analogous to the flow 
-in a river network with traversal visists measuring the accumulation of water near point A. 
-By constructing a traversal visits matrix, we can similarly compute the number of traversal 
-visits by summing along the corresponding row for our node. The traversal visits matrix for 
-Wikipedia consists 121 million entries to account for each path through the FLN. 
-We measure accumulation of first link references by summing the corresponding row for the 
-article of interest.
-
-In addition to accumulation, the traversal visits algorithm creates paths 
-which uncover depth in the FLN. By measuring the number of first links between
-two articles, we get an additional piece of information we call 
-{\it path length.} We can compute the length a path by summing 
-along a column in our traversal visits matrix.
-The path length describes how closely related ideas are, guaging the FLN's depth.
-The greater the number of first links separating two articles, the greater the number of ideas needed to relate the ideas.
-Although,
-"Train" is related to "Economics," there are several more specific articles
-bridging the connection:
-"Train" is more specifically related to transportation, whose object
-is often goods. Goods are one of the fundamental objects of study in Economics.
-Described in links, this relationship is captured by a relatedness of $4$ 
-first links ultimately connecting "Train" to "Economics." 
-We can examine path length as a relative measure of how closely connected ideas
-may be, but also as an aggregate measure describing the overall separation or 
-connectedness in the FLN.
-We can characterize whether ideas a closely connected by a handful of links, 
-many, or whether a more nuianced organization exists.
-
-One possible path through the FLN is a {\it cycle} or a group of articles 
-linking to one another inside a loop. In our sample network 
-(see figure~\ref{fig:Traversal Visits}) 
-a cycle exists among nodes A, B, and C. Since node C links back to the starting node A, the path starting at node A forms a {\it 3-cycle}. 
-By recording the history of the articles along a path in the FLN, we can identify
-the types of cycle structures within the FLN. Furthermore, since each article 
-in the a cycle has an associated traversal visit count, we can rank cycles by 
-the number of references directed towards a cycle. 
-We can also form and rank {\it basins} in the network by identifying groups of 
-path-connected articles, not necessarily forming a perfect cycle.
-A basin connects a group of articles which ultimately reference other articles
-in the FLN. By identifying basins and cycles, we pinpoint group of closely related articles.
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/traversal_funnel_algo_figure.pdf}
-  \caption{
-    \textbf{Traversal Funnel Algorithm on a sample network}
-  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
-We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.
-For example
-  the number of traversal funnels for article $E$ is 
-  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
-  \label{fig:Traversal Funnels}
-
-\end{figure*}
-
-While traversal visits measure accumulation, each article's first link also 
-influences the shape of the FLN. 
-At a large point of accumulation, a single article's first link 
-can exert great influence over the shape of the FLN by directing many
-references towards a particular path. To distinguish between an article 
-that simply happened to fall within a cycle from an article funneling 
-many first links, we develop a second metric called {\it traversal funnels}.
-
-To measure traversal funnels, we traverse the FLN in the same manner as we 
-did for traversal visits, but end a path once we enter a cycle.
-We are then able to distinguish between an article related to many other ideas
-only by virtue of its place in a cycle from an article exerting influence over where the first links flow. 
-An article with a large number of traversal funnels directs many references
-towards a particular path. In our sample network 
-(see figure~\ref{fig:Traversal Funnels}) article C 
-directs the flow of links towards the 3-cycle, while articles A and B are 
-receipients of the flow---without exerting any influence themselves. 
-While articles A, B, and C all have the same number of traversal visits, 
-article C has 4 traversal funnels (A and B have none). By incrementing the count of 
-first link references on paths only up to a cycle, we can identifying which articles
-exert the greatest influence over the network's structure.
-
-By studying the FLN not only as collection of directly linked pairs of articles, but
-as a flow of paths, 
-we build a powerful arsenal of information with which to study the FLN. 
-From accumulated references, cycles, basins to influence, we can measure how the many articles in Wikipedia are organized and related.
-
 \section{Discoveries}
@@
 
-To distinguish articles exerting great influence 
-To distinguish among articles in a cycle, we also measured traversal funnels, or the number of 
-paths an article directs towards a cycle. Here we count the number of traversal visits up to a cycle, 
-so that accumulation does not flow to all articles in a cycle.
+To analyze the influence of an article exerts in shaping the 
+struture of the FLN, we measure the number of traversal 
+funnels, or paths the article directs towards a cycle (or invalid links).
+Articles directing more paths exert a greater influence over the structure
+of the FLN by increasing the accumulation of first links in a particular
+basin. To distinguish between an article 
+that simply happened to fall within a cycle from an article funneling 
+many first links, we
 
-Measuring traversal funnels reveals a dramatically different structure where "Philosophy" stands unmatched by orders of magnitude:
+To measure traversal funnels, we traverse the FLN in the same manner as we 
+did for traversal visits, but end a path once we enter a cycle.
+We are then able to distinguish between an article related to many other ideas
+only by virtue of its place in a cycle from an article exerting influence over where the first links flow. 
+An article with a large number of traversal funnels directs many references
+towards a particular path. In our sample network 
+(see figure~\ref{fig:Traversal Funnels}) article C 
+directs the flow of links towards the 3-cycle, while articles A and B are 
+receipients of the flow---without exerting any influence themselves. 
+While articles A, B, and C all have the same number of traversal visits, 
+article C has 4 traversal funnels (A and B have none). By incrementing the count of 
+first link references on paths only up to a cycle, we can identifying which articles
+exert the greatest influence over the network's structure.
 
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\textwidth]{graphics/funnels.pdf}
-  \caption{
-    \textbf{Funnels}
-  }
-  \label{fig:Funnels}
-\end{figure*}
-
-Philosophy is not only a stand out the number of traversal visits, but also by the number articles "Philosophy" funnels into
-its cycle. Of any article, the number of traversal funnels Philosophy holds exceeds 
-all others by at least two orders of magnitude.
-Next to the contribution of the largest funnels, "Philosophy" is a singularity. 
-In proportion, "Presentation" holds only $0.4\%$ next to the number of traversal feeds for "Philosophy".
-The second contributor to the Philosophy cycle is "Reality", funneling in a mere $0.2\%$ of traversals visits in the cycle.
-Nevertheless, the other high ranking funnels are remarkably topical, culturally and politically important ideas.
-For example, "Health Care", a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August-2009 and February-2010.
-Other high ranking articles include key historical events such as the "Cold War" or critical scarce resource with recent 
-media discussion such as "Fossil Fuel".
-This coincidence of recent relevance and traversal feed rank suggests the First Link Network measurably represents
-meaningful relationships not only among ideas, but also to English-speaking society. 
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
-  \caption{
-    \textbf{Distribution of Traversal Funnels}
-  }
-  \label{fig:Funnels Distribution}
-\end{figure*}
-
-\subsection{Article Popularity}
-
-Wikipedia released an API to measure article popularity by page views
-starting November 2015. The page views adds another dimension to our
-findings about structure by revealing 
-how users view the information on Wikipedia. 
-We measure popularity as the total number 
-of page views in the English edition of Wikipedia in the month of 
-October 2015---the earliest full month for which the data is available. 
-We find the highest ranking 1000 articles by traversal visits have an average of
-$70$ thousand page views in October with high variation: the standard deviation 
-is $1.1$ million page views. 
-The number of page views has a skewed distribution with $75\%$ of articles
-reaching fewer than $73$ thousand page views in October.
-The article for "United States" has the most page 
-views of the highest ranking $1000$ articles by traversal visits with 
-$1.4$ million views in October. The next most popular articles are 
-"India," "World War II," and the "United Kingdom" each with $0.7-1.03$ million page views.
-
-The "Philosophy" article despite outranking every article in traversal visits,
-has only $240$ thousand page views in October compared to the $1.4$ million
-page views for "United States".
-While "Philosophy" is not itself the most popular article, "Philosophy" serves as 
-an accumulation points for the ideas. Wikipedia users do not login specifically
-to read about "Philosophy," but the authors relate ideas to "Philosophy" 
-by ultimately anchoring the flow of first links to "Philosophy."
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/views_visits.png}
-  \caption{
-    \textbf{Article Popularity (by page views) compared to Traversal Visits 
-    for Highest ranking 1000 articles}
-    We use the total number of page views provided by Wikipedia for the month
-    of October 2015 to compare each article's popularity to traversal visits.
-    While the most popular articles do not necessarily correspond to the articles
-    with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-  }
-  \label{fig:Views and Visits}
-
-\end{figure*}
-
-We also analyze article popularity for the highest ranking 1000 articles by 
-traversal funnels. In October, the average page views per article is 
-$22$ thousand with a standard deviation of $95$ thousand views. The distribution
-is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
-"Halloween" is the most popular article with $2.8$ million views in October,
-likely a result of Halloween falling in the month of October. 
-Other popular articles include "Scientology" ($463$ thousand views), "Clint Eastwood" 
-($341$ thousand views),
-and the "Cold War" ($298$ thousand views) although each has significantly fewer views compared to the views for "Halloween."
-Other standout popular October articles include "24-hour" and "12-hour" clocks likely due
-to Daylight savings in October and "Marriage" likely due to the drop in the number of
-weddings in the months following October
-\cite{weddings}.
-"Philosophy" is ranks is the seventh most popular of the highest ranking articles by traversal funnels, but appears no where in the top 20 among the highest ranking
-articles by traversal visits. "Philosophy" is a relatively popular article 
-among articles influencing the shape of the network, but less popular among 
-To distinguish among articles in a cycle, we also measured traversal funnels, or the number of 
-paths an article directs towards a cycle. Here we count the number of traversal visits up to a cycle, 
-so that accumulation does not flow to all articles in a cycle.
-
-Measuring traversal funnels reveals a dramatically different structure where "Philosophy" stands unmatched by orders of magnitude:
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\textwidth]{graphics/funnels.pdf}
-  \caption{
-    \textbf{Funnels}
-  }
-  \label{fig:Funnels}
-\end{figure*}
-
-Philosophy is not only a stand out the number of traversal visits, but also by the number articles "Philosophy" funnels into
-its cycle. Of any article, the number of traversal funnels Philosophy holds exceeds 
-all others by at least two orders of magnitude.
-Next to the contribution of the largest funnels, "Philosophy" is a singularity. 
-In proportion, "Presentation" holds only $0.4\%$ next to the number of traversal feeds for "Philosophy".
-The second contributor to the Philosophy cycle is "Reality", funneling in a mere $0.2\%$ of traversals visits in the cycle.
-Nevertheless, the other high ranking funnels are remarkably topical, culturally and politically important ideas.
-For example, "Health Care", a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August-2009 and February-2010.
-Other high ranking articles include key historical events such as the "Cold War" or critical scarce resource with recent 
-media discussion such as "Fossil Fuel".
-This coincidence of recent relevance and traversal feed rank suggests the First Link Network measurably represents
-meaningful relationships not only among ideas, but also to English-speaking society. 
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
-  \caption{
-    \textbf{Distribution of Traversal Funnels}
-  }
-  \label{fig:Funnels Distribution}
-\end{figure*}
-
-\subsection{Article Popularity}
-
-Wikipedia released an API to measure article popularity by page views
-starting November 2015. The page views adds another dimension to our
-findings about structure by revealing 
-how users view the information on Wikipedia. 
-We measure popularity as the total number 
-of page views in the English edition of Wikipedia in the month of 
-October 2015---the earliest full month for which the data is available. 
-We find the highest ranking 1000 articles by traversal visits have an average of
-$70$ thousand page views in October with high variation: the standard deviation 
-is $1.1$ million page views. 
-The number of page views has a skewed distribution with $75\%$ of articles
-reaching fewer than $73$ thousand page views in October.
-The article for "United States" has the most page 
-views of the highest ranking $1000$ articles by traversal visits with 
-$1.4$ million views in October. The next most popular articles are 
-"India," "World War II," and the "United Kingdom" each with $0.7-1.03$ million page views.
-
-The "Philosophy" article despite outranking every article in traversal visits,
-has only $240$ thousand page views in October compared to the $1.4$ million
-page views for "United States".
-While "Philosophy" is not itself the most popular article, "Philosophy" serves as 
-an accumulation points for the ideas. Wikipedia users do not login specifically
-to read about "Philosophy," but the authors relate ideas to "Philosophy" 
-by ultimately anchoring the flow of first links to "Philosophy."
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/views_visits.png}
-  \caption{
-    \textbf{Article Popularity (by page views) compared to Traversal Visits 
-    for Highest ranking 1000 articles}
-    We use the total number of page views provided by Wikipedia for the month
-    of October 2015 to compare each article's popularity to traversal visits.
-    While the most popular articles do not necessarily correspond to the articles
-    with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-  }
-  \label{fig:Views and Visits}
-
-\end{figure*}
-
-We also analyze article popularity for the highest ranking 1000 articles by 
-traversal funnels. In October, the average page views per article is 
-$22$ thousand with a standard deviation of $95$ thousand views. The distribution
-is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
-"Halloween" is the most popular article with $2.8$ million views in October,
-likely a result of Halloween falling in the month of October. 
-Other popular articles include "Scientology" ($463$ thousand views), "Clint Eastwood" 
-($341$ thousand views),
-and the "Cold War" ($298$ thousand views) although each has significantly fewer views compared to the views for "Halloween."
-Other standout popular October articles include "24-hour" and "12-hour" clocks likely due
-to Daylight savings in October and "Marriage" likely due to the drop in the number of
-weddings in the months following October
-\cite{weddings}.
-"Philosophy" is ranks is the seventh most popular of the highest ranking articles by traversal funnels, but appears no where in the top 20 among the highest ranking
-articles by traversal visits. "Philosophy" is a relatively popular article 
-among articles influencing the shape of the network, but less popular among 
-articles with a large accumulation among the flow of first links.
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/funnels_visits.png}
-  \caption{
-    \textbf{Article Popularity (by page views) compared to Traversal Funnels
-    for Highest ranking 1000 articles}
-  }
-    We use the total number of page views provided by Wikipedia for the month
-    of October 2015 to compare each article's popularity to traversal funnels.
-    In the top left is "Halloween" with a significantly greater 
-    number of page views in the month of October.
-  \label{fig:Views and Funnels}
-\end{figure*}
-
-%========================Conclusion====================================
-
-\section{Reflections}
-
-The findings here should only be considered within the limitations of their context.
-We examined only the English version of Wikipedia at a particular moment in time.
-Furthermore, we only studied the First Link in the main body  of each article
-as a means to related one article to another. Finally, Wikipedia, while the largest 
-collection of human knowledge, is rife with the biases of the many contributing editors ((cite))
-. Nevertheless, the findings do reveal
-generalizable relationships, point to foundational notions, and uncover many curiosities.
-
-Among the curiosities is the multiple appearance of scale-free distributions within the network. 
-The three metrics we developed: path length, traversal visits, and traversal funnels are all marked 
-by scale-free distributions. Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
-articles are responsible for funneling most paths. When measured against the traversal funnels, 
-"Philosophy" emerges as an exceptional article by orders of magnitude. 
-Nevertheless, many other foundational ideas emerged naturally within the First Link Network. 
-Basins around "Community", "State", and "Science" reveal a foundational structure within the network. 
-More curious is the emergence of recently prominent political and economics topics such as "Fossil Fuel" and "Health Care" 
-within the highest ranking funnels. 
-Wikipedia seems to reflect not only timeless foundations, but also the topical (at least within English speaking society).
-
-Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the First Link.
-These findings also form the basis for the creation of a taxonomy where 
-every idea, event, or object sits within a hierarchy of connected notions.
-The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
-Applications could range from an enhanced thesaurus of ideas to psychological insights into how humans form associations.
-Specifically, an ever-evolving reference of related hierarchical concepts can be applied to search engine algorithms 
-or natural language processing.
-
-\acknowledgments
-Thank you to my friend RJ for pointing out the xkcd comic describing how the majority of links lead to "Philosophy," inspiring this research.
-
-\newpage
-
-\section{Appendix}
-
-\subsection{Constructing The First Link Network}
-
-To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
-Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
-eliminating any statistical error due to sampling.
-We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
-The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
-and disambiguations.
-Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
-((cite)), our aim is to characterize the dynamics of the First Link Network, not record a particular link between one
-article and another.
-
-Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
-Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
-information.
-While a human could accurately identify the First Link, to map the entire First Link Network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
-
-While some libraries exist for MediaWiki Markup,
-Approaches using existing libraries led to several bugs 
-including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
-as well as compatibility with other libraries used to parse the XML.
-Consequently,  we developed an algorithm for parsing the First Link in the XML version of each article.
-Our parsing algorithm aimed to: 
-1) accurately identify the First Link among other page elements, and 
-2) efficiently do so---that is without needing for several passes through the data.
-To process an article in a single pass, we developed a hierarchical system of flags:
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/flags.pdf}  
-  \caption{
-    \textbf{Parsing Algorithm of Wikipedia's XML dump}
-     The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
-Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
-Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.
-  }
-  \label{fig:parsing algorithm}
-\end{figure*}
-
-The algorithm loops in three-character chunks to account for potentially nested elements, 
-shifting by one character steps through the article markup.
-If any markup triggers for a flag are detected, a flag is raised. 
-Once a flag is raised, we stop processing and proceed to the next character
-until the flag's closing markup.
-A First Link is identified only if Flags 1, 2, and 3 are all off.
-In this case, the entire link is retrieved. 
-We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
-as well as common file extensions for 
-((cite))
-images, audio files, and the like.
-The First Link of an article is then the earliest valid link with unraised flags.
-
-To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
-across 112 cores of the UVM supercomputer cluster
-\cite{vacc}
-We then joined the results to form a hash table containing every Wikipedia article and its corresponding
-First Link. The resulting network map is the basis of our analysis.
-
-\subsection{Supplementary Tables}
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/top_2loops.png}
-  \caption{
-    \textbf{highest ranking 2-Cycles}
-  }
-  \label{fig:Highest Ranking 2-Cyles}
-\end{figure*}
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/top_3loops.png}
-  \caption{
-    \textbf{highest ranking 3-cycles}
-  }
-  \label{fig:Highest Ranking 3-cyles}
-\end{figure*}
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
-  \caption{
-    \textbf{Highest Ranking Articles by Traversal Funnels}
-  }
-  \label{fig:Highest Ranking by Traversal Funnels}
-\end{figure*}
-
-\begin{thebibliography}{1}
-
-      \bibitem{editors} Iba, Takashi, et al.
-"Analyzing the creative editing behavior of Wikipedia editors: Through dynamic social network analysis." Procedia-Social and Behavioral Sciences 2.4 (2010): 6441-6456.
-
-      \bibitem{accuracy1} Holman Rector, Lucy. "Comparison of Wikipedia and other encyclopedias for accuracy, breadth, and depth in historical articles." Reference services review 36.1 (2008): 7-22.
-
-      \bibitem{accuracy2} Giles, Jim. "Internet encyclopaedias go head to head." Nature 438.7070 (2005): 900-901.
-
-       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. "An analysis of topical coverage of Wikipedia." Journal of Computer‐Mediated Communication 13.2 (2008): 429-440.
-
-        \bibitem{bias_women} Hill, Benjamin Mako, and Aaron Shaw.
-"The Wikipedia gender gap revisited: characterizing survey response bias with propensity score estimation." PloS one 8.6 (2013): e65782.
-
-      \bibitem{missing_entries} Williams, Jake Ryland, Clark, Eric M., Bagrow, James P., Danforth, Christopher M.
-\& Dodds, Peter Sheridan (2015).
-Identifying missing dictionary entries with frequency-conserving context models.
-Phys.
-Rev.
-E, 92, 042808.
-
-      \bibitem{clustering} Banerjee, Somnath, Krishnan Ramanathan, and Ajay Gupta.
-"Clustering short texts using wikipedia." Proceedings of the 30th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2007.
-
-      \bibitem{semantic_relatedness} Gabrilovich, Evgeniy, and Shaul Markovitch. "Computing Semantic Relatedness Using Wikipedia-based Explicit Semantic Analysis." IJCAI. Vol. 7. 2007.
-
-      \bibitem{disambiguating} Cucerzan, Silviu. "Large-Scale Named Entity Disambiguation Based on Wikipedia Data." EMNLP-CoNLL. Vol. 7. 2007.
-
-      \bibitem{drugs} Clauson, Kevin A., et al. "Scope, completeness, and accuracy of drug information in Wikipedia." Annals of Pharmacotherapy 42.12 (2008): 1814-1821.
-
-      \bibitem{stats} Wikipedia contributors. "Wikipedia:Statistics." Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 9 Nov. 2015. Web. 13 Nov. 2015.
-
-      \bibitem{pagerank} Page, Lawrence, et al. "The PageRank citation ranking: bringing order to the Web." (1999).
-
-      \bibitem{classifier} Chakrabarti, Soumen, Mukul Joshi, and Vivek Tawde.
-"Enhanced topic distillation using text, markup tags, and hyperlinks." Proceedings of the 24th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2001.
-
-      \bibitem{relevance} Kamps, Jaap, and Marijn Koolen. "Is Wikipedia link structure different?." Proceedings of the second ACM international conference on Web search and data mining. ACM, 2009.
-
-    \bibitem{locke} Bolton, Martha Brandt.
-"The Taxonomy of Ideas in Locke’s {\it Essay} ", The Cambridge Companion to Locke's 'Essay Concerning Human Understanding'.
-Ed.Lex Newman..
-1st ed.
-Cambridge: Cambridge University Press, 2007.
-67-100.
-Cambridge Companions Online.
-Web.
-13 November 2015.
-http://dx.doi.org/10.1017/CCOL0521834333.004.
-
-    \bibitem{aristotle} Studtmann, Paul, "Aristotle's Categories", The Stanford Encyclopedia of Philosophy (Summer 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/sum2014/entries/aristotle-categories/}
-
-    \bibitem{descartes} Smith, Kurt, "Descartes' Theory of Ideas", The Stanford Encyclopedia of Philosophy (Spring 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/spr2014/entries/descartes-ideas/}
-
-    \bibitem{hist_thesaurus} What is the Historical Thesaurus of the OED - Oxford English Dictionary (Oxford English Dictionary)
-    http://public.oed.com/historical-thesaurus-of-the-oed/what-is-the-historical-thesaurus-of-the-oed/
-
-    \bibitem{vacc} The authors acknowledge the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
-    \url{http://www.uvm.edu/vacc}
-
-    \bibitem{wiki_views} Page Views for Wikipedia, Non-mobile site, Normalized (Page Views for Wikipedia, Non-mobile site, Normalized)
-    \url{https://stats.wikimedia.org/EN/TablesPageViewsMonthly.htm}
-
-    \bibitem{weddings} Wedding statistics in the United States (Wedding statistics in the United States)
-    \url{http://www.soundvision.com/article/wedding-statistics-in-the-united-states}
-
-    \bibitem{six_degrees} Elmacioglu, Ergin, and Dongwon Lee. "On six degrees of separation in DBLP-DB and more." ACM SIGMOD Record 34.2 (2005): 33-40.
-
-    \bibitem{wordnet} Princeton University "About WordNet." WordNet. Princeton University. 2010. 
-    \url{http://wordnet.princeton.edu}
-
-    \bibitem{newman} Newman, Mark EJ. "The structure and function of complex networks." SIAM review 45.2 (2003): 167-256.
-
-    \bibitem{disease} Eubank, Stephen, et al. "Modelling disease outbreaks in realistic urban social networks." Nature 429.6988 (2004): 180-184.
-
-    \bibitem{social_nets} Newman, Mark EJ, Duncan J.
-Watts, and Steven H.
-Strogatz.
-"Random graph models of social networks." Proceedings of the National Academy of Sciences 99.suppl 1 (2002): 2566-2572.
-
-    \bibitem{geo_basins} Horton, Robert E.
-"Erosional development of streams and their drainage basins; hydrophysical approach to quantitative morphology." Geological society of America bulletin 56.3 (1945): 275-370.
-    APA
-
-    \bibitem{dodds} Dodds, Peter Sheridan, and Daniel H. Rothman. "Unified view of scaling laws for river networks." Physical Review E 59.5 (1999): 4865.
-
-    \bibitem{food_webs} Garlaschelli, Diego, Guido Caldarelli, and Luciano Pietronero. "Universal scaling relations in food webs." Nature 423.6936 (2003): 165-168.
-
-\end{thebibliography}
-
-\end{document}
+By studying the FLN not only as collection of directly linked pairs of articles, but
+as a flow of paths, 
+we build a powerful arsenal of information with which to study the FLN. 
+From accumulated references, cycles, basins to influence, we can measure how the many articles in Wikipedia are organized and related.

### 2015-11-29 writeup/test_com2.tex
backup

--- 2015-11-29 writeup/2015-11-wikipedia-network.tex
@@
 %========================Conclusion====================================
-
-\section{Reflections}
-
-The findings here should only be considered within the limitations of their context.
-We examined only the English version of Wikipedia at a particular moment in time.
-Furthermore, we only studied the First Link in the main body  of each article
-as a means to related one article to another. Finally, Wikipedia, while the largest 
-collection of human knowledge, is rife with the biases of the many contributing editors ((cite))
-. Nevertheless, the findings do reveal
-generalizable relationships, point to foundational notions, and uncover many curiosities.
-
-Among the curiosities is the multiple appearance of scale-free distributions within the network. 
-The three metrics we developed: path length, traversal visits, and traversal funnels are all marked 
-by scale-free distributions. Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
-articles are responsible for funneling most paths. When measured against the traversal funnels, 
-"Philosophy" emerges as an exceptional article by orders of magnitude. 
-Nevertheless, many other foundational ideas emerged naturally within the First Link Network. 
-Basins around "Community", "State", and "Science" reveal a foundational structure within the network. 
-More curious is the emergence of recently prominent political and economics topics such as "Fossil Fuel" and "Health Care" 
-within the highest ranking funnels. 
-Wikipedia seems to reflect not only timeless foundations, but also the topical (at least within English speaking society).
-
-Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the First Link.
-These findings also form the basis for the creation of a taxonomy where 
-every idea, event, or object sits within a hierarchy of connected notions.
-The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
-Applications could range from an enhanced thesaurus of ideas to psychological insights into how humans form associations.
-Specifically, an ever-evolving reference of related hierarchical concepts can be applied to search engine algorithms 
-or natural language processing.
-
-\acknowledgments
-Thank you to my friend RJ for pointing out the xkcd comic describing how the majority of links lead to "Philosophy," inspiring this research.
-
-\newpage
-
-\section{Appendix}
-
-\subsection{Constructing The First Link Network}
-
-To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
-Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
-eliminating any statistical error due to sampling.
-We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
-The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
-and disambiguations.
-Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
-((cite)), our aim is to characterize the dynamics of the First Link Network, not record a particular link between one
-article and another.
-
-Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
-Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
-information.
-While a human could accurately identify the First Link, to map the entire First Link Network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
-
-While some libraries exist for MediaWiki Markup,
-Approaches using existing libraries led to several bugs 
-including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
-as well as compatibility with other libraries used to parse the XML.
-Consequently,  we developed an algorithm for parsing the First Link in the XML version of each article.
-Our parsing algorithm aimed to: 
-1) accurately identify the First Link among other page elements, and 
-2) efficiently do so---that is without needing for several passes through the data.
-To process an article in a single pass, we developed a hierarchical system of flags:
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/flags.pdf}  
-  \caption{
-    \textbf{Parsing Algorithm of Wikipedia's XML dump}
-     The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
-Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
-Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.
-  }
-  \label{fig:parsing algorithm}
-\end{figure*}
-
-The algorithm loops in three-character chunks to account for potentially nested elements, 
-shifting by one character steps through the article markup.
-If any markup triggers for a flag are detected, a flag is raised. 
-Once a flag is raised, we stop processing and proceed to the next character
-until the flag's closing markup.
-A First Link is identified only if Flags 1, 2, and 3 are all off.
-In this case, the entire link is retrieved. 
-We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
-as well as common file extensions for 
-((cite))
-images, audio files, and the like.
-The First Link of an article is then the earliest valid link with unraised flags.
-
-To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
-across 112 cores of the UVM supercomputer cluster
-\cite{vacc}
-We then joined the results to form a hash table containing every Wikipedia article and its corresponding
-First Link. The resulting network map is the basis of our analysis.
-
-\subsection{Supplementary Tables}
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/top_2loops.png}
-  \caption{
-    \textbf{highest ranking 2-Cycles}
-  }
-  \label{fig:Highest Ranking 2-Cyles}
-\end{figure*}
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/top_3loops.png}
-  \caption{
-    \textbf{highest ranking 3-cycles}
-  }
-  \label{fig:Highest Ranking 3-cyles}
-\end{figure*}
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
-  \caption{
-    \textbf{Highest Ranking Articles by Traversal Funnels}
-  }
-  \label{fig:Highest Ranking by Traversal Funnels}
-\end{figure*}
-
-\begin{thebibliography}{1}
-
-      \bibitem{editors} Iba, Takashi, et al.
-"Analyzing the creative editing behavior of Wikipedia editors: Through dynamic social network analysis." Procedia-Social and Behavioral Sciences 2.4 (2010): 6441-6456.
-
-      \bibitem{accuracy1} Holman Rector, Lucy. "Comparison of Wikipedia and other encyclopedias for accuracy, breadth, and depth in historical articles." Reference services review 36.1 (2008): 7-22.
-
-      \bibitem{accuracy2} Giles, Jim. "Internet encyclopaedias go head to head." Nature 438.7070 (2005): 900-901.
-
-       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. "An analysis of topical coverage of Wikipedia." Journal of Computer‐Mediated Communication 13.2 (2008): 429-440.
-
-        \bibitem{bias_women} Hill, Benjamin Mako, and Aaron Shaw.
-"The Wikipedia gender gap revisited: characterizing survey response bias with propensity score estimation." PloS one 8.6 (2013): e65782.
-
-      \bibitem{missing_entries} Williams, Jake Ryland, Clark, Eric M., Bagrow, James P., Danforth, Christopher M.
-\& Dodds, Peter Sheridan (2015).
-Identifying missing dictionary entries with frequency-conserving context models.
-Phys.
-Rev.
-E, 92, 042808.
-
-      \bibitem{clustering} Banerjee, Somnath, Krishnan Ramanathan, and Ajay Gupta.
-"Clustering short texts using wikipedia." Proceedings of the 30th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2007.
-
-      \bibitem{semantic_relatedness} Gabrilovich, Evgeniy, and Shaul Markovitch. "Computing Semantic Relatedness Using Wikipedia-based Explicit Semantic Analysis." IJCAI. Vol. 7. 2007.
-
-      \bibitem{disambiguating} Cucerzan, Silviu. "Large-Scale Named Entity Disambiguation Based on Wikipedia Data." EMNLP-CoNLL. Vol. 7. 2007.
-
-      \bibitem{drugs} Clauson, Kevin A., et al. "Scope, completeness, and accuracy of drug information in Wikipedia." Annals of Pharmacotherapy 42.12 (2008): 1814-1821.
-
-      \bibitem{stats} Wikipedia contributors. "Wikipedia:Statistics." Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 9 Nov. 2015. Web. 13 Nov. 2015.
-
-      \bibitem{pagerank} Page, Lawrence, et al. "The PageRank citation ranking: bringing order to the Web." (1999).
-
-      \bibitem{classifier} Chakrabarti, Soumen, Mukul Joshi, and Vivek Tawde.
-"Enhanced topic distillation using text, markup tags, and hyperlinks." Proceedings of the 24th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2001.
-
-      \bibitem{relevance} Kamps, Jaap, and Marijn Koolen. "Is Wikipedia link structure different?." Proceedings of the second ACM international conference on Web search and data mining. ACM, 2009.
-
-    \bibitem{locke} Bolton, Martha Brandt.
-"The Taxonomy of Ideas in Locke’s {\it Essay} ", The Cambridge Companion to Locke's 'Essay Concerning Human Understanding'.
-Ed.Lex Newman..
-1st ed.
-Cambridge: Cambridge University Press, 2007.
-67-100.
-Cambridge Companions Online.
-Web.
-13 November 2015.
-http://dx.doi.org/10.1017/CCOL0521834333.004.
-
-    \bibitem{aristotle} Studtmann, Paul, "Aristotle's Categories", The Stanford Encyclopedia of Philosophy (Summer 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/sum2014/entries/aristotle-categories/}
-
-    \bibitem{descartes} Smith, Kurt, "Descartes' Theory of Ideas", The Stanford Encyclopedia of Philosophy (Spring 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/spr2014/entries/descartes-ideas/}
-
-    \bibitem{hist_thesaurus} What is the Historical Thesaurus of the OED - Oxford English Dictionary (Oxford English Dictionary)
-    http://public.oed.com/historical-thesaurus-of-the-oed/what-is-the-historical-thesaurus-of-the-oed/
-
-    \bibitem{vacc} The authors acknowledge the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
-    \url{http://www.uvm.edu/vacc}
-
-    \bibitem{wiki_views} Page Views for Wikipedia, Non-mobile site, Normalized (Page Views for Wikipedia, Non-mobile site, Normalized)
-    \url{https://stats.wikimedia.org/EN/TablesPageViewsMonthly.htm}
-
-    \bibitem{weddings} Wedding statistics in the United States (Wedding statistics in the United States)
-    \url{http://www.soundvision.com/article/wedding-statistics-in-the-united-states}
-
-    \bibitem{six_degrees} Elmacioglu, Ergin, and Dongwon Lee. "On six degrees of separation in DBLP-DB and more." ACM SIGMOD Record 34.2 (2005): 33-40.
-
-    \bibitem{wordnet} Princeton University "About WordNet." WordNet. Princeton University. 2010. 
-    \url{http://wordnet.princeton.edu}
-
-    \bibitem{newman} Newman, Mark EJ. "The structure and function of complex networks." SIAM review 45.2 (2003): 167-256.
-
-    \bibitem{disease} Eubank, Stephen, et al. "Modelling disease outbreaks in realistic urban social networks." Nature 429.6988 (2004): 180-184.
-
-    \bibitem{social_nets} Newman, Mark EJ, Duncan J.
-Watts, and Steven H.
-Strogatz.
-"Random graph models of social networks." Proceedings of the National Academy of Sciences 99.suppl 1 (2002): 2566-2572.
-
-    \bibitem{geo_basins} Horton, Robert E.
-"Erosional development of streams and their drainage basins; hydrophysical approach to quantitative morphology." Geological society of America bulletin 56.3 (1945): 275-370.
-    APA
-
-    \bibitem{dodds} Dodds, Peter Sheridan, and Daniel H. Rothman. "Unified view of scaling laws for river networks." Physical Review E 59.5 (1999): 4865.
-
-    \bibitem{food_webs} Garlaschelli, Diego, Guido Caldarelli, and Luciano Pietronero. "Universal scaling relations in food webs." Nature 423.6936 (2003): 165-168.
-
-\end{thebibliography}
-
-\end{document}

### 2015-11-29 writeup/2015-11-wikipedia-network.tex
basins updated

@@
 
-To distinguish articles exerting great influence 
+Measuring only traversal visits and indegree is limiting as it does not distinguish whether a particular article in a cycle 
+is a funnel, directing many more paths inside the cycle than others. 
 To distinguish among articles in a cycle, we also measured traversal funnels, or the number of 
@@
 so that accumulation does not flow to all articles in a cycle.
-
-Measuring traversal funnels reveals a dramatically different structure where "Philosophy" stands unmatched by orders of magnitude:
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\textwidth]{graphics/funnels.pdf}
-  \caption{
-    \textbf{Funnels}
-  }
-  \label{fig:Funnels}
-\end{figure*}
-
-Philosophy is not only a stand out the number of traversal visits, but also by the number articles "Philosophy" funnels into
-its cycle. Of any article, the number of traversal funnels Philosophy holds exceeds 
-all others by at least two orders of magnitude.
-Next to the contribution of the largest funnels, "Philosophy" is a singularity. 
-In proportion, "Presentation" holds only $0.4\%$ next to the number of traversal feeds for "Philosophy".
-The second contributor to the Philosophy cycle is "Reality", funneling in a mere $0.2\%$ of traversals visits in the cycle.
-Nevertheless, the other high ranking funnels are remarkably topical, culturally and politically important ideas.
-For example, "Health Care", a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August-2009 and February-2010.
-Other high ranking articles include key historical events such as the "Cold War" or critical scarce resource with recent 
-media discussion such as "Fossil Fuel".
-This coincidence of recent relevance and traversal feed rank suggests the First Link Network measurably represents
-meaningful relationships not only among ideas, but also to English-speaking society. 
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
-  \caption{
-    \textbf{Distribution of Traversal Funnels}
-  }
-  \label{fig:Funnels Distribution}
-\end{figure*}
-
-\subsection{Article Popularity}
-
-Wikipedia released an API to measure article popularity by page views
-starting November 2015. The page views adds another dimension to our
-findings about structure by revealing 
-how users view the information on Wikipedia. 
-We measure popularity as the total number 
-of page views in the English edition of Wikipedia in the month of 
-October 2015---the earliest full month for which the data is available. 
-We find the highest ranking 1000 articles by traversal visits have an average of
-$70$ thousand page views in October with high variation: the standard deviation 
-is $1.1$ million page views. 
-The number of page views has a skewed distribution with $75\%$ of articles
-reaching fewer than $73$ thousand page views in October.
-The article for "United States" has the most page 
-views of the highest ranking $1000$ articles by traversal visits with 
-$1.4$ million views in October. The next most popular articles are 
-"India," "World War II," and the "United Kingdom" each with $0.7-1.03$ million page views.
-
-The "Philosophy" article despite outranking every article in traversal visits,
-has only $240$ thousand page views in October compared to the $1.4$ million
-page views for "United States".
-While "Philosophy" is not itself the most popular article, "Philosophy" serves as 
-an accumulation points for the ideas. Wikipedia users do not login specifically
-to read about "Philosophy," but the authors relate ideas to "Philosophy" 
-by ultimately anchoring the flow of first links to "Philosophy."
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/views_visits.png}
-  \caption{
-    \textbf{Article Popularity (by page views) compared to Traversal Visits 
-    for Highest ranking 1000 articles}
-    We use the total number of page views provided by Wikipedia for the month
-    of October 2015 to compare each article's popularity to traversal visits.
-    While the most popular articles do not necessarily correspond to the articles
-    with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-  }
-  \label{fig:Views and Visits}
-
-\end{figure*}
-
-We also analyze article popularity for the highest ranking 1000 articles by 
-traversal funnels. In October, the average page views per article is 
-$22$ thousand with a standard deviation of $95$ thousand views. The distribution
-is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
-"Halloween" is the most popular article with $2.8$ million views in October,
-likely a result of Halloween falling in the month of October. 
-Other popular articles include "Scientology" ($463$ thousand views), "Clint Eastwood" 
-($341$ thousand views),
-and the "Cold War" ($298$ thousand views) although each has significantly fewer views compared to the views for "Halloween."
-Other standout popular October articles include "24-hour" and "12-hour" clocks likely due
-to Daylight savings in October and "Marriage" likely due to the drop in the number of
-weddings in the months following October
-\cite{weddings}.
-"Philosophy" is ranks is the seventh most popular of the highest ranking articles by traversal funnels, but appears no where in the top 20 among the highest ranking
-articles by traversal visits. "Philosophy" is a relatively popular article 
-among articles influencing the shape of the network, but less popular among 
-To distinguish among articles in a cycle, we also measured traversal funnels, or the number of 
-paths an article directs towards a cycle. Here we count the number of traversal visits up to a cycle, 
-so that accumulation does not flow to all articles in a cycle.
+The importance here is to distinguish an article that happens to be connected to another article with many traversal visits 
+from an article directly funneling many paths.
 

### 2015-11-29 writeup/2015-11-wikipedia-network.tex
edits

--- 2015-11-29 writeup/2015-11-wikipedia-network.tex
@@
 relating ideas. 
-Within the body text of Wikipedia articles hyperlinks reference
+Within the body text of Wikipedia articles, hyperlinks reference
 other Wikipedia articles.
@@
 "Philosophy" article---popularized by an xkcd comic and subsequently
-discussed in blog posts ---we holistically study 
+discussed in blog posts 
+\cite{mat_blog}
+\cite{Ilmari_first_links}
+\cite{xkcd} 
+---we holistically study 
 Wikipedia's FLN as a map relating areas of human knowledge. 
@@
 A higher indegree indicates more authors selected the particular article
-as a relevant reference in describing their topic. On the other hand, articles with zero 
-indegree have no references---they are outer leaves in the network. 
+as a relevant reference. On the other hand, articles with zero 
+indegree have no references---they are outer leaves in the FLN. 
 Similar to PageRank, the indegree is also a way to rank the articles in Wikipedia.
@@
 in a linked list data structure. 
-By following the chain of links we can construct a path through the network with each article 
+By following the chain of links we can construct a path through the FLN with each article 
 leading to the next via the first link. 
@@
 an article with a first link to "Goods" and so on. The path starting at "Train" 
-is the collection of articles obtained by following the first links starting at "Train."
+is the collection of articles obtained by following the first links starting at "Train"
+(see figure~\ref{fig:Train First Links}).
 The path contains "Goods," "Economics," "Social Science," and other articles related to 
@@
 The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
-By constructing paths through the FLN, we 
-map how the first links flow through the FLN.
+Starting at each article, we construct a path through the FLN for each article, 
+ultimately mapping how the first links flow through the FLN.
 
@@
 With paths to mark flow through the FLN, we develop new metrics to 
-guage the accumulation of references, 
+gauge the accumulation of references, 
 the length of the path relating articles, and the influence a particular
@@
 associated with the article. 
-We repeat until the first link is invalid (outside the FLN) or repeated. The collection of articles is path-connected conveying a flow of concepts from one article to another. 
+We repeat until the first link is repeated or invalid (outside the FLN). 
+The collection of articles is path-connected conveying a flow of concepts from one article to another. 
 We select a second article and repeat the process until we have 
@@
 To compute the number traversal visits for article A we sum the corresponding row in our matrix,
-which totals the number of paths containing article A. The total is a analogous to the flow 
+which totals the number of paths containing article A. The total is analogous to the flow 
 in a river network with traversal visists measuring the accumulation of water near point A. 
@@
 Wikipedia consists 121 million entries to account for each path through the FLN. 
-We measure accumulation of first link references by summing the corresponding row for the 
+We measure the accumulation of first link references by summing the corresponding row for the 
 article of interest.
@@
 In addition to accumulation, the traversal visits algorithm creates paths 
-which uncover depth in the FLN. By measuring the number of first links between
+which uncover depth in the FLN. 
+By measuring the number of first links between
 two articles, we get an additional piece of information we call 
@@
 along a column in our traversal visits matrix. 
-The path length describes how closely related ideas are, guaging the FLN's depth.
-The greater the number of first links separating two articles, the greater the number of ideas needed to relate the ideas.
-Although,
+The path length describes how closely related ideas are, gauging the FLN's depth.
+The greater the number of first links separating two articles, the greater the number of ideas needed to relate the ideas. Although, 
 "Train" is related to "Economics," there are several more specific articles
@@
 We can characterize whether ideas a closely connected by a handful of links, 
-many, or whether a more nuianced organization exists.
+many, or whether a more nuanced organization exists.
 
@@
 (see figure~\ref{fig:Traversal Visits}) 
-a cycle exists among nodes A, B, and C. Since node C links back to the starting node A, the path starting at node A forms a {\it 3-cycle}. 
+a cycle exists among nodes A, B, and C. 
+Since node C links back to the starting node A, the path starting at node A forms a {\it 3-cycle}. 
 By recording the history of the articles along a path in the FLN, we can identify
@@
 first link references on paths only up to a cycle, we can identifying which articles
-exert the greatest influence over the network's structure.
+exert the greatest influence over the FLN's structure.
 
@@
 \section{Discoveries}
-
-\green{TODO:
-rewrite \\
-introduce metric\\
-find two or three salient points to discuss in each section \\
-tie into bigger picture
-\begin{itemize}
-    \item reorder so paths and flow are described first then funnels and visits
-    \item Basins expand explanation and break up into 2 paragraphs
-    \item add more captions to figures
-\end{itemize}
-}
 
@@
 The average indegree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5 links.
-Only 4826 articles have more than 100 direct first links and $75\%$ of articles
+Only 4,826 articles have more than 100 direct first links and $75\%$ of articles
 have fewer than 9. 
@@
 corresponding to the yearly calendar of Orthodox Liturgics.
-Each day's Liturgics link to the next day's. Curiously, on the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
+Each day's Liturgics links to the next day's. Curiously, on the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
 (see discussion of cycles).
@@
 (see figure~\ref{fig:Path Length Distribution}).
+
 How can we characterize the flow of ideas along a path? Next, we develop 
@@
 To more accurately gauge the distribution, we construct a log-log plot of the entire dataset: log(traversal visits) against log(rank). 
-We observe a fairly linear fit, as is characteristic of scale-free networks, with an r-value of -.93 and
+We observe a strong linear fit, as is characteristic of scale-free networks, with an r-value of -.93 and 
 a power-law exponent of $-0.636$.
@@
 The highest ranking articles by traversal visits are 
-broad, global topics: 
+broad, global topics
+(see figure~\ref{fig:highest visits}):
 many are academic disciplines such as "Science," "Math," 
-"Geography," "Biology," and "Physics"; others are abstract fundamental concepts such as 
-"Community", "State", "Earth", "Information", "Communication", and "Power."
+"Geography," "Philosophy," "Biology," and "Physics"; others are abstract fundamental concepts such as 
+"Community", "State", "Earth", "Information", 
+"Existence," "Communication", and "Power."
 Since traversal visits measure the accumulation of First Link visits in Wikipedia, 
@@
 Of the 11 million articles, 84,000 are 2-cycles. 
-The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than different, yet connected ideas:
-"Health Care" and "Medical Case Management", "Broadcasting House" and "BBC", "Secondary Education" and "Secondary School".
+The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected ideas:
+"Health Care" and "Medical Case Management", "Broadcasting House" and "BBC", "Secondary Education" and "Secondary School" 
+(see figure~\ref{fig:2-cycles}).
 
@@
 
-Outside of the highest ranking 2-cycles, the typical 2-cycle signals a connection between different, yet very closely related ideas. 
+Outside of the highest ranking 2-cycles, the typical 2-cycle signals a connection between distinct, yet very closely related ideas. 
 Link patterns such as inventor to product ("Voere" to "VEC-91"), event to organizer ("Poetry Bus Tour" to "Weave Books"), and book to author ("Anatomy of Britain" to "Anthony Sampson").
@@
 Similarly, 3-cycles captured a synonymous or close relation among 3 articles: "Tree of life (Biology)", "Tree of life (disambiguation)", 
-and "Tree of life"; "Cinema of India", "Indian Cinema", and "Telugu Cinema". Once we extend our cycle size beyond a length of 6 however, 
+and "Tree of life"; "Cinema of India", "Indian Cinema", and "Telugu Cinema"
+(see figure~\ref{fig:3-cycles}).
+Once we extend our cycle size beyond a length of 6 however, 
 "Philosophy" along with the remaining list of high ranking articles by traversal visits dominate.
@@
 
+Extending beyond 3-cycles we find the highest ranking cycles of length 
+up to 10 links are dominated by "Philosophy," high ranking 3-cycles
+around the "Balkans" and "Tree of life" as well as popular 2-cycles.
 The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day. 
@@
 
-To distinguish articles exerting great influence 
-To distinguish among articles in a cycle, we also measured traversal funnels, or the number of 
-paths an article directs towards a cycle. Here we count the number of traversal visits up to a cycle, 
-so that accumulation does not flow to all articles in a cycle.
-
-Measuring traversal funnels reveals a dramatically different structure where "Philosophy" stands unmatched by orders of magnitude:
-
+To analyze the influence an article exerts in shaping the 
+structure of the FLN, we compute the number of traversal funnels for each.
+Articles directing more paths exert a greater influence over the structure
+of the FLN by increasing the accumulation of first links
+on a particular path. By measuring traversal funnels, we distinguish between an article that simply happened to fall within a cycle from an article funneling 
+many first links.
+
+Ranking articles by the number of traversal funnels we find 
+"Philosophy" as by far the highest-ranking article with 
+$7.37$ million paths
+(see figure~\ref{fig:Funnels}).
 \begin{figure*}[tp!]
@@
 \end{figure*}
-
-Philosophy is not only a stand out the number of traversal visits, but also by the number articles "Philosophy" funnels into
-its cycle. Of any article, the number of traversal funnels Philosophy holds exceeds 
+Of any article, the number of traversal funnels Philosophy holds exceeds 
 all others by at least two orders of magnitude.
-Next to the contribution of the largest funnels, "Philosophy" is a singularity. 
-In proportion, "Presentation" holds only $0.4\%$ next to the number of traversal feeds for "Philosophy".
-The second contributor to the Philosophy cycle is "Reality", funneling in a mere $0.2\%$ of traversals visits in the cycle.
-Nevertheless, the other high ranking funnels are remarkably topical, culturally and politically important ideas.
+The "Philosophy" cycle which contains "Existence," "Awareness," "Reality," 
+and similar articles accumulates the overwhelming proportion of its 
+references through "Philosophy": $7.37$ million of the $7.4$ million references
+are funneled through "Philosophy."
+Second on the list of highest-ranking articles by traversal funnels is 
+"Presentation" with only $30$ thousand paths. Similarly abstract 
+ideas also rank highly such as "Tree of life" (30 thousand), 
+"Reality" (13 thousand), "Jurisdiction" (3 thousand) also appear of the list.
+
+Curiously, many high-ranking articles are remarkably topical, culturally and politically important ideas.
 For example, "Health Care", a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August-2009 and February-2010.
@@
 media discussion such as "Fossil Fuel". 
+The highest-ranking list even includes "Hip Hop," "Cancer," and "Web Page."
 This coincidence of recent relevance and traversal feed rank suggests the First Link Network measurably represents
@@
     \textbf{Distribution of Traversal Funnels}
+  We fit the distribution to a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal funnels. 
+  We find two regimes with the top regime (log(rank) < 9) 
+  well-explained by a linear fit, yielding an r-value of $-0.99$ and a 
+  power-law exponent of $-1.08$. The top regime 
+  explains the distribution of traversal funnels for the $17, 821$ 
+  articles with at least one traversal funnel. The bottom 
+  horizontal regime corresponds to the more than $99\%$ of articles
+  which hold zero traversal funnels.
   }
@@
 \end{figure*}
+
+As a distribution, we find few articles influence the structure of the 
+FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
+an more than $99\%$ with none---most articles are recipients of 
+the references flowing through the articles with at least one traversal funnel.
+When fit to a log-log linear model we find the $99\%$ of articles with zero
+traversal funnels forming one regime (with log(rank) less than 9).
+The top regime, corresponding to the $17, 821$ articles with at least one 
+traversal funnel strongly fits a linear model with an r-value of $-0.99$ 
+to yield a power-law exponent of $-1.08$. Even within the few articles
+influencing the structure of the FLN, only a handful of these exert most of the 
+control.
+
+The flow of ideas in the FLN is overwhelmingly influenced by "Philosophy,"
+with a small proportion guided by "Presentation," "Tree of life (biology)," "Southeastern Europe," 
+"Feudalism," "Census-Designated Place," "United States Constitution," "Reality," and "Health Care."
 
@@
 starting November 2015. The page views adds another dimension to our
-findings about structure by revealing 
+findings by revealing 
 how users view the information on Wikipedia. 
@@
 \cite{weddings}.
-"Philosophy" is ranks is the seventh most popular of the highest ranking articles by traversal funnels, but appears no where in the top 20 among the highest ranking
-articles by traversal visits. "Philosophy" is a relatively popular article 
-among articles influencing the shape of the network, but less popular among 
-To distinguish among articles in a cycle, we also measured traversal funnels, or the number of 
-paths an article directs towards a cycle. Here we count the number of traversal visits up to a cycle, 
-so that accumulation does not flow to all articles in a cycle.
-
-Measuring traversal funnels reveals a dramatically different structure where "Philosophy" stands unmatched by orders of magnitude:
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\textwidth]{graphics/funnels.pdf}
-  \caption{
-    \textbf{Funnels}
-  }
-  \label{fig:Funnels}
-\end{figure*}
-
-Philosophy is not only a stand out the number of traversal visits, but also by the number articles "Philosophy" funnels into
-its cycle. Of any article, the number of traversal funnels Philosophy holds exceeds 
-all others by at least two orders of magnitude.
-Next to the contribution of the largest funnels, "Philosophy" is a singularity. 
-In proportion, "Presentation" holds only $0.4\%$ next to the number of traversal feeds for "Philosophy".
-The second contributor to the Philosophy cycle is "Reality", funneling in a mere $0.2\%$ of traversals visits in the cycle.
-Nevertheless, the other high ranking funnels are remarkably topical, culturally and politically important ideas.
-For example, "Health Care", a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August-2009 and February-2010.
-Other high ranking articles include key historical events such as the "Cold War" or critical scarce resource with recent 
-media discussion such as "Fossil Fuel".
-This coincidence of recent relevance and traversal feed rank suggests the First Link Network measurably represents
-meaningful relationships not only among ideas, but also to English-speaking society. 
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
-  \caption{
-    \textbf{Distribution of Traversal Funnels}
-  }
-  \label{fig:Funnels Distribution}
-\end{figure*}
-
-\subsection{Article Popularity}
-
-Wikipedia released an API to measure article popularity by page views
-starting November 2015. The page views adds another dimension to our
-findings about structure by revealing 
-how users view the information on Wikipedia. 
-We measure popularity as the total number 
-of page views in the English edition of Wikipedia in the month of 
-October 2015---the earliest full month for which the data is available. 
-We find the highest ranking 1000 articles by traversal visits have an average of
-$70$ thousand page views in October with high variation: the standard deviation 
-is $1.1$ million page views. 
-The number of page views has a skewed distribution with $75\%$ of articles
-reaching fewer than $73$ thousand page views in October.
-The article for "United States" has the most page 
-views of the highest ranking $1000$ articles by traversal visits with 
-$1.4$ million views in October. The next most popular articles are 
-"India," "World War II," and the "United Kingdom" each with $0.7-1.03$ million page views.
-
-The "Philosophy" article despite outranking every article in traversal visits,
-has only $240$ thousand page views in October compared to the $1.4$ million
-page views for "United States".
-While "Philosophy" is not itself the most popular article, "Philosophy" serves as 
-an accumulation points for the ideas. Wikipedia users do not login specifically
-to read about "Philosophy," but the authors relate ideas to "Philosophy" 
-by ultimately anchoring the flow of first links to "Philosophy."
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/views_visits.png}
-  \caption{
-    \textbf{Article Popularity (by page views) compared to Traversal Visits 
-    for Highest ranking 1000 articles}
-    We use the total number of page views provided by Wikipedia for the month
-    of October 2015 to compare each article's popularity to traversal visits.
-    While the most popular articles do not necessarily correspond to the articles
-    with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-  }
-  \label{fig:Views and Visits}
-
-\end{figure*}
-
-We also analyze article popularity for the highest ranking 1000 articles by 
-traversal funnels. In October, the average page views per article is 
-$22$ thousand with a standard deviation of $95$ thousand views. The distribution
-is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
-"Halloween" is the most popular article with $2.8$ million views in October,
-likely a result of Halloween falling in the month of October. 
-Other popular articles include "Scientology" ($463$ thousand views), "Clint Eastwood" 
-($341$ thousand views),
-and the "Cold War" ($298$ thousand views) although each has significantly fewer views compared to the views for "Halloween."
-Other standout popular October articles include "24-hour" and "12-hour" clocks likely due
-to Daylight savings in October and "Marriage" likely due to the drop in the number of
-weddings in the months following October
-\cite{weddings}.
-"Philosophy" is ranks is the seventh most popular of the highest ranking articles by traversal funnels, but appears no where in the top 20 among the highest ranking
-articles by traversal visits. "Philosophy" is a relatively popular article 
+"Philosophy" which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears no where in the top 20
+by traversal visits. "Philosophy" is a relatively popular article 
 among articles influencing the shape of the network, but less popular among 
@@
 as a means to related one article to another. Finally, Wikipedia, while the largest 
-collection of human knowledge, is rife with the biases of the many contributing editors ((cite))
+collection of human knowledge, is rife with the biases of the many contributing editors
+\cite{bias}
 . Nevertheless, the findings do reveal
@@
 \acknowledgments
-Thank you to my friend RJ for pointing out the xkcd comic describing how the majority of links lead to "Philosophy," inspiring this research.
+Thank you to my friend RJ for pointing out the Reddit post 
+\cite{reddit}
+describing how the majority of links lead to "Philosophy," inspiring this research.
 
@@
 Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
-((cite)), our aim is to characterize the dynamics of the First Link Network, not record a particular link between one
+\cite{wiki_edits},
+our aim is to characterize the dynamics of the First Link Network, not record a particular link between one
 article and another.
@@
 as well as common file extensions for 
-((cite))
-images, audio files, and the like.
+images, audio files, and the like 
+\cite{media_wiki_templates}.
 The First Link of an article is then the earliest valid link with unraised flags.
@@
 First Link. The resulting network map is the basis of our analysis.
-
-\subsection{Supplementary Tables}
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/top_2loops.png}
-  \caption{
-    \textbf{highest ranking 2-Cycles}
-  }
-  \label{fig:Highest Ranking 2-Cyles}
-\end{figure*}
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/top_3loops.png}
-  \caption{
-    \textbf{highest ranking 3-cycles}
-  }
-  \label{fig:Highest Ranking 3-cyles}
-\end{figure*}
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
-  \caption{
-    \textbf{Highest Ranking Articles by Traversal Funnels}
-  }
-  \label{fig:Highest Ranking by Traversal Funnels}
-\end{figure*}
 
@@
 
+    \bibitem{mat_blog} brain of mat kelcey (brain of mat kelcey). 
+    "Do all first links lead to Philosophy?"
+    \url{http://matpalm.com/blog/2011/08/13/wikipedia-philosophy/}
+
+    \bibitem{Ilmari_first_links} User:Ilmari Karonen/First link (Wikipedia)
+    \url{https://en.wikipedia.org/wiki/User:Ilmari_Karonen/First_link}
+
+    \bibitem{xkcd} xkcd: Extended Mind (xkcd: Extended Mind)
+    \url{http://xkcd.com/903/}
+
+    \bibitem{reddit} Almost every Wikipedia page links to Philosophy if you keep clicking the first link in every article. 
+    \url{https://www.reddit.com/r/InternetIsBeautiful/comments/35asgt/almost_every_wikipedia_page_links_to_philosophy/}
+
+    \bibitem{bias} Wagner, Claudia, et al. "It's a Man's Wikipedia? Assessing Gender Inequality in an Online Encyclopedia." arXiv preprint arXiv:1501.06307 (2015).
+
+    \bibitem{wiki_edits} Wikipedia:Statistics 
+    \url{https://en.wikipedia.org/wiki/Wikipedia:Statistics}
+
+    \bibitem{media_wiki_templates} MediaWiki: "Help: Templates"
+    \url{https://www.mediawiki.org/wiki/Help:Templates}
 \end{thebibliography}

### 2015-11-30 writeup/2015-11-wikipedia-network.tex
synonyms figure + edits

@@
 \title{\protect
-A Web of Human Knowledge: 
-Unravelling Wikipedia's First Link Network
+Connecting Every Bit of Knowledge: \\
+Wikipedia's First Link Network Unravelled
 }
@@
   an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
-  The distribution appears scale free with most articles having fewer than 9 
+  The distribution appears to be scale free with most articles having fewer than 9 
   direct first links, while few hold most.
@@
 \end{figure*}
-The FLN's indegree exihibits a scale-free distribution where a few articles 
+The FLN's indegree exhibits a scale-free distribution where a few articles 
 receive a most direct first references, while most articles receive few or none.
@@
 
-As a distribution, the number traversal visits per article appears scale-free. The majority of articles have fewer than 30 traversal visits, while few 
+As a distribution, the number traversal visits per article appears to be scale-free. The majority of articles have fewer than 30 traversal visits, while few 
 have 5 order of magnitude more traversal visits. 
@@
 
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/synonyms.png}
+  \caption{
+    \textbf{Number of Synonyms for the highest-ranking articles}
+    We use WordNet (the largest lexical database for the English language)
+    to identify the number of synonyms for an article's title. We compare 
+    the distribution of synonyms of the highest-ranking 100 articles
+    against all other Wikipedia articles. As the distribution demonstrates
+    the highest ranking concepts have more synonyms than other articles.
+  }
+  \label{fig:synonyms}
+\end{figure*}
+
 One means to measure the specificity of an article is to identify the number of synonyms available for a word or topic.
@@
 Banana has fewer synonyms than botanical for example.
-To quantify the observed flow, we measured the number of synonyms the article title contained in WordNet---the largest lexical database of the English language
+To quantify the observed flow, we measured the number of synonyms each title contained in WordNet---the largest lexical database of the English language
 \cite{wordnet}. 
-We find the highest ranking 100 articles have on average 5 more synonyms than the typical article; a difference of 2.5 synonyms if we compare the median number of synonyms in each group. 
+We find the highest ranking 100 articles have on average 5 more synonyms than the typical article; a difference of 2.5 synonyms if we compare the median number of synonyms in each group
+(see figure~\ref{fig:synonyms})
+. 
 As suggested by the median, many articles have no synonyms as we might expect, because titles with more than one word are not likely to appear in a thesaurus. 
@@
     \textbf{highest ranking 2-cycles}
+    We identify pairs of articles whose first links point to one another, forming
+    a 2-cycle. We then rank each pair of articles by the total number of 
+    traversal visits to gauge the most referenced groups of two articles linked
+    to each other. We find are synonyms representing the same or nearly the 
+    same concepts as opposed to distinct ideas.
   }
   \label{fig:2-cycles}
-
 \end{figure*}
@@
     \textbf{highest ranking 3-cycles}
+    We identify groups of three articles whose links form a 3-cycle with the 
+    last article in the group linking back to the first article.
+    We rank each 3-cycle by the total number of traversal visits to gauge
+    the most referenced group of three linked articles.
   }
@@
     \textbf{Funnels}
+    We represent the highest-ranking articles by the number of traversal 
+    funnels to gauge the influence each article exerts in shaping the 
+    structure of the FLN. We find "Philosophy" exerts an overwhelming proportion
+    of the influence with other abstract notions and topical concepts ranking
+    next.
   }

### 2015-12-09 writeup/2015-11-wikipedia-network.tex
directory cleanup
fixed figure placement

@@
 Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
-More than information about a topic though, Wikipedia is a marvelous web of naturally emerging relationships.  
+More than information about a topic though, Wikipedia is a web of naturally emerging relationships.  
 By following the First Link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
-Here we study the English edition of Wikipedia's First Link Network for insight into how we relate topics, ideas, people, objects, and events.
+Here we study the English edition of Wikipedia's First Link Network for insight into how the many topics, ideas, people, objects, and events
+are related and organized.
 
@@
 In a novel approach to uncover structure, we traverse every possible path through the network, 
-measuring the accumulation of First Links, path lengths, basins, cycles, and even particular articles funneling links into the cycles.
+measuring the accumulation of First Links, path lengths, basins, cycles, and even the influence a particular article exerts in shaping the 
+network.
 We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general with 
-basins around fundamental notions such as Community, State, and Science. 
+culminating around fundamental notions such as Community, State, and Science. 
 Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
@@
 It's not surprising then to see Wikipedia as the object of many studies. 
-Researchers have examined the cultural dynamics among editors,
-\cite{editors}
-the accuracy of the content relative to traditional encyclopedias,
-\cite{accuracy1}
-\cite{accuracy2}
+Researchers have examined the cultural dynamics among editors
+\cite{editors},
+the accuracy of the content relative to traditional encyclopedias
+\cite{accuracy1},
+\cite{accuracy2},
 the disciplines and topics covered 
-\cite{coverage}
+\cite{coverage},
 and bias against portions of the population
@@
 by including hyperlinks in the article's markup.
-In describing a particular topic, each set of authors decides which articles are the most
+In describing a particular topic, the authors decide which other articles are most
 relevant references.
-For example, the hyperlinks in the "Train" article depend only on what its authors 
-decide are pertinent articles to reference.
+For example, the authors of the "Train" article chose
+"Rail Transport," "Cargo," and "Passengers" 
+as pertinent articles to reference in describing trains.
 The collection of hyperlinks emerges from the independent choice of each article's authors. 
@@
     \textbf{First Link Path For "Train"}
-    We follow the first link to another Wikipedia article
+    We follow the First Link to another Wikipedia article
     in the main body of the article---the area
@@
     side bar elements, the navigation bar and title. 
-    In this example, the first link to another Wikipedia article is "Rail Transport." 
-    We can again select the first link on the "Rail Transport"
+    In this example, the First Link to another Wikipedia article is "Rail Transport." 
+    We can again select the First Link on the "Rail Transport"
     article, repeating the process to 
@@
 
-What information could the first links possibly reveal?
-Inspired by the claim that the majority of first links lead to the 
+What information could the First Links possibly reveal?
+Inspired by the claim that the majority of First Links lead to the 
 "Philosophy" article---popularized by an xkcd comic and subsequently
@@
 the organization of ideas in the FLN 
-does not emerge from a concerted centralized effort, 
-but rather from the independent choice of every article's authors.
-The FLN is a wealth of relations among all the inventions, places,
-figures, objects, and events described in Wikipedia, across space and time.
+does not emerge from a concerted or a centralized effort. 
+Rather each link in the FLN emerges from the independent choice of an article's authors.
+Collectively, the FLN is a wealth of relations among inventions, places,
+figures, objects, and events across space and time.
 
@@
 We measure dynamics of the FLN as a flow, quantifying 
-aspects of the FLN from the accumulation of first links around articles 
+aspects of the FLN from the accumulation of First Links around articles 
 to the influence an article exerts in shaping the FLN.
@@
 \cite{social_nets}.
-The degree distribution in the FLN describes how many first links point to a particular article. 
-We measure the indegree in the FLN to understand which articles authors choose to reference in 
-the first link.
+The degree distribution in the FLN describes how many First Links point to a particular article. 
+We measure the indegree in the FLN to guage the number of articles referencing a particular article.
 A higher indegree indicates more authors selected the particular article
@@
 the destination of a given link. 
-Indegree measures only the particular set of first links between two articles, 
+Indegree measures only the particular set of First Links between two articles, 
 rather than the richer dynamics of how many articles fit together---how links 
@@
 One way to expand the analysis beyond a single step is to map the full path of 
-an article's first link. 
+an article's First Link. 
 Each article is both a destination and a link to another article, similar to an element 
@@
 By following the chain of links we can construct a path through the FLN with each article 
-leading to the next via the first link. 
-For example "Train" has a first link to "Conveyance of Passengers and Goods," which is itself
-an article with a first link to "Goods" and so on. The path starting at "Train" 
-is the collection of articles obtained by following the first links starting at "Train"
+leading to the next via the First Link. 
+For example "Train" has a First Link to "Conveyance of Passengers and Goods," which is itself
+an article with a First Link to "Goods" and so on. The path starting at "Train" 
+is the collection of articles obtained by following the First Links starting at "Train"
 (see figure~\ref{fig:Train First Links}).
 The path contains "Goods," "Economics," "Social Science," and other articles related to 
-"Train," but not directly referenced by a first link.
+"Train," but not directly referenced by a First Link.
+
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\textwidth]{graphics/traversal_visit_algo_figure.pdf}  
+  \caption{
+    \textbf{Traversal Visit Algorithm on a sample network}
+     The traversal visit vectors are an adjacency matrix for the paths through the network: the first column is indicates the path formed starting with article A.
+The number of traversal visits for article A is then the number of paths containing A or the sum of the first row in our matrix:
+     $\sum_{i=1}^7 A_{\text{visit, i}} = 7$.
+  }
+  \label{fig:Traversal Visits}
+\end{figure*}
 
@@
 Starting at each article, we construct a path through the FLN for each article, 
-ultimately mapping how the first links flow through the FLN.
+ultimately mapping how the First Links flow through the FLN.
 
@@
 the structure of the FLN as a flow relating ideas. 
-Grounded in the notion of flow through river networks,
+Grounded in the dynamics of river networks,
 \cite{geo_basins}
-we characterize features of the FLN as a flow via first links.
+we characterize features of the FLN as a flow through First Links.
 Previous studies have used flow to characterize the structure of river networks
@@
 
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\textwidth]{graphics/traversal_visit_algo_figure.pdf}  
-  \caption{
-    \textbf{Traversal Visit Algorithm on a sample network}
-     The traversal visit vectors are an adjacency matrix for the paths through the network: the first column is indicates the path formed starting with article A.
-The number of traversal visits for article A is then the number of paths containing A or the sum of the first row in our matrix:
-     $\sum_{i=1}^7 A_{\text{visit, i}} = 7$.
-  }
-  \label{fig:Traversal Visits}
-\end{figure*}
-
-Our goal is to map the flow of first links by following the paths of first links 
+Our goal is to map the flow of First Links by following the paths of First Links 
 starting at each article. 
 Since a path reveals how the ideas in one article may eventually be 
-anchored in another, the first metric we develop quantifies the accumulation of first links 
+anchored in another, the first metric we develop quantifies the accumulation of First Links 
 within the FLN. 
 The algorithm begins by selecting an article then traversing the path formed
-by following the first links. Each time a first link references an article, we increment a count
+by following the First Links. Each time a First Link references an article, we increment a count
 associated with the article. 
-We repeat until the first link is repeated or invalid (outside the FLN). 
-The collection of articles is path-connected conveying a flow of concepts from one article to another. 
+We repeat until the First Link is repeated or invalid (outside the FLN). 
 We select a second article and repeat the process until we have 
-constructed a path for each article in the network. We call the resulting count associated with each article the number of {\it traversal visits}. The number of traversal visits of an article 
+constructed a path for each article in the network. We call the resulting count for each article the number of {\it traversal visits}. The number of traversal visits of an article 
 measures the number of references flowing to the ideas in the article.
+
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/traversal_funnel_algo_figure.pdf}
+  \caption{
+    \textbf{Traversal Funnel Algorithm on a sample network}
+  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
+We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.
+For example
+  the number of traversal funnels for article $E$ is 
+  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
+  \label{fig:Traversal Funnels}
+
+\end{figure*}
 
@@
 (see figure~\ref{fig:Traversal Visits}), the path starting at article A is the 
-first column in the traversal vists matrix with $1$ if the path contains a given article and 
-$0$ if the path doesn't. The second column indicates the path starting at article B and so on. 
-To compute the number traversal visits for article A we sum the corresponding row in our matrix,
-which totals the number of paths containing article A. The total is analogous to the flow 
-in a river network with traversal visists measuring the accumulation of water near point A. 
+first column in the traversal vists matrix. 
+An entry of $1$ indicates the path contains a given article and 
+$0$ indicates the path doesn't. The columns identify the starting article of the path. 
+For example, the second column corresponds to the path starting at article B. 
+To compute the number traversal visits for article A we sum the corresponding row in our matrix.
+The total is the number of paths containing article A. 
+Analogous to the flow 
+in a river network, traversal visits measure the accumulation of water near point A. 
 By constructing a traversal visits matrix, we can similarly compute the number of traversal 
@@
 Wikipedia consists 121 million entries to account for each path through the FLN. 
-We measure the accumulation of first link references by summing the corresponding row for the 
+We measure the accumulation of First Link references by summing the corresponding row for the 
 article of interest.
@@
 which uncover depth in the FLN. 
-By measuring the number of first links between
+By measuring the number of First Links between
 two articles, we get an additional piece of information we call 
@@
 The path length describes how closely related ideas are, gauging the FLN's depth.
-The greater the number of first links separating two articles, the greater the number of ideas needed to relate the ideas. Although, 
+The greater the number of First Links separating two articles, the greater the number of ideas needed to relate the ideas. Although, 
 "Train" is related to "Economics," there are several more specific articles
@@
 Described in links, this relationship is captured by a relatedness of $4$ 
-first links ultimately connecting "Train" to "Economics." 
+First Links ultimately connecting "Train" to "Economics." 
 We can examine path length as a relative measure of how closely connected ideas
@@
 A basin connects a group of articles which ultimately reference other articles
-in the FLN. By identifying basins and cycles, we pinpoint group of closely related articles.
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/traversal_funnel_algo_figure.pdf}
-  \caption{
-    \textbf{Traversal Funnel Algorithm on a sample network}
-  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
-We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.
-For example
-  the number of traversal funnels for article $E$ is 
-  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
-  \label{fig:Traversal Funnels}
-
-\end{figure*}
-
-While traversal visits measure accumulation, each article's first link also 
+in the FLN. By identifying basins and cycles, we pinpoint groups of closely related articles.
+
+While traversal visits measure accumulation, each article's First Link also 
 influences the shape of the FLN. 
-At a large point of accumulation, a single article's first link 
+At a large point of accumulation, a single article's First Link 
 can exert great influence over the shape of the FLN by directing many
@@
 that simply happened to fall within a cycle from an article funneling 
-many first links, we develop a second metric called {\it traversal funnels}.
+many First Links, we develop a second metric called {\it traversal funnels}.
 
@@
 We are then able to distinguish between an article related to many other ideas
-only by virtue of its place in a cycle from an article exerting influence over where the first links flow. 
+only by virtue of its place in a cycle from an article exerting influence over where the First Links flow. 
 An article with a large number of traversal funnels directs many references
@@
 article C has 4 traversal funnels (A and B have none). By incrementing the count of 
-first link references on paths only up to a cycle, we can identifying which articles
+First Link references on paths only up to a cycle, we can identifying which articles
 exert the greatest influence over the FLN's structure.
@@
 
-We first identify the set of first links directly referencing a particular article 
+\begin{figure}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
+  \caption{
+    \textbf{Highest Ranking Articles by indegree}
+  }
+  We rank each article by the number of direct First Links to the article (indegree). The highest-ranking articles tend to represent abstraction for concepts
+  comprised of various facets and parts.
+  \label{fig:ndegree list}
+\end{figure}
+
+We first identify the set of First Links directly referencing a particular article 
 by studying the indegree of the FLN. 
 A higher indegree suggests more authors found the ideas in the article 
-a pertinent direct reference. 
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
-  \caption{
-    \textbf{Highest Ranking Articles by indegree}
-  }
-  We rank each article by the number of direct first links to the article (indegree). The highest-ranking articles tend to represent abstraction for concepts
-  comprised of various facets and parts.
-  \label{fig:ndegree list}
-\end{figure*}
+pertinent to reference. 
 We rank all 11 million articles by indegree to find 
-the "United States" with 80,249 direct first links as the most referenced
+the "United States" with 80,249 direct First Links as the most referenced
 Wikipedia article. 
@@
 are not among the highest-ranking articles by indegree.
-"Philosophy" has an indegree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W.
+"Philosophy" has an indegree of 581, with direct First Links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W.
 Morris."
@@
 result of many articles directly referencing "Philosophy." 
-Instead, the accumulation of first links, as we argue in our 
+Instead, the accumulation of First Links, as we argue in our 
 discussion of traversal visits, flows towards Philosophy as the 
 ultimate anchor when genralizing from specific articles towards broader notions.
-\begin{figure*}[tp!]
+
+\begin{figure}[tp!]
   \centering	
@@
   The distribution appears to be scale free with most articles having fewer than 9 
-  direct first links, while few hold most.
+  direct First Links, while few hold most.
   \label{fig:degree distribution}
-\end{figure*}
+\end{figure}
+
 The FLN's indegree exhibits a scale-free distribution where a few articles 
 receive a most direct first references, while most articles receive few or none.
-The average indegree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5 links.
-Only 4,826 articles have more than 100 direct first links and $75\%$ of articles
+The average indegree for all 11 million articles is 3.6 direct First Links with a standard deviation of 89.5 links.
+Only 4,826 articles have more than 100 direct First Links and $75\%$ of articles
 have fewer than 9. 
@@
   }
-  The median network depth measured by the number of first links in a path
+  The median network depth measured by the number of First Links in a path
   is 29. The 365 cycle for "Orthodox" liturgics is the outlier to the right, while
@@
 Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
-This roughly corresponds to the official number of articles on Wikipedia 
+This roughly corresponds to the official number of articles on Wikipedia: 
 $~4.7$ million as of November 2014---approximately half of the 11 million 
@@
 example the average number of academic publications separating scholars 
-\cite{six_degrees}), indicating a greater network depth. The network depth suggests a large network of spanning a variety of ideas.
+\cite{six_degrees}), indicating a greater network depth. 
+The network depth's depth suggests ideas are related in small independent clusters. 
+Instead the many diverse ideas fit together in long-chained connections.
 As a distribution, more than $75\%$ of articles have a path length below 
-$50$ first links 
+$50$ First Links 
 while a few temporally organized paths exceed 50 links 
@@
 How can we characterize the flow of ideas along a path? Next, we develop 
-a metric to gauge the accumulation first link references
+a metric to gauge the accumulation First Link references
 and characterize a global organization 
@@
 We followed every possible path through the network, incrementing
-the number of traversal visits for every first link reference along a path 
+the number of traversal visits for every First Link reference along a path 
 to generate a total of 232 million traversal visits.
 
-As a distribution, the number traversal visits per article appears to be scale-free. The majority of articles have fewer than 30 traversal visits, while few 
-have 5 order of magnitude more traversal visits. 
+As a distribution, the number traversal visits per article appears to be scale-free. 
+The majority of articles have fewer than 30 traversal visits. 
+First Link references accumulate at a few articles.
 Specifically, $99.76\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
@@
 
-\begin{figure*}[tp!]
+\begin{figure}[tp!]
   \centering	
@@
 
-\end{figure*}
+\end{figure}
 
@@
 
-The highest ranking articles include Philosophy alongside related articles such as "Existence", "Quality", and "Reality".
-
-\begin{figure*}[tp!]
+\begin{figure}[tp!]
   \centering	
@@
   Traversal Algorithm section for details). In doing so, we can rank each article
-  by the accumulation of first links. Articles with a greater number of traversal visits
-  mark greater points of first link accumulation. The highest ranking articles by traversal visits reveals where the greatest accumulation occurs.
+  by the accumulation of First Links. Articles with a greater number of traversal visits
+  mark greater points of First Link accumulation. The highest ranking articles by traversal visits reveals where the greatest accumulation occurs.
   \label{fig:highest visits}
-\end{figure*}
+\end{figure}
 
@@
   \label{fig:synonyms}
-\end{figure*}
-
-One means to measure the specificity of an article is to identify the number of synonyms available for a word or topic.
-The reasoning here is that a  broader topic likely has many more synonyms relative to a specific, concrete topic.
+\end{figure}
+
+To measure the specificity of an article, we identify the number of synonyms available for a word or topic. 
+A  broader topic likely has many more synonyms relative to a specific, concrete topic. 
 Banana has fewer synonyms than botanical for example. 
@@
     \textbf{highest ranking 2-cycles}
-    We identify pairs of articles whose first links point to one another, forming
+    We identify pairs of articles whose First Links point to one another, forming
     a 2-cycle. We then rank each pair of articles by the total number of 
@@
 Outside of the highest ranking 2-cycles, the typical 2-cycle signals a connection between distinct, yet very closely related ideas. 
-Link patterns such as inventor to product ("Voere" to "VEC-91"), event to organizer ("Poetry Bus Tour" to "Weave Books"), and book to author ("Anatomy of Britain" to "Anthony Sampson").
+We also observe link patterns such as inventor to product ("Voere" to "VEC-91"), event to organizer ("Poetry Bus Tour" to "Weave Books"), and book to author ("Anatomy of Britain" to "Anthony Sampson").
 
@@
 articles branching outside of a cycle into the rest of the FLN.
-We rank basins by the total number of traversal visits in each of the articles
-along the path forming the path.
+We rank basins by the total number of traversal visits for each article in the path. 
 Akin to river networks, these basins are areas of accumulation with a path 
@@
 around "Philosophy." 
-Since the accumulation of first links to "Philosophy" is so high, 
+Since the accumulation of First Links to "Philosophy" is so high, 
 the paths leading to "Philosophy" carry many references.
@@
 such as "Eastern Europe" building up to "Continent" and finally "Landmass."
-Each basin accumulates first link references through various 
+Each basin groups First Link references through various 
 natural paths linking general themes to the foundational concept.
@@
 Next we study the FLN using traversal funnels to understand the influence
-a particular article exerts in directing the flow of first link references.
+a particular article exerts in directing the flow of First Link references.
 
@@
 Articles directing more paths exert a greater influence over the structure
-of the FLN by increasing the accumulation of first links
+of the FLN by increasing the accumulation of First Links
 on a particular path. By measuring traversal funnels, we distinguish between an article that simply happened to fall within a cycle from an article funneling 
-many first links.
+many First Links.
 
@@
 meaningful relationships not only among ideas, but also to English-speaking society. 
-\begin{figure*}[tp!]
+\begin{figure}[tp!]
   \centering	
@@
   \label{fig:Funnels Distribution}
-\end{figure*}
+\end{figure}
 
@@
 When fit to a log-log linear model we find the $99\%$ of articles with zero
-traversal funnels forming one regime (with log(rank) less than 9).
+traversal funnels form one regime (with log(rank) less than 9).
 The top regime, corresponding to the $17, 821$ articles with at least one 
-traversal funnel strongly fits a linear model with an r-value of $-0.99$ 
-to yield a power-law exponent of $-1.08$. Even within the few articles
+traversal funnel strongly fits a linear model with an r-value of $-0.99$. 
+The resulting power-law exponent is $-1.08$. Even within the few articles
 influencing the structure of the FLN, only a handful of these exert most of the 
@@
 to read about "Philosophy," but the authors relate ideas to "Philosophy" 
-by ultimately anchoring the flow of first links to "Philosophy."
-
-\begin{figure*}[tp!]
+by ultimately anchoring the flow of First Links to "Philosophy."
+
+\begin{figure}[tp!]
   \centering	
@@
 among articles influencing the shape of the network, but less popular among 
-articles with a large accumulation among the flow of first links.
-
-\begin{figure*}[tp!]
+articles with a large accumulation among the flow of First Links.
+
+\begin{figure}[tp!]
   \centering	
@@
   \label{fig:Views and Funnels}
-\end{figure*}
+\end{figure}
 
@@
   \label{fig:parsing algorithm}
-\end{figure*}
+\end{figure}
 

### 2015-12-30 writeup/2015-11-wikipedia-network.tex
edits to intro + method

@@
 
+\usepackage{hyperref}
 \usepackage{color}
@@
     Computational Story Lab,  \\
+    Vermont Complex Systems Center,
+    Vermont Advanced Computing Core,
     The University of Vermont, Burlington, VT 05401.}
@@
   \protect
-Apples, oranges, and the most obscure Dylan song too---is everything a few clicks from Philosophy? 
+Apples, oranges, and the most obscure Dylan song too---is every subject a few clicks from Philosophy? 
 Within Wikipedia, the surprising answer is yes: nearly all 
@@
 More than information about a topic though, Wikipedia is a web of naturally emerging relationships.  
-By following the First Link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
+By following the First Link in an article, we connect entries to form a directed network: Wikipedia's First Link Network. 
 Here we study the English edition of Wikipedia's First Link Network for insight into how the many topics, ideas, people, objects, and events
@@
 network.
-We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general with 
+We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general 
 culminating around fundamental notions such as Community, State, and Science. 
@@
 with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
-\cite{stats} \cite{drugs}.
+\cite{drugs} 
+\cite{stats}.
 All this information does not lie dormant. 
 Wikipedia is the sixth most visited
-site in the world, surpassing 18 billion page views in a single month
+site in the world, surpassing 18 billion page views and 26 million edits in a single month
+\cite{wiki_edits}
 \cite{wiki_views}.
@@
 the accuracy of the content relative to traditional encyclopedias
-\cite{accuracy1},
+\cite{accuracy1}
 \cite{accuracy2},
-the disciplines and topics covered 
+the topics covered 
 \cite{coverage},
@@
 other Wikipedia articles.
-A hyperlink from one Wikipedia article to another is a way to relate the ideas
-conveyed in the two articles 
+A hyperlink from one Wikipedia article to another
+indicates a relationship between the two articles
 \cite{relevance}.
@@
 \cite{classifier}.
-We treat a hyperlink as a mechanism to connect two ideas.
+We treat a hyperlink as a mechanism connecting two ideas.
 
 The authors of a Wikipedia article choose where and whether to include a 
-hyperlink reference to another Wikipedia article
-by including hyperlinks in the article's markup.
+reference to another Wikipedia article
+in the HTML markup.
 In describing a particular topic, the authors decide which other articles are most
 relevant references.
-For example, the authors of the "Train" article chose
+For example, the authors of the 
+\href{https://en.wikipedia.org/wiki/Train}{"Train"}
+article chose
 "Rail Transport," "Cargo," and "Passengers" 
-as pertinent articles to reference in describing trains.
+as pertinent articles to reference in describing trains
+\cite{wiki_train}
+.
 The collection of hyperlinks emerges from the independent choice of each article's authors. 
@@
   \caption{
-    \textbf{First Link Path For "Train"}
+    \textbf{First Link Path For "Train".}
     We follow the First Link to another Wikipedia article
@@
 Inspired by the claim that the majority of First Links lead to the 
-"Philosophy" article---popularized by an xkcd comic and subsequently
+"Philosophy" article---popularized by an 
+\href{https://xkcd.com/903/}{xkcd}
+comic and subsequently
 discussed in blog posts 
-\cite{mat_blog}
-\cite{Ilmari_first_links}
-\cite{xkcd} 
+\cite{mat_blog, Ilmari_first_links, xkcd}
 ---we holistically study 
 Wikipedia's FLN as a map relating areas of human knowledge. 
-The map is a hierarchical taxonomy where "Train" links to a parent node, 
-"Rail Transport," while many child nodes, "Steel" and "Horsepower," link up to 
+The map is a hierarchical taxonomy where, for example, "Train" links to a parent node, 
+"Rail Transport," while many child nodes, "Steel" and "Horsepower," link to 
 "Train." Unlike previous taxonomies
 created by individuals
-\cite{locke}
-\cite{descartes}
-\cite{aristotle} 
+\cite{locke, descartes, aristotle}
 or a select group of individuals 
@@
 does not emerge from a concerted or a centralized effort. 
-Rather each link in the FLN emerges from the independent choice of an article's authors.
+Rather each link in the FLN emerges from the choice of each author.
 Collectively, the FLN is a wealth of relations among inventions, places,
@@
 
-Our goal is to gain insight into how the information on Wikipedia is organized and related
-by studying the structure of the FLN.
+Our goal is to study the structure of the FLN for insight into how the information on Wikipedia is organized and related.
 In a novel approach, we develop metrics to capture 
@@
 
-One essential measure of a directed network structure is the degree distribution 
+One essential measure of a directed network's structure is the degree distribution 
 \cite{newman}. 
@@
 \cite{social_nets}.
-The degree distribution in the FLN describes how many First Links point to a particular article. 
-We measure the indegree in the FLN to guage the number of articles referencing a particular article.
+The degree distribution in the FLN describes how many First Links point to a 
+particular article. 
+We measure the indegree in the FLN to gauge the number of articles referencing a particular article.
 A higher indegree indicates more authors selected the particular article
@@
 
-The degree distribution captures one step in the FLN, measuring   
+The degree distribution captures one step in the FLN by measuring   
 the destination of a given link. 
@@
 By following the chain of links we can construct a path through the FLN with each article 
-leading to the next via the First Link. 
+leading to the next through the First Link. 
 For example "Train" has a First Link to "Conveyance of Passengers and Goods," which is itself
@@
   \caption{
-    \textbf{Traversal Visit Algorithm on a sample network}
+    \textbf{Traversal Visit Algorithm on a sample network.}
      The traversal visit vectors are an adjacency matrix for the paths through the network: the first column is indicates the path formed starting with article A.
@@
 
-Applying a similar concept with a 
-different starting article, we form 
-a path through the FLN moving from one article to the next.
-A path in the FLN connects a group of articles allowing us to measure how two articles relate,
+Applying a similar concept with  
+varying starting articles, we form 
+paths through the FLN moving from one article to the next.
+A path in the FLN connects a group of articles, allowing us to measure how two articles relate,
 even if the articles are not directly linked. 
-The collection of articles is path-connected conveying a flow of concepts from one article to another. 
+Path-connected articles express a flow of concepts.
 The notion of paths allows us to map not only which articles link directly to "Philosophy" 
-for example, but also which ideas are eventually anchored in "Philosophy" even if the articles
-are not directly linked.
+for example, but also which ideas flow to "Philosophy":
+path-connected articles to "Philosophy."
 The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
-Starting at each article, we construct a path through the FLN for each article, 
-ultimately mapping how the First Links flow through the FLN.
+Starting at each article, we construct a path through the FLN, 
+ultimately mapping a flow of connections among articles.
 
@@
 Since a path reveals how the ideas in one article may eventually be 
-anchored in another, the first metric we develop quantifies the accumulation of First Links 
-within the FLN. 
-The algorithm begins by selecting an article then traversing the path formed
+anchored in another, the first metric we develop quantifies the accumulation of First Links.
+The algorithm begins by selecting an article, then traversing the path formed
 by following the First Links. Each time a First Link references an article, we increment a count
 associated with the article. 
-We repeat until the First Link is repeated or invalid (outside the FLN). 
+We continue until the First Link is repeated or is invalid---defined here to mean outside of Wikipedia.
 We select a second article and repeat the process until we have 
@@
   \caption{
-    \textbf{Traversal Funnel Algorithm on a sample network}
+    \textbf{Traversal Funnel Algorithm on a sample network.}
   The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
@@
 
-\end{figure*}
-
-We can map the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
+\end{figure}
+
+We can characterize the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
 (see figure~\ref{fig:Traversal Visits}), the path starting at article A is the 
@@
 Analogous to the flow 
-in a river network, traversal visits measure the accumulation of water near point A. 
+in a river network, traversal visits measure the accumulation of water at point A. 
 By constructing a traversal visits matrix, we can similarly compute the number of traversal 
-visits by summing along the corresponding row for our node. The traversal visits matrix for 
-Wikipedia consists 121 million entries to account for each path through the FLN. 
+visits by summing along the corresponding row. The traversal visits matrix for 
+Wikipedia consists of 121 million entries to account for each path through the FLN. 
 We measure the accumulation of First Link references by summing the corresponding row for the 
@@
 two articles, we get an additional piece of information we call 
-{\it path length.} We can compute the length a path by summing 
+{\it path length.} We can compute the length of a path by summing 
 along a column in our traversal visits matrix. 
 The path length describes how closely related ideas are, gauging the FLN's depth.
-The greater the number of First Links separating two articles, the greater the number of ideas needed to relate the ideas. Although, 
-"Train" is related to "Economics," there are several more specific articles
+The greater the number of First Links separating two articles, the greater the number of ideas needed to relate their corresponding ideas. Although
+"Train" is related to "Economics" for example, there are several articles
 bridging the connection:
@@
 connectedness in the FLN.
-We can characterize whether ideas a closely connected by a handful of links, 
+We can characterize whether ideas are closely connected by a handful of links, 
 many, or whether a more nuanced organization exists.
@@
 the types of cycle structures within the FLN. Furthermore, since each article 
-in the a cycle has an associated traversal visit count, we can rank cycles by 
+has an associated traversal visit count, we can rank cycles by 
 the number of references directed towards a cycle. 
@@
 path-connected articles, not necessarily forming a perfect cycle.
-A basin connects a group of articles which ultimately reference other articles
-in the FLN. By identifying basins and cycles, we pinpoint groups of closely related articles.
+A basin connects a group of articles and identifies the paths 
+to a particular article.
+By identifying basins and cycles, we pinpoint groups of closely related articles.
 
@@
 can exert great influence over the shape of the FLN by directing many
-references towards a particular path. To distinguish between an article 
+references on a particular path. To distinguish between an article 
 that simply happened to fall within a cycle from an article funneling 
@@
 We are then able to distinguish between an article related to many other ideas
-only by virtue of its place in a cycle from an article exerting influence over where the First Links flow. 
+only by virtue of its place in a cycle, from an article exerting influence over where the First Links flow. 
 An article with a large number of traversal funnels directs many references
-towards a particular path. In our sample network 
+towards a particular cycle. In our sample network 
 (see figure~\ref{fig:Traversal Funnels}) article C 
 directs the flow of links towards the 3-cycle, while articles A and B are 
-receipients of the flow---without exerting any influence themselves. 
+recipients of the flow---without exerting any direct influence themselves. 
 While articles A, B, and C all have the same number of traversal visits, 
@@
 By studying the FLN not only as collection of directly linked pairs of articles, but
-as a flow of paths, 
+as a flow, 
 we build a powerful arsenal of information with which to study the FLN. 
-From accumulated references, cycles, basins to influence, we can measure how the many articles in Wikipedia are organized and related.
+From accumulated references, cycles and basins to influence, we can measure how the many articles in Wikipedia are organized and related.
 
@@
   \caption{
-    \textbf{Highest Ranking Articles by indegree}
+    \textbf{Highest Ranking Articles by indegree.}
   }
   We rank each article by the number of direct First Links to the article (indegree). The highest-ranking articles tend to represent abstraction for concepts
-  comprised of various facets and parts.
+  comprised of various facets.
   \label{fig:ndegree list}
@@
 
-We first identify the set of First Links directly referencing a particular article 
+We identify the set of First Links directly referencing a particular article 
 by studying the indegree of the FLN. 
@@
 the "United States" with 80,249 direct First Links as the most referenced
-Wikipedia article. 
+Wikipedia article 
+(see figure~\ref{fig:ndegree list}). 
 Other high-ranking articles
@@
 Morris."
-While many articles accumulate at "Philosophy" (see traversal visits discussion), 
+While many articles accumulate at "Philosophy" (see the coming traversal visits discussion), 
 the accumulation is not the 
@@
 discussion of traversal visits, flows towards Philosophy as the 
-ultimate anchor when genralizing from specific articles towards broader notions.
+ultimate anchor when generalizing from specific articles towards broader notions.
 
@@
   \caption{
-    \textbf{FLN Degree Distribution}
-  }
-  We construct a log-log plot fit our results to a linear model. The result is a 
+    \textbf{FLN Degree Distribution.}
+  }
+  We construct a log-log plot and fit our results with a linear model. The result is a 
   an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
-  The distribution appears to be scale free with most articles having fewer than 9 
-  direct First Links, while few hold most.
+  The distribution appears to be scale free.
   \label{fig:degree distribution}
@@
 have fewer than 9. 
-When fit to a linear model on a log-log scale (log(rank) versus log(indegree)), 
+When fit with a linear model on a log-log scale (log(rank) versus log(indegree)), 
 the model's r value is $-0.98$ suggesting a strong log-log linear fit 
@@
 
-Beyond indegree, our work expands the analysis of relations to understand 
-how many more than two articles are organized and related. 
-We develop the notion of a path to describe the structure of the FLN as a 
+Beyond indegree, we develop the notion of a path to describe the structure of the FLN as a 
 flow relating ideas. We traverse every possible path through the FLN, taking 
@@
 connection of ideas span? We gauge the FLN's depth via path length or the number of links traversed until a repeated or invalid link (one outside the FLN). 
-We discover the longest path length is 365
+We discover the longest path length is 365,
 corresponding to the yearly calendar of Orthodox Liturgics.
-Each day's Liturgics links to the next day's. Curiously, on the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
+Each day's Liturgics links to the next day's. On the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
 (see discussion of cycles).
@@
   \caption{
-    \textbf{Path Length Distribution}
+    \textbf{Path Length Distribution.}
   }
@@
 \cite{six_degrees}), indicating a greater network depth. 
-The network depth's depth suggests ideas are related in small independent clusters. 
+The network's depth suggests ideas are not related in small independent clusters. 
 Instead the many diverse ideas fit together in long-chained connections.
@@
 How can we characterize the flow of ideas along a path? Next, we develop 
-a metric to gauge the accumulation First Link references
+a metric to gauge the accumulation of First Link references
 and characterize a global organization 
@@
   \caption{
-    \textbf{Distribution of Traversal Visits}
+    \textbf{Distribution of Traversal Visits.}
   }
@@
   \caption{
-    \textbf{highest ranking articles by number of traversal visits}
+    \textbf{Highest ranking articles by number of traversal visits.}
   }
@@
 For example the flow of First Links for the "Banana" article begins at a very concrete
-and specific topic then flows into progressively broader and broader disciplines, eventually 
-culminating at "Philosophy: "Banana" links to the broader category of "Fruit," which then links to 
+and specific topic, then flows into progressively broader disciplines, eventually 
+culminating at "Philosophy": "Banana" links to the broader category of "Fruit," which then links to 
 "Botany," eventually "Biology," then "Science" and ultimately "Philosophy."
@@
   \caption{
-    \textbf{Number of Synonyms for the highest-ranking articles}
+    \textbf{Number of Synonyms for the highest-ranking articles.}
     We use WordNet (the largest lexical database for the English language)
@@
 We find the highest ranking 100 articles have on average 5 more synonyms than the typical article; a difference of 2.5 synonyms if we compare the median number of synonyms in each group
-(see figure~\ref{fig:synonyms})
-. 
+(see figure~\ref{fig:synonyms}). 
 As suggested by the median, many articles have no synonyms as we might expect, because titles with more than one word are not likely to appear in a thesaurus. 
 Since many articles have no synonyms, we also compared the number of synonyms in the highest-ranking versus typical article, this time excluding all articles without at least 1 synonym. 
-We still find the highest ranking 100 articles with an average of 9.0 (median of 7.0) synonyms whereas the remaining articles on average 5.8 synonyms (median of 7.0), even with the exclusion of all articles without any synonyms.
+We still find the highest ranking 100 articles with an average of 9.0 (median of 7.0) synonyms whereas the remaining articles have on average 5.8 synonyms (median of 7.0), even with the exclusion of all articles without any synonyms.
 The quantifiable difference in synonyms corroborates the flow of links from concrete, specific articles to broader disciplines or fundamental notions.
@@
   \caption{
-    \textbf{highest ranking 2-cycles}
+    \textbf{Highest ranking 2-cycles.}
     We identify pairs of articles whose First Links point to one another, forming
@@
     traversal visits to gauge the most referenced groups of two articles linked
-    to each other. We find are synonyms representing the same or nearly the 
+    to each other. The area of the circles in the figure indicates the number of traversal visits. We find 2-cycles often capture synonyms or articles representing nearly the 
     same concepts as opposed to distinct ideas.
@@
   \caption{
-    \textbf{highest ranking 3-cycles}
+    \textbf{Highest ranking 3-cycles.}
     We identify groups of three articles whose links form a 3-cycle with the 
@@
     the most referenced group of three linked articles. 
+    The area of the circles in the figure indicates the number of traversal visits. 
   }
@@
 The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day.
-Curiously, on the last calendar day, the last article simply links back to January 1, forming a 365-cycle.
 Other lengthy cycles span 60-75 articles including collections of articles on national histories such as "Japanese Eras" 
@@
 
-Excluding basins around "Philosophy" we find other basins around 
+Excluding basins around "Philosophy," we find other basins around 
 foundational ideas such as "Community," "Landmass," "Federal Government," 
@@
 "Dictatorship," and so on. The ideas build from specific means of organizing
-a community (or society) then build up to "Community." 
+a community (or society) and then build up to "Community." 
 Other basins around landmass for example begin at specific geographical regions
@@
   \caption{
-    \textbf{Funnels}
+    \textbf{Funnels.}
     We represent the highest-ranking articles by the number of traversal 
@@
     structure of the FLN. We find "Philosophy" exerts an overwhelming proportion
-    of the influence with other abstract notions and topical concepts ranking
+    of the influence, with other abstract notions and topical concepts ranking
     next.
@@
 Of any article, the number of traversal funnels Philosophy holds exceeds 
-all others by at least two orders of magnitude.
+all others by more than two orders of magnitude.
 The "Philosophy" cycle which contains "Existence," "Awareness," "Reality," 
@@
   \caption{
-    \textbf{Distribution of Traversal Funnels}
+    \textbf{Distribution of Traversal Funnels.}
   We fit the distribution to a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal funnels. 
-  We find two regimes with the top regime (log(rank) < 9) 
+  We find two regimes with the top regime (log(rank) < 4) 
   well-explained by a linear fit, yielding an r-value of $-0.99$ and a 
@@
 When fit to a log-log linear model we find the $99\%$ of articles with zero
-traversal funnels form one regime (with log(rank) less than 9).
+traversal funnels form one regime (with log(rank) less than 4).
 The top regime, corresponding to the $17, 821$ articles with at least one 
@@
 Wikipedia released an API to measure article popularity by page views
-starting November 2015. The page views adds another dimension to our
+starting November 2015. Page views add another dimension to our
 findings by revealing 
@@
 While "Philosophy" is not itself the most popular article, "Philosophy" serves as 
-an accumulation points for the ideas. Wikipedia users do not login specifically
+an accumulation point for ideas. Wikipedia users do not login specifically
 to read about "Philosophy," but the authors relate ideas to "Philosophy" 
@@
     \textbf{Article Popularity (by page views) compared to Traversal Visits 
-    for Highest ranking 1000 articles}
+    for Highest ranking 1000 articles.}
     We use the total number of page views provided by Wikipedia for the month
@@
     \textbf{Article Popularity (by page views) compared to Traversal Funnels
-    for Highest ranking 1000 articles}
+    for Highest ranking 1000 articles.}
   }
@@
   \caption{
-    \textbf{Parsing Algorithm of Wikipedia's XML dump}
+    \textbf{Parsing Algorithm of Wikipedia's XML dump.}
      The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
@@
 
+      \bibitem{wiki_train} Wikipedia contributors. "Train." Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 
+          \url{https://en.wikipedia.org/wiki/Train}
+
     \bibitem{locke} Bolton, Martha Brandt.

### 2016-01-03 writeup/2015-11-wikipedia-network.tex
edits to discussion section

@@
   \protect
-Apples, oranges, and the most obscure Dylan song too---is every subject a few clicks from Philosophy? 
+Apples, oranges, and the most obscure Dylan song too---is every topic a few clicks from Philosophy? 
 Within Wikipedia, the surprising answer is yes: nearly all 
@@
 We identify the set of First Links directly referencing a particular article 
-by studying the indegree of the FLN. 
+by studying indegree in the FLN.
 A higher indegree suggests more authors found the ideas in the article 
@@
 the "Netherlands." These high ranking articles are useful abstractions: nations
-describe a collection of individuals, a common culture, language, or 
+describe a collection of individuals with a common culture, language, or 
 geographical proximity; sports teams describe an ever changing collection of 
-sports players often associated with a cultural identity and a geographical 
-region. Since abstractions such as nations and teams are inherently comprised
-of many parts, authors reference the larger abstraction when describing 
-a part. For example, in describing the geographical region of north America, 
-the United States, Canada, or Mexico are natural references anchoring the 
-geographical location to a larger national identity.
-
-"Philosophy" and other philosophical concepts with many traversal visits
+sports players often associated with a cultural identity or a geographical 
+region. 
+Since abstractions such as nations and teams are inherently comprised
+of many parts, authors reference the abstraction when describing a part.
+In describing the New England Patriots or the New York Giants
+for example, "American Football" is the natural abstraction, which 
+anchors many specific teams.
+
+"Philosophy" and other philosophical concepts
 are not among the highest-ranking articles by indegree.
@@
 discussion of traversal visits, flows towards Philosophy as the 
-ultimate anchor when generalizing from specific articles towards broader notions.
+ultimate anchor when generalizing from specific articles to broader notions.
 
@@
 The FLN's indegree exhibits a scale-free distribution where a few articles 
-receive a most direct first references, while most articles receive few or none.
+receive most direct first references; most articles receive few or none.
 The average indegree for all 11 million articles is 3.6 direct First Links with a standard deviation of 89.5 links.
@@
 the model's r value is $-0.98$ suggesting a strong log-log linear fit 
-with a power law exponent of $-0.79$.
+with a power law exponent of $-0.79$.\\
 
@@
 We first seek to describe the depth of the FLN: how many links does a 
-connection of ideas span? We gauge the FLN's depth via path length or the number of links traversed until a repeated or invalid link (one outside the FLN). 
+connection of ideas span? We gauge the FLN's depth by 
+measuring path length or the number of links traversed until a repeated or invalid link (one outside the FLN). 
 We discover the longest path length is 365,
@@
 (see discussion of cycles).
-We also found similarly lengthy paths following the evolution of a place or topic through time: 
+We also find similarly lengthy paths following the evolution of a place or topic through time: 
 "1953 in Scotland" or "1560s Architecture", with articles sequentially proceeding by year, decade or era.
-The longest paths connecting ideas are organized temporally.
+The longest paths connect temporally organized ideas.
 
@@
 articles in the XML dump are redirects or disambiguations, not full articles.
-The most common path length is 29, with an interquartile range (26, 30).
+The most common path length is 29, with an interquartile range of 4---path length
+is typically between 26 and 29 articles.
 The median path length far exceeds the popular 6-degrees of separation (see for 
@@
 a metric to gauge the accumulation of First Link references
-and characterize a global organization 
-of ideas among articles.
+and describe a pattern for how ideas flow through the FLN.
 
@@
 
-We followed every possible path through the network, incrementing
-the number of traversal visits for every First Link reference along a path 
-to generate a total of 232 million traversal visits.
-
-As a distribution, the number traversal visits per article appears to be scale-free. 
+We follow every possible path through the network, incrementing
+the number of traversal visits for every First Link reference.
+We generate a total of 232 million traversal visits.
+
+As a distribution, the number traversal visits by article appears to be scale-free. 
 The majority of articles have fewer than 30 traversal visits. 
@@
   The model explains $86\%$ of the variation in the data, yielding an r-value of $-0.93$ 
-  and a power law exponent of $-0.636$. The horizontal flattening around the highest
+  and a power law exponent of $-0.64$. The horizontal flattening around the highest
   ranking articles is a result of the cyclic structure (see discussion on cycles).
@@
   by the accumulation of First Links. Articles with a greater number of traversal visits
-  mark greater points of First Link accumulation. The highest ranking articles by traversal visits reveals where the greatest accumulation occurs.
+  mark greater points of First Link accumulation. The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.
   \label{fig:highest visits}
@@
 Since traversal visits measure the accumulation of First Link visits in Wikipedia, 
-the highest ranking articles suggest a flow from specific to general. 
+the highest ranking articles suggest First Links flow from specific to general articles. 
 For example the flow of First Links for the "Banana" article begins at a very concrete
-and specific topic, then flows into progressively broader disciplines, eventually 
+and specific topic, then flows towards progressively broader disciplines, eventually 
 culminating at "Philosophy": "Banana" links to the broader category of "Fruit," which then links to 
@@
     the distribution of synonyms of the highest-ranking 100 articles
-    against all other Wikipedia articles. As the distribution demonstrates
+    against all other Wikipedia articles (with the top 100 articles removed). As the distribution demonstrates
     the highest ranking concepts have more synonyms than other articles.
@@
 \cite{wordnet}. 
+We compare the highest ranking 100 articles against the all other Wikipedia articles, with the top 100 articles removed.
 We find the highest ranking 100 articles have on average 5 more synonyms than the typical article; a difference of 2.5 synonyms if we compare the median number of synonyms in each group
@@
 As suggested by the median, many articles have no synonyms as we might expect, because titles with more than one word are not likely to appear in a thesaurus.
-Since many articles have no synonyms, we also compared the number of synonyms in the highest-ranking versus typical article, this time excluding all articles without at least 1 synonym. 
-We still find the highest ranking 100 articles with an average of 9.0 (median of 7.0) synonyms whereas the remaining articles have on average 5.8 synonyms (median of 7.0), even with the exclusion of all articles without any synonyms.
+
+Since many articles have no synonyms, we also compare the number of synonyms for the highest-ranking versus typical article, this time excluding all articles without at least 1 synonym. 
+We still find the highest ranking 100 articles have on average 9.0 
+(median of 7.0) synonyms whereas the remaining articles have 
+on average 5.8 synonyms (median of 3.0)---even when all articles without
+synonyms are excluded.
 The quantifiable difference in synonyms corroborates the flow of links from concrete, specific articles to broader disciplines or fundamental notions.
@@
 cycle of 6 other articles including "Experiment", "Fact", and "Knowledge".
-To confirm the suggested cyclic structure, we record the history of articles traversed along a path. For example starting on "Train" we construct an path of articles: 
+To confirm the suggested cyclic structure, we record the history of articles traversed along a path. For example starting on "Train" we construct a path of articles: 
 "Rain Transport,"
@@
 "Quality,"
-"Philosophy."
+"Philosophy," and so on until an article is repeated.
 
@@
 
-Similarly, 3-cycles captured a synonymous or close relation among 3 articles: "Tree of life (Biology)", "Tree of life (disambiguation)", 
+Similarly, 3-cycles capture a synonymous or close relation among 3 articles: "Tree of life (Biology)", "Tree of life (disambiguation)", 
 and "Tree of life"; "Cinema of India", "Indian Cinema", and "Telugu Cinema"
@@
 "Philosophy" along with the remaining list of high ranking articles by traversal visits dominate.
+The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day.
+Other lengthy cycles span 60-75 articles including collections of articles on national histories such as "Japanese Eras" 
+or judicial bodies such as the "Legislative Assembly of Ontario".
 
@@
 \end{figure*}
-
-Extending beyond 3-cycles we find the highest ranking cycles of length 
-up to 10 links are dominated by "Philosophy," high ranking 3-cycles
-around the "Balkans" and "Tree of life" as well as popular 2-cycles.
-The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day.
-Other lengthy cycles span 60-75 articles including collections of articles on national histories such as "Japanese Eras" 
-or judicial bodies such as the "Legislative Assembly of Ontario".
 
@@
 ideas also rank highly such as "Tree of life" (30 thousand), 
-"Reality" (13 thousand), "Jurisdiction" (3 thousand) also appear of the list.
-
-Curiously, many high-ranking articles are remarkably topical, culturally and politically important ideas.
+"Reality" (13 thousand), and "Jurisdiction" (3 thousand).
+
+Many high-ranking articles are remarkably topical, culturally and politically important ideas.
 For example, "Health Care", a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August-2009 and February-2010.
@@
 The highest-ranking list even includes "Hip Hop," "Cancer," and "Web Page."
-This coincidence of recent relevance and traversal feed rank suggests the First Link Network measurably represents
+This coincidence of recent relevance and traversal funnel rank suggests the First Link Network measurably represents
 meaningful relationships not only among ideas, but also to English-speaking society. 
@@
     \textbf{Distribution of Traversal Funnels.}
-  We fit the distribution to a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal funnels. 
+  We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal funnels. 
   We find two regimes with the top regime (log(rank) < 4) 
@@
 October 2015---the earliest full month for which the data is available. 
-We find the highest ranking 1000 articles by traversal visits have an average of
+We find the highest ranking 1000 articles (by traversal visits) have an average of
 $70$ thousand page views in October with high variation: the standard deviation 
@@
 
-The "Philosophy" article despite outranking every article in traversal visits,
+The "Philosophy" article despite outranking every article by traversal visits,
 has only $240$ thousand page views in October compared to the $1.4$ million
@@
 among articles influencing the shape of the network, but less popular among 
-articles with a large accumulation among the flow of First Links.
+highly-ranked articles by accumulation.
 

### 2016-02-07 writeup/wikipedia-network.tex
cleaned writeup directory pre peter round 1 edits
edits up to article popularity

--- 2016-01-03 writeup/2015-11-wikipedia-network.tex
@@
 \title{\protect
-Connecting Every Bit of Knowledge: \\
-Wikipedia's First Link Network Unravelled
+Connecting every bit of knowledge: \\
+The Structure of Wikipedia's first link network
 }
+
+\date{\today}
 
@@
   \protect
-Apples, oranges, and the most obscure Dylan song too---is every topic a few clicks from Philosophy? 
+Apples, porcupines, and the most obscure Dylan song---is every topic a few clicks from Philosophy? 
 Within Wikipedia, the surprising answer is yes: nearly all 
@@
 Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
-More than information about a topic though, Wikipedia is a web of naturally emerging relationships.  
-By following the First Link in an article, we connect entries to form a directed network: Wikipedia's First Link Network. 
-Here we study the English edition of Wikipedia's First Link Network for insight into how the many topics, ideas, people, objects, and events
+More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
+By following the first link in an article, we connect entries to form a directed network: Wikipedia's First Link Network. 
+Here, we study the English edition of Wikipedia's First Link Network for insight into how the many topics, ideas, people, objects, and events
 are related and organized.  
-
 We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
-In a novel approach to uncover structure, we traverse every possible path through the network, 
-measuring the accumulation of First Links, path lengths, basins, cycles, and even the influence a particular article exerts in shaping the 
+We traverse every possible path through the network, 
+measuring the accumulation of first links, path lengths, basins, cycles, and the influence a particular article exerts in shaping the 
 network.
@@
 These findings enrich our view of how we connect and structure
-an ever growing load of information.
+Wikipedia's ever growing store of knowledge.
 
@@
 
-\section{Introduction}
-
-Wikipedia is a towering achievement of the modern era. 
-At no point in history has a larger or more meticulously indexed collection of human knowledge 
-existed.
-Wikipedia contains 37 million articles in 283 languages, 
-with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
-\cite{drugs} 
-\cite{stats}.
-All this information does not lie dormant. 
-Wikipedia is the sixth most visited
-site in the world, surpassing 18 billion page views and 26 million edits in a single month
-\cite{wiki_edits}
-\cite{wiki_views}.
-The most widely used general reference, Wikipedia has amassed an awe-inspiring collection of human knowledge.
-
-It's not surprising then to see Wikipedia as the object of many studies. 
-Researchers have examined the cultural dynamics among editors
-\cite{editors},
-the accuracy of the content relative to traditional encyclopedias
-\cite{accuracy1}
-\cite{accuracy2},
-the topics covered 
-\cite{coverage},
-and bias against portions of the population
-\cite{bias_women}.
-Wikipedia's content has also proven a powerful tool. 
-Researchers have used Wikipedia to identify missing dictionary entries
-\cite{missing_entries},
-cluster short text
-\cite{clustering},
-compute semantic relatedness
-\cite{semantic_relatedness},
-and disambiguate meaning \cite{disambiguating}.
-
-While these many studies have dissected and fruitfully applied Wikipedia's content,
-few have examined the connections among the many articles.
-Beyond information, Wikipedia is a web 
-relating ideas. 
-Within the body text of Wikipedia articles, hyperlinks reference
-other Wikipedia articles.
-A hyperlink from one Wikipedia article to another
-indicates a relationship between the two articles
-\cite{relevance}.
-The notion that hyperlinks convey information about the content of a
-page has proved successful in numerous domains from search engine algorithms 
-such as PageRank 
-\cite{pagerank} 
-to topic classification
-\cite{classifier}.
-We treat a hyperlink as a mechanism connecting two ideas.
-
-The authors of a Wikipedia article choose where and whether to include a 
-reference to another Wikipedia article
-in the HTML markup.
-In describing a particular topic, the authors decide which other articles are most
-relevant references.
-For example, the authors of the 
-\href{https://en.wikipedia.org/wiki/Train}{"Train"}
-article chose
-"Rail Transport," "Cargo," and "Passengers" 
-as pertinent articles to reference in describing trains
-\cite{wiki_train}
-.
-The collection of hyperlinks emerges from the independent choice of each article's authors. 
-The first hyperlink to another Wikipedia article marks the earliest moment in the body text where the authors
-choose to connect the information discussed to that of another article.
-By following the first hyperlink to another article in the English edition of
-Wikipedia, we connect the ideas in one article to those in another, ultimately forming a directed network: 
-{\it Wikipedia's First Link Network} (FLN).
-
-\begin{figure*}[tp!]
-  \centering	
-  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
-  \caption{
-    \textbf{First Link Path For "Train".}
-    We follow the First Link to another Wikipedia article
-    in the main body of the article---the area
-    inside the green rectangle, which excludes 
-    side bar elements, the navigation bar and title. 
-    In this example, the First Link to another Wikipedia article is "Rail Transport." 
-    We can again select the First Link on the "Rail Transport"
-    article, repeating the process to 
-    form a path of First Links.
-  }
-  \label{fig:Train First Links}
-\end{figure*}
-
-What information could the First Links possibly reveal?
-Inspired by the claim that the majority of First Links lead to the 
-"Philosophy" article---popularized by an 
-\href{https://xkcd.com/903/}{xkcd}
-comic and subsequently
-discussed in blog posts 
-\cite{mat_blog, Ilmari_first_links, xkcd}
----we holistically study 
-Wikipedia's FLN as a map relating areas of human knowledge. 
-The map is a hierarchical taxonomy where, for example, "Train" links to a parent node, 
-"Rail Transport," while many child nodes, "Steel" and "Horsepower," link to 
-"Train." Unlike previous taxonomies
-created by individuals
-\cite{locke, descartes, aristotle}
-or a select group of individuals 
-\cite{hist_thesaurus}, 
-the organization of ideas in the FLN 
-does not emerge from a concerted or a centralized effort. 
-Rather each link in the FLN emerges from the choice of each author.
-Collectively, the FLN is a wealth of relations among inventions, places,
-figures, objects, and events across space and time.
-
-Our goal is to study the structure of the FLN for insight into how the information on Wikipedia is organized and related.
-In a novel approach, we develop metrics to capture 
-the dominant features of the FLN's structure.
-We measure dynamics of the FLN as a flow, quantifying 
-aspects of the FLN from the accumulation of First Links around articles 
-to the influence an article exerts in shaping the FLN.
-Together with cycles, indegree, depth, and the content of the articles, 
-we build our analysis of the relations among the ideas in Wikipedia.
+TODO: 
+* fix end of second paragraph in Results Section
+    * power-law biz and total peter is asking for
+* fix exponent reporting (with alpha) 
+    * figure out what gamma size freq means
 
@@
 
-One essential measure of a directed network's structure is the degree distribution 
+An essential feature of a directed network's structure is the degree distribution 
 \cite{newman}. 
-Degree distribution has been used to study many phenomena from disease outbreak 
+The degree distribution has been used to study many phenomena from disease outbreak 
 \cite{disease} 
@@
 \cite{social_nets}.
-The degree distribution in the FLN describes how many First Links point to a 
+The in-degree distribution in the FLN describes how many first links point to a 
 particular article. 
-We measure the indegree in the FLN to gauge the number of articles referencing a particular article.
-A higher indegree indicates more authors selected the particular article
-as a relevant reference. On the other hand, articles with zero 
-indegree have no references---they are outer leaves in the FLN. 
-Similar to PageRank, the indegree is also a way to rank the articles in Wikipedia.
-We can rank articles by indegree to see which articles are most and least referenced in the FLN. 
-Additionally, we can assess the degree distribution to see whether authors tend to reference all articles
-equally or only a few, as observed in many social networks \cite{social_nets}.
+We measure the in-degree in the FLN to gauge the number of articles referencing a particular article.
+Articles with zero in-degree have no references---they are outer leaves in the FLN. 
+Similar to PageRank, the in-degree 
+provides a way to rank the articles in Wikipedia, allowing us to see 
+whether authors tend to reference all articles equally or only a few.
 
-The degree distribution captures one step in the FLN by measuring   
-the destination of a given link. 
-Indegree measures only the particular set of First Links between two articles, 
-rather than the richer dynamics of how many articles fit together---how links 
-create a flow through the FLN.
-One way to expand the analysis beyond a single step is to map the full path of 
-an article's First Link. 
-Each article is both a destination and a link to another article, similar to an element 
-in a linked list data structure. 
-By following the chain of links we can construct a path through the FLN with each article 
-leading to the next through the First Link. 
-For example "Train" has a First Link to "Conveyance of Passengers and Goods," which is itself
-an article with a First Link to "Goods" and so on. The path starting at "Train" 
-is the collection of articles obtained by following the First Links starting at "Train"
-(see figure~\ref{fig:Train First Links}).
-The path contains "Goods," "Economics," "Social Science," and other articles related to 
-"Train," but not directly referenced by a First Link.
+Our greater interest is lies in the richer dynamics of how many articles fit together---how links create a flow through the FLN.
+We construct a path through the FLN with each article 
+leading to the next through the first link. 
+For example ``Train'' has a first link to ``Conveyance of Passengers and Goods,'' which is itself
+an article with a first link to ``Goods'' and so on. 
+As shown in Fig.~\ref{fig:Train First Links}, the path starting at ``Train'' 
+contains ``Goods,'' ``Economics,'' ``Social Science,''
+ultimately leading to ``Philosophy''.
 
@@
 
-Applying a similar concept with  
-varying starting articles, we form 
-paths through the FLN moving from one article to the next.
-A path in the FLN connects a group of articles, allowing us to measure how two articles relate,
-even if the articles are not directly linked. 
-Path-connected articles express a flow of concepts.
-The notion of paths allows us to map not only which articles link directly to "Philosophy" 
-for example, but also which ideas flow to "Philosophy":
-path-connected articles to "Philosophy."
-The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
 Starting at each article, we construct a path through the FLN, 
 ultimately mapping a flow of connections among articles.
+The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
 
-In a novel approach, we use the notion of a path through the FLN to describe 
-the structure of the FLN as a flow relating ideas. 
-Grounded in the dynamics of river networks,
-\cite{geo_basins}
-we characterize features of the FLN as a flow through First Links.
+We use the notion of a path through the FLN to describe 
+the structure of the FLN as a flow relating topics. 
+Grounded in the dynamics of river networks
+\cite{geo_basins},
+we characterize features of the FLN as a flow through first links.
 Previous studies have used flow to characterize the structure of river networks
@@
 \cite{food_webs}.
-With paths to mark flow through the FLN, we develop new metrics to 
+With paths to mark flow through the FLN, we develop metrics to 
 gauge the accumulation of references, 
@@
 
-Our goal is to map the flow of First Links by following the paths of First Links 
-starting at each article. 
-Since a path reveals how the ideas in one article may eventually be 
-anchored in another, the first metric we develop quantifies the accumulation of First Links.
+The first metric we develop quantifies the accumulation of first links.
 The algorithm begins by selecting an article, then traversing the path formed
-by following the First Links. Each time a First Link references an article, we increment a count
+by following the first links. Each time a first link references an article, we increment a count
 associated with the article. 
-We continue until the First Link is repeated or is invalid---defined here to mean outside of Wikipedia.
+We continue until the first link is retraced or is invalid---defined here to mean outside of Wikipedia.
 We select a second article and repeat the process until we have 
 constructed a path for each article in the network. We call the resulting count for each article the number of {\it traversal visits}. The number of traversal visits of an article 
-measures the number of references flowing to the ideas in the article.
+measures the number of references flowing to the ideas in the article---equivalent
+to basin area in geomorphology.
 
@@
 We can characterize the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
-(see figure~\ref{fig:Traversal Visits}), the path starting at article A is the 
-first column in the traversal vists matrix. 
+(Fig.~\ref{fig:Traversal Visits}), the path starting at article A is the 
+first column in the traversal visits matrix. 
 An entry of $1$ indicates the path contains a given article and 
-$0$ indicates the path doesn't. The columns identify the starting article of the path. 
-For example, the second column corresponds to the path starting at article B. 
-To compute the number traversal visits for article A we sum the corresponding row in our matrix.
-The total is the number of paths containing article A. 
-Analogous to the flow 
-in a river network, traversal visits measure the accumulation of water at point A. 
-By constructing a traversal visits matrix, we can similarly compute the number of traversal 
-visits by summing along the corresponding row. The traversal visits matrix for 
-Wikipedia consists of 121 million entries to account for each path through the FLN. 
-We measure the accumulation of First Link references by summing the corresponding row for the 
-article of interest.
+$0$ indicates the path does not.
+To compute the number of traversal visits for an article, we sum the corresponding row in our matrix.
+The traversal visits matrix for 
+our Wikipedia dataset consists of 121 million entries to account for each path through the FLN.
 
-In addition to accumulation, the traversal visits algorithm creates paths 
-which uncover depth in the FLN. 
-By measuring the number of First Links between
-two articles, we get an additional piece of information we call 
+By measuring the number of first links between
+two articles, we obtain an additional piece of information we call 
 {\it path length.} We can compute the length of a path by summing 
 along a column in our traversal visits matrix. 
-The path length describes how closely related ideas are, gauging the FLN's depth.
-The greater the number of First Links separating two articles, the greater the number of ideas needed to relate their corresponding ideas. Although
-"Train" is related to "Economics" for example, there are several articles
+In our Wikipedia dataset, the sum of all path lengths is 232 million first links.
+The path length describes how closely related topics are.
+The greater the number of first links separating two articles, the greater the number of ideas needed to relate their corresponding ideas. Although
+``Train'' is related to ``Economics'' for example, there are several articles
 bridging the connection:
-"Train" is more specifically related to transportation, whose object
+``Train'' is more specifically related to transportation, whose object
 is often goods. Goods are one of the fundamental objects of study in Economics.
 Described in links, this relationship is captured by a relatedness of $4$ 
-First Links ultimately connecting "Train" to "Economics." 
-We can examine path length as a relative measure of how closely connected ideas
-may be, but also as an aggregate measure describing the overall separation or 
-connectedness in the FLN.
-We can characterize whether ideas are closely connected by a handful of links, 
-many, or whether a more nuanced organization exists.
+first links ultimately connecting ``Train'' to ``Economics.''
 
@@
 linking to one another inside a loop. In our sample network 
-(see figure~\ref{fig:Traversal Visits}) 
-a cycle exists among nodes A, B, and C. 
-Since node C links back to the starting node A, the path starting at node A forms a {\it 3-cycle}. 
-By recording the history of the articles along a path in the FLN, we can identify
-the types of cycle structures within the FLN. Furthermore, since each article 
-has an associated traversal visit count, we can rank cycles by 
+(Fig.~\ref{fig:Traversal Visits}) 
+a 3-cycle exists among nodes A, B, and C. 
+We can readily identify
+the types of cycle structures within the FLN
+and rank each by 
 the number of references directed towards a cycle.
+
 We can also form and rank {\it basins} in the network by identifying groups of 
@@
 to a particular article.
-By identifying basins and cycles, we pinpoint groups of closely related articles.
 
-While traversal visits measure accumulation, each article's First Link also 
+While traversal visits measure accumulation, each article's first link also 
 influences the shape of the FLN. 
-At a large point of accumulation, a single article's First Link 
+At a large point of accumulation, a single article's first link 
 can exert great influence over the shape of the FLN by directing many
@@
 that simply happened to fall within a cycle from an article funneling 
-many First Links, we develop a second metric called {\it traversal funnels}.
+many first links, we develop a second metric called {\it traversal funnels}.
 
@@
 We are then able to distinguish between an article related to many other ideas
-only by virtue of its place in a cycle, from an article exerting influence over where the First Links flow. 
+only by virtue of its place in a cycle, from an article exerting influence over where the first links flow. 
 An article with a large number of traversal funnels directs many references
-towards a particular cycle. In our sample network 
+into a particular cycle. In our sample network 
 (see figure~\ref{fig:Traversal Funnels}) article C 
@@
 recipients of the flow---without exerting any direct influence themselves.
-While articles A, B, and C all have the same number of traversal visits, 
-article C has 4 traversal funnels (A and B have none). By incrementing the count of 
-First Link references on paths only up to a cycle, we can identifying which articles
-exert the greatest influence over the FLN's structure.
 
@@
 as a flow, 
-we build a powerful arsenal of information with which to study the FLN. 
-From accumulated references, cycles and basins to influence, we can measure how the many articles in Wikipedia are organized and related.
+we build a powerful arsenal of information.
+From accumulated references, cycles, and basins, we can measure how the many articles in Wikipedia are organized and related.
 
-\section{Discoveries}
+\section{Results}
 
@@
   \caption{
-    \textbf{Highest Ranking Articles by indegree.}
+    \textbf{Highest Ranking Articles by in-degree.}
   }
-  We rank each article by the number of direct First Links to the article (indegree). The highest-ranking articles tend to represent abstraction for concepts
+  We rank each article by the number of direct first links to the article (in-degree). The highest-ranking articles tend to represent abstraction for concepts
   comprised of various facets.
@@
 
-We identify the set of First Links directly referencing a particular article 
-by studying indegree in the FLN.
-A higher indegree suggests more authors found the ideas in the article 
-pertinent to reference. 
-We rank all 11 million articles by indegree to find 
-the "United States" with 80,249 direct First Links as the most referenced
+We rank all 11 million articles by in-degree to find 
+the ``United States'' with 80,249 direct first links as the most referenced
 Wikipedia article 
-(see figure~\ref{fig:ndegree list}). 
+(Fig.~\ref{fig:ndegree list}). 
 Other high-ranking articles
-include foundational abstract concepts such as "village," "species,"; 
-sports associations such as "American Football," "Association Football"; 
-and developed nations such as "France," "Japan," "Russia," "Australia," and 
-the "Netherlands." These high ranking articles are useful abstractions: nations
+include foundational abstract concepts such as ``village,'' ``species,''; 
+sports associations such as ``American Football,'' ``Association Football''; 
+and developed nations such as ``France,'' ``Japan'', ``Russia,'' ``Australia,'' and 
+the ``Netherlands.'' These high ranking articles are useful abstractions: nations
 describe a collection of individuals with a common culture, language, or 
@@
 In describing the New England Patriots or the New York Giants
-for example, "American Football" is the natural abstraction, which 
+for example, ``American Football'' is the natural abstraction, which 
 anchors many specific teams.
 
-"Philosophy" and other philosophical concepts
-are not among the highest-ranking articles by indegree.
-"Philosophy" has an indegree of 581, with direct First Links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W.
-Morris."
-While many articles accumulate at "Philosophy" (see the coming traversal visits discussion), 
+``Philosophy'' and other philosophical concepts
+are not among the highest-ranking articles by in-degree.
+``Philosophy'' has an in-degree of only 581, with direct first links from articles about Philosophers and areas of Philosophy: ``Existentialism and Humanism,'' ``Predeterminism,'' ''Synoptic Philosophy,'' ``Qualia,'' ``Dorothy Emmet,'' and ``Christopher W.
+Morris.''
+While many articles accumulate at ``Philosophy'' (see traversal visits discussion below), 
 the accumulation is not the 
-result of many articles directly referencing "Philosophy." 
-Instead, the accumulation of First Links, as we argue in our 
-discussion of traversal visits, flows towards Philosophy as the 
+result of many articles directly referencing ``Philosophy.'' 
+Instead, flows towards Philosophy as the 
 ultimate anchor when generalizing from specific articles to broader notions.
@@
 
-The FLN's indegree exhibits a scale-free distribution where a few articles 
-receive most direct first references; most articles receive few or none.
-The average indegree for all 11 million articles is 3.6 direct First Links with a standard deviation of 89.5 links.
-Only 4,826 articles have more than 100 direct First Links and $75\%$ of articles
+The FLN's in-degree exhibits a decaying power-law distribution where a few articles 
+receive most direct first references, while most articles receive few or none.
+The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5 links.
+Only 4,826 articles have more than 100 direct first links and $75\%$ of articles
 have fewer than 9. 
-When fit with a linear model on a log-log scale (log(rank) versus log(indegree)), 
+When fit with a linear model on a log-log scale (log(rank) versus log(in-degree)), 
 the model's r value is $-0.98$ suggesting a strong log-log linear fit 
 with a power law exponent of $-0.79$.\\
-
-\centerline{***}
-
-Beyond indegree, we develop the notion of a path to describe the structure of the FLN as a 
-flow relating ideas. We traverse every possible path through the FLN, taking 
-$232$ million steps along the way.
 
@@
 We first seek to describe the depth of the FLN: how many links does a 
-connection of ideas span? We gauge the FLN's depth by 
-measuring path length or the number of links traversed until a repeated or invalid link (one outside the FLN). 
-We discover the longest path length is 365,
+connection of among articles span? 
+We find the longest path length is 365,
 corresponding to the yearly calendar of Orthodox Liturgics.
@@
 We also find similarly lengthy paths following the evolution of a place or topic through time: 
-"1953 in Scotland" or "1560s Architecture", with articles sequentially proceeding by year, decade or era.
-The longest paths connect temporally organized ideas.
+``1953 in Scotland'' or ``1560s Architecture,'' with articles sequentially proceeding by year, decade or era.
+In general, the longest paths connect temporally organized ideas.
 
@@
   }
-  The median network depth measured by the number of First Links in a path
-  is 29. The 365 cycle for "Orthodox" liturgics is the outlier to the right, while
-  other historical articles about "Scotland," "UK" and so on are slightly outside
+  The median network depth measured by the number of first links in a path
+  is 29. The 365 cycle for ``Orthodox'' liturgics is the outlier to the right, while
+  other historical articles about ``Scotland,'' ``UK'' and so on are slightly outside
   the third quartile. More than $75\%$ of articles have path lengths between 
@@
 is typically between 26 and 29 articles.
-The median path length far exceeds the popular 6-degrees of separation (see for 
-example the average number of academic publications separating scholars 
-\cite{six_degrees}), indicating a greater network depth.
-The network's depth suggests ideas are not related in small independent clusters. 
-Instead the many diverse ideas fit together in long-chained connections.
 As a distribution, more than $75\%$ of articles have a path length below 
-$50$ First Links 
+$50$ first links 
 while a few temporally organized paths exceed 50 links 
@@
 
-How can we characterize the flow of ideas along a path? Next, we develop 
-a metric to gauge the accumulation of First Link references
-and describe a pattern for how ideas flow through the FLN.
-
 \subsection{Traversal Visits}
 
-We follow every possible path through the network, incrementing
-the number of traversal visits for every First Link reference.
-We generate a total of 232 million traversal visits.
-
-As a distribution, the number traversal visits by article appears to be scale-free. 
-The majority of articles have fewer than 30 traversal visits. 
-First Link references accumulate at a few articles.
+As a distribution, the number of traversal visits by article appears to follow a decaying power law. 
+The majority of articles have fewer than 30 traversal visits and
+first link references accumulate at a few articles.
 Specifically, $99.76\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
@@
 
-To more accurately gauge the distribution, we construct a log-log plot of the entire dataset: log(traversal visits) against log(rank). 
-We observe a strong linear fit, as is characteristic of scale-free networks, with an r-value of -.93 and 
+In Fig., we show a log-log plot for the entire dataset:traversal visits against rank. 
+We observe a strong linear fit, with a Pearson's Correlation Coefficient of -.93 and 
 a power-law exponent of $-0.636$.
@@
   Traversal Algorithm section for details). In doing so, we can rank each article
-  by the accumulation of First Links. Articles with a greater number of traversal visits
-  mark greater points of First Link accumulation. The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.
+  by the accumulation of first links. Articles with a greater number of traversal visits
+  mark greater points of first link accumulation. The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.
   \label{fig:highest visits}
@@
 
-\subsubsection{A flow from general to specific}
-
-The highest ranking articles by traversal visits are 
-broad, global topics
-(see figure~\ref{fig:highest visits}):
-many are academic disciplines such as "Science," "Math," 
-"Geography," "Philosophy," "Biology," and "Physics"; others are abstract fundamental concepts such as 
-"Community", "State", "Earth", "Information", 
-"Existence," "Communication", and "Power."
-Since traversal visits measure the accumulation of First Link visits in Wikipedia, 
-the highest ranking articles suggest First Links flow from specific to general articles. 
-For example the flow of First Links for the "Banana" article begins at a very concrete
-and specific topic, then flows towards progressively broader disciplines, eventually 
-culminating at "Philosophy": "Banana" links to the broader category of "Fruit," which then links to 
-"Botany," eventually "Biology," then "Science" and ultimately "Philosophy."
-
-\begin{figure}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/synonyms.png}
-  \caption{
-    \textbf{Number of Synonyms for the highest-ranking articles.}
-    We use WordNet (the largest lexical database for the English language)
-    to identify the number of synonyms for an article's title. We compare 
-    the distribution of synonyms of the highest-ranking 100 articles
-    against all other Wikipedia articles (with the top 100 articles removed). As the distribution demonstrates
-    the highest ranking concepts have more synonyms than other articles.
-  }
-  \label{fig:synonyms}
-\end{figure}
-
-To measure the specificity of an article, we identify the number of synonyms available for a word or topic. 
-A  broader topic likely has many more synonyms relative to a specific, concrete topic. 
-Banana has fewer synonyms than botanical for example. 
-To quantify the observed flow, we measured the number of synonyms each title contained in WordNet---the largest lexical database of the English language 
-\cite{wordnet}. 
-We compare the highest ranking 100 articles against the all other Wikipedia articles, with the top 100 articles removed.
-We find the highest ranking 100 articles have on average 5 more synonyms than the typical article; a difference of 2.5 synonyms if we compare the median number of synonyms in each group
-(see figure~\ref{fig:synonyms}). 
-As suggested by the median, many articles have no synonyms as we might expect, because titles with more than one word are not likely to appear in a thesaurus.
-
-Since many articles have no synonyms, we also compare the number of synonyms for the highest-ranking versus typical article, this time excluding all articles without at least 1 synonym. 
-We still find the highest ranking 100 articles have on average 9.0 
-(median of 7.0) synonyms whereas the remaining articles have 
-on average 5.8 synonyms (median of 3.0)---even when all articles without
-synonyms are excluded.
-The quantifiable difference in synonyms corroborates the flow of links from concrete, specific articles to broader disciplines or fundamental notions.
-
 \subsection{Network Cycles}
 
-The recurrence of an exact number of traversal visits suggests some articles are part of a cycle. 
-The "Philosophy" article for example sits in what seems like a cycle of seven other articles; "Hypothesis" appears to sit in a 
-cycle of 6 other articles including "Experiment", "Fact", and "Knowledge".
-To confirm the suggested cyclic structure, we record the history of articles traversed along a path. For example starting on "Train" we construct a path of articles: 
-"Rain Transport,"
-"Conveyance of Passengers and Goods," 
-"Goods," 
-"Economics,"
-"Social Science,"
-"Academic Disciplines,"
-"Knowledge,"
-"Awareness,"
-"Consciousness,"
-"Quality,"
-"Philosophy," and so on until an article is repeated.
-
-We first identified 2-cycles, meaning a pair of articles with First Link pointing to one another.
+We first identified 2-cycles, meaning a pair of articles with first link pointing to one another.
 Of the 11 million articles, 84,000 are 2-cycles. 
 The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected ideas:
-"Health Care" and "Medical Case Management", "Broadcasting House" and "BBC", "Secondary Education" and "Secondary School" 
-(see figure~\ref{fig:2-cycles}).
+``Health Care'' and ``Medical Case Management,'' ``Broadcasting House,'' and ``BBC,'' ``Secondary Education'' and ``Secondary School'' 
+(see Fig.~\ref{fig:2-cycles}).
 
@@
     \textbf{Highest ranking 2-cycles.}
-    We identify pairs of articles whose First Links point to one another, forming
+    We identify pairs of articles whose first links point to one another, forming
     a 2-cycle. We then rank each pair of articles by the total number of 
@@
 Outside of the highest ranking 2-cycles, the typical 2-cycle signals a connection between distinct, yet very closely related ideas. 
-We also observe link patterns such as inventor to product ("Voere" to "VEC-91"), event to organizer ("Poetry Bus Tour" to "Weave Books"), and book to author ("Anatomy of Britain" to "Anthony Sampson").
+We also observe link patterns such as inventor to product (``Voere'' to ``VEC-91''), event to organizer (``Poetry Bus Tour'' to ``Weave Books''), and book to author (``Anatomy of Britain'' to ``Anthony Sampson'').
 
-Similarly, 3-cycles capture a synonymous or close relation among 3 articles: "Tree of life (Biology)", "Tree of life (disambiguation)", 
-and "Tree of life"; "Cinema of India", "Indian Cinema", and "Telugu Cinema"
+Similarly, 3-cycles capture a synonymous or close relation among 3 articles: ``Tree of life (Biology),'' ``Tree of life (disambiguation),'' 
+and ``Tree of life''; ``Cinema of India,'' ``Indian Cinema,'' and ``Telugu Cinema''
 (see figure~\ref{fig:3-cycles}).
 Once we extend our cycle size beyond a length of 6 however, 
-"Philosophy" along with the remaining list of high ranking articles by traversal visits dominate.
+``Philosophy'' along with the remaining list of high ranking articles by traversal visits dominate.
 The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day.
-Other lengthy cycles span 60-75 articles including collections of articles on national histories such as "Japanese Eras" 
-or judicial bodies such as the "Legislative Assembly of Ontario".
+Other lengthy cycles span 60--75 articles including collections of articles on national histories such as ``Japanese Eras'' 
+or judicial bodies such as the ``Legislative Assembly of Ontario.''
 
@@
 The highest ranking basins by the number of traversal visits are groups of articles
-around "Philosophy." 
-Since the accumulation of First Links to "Philosophy" is so high, 
-the paths leading to "Philosophy" carry many references.
+around ``Philosophy.'' 
 The highest ranking paths include branches of philosophy flowing through 
-"Awareness,"Existence," and "Consciousness" to "Philosophy." Other paths
-include concepts around "Mathematics," scientific "Experiments," 
-"Biology," and "Fact."
-These paths link many specific articles to "Philosophy," each funneled through a particular domain.
+``Awareness,'' ``Existence,'' and ``Consciousness'' to ``Philosophy.'' Other paths
+include concepts around ``Mathematics,'' ``Scientific,'' ``Experiments,'' 
+``Biology,'' and ``Fact.''
+These paths link many specific articles to ``Philosophy,'' each funneled through a particular domain.
 
-Excluding basins around "Philosophy," we find other basins around 
-foundational ideas such as "Community," "Landmass," "Federal Government," 
-"Presentation," and "Belief System." 
+Excluding basins around ``Philosophy,'' we find other basins around 
+foundational concepts such as ``Community,'' ``Landmass,'' ``Federal Government,'' 
+``Presentation,'' and ``Belief System.'' 
 The basins around each of these foundational notions are 
 various paths containing related articles. For example around 
-"Community" we find basins
-flowing from "United States" to "Federal Republic" to "Political Union" to "State" culminating at "Community"; we also find basins flowing from 
-"Public Policy" to "Executive (government)" to "Government" to State" and then 
-to "Community"; we also find paths flowing through a similar chain beginning
-with "Democracy," another beginning with "Constitution," another at 
-"Dictatorship," and so on. The ideas build from specific means of organizing
-a community (or society) and then build up to "Community." 
+``Community'' we find basins
+flowing from ``United States'' to ``Federal Republic'' to ``Political Union'' to ``State'' culminating at ``Community''; we also find basins flowing from 
+``Public Policy'' to ``Executive (government)'' to ``Government'' to ``State'' and then 
+to ``Community''; we also find paths flowing through a similar chain beginning
+with ``Democracy,'' another beginning with ``Constitution,'' another at 
+``Dictatorship,'' and so on. The ideas build from specific means of organizing
+a community (or society) and then build up to ``Community.'' 
 Other basins around landmass for example begin at specific geographical regions
-such as "Eastern Europe" building up to "Continent" and finally "Landmass."
-Each basin groups First Link references through various 
-natural paths linking general themes to the foundational concept.
-
-Which specific articles direct the path towards a particular concepts? 
-Next we study the FLN using traversal funnels to understand the influence
-a particular article exerts in directing the flow of First Link references.
+such as ``Eastern Europe'' building up to ``Continent'' and finally ``Landmass''.
 
@@
 Articles directing more paths exert a greater influence over the structure
-of the FLN by increasing the accumulation of First Links
+of the FLN by increasing the accumulation of first links
 on a particular path. By measuring traversal funnels, we distinguish between an article that simply happened to fall within a cycle from an article funneling 
-many First Links.
+many first links.
 
 Ranking articles by the number of traversal funnels we find 
-"Philosophy" as by far the highest-ranking article with 
+``Philosophy'' to be by far the highest-ranking article with 
 $7.37$ million paths
@@
     funnels to gauge the influence each article exerts in shaping the 
-    structure of the FLN. We find "Philosophy" exerts an overwhelming proportion
+    structure of the FLN. 
+    The area of each circle is sized in proportion to the number of 
+    traversal funnels each article has. 
+    We find ``Philosophy'' exerts an overwhelming proportion
     of the influence, with other abstract notions and topical concepts ranking
@@
 all others by more than two orders of magnitude.
-The "Philosophy" cycle which contains "Existence," "Awareness," "Reality," 
+The ``Philosophy'' cycle which contains ``Existence,'' ``Awareness,'' ``Reality,'' 
 and similar articles accumulates the overwhelming proportion of its 
-references through "Philosophy": $7.37$ million of the $7.4$ million references
-are funneled through "Philosophy."
+references through ``Philosophy'': $7.37$ million of the $7.4$ million references
+are funneled through ``Philosophy''.
 Second on the list of highest-ranking articles by traversal funnels is 
-"Presentation" with only $30$ thousand paths. Similarly abstract 
-ideas also rank highly such as "Tree of life" (30 thousand), 
-"Reality" (13 thousand), and "Jurisdiction" (3 thousand).
+``Presentation'' with only $30$ thousand paths. Similarly abstract 
+ideas also rank highly such as ``Tree of life'' (30 thousand), 
+``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).
 
 Many high-ranking articles are remarkably topical, culturally and politically important ideas.
-For example, "Health Care", a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August-2009 and February-2010.
-Other high ranking articles include key historical events such as the "Cold War" or critical scarce resource with recent 
-media discussion such as "Fossil Fuel". 
-The highest-ranking list even includes "Hip Hop," "Cancer," and "Web Page."
-This coincidence of recent relevance and traversal funnel rank suggests the First Link Network measurably represents
-meaningful relationships not only among ideas, but also to English-speaking society. 
+For example, ``Health Care'', a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August, 2009 and February, 2010.
+Other high ranking articles include key historical events such as the ``Cold War'' or critical scarce resource with recent 
+media discussion such as ``Fossil Fuel''. 
+The highest-ranking list also includes ``Hip Hop,'' ``Cancer,'' and ``Web Page''.
 \begin{figure}[tp!]
@@
 
-The flow of ideas in the FLN is overwhelmingly influenced by "Philosophy,"
-with a small proportion guided by "Presentation," "Tree of life (biology)," "Southeastern Europe," 
-"Feudalism," "Census-Designated Place," "United States Constitution," "Reality," and "Health Care."
-
-\subsection{Article Popularity}
-
-Wikipedia released an API to measure article popularity by page views
-starting November 2015. Page views add another dimension to our
-findings by revealing 
-how users view the information on Wikipedia. 
-We measure popularity as the total number 
-of page views in the English edition of Wikipedia in the month of 
-October 2015---the earliest full month for which the data is available. 
-We find the highest ranking 1000 articles (by traversal visits) have an average of
-$70$ thousand page views in October with high variation: the standard deviation 
-is $1.1$ million page views. 
-The number of page views has a skewed distribution with $75\%$ of articles
-reaching fewer than $73$ thousand page views in October.
-The article for "United States" has the most page 
-views of the highest ranking $1000$ articles by traversal visits with 
-$1.4$ million views in October. The next most popular articles are 
-"India," "World War II," and the "United Kingdom" each with $0.7-1.03$ million page views.
-
-The "Philosophy" article despite outranking every article by traversal visits,
-has only $240$ thousand page views in October compared to the $1.4$ million
-page views for "United States".
-While "Philosophy" is not itself the most popular article, "Philosophy" serves as 
-an accumulation point for ideas. Wikipedia users do not login specifically
-to read about "Philosophy," but the authors relate ideas to "Philosophy" 
-by ultimately anchoring the flow of First Links to "Philosophy."
-
-\begin{figure}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/views_visits.png}
-  \caption{
-    \textbf{Article Popularity (by page views) compared to Traversal Visits 
-    for Highest ranking 1000 articles.}
-    We use the total number of page views provided by Wikipedia for the month
-    of October 2015 to compare each article's popularity to traversal visits.
-    While the most popular articles do not necessarily correspond to the articles
-    with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-  }
-  \label{fig:Views and Visits}
-
-\end{figure}
-
-We also analyze article popularity for the highest ranking 1000 articles by 
-traversal funnels. In October, the average page views per article is 
-$22$ thousand with a standard deviation of $95$ thousand views. The distribution
-is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
-"Halloween" is the most popular article with $2.8$ million views in October,
-likely a result of Halloween falling in the month of October. 
-Other popular articles include "Scientology" ($463$ thousand views), "Clint Eastwood" 
-($341$ thousand views),
-and the "Cold War" ($298$ thousand views) although each has significantly fewer views compared to the views for "Halloween."
-Other standout popular October articles include "24-hour" and "12-hour" clocks likely due
-to Daylight savings in October and "Marriage" likely due to the drop in the number of
-weddings in the months following October
-\cite{weddings}.
-"Philosophy" which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears no where in the top 20
-by traversal visits. "Philosophy" is a relatively popular article 
-among articles influencing the shape of the network, but less popular among 
-highly-ranked articles by accumulation.
-
-\begin{figure}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/funnels_visits.png}
-  \caption{
-    \textbf{Article Popularity (by page views) compared to Traversal Funnels
-    for Highest ranking 1000 articles.}
-  }
-    We use the total number of page views provided by Wikipedia for the month
-    of October 2015 to compare each article's popularity to traversal funnels.
-    In the top left is "Halloween" with a significantly greater 
-    number of page views in the month of October.
-  \label{fig:Views and Funnels}
-\end{figure}
-
-%========================Conclusion====================================
-
-\section{Reflections}
-
-The findings here should only be considered within the limitations of their context.
-We examined only the English version of Wikipedia at a particular moment in time.
-Furthermore, we only studied the First Link in the main body  of each article
-as a means to related one article to another. Finally, Wikipedia, while the largest 
-collection of human knowledge, is rife with the biases of the many contributing editors
-\cite{bias}
-. Nevertheless, the findings do reveal
-generalizable relationships, point to foundational notions, and uncover many curiosities.
-
-Among the curiosities is the multiple appearance of scale-free distributions within the network. 
-The three metrics we developed: path length, traversal visits, and traversal funnels are all marked 
-by scale-free distributions. Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
-articles are responsible for funneling most paths. When measured against the traversal funnels, 
-"Philosophy" emerges as an exceptional article by orders of magnitude. 
-Nevertheless, many other foundational ideas emerged naturally within the First Link Network. 
-Basins around "Community", "State", and "Science" reveal a foundational structure within the network. 
-More curious is the emergence of recently prominent political and economics topics such as "Fossil Fuel" and "Health Care" 
-within the highest ranking funnels. 
-Wikipedia seems to reflect not only timeless foundations, but also the topical (at least within English speaking society).
-
-Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the First Link.
-These findings also form the basis for the creation of a taxonomy where 
-every idea, event, or object sits within a hierarchy of connected notions.
-The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
-Applications could range from an enhanced thesaurus of ideas to psychological insights into how humans form associations.
-Specifically, an ever-evolving reference of related hierarchical concepts can be applied to search engine algorithms 
-or natural language processing.
-
-\acknowledgments
-Thank you to my friend RJ for pointing out the Reddit post 
-\cite{reddit}
-describing how the majority of links lead to "Philosophy," inspiring this research.
-
-\newpage
-
-\section{Appendix}
-
-\subsection{Constructing The First Link Network}
-
-To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
-Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
-eliminating any statistical error due to sampling.
-We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
-The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
-and disambiguations.
-Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
-\cite{wiki_edits},
-our aim is to characterize the dynamics of the First Link Network, not record a particular link between one
-article and another.
-
-Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
-Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
-information.
-While a human could accurately identify the First Link, to map the entire First Link Network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
-
-While some libraries exist for MediaWiki Markup,
-Approaches using existing libraries led to several bugs 
-including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
-as well as compatibility with other libraries used to parse the XML.
-Consequently,  we developed an algorithm for parsing the First Link in the XML version of each article.
-Our parsing algorithm aimed to: 
-1) accurately identify the First Link among other page elements, and 
-2) efficiently do so---that is without needing for several passes through the data.
-To process an article in a single pass, we developed a hierarchical system of flags:
-
-\begin{figure}[tp!]
-  \centering	
-  \includegraphics[width=\columnwidth]{graphics/flags.pdf}  
-  \caption{
-    \textbf{Parsing Algorithm of Wikipedia's XML dump.}
-     The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
-Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
-Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.
-  }
-  \label{fig:parsing algorithm}
-\end{figure}
-
-The algorithm loops in three-character chunks to account for potentially nested elements, 
-shifting by one character steps through the article markup.
-If any markup triggers for a flag are detected, a flag is raised. 
-Once a flag is raised, we stop processing and proceed to the next character
-until the flag's closing markup.
-A First Link is identified only if Flags 1, 2, and 3 are all off.
-In this case, the entire link is retrieved. 
-We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
-as well as common file extensions for 
-images, audio files, and the like 
-\cite{media_wiki_templates}.
-The First Link of an article is then the earliest valid link with unraised flags.
-
-To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
-across 112 cores of the UVM supercomputer cluster
-\cite{vacc}
-We then joined the results to form a hash table containing every Wikipedia article and its corresponding
-First Link. The resulting network map is the basis of our analysis.
-
-\begin{thebibliography}{1}
-
-      \bibitem{editors} Iba, Takashi, et al.
-"Analyzing the creative editing behavior of Wikipedia editors: Through dynamic social network analysis." Procedia-Social and Behavioral Sciences 2.4 (2010): 6441-6456.
-
-      \bibitem{accuracy1} Holman Rector, Lucy. "Comparison of Wikipedia and other encyclopedias for accuracy, breadth, and depth in historical articles." Reference services review 36.1 (2008): 7-22.
-
-      \bibitem{accuracy2} Giles, Jim. "Internet encyclopaedias go head to head." Nature 438.7070 (2005): 900-901.
-
-       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. "An analysis of topical coverage of Wikipedia." Journal of Computer‐Mediated Communication 13.2 (2008): 429-440.
-
-        \bibitem{bias_women} Hill, Benjamin Mako, and Aaron Shaw.
-"The Wikipedia gender gap revisited: characterizing survey response bias with propensity score estimation." PloS one 8.6 (2013): e65782.
-
-      \bibitem{missing_entries} Williams, Jake Ryland, Clark, Eric M., Bagrow, James P., Danforth, Christopher M.
-\& Dodds, Peter Sheridan (2015).
-Identifying missing dictionary entries with frequency-conserving context models.
-Phys.
-Rev.
-E, 92, 042808.
-
-      \bibitem{clustering} Banerjee, Somnath, Krishnan Ramanathan, and Ajay Gupta.
-"Clustering short texts using wikipedia." Proceedings of the 30th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2007.
-
-      \bibitem{semantic_relatedness} Gabrilovich, Evgeniy, and Shaul Markovitch. "Computing Semantic Relatedness Using Wikipedia-based Explicit Semantic Analysis." IJCAI. Vol. 7. 2007.
-
-      \bibitem{disambiguating} Cucerzan, Silviu. "Large-Scale Named Entity Disambiguation Based on Wikipedia Data." EMNLP-CoNLL. Vol. 7. 2007.
-
-      \bibitem{drugs} Clauson, Kevin A., et al. "Scope, completeness, and accuracy of drug information in Wikipedia." Annals of Pharmacotherapy 42.12 (2008): 1814-1821.
-
-      \bibitem{stats} Wikipedia contributors. "Wikipedia:Statistics." Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 9 Nov. 2015. Web. 13 Nov. 2015.
-
-      \bibitem{pagerank} Page, Lawrence, et al. "The PageRank citation ranking: bringing order to the Web." (1999).
-
-      \bibitem{classifier} Chakrabarti, Soumen, Mukul Joshi, and Vivek Tawde.
-"Enhanced topic distillation using text, markup tags, and hyperlinks." Proceedings of the 24th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2001.
-
-      \bibitem{relevance} Kamps, Jaap, and Marijn Koolen. "Is Wikipedia link structure different?." Proceedings of the second ACM international conference on Web search and data mining. ACM, 2009.
-
-      \bibitem{wiki_train} Wikipedia contributors. "Train." Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 
-          \url{https://en.wikipedia.org/wiki/Train}
-
-    \bibitem{locke} Bolton, Martha Brandt.
-"The Taxonomy of Ideas in Locke’s {\it Essay} ", The Cambridge Companion to Locke's 'Essay Concerning Human Understanding'.
-Ed.Lex Newman..
-1st ed.
-Cambridge: Cambridge University Press, 2007.
-67-100.
-Cambridge Companions Online.
-Web.
-13 November 2015.
-http://dx.doi.org/10.1017/CCOL0521834333.004.
-
-    \bibitem{aristotle} Studtmann, Paul, "Aristotle's Categories", The Stanford Encyclopedia of Philosophy (Summer 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/sum2014/entries/aristotle-categories/}
-
-    \bibitem{descartes} Smith, Kurt, "Descartes' Theory of Ideas", The Stanford Encyclopedia of Philosophy (Spring 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/spr2014/entries/descartes-ideas/}
-
-    \bibitem{hist_thesaurus} What is the Historical Thesaurus of the OED - Oxford English Dictionary (Oxford English Dictionary)
-    http://public.oed.com/historical-thesaurus-of-the-oed/what-is-the-historical-thesaurus-of-the-oed/
-
-    \bibitem{vacc} The authors acknowledge the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
-    \url{http://www.uvm.edu/vacc}
-
-    \bibitem{wiki_views} Page Views for Wikipedia, Non-mobile site, Normalized (Page Views for Wikipedia, Non-mobile site, Normalized)
-    \url{https://stats.wikimedia.org/EN/TablesPageViewsMonthly.htm}
-
-    \bibitem{weddings} Wedding statistics in the United States (Wedding statistics in the United States)
-    \url{http://www.soundvision.com/article/wedding-statistics-in-the-united-states}
-
-    \bibitem{six_degrees} Elmacioglu, Ergin, and Dongwon Lee. "On six degrees of separation in DBLP-DB and more." ACM SIGMOD Record 34.2 (2005): 33-40.
-
-    \bibitem{wordnet} Princeton University "About WordNet." WordNet. Princeton University. 2010. 
-    \url{http://wordnet.princeton.edu}
-
-    \bibitem{newman} Newman, Mark EJ. "The structure and function of complex networks." SIAM review 45.2 (2003): 167-256.
-
-    \bibitem{disease} Eubank, Stephen, et al. "Modelling disease outbreaks in realistic urban social networks." Nature 429.6988 (2004): 180-184.
-
-    \bibitem{social_nets} Newman, Mark EJ, Duncan J.
-Watts, and Steven H.
-Strogatz.
-"Random graph models of social networks." Proceedings of the National Academy of Sciences 99.suppl 1 (2002): 2566-2572.
-
-    \bibitem{geo_basins} Horton, Robert E.
-"Erosional development of streams and their drainage basins; hydrophysical approach to quantitative morphology." Geological society of America bulletin 56.3 (1945): 275-370.
-    APA
-
-    \bibitem{dodds} Dodds, Peter Sheridan, and Daniel H. Rothman. "Unified view of scaling laws for river networks." Physical Review E 59.5 (1999): 4865.
-
-    \bibitem{food_webs} Garlaschelli, Diego, Guido Caldarelli, and Luciano Pietronero. "Universal scaling relations in food webs." Nature 423.6936 (2003): 165-168.
-
-    \bibitem{mat_blog} brain of mat kelcey (brain of mat kelcey). 
-    "Do all first links lead to Philosophy?"
-    \url{http://matpalm.com/blog/2011/08/13/wikipedia-philosophy/}
-
-    \bibitem{Ilmari_first_links} User:Ilmari Karonen/First link (Wikipedia)
-    \url{https://en.wikipedia.org/wiki/User:Ilmari_Karonen/First_link}
-
-    \bibitem{xkcd} xkcd: Extended Mind (xkcd: Extended Mind)
-    \url{http://xkcd.com/903/}
-
-    \bibitem{reddit} Almost every Wikipedia page links to Philosophy if you keep clicking the first link in every article. 
-    \url{https://www.reddit.com/r/InternetIsBeautiful/comments/35asgt/almost_every_wikipedia_page_links_to_philosophy/}
-
-    \bibitem{bias} Wagner, Claudia, et al. "It's a Man's Wikipedia? Assessing Gender Inequality in an Online Encyclopedia." arXiv preprint arXiv:1501.06307 (2015).
-
-    \bibitem{wiki_edits} Wikipedia:Statistics 
-    \url{https://en.wikipedia.org/wiki/Wikipedia:Statistics}
-
-    \bibitem{media_wiki_templates} MediaWiki: "Help: Templates"
-    \url{https://www.mediawiki.org/wiki/Help:Templates}
-\end{thebibliography}
-
 \end{document}

### 2016-02-07 writeup/wikipedia-network.tex
recovered last portion

@@
 
+\subsection{Article Popularity}
+
+Wikipedia released an API to measure article popularity by page views
+starting November 2015. Page views add another dimension to our
+findings by revealing 
+how users view the information on Wikipedia. 
+We measure popularity as the total number 
+of page views in the English edition of Wikipedia in the month of 
+October 2015---the earliest full month for which the data is available. 
+We find the highest ranking 1000 articles (by traversal visits) have an average of
+$70$ thousand page views in October with high variation: the standard deviation 
+is $1.1$ million page views. 
+The number of page views has a skewed distribution with $75\%$ of articles
+reaching fewer than $73$ thousand page views in October.
+The article for "United States" has the most page 
+views of the highest ranking $1000$ articles by traversal visits with 
+$1.4$ million views in October. The next most popular articles are 
+"India," "World War II," and the "United Kingdom" each with $0.7-1.03$ million page views.
+
+The "Philosophy" article despite outranking every article by traversal visits,
+has only $240$ thousand page views in October compared to the $1.4$ million
+page views for "United States".
+While "Philosophy" is not itself the most popular article, "Philosophy" serves as 
+an accumulation point for ideas. Wikipedia users do not login specifically
+to read about "Philosophy," but the authors relate ideas to "Philosophy" 
+by ultimately anchoring the flow of First Links to "Philosophy."
+
+\begin{figure}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/views_visits.png}
+  \caption{
+    \textbf{Article Popularity (by page views) compared to Traversal Visits 
+    for Highest ranking 1000 articles.}
+    We use the total number of page views provided by Wikipedia for the month
+    of October 2015 to compare each article's popularity to traversal visits.
+    While the most popular articles do not necessarily correspond to the articles
+    with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
+  }
+  \label{fig:Views and Visits}
+
+\end{figure}
+
+We also analyze article popularity for the highest ranking 1000 articles by 
+traversal funnels. In October, the average page views per article is 
+$22$ thousand with a standard deviation of $95$ thousand views. The distribution
+is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
+"Halloween" is the most popular article with $2.8$ million views in October,
+likely a result of Halloween falling in the month of October. 
+Other popular articles include "Scientology" ($463$ thousand views), "Clint Eastwood" 
+($341$ thousand views),
+and the "Cold War" ($298$ thousand views) although each has significantly fewer views compared to the views for "Halloween."
+Other standout popular October articles include "24-hour" and "12-hour" clocks likely due
+to Daylight savings in October and "Marriage" likely due to the drop in the number of
+weddings in the months following October
+\cite{weddings}.
+"Philosophy" which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears no where in the top 20
+by traversal visits. "Philosophy" is a relatively popular article 
+among articles influencing the shape of the network, but less popular among 
+highly-ranked articles by accumulation.
+
+\begin{figure}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/funnels_visits.png}
+  \caption{
+    \textbf{Article Popularity (by page views) compared to Traversal Funnels
+    for Highest ranking 1000 articles.}
+  }
+    We use the total number of page views provided by Wikipedia for the month
+    of October 2015 to compare each article's popularity to traversal funnels.
+    In the top left is "Halloween" with a significantly greater 
+    number of page views in the month of October.
+  \label{fig:Views and Funnels}
+\end{figure}
+
+%========================Conclusion====================================
+
+\section{Reflections}
+
+The findings here should only be considered within the limitations of their context.
+We examined only the English version of Wikipedia at a particular moment in time.
+Furthermore, we only studied the First Link in the main body  of each article
+as a means to related one article to another. Finally, Wikipedia, while the largest 
+collection of human knowledge, is rife with the biases of the many contributing editors
+\cite{bias}
+. Nevertheless, the findings do reveal
+generalizable relationships, point to foundational notions, and uncover many curiosities.
+
+Among the curiosities is the multiple appearance of scale-free distributions within the network. 
+The three metrics we developed: path length, traversal visits, and traversal funnels are all marked 
+by scale-free distributions. Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
+articles are responsible for funneling most paths. When measured against the traversal funnels, 
+"Philosophy" emerges as an exceptional article by orders of magnitude. 
+Nevertheless, many other foundational ideas emerged naturally within the First Link Network. 
+Basins around "Community", "State", and "Science" reveal a foundational structure within the network. 
+More curious is the emergence of recently prominent political and economics topics such as "Fossil Fuel" and "Health Care" 
+within the highest ranking funnels. 
+Wikipedia seems to reflect not only timeless foundations, but also the topical (at least within English speaking society).
+
+Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the First Link.
+These findings also form the basis for the creation of a taxonomy where 
+every idea, event, or object sits within a hierarchy of connected notions.
+The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
+Applications could range from an enhanced thesaurus of ideas to psychological insights into how humans form associations.
+Specifically, an ever-evolving reference of related hierarchical concepts can be applied to search engine algorithms 
+or natural language processing.
+
+\acknowledgments
+Thank you to my friend RJ for pointing out the Reddit post 
+\cite{reddit}
+describing how the majority of links lead to "Philosophy," inspiring this research.
+
+\newpage
+
+\section{Appendix}
+
+\subsection{Constructing The First Link Network}
+
+To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
+Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
+eliminating any statistical error due to sampling.
+We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
+The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
+and disambiguations.
+Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
+\cite{wiki_edits},
+our aim is to characterize the dynamics of the First Link Network, not record a particular link between one
+article and another.
+
+Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
+Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
+information.
+While a human could accurately identify the First Link, to map the entire First Link Network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
+
+While some libraries exist for MediaWiki Markup,
+Approaches using existing libraries led to several bugs 
+including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
+as well as compatibility with other libraries used to parse the XML.
+Consequently,  we developed an algorithm for parsing the First Link in the XML version of each article.
+Our parsing algorithm aimed to: 
+1) accurately identify the First Link among other page elements, and 
+2) efficiently do so---that is without needing for several passes through the data.
+To process an article in a single pass, we developed a hierarchical system of flags:
+
+\begin{figure}[tp!]
+  \centering	
+  \includegraphics[width=\columnwidth]{graphics/flags.pdf}  
+  \caption{
+    \textbf{Parsing Algorithm of Wikipedia's XML dump.}
+     The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
+Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
+Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.
+  }
+  \label{fig:parsing algorithm}
+\end{figure}
+
+The algorithm loops in three-character chunks to account for potentially nested elements, 
+shifting by one character steps through the article markup.
+If any markup triggers for a flag are detected, a flag is raised. 
+Once a flag is raised, we stop processing and proceed to the next character
+until the flag's closing markup.
+A First Link is identified only if Flags 1, 2, and 3 are all off.
+In this case, the entire link is retrieved. 
+We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
+as well as common file extensions for 
+images, audio files, and the like 
+\cite{media_wiki_templates}.
+The First Link of an article is then the earliest valid link with unraised flags.
+
+To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
+across 112 cores of the UVM supercomputer cluster
+\cite{vacc}
+We then joined the results to form a hash table containing every Wikipedia article and its corresponding
+First Link. The resulting network map is the basis of our analysis.
+
+\begin{thebibliography}{1}
+
+      \bibitem{editors} Iba, Takashi, et al.
+"Analyzing the creative editing behavior of Wikipedia editors: Through dynamic social network analysis." Procedia-Social and Behavioral Sciences 2.4 (2010): 6441-6456.
+
+      \bibitem{accuracy1} Holman Rector, Lucy. "Comparison of Wikipedia and other encyclopedias for accuracy, breadth, and depth in historical articles." Reference services review 36.1 (2008): 7-22.
+
+      \bibitem{accuracy2} Giles, Jim. "Internet encyclopaedias go head to head." Nature 438.7070 (2005): 900-901.
+
+       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. "An analysis of topical coverage of Wikipedia." Journal of Computer‐Mediated Communication 13.2 (2008): 429-440.
+
+        \bibitem{bias_women} Hill, Benjamin Mako, and Aaron Shaw.
+"The Wikipedia gender gap revisited: characterizing survey response bias with propensity score estimation." PloS one 8.6 (2013): e65782.
+
+      \bibitem{missing_entries} Williams, Jake Ryland, Clark, Eric M., Bagrow, James P., Danforth, Christopher M.
+\& Dodds, Peter Sheridan (2015).
+Identifying missing dictionary entries with frequency-conserving context models.
+Phys.
+Rev.
+E, 92, 042808.
+
+      \bibitem{clustering} Banerjee, Somnath, Krishnan Ramanathan, and Ajay Gupta.
+"Clustering short texts using wikipedia." Proceedings of the 30th annual international ACM SIGIR conference on Research and development in information retrieval.
+ACM, 2007.
+
+      \bibitem{semantic_relatedness} Gabrilovich, Evgeniy, and Shaul Markovitch. "Computing Semantic Relatedness Using Wikipedia-based Explicit Semantic Analysis." IJCAI. Vol. 7. 2007.
+
+      \bibitem{disambiguating} Cucerzan, Silviu. "Large-Scale Named Entity Disambiguation Based on Wikipedia Data." EMNLP-CoNLL. Vol. 7. 2007.
+
+      \bibitem{drugs} Clauson, Kevin A., et al. "Scope, completeness, and accuracy of drug information in Wikipedia." Annals of Pharmacotherapy 42.12 (2008): 1814-1821.
+
+      \bibitem{stats} Wikipedia contributors. "Wikipedia:Statistics." Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 9 Nov. 2015. Web. 13 Nov. 2015.
+
+      \bibitem{pagerank} Page, Lawrence, et al. "The PageRank citation ranking: bringing order to the Web." (1999).
+
+      \bibitem{classifier} Chakrabarti, Soumen, Mukul Joshi, and Vivek Tawde.
+"Enhanced topic distillation using text, markup tags, and hyperlinks." Proceedings of the 24th annual international ACM SIGIR conference on Research and development in information retrieval.
+ACM, 2001.
+
+      \bibitem{relevance} Kamps, Jaap, and Marijn Koolen. "Is Wikipedia link structure different?." Proceedings of the second ACM international conference on Web search and data mining. ACM, 2009.
+
+      \bibitem{wiki_train} Wikipedia contributors. "Train." Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 
+          \url{https://en.wikipedia.org/wiki/Train}
+
+    \bibitem{locke} Bolton, Martha Brandt.
+"The Taxonomy of Ideas in Locke’s {\it Essay} ", The Cambridge Companion to Locke's 'Essay Concerning Human Understanding'.
+Ed.Lex Newman..
+1st ed.
+Cambridge: Cambridge University Press, 2007.
+67-100.
+Cambridge Companions Online.
+Web.
+13 November 2015.
+http://dx.doi.org/10.1017/CCOL0521834333.004.
+
+    \bibitem{aristotle} Studtmann, Paul, "Aristotle's Categories", The Stanford Encyclopedia of Philosophy (Summer 2014 Edition), Edward N.
+Zalta (ed.), \url{http://plato.stanford.edu/archives/sum2014/entries/aristotle-categories/}
+
+    \bibitem{descartes} Smith, Kurt, "Descartes' Theory of Ideas", The Stanford Encyclopedia of Philosophy (Spring 2014 Edition), Edward N.
+Zalta (ed.), \url{http://plato.stanford.edu/archives/spr2014/entries/descartes-ideas/}
+
+    \bibitem{hist_thesaurus} What is the Historical Thesaurus of the OED - Oxford English Dictionary (Oxford English Dictionary)
+    http://public.oed.com/historical-thesaurus-of-the-oed/what-is-the-historical-thesaurus-of-the-oed/
+
+    \bibitem{vacc} The authors acknowledge the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
+    \url{http://www.uvm.edu/vacc}
+
+    \bibitem{wiki_views} Page Views for Wikipedia, Non-mobile site, Normalized (Page Views for Wikipedia, Non-mobile site, Normalized)
+    \url{https://stats.wikimedia.org/EN/TablesPageViewsMonthly.htm}
+
+    \bibitem{weddings} Wedding statistics in the United States (Wedding statistics in the United States)
+    \url{http://www.soundvision.com/article/wedding-statistics-in-the-united-states}
+
+    \bibitem{six_degrees} Elmacioglu, Ergin, and Dongwon Lee. "On six degrees of separation in DBLP-DB and more." ACM SIGMOD Record 34.2 (2005): 33-40.
+
+    \bibitem{wordnet} Princeton University "About WordNet." WordNet. Princeton University. 2010. 
+    \url{http://wordnet.princeton.edu}
+
+    \bibitem{newman} Newman, Mark EJ. "The structure and function of complex networks." SIAM review 45.2 (2003): 167-256.
+
+    \bibitem{disease} Eubank, Stephen, et al. "Modelling disease outbreaks in realistic urban social networks." Nature 429.6988 (2004): 180-184.
+
+    \bibitem{social_nets} Newman, Mark EJ, Duncan J.
+Watts, and Steven H.
+Strogatz.
+"Random graph models of social networks." Proceedings of the National Academy of Sciences 99.suppl 1 (2002): 2566-2572.
+
+    \bibitem{geo_basins} Horton, Robert E.
+"Erosional development of streams and their drainage basins; hydrophysical approach to quantitative morphology." Geological society of America bulletin 56.3 (1945): 275-370.
+    APA
+
+    \bibitem{dodds} Dodds, Peter Sheridan, and Daniel H. Rothman. "Unified view of scaling laws for river networks." Physical Review E 59.5 (1999): 4865.
+
+    \bibitem{food_webs} Garlaschelli, Diego, Guido Caldarelli, and Luciano Pietronero. "Universal scaling relations in food webs." Nature 423.6936 (2003): 165-168.
+
+    \bibitem{mat_blog} brain of mat kelcey (brain of mat kelcey). 
+    "Do all first links lead to Philosophy?"
+    \url{http://matpalm.com/blog/2011/08/13/wikipedia-philosophy/}
+
+    \bibitem{Ilmari_first_links} User:Ilmari Karonen/First link (Wikipedia)
+    \url{https://en.wikipedia.org/wiki/User:Ilmari_Karonen/First_link}
+
+    \bibitem{xkcd} xkcd: Extended Mind (xkcd: Extended Mind)
+    \url{http://xkcd.com/903/}
+
+    \bibitem{reddit} Almost every Wikipedia page links to Philosophy if you keep clicking the first link in every article. 
+    \url{https://www.reddit.com/r/InternetIsBeautiful/comments/35asgt/almost_every_wikipedia_page_links_to_philosophy/}
+
+    \bibitem{bias} Wagner, Claudia, et al. "It's a Man's Wikipedia? Assessing Gender Inequality in an Online Encyclopedia." arXiv preprint arXiv:1501.06307 (2015).
+
+    \bibitem{wiki_edits} Wikipedia:Statistics 
+    \url{https://en.wikipedia.org/wiki/Wikipedia:Statistics}
+
+    \bibitem{media_wiki_templates} MediaWiki: "Help: Templates"
+    \url{https://www.mediawiki.org/wiki/Help:Templates}
+\end{thebibliography}
+
 \end{document}

### 2016-02-07 writeup/wikipedia-network.tex
sanity restored, edits up to article popularity

@@
     * figure out what gamma size freq means
+
+\section{Introduction}
+
+Wikipedia is a towering achievement of the modern era. 
+At no point in history has a larger or more meticulously indexed collection of human knowledge 
+existed.
+Wikipedia contains 37 million articles in 283 languages, 
+with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
+\cite{drugs, stats}.
+All this information does not lie dormant. 
+Wikipedia is the sixth most visited
+site in the world, surpassing 18 billion page views and 26 million edits in a single month
+\cite{wiki_edits, wiki_views}.
+
+Wikipedia has naturally become the object of many studies. 
+Researchers have examined the cultural dynamics among editors
+\cite{editors},
+the accuracy of the content relative to traditional encyclopedias
+\cite{accuracy1, accuracy2},
+the topics covered 
+\cite{coverage},
+and bias against portions of the population
+\cite{bias_women}.
+Wikipedia's content has also proven a powerful tool. 
+Researchers have used Wikipedia to identify missing dictionary entries
+\cite{missing_entries},
+cluster short text
+\cite{clustering},
+compute semantic relatedness
+\cite{semantic_relatedness},
+and disambiguate meaning \cite{disambiguating}.
+
+While these many studies have dissected and fruitfully applied Wikipedia's content,
+few have examined the connections among the many articles.
+A hyperlink from one Wikipedia article to another
+indicates a relationship between the two articles
+\cite{relevance}.
+Naturally, the notion that hyperlinks convey information about the content of a
+page has proved enormously successful in multiple domains from search engine algorithms 
+such as PageRank 
+\cite{pagerank} 
+to topic classification
+\cite{classifier}.
+We treat a hyperlink as a mechanism connecting two topics.
+
+The authors of a Wikipedia article choose where and whether to include a 
+reference to another Wikipedia article
+in the HTML markup.
+For example, the authors of the 
+\href{https://en.wikipedia.org/wiki/Train}{"Train"}
+article chose
+"Rail Transport," "Cargo," and "Passengers" 
+as pertinent articles to reference in describing trains
+\cite{wiki_train}
+.
+Here we focus on the first hyperlink to another Wikipedia article, which 
+marks the earliest moment in the body text where the authors
+chose to connect the information discussed to that of another article.
+By following the first hyperlink to another article in the English edition of
+Wikipedia as of November, 2014, we connect the ideas in one article to those in another, ultimately forming a directed network: 
+{\it Wikipedia's First Link Network} (FLN).
+
+\begin{figure*}[tp!]
+  \centering	
+  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
+  \caption{
+    \textbf{First Link Path For "Train".}
+    We follow the First Link to another Wikipedia article
+    in the main body of the article---the area
+    inside the green rectangle, which excludes 
+    side bar elements, the navigation bar and title. 
+    In this example, the First Link to another Wikipedia article is "Rail Transport." 
+    We can again select the First Link on the "Rail Transport"
+    article, repeating the process to 
+    form a path of First Links.
+  }
+  \label{fig:Train First Links}
+\end{figure*}
+
+What information could the First Links possibly reveal?
+Inspired by the claim that the majority of First Links lead to the 
+"Philosophy" article---popularized by an 
+\href{https://xkcd.com/903/}{xkcd}
+comic and subsequently
+discussed in blog posts 
+\cite{mat_blog, Ilmari_first_links, xkcd}
+---we holistically study 
+Wikipedia's FLN as a map relating areas of human knowledge. 
+The map is a hierarchical taxonomy where, for example, "Train" links to a parent node, 
+"Rail Transport," while many child nodes, "Steel" and "Horsepower," link to 
+"Train." Unlike previous taxonomies
+created by individuals
+\cite{locke, descartes, aristotle}
+or a select group of individuals 
+\cite{hist_thesaurus}, 
+the organization of ideas in the FLN 
+does not emerge from a centralized effort. 
+Rather each link in the FLN emerges from the choice of each author and ensuing edits.
+Collectively, the FLN is a wealth of relations among inventions, places,
+figures, objects, and events across space and time.
+
+Our goal is to study the structure of the FLN for insight into how the information on Wikipedia is organized and related.
+We develop metrics to capture 
+the dominant features of the FLN's structure.
+We measure dynamics of the FLN as a flow, quantifying 
+aspects of the FLN from the accumulation of First Links around articles 
+to the influence an article exerts in shaping the FLN.
+Together with cycles, in-degree, depth, and the content of the articles, 
+we build our analysis of the relations among the ideas in Wikipedia.
 

### 2016-02-07 writeup/wikipedia-network.tex
first pass syntax edits complete

@@
 TODO: 
-* fix end of second paragraph in Results Section
+\begin{itemize}
+    \item fix end of second paragraph in Results Section\\
     * power-law biz and total peter is asking for
-* fix exponent reporting (with alpha) 
+\item fix exponent reporting (with alpha) \\
     * figure out what gamma size freq means
+\end{itemize}
 
@@
 For example, the authors of the 
-\href{https://en.wikipedia.org/wiki/Train}{"Train"}
+\href{https://en.wikipedia.org/wiki/Train}{``Train''}
 article chose
-"Rail Transport," "Cargo," and "Passengers" 
+``Rail Transport,'' ``Cargo,'' and ``Passengers'' 
 as pertinent articles to reference in describing trains
@@
   \caption{
-    \textbf{First Link Path For "Train".}
+    \textbf{First Link Path For ``Train.''}
     We follow the First Link to another Wikipedia article
@@
     side bar elements, the navigation bar and title. 
-    In this example, the First Link to another Wikipedia article is "Rail Transport." 
-    We can again select the First Link on the "Rail Transport"
+    In this example, the First Link to another Wikipedia article is ``Rail Transport.'' 
+    We can again select the First Link on the ``Rail Transport''
     article, repeating the process to 
@@
 Inspired by the claim that the majority of First Links lead to the 
-"Philosophy" article---popularized by an 
+``Philosophy'' article---popularized by an 
 \href{https://xkcd.com/903/}{xkcd}
@@
 Wikipedia's FLN as a map relating areas of human knowledge. 
-The map is a hierarchical taxonomy where, for example, "Train" links to a parent node, 
+The map is a hierarchical taxonomy where, for example, ``Train'' links to a parent node, 
 "Rail Transport," while many child nodes, "Steel" and "Horsepower," link to 
@@
 
-The "Philosophy" article despite outranking every article by traversal visits,
+The "Philosophy" article, despite outranking every article by traversal visits,
 has only $240$ thousand page views in October compared to the $1.4$ million
 page views for "United States".
-While "Philosophy" is not itself the most popular article, "Philosophy" serves as 
-an accumulation point for ideas. Wikipedia users do not login specifically
-to read about "Philosophy," but the authors relate ideas to "Philosophy" 
-by ultimately anchoring the flow of First Links to "Philosophy."
 
@@
 
-\section{Reflections}
+\section{Concluding Remarks}
 
 The findings here should only be considered within the limitations of their context.
-We examined only the English version of Wikipedia at a particular moment in time.
-Furthermore, we only studied the First Link in the main body  of each article
-as a means to related one article to another. Finally, Wikipedia, while the largest 
+We examined only the English editiion of Wikipedia at a particular moment in time, 
+considering only the first link in in each article as a means to related one article to another.
+Finally, Wikipedia, while the largest 
 collection of human knowledge, is rife with the biases of the many contributing editors
@@
 . Nevertheless, the findings do reveal
-generalizable relationships, point to foundational notions, and uncover many curiosities.
+generalizable relationships and point to foundational notions.
 
-Among the curiosities is the multiple appearance of scale-free distributions within the network. 
-The three metrics we developed: path length, traversal visits, and traversal funnels are all marked 
-by scale-free distributions. Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
+Among our observations is the multiple appearance of scale-free distributions within the network. 
+Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
 articles are responsible for funneling most paths. When measured against the traversal funnels, 
@@
 The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
-Applications could range from an enhanced thesaurus of ideas to psychological insights into how humans form associations.
-Specifically, an ever-evolving reference of related hierarchical concepts can be applied to search engine algorithms 
+Applications could range from an enhanced network of ideas to psychological insights into how humans form associations.
+Specifically, an ever-evolving reference of related hierarchical concepts could be applied to search engine algorithms 
 or natural language processing.
@@
 \acknowledgments
-Thank you to my friend RJ for pointing out the Reddit post 
+MI is grateful to RJ for pointing out the Reddit post 
 \cite{reddit}

### 2016-02-07 writeup/wikipedia-network.tex
first pass syntax edits complete + quotes fixed

@@
 The map is a hierarchical taxonomy where, for example, ``Train'' links to a parent node, 
-"Rail Transport," while many child nodes, "Steel" and "Horsepower," link to 
-"Train." Unlike previous taxonomies
+``Rail Transport,'' while many child nodes, ``Steel'' and ``Horsepower,'' link to 
+``Train.'' Unlike previous taxonomies
 created by individuals
@@
 reaching fewer than $73$ thousand page views in October.
-The article for "United States" has the most page 
+The article for ``United States'' has the most page 
 views of the highest ranking $1000$ articles by traversal visits with 
 $1.4$ million views in October. The next most popular articles are 
-"India," "World War II," and the "United Kingdom" each with $0.7-1.03$ million page views.
+``India,'' ``World War II,'' and the ``United Kingdom'' each with $0.7-1.03$ million page views.
 
-The "Philosophy" article, despite outranking every article by traversal visits,
+The ``Philosophy'' article, despite outranking every article by traversal visits,
 has only $240$ thousand page views in October compared to the $1.4$ million
-page views for "United States".
+page views for ``United States''.
 
@@
 is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
-"Halloween" is the most popular article with $2.8$ million views in October,
+``Halloween'' is the most popular article with $2.8$ million views in October,
 likely a result of Halloween falling in the month of October. 
-Other popular articles include "Scientology" ($463$ thousand views), "Clint Eastwood" 
+Other popular articles include ``Scientology'' ($463$ thousand views), ``Clint Eastwood'' 
 ($341$ thousand views),
-and the "Cold War" ($298$ thousand views) although each has significantly fewer views compared to the views for "Halloween."
-Other standout popular October articles include "24-hour" and "12-hour" clocks likely due
-to Daylight savings in October and "Marriage" likely due to the drop in the number of
+and the ``Cold War'' ($298$ thousand views) although each has significantly fewer views compared to the views for ``Halloween''.
+Other standout popular October articles include ``24-hour'' and ``12-hour'' clocks likely due
+to Daylight savings in October and ``Marriage'' likely due to the drop in the number of
 weddings in the months following October
 \cite{weddings}.
-"Philosophy" which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears no where in the top 20
-by traversal visits. "Philosophy" is a relatively popular article 
+``Philosophy'' which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears no where in the top 20
+by traversal visits. ``Philosophy'' is a relatively popular article 
 among articles influencing the shape of the network, but less popular among 
@@
     of October 2015 to compare each article's popularity to traversal funnels.
-    In the top left is "Halloween" with a significantly greater 
+    In the top left is ``Halloween'' with a significantly greater 
     number of page views in the month of October.
@@
 articles are responsible for funneling most paths. When measured against the traversal funnels, 
-"Philosophy" emerges as an exceptional article by orders of magnitude. 
+``Philosophy'' emerges as an exceptional article by orders of magnitude. 
 Nevertheless, many other foundational ideas emerged naturally within the First Link Network. 
-Basins around "Community", "State", and "Science" reveal a foundational structure within the network. 
-More curious is the emergence of recently prominent political and economics topics such as "Fossil Fuel" and "Health Care" 
+Basins around ``Community'', ``State'', and ``Science'' reveal a foundational structure within the network. 
+More curious is the emergence of recently prominent political and economics topics such as ``Fossil Fuel'' and ``Health Care'' 
 within the highest ranking funnels. 
@@
 \cite{reddit}
-describing how the majority of links lead to "Philosophy," inspiring this research.
+describing how the majority of links lead to ``Philosophy,'' inspiring this research.
 
@@
       \bibitem{editors} Iba, Takashi, et al.
-"Analyzing the creative editing behavior of Wikipedia editors: Through dynamic social network analysis." Procedia-Social and Behavioral Sciences 2.4 (2010): 6441-6456.
+``Analyzing the creative editing behavior of Wikipedia editors: Through dynamic social network analysis.'' Procedia-Social and Behavioral Sciences 2.4 (2010): 6441-6456.
 
-      \bibitem{accuracy1} Holman Rector, Lucy. "Comparison of Wikipedia and other encyclopedias for accuracy, breadth, and depth in historical articles." Reference services review 36.1 (2008): 7-22.
+      \bibitem{accuracy1} Holman Rector, Lucy. ``Comparison of Wikipedia and other encyclopedias for accuracy, breadth, and depth in historical articles.'' Reference services review 36.1 (2008): 7-22.
 
-      \bibitem{accuracy2} Giles, Jim. "Internet encyclopaedias go head to head." Nature 438.7070 (2005): 900-901.
+      \bibitem{accuracy2} Giles, Jim. ``Internet encyclopaedias go head to head.'' Nature 438.7070 (2005): 900-901.
 
-       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. "An analysis of topical coverage of Wikipedia." Journal of Computer‐Mediated Communication 13.2 (2008): 429-440.
+       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. ``An analysis of topical coverage of Wikipedia.'' Journal of Computer‐Mediated Communication 13.2 (2008): 429-440.
 
         \bibitem{bias_women} Hill, Benjamin Mako, and Aaron Shaw.
-"The Wikipedia gender gap revisited: characterizing survey response bias with propensity score estimation." PloS one 8.6 (2013): e65782.
+``The Wikipedia gender gap revisited: characterizing survey response bias with propensity score estimation.'' PloS one 8.6 (2013): e65782.
 
@@
       \bibitem{clustering} Banerjee, Somnath, Krishnan Ramanathan, and Ajay Gupta.
-"Clustering short texts using wikipedia." Proceedings of the 30th annual international ACM SIGIR conference on Research and development in information retrieval.
+``Clustering short texts using wikipedia.'' Proceedings of the 30th annual international ACM SIGIR conference on Research and development in information retrieval.
 ACM, 2007.
 
-      \bibitem{semantic_relatedness} Gabrilovich, Evgeniy, and Shaul Markovitch. "Computing Semantic Relatedness Using Wikipedia-based Explicit Semantic Analysis." IJCAI. Vol. 7. 2007.
+      \bibitem{semantic_relatedness} Gabrilovich, Evgeniy, and Shaul Markovitch. ``Computing Semantic Relatedness Using Wikipedia-based Explicit Semantic Analysis.'' IJCAI. Vol. 7. 2007.
 
-      \bibitem{disambiguating} Cucerzan, Silviu. "Large-Scale Named Entity Disambiguation Based on Wikipedia Data." EMNLP-CoNLL. Vol. 7. 2007.
+      \bibitem{disambiguating} Cucerzan, Silviu. ``Large-Scale Named Entity Disambiguation Based on Wikipedia Data.'' EMNLP-CoNLL. Vol. 7. 2007.
 
-      \bibitem{drugs} Clauson, Kevin A., et al. "Scope, completeness, and accuracy of drug information in Wikipedia." Annals of Pharmacotherapy 42.12 (2008): 1814-1821.
+      \bibitem{drugs} Clauson, Kevin A., et al. ``Scope, completeness, and accuracy of drug information in Wikipedia.'' Annals of Pharmacotherapy 42.12 (2008): 1814-1821.
 
-      \bibitem{stats} Wikipedia contributors. "Wikipedia:Statistics." Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 9 Nov. 2015. Web. 13 Nov. 2015.
+      \bibitem{stats} Wikipedia contributors. ``Wikipedia:Statistics.'' Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 9 Nov. 2015. Web. 13 Nov. 2015.
 
-      \bibitem{pagerank} Page, Lawrence, et al. "The PageRank citation ranking: bringing order to the Web." (1999).
+      \bibitem{pagerank} Page, Lawrence, et al. ``The PageRank citation ranking: bringing order to the Web.'' (1999).
 
       \bibitem{classifier} Chakrabarti, Soumen, Mukul Joshi, and Vivek Tawde.
-"Enhanced topic distillation using text, markup tags, and hyperlinks." Proceedings of the 24th annual international ACM SIGIR conference on Research and development in information retrieval.
+``Enhanced topic distillation using text, markup tags, and hyperlinks.'' Proceedings of the 24th annual international ACM SIGIR conference on Research and development in information retrieval.
 ACM, 2001.
 
-      \bibitem{relevance} Kamps, Jaap, and Marijn Koolen. "Is Wikipedia link structure different?." Proceedings of the second ACM international conference on Web search and data mining. ACM, 2009.
+      \bibitem{relevance} Kamps, Jaap, and Marijn Koolen. ``Is Wikipedia link structure different?.'' Proceedings of the second ACM international conference on Web search and data mining. ACM, 2009.
 
-      \bibitem{wiki_train} Wikipedia contributors. "Train." Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 
+      \bibitem{wiki_train} Wikipedia contributors. ``Train.'' Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 
           \url{https://en.wikipedia.org/wiki/Train}
@@
     \bibitem{locke} Bolton, Martha Brandt.
-"The Taxonomy of Ideas in Locke’s {\it Essay} ", The Cambridge Companion to Locke's 'Essay Concerning Human Understanding'.
+``The Taxonomy of Ideas in Locke’s {\it Essay},'' The Cambridge Companion to Locke's 'Essay Concerning Human Understanding'.
 Ed.Lex Newman..
@@
 
-    \bibitem{aristotle} Studtmann, Paul, "Aristotle's Categories", The Stanford Encyclopedia of Philosophy (Summer 2014 Edition), Edward N.
+    \bibitem{aristotle} Studtmann, Paul, ``Aristotle's Categories,'' The Stanford Encyclopedia of Philosophy (Summer 2014 Edition), Edward N.
 Zalta (ed.), \url{http://plato.stanford.edu/archives/sum2014/entries/aristotle-categories/}
 
-    \bibitem{descartes} Smith, Kurt, "Descartes' Theory of Ideas", The Stanford Encyclopedia of Philosophy (Spring 2014 Edition), Edward N.
+    \bibitem{descartes} Smith, Kurt, ``Descartes' Theory of Ideas,'' The Stanford Encyclopedia of Philosophy (Spring 2014 Edition), Edward N.
 Zalta (ed.), \url{http://plato.stanford.edu/archives/spr2014/entries/descartes-ideas/}
@@
 
-    \bibitem{six_degrees} Elmacioglu, Ergin, and Dongwon Lee. "On six degrees of separation in DBLP-DB and more." ACM SIGMOD Record 34.2 (2005): 33-40.
+    \bibitem{six_degrees} Elmacioglu, Ergin, and Dongwon Lee. ``On six degrees of separation in DBLP-DB and more.'' ACM SIGMOD Record 34.2 (2005): 33-40.
 
-    \bibitem{wordnet} Princeton University "About WordNet." WordNet. Princeton University. 2010. 
+    \bibitem{wordnet} Princeton University ``About WordNet.'' WordNet. Princeton University. 2010. 
     \url{http://wordnet.princeton.edu}
 
-    \bibitem{newman} Newman, Mark EJ. "The structure and function of complex networks." SIAM review 45.2 (2003): 167-256.
+    \bibitem{newman} Newman, Mark EJ. ``The structure and function of complex networks.'' SIAM review 45.2 (2003): 167-256.
 
-    \bibitem{disease} Eubank, Stephen, et al. "Modelling disease outbreaks in realistic urban social networks." Nature 429.6988 (2004): 180-184.
+    \bibitem{disease} Eubank, Stephen, et al. ``Modelling disease outbreaks in realistic urban social networks.'' Nature 429.6988 (2004): 180-184.
 
@@
 Strogatz.
-"Random graph models of social networks." Proceedings of the National Academy of Sciences 99.suppl 1 (2002): 2566-2572.
+``Random graph models of social networks.'' Proceedings of the National Academy of Sciences 99.suppl 1 (2002): 2566-2572.
 
     \bibitem{geo_basins} Horton, Robert E.
-"Erosional development of streams and their drainage basins; hydrophysical approach to quantitative morphology." Geological society of America bulletin 56.3 (1945): 275-370.
+``Erosional development of streams and their drainage basins; hydrophysical approach to quantitative morphology.'' Geological society of America bulletin 56.3 (1945): 275-370.
     APA
 
-    \bibitem{dodds} Dodds, Peter Sheridan, and Daniel H. Rothman. "Unified view of scaling laws for river networks." Physical Review E 59.5 (1999): 4865.
+    \bibitem{dodds} Dodds, Peter Sheridan, and Daniel H. Rothman. ``Unified view of scaling laws for river networks.'' Physical Review E 59.5 (1999): 4865.
 
-    \bibitem{food_webs} Garlaschelli, Diego, Guido Caldarelli, and Luciano Pietronero. "Universal scaling relations in food webs." Nature 423.6936 (2003): 165-168.
+    \bibitem{food_webs} Garlaschelli, Diego, Guido Caldarelli, and Luciano Pietronero. ``Universal scaling relations in food webs.'' Nature 423.6936 (2003): 165-168.
 
     \bibitem{mat_blog} brain of mat kelcey (brain of mat kelcey). 
-    "Do all first links lead to Philosophy?"
+    ``Do all first links lead to Philosophy?''
     \url{http://matpalm.com/blog/2011/08/13/wikipedia-philosophy/}
@@
 
-    \bibitem{bias} Wagner, Claudia, et al. "It's a Man's Wikipedia? Assessing Gender Inequality in an Online Encyclopedia." arXiv preprint arXiv:1501.06307 (2015).
+    \bibitem{bias} Wagner, Claudia, et al. ``It's a Man's Wikipedia? Assessing Gender Inequality in an Online Encyclopedia.'' arXiv preprint arXiv:1501.06307 (2015).
 
@@
 
-    \bibitem{media_wiki_templates} MediaWiki: "Help: Templates"
+    \bibitem{media_wiki_templates} MediaWiki: ``Help: Templates''
     \url{https://www.mediawiki.org/wiki/Help:Templates}

### 2016-02-07 writeup/wikipedia-network.tex
good syntax, after second pass

@@
 \begin{itemize}
+    \item add the what to scale-free distribution in the abstract
     \item fix end of second paragraph in Results Section\\
@@
 While these many studies have dissected and fruitfully applied Wikipedia's content,
-few have examined the connections among the many articles.
+few have examined the connections among the many articles
+\cite{links_that_speak}.
 A hyperlink from one Wikipedia article to another
-indicates a relationship between the two articles
+naturally indicates a relationship between the two articles
 \cite{relevance}.
-Naturally, the notion that hyperlinks convey information about the content of a
+The notion that hyperlinks convey information about the content of a
 page has proved enormously successful in multiple domains from search engine algorithms 
@@
 contains ``Goods,'' ``Economics,'' ``Social Science,''
-ultimately leading to ``Philosophy''.
+leading ultimately to ``Philosophy''.
 
@@
 Starting at each article, we construct a path through the FLN, 
-ultimately mapping a flow of connections among articles.
+to map the flow of connections among articles.
 The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
@@
 Previous studies have used flow to characterize the structure of river networks
-\cite{dodds} and describe the organization of food systems as transportation networks
+\cite{dodds} and describe the organization of food systems through transportation networks
 \cite{food_webs}.
@@
 into a particular cycle. In our sample network 
-(see figure~\ref{fig:Traversal Funnels}) article C 
+(see Fig.~\ref{fig:Traversal Funnels}) article C 
 directs the flow of links towards the 3-cycle, while articles A and B are 
@@
 sports associations such as ``American Football,'' ``Association Football''; 
-and developed nations such as ``France,'' ``Japan'', ``Russia,'' ``Australia,'' and 
+and developed nations such as ``France,'' ``Japan,'' ``Russia,'' ``Australia,'' and 
 the ``Netherlands.'' These high ranking articles are useful abstractions: nations
@@
 result of many articles directly referencing ``Philosophy.'' 
-Instead, flows towards Philosophy as the 
-ultimate anchor when generalizing from specific articles to broader notions.
+Instead, first links flow towards ``Philosophy'' as the 
+ultimate anchor, by generalizing from specific articles to broader notions.
 
@@
 
-The FLN's in-degree exhibits a decaying power-law distribution where a few articles 
+The FLN's in-degree exhibits a decaying power law distribution where a few articles 
 receive most direct first references, while most articles receive few or none.
 The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5 links.
-Only 4,826 articles have more than 100 direct first links and $75\%$ of articles
+Less than $1\%$ of articles have more than 100 direct first links and $75\%$ of articles
 have fewer than 9. 
@@
 We first seek to describe the depth of the FLN: how many links does a 
-connection of among articles span? 
+connection among articles span? 
 We find the longest path length is 365,
@@
 while a few temporally organized paths exceed 50 links 
-(see figure~\ref{fig:Path Length Distribution}).
+(see Fig.~\ref{fig:Path Length Distribution}).
 
@@
 
-In Fig., we show a log-log plot for the entire dataset:traversal visits against rank. 
+In Fig.~\ref{fig:Distribution of Visits}, we show a log-log plot for the entire dataset:traversal visits against rank. 
 We observe a strong linear fit, with a Pearson's Correlation Coefficient of -.93 and 
@@
 Of the 11 million articles, 84,000 are 2-cycles. 
-The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected ideas:
+The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
 ``Health Care'' and ``Medical Case Management,'' ``Broadcasting House,'' and ``BBC,'' ``Secondary Education'' and ``Secondary School'' 
@@
     to each other. The area of the circles in the figure indicates the number of traversal visits. We find 2-cycles often capture synonyms or articles representing nearly the 
-    same concepts as opposed to distinct ideas.
+    same concepts as opposed to distinct concepts.
   }
@@
 
-Outside of the highest ranking 2-cycles, the typical 2-cycle signals a connection between distinct, yet very closely related ideas. 
+Outside of the highest ranking 2-cycles, the typical 2-cycle signals a connection between distinct, yet very closely related concepts. 
 We also observe link patterns such as inventor to product (``Voere'' to ``VEC-91''), event to organizer (``Poetry Bus Tour'' to ``Weave Books''), and book to author (``Anatomy of Britain'' to ``Anthony Sampson'').
@@
 and ``Tree of life''; ``Cinema of India,'' ``Indian Cinema,'' and ``Telugu Cinema''
-(see figure~\ref{fig:3-cycles}).
+(see Fig.~\ref{fig:3-cycles}).
 Once we extend our cycle size beyond a length of 6 however, 
@@
 The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day.
-Other lengthy cycles span 60--75 articles including collections of articles on national histories such as ``Japanese Eras'' 
+Other lengthy cycles span $60-75$ articles including collections of articles on national histories such as ``Japanese Eras'' 
 or judicial bodies such as the ``Legislative Assembly of Ontario.''
@@
 with ``Democracy,'' another beginning with ``Constitution,'' another at 
-``Dictatorship,'' and so on. The ideas build from specific means of organizing
+``Dictatorship,'' and so on. The articles build from specific means of organizing
 a community (or society) and then build up to ``Community.'' 
@@
 $7.37$ million paths
-(see figure~\ref{fig:Funnels}).
+(see Fig.~\ref{fig:Funnels}).
 \begin{figure*}[tp!]
@@
 ``Presentation'' with only $30$ thousand paths. Similarly abstract 
-ideas also rank highly such as ``Tree of life'' (30 thousand), 
+concepts also rank highly such as ``Tree of life'' (30 thousand), 
 ``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).
 
-Many high-ranking articles are remarkably topical, culturally and politically important ideas.
+Many high-ranking articles are remarkably topical, culturally and politically important concepts.
 For example, ``Health Care'', a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August, 2009 and February, 2010.
@@
 ``Philosophy'' emerges as an exceptional article by orders of magnitude. 
-Nevertheless, many other foundational ideas emerged naturally within the First Link Network. 
+Nevertheless, many other foundational concepts emerged naturally within the First Link Network. 
 Basins around ``Community'', ``State'', and ``Science'' reveal a foundational structure within the network. 
@@
     \url{https://www.mediawiki.org/wiki/Help:Templates}
+
+    \bibitem{links_that_speak} Ronen, Shahar, Bruno Gonçalves, Kevin Z.
+Hu, Alessandro Vespignani, Steven Pinker, and César A.
+Hidalgo.
+``'Links That Speak: The Global Language Network and Its Association with Global Fame.'' Proceedings of the National Academy of Sciences Proc Natl Acad Sci USA 111.52 (2014).
+
 \end{thebibliography}

### 2016-02-10 writeup/wikipedia-network.tex
updated figures + caption alignment

@@
     \item fix end of second paragraph in Results Section\\
-    * power-law biz and total peter is asking for
+    * power-law biz: rank exponent (slope of log-log), denoted $\alpha$ \\
+    * size exponent corresponds to the CDF (how many people have an income greater than x)
+    = $\gamma = \frac{1}{\alpha} + 1$.
 \item fix exponent reporting (with alpha) \\
@@
 \begin{figure*}[tp!]
-  \centering	
   \includegraphics[width=\textwidth]{graphics/Train.pdf}  
@@
 \begin{figure*}[tp!]
-  \centering	
   \includegraphics[width=\textwidth]{graphics/traversal_visit_algo_figure.pdf}  
@@
 \begin{figure}[tp!]
-  \centering	
   \includegraphics[width=\columnwidth]{graphics/traversal_funnel_algo_figure.pdf}
@@
 \begin{figure}[tp!]
-  \centering	
   \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
@@
     \textbf{Highest Ranking Articles by in-degree.}
-  }
-  We rank each article by the number of direct first links to the article (in-degree). The highest-ranking articles tend to represent abstraction for concepts
-  comprised of various facets.
+We rank each article by the number of direct first links to the article (in-degree). The highest-ranking articles tend to represent abstraction for concepts comprised of various facets.}
   \label{fig:ndegree list}
@@
 \begin{figure}[tp!]
-  \centering	
   \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
@@
     \textbf{FLN Degree Distribution.}
-  }
 We construct a log-log plot and fit our results with a linear model. The result is a 
 an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
-  The distribution appears to be scale free.
+The distribution appears to be scale free.}
   \label{fig:degree distribution}
@@
 \begin{figure}[tp!]
-  \centering	
   \includegraphics[width=\columnwidth]{graphics/path_lengths_boxplot.png}
@@
     \textbf{Path Length Distribution.}
-  }
 The median network depth measured by the number of first links in a path
@@
 the third quartile. More than $75\%$ of articles have path lengths between 
-  $0$ and $50$ links.
-
+$0$ and $50$ links.}
   \label{fig:Path Length Distribution}
@@
 \begin{figure}[tp!]
-  \centering	
   \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
@@
     \textbf{Distribution of Traversal Visits.}
-  }
 We fit the distribution to a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal visits. 
@@
 and a power law exponent of $-0.64$. The horizontal flattening around the highest
-  ranking articles is a result of the cyclic structure (see discussion on cycles).
+ranking articles is a result of the cyclic structure (see discussion on cycles).}
   \label{fig:Distribution of Visits}
@@
 \begin{figure}[tp!]
-  \centering	
   \includegraphics[width=\columnwidth]{graphics/articles_ranked.png}
@@
     \textbf{Highest ranking articles by number of traversal visits.}
-  }
 We compute the number of traversal visits for each article in the FLN (see 
@@
 by the accumulation of first links. Articles with a greater number of traversal visits
-  mark greater points of first link accumulation. The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.
+mark greater points of first link accumulation. The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.}
   \label{fig:highest visits}
@@
 \begin{figure*}[tp!]
-  \centering	
   \includegraphics[width=\textwidth]{graphics/2_cycles.pdf}
@@
 to each other. The area of the circles in the figure indicates the number of traversal visits. We find 2-cycles often capture synonyms or articles representing nearly the 
-    same concepts as opposed to distinct concepts.
-  }
+same concepts as opposed to distinct concepts.}
   \label{fig:2-cycles}
@@
 \begin{figure*}[tp!]
-  \centering	
   \includegraphics[width=\textwidth]{graphics/3_cycles.pdf}
@@
 the most referenced group of three linked articles. 
-    The area of the circles in the figure indicates the number of traversal visits. 
-  }
+The area of the circles in the figure indicates the number of traversal visits.}
   \label{fig:3-cycles}
@@
 \begin{figure*}[tp!]
-  \centering	
   \includegraphics[width=\textwidth]{graphics/funnels.pdf}
@@
 of the influence, with other abstract notions and topical concepts ranking
-    next.
-  }
+next.}
   \label{fig:Funnels}
@@
 \begin{figure}[tp!]
-  \centering	
   \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
@@
 horizontal regime corresponds to the more than $99\%$ of articles
-  which hold zero traversal funnels.
-  }
+which hold zero traversal funnels.}
   \label{fig:Funnels Distribution}
@@
 \begin{figure}[tp!]
-  \centering	
   \includegraphics[width=\columnwidth]{graphics/views_visits.png}
@@
 While the most popular articles do not necessarily correspond to the articles
-    with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-  }
+with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.}
   \label{fig:Views and Visits}
@@
 \begin{figure}[tp!]
-  \centering	
   \includegraphics[width=\columnwidth]{graphics/funnels_visits.png}
@@
     for Highest ranking 1000 articles.}
-  }
 We use the total number of page views provided by Wikipedia for the month
@@
 In the top left is ``Halloween'' with a significantly greater 
-    number of page views in the month of October.
+number of page views in the month of October.}
   \label{fig:Views and Funnels}
@@
 \begin{figure}[tp!]
-  \centering	
   \includegraphics[width=\columnwidth]{graphics/flags.pdf}  
@@
 Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
-Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.
-  }
+Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.}
   \label{fig:parsing algorithm}

### 2016-02-19 writeup/wikipedia-network.tex
traversal funnels figure changed to log horz bar chart

@@
     \textbf{Highest Ranking Articles by in-degree.}
-We rank each article by the number of direct first links to the article (in-degree). The highest-ranking articles tend to represent abstraction for concepts comprised of various facets.}
-  \label{fig:ndegree list}
+We rank each article by the number of direct first links to the article (in-degree). The highest-ranking articles tend to represent geographical and biological abstractions.}
+  \label{fig:indegree list}
 \end{figure}
@@
 Wikipedia article 
-(Fig.~\ref{fig:ndegree list}). 
+(Fig.~\ref{fig:indegree list}). 
 Other high-ranking articles
@@
     \textbf{FLN Degree Distribution.}
-We construct a log-log plot and fit our results with a linear model. The result is a 
-an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
-The distribution appears to be scale free.}
+We fit the in-degree distribution to a linear $\log_{10}-\log_{10}$ model. The result is a 
+an excellent fit with a Pearson's Correlation Coefficient of $-0.98$, yielding a 
+power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$. 
+}
   \label{fig:degree distribution}
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/path_lengths_boxplot.png}
+  \includegraphics[width=\columnwidth]{graphics/path_lengths_dist.png}
   \caption{
     \textbf{Path Length Distribution.}
-The median network depth measured by the number of first links in a path
-is 29. The 365 cycle for ``Orthodox'' liturgics is the outlier to the right, while
-other historical articles about ``Scotland,'' ``UK'' and so on are slightly outside
-the third quartile. More than $75\%$ of articles have path lengths between 
+The network depth appears to have a bimodal distribution with a median of 29 first links in a path.
+The 365 cycle for ``Orthodox'' liturgics is the outlier to the right, with historical articles 
+describing nations such as ``Scotland'' and ``UK'' also lying outside the third quartile.
+More than $75\%$ of articles have path lengths between 
 $0$ and $50$ links.}
@@
 
-We first identified 2-cycles, meaning a pair of articles with first link pointing to one another.
-Of the 11 million articles, 84,000 are 2-cycles. 
-The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
-``Health Care'' and ``Medical Case Management,'' ``Broadcasting House,'' and ``BBC,'' ``Secondary Education'' and ``Secondary School'' 
-(see Fig.~\ref{fig:2-cycles}).
-
 \begin{figure*}[tp!]
@@
   \label{fig:2-cycles}
+\end{figure*}
+We first identified 2-cycles, meaning a pair of articles with first link pointing to one another.
+Of the 11 million articles, 84,000 are 2-cycles. 
+The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
+``Health Care'' and ``Medical Case Management,'' ``Broadcasting House,'' and ``BBC,'' ``Secondary Education'' and ``Secondary School'' 
+(see Fig.~\ref{fig:2-cycles}).
+
+\begin{figure*}[tp!]
+  \includegraphics[width=\textwidth]{graphics/3_cycles.pdf}
+  \caption{
+\textbf{Highest ranking 3-cycles.}
+We identify groups of three articles whose links form a 3-cycle with the 
+last article in the group linking back to the first article.
+We rank each 3-cycle by the total number of traversal visits to gauge
+the most referenced group of three linked articles. 
+The area of the circles in the figure indicates the number of traversal visits.}
+  \label{fig:3-cycles}
 \end{figure*}
@@
 or judicial bodies such as the ``Legislative Assembly of Ontario.''
-
-\begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/3_cycles.pdf}
-  \caption{
-\textbf{Highest ranking 3-cycles.}
-We identify groups of three articles whose links form a 3-cycle with the 
-last article in the group linking back to the first article.
-We rank each 3-cycle by the total number of traversal visits to gauge
-the most referenced group of three linked articles. 
-The area of the circles in the figure indicates the number of traversal visits.}
-  \label{fig:3-cycles}
-\end{figure*}
 
@@
 
-To analyze the influence an article exerts in shaping the 
-structure of the FLN, we compute the number of traversal funnels for each.
-Articles directing more paths exert a greater influence over the structure
-of the FLN by increasing the accumulation of first links
-on a particular path. By measuring traversal funnels, we distinguish between an article that simply happened to fall within a cycle from an article funneling 
-many first links.
-
-Ranking articles by the number of traversal funnels we find 
-``Philosophy'' to be by far the highest-ranking article with 
-$7.37$ million paths
-(see Fig.~\ref{fig:Funnels}).
-\begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/funnels.pdf}
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
   \caption{
@@
   \label{fig:Funnels}
-\end{figure*}
-Of any article, the number of traversal funnels Philosophy holds exceeds 
-all others by more than two orders of magnitude.
-The ``Philosophy'' cycle which contains ``Existence,'' ``Awareness,'' ``Reality,'' 
-and similar articles accumulates the overwhelming proportion of its 
-references through ``Philosophy'': $7.37$ million of the $7.4$ million references
-are funneled through ``Philosophy''.
-Second on the list of highest-ranking articles by traversal funnels is 
-``Presentation'' with only $30$ thousand paths. Similarly abstract 
-concepts also rank highly such as ``Tree of life'' (30 thousand), 
-``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).
+\end{figure}
+To analyze the influence an article exerts in shaping the 
+structure of the FLN, we compute the number of traversal funnels for each.
+Articles directing more paths exert a greater influence over the structure
+of the FLN by increasing the accumulation of first links
+on a particular path. By measuring traversal funnels, we distinguish between an article that simply happened to fall within a cycle from an article funneling 
+many first links.
 
-Many high-ranking articles are remarkably topical, culturally and politically important concepts.
-For example, ``Health Care'', a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August, 2009 and February, 2010.
-Other high ranking articles include key historical events such as the ``Cold War'' or critical scarce resource with recent 
-media discussion such as ``Fossil Fuel''. 
-The highest-ranking list also includes ``Hip Hop,'' ``Cancer,'' and ``Web Page''.
 \begin{figure}[tp!]
@@
 \end{figure}
+Ranking articles by the number of traversal funnels we find 
+``Philosophy'' to be by far the highest-ranking article with 
+$7.37$ million paths
+(see Fig.~\ref{fig:Funnels}).
+Of any article, the number of traversal funnels Philosophy holds exceeds 
+all others by more than two orders of magnitude.
+The ``Philosophy'' cycle which contains ``Existence,'' ``Awareness,'' ``Reality,'' 
+and similar articles accumulates the overwhelming proportion of its 
+references through ``Philosophy'': $7.37$ million of the $7.4$ million references
+are funneled through ``Philosophy''.
+Second on the list of highest-ranking articles by traversal funnels is 
+``Presentation'' with only $30$ thousand paths. Similarly abstract 
+concepts also rank highly such as ``Tree of life'' (30 thousand), 
+``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).
+
+Many high-ranking articles are remarkably topical, culturally and politically important concepts.
+For example, ``Health Care,'' a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August, 2009 and February, 2010.
+Other high ranking articles include key historical events such as the ``Cold War'' or critical scarce resource with recent 
+media discussion such as ``Fossil Fuel.'' 
+The highest-ranking list also includes ``Hip Hop,'' ``Cancer,'' and ``Web Page.''
 
@@
 
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{graphics/views_visits.png}
+  \caption{
+\textbf{Article Popularity (by page views) compared to Traversal Visits 
+for Highest ranking 1000 articles.}
+We use the total number of page views provided by Wikipedia for the month
+of October 2015 to compare each article's popularity to traversal visits.
+While the most popular articles do not necessarily correspond to the articles
+with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
+The greatest number of articles fall within $10^{3.8}-10^5$ traversal visits and page views.}
+  \label{fig:Views and Visits}
+
+\end{figure}
 Wikipedia released an API to measure article popularity by page views
@@
 
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{graphics/funnels_visits.png}
+  \caption{
+    \textbf{Article Popularity (by page views) compared to Traversal Funnels
+    for Highest ranking 1000 articles.}
+We use the total number of page views provided by Wikipedia for the month
+of October 2015 to compare each article's popularity to traversal funnels.
+In the top left is ``Halloween'' with a significantly greater 
+number of page views in the month of October. 
+The greatest number of articles falls within $10^3-10^5$ page views, 
+but only $10^{1.2}-10^2$ traversal funnels.}
+  \label{fig:Views and Funnels}
+\end{figure}
 The ``Philosophy'' article, despite outranking every article by traversal visits,
@@
 page views for ``United States''.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/views_visits.png}
-  \caption{
-\textbf{Article Popularity (by page views) compared to Traversal Visits 
-for Highest ranking 1000 articles.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal visits.
-While the most popular articles do not necessarily correspond to the articles
-with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.}
-  \label{fig:Views and Visits}
-
-\end{figure}
 
@@
 highly-ranked articles by accumulation.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/funnels_visits.png}
-  \caption{
-    \textbf{Article Popularity (by page views) compared to Traversal Funnels
-    for Highest ranking 1000 articles.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal funnels.
-In the top left is ``Halloween'' with a significantly greater 
-number of page views in the month of October.}
-  \label{fig:Views and Funnels}
-\end{figure}
 
@@
 To process an article in a single pass, we developed a hierarchical system of flags:
-
 \begin{figure}[tp!]

### 2016-02-20 writeup/wikipedia-network.tex
graphics and language ok

@@
 network.
-We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general 
+We discover many scale-free distributions describing path length, 
+accumulation, and influence; find Philosophy at a salient center; and uncover a flow from specific to general 
 culminating around fundamental notions such as Community, State, and Science. 
 Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
-These findings enrich our view of how we connect and structure
+These findings enrich our view of the connections and structure of
 Wikipedia's ever growing store of knowledge.
@@
 \maketitle
-
-TODO: 
-\begin{itemize}
-    \item add the what to scale-free distribution in the abstract
-    \item fix end of second paragraph in Results Section\\
-    * power-law biz: rank exponent (slope of log-log), denoted $\alpha$ \\
-    * size exponent corresponds to the CDF (how many people have an income greater than x)
-    = $\gamma = \frac{1}{\alpha} + 1$.
-\item fix exponent reporting (with alpha) \\
-    * figure out what gamma size freq means
-\end{itemize}
 
@@
     \textbf{FLN Degree Distribution.}
-We fit the in-degree distribution to a linear $\log_{10}-\log_{10}$ model. The result is a 
+We fit the in-degree distribution with a linear $\log_{10}-\log_{10}$ model. 
+The result is a 
 an excellent fit with a Pearson's Correlation Coefficient of $-0.98$, yielding a 
@@
 have fewer than 9. 
-When fit with a linear model on a log-log scale (log(rank) versus log(in-degree)), 
-the model's r value is $-0.98$ suggesting a strong log-log linear fit 
-with a power law exponent of $-0.79$.\\
+We fit the in-degree distribution against each article's rank with a log-log (base 10) linear regression model.
+We obtain a Pearson's Correlation Coefficient of $-0.98$
+with a corresponding  power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$.
 
@@
     \textbf{Distribution of Traversal Visits.}
-We fit the distribution to a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal visits. 
-The model explains $86\%$ of the variation in the data, yielding an r-value of $-0.93$ 
-and a power law exponent of $-0.64$. The horizontal flattening around the highest
+We fit the distribution of traversal visits against article rank with a log-log (base 10) linear regression model log-log model and uncover two regimes: 
+with a cutoff rank of $10^{5}$.
+The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
+Coefficient of $-0.99$ with a corresponding power law rank exponent 
+$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
+The horizontal flattening around the highest
 ranking articles is a result of the cyclic structure (see discussion on cycles).}
@@
 
-In Fig.~\ref{fig:Distribution of Visits}, we show a log-log plot for the entire dataset:traversal visits against rank. 
-We observe a strong linear fit, with a Pearson's Correlation Coefficient of -.93 and 
-a power-law exponent of $-0.636$.
-A handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none.
+In Fig.~\ref{fig:Distribution of Visits}, we show a log-log (base 10) plot for the entire dataset: traversal visits against rank. 
+We fit the data to a two regime log-log linear regression with a cutoff rank 
+of $10^{5}$.
+The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
+Coefficient of $-0.99$ with a corresponding power law rank exponent 
+$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
+On the aggregate, the distribution suggests
+a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none.
 The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.
@@
 We find two regimes with the top regime (log(rank) < 4) 
-well-explained by a linear fit, yielding an r-value of $-0.99$ and a 
-power-law exponent of $-1.08$. The top regime 
+well-explained by a linear fit, yielding Pearson's 
+Correlation Coefficient of $-0.99$ and a 
+corresponding power law rank exponent $\alpha = -1.08$
+and a size exponent $\gamma = 0.074$. The top regime 
 explains the distribution of traversal funnels for the $17, 821$ 
@@
 The top regime, corresponding to the $17, 821$ articles with at least one 
-traversal funnel strongly fits a linear model with an r-value of $-0.99$. 
-The resulting power-law exponent is $-1.08$. Even within the few articles
+traversal funnel strongly fits a linear model with
+a Pearson's Correlation Coefficient of $-0.99$ and a 
+corresponding power law rank exponent $\alpha = -1.08$
+and a size exponent $\gamma = 0.074$. Even within the few articles
 influencing the structure of the FLN, only a handful of these exert most of the 

### 2016-02-22 writeup/wikipedia-network.tex
first link justfication added

@@
 in the HTML markup.
-For example, the authors of the 
+For example, the authors of 
 \href{https://en.wikipedia.org/wiki/Train}{``Train''}
-article chose
-``Rail Transport,'' ``Cargo,'' and ``Passengers'' 
-as pertinent articles to reference in describing trains
+chose
+``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
+among others as relevant articles to reference in describing ``Train''
 \cite{wiki_train}
 .
-Here we focus on the first hyperlink to another Wikipedia article, which 
-marks the earliest moment in the body text where the authors
-chose to connect the information discussed to that of another article.
+By focusing our attention on the main body of an article---excluding
+elements in the side bars and headings---we capture the introduction to the topic.
+The first link marks the earliest moment in a topic's introduction where the 
+authors choose to directly reference another article. 
+While the many links in an article may reference relevant details to a topic, 
+the first link is an association within the introductory 
+description. While 
+``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
+are links detailing particulars, the first link, ``rail transport,''
+is the topic the authors associate with ``Train'' in the introduction. 
+Collectively, the first links provide a pragmatic and interpretable 
+means to connect each article to another.
+
 By following the first hyperlink to another article in the English edition of

### 2016-02-28 writeup/thesis format/thesis.tex
edits post peter comments

--- 2016-02-22 writeup/wikipedia-network.tex
@@
-\documentclass[pre,twocolumn,twoside,superscriptaddress,floatfix, aps, 10pt]{revtex4-1}
+\documentclass[11pt]{report}
 
-\usepackage{currfile}
-\lefthyphenmin=3
-\righthyphenmin=2
+\usepackage[latin2]{inputenc}
+\usepackage{amsmath}
+\usepackage[pdftex]{graphicx}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+\usepackage{fancyhdr}
+\usepackage[left=1.5in]{geometry}
+\usepackage{geometry}
+\pagestyle{fancyplain}
+\usepackage{wrapfig}
+\usepackage{setspace}
+\usepackage{epsfig}
+\usepackage{color}
+\usepackage{tikz}
+\usepackage{thesis_MTG_3}
+\usepackage{cancel}
+\usepackage{cite}
 
-\usepackage{graphicx,epsfig,verbatim,enumerate}
-\usepackage{amssymb,amsmath}
-\usepackage{ifthen}
-
-% Fonts and typesetting settings
-\usepackage[sc]{mathpazo}
-\usepackage[T1]{fontenc}
-\linespread{1.05} % Palatino needs more space between lines
-\usepackage{microtype}
-\usepackage{textcomp}
-
-\usepackage{longtable}
-
-\usepackage{mathtools}
-
-\newboolean{twocolswitch}
-
-\newcommand{\www}[1]{\url{#1}}
-\newcommand{\req}[1]{(\ref{#1})}
-\newcommand{\Req}[1]{Eq.~(\ref{#1})}
-
-% Lettrines
-\usepackage{lettrine}
-
-%useful shortcuts
-\def\R{\ensuremath{\mathbb{R}}} %\ensuremath adds math mode, if forgotten
-\def\Q{\ensuremath{\mathbb{Q}}}
-\def\N{\ensuremath{\mathbb{N}}}
-\def\Z{\ensuremath{\mathbb{Z}}}
-\def\C{\ensuremath{\mathbb{C}}}
-
-%shorcuts with arguments
-\newcommand{\abs}[1]{\left\vert#1\right\vert} %nice absolute values
-\newcommand{\bt}[1]{\textbf{#1}} %bold
-\newcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert} %vector norm
-\renewcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-
-%piecewise function
-
-%\begin{displaymath}
-%   f(x) = \left\{
-%     \begin{array}{lr}
-%       1 & : x \in \mathbb{Q}\\
-%       0 & : x \notin \mathbb{Q}
-%     \end{array}
-%   \right.
-%\end{displaymath}
-
-%environment
-\newcommand{\tab}{\phantom{ssss}}
-
-\usepackage{hyperref}
-\usepackage{color}
-\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}
-
-%colors
-\definecolor{javagreen}{rgb}{0.25,0.5,0.35} %dark green color
-\definecolor{lightblue}{rgb}{0.149,0.545,0.824} %solarized blue
-\definecolor{sred}{rgb}{0.863, 0.196, 0.184} %solarized red
-
-\newcommand{\blue}[1]{{\leavevmode\color{lightblue}{#1}}} %solarized blue 
-\newcommand{\green}[1]{{\leavevmode\color{javagreen}{#1}}} %command for green
-\newcommand{\red}[1]{{\leavevmode\color{sred}{#1}}} %solarized red
-\newcommand{\gray}[1]{{\leavevmode\color[gray]{0.5}{#1}}} %gray text
-\newcommand{\Prob}[1]{{\rm Pr}\left(#1\right)}
-
-\setboolean{twocolswitch}{true}
+% ==================================
+% Set up title page
+% ==================================
+\title{
+Connecting every bit of knowledge: \\
+The Structure of Wikipedia's first link network
+}
+\author{Mark Ibrahim}
+\defensedate{March, 2016}
+\thesis
+\masterscience
+\matsci
+\maygrad
+\advisor{Christopher M. Danforth, Ph.D.} %advisor
+\readerone{Christopher M. Danforth, Ph.D.}
+\readertwo{Peter Sheridan Dodds, Ph.D.}
+\chair{Randall Harp, Ph.D.}
 
@@
 
-\title{\protect
-Connecting every bit of knowledge: \\
-The Structure of Wikipedia's first link network
-}
-
-\date{\today}
-
-\author{
-\firstname{Mark}
-\surname{Ibrahim}
-}
-\email{mark.s.ibrahim@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Christopher}
-\surname{M. Danforth}
-}
-\email{chris.danforth@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Peter}
-\surname{Sheridan Dodds}
-}
-
-\email{peter.dodds@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
+\maketitle
+\makeacceptance
+\pagenumbering{roman}
 
 \begin{abstract}
-  \protect
 Apples, porcupines, and the most obscure Dylan song---is every topic a few clicks from Philosophy? 
@@
 
-\maketitle
+\begin{acknowledgements}
+    Thank you to my advisors \dots
+\end{acknowledgements}
+
+\tableofcontents
+\newpage
+
+\listoffigures
+\newpage
+
+\doublespacing
 
@@
 authors choose to directly reference another article. 
-While the many links in an article may reference relevant details to a topic, 
-the first link is an association within the introductory 
+While the many links in an article reference relevant details, 
+the first link is an association within a topic's introductory 
 description. While 
@@
 is the topic the authors associate with ``Train'' in the introduction. 
+[TODO:add a few more examples]
 Collectively, the first links provide a pragmatic and interpretable 
@@
 By following the first hyperlink to another article in the English edition of
-Wikipedia as of November, 2014, we connect the ideas in one article to those in another, ultimately forming a directed network: 
+Wikipedia as of November, 2014, we connect the topic of one article to that of another, ultimately forming a directed network: 
 {\it Wikipedia's First Link Network} (FLN).
@@
 
-What information could the First Links possibly reveal?
-Inspired by the claim that the majority of First Links lead to the 
-``Philosophy'' article---popularized by an 
-\href{https://xkcd.com/903/}{xkcd}
-comic and subsequently
-discussed in blog posts 
-\cite{mat_blog, Ilmari_first_links, xkcd}
----we holistically study 
-Wikipedia's FLN as a map relating areas of human knowledge. 
-The map is a hierarchical taxonomy where, for example, ``Train'' links to a parent node, 
-``Rail Transport,'' while many child nodes, ``Steel'' and ``Horsepower,'' link to 
-``Train.'' Unlike previous taxonomies
-created by individuals
-\cite{locke, descartes, aristotle}
-or a select group of individuals 
-\cite{hist_thesaurus}, 
-the organization of ideas in the FLN 
-does not emerge from a centralized effort. 
-Rather each link in the FLN emerges from the choice of each author and ensuing edits.
-Collectively, the FLN is a wealth of relations among inventions, places,
+The FLN is a wealth of relations among inventions, places,
 figures, objects, and events across space and time.
+``Train'' for example, links to a parent node, ``Rail Transport,''
+while many child nodes such as ``Steel'' and ``Horsepower'' link to ``Train.''
+Unlike previous taxonomies created by individuals
+\cite{locke, descartes, aristotle, hist_thesaurus},
+the relations in the FLN emerge without a centralized effort 
+as the aggregate of each article's authors choice of first link.
 
@@
 
-Our greater interest is lies in the richer dynamics of how many articles fit together---how links create a flow through the FLN.
+Our greater interest is lies in the richer dynamics of how many articles fit together.
+Inspired by the claim that the majority of First Links lead to
+``Philosophy''---popularized by an 
+\href{https://xkcd.com/903/}{xkcd}
+comic and subsequently
+discussed in blog posts 
+\cite{mat_blog, Ilmari_first_links, xkcd}
+---
+we holistically study how the FLN  yields a flow of connections.
 We construct a path through the FLN with each article 
@@
 The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
-
-We use the notion of a path through the FLN to describe 
-the structure of the FLN as a flow relating topics. 
 Grounded in the dynamics of river networks
@@
 
-       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. ``An analysis of topical coverage of Wikipedia.'' Journal of Computer‐Mediated Communication 13.2 (2008): 429-440.
+       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. ``An analysis of topical coverage of Wikipedia.'' Journal of Computer Mediated Communication 13.2 (2008): 429-440.
 
@@
     \bibitem{locke} Bolton, Martha Brandt.
-``The Taxonomy of Ideas in Locke’s {\it Essay},'' The Cambridge Companion to Locke's 'Essay Concerning Human Understanding'.
-Ed.Lex Newman..
+``The Taxonomy of Ideas in Locke\textquotesingle s {\it Essay},'' The Cambridge Companion to Locke's ``Essay Concerning Human Understanding.'' Ed.Lex Newman..
 1st ed.

### 2016-02-28 writeup/wikipedia-network.tex
abstract wording tweaks

@@
 By following the first link in an article, we connect entries to form a directed network: Wikipedia's First Link Network. 
-Here, we study the English edition of Wikipedia's First Link Network for insight into how the many topics, ideas, people, objects, and events
+Here, we study the English edition of Wikipedia's First Link Network for insight into how the many 
+inventions, places, people, objects, and events
 are related and organized.  
@@
 network.
-We discover many scale-free distributions describing path length, 
-accumulation, and influence; find Philosophy at a salient center; and uncover a flow from specific to general 
-culminating around fundamental notions such as Community, State, and Science. 
-Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
+We discover scale-free distributions describe path length, 
+accumulation, and influence. Far from dispersed, first links disproportionately accumulate 
+at a few articles---flowing from specific to general and culminating around fundamental notions such as
+Community, State, and Science. Even fewer articles most influence the network's structure with Philosophy 
+shaping more paths than any other article by two orders of magnitude. 
+Curiously, we also observe a gravitation towards topical articles such as 
+Health Care and Fossil Fuel. 
 These findings enrich our view of the connections and structure of
@@
 .
+
 By focusing our attention on the main body of an article---excluding
-elements in the side bars and headings---we capture the introduction to the topic.
-The first link marks the earliest moment in a topic's introduction where the 
+elements in the side bars and headings---we 
+systematically capture the core description of a topic.
+Within the body text, the first link marks the earliest moment in a topic's introduction where the 
 authors choose to directly reference another article. 
-While the many links in an article may reference relevant details to a topic, 
-the first link is an association within the introductory 
+While many links reference relevant details, 
+the first link is an association within a topic's initial 
 description. While 
@@
 is the topic the authors associate with ``Train'' in the introduction. 
+``Banana'' has a first link to ``fruit,'' ``Bob Dylan'' has a first link to 
+``Blowing in the Wind,'' and ``Physics'' has a first link to ``natural science.''
 Collectively, the first links provide a pragmatic and interpretable 
@@
     \textbf{First Link Path For ``Train.''}
-    We follow the First Link to another Wikipedia article
+    We follow the first link to another Wikipedia article
     in the main body of the article---the area
@@
     side bar elements, the navigation bar and title. 
-    In this example, the First Link to another Wikipedia article is ``Rail Transport.'' 
-    We can again select the First Link on the ``Rail Transport''
+    In this example, the first link to another Wikipedia article is ``Rail Transport.'' 
+    We can again select the first link on the ``Rail Transport''
     article, repeating the process to 
-    form a path of First Links.
+    form a path of first links.
   }
@@
 We measure dynamics of the FLN as a flow, quantifying 
-aspects of the FLN from the accumulation of First Links around articles 
+aspects of the FLN from the accumulation of first links around articles 
 to the influence an article exerts in shaping the FLN.
@@
 
-Our greater interest is lies in the richer dynamics of how many articles fit together---how links create a flow through the FLN.
+Our greater interest is lies in the richer dynamics of how many articles fit together.
+Inspired by the claim that the majority of first links lead to
+``Philosophy''---popularized by an 
+\href{https://xkcd.com/903/}{xkcd}
+comic and subsequently
+discussed in blog posts 
+\cite{mat_blog, Ilmari_first_links, xkcd}
+---
+we holistically study how the FLN  yields a flow of connections.
 We construct a path through the FLN with each article 
 leading to the next through the first link. 
-For example ``Train'' has a first link to ``Conveyance of Passengers and Goods,'' which is itself
-an article with a first link to ``Goods'' and so on. 
+For example ``Train'' has a first link to 
+``Rail Transport,''
+ which is itself
+an article with a first link to 
+``Conveyance of Passengers and Goods,''
+and so on.
 As shown in Fig.~\ref{fig:Train First Links}, the path starting at ``Train'' 
@@
 The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
-Coefficient of $-0.99$ with a corresponding power law rank exponent 
+Coefficient of $-0.99$ 
+with a corresponding power law rank exponent 
 $\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
 The horizontal flattening around the highest
-ranking articles is a result of the cyclic structure (see discussion on cycles).}
+ranking articles is a result of the cyclic structure (see discussion on cycles).
+The bottom regime has a Pearson's Correlation Coefficient of $-0.93$
+with a corresponding power law rank exponent 
+$\alpha = -0.579$ and a size exponent $\gamma = -0.727$.}
   \label{fig:Distribution of Visits}
@@
 In Fig.~\ref{fig:Distribution of Visits}, we show a log-log (base 10) plot for the entire dataset: traversal visits against rank. 
-We fit the data to a two regime log-log linear regression with a cutoff rank 
-of $10^{5}$.
-The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
-Coefficient of $-0.99$ with a corresponding power law rank exponent 
-$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
+We obtain an excellent fit with regimes corresponding to power law rank exponents of 
+$\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
 On the aggregate, the distribution suggests
@@
 
+The highest ranking articles by traversal visits are broad topics spanning academic disciplines 
+or notions fundamental to society~(Fig. \ref{fig:highest visits}): ``Science,'' ``Mathematics,'' `` Geography,'' and ``Philosophy''
+as well as ``Community,'' ``State,'' ``Earth,'' ``Information,'' ``Power,'' and ``Communication.''
+The first links flow towards broader topics, accumulating at of these foundational notions.
+While ``Banana'' is a concrete fruit for example, 
+the first links flow from ``Fruit'' to ``Botany'' to ``Biology,'' and ultimately 
+culminate with ``Science'' and ``Philosophy.'' 
+The first links anchor the wealth of specific knowledge on Wikipedia to 
+a few notions foundational to society.
+
 \subsection{Network Cycles}
@@
 \end{figure*}
-We first identified 2-cycles, meaning a pair of articles with first link pointing to one another.
+We identify 2-cycles, meaning a pair of articles with first link pointing to one another.
 Of the 11 million articles, 84,000 are 2-cycles. 
@@
 structure of the FLN. 
-The area of each circle is sized in proportion to the number of 
-traversal funnels each article has. 
 We find ``Philosophy'' exerts an overwhelming proportion
@@
     \textbf{Distribution of Traversal Funnels.}
-We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal funnels. 
+We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed number of traversal funnels. 
 We find two regimes with the top regime (log(rank) < 4) 
@@
 corresponding power law rank exponent $\alpha = -1.08$
-and a size exponent $\gamma = 0.074$. The top regime 
-explains the distribution of traversal funnels for the $17, 821$ 
-articles with at least one traversal funnel. The bottom 
-horizontal regime corresponds to the more than $99\%$ of articles
+and a size exponent $\gamma = 0.074$. 
+The bottom horizontal regime corresponds to the more than $99\%$ of articles
 which hold zero traversal funnels.}
@@
 Wikipedia released an API to measure article popularity by page views
-starting November 2015. Page views add another dimension to our
-findings by revealing 
-how users view the information on Wikipedia. 
+starting November 2015. 
+Page views add another dimension to our
+findings by constrasting
+the number of users who access a particular article against
+the number of traversal visits and funnels.
 We measure popularity as the total number 
@@
 ``Philosophy'' emerges as an exceptional article by orders of magnitude. 
-Nevertheless, many other foundational concepts emerged naturally within the First Link Network. 
+Nevertheless, many other foundational concepts emerged naturally within FLN. 
 Basins around ``Community'', ``State'', and ``Science'' reveal a foundational structure within the network. 
@@
 
-Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the First Link.
+Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the first link.
 These findings also form the basis for the creation of a taxonomy where 
@@
 information.
-While a human could accurately identify the First Link, to map the entire First Link Network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
+While a human could accurately identify the first link, to map the entire First Link network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
 
@@
 as well as compatibility with other libraries used to parse the XML.
-Consequently,  we developed an algorithm for parsing the First Link in the XML version of each article.
+Consequently,  we developed an algorithm for parsing the first link in the XML version of each article.
 Our parsing algorithm aimed to: 
-1) accurately identify the First Link among other page elements, and 
+1) accurately identify the first link among other page elements, and 
 2) efficiently do so---that is without needing for several passes through the data.
@@
 until the flag's closing markup.
-A First Link is identified only if Flags 1, 2, and 3 are all off.
+A first link is identified only if Flags 1, 2, and 3 are all off.
 In this case, the entire link is retrieved. 
@@
 \cite{media_wiki_templates}.
-The First Link of an article is then the earliest valid link with unraised flags.
+The first link of an article is then the earliest valid link with unraised flags.
 
@@
 We then joined the results to form a hash table containing every Wikipedia article and its corresponding
-First Link. The resulting network map is the basis of our analysis.
+first link. The resulting network map is the basis of our analysis.
 

### 2016-03-01 writeup/wikipedia-network.tex
link to online appendix and enhanced figure placement

@@
     \textbf{Highest Ranking Articles by in-degree.}
-We rank each article by the number of direct first links to the article (in-degree). The highest-ranking articles tend to represent geographical and biological abstractions.}
+    We rank each article by the number of direct first links to the article (in-degree).
+The highest-ranking articles tend to represent geographical and biological abstractions.
+A full online appendix of the results and data is available \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.}
   \label{fig:indegree list}
@@
 
-\subsection{Network Cycles}
-
 \begin{figure*}[tp!]
@@
 \end{figure*}
-We identify 2-cycles, meaning a pair of articles with first link pointing to one another.
-Of the 11 million articles, 84,000 are 2-cycles. 
-The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
-``Health Care'' and ``Medical Case Management,'' ``Broadcasting House,'' and ``BBC,'' ``Secondary Education'' and ``Secondary School'' 
-(see Fig.~\ref{fig:2-cycles}).
+\subsection{Network Cycles}
 
@@
 \end{figure*}
+We identify 2-cycles, meaning a pair of articles with first link pointing to one another.
+Of the 11 million articles, 84,000 are 2-cycles. 
+The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
+``Health Care'' and ``Medical Case Management,'' ``Broadcasting House,'' and ``BBC,'' ``Secondary Education'' and ``Secondary School'' 
+(see Fig.~\ref{fig:2-cycles}).
 
@@
 
+An online appendix containing all of the code and data used can be accessed \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.
+
 \begin{thebibliography}{1}

### 2016-03-07 writeup/wikipedia-network.tex
chris edits incorporated up to results section

@@
 Connecting every bit of knowledge: \\
-The Structure of Wikipedia's first link network
+The structure of Wikipedia's first link network
 }
@@
   \protect
-Apples, porcupines, and the most obscure Dylan song---is every topic a few clicks from Philosophy? 
+Apples, porcupines, and the most obscure Bob Dylan song---is every topic a few clicks from Philosophy? 
 Within Wikipedia, the surprising answer is yes: nearly all 
@@
 More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
-By following the first link in an article, we connect entries to form a directed network: Wikipedia's First Link Network. 
+By following the first link in each article, we algorithmically construct a directed network of 
+all 4.7 million articles: Wikipedia's First Link Network.
 Here, we study the English edition of Wikipedia's First Link Network for insight into how the many 
-inventions, places, people, objects, and events
-are related and organized.  
-We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
-We traverse every possible path through the network, 
-measuring the accumulation of first links, path lengths, basins, cycles, and the influence a particular article exerts in shaping the 
-network.
+inventions, places, people, objects, and events are related and organized.  
+We traverse every path, measuring the accumulation of first links, path lengths, basins, cycles, 
+and the influence each article exerts in shaping the network.
 We discover scale-free distributions describe path length, 
@@
 at a few articles---flowing from specific to general and culminating around fundamental notions such as
-Community, State, and Science. Even fewer articles most influence the network's structure with Philosophy 
-shaping more paths than any other article by two orders of magnitude. 
+Community, State, and Science. 
+Philosophy shapes more paths than any other article by two orders of magnitude. 
 Curiously, we also observe a gravitation towards topical articles such as 
@@
 \cite{drugs, stats}.
-All this information does not lie dormant. 
+Demonstrating its relevance to modern inquiry,
 Wikipedia is the sixth most visited
@@
 \cite{bias_women}.
-Wikipedia's content has also proven a powerful tool. 
+Wikipedia's content has also proven to be a powerful tool. 
 Researchers have used Wikipedia to identify missing dictionary entries
@@
 in the HTML markup.
-For example, the authors of 
+For example, the authors of the 
 \href{https://en.wikipedia.org/wiki/Train}{``Train''}
+article
 chose
@@
 {\it Wikipedia's First Link Network} (FLN).
+For methodological details, see section~\ref{Appendix}.
 
@@
     inside the green rectangle, which excludes 
-    side bar elements, the navigation bar and title. 
+    side bar elements, the navigation bar and title; the first link is circled in red. 
     In this example, the first link to another Wikipedia article is ``Rail Transport.'' 
@@
     article, repeating the process to 
-    form a path of first links.
+    form a path of first links. After 11 links, we arrive at ``Philosophy.''
   }
@@
 We measure dynamics of the FLN as a flow, quantifying 
-aspects of the FLN from the accumulation of first links around articles 
-to the influence an article exerts in shaping the FLN.
+the accumulation of first links around articles 
+and the influence an article exerts in shaping the FLN.
 Together with cycles, in-degree, depth, and the content of the articles, 
@@
 particular article. 
-We measure the in-degree in the FLN to gauge the number of articles referencing a particular article.
 Articles with zero in-degree have no references---they are outer leaves in the FLN. 
-Similar to PageRank, the in-degree 
-provides a way to rank the articles in Wikipedia, allowing us to see 
-whether authors tend to reference all articles equally or only a few.
+The in-degree 
+provides a way to rank the articles in Wikipedia by the number of first link references.
 
-Our greater interest is lies in the richer dynamics of how many articles fit together.
 Inspired by the claim that the majority of first links lead to
@@
 discussed in blog posts 
-\cite{mat_blog, Ilmari_first_links, xkcd}
+\cite{xkcd, mat_blog, Ilmari_first_links}
 ---
 we holistically study how the FLN  yields a flow of connections.
-We construct a path through the FLN with each article 
-leading to the next through the first link. 
 For example ``Train'' has a first link to 
@@
 The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
-Grounded in the dynamics of river networks
-\cite{geo_basins},
-we characterize features of the FLN as a flow through first links.
 Previous studies have used flow to characterize the structure of river networks
-\cite{dodds} and describe the organization of food systems through transportation networks
+\cite{geo_basins, dodds} and describe the organization of food systems through transportation networks
 \cite{food_webs}.
@@
 The traversal visits matrix for 
-our Wikipedia dataset consists of 121 million entries to account for each path through the FLN.
+our Wikipedia dataset consists of 121 million entries encoding each path out of the more than 
+googol possible paths ($4.7*10^6*2^{4.7*10^6}$) through the FLN .
 
@@
 The path length describes how closely related topics are.
-The greater the number of first links separating two articles, the greater the number of ideas needed to relate their corresponding ideas. Although
-``Train'' is related to ``Economics'' for example, there are several articles
+Although ``Train'' is related to ``Economics'' for example, there are several articles
 bridging the connection:
@@
 influences the shape of the FLN. 
-At a large point of accumulation, a single article's first link 
+At a point of accumulation, a single article's first link 
 can exert great influence over the shape of the FLN by directing many
@@
 directs the flow of links towards the 3-cycle, while articles A and B are 
-recipients of the flow---without exerting any direct influence themselves.
+recipients of the flow---without exerting direct influence themselves.
 
@@
 \section{Appendix}
+\label{Appendix}
 

### 2016-03-07 writeup/wikipedia-network.tex
edits to text incorporated, figures remain

@@
 
+\begin{figure*}[tp!]
+  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
+  \caption{
+    \textbf{First Link Path For ``Train.''}
+    We follow the first link  to another Wikipedia article
+    in the main body of the article---the area
+    inside the green rectangle, which excludes 
+    side bar elements, the navigation bar and title; the first link is circled in red. 
+    In this example, the first link to another Wikipedia article is ``Rail Transport.'' 
+    We can again select the first link on the ``Rail Transport''
+    article, repeating the process to 
+    form a path of first links. After 11 links, we arrive at ``Philosophy.''
+  }
+  \label{fig:Train First Links}
+\end{figure*}
+
 By focusing our attention on the main body of an article---excluding
@@
 For methodological details, see section~\ref{Appendix}.
-
-\begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
-  \caption{
-    \textbf{First Link Path For ``Train.''}
-    We follow the first link  to another Wikipedia article
-    in the main body of the article---the area
-    inside the green rectangle, which excludes 
-    side bar elements, the navigation bar and title; the first link is circled in red. 
-    In this example, the first link to another Wikipedia article is ``Rail Transport.'' 
-    We can again select the first link on the ``Rail Transport''
-    article, repeating the process to 
-    form a path of first links. After 11 links, we arrive at ``Philosophy.''
-  }
-  \label{fig:Train First Links}
-\end{figure*}
 
@@
     \textbf{Traversal Visit Algorithm on a sample network.}
-     The traversal visit vectors are an adjacency matrix for the paths through the network: the first column is indicates the path formed starting with article A.
-The number of traversal visits for article A is then the number of paths containing A or the sum of the first row in our matrix:
+     The traversal visit vectors are an adjacency matrix for the paths through the network: 
+     the first column indicates the path formed starting with article A. 
+     The number of traversal visits for article A is then the number of paths containing A 
+     or the sum of the first row in our matrix:
      $\sum_{i=1}^7 A_{\text{visit, i}} = 7$.
@@
 geographical proximity; sports teams describe an ever changing collection of 
-sports players often associated with a cultural identity or a geographical 
+players often associated with a cultural identity or a geographical 
 region. 
@@
 Instead, first links flow towards ``Philosophy'' as the 
-ultimate anchor, by generalizing from specific articles to broader notions.
+ultimate anchor, by generalizing from specific to broad.
 
@@
 receive most direct first references, while most articles receive few or none.
-The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5 links.
+The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5.
 Less than $1\%$ of articles have more than 100 direct first links and $75\%$ of articles
 have fewer than 9. 
-We fit the in-degree distribution against each article's rank with a log-log (base 10) linear regression model.
+We fit the in-degree distribution against each article's rank with a log-log (base 10) linear regression model
+(see Fig.~\ref{fig:degree distribution}).
 We obtain a Pearson's Correlation Coefficient of $-0.98$
@@
 
-We first seek to describe the depth of the FLN: how many links does a 
-connection among articles span? 
+How many links does a connection among articles span? 
 We find the longest path length is 365,
@@
 Each day's Liturgics links to the next day's. On the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
-(see discussion of cycles).
+(see discussion of cycles ~\ref{cycles}).
 We also find similarly lengthy paths following the evolution of a place or topic through time: 
@@
 
-In Fig.~\ref{fig:Distribution of Visits}, we show a log-log (base 10) plot for the entire dataset: traversal visits against rank. 
-We obtain an excellent fit with regimes corresponding to power law rank exponents of 
+In Fig.~\ref{fig:Distribution of Visits}, we show traversal visits against rank, yielding
+an excellent fit with regimes corresponding to power law rank exponents of 
 $\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
@@
 Traversal Algorithm section for details). In doing so, we can rank each article
-by the accumulation of first links. Articles with a greater number of traversal visits
-mark greater points of first link accumulation. The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.}
+by the accumulation of first links.
+The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.}
   \label{fig:highest visits}
@@
 traversal visits to gauge the most referenced groups of two articles linked
-to each other. The area of the circles in the figure indicates the number of traversal visits. We find 2-cycles often capture synonyms or articles representing nearly the 
+to each other. We find 2-cycles often capture synonyms or articles representing nearly the 
 same concepts as opposed to distinct concepts.}
@@
 \subsection{Network Cycles}
+\label{cycles}
 
@@
 We identify 2-cycles, meaning a pair of articles with first link pointing to one another.
-Of the 11 million articles, 84,000 are 2-cycles. 
+Of the 11 million articles, roughly 84,000 are memebers of 2-cycles. 
 The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
@@
 
-Excluding basins around ``Philosophy,'' we find other basins around 
+Moving beyon basins around ``Philosophy,'' we find other basins around 
 foundational concepts such as ``Community,'' ``Landmass,'' ``Federal Government,'' 
@@
 FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
-an more than $99\%$ with none---most articles are recipients of 
+more than $99\%$ with none---most articles are recipients of 
 the references flowing through the articles with at least one traversal funnel.
@@
 corresponding power law rank exponent $\alpha = -1.08$
-and a size exponent $\gamma = 0.074$. Even within the few articles
+and a size exponent $\gamma = 0.074$ 
+(see Fig.~\ref{fig:Funnels Distribution}). Even within the few articles
 influencing the structure of the FLN, only a handful of these exert most of the 
@@
 with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-The greatest number of articles fall within $10^{3.8}-10^5$ traversal visits and page views.}
+The greatest number of articles fall within roughly 1 million traversal visits and page views.}
   \label{fig:Views and Visits}
@@
 $1.4$ million views in October. The next most popular articles are 
-``India,'' ``World War II,'' and the ``United Kingdom'' each with $0.7-1.03$ million page views.
+``India,'' ``World War II,'' and the ``United Kingdom'' each with roughly 1 million page views. 
+The ``Philosophy'' article, despite outranking every article by traversal visits,
+has only $240$ thousand page views in October.
 
@@
 \end{figure}
-The ``Philosophy'' article, despite outranking every article by traversal visits,
-has only $240$ thousand page views in October compared to the $1.4$ million
-page views for ``United States''.
 
@@
 Other standout popular October articles include ``24-hour'' and ``12-hour'' clocks likely due
-to Daylight savings in October and ``Marriage'' likely due to the drop in the number of
+to Daylight savings in October and ``Marriage'' possibly due to the drop in the number of
 weddings in the months following October
 \cite{weddings}.
-``Philosophy'' which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears no where in the top 20
+``Philosophy'' which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears nowhere in the top 20
 by traversal visits. ``Philosophy'' is a relatively popular article 
@@
 The findings here should only be considered within the limitations of their context.
-We examined only the English editiion of Wikipedia at a particular moment in time, 
-considering only the first link in in each article as a means to related one article to another.
-Finally, Wikipedia, while the largest 
+We examined only the English edition of Wikipedia at a particular moment in time, 
+considering only the first link in in each article as a means to relate articles.
+Finally, Wikipedia, while the largest digital 
 collection of human knowledge, is rife with the biases of the many contributing editors
-\cite{bias}
-. Nevertheless, the findings do reveal
+\cite{bias}.
+Nevertheless, the findings do reveal
 generalizable relationships and point to foundational notions.
 
-Among our observations is the multiple appearance of scale-free distributions within the network. 
+Among our observations is the appearance of multiple scale-free distributions within the network. 
 Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
@@
 Applications could range from an enhanced network of ideas to psychological insights into how humans form associations.
-Specifically, an ever-evolving reference of related hierarchical concepts could be applied to search engine algorithms 
-or natural language processing.
+Specifically, an ever-evolving reference of related hierarchical concepts could be 
+used to improve search engine algorithms or natural language processing.
 
@@
 describing how the majority of links lead to ``Philosophy,'' inspiring this research.
+
+PSD and CMD acknowledge support from NSF Big Data Grand $\#1447634$.
+
+The authors also acknowledge support from the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
 
@@
 \cite{wiki_edits},
-our aim is to characterize the dynamics of the First Link Network, not record a particular link between one
-article and another.
+our aim is to characterize the structure of the First Link Network.
 
@@
 information.
-While a human could accurately identify the first link, to map the entire First Link network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
+While a human could manually identify the first link, to map the entire First Link network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
 
 While some libraries exist for MediaWiki Markup,
-Approaches using existing libraries led to several bugs 
+approaches using existing libraries led to several bugs 
 including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
@@
 across 112 cores of the UVM supercomputer cluster
-\cite{vacc}
+\cite{vacc}.
 We then joined the results to form a hash table containing every Wikipedia article and its corresponding
@@
 
-    \bibitem{vacc} The authors acknowledge the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
-    \url{http://www.uvm.edu/vacc}
+    \bibitem{vacc}  Vermont Advanced Computing Core, \url{http://www.uvm.edu/vacc}.
 

### 2016-03-09 writeup/wikipedia-network.tex
chris suggestions incorporated

@@
 \end{figure}
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
+  \caption{
+    \textbf{FLN Degree Distribution.}
+We fit the in-degree distribution with a linear $\log_{10}-\log_{10}$ model. 
+The result is a 
+an excellent fit with a Pearson's Correlation Coefficient of $-0.98$, yielding a 
+power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$. 
+}
+  \label{fig:degree distribution}
+\end{figure}
 
@@
 ultimate anchor, by generalizing from specific to broad.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
-  \caption{
-    \textbf{FLN Degree Distribution.}
-We fit the in-degree distribution with a linear $\log_{10}-\log_{10}$ model. 
-The result is a 
-an excellent fit with a Pearson's Correlation Coefficient of $-0.98$, yielding a 
-power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$. 
-}
-  \label{fig:degree distribution}
-\end{figure}
 
@@
 Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
-This roughly corresponds to the official number of articles on Wikipedia: 
-$~4.7$ million as of November 2014---approximately half of the 11 million 
-articles in the XML dump are redirects or disambiguations, not full articles.
-The most common path length is 29, with an interquartile range of 4---path length
-is typically between 26 and 29 articles.
-As a distribution, more than $75\%$ of articles have a path length below 
-$50$ first links 
-while a few temporally organized paths exceed 50 links 
-(see Fig.~\ref{fig:Path Length Distribution}).
-
-\subsection{Traversal Visits}
-
-As a distribution, the number of traversal visits by article appears to follow a decaying power law. 
-The majority of articles have fewer than 30 traversal visits and
-first link references accumulate at a few articles.
-Specifically, $99.76\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
-Meanwhile, the highest ranking 30 articles have an extremely disproportionate number of traversal visits.
-
 \begin{figure}[tp!]
@@
 
-In Fig.~\ref{fig:Distribution of Visits}, we show traversal visits against rank, yielding
-an excellent fit with regimes corresponding to power law rank exponents of 
-$\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
-On the aggregate, the distribution suggests
-a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none.
-The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.
-
 \begin{figure}[tp!]
@@
 
+This roughly corresponds to the official number of articles on Wikipedia: 
+$~4.7$ million as of November 2014---approximately half of the 11 million 
+articles in the XML dump are redirects or disambiguations, not full articles.
+The most common path length is 29, with an interquartile range of 4---path length
+is typically between 26 and 29 articles.
+As a distribution, more than $75\%$ of articles have a path length below 
+$50$ first links 
+while a few temporally organized paths exceed 50 links 
+(see Fig.~\ref{fig:Path Length Distribution}).
+
+\subsection{Traversal Visits}
+
+As a distribution, the number of traversal visits by article appears to follow a decaying power law. 
+The majority of articles have fewer than 30 traversal visits and
+first link references accumulate at a few articles.
+Specifically, $99.76\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
+Meanwhile, the highest ranking 30 articles have an extremely disproportionate number of traversal visits.
+
+\begin{figure*}[tph!]
+  \includegraphics[width=\textwidth]{graphics/2_and_3_cycles.pdf}
+  \caption{
+    \textbf{Highest ranking 2-cycles and 3-cycles.}
+We identify pairs of articles whose first links point to one another, forming
+a 2-cycle. We then rank each pair of articles by the total number of 
+traversal visits to gauge the most referenced groups of two articles linked
+to each other. We find 2-cycles often capture synonyms or articles representing nearly the 
+same concepts as opposed to distinct concepts. Similarly, we identify and 
+rank 3-cycles to find they appear to capture three closely related ideas or synonyms.}
+  \label{fig:2-cycles}
+\end{figure*}
+In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against rank
+to two regimes corresponding to power law rank exponents of 
+$\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
+On the aggregate, the distribution suggests
+a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none.
+The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.
+
 The highest ranking articles by traversal visits are broad topics spanning academic disciplines 
@@
 
-\begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/2_cycles.pdf}
-  \caption{
-    \textbf{Highest ranking 2-cycles.}
-We identify pairs of articles whose first links point to one another, forming
-a 2-cycle. We then rank each pair of articles by the total number of 
-traversal visits to gauge the most referenced groups of two articles linked
-to each other. We find 2-cycles often capture synonyms or articles representing nearly the 
-same concepts as opposed to distinct concepts.}
-  \label{fig:2-cycles}
-\end{figure*}
 \subsection{Network Cycles}
 \label{cycles}
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
+  \caption{
+    \textbf{Funnels.}
+We represent the highest-ranking articles by the number of traversal 
+funnels to gauge the influence each article exerts in shaping the 
+structure of the FLN. 
+We find ``Philosophy'' exerts an overwhelming proportion
+of the influence, with other abstract notions and topical concepts ranking
+next.}
+  \label{fig:Funnels}
+\end{figure}
 
-\begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/3_cycles.pdf}
-  \caption{
-\textbf{Highest ranking 3-cycles.}
-We identify groups of three articles whose links form a 3-cycle with the 
-last article in the group linking back to the first article.
-We rank each 3-cycle by the total number of traversal visits to gauge
-the most referenced group of three linked articles. 
-The area of the circles in the figure indicates the number of traversal visits.}
-  \label{fig:3-cycles}
-\end{figure*}
 We identify 2-cycles, meaning a pair of articles with first link pointing to one another.
@@
 
-\subsection{Traversal Funnels}
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
-  \caption{
-    \textbf{Funnels.}
-We represent the highest-ranking articles by the number of traversal 
-funnels to gauge the influence each article exerts in shaping the 
-structure of the FLN. 
-We find ``Philosophy'' exerts an overwhelming proportion
-of the influence, with other abstract notions and topical concepts ranking
-next.}
-  \label{fig:Funnels}
-\end{figure}
-To analyze the influence an article exerts in shaping the 
-structure of the FLN, we compute the number of traversal funnels for each.
-Articles directing more paths exert a greater influence over the structure
-of the FLN by increasing the accumulation of first links
-on a particular path. By measuring traversal funnels, we distinguish between an article that simply happened to fall within a cycle from an article funneling 
-many first links.
-
 \begin{figure}[tp!]
@@
 \end{figure}
+
+\subsection{Traversal Funnels}
+
+To analyze the influence an article exerts in shaping the 
+structure of the FLN, we compute the number of traversal funnels for each.
+Articles directing more paths exert a greater influence over the structure
+of the FLN by increasing the accumulation of first links
+on a particular path. By measuring traversal funnels, we distinguish between an article that simply happened to fall within a cycle from an article funneling 
+many first links.
+
 Ranking articles by the number of traversal funnels we find 
@@
 
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{graphics/popularity_visits.pdf}
+  \caption{
+\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal visits.}
+We use the total number of page views provided by Wikipedia for the month
+of October 2015 to compare each article's popularity to traversal visits.
+While the most popular articles do not necessarily correspond to the articles
+with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
+The greatest number of articles fall within roughly 1 million traversal visits and page views.
+Overall, the page views for the top 1000 articles articles by traversal visits appear to be log-normally distributed.
+}
+  \label{fig:Views and Visits}
+
+\end{figure}
 As a distribution, we find few articles influence the structure of the 
@@
 
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{graphics/popularity_funnels.pdf}
+  \caption{
+\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal funnels.}
+We use the total number of page views provided by Wikipedia for the month
+of October 2015 to compare each article's popularity to traversal funnels.
+}
+  \label{fig:Views and Funnels}
+\end{figure}
 \subsection{Article Popularity}
 
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/views_visits.png}
-  \caption{
-\textbf{Article Popularity (by page views) compared to Traversal Visits 
-for Highest ranking 1000 articles.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal visits.
-While the most popular articles do not necessarily correspond to the articles
-with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-The greatest number of articles fall within roughly 1 million traversal visits and page views.}
-  \label{fig:Views and Visits}
-
-\end{figure}
 Wikipedia released an API to measure article popularity by page views
@@
 has only $240$ thousand page views in October.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/funnels_visits.png}
-  \caption{
-    \textbf{Article Popularity (by page views) compared to Traversal Funnels
-    for Highest ranking 1000 articles.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal funnels.
-In the top left is ``Halloween'' with a significantly greater 
-number of page views in the month of October. 
-The greatest number of articles falls within $10^3-10^5$ page views, 
-but only $10^{1.2}-10^2$ traversal funnels.}
-  \label{fig:Views and Funnels}
-\end{figure}
 
@@
 first link. The resulting network map is the basis of our analysis.
-
 An online appendix containing all of the code and data used can be accessed \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.

### 2016-04-12 writeup/wikipedia-network.tex
edits up to page 3

@@
 Connecting every bit of knowledge: \\
-The structure of Wikipedia's first link network
+The structure of Wikipedia's Flrst Link Network
 }
@@
 Here, we study the English edition of Wikipedia's First Link Network for insight into how the many 
-inventions, places, people, objects, and events are related and organized.  
+articles on inventions, places, people, objects, and events are related and organized.  
 We traverse every path, measuring the accumulation of first links, path lengths, basins, cycles, 
 and the influence each article exerts in shaping the network.
-We discover scale-free distributions describe path length, 
+We find scale-free distributions describe path length, 
 accumulation, and influence. Far from dispersed, first links disproportionately accumulate 
@@
 Community, State, and Science. 
-Philosophy shapes more paths than any other article by two orders of magnitude. 
-Curiously, we also observe a gravitation towards topical articles such as 
+Philosophy funnels more paths than any other article by two orders of magnitude. 
+We also observe a gravitation towards topical articles such as 
 Health Care and Fossil Fuel. 
@@
 
+%TODO:
+% Abstract 
+    % In traversing every path, we measure ?
+    % Our findings... braided network (what does this mean?)
+% Formatting
+    % incorporate into template
+% Intro
+    % add month to monthly view stats
+% Method
+    % what to cite for drainage basin in geomorphology
+
+% continue on Page 3
+
 \section{Introduction}
 
-Wikipedia is a towering achievement of the modern era. 
+Wikipedia is a towering achievement of the Internet age. 
 At no point in history has a larger or more meticulously indexed collection of human knowledge 
@@
 Wikipedia is the sixth most visited
-site in the world, surpassing 18 billion page views and 26 million edits in a single month
+site in the world, surpassing 18 billion page views and 11 million edits in a single month.
 \cite{wiki_edits, wiki_views}.
@@
 While these many studies have dissected and fruitfully applied Wikipedia's content,
-few have examined the connections among the many articles
-\cite{links_that_speak}.
+here we examine the connections among the many articles.
 A hyperlink from one Wikipedia article to another
@@
 \cite{classifier}.
-We treat a hyperlink as a mechanism connecting two topics.
+Here, we treat a hyperlink as a mechanism connecting two topics.
 
@@
 in the HTML markup.
-For example, the authors of the 
+For example, as of November 2014, the authors of the 
 \href{https://en.wikipedia.org/wiki/Train}{``Train''}
 article
-chose
+had collectively chosen
 ``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
@@
 By focusing our attention on the main body of an article---excluding
-elements in the side bars and headings---we 
-systematically capture the core description of a topic.
+elements in side bars and headings---we 
+attempt to systematically capture the core description of a topic.
 Within the body text, the first link marks the earliest moment in a topic's introduction where the 
@@
 are links detailing particulars, the first link, ``rail transport,''
-is the topic the authors associate with ``Train'' in the introduction. 
+is the topic the authors associate with ``Train'' in the introduction
+(Fig.~\ref{fig:Train First Links}).
 ``Banana'' has a first link to ``fruit,'' ``Bob Dylan'' has a first link to 
 ``Blowing in the Wind,'' and ``Physics'' has a first link to ``natural science.''
-Collectively, the first links provide a pragmatic and interpretable 
+Collectively, first links provide a pragmatic and interpretable 
 means to connect each article to another.
 
-By following the first hyperlink to another article in the English edition of
-Wikipedia as of November, 2014, we connect the topic of one article to that of another, ultimately forming a directed network: 
+By following the first hyperlink contained in all articles of the English edition of
+Wikipedia as of November, 2014, we form a directed network:
 {\it Wikipedia's First Link Network} (FLN).
-For methodological details, see section~\ref{Appendix}.
+Inspired by the claim that the majority of first links lead to
+``Philosophy''---popularized by an 
+\href{https://xkcd.com/903/}{xkcd}
+comic and subsequently
+discussed in blog posts 
+\cite{xkcd, mat_blog, Ilmari_first_links}
+---
+we holistically study how the FLN  yields a flow of connections.
+For methodological details, see Sec.~\ref{Appendix}.
 
@@
 while many child nodes such as ``Steel'' and ``Horsepower'' link to ``Train.''
+As shown in Fig.~\ref{fig:Train First Links}, the path starting at ``Train'' 
+contains ``Goods,'' ``Economics,'' ``Social Science,''
+leading ultimately to ``Philosophy''.
 Unlike previous taxonomies created by individuals
@@
 Our goal is to study the structure of the FLN for insight into how the information on Wikipedia is organized and related.
-We develop metrics to capture 
-the dominant features of the FLN's structure.
-We measure dynamics of the FLN as a flow, quantifying 
+Informed by river network metrics,
+we consider the FLN as a flow, quantifying 
 the accumulation of first links around articles 
@@
 particular article. 
-Articles with zero in-degree have no references---they are outer leaves in the FLN. 
-The in-degree 
-provides a way to rank the articles in Wikipedia by the number of first link references.
-
-Inspired by the claim that the majority of first links lead to
-``Philosophy''---popularized by an 
-\href{https://xkcd.com/903/}{xkcd}
-comic and subsequently
-discussed in blog posts 
-\cite{xkcd, mat_blog, Ilmari_first_links}
----
-we holistically study how the FLN  yields a flow of connections.
-For example ``Train'' has a first link to 
-``Rail Transport,''
- which is itself
-an article with a first link to 
-``Conveyance of Passengers and Goods,''
-and so on.
-As shown in Fig.~\ref{fig:Train First Links}, the path starting at ``Train'' 
-contains ``Goods,'' ``Economics,'' ``Social Science,''
-leading ultimately to ``Philosophy''.
+Articles with zero in-degree have no references---they are outer leaves in the FLN 
+or are disconnected entirely.
 
@@
 Previous studies have used flow to characterize the structure of river networks
-\cite{geo_basins, dodds} and describe the organization of food systems through transportation networks
+\cite{geo_basins, dodds} and describe the organization of 
+blood networks, food systems, and transportation networks
 \cite{food_webs}.
-With paths to mark flow through the FLN, we develop metrics to 
-gauge the accumulation of references, 
-the length of the path relating articles, and the influence a particular
-article exerts in shaping the flow of links through the FLN.
 
@@
 associated with the article. 
-We continue until the first link is retraced or is invalid---defined here to mean outside of Wikipedia.
+We continue until a page is revisited or is invalid---defined here to mean pointing to a page outside of Wikipedia.
 We select a second article and repeat the process until we have 
-constructed a path for each article in the network. We call the resulting count for each article the number of {\it traversal visits}. The number of traversal visits of an article 
-measures the number of references flowing to the ideas in the article---equivalent
-to basin area in geomorphology.
+constructed a path for each article in the network. 
+We define the number of {\it traversal visits} of an article 
+to be the number of references flowing to the ideas in the article---equivalent
+to drainage basin area in geomorphology.
 
@@
   \caption{
-    \textbf{Traversal Funnel Algorithm on a sample network.}
+    \textbf{Traversal Funnel Algorithm on a simple network.}
   The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
@@
 
+The authors are also grateful for the suggestions and input provided by Randall Harp.
+
 PSD and CMD acknowledge support from NSF Big Data Grand $\#1447634$.

### 2016-04-22 writeup/2016-04wikipedia-network/wikipedia-network.acknowledgments.tex
universal template and bibtex format changed

--- 2016-04-12 writeup/wikipedia-network.tex
@@
-\documentclass[pre,twocolumn,twoside,superscriptaddress,floatfix, aps, 10pt]{revtex4-1}
-
-\usepackage{currfile}
-\lefthyphenmin=3
-\righthyphenmin=2
-
-\usepackage{graphicx,epsfig,verbatim,enumerate}
-\usepackage{amssymb,amsmath}
-\usepackage{ifthen}
-
-% Fonts and typesetting settings
-\usepackage[sc]{mathpazo}
-\usepackage[T1]{fontenc}
-\linespread{1.05} % Palatino needs more space between lines
-\usepackage{microtype}
-\usepackage{textcomp}
-
-\usepackage{longtable}
-
-\usepackage{mathtools}
-
-\newboolean{twocolswitch}
-
-\newcommand{\www}[1]{\url{#1}}
-\newcommand{\req}[1]{(\ref{#1})}
-\newcommand{\Req}[1]{Eq.~(\ref{#1})}
-
-% Lettrines
-\usepackage{lettrine}
-
-%useful shortcuts
-\def\R{\ensuremath{\mathbb{R}}} %\ensuremath adds math mode, if forgotten
-\def\Q{\ensuremath{\mathbb{Q}}}
-\def\N{\ensuremath{\mathbb{N}}}
-\def\Z{\ensuremath{\mathbb{Z}}}
-\def\C{\ensuremath{\mathbb{C}}}
-
-%shorcuts with arguments
-\newcommand{\abs}[1]{\left\vert#1\right\vert} %nice absolute values
-\newcommand{\bt}[1]{\textbf{#1}} %bold
-\newcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert} %vector norm
-\renewcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-
-%piecewise function
-
-%\begin{displaymath}
-%   f(x) = \left\{
-%     \begin{array}{lr}
-%       1 & : x \in \mathbb{Q}\\
-%       0 & : x \notin \mathbb{Q}
-%     \end{array}
-%   \right.
-%\end{displaymath}
-
-%environment
-\newcommand{\tab}{\phantom{ssss}}
-
-\usepackage{hyperref}
-\usepackage{color}
-\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}
-
-%colors
-\definecolor{javagreen}{rgb}{0.25,0.5,0.35} %dark green color
-\definecolor{lightblue}{rgb}{0.149,0.545,0.824} %solarized blue
-\definecolor{sred}{rgb}{0.863, 0.196, 0.184} %solarized red
-
-\newcommand{\blue}[1]{{\leavevmode\color{lightblue}{#1}}} %solarized blue 
-\newcommand{\green}[1]{{\leavevmode\color{javagreen}{#1}}} %command for green
-\newcommand{\red}[1]{{\leavevmode\color{sred}{#1}}} %solarized red
-\newcommand{\gray}[1]{{\leavevmode\color[gray]{0.5}{#1}}} %gray text
-\newcommand{\Prob}[1]{{\rm Pr}\left(#1\right)}
-
-\setboolean{twocolswitch}{true}
-
-\begin{document}
-
-\title{\protect
-Connecting every bit of knowledge: \\
-The structure of Wikipedia's Flrst Link Network
-}
-
-\date{\today}
-
-\author{
-\firstname{Mark}
-\surname{Ibrahim}
-}
-\email{mark.s.ibrahim@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Christopher}
-\surname{M. Danforth}
-}
-\email{chris.danforth@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Peter}
-\surname{Sheridan Dodds}
-}
-
-\email{peter.dodds@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\begin{abstract}
-  \protect
-Apples, porcupines, and the most obscure Bob Dylan song---is every topic a few clicks from Philosophy? 
-Within Wikipedia, the surprising answer is yes: nearly all 
-paths lead to Philosophy.
-Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
-More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
-By following the first link in each article, we algorithmically construct a directed network of 
-all 4.7 million articles: Wikipedia's First Link Network.
-Here, we study the English edition of Wikipedia's First Link Network for insight into how the many 
-articles on inventions, places, people, objects, and events are related and organized.  
-We traverse every path, measuring the accumulation of first links, path lengths, basins, cycles, 
-and the influence each article exerts in shaping the network.
-We find scale-free distributions describe path length, 
-accumulation, and influence. Far from dispersed, first links disproportionately accumulate 
-at a few articles---flowing from specific to general and culminating around fundamental notions such as
-Community, State, and Science. 
-Philosophy funnels more paths than any other article by two orders of magnitude. 
-We also observe a gravitation towards topical articles such as 
-Health Care and Fossil Fuel. 
-These findings enrich our view of the connections and structure of
-Wikipedia's ever growing store of knowledge.
-
-\end{abstract}
-
-\maketitle
-
-%TODO:
-% Abstract 
-    % In traversing every path, we measure ?
-    % Our findings... braided network (what does this mean?)
-% Formatting
-    % incorporate into template
-% Intro
-    % add month to monthly view stats
-% Method
-    % what to cite for drainage basin in geomorphology
-
-% continue on Page 3
-
-\section{Introduction}
-
-Wikipedia is a towering achievement of the Internet age. 
-At no point in history has a larger or more meticulously indexed collection of human knowledge 
-existed.
-Wikipedia contains 37 million articles in 283 languages, 
-with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
-\cite{drugs, stats}.
-Demonstrating its relevance to modern inquiry,
-Wikipedia is the sixth most visited
-site in the world, surpassing 18 billion page views and 11 million edits in a single month.
-\cite{wiki_edits, wiki_views}.
-
-Wikipedia has naturally become the object of many studies. 
-Researchers have examined the cultural dynamics among editors
-\cite{editors},
-the accuracy of the content relative to traditional encyclopedias
-\cite{accuracy1, accuracy2},
-the topics covered 
-\cite{coverage},
-and bias against portions of the population
-\cite{bias_women}.
-Wikipedia's content has also proven to be a powerful tool. 
-Researchers have used Wikipedia to identify missing dictionary entries
-\cite{missing_entries},
-cluster short text
-\cite{clustering},
-compute semantic relatedness
-\cite{semantic_relatedness},
-and disambiguate meaning \cite{disambiguating}.
-
-While these many studies have dissected and fruitfully applied Wikipedia's content,
-here we examine the connections among the many articles.
-A hyperlink from one Wikipedia article to another
-naturally indicates a relationship between the two articles
-\cite{relevance}.
-The notion that hyperlinks convey information about the content of a
-page has proved enormously successful in multiple domains from search engine algorithms 
-such as PageRank 
-\cite{pagerank} 
-to topic classification
-\cite{classifier}.
-Here, we treat a hyperlink as a mechanism connecting two topics.
-
-The authors of a Wikipedia article choose where and whether to include a 
-reference to another Wikipedia article
-in the HTML markup.
-For example, as of November 2014, the authors of the 
-\href{https://en.wikipedia.org/wiki/Train}{``Train''}
-article
-had collectively chosen
-``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
-among others as relevant articles to reference in describing ``Train''
-\cite{wiki_train}
-.
-
-\begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
-  \caption{
-    \textbf{First Link Path For ``Train.''}
-    We follow the first link  to another Wikipedia article
-    in the main body of the article---the area
-    inside the green rectangle, which excludes 
-    side bar elements, the navigation bar and title; the first link is circled in red. 
-    In this example, the first link to another Wikipedia article is ``Rail Transport.'' 
-    We can again select the first link on the ``Rail Transport''
-    article, repeating the process to 
-    form a path of first links. After 11 links, we arrive at ``Philosophy.''
-  }
-  \label{fig:Train First Links}
-\end{figure*}
-
-By focusing our attention on the main body of an article---excluding
-elements in side bars and headings---we 
-attempt to systematically capture the core description of a topic.
-Within the body text, the first link marks the earliest moment in a topic's introduction where the 
-authors choose to directly reference another article. 
-While many links reference relevant details, 
-the first link is an association within a topic's initial 
-description. While 
-``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
-are links detailing particulars, the first link, ``rail transport,''
-is the topic the authors associate with ``Train'' in the introduction
-(Fig.~\ref{fig:Train First Links}).
-``Banana'' has a first link to ``fruit,'' ``Bob Dylan'' has a first link to 
-``Blowing in the Wind,'' and ``Physics'' has a first link to ``natural science.''
-Collectively, first links provide a pragmatic and interpretable 
-means to connect each article to another.
-
-By following the first hyperlink contained in all articles of the English edition of
-Wikipedia as of November, 2014, we form a directed network:
-{\it Wikipedia's First Link Network} (FLN).
-Inspired by the claim that the majority of first links lead to
-``Philosophy''---popularized by an 
-\href{https://xkcd.com/903/}{xkcd}
-comic and subsequently
-discussed in blog posts 
-\cite{xkcd, mat_blog, Ilmari_first_links}
----
-we holistically study how the FLN  yields a flow of connections.
-For methodological details, see Sec.~\ref{Appendix}.
-
-The FLN is a wealth of relations among inventions, places,
-figures, objects, and events across space and time.
-``Train'' for example, links to a parent node, ``Rail Transport,''
-while many child nodes such as ``Steel'' and ``Horsepower'' link to ``Train.''
-As shown in Fig.~\ref{fig:Train First Links}, the path starting at ``Train'' 
-contains ``Goods,'' ``Economics,'' ``Social Science,''
-leading ultimately to ``Philosophy''.
-Unlike previous taxonomies created by individuals
-\cite{locke, descartes, aristotle, hist_thesaurus},
-the relations in the FLN emerge without a centralized effort 
-as the aggregate of each article's authors choice of first link.
-
-Our goal is to study the structure of the FLN for insight into how the information on Wikipedia is organized and related.
-Informed by river network metrics,
-we consider the FLN as a flow, quantifying 
-the accumulation of first links around articles 
-and the influence an article exerts in shaping the FLN.
-Together with cycles, in-degree, depth, and the content of the articles, 
-we build our analysis of the relations among the ideas in Wikipedia.
-
-\section{Traversing the First Link Network}
-
-An essential feature of a directed network's structure is the degree distribution 
-\cite{newman}. 
-The degree distribution has been used to study many phenomena from disease outbreak 
-\cite{disease} 
-to the dynamics of social networks 
-\cite{social_nets}.
-The in-degree distribution in the FLN describes how many first links point to a 
-particular article. 
-Articles with zero in-degree have no references---they are outer leaves in the FLN 
-or are disconnected entirely.
-
-\begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/traversal_visit_algo_figure.pdf}  
-  \caption{
-    \textbf{Traversal Visit Algorithm on a sample network.}
-     The traversal visit vectors are an adjacency matrix for the paths through the network: 
-     the first column indicates the path formed starting with article A. 
-     The number of traversal visits for article A is then the number of paths containing A 
-     or the sum of the first row in our matrix:
-     $\sum_{i=1}^7 A_{\text{visit, i}} = 7$.
-  }
-  \label{fig:Traversal Visits}
-\end{figure*}
-
-Starting at each article, we construct a path through the FLN, 
-to map the flow of connections among articles.
-The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
-Previous studies have used flow to characterize the structure of river networks
-\cite{geo_basins, dodds} and describe the organization of 
-blood networks, food systems, and transportation networks
-\cite{food_webs}.
-
-The first metric we develop quantifies the accumulation of first links.
-The algorithm begins by selecting an article, then traversing the path formed
-by following the first links. Each time a first link references an article, we increment a count
-associated with the article. 
-We continue until a page is revisited or is invalid---defined here to mean pointing to a page outside of Wikipedia.
-We select a second article and repeat the process until we have 
-constructed a path for each article in the network. 
-We define the number of {\it traversal visits} of an article 
-to be the number of references flowing to the ideas in the article---equivalent
-to drainage basin area in geomorphology.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/traversal_funnel_algo_figure.pdf}
-  \caption{
-    \textbf{Traversal Funnel Algorithm on a simple network.}
-  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
-We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.
-For example
-  the number of traversal funnels for article $E$ is 
-  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
-  \label{fig:Traversal Funnels}
-
-\end{figure}
-
-We can characterize the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
-(Fig.~\ref{fig:Traversal Visits}), the path starting at article A is the 
-first column in the traversal visits matrix. 
-An entry of $1$ indicates the path contains a given article and 
-$0$ indicates the path does not.
-To compute the number of traversal visits for an article, we sum the corresponding row in our matrix.
-The traversal visits matrix for 
-our Wikipedia dataset consists of 121 million entries encoding each path out of the more than 
-googol possible paths ($4.7*10^6*2^{4.7*10^6}$) through the FLN .
-
-By measuring the number of first links between
-two articles, we obtain an additional piece of information we call 
-{\it path length.} We can compute the length of a path by summing 
-along a column in our traversal visits matrix. 
-In our Wikipedia dataset, the sum of all path lengths is 232 million first links.
-The path length describes how closely related topics are.
-Although ``Train'' is related to ``Economics'' for example, there are several articles
-bridging the connection:
-``Train'' is more specifically related to transportation, whose object
-is often goods. Goods are one of the fundamental objects of study in Economics.
-Described in links, this relationship is captured by a relatedness of $4$ 
-first links ultimately connecting ``Train'' to ``Economics.''
-
-One possible path through the FLN is a {\it cycle} or a group of articles 
-linking to one another inside a loop. In our sample network 
-(Fig.~\ref{fig:Traversal Visits}) 
-a 3-cycle exists among nodes A, B, and C. 
-We can readily identify
-the types of cycle structures within the FLN
-and rank each by 
-the number of references directed towards a cycle.
-
-We can also form and rank {\it basins} in the network by identifying groups of 
-path-connected articles, not necessarily forming a perfect cycle.
-A basin connects a group of articles and identifies the paths 
-to a particular article.
-
-While traversal visits measure accumulation, each article's first link also 
-influences the shape of the FLN. 
-At a point of accumulation, a single article's first link 
-can exert great influence over the shape of the FLN by directing many
-references on a particular path. To distinguish between an article 
-that simply happened to fall within a cycle from an article funneling 
-many first links, we develop a second metric called {\it traversal funnels}.
-
-To measure traversal funnels, we traverse the FLN in the same manner as we 
-did for traversal visits, but end a path once we enter a cycle.
-We are then able to distinguish between an article related to many other ideas
-only by virtue of its place in a cycle, from an article exerting influence over where the first links flow. 
-An article with a large number of traversal funnels directs many references
-into a particular cycle. In our sample network 
-(see Fig.~\ref{fig:Traversal Funnels}) article C 
-directs the flow of links towards the 3-cycle, while articles A and B are 
-recipients of the flow---without exerting direct influence themselves.
-
-By studying the FLN not only as collection of directly linked pairs of articles, but
-as a flow, 
-we build a powerful arsenal of information.
-From accumulated references, cycles, and basins, we can measure how the many articles in Wikipedia are organized and related.
-
-\section{Results}
-
-\subsection{Degree Distribution}
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
-  \caption{
-    \textbf{Highest Ranking Articles by in-degree.}
-    We rank each article by the number of direct first links to the article (in-degree).
-The highest-ranking articles tend to represent geographical and biological abstractions.
-A full online appendix of the results and data is available \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.}
-  \label{fig:indegree list}
-\end{figure}
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
-  \caption{
-    \textbf{FLN Degree Distribution.}
-We fit the in-degree distribution with a linear $\log_{10}-\log_{10}$ model. 
-The result is a 
-an excellent fit with a Pearson's Correlation Coefficient of $-0.98$, yielding a 
-power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$. 
-}
-  \label{fig:degree distribution}
-\end{figure}
-
-We rank all 11 million articles by in-degree to find 
-the ``United States'' with 80,249 direct first links as the most referenced
-Wikipedia article 
-(Fig.~\ref{fig:indegree list}). 
-Other high-ranking articles
-include foundational abstract concepts such as ``village,'' ``species,''; 
-sports associations such as ``American Football,'' ``Association Football''; 
-and developed nations such as ``France,'' ``Japan,'' ``Russia,'' ``Australia,'' and 
-the ``Netherlands.'' These high ranking articles are useful abstractions: nations
-describe a collection of individuals with a common culture, language, or 
-geographical proximity; sports teams describe an ever changing collection of 
-players often associated with a cultural identity or a geographical 
-region. 
-Since abstractions such as nations and teams are inherently comprised
-of many parts, authors reference the abstraction when describing a part.
-In describing the New England Patriots or the New York Giants
-for example, ``American Football'' is the natural abstraction, which 
-anchors many specific teams.
-
-``Philosophy'' and other philosophical concepts
-are not among the highest-ranking articles by in-degree.
-``Philosophy'' has an in-degree of only 581, with direct first links from articles about Philosophers and areas of Philosophy: ``Existentialism and Humanism,'' ``Predeterminism,'' ''Synoptic Philosophy,'' ``Qualia,'' ``Dorothy Emmet,'' and ``Christopher W.
-Morris.''
-While many articles accumulate at ``Philosophy'' (see traversal visits discussion below), 
-the accumulation is not the 
-result of many articles directly referencing ``Philosophy.'' 
-Instead, first links flow towards ``Philosophy'' as the 
-ultimate anchor, by generalizing from specific to broad.
-
-The FLN's in-degree exhibits a decaying power law distribution where a few articles 
-receive most direct first references, while most articles receive few or none.
-The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5.
-Less than $1\%$ of articles have more than 100 direct first links and $75\%$ of articles
-have fewer than 9. 
-We fit the in-degree distribution against each article's rank with a log-log (base 10) linear regression model
-(see Fig.~\ref{fig:degree distribution}).
-We obtain a Pearson's Correlation Coefficient of $-0.98$
-with a corresponding  power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$.
-
-\subsection{Depth of the FLN}
-
-How many links does a connection among articles span? 
-We find the longest path length is 365,
-corresponding to the yearly calendar of Orthodox Liturgics.
-Each day's Liturgics links to the next day's. On the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
-(see discussion of cycles ~\ref{cycles}).
-We also find similarly lengthy paths following the evolution of a place or topic through time: 
-``1953 in Scotland'' or ``1560s Architecture,'' with articles sequentially proceeding by year, decade or era.
-In general, the longest paths connect temporally organized ideas.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/path_lengths_dist.png}
-  \caption{
-    \textbf{Path Length Distribution.}
-The network depth appears to have a bimodal distribution with a median of 29 first links in a path.
-The 365 cycle for ``Orthodox'' liturgics is the outlier to the right, with historical articles 
-describing nations such as ``Scotland'' and ``UK'' also lying outside the third quartile.
-More than $75\%$ of articles have path lengths between 
-$0$ and $50$ links.}
-  \label{fig:Path Length Distribution}
-\end{figure}
-
-Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
-  \caption{
-    \textbf{Distribution of Traversal Visits.}
-We fit the distribution of traversal visits against article rank with a log-log (base 10) linear regression model log-log model and uncover two regimes: 
-with a cutoff rank of $10^{5}$.
-The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
-Coefficient of $-0.99$ 
-with a corresponding power law rank exponent 
-$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
-The horizontal flattening around the highest
-ranking articles is a result of the cyclic structure (see discussion on cycles).
-The bottom regime has a Pearson's Correlation Coefficient of $-0.93$
-with a corresponding power law rank exponent 
-$\alpha = -0.579$ and a size exponent $\gamma = -0.727$.}
-  \label{fig:Distribution of Visits}
-
-\end{figure}
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/articles_ranked.png}
-  \caption{
-    \textbf{Highest ranking articles by number of traversal visits.}
-We compute the number of traversal visits for each article in the FLN (see 
-Traversal Algorithm section for details). In doing so, we can rank each article
-by the accumulation of first links.
-The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.}
-  \label{fig:highest visits}
-\end{figure}
-
-This roughly corresponds to the official number of articles on Wikipedia: 
-$~4.7$ million as of November 2014---approximately half of the 11 million 
-articles in the XML dump are redirects or disambiguations, not full articles.
-The most common path length is 29, with an interquartile range of 4---path length
-is typically between 26 and 29 articles.
-As a distribution, more than $75\%$ of articles have a path length below 
-$50$ first links 
-while a few temporally organized paths exceed 50 links 
-(see Fig.~\ref{fig:Path Length Distribution}).
-
-\subsection{Traversal Visits}
-
-As a distribution, the number of traversal visits by article appears to follow a decaying power law. 
-The majority of articles have fewer than 30 traversal visits and
-first link references accumulate at a few articles.
-Specifically, $99.76\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
-Meanwhile, the highest ranking 30 articles have an extremely disproportionate number of traversal visits.
-
-\begin{figure*}[tph!]
-  \includegraphics[width=\textwidth]{graphics/2_and_3_cycles.pdf}
-  \caption{
-    \textbf{Highest ranking 2-cycles and 3-cycles.}
-We identify pairs of articles whose first links point to one another, forming
-a 2-cycle. We then rank each pair of articles by the total number of 
-traversal visits to gauge the most referenced groups of two articles linked
-to each other. We find 2-cycles often capture synonyms or articles representing nearly the 
-same concepts as opposed to distinct concepts. Similarly, we identify and 
-rank 3-cycles to find they appear to capture three closely related ideas or synonyms.}
-  \label{fig:2-cycles}
-\end{figure*}
-In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against rank
-to two regimes corresponding to power law rank exponents of 
-$\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
-On the aggregate, the distribution suggests
-a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none.
-The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.
-
-The highest ranking articles by traversal visits are broad topics spanning academic disciplines 
-or notions fundamental to society~(Fig. \ref{fig:highest visits}): ``Science,'' ``Mathematics,'' `` Geography,'' and ``Philosophy''
-as well as ``Community,'' ``State,'' ``Earth,'' ``Information,'' ``Power,'' and ``Communication.''
-The first links flow towards broader topics, accumulating at of these foundational notions.
-While ``Banana'' is a concrete fruit for example, 
-the first links flow from ``Fruit'' to ``Botany'' to ``Biology,'' and ultimately 
-culminate with ``Science'' and ``Philosophy.'' 
-The first links anchor the wealth of specific knowledge on Wikipedia to 
-a few notions foundational to society.
-
-\subsection{Network Cycles}
-\label{cycles}
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
-  \caption{
-    \textbf{Funnels.}
-We represent the highest-ranking articles by the number of traversal 
-funnels to gauge the influence each article exerts in shaping the 
-structure of the FLN. 
-We find ``Philosophy'' exerts an overwhelming proportion
-of the influence, with other abstract notions and topical concepts ranking
-next.}
-  \label{fig:Funnels}
-\end{figure}
-
-We identify 2-cycles, meaning a pair of articles with first link pointing to one another.
-Of the 11 million articles, roughly 84,000 are memebers of 2-cycles. 
-The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
-``Health Care'' and ``Medical Case Management,'' ``Broadcasting House,'' and ``BBC,'' ``Secondary Education'' and ``Secondary School'' 
-(see Fig.~\ref{fig:2-cycles}).
-
-Outside of the highest ranking 2-cycles, the typical 2-cycle signals a connection between distinct, yet very closely related concepts. 
-We also observe link patterns such as inventor to product (``Voere'' to ``VEC-91''), event to organizer (``Poetry Bus Tour'' to ``Weave Books''), and book to author (``Anatomy of Britain'' to ``Anthony Sampson'').
-
-Similarly, 3-cycles capture a synonymous or close relation among 3 articles: ``Tree of life (Biology),'' ``Tree of life (disambiguation),'' 
-and ``Tree of life''; ``Cinema of India,'' ``Indian Cinema,'' and ``Telugu Cinema''
-(see Fig.~\ref{fig:3-cycles}).
-Once we extend our cycle size beyond a length of 6 however, 
-``Philosophy'' along with the remaining list of high ranking articles by traversal visits dominate.
-The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day.
-Other lengthy cycles span $60-75$ articles including collections of articles on national histories such as ``Japanese Eras'' 
-or judicial bodies such as the ``Legislative Assembly of Ontario.''
-
-\subsection{Basins}
-
-We can group articles lying on the same path to identify {\it basins}: 
-a group of path-connected articles.
-Since cycles identify only groups of articles with a closed set of links, 
-we additionally measure and rank basins to capture groups of closely related
-articles branching outside of a cycle into the rest of the FLN.
-We rank basins by the total number of traversal visits for each article in the path. 
-Akin to river networks, these basins are areas of accumulation with a path 
-flowing outwards to the rest of the FLN.
-
-The highest ranking basins by the number of traversal visits are groups of articles
-around ``Philosophy.'' 
-The highest ranking paths include branches of philosophy flowing through 
-``Awareness,'' ``Existence,'' and ``Consciousness'' to ``Philosophy.'' Other paths
-include concepts around ``Mathematics,'' ``Scientific,'' ``Experiments,'' 
-``Biology,'' and ``Fact.''
-These paths link many specific articles to ``Philosophy,'' each funneled through a particular domain.
-
-Moving beyon basins around ``Philosophy,'' we find other basins around 
-foundational concepts such as ``Community,'' ``Landmass,'' ``Federal Government,'' 
-``Presentation,'' and ``Belief System.'' 
-The basins around each of these foundational notions are 
-various paths containing related articles. For example around 
-``Community'' we find basins
-flowing from ``United States'' to ``Federal Republic'' to ``Political Union'' to ``State'' culminating at ``Community''; we also find basins flowing from 
-``Public Policy'' to ``Executive (government)'' to ``Government'' to ``State'' and then 
-to ``Community''; we also find paths flowing through a similar chain beginning
-with ``Democracy,'' another beginning with ``Constitution,'' another at 
-``Dictatorship,'' and so on. The articles build from specific means of organizing
-a community (or society) and then build up to ``Community.'' 
-Other basins around landmass for example begin at specific geographical regions
-such as ``Eastern Europe'' building up to ``Continent'' and finally ``Landmass''.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
-  \caption{
-    \textbf{Distribution of Traversal Funnels.}
-We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed number of traversal funnels. 
-We find two regimes with the top regime (log(rank) < 4) 
-well-explained by a linear fit, yielding Pearson's 
-Correlation Coefficient of $-0.99$ and a 
-corresponding power law rank exponent $\alpha = -1.08$
-and a size exponent $\gamma = 0.074$. 
-The bottom horizontal regime corresponds to the more than $99\%$ of articles
-which hold zero traversal funnels.}
-  \label{fig:Funnels Distribution}
-\end{figure}
-
-\subsection{Traversal Funnels}
-
-To analyze the influence an article exerts in shaping the 
-structure of the FLN, we compute the number of traversal funnels for each.
-Articles directing more paths exert a greater influence over the structure
-of the FLN by increasing the accumulation of first links
-on a particular path. By measuring traversal funnels, we distinguish between an article that simply happened to fall within a cycle from an article funneling 
-many first links.
-
-Ranking articles by the number of traversal funnels we find 
-``Philosophy'' to be by far the highest-ranking article with 
-$7.37$ million paths
-(see Fig.~\ref{fig:Funnels}).
-Of any article, the number of traversal funnels Philosophy holds exceeds 
-all others by more than two orders of magnitude.
-The ``Philosophy'' cycle which contains ``Existence,'' ``Awareness,'' ``Reality,'' 
-and similar articles accumulates the overwhelming proportion of its 
-references through ``Philosophy'': $7.37$ million of the $7.4$ million references
-are funneled through ``Philosophy''.
-Second on the list of highest-ranking articles by traversal funnels is 
-``Presentation'' with only $30$ thousand paths. Similarly abstract 
-concepts also rank highly such as ``Tree of life'' (30 thousand), 
-``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).
-
-Many high-ranking articles are remarkably topical, culturally and politically important concepts.
-For example, ``Health Care,'' a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August, 2009 and February, 2010.
-Other high ranking articles include key historical events such as the ``Cold War'' or critical scarce resource with recent 
-media discussion such as ``Fossil Fuel.'' 
-The highest-ranking list also includes ``Hip Hop,'' ``Cancer,'' and ``Web Page.''
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/popularity_visits.pdf}
-  \caption{
-\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal visits.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal visits.
-While the most popular articles do not necessarily correspond to the articles
-with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-The greatest number of articles fall within roughly 1 million traversal visits and page views.
-Overall, the page views for the top 1000 articles articles by traversal visits appear to be log-normally distributed.
-}
-  \label{fig:Views and Visits}
-
-\end{figure}
-As a distribution, we find few articles influence the structure of the 
-FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
-more than $99\%$ with none---most articles are recipients of 
-the references flowing through the articles with at least one traversal funnel.
-When fit to a log-log linear model we find the $99\%$ of articles with zero
-traversal funnels form one regime (with log(rank) less than 4).
-The top regime, corresponding to the $17, 821$ articles with at least one 
-traversal funnel strongly fits a linear model with
-a Pearson's Correlation Coefficient of $-0.99$ and a 
-corresponding power law rank exponent $\alpha = -1.08$
-and a size exponent $\gamma = 0.074$ 
-(see Fig.~\ref{fig:Funnels Distribution}). Even within the few articles
-influencing the structure of the FLN, only a handful of these exert most of the 
-control.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/popularity_funnels.pdf}
-  \caption{
-\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal funnels.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal funnels.
-}
-  \label{fig:Views and Funnels}
-\end{figure}
-\subsection{Article Popularity}
-
-Wikipedia released an API to measure article popularity by page views
-starting November 2015. 
-Page views add another dimension to our
-findings by constrasting
-the number of users who access a particular article against
-the number of traversal visits and funnels.
-We measure popularity as the total number 
-of page views in the English edition of Wikipedia in the month of 
-October 2015---the earliest full month for which the data is available. 
-We find the highest ranking 1000 articles (by traversal visits) have an average of
-$70$ thousand page views in October with high variation: the standard deviation 
-is $1.1$ million page views. 
-The number of page views has a skewed distribution with $75\%$ of articles
-reaching fewer than $73$ thousand page views in October.
-The article for ``United States'' has the most page 
-views of the highest ranking $1000$ articles by traversal visits with 
-$1.4$ million views in October. The next most popular articles are 
-``India,'' ``World War II,'' and the ``United Kingdom'' each with roughly 1 million page views. 
-The ``Philosophy'' article, despite outranking every article by traversal visits,
-has only $240$ thousand page views in October.
-
-We also analyze article popularity for the highest ranking 1000 articles by 
-traversal funnels. In October, the average page views per article is 
-$22$ thousand with a standard deviation of $95$ thousand views. The distribution
-is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
-``Halloween'' is the most popular article with $2.8$ million views in October,
-likely a result of Halloween falling in the month of October. 
-Other popular articles include ``Scientology'' ($463$ thousand views), ``Clint Eastwood'' 
-($341$ thousand views),
-and the ``Cold War'' ($298$ thousand views) although each has significantly fewer views compared to the views for ``Halloween''.
-Other standout popular October articles include ``24-hour'' and ``12-hour'' clocks likely due
-to Daylight savings in October and ``Marriage'' possibly due to the drop in the number of
-weddings in the months following October
-\cite{weddings}.
-``Philosophy'' which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears nowhere in the top 20
-by traversal visits. ``Philosophy'' is a relatively popular article 
-among articles influencing the shape of the network, but less popular among 
-highly-ranked articles by accumulation.
-
-%========================Conclusion====================================
-
-\section{Concluding Remarks}
-
-The findings here should only be considered within the limitations of their context.
-We examined only the English edition of Wikipedia at a particular moment in time, 
-considering only the first link in in each article as a means to relate articles.
-Finally, Wikipedia, while the largest digital 
-collection of human knowledge, is rife with the biases of the many contributing editors
-\cite{bias}.
-Nevertheless, the findings do reveal
-generalizable relationships and point to foundational notions.
-
-Among our observations is the appearance of multiple scale-free distributions within the network. 
-Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
-articles are responsible for funneling most paths. When measured against the traversal funnels, 
-``Philosophy'' emerges as an exceptional article by orders of magnitude. 
-Nevertheless, many other foundational concepts emerged naturally within FLN. 
-Basins around ``Community'', ``State'', and ``Science'' reveal a foundational structure within the network. 
-More curious is the emergence of recently prominent political and economics topics such as ``Fossil Fuel'' and ``Health Care'' 
-within the highest ranking funnels. 
-Wikipedia seems to reflect not only timeless foundations, but also the topical (at least within English speaking society).
-
-Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the first link.
-These findings also form the basis for the creation of a taxonomy where 
-every idea, event, or object sits within a hierarchy of connected notions.
-The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
-Applications could range from an enhanced network of ideas to psychological insights into how humans form associations.
-Specifically, an ever-evolving reference of related hierarchical concepts could be 
-used to improve search engine algorithms or natural language processing.
-
-\acknowledgments
 MI is grateful to RJ for pointing out the Reddit post 
@@
 
-PSD and CMD acknowledge support from NSF Big Data Grand $\#1447634$.
+PSD and CMD acknowledge support from NSF Big Data Grant $\#1447634$.
 
 The authors also acknowledge support from the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
-
-\newpage
-
-\section{Appendix}
-\label{Appendix}
-
-\subsection{Constructing The First Link Network}
-
-To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
-Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
-eliminating any statistical error due to sampling.
-We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
-The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
-and disambiguations.
-Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
-\cite{wiki_edits},
-our aim is to characterize the structure of the First Link Network.
-
-Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
-Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
-information.
-While a human could manually identify the first link, to map the entire First Link network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
-
-While some libraries exist for MediaWiki Markup,
-approaches using existing libraries led to several bugs 
-including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
-as well as compatibility with other libraries used to parse the XML.
-Consequently,  we developed an algorithm for parsing the first link in the XML version of each article.
-Our parsing algorithm aimed to: 
-1) accurately identify the first link among other page elements, and 
-2) efficiently do so---that is without needing for several passes through the data.
-To process an article in a single pass, we developed a hierarchical system of flags:
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/flags.pdf}  
-  \caption{
-    \textbf{Parsing Algorithm of Wikipedia's XML dump.}
-The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
-Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
-Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.}
-  \label{fig:parsing algorithm}
-\end{figure}
-
-The algorithm loops in three-character chunks to account for potentially nested elements, 
-shifting by one character steps through the article markup.
-If any markup triggers for a flag are detected, a flag is raised. 
-Once a flag is raised, we stop processing and proceed to the next character
-until the flag's closing markup.
-A first link is identified only if Flags 1, 2, and 3 are all off.
-In this case, the entire link is retrieved. 
-We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
-as well as common file extensions for 
-images, audio files, and the like 
-\cite{media_wiki_templates}.
-The first link of an article is then the earliest valid link with unraised flags.
-
-To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
-across 112 cores of the UVM supercomputer cluster
-\cite{vacc}.
-We then joined the results to form a hash table containing every Wikipedia article and its corresponding
-first link. The resulting network map is the basis of our analysis.
-An online appendix containing all of the code and data used can be accessed \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.
-
-\begin{thebibliography}{1}
-
-      \bibitem{editors} Iba, Takashi, et al.
-``Analyzing the creative editing behavior of Wikipedia editors: Through dynamic social network analysis.'' Procedia-Social and Behavioral Sciences 2.4 (2010): 6441-6456.
-
-      \bibitem{accuracy1} Holman Rector, Lucy. ``Comparison of Wikipedia and other encyclopedias for accuracy, breadth, and depth in historical articles.'' Reference services review 36.1 (2008): 7-22.
-
-      \bibitem{accuracy2} Giles, Jim. ``Internet encyclopaedias go head to head.'' Nature 438.7070 (2005): 900-901.
-
-       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. ``An analysis of topical coverage of Wikipedia.'' Journal of Computer Mediated Communication 13.2 (2008): 429-440.
-
-        \bibitem{bias_women} Hill, Benjamin Mako, and Aaron Shaw.
-``The Wikipedia gender gap revisited: characterizing survey response bias with propensity score estimation.'' PloS one 8.6 (2013): e65782.
-
-      \bibitem{missing_entries} Williams, Jake Ryland, Clark, Eric M., Bagrow, James P., Danforth, Christopher M.
-\& Dodds, Peter Sheridan (2015).
-Identifying missing dictionary entries with frequency-conserving context models.
-Phys.
-Rev.
-E, 92, 042808.
-
-      \bibitem{clustering} Banerjee, Somnath, Krishnan Ramanathan, and Ajay Gupta.
-``Clustering short texts using wikipedia.'' Proceedings of the 30th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2007.
-
-      \bibitem{semantic_relatedness} Gabrilovich, Evgeniy, and Shaul Markovitch. ``Computing Semantic Relatedness Using Wikipedia-based Explicit Semantic Analysis.'' IJCAI. Vol. 7. 2007.
-
-      \bibitem{disambiguating} Cucerzan, Silviu. ``Large-Scale Named Entity Disambiguation Based on Wikipedia Data.'' EMNLP-CoNLL. Vol. 7. 2007.
-
-      \bibitem{drugs} Clauson, Kevin A., et al. ``Scope, completeness, and accuracy of drug information in Wikipedia.'' Annals of Pharmacotherapy 42.12 (2008): 1814-1821.
-
-      \bibitem{stats} Wikipedia contributors. ``Wikipedia:Statistics.'' Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 9 Nov. 2015. Web. 13 Nov. 2015.
-
-      \bibitem{pagerank} Page, Lawrence, et al. ``The PageRank citation ranking: bringing order to the Web.'' (1999).
-
-      \bibitem{classifier} Chakrabarti, Soumen, Mukul Joshi, and Vivek Tawde.
-``Enhanced topic distillation using text, markup tags, and hyperlinks.'' Proceedings of the 24th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2001.
-
-      \bibitem{relevance} Kamps, Jaap, and Marijn Koolen. ``Is Wikipedia link structure different?.'' Proceedings of the second ACM international conference on Web search and data mining. ACM, 2009.
-
-      \bibitem{wiki_train} Wikipedia contributors. ``Train.'' Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 
-          \url{https://en.wikipedia.org/wiki/Train}
-
-    \bibitem{locke} Bolton, Martha Brandt.
-``The Taxonomy of Ideas in Locke\textquotesingle s {\it Essay},'' The Cambridge Companion to Locke's ``Essay Concerning Human Understanding.'' Ed.Lex Newman..
-1st ed.
-Cambridge: Cambridge University Press, 2007.
-67-100.
-Cambridge Companions Online.
-Web.
-13 November 2015.
-http://dx.doi.org/10.1017/CCOL0521834333.004.
-
-    \bibitem{aristotle} Studtmann, Paul, ``Aristotle's Categories,'' The Stanford Encyclopedia of Philosophy (Summer 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/sum2014/entries/aristotle-categories/}
-
-    \bibitem{descartes} Smith, Kurt, ``Descartes' Theory of Ideas,'' The Stanford Encyclopedia of Philosophy (Spring 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/spr2014/entries/descartes-ideas/}
-
-    \bibitem{hist_thesaurus} What is the Historical Thesaurus of the OED - Oxford English Dictionary (Oxford English Dictionary)
-    http://public.oed.com/historical-thesaurus-of-the-oed/what-is-the-historical-thesaurus-of-the-oed/
-
-    \bibitem{vacc}  Vermont Advanced Computing Core, \url{http://www.uvm.edu/vacc}.
-
-    \bibitem{wiki_views} Page Views for Wikipedia, Non-mobile site, Normalized (Page Views for Wikipedia, Non-mobile site, Normalized)
-    \url{https://stats.wikimedia.org/EN/TablesPageViewsMonthly.htm}
-
-    \bibitem{weddings} Wedding statistics in the United States (Wedding statistics in the United States)
-    \url{http://www.soundvision.com/article/wedding-statistics-in-the-united-states}
-
-    \bibitem{six_degrees} Elmacioglu, Ergin, and Dongwon Lee. ``On six degrees of separation in DBLP-DB and more.'' ACM SIGMOD Record 34.2 (2005): 33-40.
-
-    \bibitem{wordnet} Princeton University ``About WordNet.'' WordNet. Princeton University. 2010. 
-    \url{http://wordnet.princeton.edu}
-
-    \bibitem{newman} Newman, Mark EJ. ``The structure and function of complex networks.'' SIAM review 45.2 (2003): 167-256.
-
-    \bibitem{disease} Eubank, Stephen, et al. ``Modelling disease outbreaks in realistic urban social networks.'' Nature 429.6988 (2004): 180-184.
-
-    \bibitem{social_nets} Newman, Mark EJ, Duncan J.
-Watts, and Steven H.
-Strogatz.
-``Random graph models of social networks.'' Proceedings of the National Academy of Sciences 99.suppl 1 (2002): 2566-2572.
-
-    \bibitem{geo_basins} Horton, Robert E.
-``Erosional development of streams and their drainage basins; hydrophysical approach to quantitative morphology.'' Geological society of America bulletin 56.3 (1945): 275-370.
-    APA
-
-    \bibitem{dodds} Dodds, Peter Sheridan, and Daniel H. Rothman. ``Unified view of scaling laws for river networks.'' Physical Review E 59.5 (1999): 4865.
-
-    \bibitem{food_webs} Garlaschelli, Diego, Guido Caldarelli, and Luciano Pietronero. ``Universal scaling relations in food webs.'' Nature 423.6936 (2003): 165-168.
-
-    \bibitem{mat_blog} brain of mat kelcey (brain of mat kelcey). 
-    ``Do all first links lead to Philosophy?''
-    \url{http://matpalm.com/blog/2011/08/13/wikipedia-philosophy/}
-
-    \bibitem{Ilmari_first_links} User:Ilmari Karonen/First link (Wikipedia)
-    \url{https://en.wikipedia.org/wiki/User:Ilmari_Karonen/First_link}
-
-    \bibitem{xkcd} xkcd: Extended Mind (xkcd: Extended Mind)
-    \url{http://xkcd.com/903/}
-
-    \bibitem{reddit} Almost every Wikipedia page links to Philosophy if you keep clicking the first link in every article. 
-    \url{https://www.reddit.com/r/InternetIsBeautiful/comments/35asgt/almost_every_wikipedia_page_links_to_philosophy/}
-
-    \bibitem{bias} Wagner, Claudia, et al. ``It's a Man's Wikipedia? Assessing Gender Inequality in an Online Encyclopedia.'' arXiv preprint arXiv:1501.06307 (2015).
-
-    \bibitem{wiki_edits} Wikipedia:Statistics 
-    \url{https://en.wikipedia.org/wiki/Wikipedia:Statistics}
-
-    \bibitem{media_wiki_templates} MediaWiki: ``Help: Templates''
-    \url{https://www.mediawiki.org/wiki/Help:Templates}
-
-    \bibitem{links_that_speak} Ronen, Shahar, Bruno Gonçalves, Kevin Z.
-Hu, Alessandro Vespignani, Steven Pinker, and César A.
-Hidalgo.
-``'Links That Speak: The Global Language Network and Its Association with Global Fame.'' Proceedings of the National Academy of Sciences Proc Natl Acad Sci USA 111.52 (2014).
-
-\end{thebibliography}
-
-\end{document}

### 2016-04-22 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
universal template and bibtex format changed

--- 2016-04-12 writeup/wikipedia-network.tex
@@
-\documentclass[pre,twocolumn,twoside,superscriptaddress,floatfix, aps, 10pt]{revtex4-1}
-
-\usepackage{currfile}
-\lefthyphenmin=3
-\righthyphenmin=2
-
-\usepackage{graphicx,epsfig,verbatim,enumerate}
-\usepackage{amssymb,amsmath}
-\usepackage{ifthen}
-
-% Fonts and typesetting settings
-\usepackage[sc]{mathpazo}
-\usepackage[T1]{fontenc}
-\linespread{1.05} % Palatino needs more space between lines
-\usepackage{microtype}
-\usepackage{textcomp}
-
-\usepackage{longtable}
-
-\usepackage{mathtools}
-
-\newboolean{twocolswitch}
-
-\newcommand{\www}[1]{\url{#1}}
-\newcommand{\req}[1]{(\ref{#1})}
-\newcommand{\Req}[1]{Eq.~(\ref{#1})}
-
-% Lettrines
-\usepackage{lettrine}
-
-%useful shortcuts
-\def\R{\ensuremath{\mathbb{R}}} %\ensuremath adds math mode, if forgotten
-\def\Q{\ensuremath{\mathbb{Q}}}
-\def\N{\ensuremath{\mathbb{N}}}
-\def\Z{\ensuremath{\mathbb{Z}}}
-\def\C{\ensuremath{\mathbb{C}}}
-
-%shorcuts with arguments
-\newcommand{\abs}[1]{\left\vert#1\right\vert} %nice absolute values
-\newcommand{\bt}[1]{\textbf{#1}} %bold
-\newcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert} %vector norm
-\renewcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-
-%piecewise function
-
-%\begin{displaymath}
-%   f(x) = \left\{
-%     \begin{array}{lr}
-%       1 & : x \in \mathbb{Q}\\
-%       0 & : x \notin \mathbb{Q}
-%     \end{array}
-%   \right.
-%\end{displaymath}
-
-%environment
-\newcommand{\tab}{\phantom{ssss}}
-
-\usepackage{hyperref}
-\usepackage{color}
-\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}
-
-%colors
-\definecolor{javagreen}{rgb}{0.25,0.5,0.35} %dark green color
-\definecolor{lightblue}{rgb}{0.149,0.545,0.824} %solarized blue
-\definecolor{sred}{rgb}{0.863, 0.196, 0.184} %solarized red
-
-\newcommand{\blue}[1]{{\leavevmode\color{lightblue}{#1}}} %solarized blue 
-\newcommand{\green}[1]{{\leavevmode\color{javagreen}{#1}}} %command for green
-\newcommand{\red}[1]{{\leavevmode\color{sred}{#1}}} %solarized red
-\newcommand{\gray}[1]{{\leavevmode\color[gray]{0.5}{#1}}} %gray text
-\newcommand{\Prob}[1]{{\rm Pr}\left(#1\right)}
-
-\setboolean{twocolswitch}{true}
-
-\begin{document}
-
-\title{\protect
-Connecting every bit of knowledge: \\
-The structure of Wikipedia's Flrst Link Network
-}
-
-\date{\today}
-
-\author{
-\firstname{Mark}
-\surname{Ibrahim}
-}
-\email{mark.s.ibrahim@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Christopher}
-\surname{M. Danforth}
-}
-\email{chris.danforth@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Peter}
-\surname{Sheridan Dodds}
-}
-
-\email{peter.dodds@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\begin{abstract}
-  \protect
-Apples, porcupines, and the most obscure Bob Dylan song---is every topic a few clicks from Philosophy? 
-Within Wikipedia, the surprising answer is yes: nearly all 
-paths lead to Philosophy.
-Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
-More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
-By following the first link in each article, we algorithmically construct a directed network of 
-all 4.7 million articles: Wikipedia's First Link Network.
-Here, we study the English edition of Wikipedia's First Link Network for insight into how the many 
-articles on inventions, places, people, objects, and events are related and organized.  
-We traverse every path, measuring the accumulation of first links, path lengths, basins, cycles, 
-and the influence each article exerts in shaping the network.
-We find scale-free distributions describe path length, 
-accumulation, and influence. Far from dispersed, first links disproportionately accumulate 
-at a few articles---flowing from specific to general and culminating around fundamental notions such as
-Community, State, and Science. 
-Philosophy funnels more paths than any other article by two orders of magnitude. 
-We also observe a gravitation towards topical articles such as 
-Health Care and Fossil Fuel. 
-These findings enrich our view of the connections and structure of
-Wikipedia's ever growing store of knowledge.
-
-\end{abstract}
-
-\maketitle
-
-%TODO:
-% Abstract 
-    % In traversing every path, we measure ?
-    % Our findings... braided network (what does this mean?)
-% Formatting
-    % incorporate into template
-% Intro
-    % add month to monthly view stats
-% Method
-    % what to cite for drainage basin in geomorphology
-
-% continue on Page 3
-
 \section{Introduction}
@@
 with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
-\cite{drugs, stats}.
+\cite{clauson2008scope, stats}.
 Demonstrating its relevance to modern inquiry,
 Wikipedia is the sixth most visited
-site in the world, surpassing 18 billion page views and 11 million edits in a single month.
+site in the world, surpassing 18 billion page views and 10 million edits in January, 2013 alone.
 \cite{wiki_edits, wiki_views}.
@@
 Researchers have examined the cultural dynamics among editors
-\cite{editors},
+\cite{iba2010analyzing, samoilenko2016linguistic},
 the accuracy of the content relative to traditional encyclopedias
-\cite{accuracy1, accuracy2},
+\cite{holman2008comparison, giles2005internet},
 the topics covered 
-\cite{coverage},
+\cite{halavais2008analysis},
 and bias against portions of the population
-\cite{bias_women}.
+\cite{hill2013wikipedia}.
 Wikipedia's content has also proven to be a powerful tool. 
 Researchers have used Wikipedia to identify missing dictionary entries
-\cite{missing_entries},
+\cite{williams2015identifying},
 cluster short text
-\cite{clustering},
+\cite{banerjee2007clustering},
 compute semantic relatedness
-\cite{semantic_relatedness},
-and disambiguate meaning \cite{disambiguating}.
+\cite{gabrilovich2007computing},
+and disambiguate meaning \cite{cucerzan2007large}.
 
@@
 naturally indicates a relationship between the two articles
-\cite{relevance}.
+\cite{kamps2009wikipedia}.
 The notion that hyperlinks convey information about the content of a
@@
 such as PageRank 
-\cite{pagerank} 
+\cite{page1999pagerank} 
 to topic classification
-\cite{classifier}.
+\cite{chakrabarti2001integrating}.
 Here, we treat a hyperlink as a mechanism connecting two topics.
@@
 discussed in blog posts 
-\cite{xkcd, mat_blog, Ilmari_first_links}
+\cite{xkcd, mat_blog, Ilmari_first_links, 
+bob_west}
 ---
@@
 Unlike previous taxonomies created by individuals
-\cite{locke, descartes, aristotle, hist_thesaurus},
+\cite{Bolton2007, descartes, aristotle, hist_thesaurus},
 the relations in the FLN emerge without a centralized effort 
@@
 \section{Traversing the First Link Network}
+\label{Traversal Algorithm}
 
 An essential feature of a directed network's structure is the degree distribution 
-\cite{newman}. 
+\cite{newman2003structure}. 
 The degree distribution has been used to study many phenomena from disease outbreak 
-\cite{disease} 
+\cite{eubank2004modelling} 
 to the dynamics of social networks 
-\cite{social_nets}.
+\cite{newman2002random}.
 The in-degree distribution in the FLN describes how many first links point to a 
@@
 Previous studies have used flow to characterize the structure of river networks
-\cite{geo_basins, dodds} and describe the organization of 
+\cite{horton1945erosional, dodds1999unified} and describe the organization of 
 blood networks, food systems, and transportation networks
-\cite{food_webs}.
+\cite{garlaschelli2003universal}.
 
@@
 our Wikipedia dataset consists of 121 million entries encoding each path out of the more than 
-googol possible paths ($4.7*10^6*2^{4.7*10^6}$) through the FLN .
+googol possible paths ($\simeq4.7\times10^6\times2^{4.7\times10^6}$) through the FLN .
 
@@
 two articles, we obtain an additional piece of information we call 
-{\it path length.} We can compute the length of a path by summing 
-along a column in our traversal visits matrix. 
-In our Wikipedia dataset, the sum of all path lengths is 232 million first links.
-The path length describes how closely related topics are.
+{\it path length}, computed by summing 
+columns in the traversal visits matrix. 
+Path length describes how closely related topics are.
 Although ``Train'' is related to ``Economics'' for example, there are several articles
@@
 many first links, we develop a second metric called {\it traversal funnels}.
-
 To measure traversal funnels, we traverse the FLN in the same manner as we 
@@
 only by virtue of its place in a cycle, from an article exerting influence over where the first links flow. 
-An article with a large number of traversal funnels directs many references
-into a particular cycle. In our sample network 
-(see Fig.~\ref{fig:Traversal Funnels}) article C 
+In our simple network 
+(see Fig.~\ref{fig:Traversal Funnels}), article C 
 directs the flow of links towards the 3-cycle, while articles A and B are 
@@
 
-By studying the FLN not only as collection of directly linked pairs of articles, but
+By considering the FLN not only as collection of directly linked pairs of articles, but
 as a flow, 
-we build a powerful arsenal of information.
-From accumulated references, cycles, and basins, we can measure how the many articles in Wikipedia are organized and related.
+With accumulated references, cycles, and basins, we can explore how the many articles in Wikipedia are organized and related.
 
@@
 
-We rank all 11 million articles by in-degree to find 
+We rank all 11 million articles in the FLN by in-degree to find 
 the ``United States'' with 80,249 direct first links as the most referenced
@@
 Since abstractions such as nations and teams are inherently comprised
-of many parts, authors reference the abstraction when describing a part.
-In describing the New England Patriots or the New York Giants
-for example, ``American Football'' is the natural abstraction, which 
-anchors many specific teams.
+of many parts, authors often reference the abstraction when describing a part.
 
@@
 Morris.''
-While many articles accumulate at ``Philosophy'' (see traversal visits discussion below), 
+So while a great many FLN paths flow to 
+``Philosophy'' (see traversal visits discussion below), 
 the accumulation is not the 
@@
 We compute the number of traversal visits for each article in the FLN (see 
-Traversal Algorithm section for details). In doing so, we can rank each article
-by the accumulation of first links.
+Sec.~\ref{Traversal Algorithm}). 
+In doing so, we can rank each article by the accumulation of first links.
 The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.}
@@
 articles in the XML dump are redirects or disambiguations, not full articles.
-The most common path length is 29, with an interquartile range of 4---path length
-is typically between 26 and 29 articles.
+The most common path length is 29, with $75\%$ of path lengths ranging between 7--30 articles.
 As a distribution, more than $75\%$ of articles have a path length below 
@@
 rank 3-cycles to find they appear to capture three closely related ideas or synonyms.}
-  \label{fig:2-cycles}
+  \label{fig:cycles}
 \end{figure*}
@@
 The highest ranking articles by traversal visits are broad topics spanning academic disciplines 
-or notions fundamental to society~(Fig. \ref{fig:highest visits}): ``Science,'' ``Mathematics,'' `` Geography,'' and ``Philosophy''
+or notions fundamental to society~(Fig. \ref{fig:highest visits}): ``Science,'' ``Mathematics,'' ``Geography,'' and ``Philosophy''
 as well as ``Community,'' ``State,'' ``Earth,'' ``Information,'' ``Power,'' and ``Communication.''
@@
 culminate with ``Science'' and ``Philosophy.'' 
-The first links anchor the wealth of specific knowledge on Wikipedia to 
-a few notions foundational to society.
+We observe that first links direct the wealth of specific knowledge on Wikipedia to 
+a few foundational notions.
 
@@
 
-We identify 2-cycles, meaning a pair of articles with first link pointing to one another.
-Of the 11 million articles, roughly 84,000 are memebers of 2-cycles. 
+We first identify 2-cycles, meaning a pair of articles with first link pointing to one another.
+Of the 11 million articles, roughly 84,000 are members of 2-cycles. 
 The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
-``Health Care'' and ``Medical Case Management,'' ``Broadcasting House,'' and ``BBC,'' ``Secondary Education'' and ``Secondary School'' 
-(see Fig.~\ref{fig:2-cycles}).
+``Health Care'' and ``Medical Case Management''; ``Broadcasting House'' and ``BBC''; and ``Secondary Education'' and ``Secondary School'' 
+(see Fig.~\ref{fig:cycles}).
 
-Outside of the highest ranking 2-cycles, the typical 2-cycle signals a connection between distinct, yet very closely related concepts. 
-We also observe link patterns such as inventor to product (``Voere'' to ``VEC-91''), event to organizer (``Poetry Bus Tour'' to ``Weave Books''), and book to author (``Anatomy of Britain'' to ``Anthony Sampson'').
+Outside of these prominent 2-cycles, the typical 2-cycle signals a connection between distinct, yet very closely related concepts. 
+For example, we observe link patterns such as inventor to product (``Voere'' to ``VEC-91''), event to organizer (``Poetry Bus Tour'' to ``Weave Books''), and book to author (``Anatomy of Britain'' to ``Anthony Sampson'').
 
@@
 and ``Tree of life''; ``Cinema of India,'' ``Indian Cinema,'' and ``Telugu Cinema''
-(see Fig.~\ref{fig:3-cycles}).
+(see Fig.~\ref{fig:cycles}).
 Once we extend our cycle size beyond a length of 6 however, 
@@
 The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day.
-Other lengthy cycles span $60-75$ articles including collections of articles on national histories such as ``Japanese Eras'' 
+Other lengthy cycles span 60--75 articles including collections of articles on national histories such as ``Japanese Eras'' 
 or judicial bodies such as the ``Legislative Assembly of Ontario.''
@@
 
-We can group articles lying on the same path to identify {\it basins}: 
-a group of path-connected articles.
-Since cycles identify only groups of articles with a closed set of links, 
-we additionally measure and rank basins to capture groups of closely related
-articles branching outside of a cycle into the rest of the FLN.
-We rank basins by the total number of traversal visits for each article in the path. 
-Akin to river networks, these basins are areas of accumulation with a path 
-flowing outwards to the rest of the FLN.
-
-The highest ranking basins by the number of traversal visits are groups of articles
-around ``Philosophy.'' 
+The highest ranking basins (Sec.~\ref{Traversal Algorithm}) by the number of traversal visits are groups of articles around ``Philosophy.'' 
 The highest ranking paths include branches of philosophy flowing through 
-``Awareness,'' ``Existence,'' and ``Consciousness'' to ``Philosophy.'' Other paths
+``Awareness,'' ``Existence,'' and ``Consciousness'' to ``Philosophy,'' 
+which form a cycle of seven articles. As of April 10, 2016 the basin flowing from 
+``Philosophy'' has changed to contain instead ``Ancient Greece,'' ``Greek,'' and 
+``Indo-European Languages'' (see network resilience Sec.~\ref{Traversal Funnels}).
+Other paths
 include concepts around ``Mathematics,'' ``Scientific,'' ``Experiments,'' 
@@
 
-Moving beyon basins around ``Philosophy,'' we find other basins around 
+Moving beyond basins around ``Philosophy,'' we find other basins around 
 foundational concepts such as ``Community,'' ``Landmass,'' ``Federal Government,'' 
@@
 \subsection{Traversal Funnels}
+\label{Traversal Funnels}
 
@@
 structure of the FLN, we compute the number of traversal funnels for each.
-Articles directing more paths exert a greater influence over the structure
-of the FLN by increasing the accumulation of first links
-on a particular path. By measuring traversal funnels, we distinguish between an article that simply happened to fall within a cycle from an article funneling 
-many first links.
-
 Ranking articles by the number of traversal funnels we find 
-``Philosophy'' to be by far the highest-ranking article with 
+``Philosophy'' to be by far the highest ranking article with 
 $7.37$ million paths
@@
 The highest-ranking list also includes ``Hip Hop,'' ``Cancer,'' and ``Web Page.''
+
+Traversal funnels are comparatively resilient to changes in the FLN over time.
+While a few links can alter a cycle, potentially changing the number of traversal visits for an article, traversal funnels do not depend on cycles.
+Even when the particular links between articles change, the overall influence measured by traversal funnels is relatively resilient.
 
@@
 When fit to a log-log linear model we find the $99\%$ of articles with zero
-traversal funnels form one regime (with log(rank) less than 4).
+traversal funnels form one regime (with $\log_{10}$(rank) less than 4).
 The top regime, corresponding to the $17, 821$ articles with at least one 
@@
 likely a result of Halloween falling in the month of October. 
-Other popular articles include ``Scientology'' ($463$ thousand views), ``Clint Eastwood'' 
-($341$ thousand views),
+The second most popular article, ``Scientology'' ($463$ thousand views) appeared in popular news as actress Leah Remini announced she's leaving the Church of Scientology 
+\cite{scientology}.
+Other popular articles include ``Clint Eastwood'' ($341$ thousand views),
 and the ``Cold War'' ($298$ thousand views) although each has significantly fewer views compared to the views for ``Halloween''.
@@
 ``Philosophy'' which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears nowhere in the top 20
-by traversal visits. ``Philosophy'' is a relatively popular article 
+by traversal visits with only $240$ thousand page views in October, 2015. ``Philosophy'' is a relatively popular article 
 among articles influencing the shape of the network, but less popular among 
@@
 collection of human knowledge, is rife with the biases of the many contributing editors
-\cite{bias}.
+\cite{wagner2015s}.
 Nevertheless, the findings do reveal
@@
 
-PSD and CMD acknowledge support from NSF Big Data Grand $\#1447634$.
+PSD and CMD acknowledge support from NSF Big Data Grant $\#1447634$.
 
@@
 \newpage
-
-\section{Appendix}
-\label{Appendix}
-
-\subsection{Constructing The First Link Network}
-
-To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
-Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
-eliminating any statistical error due to sampling.
-We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
-The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
-and disambiguations.
-Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
-\cite{wiki_edits},
-our aim is to characterize the structure of the First Link Network.
-
-Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
-Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
-information.
-While a human could manually identify the first link, to map the entire First Link network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
-
-While some libraries exist for MediaWiki Markup,
-approaches using existing libraries led to several bugs 
-including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
-as well as compatibility with other libraries used to parse the XML.
-Consequently,  we developed an algorithm for parsing the first link in the XML version of each article.
-Our parsing algorithm aimed to: 
-1) accurately identify the first link among other page elements, and 
-2) efficiently do so---that is without needing for several passes through the data.
-To process an article in a single pass, we developed a hierarchical system of flags:
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/flags.pdf}  
-  \caption{
-    \textbf{Parsing Algorithm of Wikipedia's XML dump.}
-The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
-Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
-Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.}
-  \label{fig:parsing algorithm}
-\end{figure}
-
-The algorithm loops in three-character chunks to account for potentially nested elements, 
-shifting by one character steps through the article markup.
-If any markup triggers for a flag are detected, a flag is raised. 
-Once a flag is raised, we stop processing and proceed to the next character
-until the flag's closing markup.
-A first link is identified only if Flags 1, 2, and 3 are all off.
-In this case, the entire link is retrieved. 
-We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
-as well as common file extensions for 
-images, audio files, and the like 
-\cite{media_wiki_templates}.
-The first link of an article is then the earliest valid link with unraised flags.
-
-To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
-across 112 cores of the UVM supercomputer cluster
-\cite{vacc}.
-We then joined the results to form a hash table containing every Wikipedia article and its corresponding
-first link. The resulting network map is the basis of our analysis.
-An online appendix containing all of the code and data used can be accessed \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.
-
-\begin{thebibliography}{1}
-
-      \bibitem{editors} Iba, Takashi, et al.
-``Analyzing the creative editing behavior of Wikipedia editors: Through dynamic social network analysis.'' Procedia-Social and Behavioral Sciences 2.4 (2010): 6441-6456.
-
-      \bibitem{accuracy1} Holman Rector, Lucy. ``Comparison of Wikipedia and other encyclopedias for accuracy, breadth, and depth in historical articles.'' Reference services review 36.1 (2008): 7-22.
-
-      \bibitem{accuracy2} Giles, Jim. ``Internet encyclopaedias go head to head.'' Nature 438.7070 (2005): 900-901.
-
-       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. ``An analysis of topical coverage of Wikipedia.'' Journal of Computer Mediated Communication 13.2 (2008): 429-440.
-
-        \bibitem{bias_women} Hill, Benjamin Mako, and Aaron Shaw.
-``The Wikipedia gender gap revisited: characterizing survey response bias with propensity score estimation.'' PloS one 8.6 (2013): e65782.
-
-      \bibitem{missing_entries} Williams, Jake Ryland, Clark, Eric M., Bagrow, James P., Danforth, Christopher M.
-\& Dodds, Peter Sheridan (2015).
-Identifying missing dictionary entries with frequency-conserving context models.
-Phys.
-Rev.
-E, 92, 042808.
-
-      \bibitem{clustering} Banerjee, Somnath, Krishnan Ramanathan, and Ajay Gupta.
-``Clustering short texts using wikipedia.'' Proceedings of the 30th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2007.
-
-      \bibitem{semantic_relatedness} Gabrilovich, Evgeniy, and Shaul Markovitch. ``Computing Semantic Relatedness Using Wikipedia-based Explicit Semantic Analysis.'' IJCAI. Vol. 7. 2007.
-
-      \bibitem{disambiguating} Cucerzan, Silviu. ``Large-Scale Named Entity Disambiguation Based on Wikipedia Data.'' EMNLP-CoNLL. Vol. 7. 2007.
-
-      \bibitem{drugs} Clauson, Kevin A., et al. ``Scope, completeness, and accuracy of drug information in Wikipedia.'' Annals of Pharmacotherapy 42.12 (2008): 1814-1821.
-
-      \bibitem{stats} Wikipedia contributors. ``Wikipedia:Statistics.'' Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 9 Nov. 2015. Web. 13 Nov. 2015.
-
-      \bibitem{pagerank} Page, Lawrence, et al. ``The PageRank citation ranking: bringing order to the Web.'' (1999).
-
-      \bibitem{classifier} Chakrabarti, Soumen, Mukul Joshi, and Vivek Tawde.
-``Enhanced topic distillation using text, markup tags, and hyperlinks.'' Proceedings of the 24th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2001.
-
-      \bibitem{relevance} Kamps, Jaap, and Marijn Koolen. ``Is Wikipedia link structure different?.'' Proceedings of the second ACM international conference on Web search and data mining. ACM, 2009.
-
-      \bibitem{wiki_train} Wikipedia contributors. ``Train.'' Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 
-          \url{https://en.wikipedia.org/wiki/Train}
-
-    \bibitem{locke} Bolton, Martha Brandt.
-``The Taxonomy of Ideas in Locke\textquotesingle s {\it Essay},'' The Cambridge Companion to Locke's ``Essay Concerning Human Understanding.'' Ed.Lex Newman..
-1st ed.
-Cambridge: Cambridge University Press, 2007.
-67-100.
-Cambridge Companions Online.
-Web.
-13 November 2015.
-http://dx.doi.org/10.1017/CCOL0521834333.004.
-
-    \bibitem{aristotle} Studtmann, Paul, ``Aristotle's Categories,'' The Stanford Encyclopedia of Philosophy (Summer 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/sum2014/entries/aristotle-categories/}
-
-    \bibitem{descartes} Smith, Kurt, ``Descartes' Theory of Ideas,'' The Stanford Encyclopedia of Philosophy (Spring 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/spr2014/entries/descartes-ideas/}
-
-    \bibitem{hist_thesaurus} What is the Historical Thesaurus of the OED - Oxford English Dictionary (Oxford English Dictionary)
-    http://public.oed.com/historical-thesaurus-of-the-oed/what-is-the-historical-thesaurus-of-the-oed/
-
-    \bibitem{vacc}  Vermont Advanced Computing Core, \url{http://www.uvm.edu/vacc}.
-
-    \bibitem{wiki_views} Page Views for Wikipedia, Non-mobile site, Normalized (Page Views for Wikipedia, Non-mobile site, Normalized)
-    \url{https://stats.wikimedia.org/EN/TablesPageViewsMonthly.htm}
-
-    \bibitem{weddings} Wedding statistics in the United States (Wedding statistics in the United States)
-    \url{http://www.soundvision.com/article/wedding-statistics-in-the-united-states}
-
-    \bibitem{six_degrees} Elmacioglu, Ergin, and Dongwon Lee. ``On six degrees of separation in DBLP-DB and more.'' ACM SIGMOD Record 34.2 (2005): 33-40.
-
-    \bibitem{wordnet} Princeton University ``About WordNet.'' WordNet. Princeton University. 2010. 
-    \url{http://wordnet.princeton.edu}
-
-    \bibitem{newman} Newman, Mark EJ. ``The structure and function of complex networks.'' SIAM review 45.2 (2003): 167-256.
-
-    \bibitem{disease} Eubank, Stephen, et al. ``Modelling disease outbreaks in realistic urban social networks.'' Nature 429.6988 (2004): 180-184.
-
-    \bibitem{social_nets} Newman, Mark EJ, Duncan J.
-Watts, and Steven H.
-Strogatz.
-``Random graph models of social networks.'' Proceedings of the National Academy of Sciences 99.suppl 1 (2002): 2566-2572.
-
-    \bibitem{geo_basins} Horton, Robert E.
-``Erosional development of streams and their drainage basins; hydrophysical approach to quantitative morphology.'' Geological society of America bulletin 56.3 (1945): 275-370.
-    APA
-
-    \bibitem{dodds} Dodds, Peter Sheridan, and Daniel H. Rothman. ``Unified view of scaling laws for river networks.'' Physical Review E 59.5 (1999): 4865.
-
-    \bibitem{food_webs} Garlaschelli, Diego, Guido Caldarelli, and Luciano Pietronero. ``Universal scaling relations in food webs.'' Nature 423.6936 (2003): 165-168.
-
-    \bibitem{mat_blog} brain of mat kelcey (brain of mat kelcey). 
-    ``Do all first links lead to Philosophy?''
-    \url{http://matpalm.com/blog/2011/08/13/wikipedia-philosophy/}
-
-    \bibitem{Ilmari_first_links} User:Ilmari Karonen/First link (Wikipedia)
-    \url{https://en.wikipedia.org/wiki/User:Ilmari_Karonen/First_link}
-
-    \bibitem{xkcd} xkcd: Extended Mind (xkcd: Extended Mind)
-    \url{http://xkcd.com/903/}
-
-    \bibitem{reddit} Almost every Wikipedia page links to Philosophy if you keep clicking the first link in every article. 
-    \url{https://www.reddit.com/r/InternetIsBeautiful/comments/35asgt/almost_every_wikipedia_page_links_to_philosophy/}
-
-    \bibitem{bias} Wagner, Claudia, et al. ``It's a Man's Wikipedia? Assessing Gender Inequality in an Online Encyclopedia.'' arXiv preprint arXiv:1501.06307 (2015).
-
-    \bibitem{wiki_edits} Wikipedia:Statistics 
-    \url{https://en.wikipedia.org/wiki/Wikipedia:Statistics}
-
-    \bibitem{media_wiki_templates} MediaWiki: ``Help: Templates''
-    \url{https://www.mediawiki.org/wiki/Help:Templates}
-
-    \bibitem{links_that_speak} Ronen, Shahar, Bruno Gonçalves, Kevin Z.
-Hu, Alessandro Vespignani, Steven Pinker, and César A.
-Hidalgo.
-``'Links That Speak: The Global Language Network and Its Association with Global Fame.'' Proceedings of the National Academy of Sciences Proc Natl Acad Sci USA 111.52 (2014).
-
-\end{thebibliography}
-
-\end{document}

### 2016-04-22 writeup/2016-04wikipedia-network/wikipedia-network.supplementary.tex
universal template and bibtex format changed

--- 2016-04-12 writeup/wikipedia-network.tex
@@
-\documentclass[pre,twocolumn,twoside,superscriptaddress,floatfix, aps, 10pt]{revtex4-1}
-
-\usepackage{currfile}
-\lefthyphenmin=3
-\righthyphenmin=2
-
-\usepackage{graphicx,epsfig,verbatim,enumerate}
-\usepackage{amssymb,amsmath}
-\usepackage{ifthen}
-
-% Fonts and typesetting settings
-\usepackage[sc]{mathpazo}
-\usepackage[T1]{fontenc}
-\linespread{1.05} % Palatino needs more space between lines
-\usepackage{microtype}
-\usepackage{textcomp}
-
-\usepackage{longtable}
-
-\usepackage{mathtools}
-
-\newboolean{twocolswitch}
-
-\newcommand{\www}[1]{\url{#1}}
-\newcommand{\req}[1]{(\ref{#1})}
-\newcommand{\Req}[1]{Eq.~(\ref{#1})}
-
-% Lettrines
-\usepackage{lettrine}
-
-%useful shortcuts
-\def\R{\ensuremath{\mathbb{R}}} %\ensuremath adds math mode, if forgotten
-\def\Q{\ensuremath{\mathbb{Q}}}
-\def\N{\ensuremath{\mathbb{N}}}
-\def\Z{\ensuremath{\mathbb{Z}}}
-\def\C{\ensuremath{\mathbb{C}}}
-
-%shorcuts with arguments
-\newcommand{\abs}[1]{\left\vert#1\right\vert} %nice absolute values
-\newcommand{\bt}[1]{\textbf{#1}} %bold
-\newcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert} %vector norm
-\renewcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-
-%piecewise function
-
-%\begin{displaymath}
-%   f(x) = \left\{
-%     \begin{array}{lr}
-%       1 & : x \in \mathbb{Q}\\
-%       0 & : x \notin \mathbb{Q}
-%     \end{array}
-%   \right.
-%\end{displaymath}
-
-%environment
-\newcommand{\tab}{\phantom{ssss}}
-
-\usepackage{hyperref}
-\usepackage{color}
-\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}
-
-%colors
-\definecolor{javagreen}{rgb}{0.25,0.5,0.35} %dark green color
-\definecolor{lightblue}{rgb}{0.149,0.545,0.824} %solarized blue
-\definecolor{sred}{rgb}{0.863, 0.196, 0.184} %solarized red
-
-\newcommand{\blue}[1]{{\leavevmode\color{lightblue}{#1}}} %solarized blue 
-\newcommand{\green}[1]{{\leavevmode\color{javagreen}{#1}}} %command for green
-\newcommand{\red}[1]{{\leavevmode\color{sred}{#1}}} %solarized red
-\newcommand{\gray}[1]{{\leavevmode\color[gray]{0.5}{#1}}} %gray text
-\newcommand{\Prob}[1]{{\rm Pr}\left(#1\right)}
-
-\setboolean{twocolswitch}{true}
-
-\begin{document}
-
-\title{\protect
-Connecting every bit of knowledge: \\
-The structure of Wikipedia's Flrst Link Network
-}
-
-\date{\today}
-
-\author{
-\firstname{Mark}
-\surname{Ibrahim}
-}
-\email{mark.s.ibrahim@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Christopher}
-\surname{M. Danforth}
-}
-\email{chris.danforth@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Peter}
-\surname{Sheridan Dodds}
-}
-
-\email{peter.dodds@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\begin{abstract}
-  \protect
-Apples, porcupines, and the most obscure Bob Dylan song---is every topic a few clicks from Philosophy? 
-Within Wikipedia, the surprising answer is yes: nearly all 
-paths lead to Philosophy.
-Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
-More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
-By following the first link in each article, we algorithmically construct a directed network of 
-all 4.7 million articles: Wikipedia's First Link Network.
-Here, we study the English edition of Wikipedia's First Link Network for insight into how the many 
-articles on inventions, places, people, objects, and events are related and organized.  
-We traverse every path, measuring the accumulation of first links, path lengths, basins, cycles, 
-and the influence each article exerts in shaping the network.
-We find scale-free distributions describe path length, 
-accumulation, and influence. Far from dispersed, first links disproportionately accumulate 
-at a few articles---flowing from specific to general and culminating around fundamental notions such as
-Community, State, and Science. 
-Philosophy funnels more paths than any other article by two orders of magnitude. 
-We also observe a gravitation towards topical articles such as 
-Health Care and Fossil Fuel. 
-These findings enrich our view of the connections and structure of
-Wikipedia's ever growing store of knowledge.
-
-\end{abstract}
-
-\maketitle
-
-%TODO:
-% Abstract 
-    % In traversing every path, we measure ?
-    % Our findings... braided network (what does this mean?)
-% Formatting
-    % incorporate into template
-% Intro
-    % add month to monthly view stats
-% Method
-    % what to cite for drainage basin in geomorphology
-
-% continue on Page 3
-
-\section{Introduction}
-
-Wikipedia is a towering achievement of the Internet age. 
-At no point in history has a larger or more meticulously indexed collection of human knowledge 
-existed.
-Wikipedia contains 37 million articles in 283 languages, 
-with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
-\cite{drugs, stats}.
-Demonstrating its relevance to modern inquiry,
-Wikipedia is the sixth most visited
-site in the world, surpassing 18 billion page views and 11 million edits in a single month.
-\cite{wiki_edits, wiki_views}.
-
-Wikipedia has naturally become the object of many studies. 
-Researchers have examined the cultural dynamics among editors
-\cite{editors},
-the accuracy of the content relative to traditional encyclopedias
-\cite{accuracy1, accuracy2},
-the topics covered 
-\cite{coverage},
-and bias against portions of the population
-\cite{bias_women}.
-Wikipedia's content has also proven to be a powerful tool. 
-Researchers have used Wikipedia to identify missing dictionary entries
-\cite{missing_entries},
-cluster short text
-\cite{clustering},
-compute semantic relatedness
-\cite{semantic_relatedness},
-and disambiguate meaning \cite{disambiguating}.
-
-While these many studies have dissected and fruitfully applied Wikipedia's content,
-here we examine the connections among the many articles.
-A hyperlink from one Wikipedia article to another
-naturally indicates a relationship between the two articles
-\cite{relevance}.
-The notion that hyperlinks convey information about the content of a
-page has proved enormously successful in multiple domains from search engine algorithms 
-such as PageRank 
-\cite{pagerank} 
-to topic classification
-\cite{classifier}.
-Here, we treat a hyperlink as a mechanism connecting two topics.
-
-The authors of a Wikipedia article choose where and whether to include a 
-reference to another Wikipedia article
-in the HTML markup.
-For example, as of November 2014, the authors of the 
-\href{https://en.wikipedia.org/wiki/Train}{``Train''}
-article
-had collectively chosen
-``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
-among others as relevant articles to reference in describing ``Train''
-\cite{wiki_train}
-.
-
-\begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
-  \caption{
-    \textbf{First Link Path For ``Train.''}
-    We follow the first link  to another Wikipedia article
-    in the main body of the article---the area
-    inside the green rectangle, which excludes 
-    side bar elements, the navigation bar and title; the first link is circled in red. 
-    In this example, the first link to another Wikipedia article is ``Rail Transport.'' 
-    We can again select the first link on the ``Rail Transport''
-    article, repeating the process to 
-    form a path of first links. After 11 links, we arrive at ``Philosophy.''
-  }
-  \label{fig:Train First Links}
-\end{figure*}
-
-By focusing our attention on the main body of an article---excluding
-elements in side bars and headings---we 
-attempt to systematically capture the core description of a topic.
-Within the body text, the first link marks the earliest moment in a topic's introduction where the 
-authors choose to directly reference another article. 
-While many links reference relevant details, 
-the first link is an association within a topic's initial 
-description. While 
-``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
-are links detailing particulars, the first link, ``rail transport,''
-is the topic the authors associate with ``Train'' in the introduction
-(Fig.~\ref{fig:Train First Links}).
-``Banana'' has a first link to ``fruit,'' ``Bob Dylan'' has a first link to 
-``Blowing in the Wind,'' and ``Physics'' has a first link to ``natural science.''
-Collectively, first links provide a pragmatic and interpretable 
-means to connect each article to another.
-
-By following the first hyperlink contained in all articles of the English edition of
-Wikipedia as of November, 2014, we form a directed network:
-{\it Wikipedia's First Link Network} (FLN).
-Inspired by the claim that the majority of first links lead to
-``Philosophy''---popularized by an 
-\href{https://xkcd.com/903/}{xkcd}
-comic and subsequently
-discussed in blog posts 
-\cite{xkcd, mat_blog, Ilmari_first_links}
----
-we holistically study how the FLN  yields a flow of connections.
-For methodological details, see Sec.~\ref{Appendix}.
-
-The FLN is a wealth of relations among inventions, places,
-figures, objects, and events across space and time.
-``Train'' for example, links to a parent node, ``Rail Transport,''
-while many child nodes such as ``Steel'' and ``Horsepower'' link to ``Train.''
-As shown in Fig.~\ref{fig:Train First Links}, the path starting at ``Train'' 
-contains ``Goods,'' ``Economics,'' ``Social Science,''
-leading ultimately to ``Philosophy''.
-Unlike previous taxonomies created by individuals
-\cite{locke, descartes, aristotle, hist_thesaurus},
-the relations in the FLN emerge without a centralized effort 
-as the aggregate of each article's authors choice of first link.
-
-Our goal is to study the structure of the FLN for insight into how the information on Wikipedia is organized and related.
-Informed by river network metrics,
-we consider the FLN as a flow, quantifying 
-the accumulation of first links around articles 
-and the influence an article exerts in shaping the FLN.
-Together with cycles, in-degree, depth, and the content of the articles, 
-we build our analysis of the relations among the ideas in Wikipedia.
-
-\section{Traversing the First Link Network}
-
-An essential feature of a directed network's structure is the degree distribution 
-\cite{newman}. 
-The degree distribution has been used to study many phenomena from disease outbreak 
-\cite{disease} 
-to the dynamics of social networks 
-\cite{social_nets}.
-The in-degree distribution in the FLN describes how many first links point to a 
-particular article. 
-Articles with zero in-degree have no references---they are outer leaves in the FLN 
-or are disconnected entirely.
-
-\begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/traversal_visit_algo_figure.pdf}  
-  \caption{
-    \textbf{Traversal Visit Algorithm on a sample network.}
-     The traversal visit vectors are an adjacency matrix for the paths through the network: 
-     the first column indicates the path formed starting with article A. 
-     The number of traversal visits for article A is then the number of paths containing A 
-     or the sum of the first row in our matrix:
-     $\sum_{i=1}^7 A_{\text{visit, i}} = 7$.
-  }
-  \label{fig:Traversal Visits}
-\end{figure*}
-
-Starting at each article, we construct a path through the FLN, 
-to map the flow of connections among articles.
-The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
-Previous studies have used flow to characterize the structure of river networks
-\cite{geo_basins, dodds} and describe the organization of 
-blood networks, food systems, and transportation networks
-\cite{food_webs}.
-
-The first metric we develop quantifies the accumulation of first links.
-The algorithm begins by selecting an article, then traversing the path formed
-by following the first links. Each time a first link references an article, we increment a count
-associated with the article. 
-We continue until a page is revisited or is invalid---defined here to mean pointing to a page outside of Wikipedia.
-We select a second article and repeat the process until we have 
-constructed a path for each article in the network. 
-We define the number of {\it traversal visits} of an article 
-to be the number of references flowing to the ideas in the article---equivalent
-to drainage basin area in geomorphology.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/traversal_funnel_algo_figure.pdf}
-  \caption{
-    \textbf{Traversal Funnel Algorithm on a simple network.}
-  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
-We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.
-For example
-  the number of traversal funnels for article $E$ is 
-  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
-  \label{fig:Traversal Funnels}
-
-\end{figure}
-
-We can characterize the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
-(Fig.~\ref{fig:Traversal Visits}), the path starting at article A is the 
-first column in the traversal visits matrix. 
-An entry of $1$ indicates the path contains a given article and 
-$0$ indicates the path does not.
-To compute the number of traversal visits for an article, we sum the corresponding row in our matrix.
-The traversal visits matrix for 
-our Wikipedia dataset consists of 121 million entries encoding each path out of the more than 
-googol possible paths ($4.7*10^6*2^{4.7*10^6}$) through the FLN .
-
-By measuring the number of first links between
-two articles, we obtain an additional piece of information we call 
-{\it path length.} We can compute the length of a path by summing 
-along a column in our traversal visits matrix. 
-In our Wikipedia dataset, the sum of all path lengths is 232 million first links.
-The path length describes how closely related topics are.
-Although ``Train'' is related to ``Economics'' for example, there are several articles
-bridging the connection:
-``Train'' is more specifically related to transportation, whose object
-is often goods. Goods are one of the fundamental objects of study in Economics.
-Described in links, this relationship is captured by a relatedness of $4$ 
-first links ultimately connecting ``Train'' to ``Economics.''
-
-One possible path through the FLN is a {\it cycle} or a group of articles 
-linking to one another inside a loop. In our sample network 
-(Fig.~\ref{fig:Traversal Visits}) 
-a 3-cycle exists among nodes A, B, and C. 
-We can readily identify
-the types of cycle structures within the FLN
-and rank each by 
-the number of references directed towards a cycle.
-
-We can also form and rank {\it basins} in the network by identifying groups of 
-path-connected articles, not necessarily forming a perfect cycle.
-A basin connects a group of articles and identifies the paths 
-to a particular article.
-
-While traversal visits measure accumulation, each article's first link also 
-influences the shape of the FLN. 
-At a point of accumulation, a single article's first link 
-can exert great influence over the shape of the FLN by directing many
-references on a particular path. To distinguish between an article 
-that simply happened to fall within a cycle from an article funneling 
-many first links, we develop a second metric called {\it traversal funnels}.
-
-To measure traversal funnels, we traverse the FLN in the same manner as we 
-did for traversal visits, but end a path once we enter a cycle.
-We are then able to distinguish between an article related to many other ideas
-only by virtue of its place in a cycle, from an article exerting influence over where the first links flow. 
-An article with a large number of traversal funnels directs many references
-into a particular cycle. In our sample network 
-(see Fig.~\ref{fig:Traversal Funnels}) article C 
-directs the flow of links towards the 3-cycle, while articles A and B are 
-recipients of the flow---without exerting direct influence themselves.
-
-By studying the FLN not only as collection of directly linked pairs of articles, but
-as a flow, 
-we build a powerful arsenal of information.
-From accumulated references, cycles, and basins, we can measure how the many articles in Wikipedia are organized and related.
-
-\section{Results}
-
-\subsection{Degree Distribution}
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
-  \caption{
-    \textbf{Highest Ranking Articles by in-degree.}
-    We rank each article by the number of direct first links to the article (in-degree).
-The highest-ranking articles tend to represent geographical and biological abstractions.
-A full online appendix of the results and data is available \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.}
-  \label{fig:indegree list}
-\end{figure}
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
-  \caption{
-    \textbf{FLN Degree Distribution.}
-We fit the in-degree distribution with a linear $\log_{10}-\log_{10}$ model. 
-The result is a 
-an excellent fit with a Pearson's Correlation Coefficient of $-0.98$, yielding a 
-power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$. 
-}
-  \label{fig:degree distribution}
-\end{figure}
-
-We rank all 11 million articles by in-degree to find 
-the ``United States'' with 80,249 direct first links as the most referenced
-Wikipedia article 
-(Fig.~\ref{fig:indegree list}). 
-Other high-ranking articles
-include foundational abstract concepts such as ``village,'' ``species,''; 
-sports associations such as ``American Football,'' ``Association Football''; 
-and developed nations such as ``France,'' ``Japan,'' ``Russia,'' ``Australia,'' and 
-the ``Netherlands.'' These high ranking articles are useful abstractions: nations
-describe a collection of individuals with a common culture, language, or 
-geographical proximity; sports teams describe an ever changing collection of 
-players often associated with a cultural identity or a geographical 
-region. 
-Since abstractions such as nations and teams are inherently comprised
-of many parts, authors reference the abstraction when describing a part.
-In describing the New England Patriots or the New York Giants
-for example, ``American Football'' is the natural abstraction, which 
-anchors many specific teams.
-
-``Philosophy'' and other philosophical concepts
-are not among the highest-ranking articles by in-degree.
-``Philosophy'' has an in-degree of only 581, with direct first links from articles about Philosophers and areas of Philosophy: ``Existentialism and Humanism,'' ``Predeterminism,'' ''Synoptic Philosophy,'' ``Qualia,'' ``Dorothy Emmet,'' and ``Christopher W.
-Morris.''
-While many articles accumulate at ``Philosophy'' (see traversal visits discussion below), 
-the accumulation is not the 
-result of many articles directly referencing ``Philosophy.'' 
-Instead, first links flow towards ``Philosophy'' as the 
-ultimate anchor, by generalizing from specific to broad.
-
-The FLN's in-degree exhibits a decaying power law distribution where a few articles 
-receive most direct first references, while most articles receive few or none.
-The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5.
-Less than $1\%$ of articles have more than 100 direct first links and $75\%$ of articles
-have fewer than 9. 
-We fit the in-degree distribution against each article's rank with a log-log (base 10) linear regression model
-(see Fig.~\ref{fig:degree distribution}).
-We obtain a Pearson's Correlation Coefficient of $-0.98$
-with a corresponding  power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$.
-
-\subsection{Depth of the FLN}
-
-How many links does a connection among articles span? 
-We find the longest path length is 365,
-corresponding to the yearly calendar of Orthodox Liturgics.
-Each day's Liturgics links to the next day's. On the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
-(see discussion of cycles ~\ref{cycles}).
-We also find similarly lengthy paths following the evolution of a place or topic through time: 
-``1953 in Scotland'' or ``1560s Architecture,'' with articles sequentially proceeding by year, decade or era.
-In general, the longest paths connect temporally organized ideas.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/path_lengths_dist.png}
-  \caption{
-    \textbf{Path Length Distribution.}
-The network depth appears to have a bimodal distribution with a median of 29 first links in a path.
-The 365 cycle for ``Orthodox'' liturgics is the outlier to the right, with historical articles 
-describing nations such as ``Scotland'' and ``UK'' also lying outside the third quartile.
-More than $75\%$ of articles have path lengths between 
-$0$ and $50$ links.}
-  \label{fig:Path Length Distribution}
-\end{figure}
-
-Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
-  \caption{
-    \textbf{Distribution of Traversal Visits.}
-We fit the distribution of traversal visits against article rank with a log-log (base 10) linear regression model log-log model and uncover two regimes: 
-with a cutoff rank of $10^{5}$.
-The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
-Coefficient of $-0.99$ 
-with a corresponding power law rank exponent 
-$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
-The horizontal flattening around the highest
-ranking articles is a result of the cyclic structure (see discussion on cycles).
-The bottom regime has a Pearson's Correlation Coefficient of $-0.93$
-with a corresponding power law rank exponent 
-$\alpha = -0.579$ and a size exponent $\gamma = -0.727$.}
-  \label{fig:Distribution of Visits}
-
-\end{figure}
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/articles_ranked.png}
-  \caption{
-    \textbf{Highest ranking articles by number of traversal visits.}
-We compute the number of traversal visits for each article in the FLN (see 
-Traversal Algorithm section for details). In doing so, we can rank each article
-by the accumulation of first links.
-The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.}
-  \label{fig:highest visits}
-\end{figure}
-
-This roughly corresponds to the official number of articles on Wikipedia: 
-$~4.7$ million as of November 2014---approximately half of the 11 million 
-articles in the XML dump are redirects or disambiguations, not full articles.
-The most common path length is 29, with an interquartile range of 4---path length
-is typically between 26 and 29 articles.
-As a distribution, more than $75\%$ of articles have a path length below 
-$50$ first links 
-while a few temporally organized paths exceed 50 links 
-(see Fig.~\ref{fig:Path Length Distribution}).
-
-\subsection{Traversal Visits}
-
-As a distribution, the number of traversal visits by article appears to follow a decaying power law. 
-The majority of articles have fewer than 30 traversal visits and
-first link references accumulate at a few articles.
-Specifically, $99.76\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
-Meanwhile, the highest ranking 30 articles have an extremely disproportionate number of traversal visits.
-
-\begin{figure*}[tph!]
-  \includegraphics[width=\textwidth]{graphics/2_and_3_cycles.pdf}
-  \caption{
-    \textbf{Highest ranking 2-cycles and 3-cycles.}
-We identify pairs of articles whose first links point to one another, forming
-a 2-cycle. We then rank each pair of articles by the total number of 
-traversal visits to gauge the most referenced groups of two articles linked
-to each other. We find 2-cycles often capture synonyms or articles representing nearly the 
-same concepts as opposed to distinct concepts. Similarly, we identify and 
-rank 3-cycles to find they appear to capture three closely related ideas or synonyms.}
-  \label{fig:2-cycles}
-\end{figure*}
-In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against rank
-to two regimes corresponding to power law rank exponents of 
-$\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
-On the aggregate, the distribution suggests
-a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none.
-The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.
-
-The highest ranking articles by traversal visits are broad topics spanning academic disciplines 
-or notions fundamental to society~(Fig. \ref{fig:highest visits}): ``Science,'' ``Mathematics,'' `` Geography,'' and ``Philosophy''
-as well as ``Community,'' ``State,'' ``Earth,'' ``Information,'' ``Power,'' and ``Communication.''
-The first links flow towards broader topics, accumulating at of these foundational notions.
-While ``Banana'' is a concrete fruit for example, 
-the first links flow from ``Fruit'' to ``Botany'' to ``Biology,'' and ultimately 
-culminate with ``Science'' and ``Philosophy.'' 
-The first links anchor the wealth of specific knowledge on Wikipedia to 
-a few notions foundational to society.
-
-\subsection{Network Cycles}
-\label{cycles}
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
-  \caption{
-    \textbf{Funnels.}
-We represent the highest-ranking articles by the number of traversal 
-funnels to gauge the influence each article exerts in shaping the 
-structure of the FLN. 
-We find ``Philosophy'' exerts an overwhelming proportion
-of the influence, with other abstract notions and topical concepts ranking
-next.}
-  \label{fig:Funnels}
-\end{figure}
-
-We identify 2-cycles, meaning a pair of articles with first link pointing to one another.
-Of the 11 million articles, roughly 84,000 are memebers of 2-cycles. 
-The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
-``Health Care'' and ``Medical Case Management,'' ``Broadcasting House,'' and ``BBC,'' ``Secondary Education'' and ``Secondary School'' 
-(see Fig.~\ref{fig:2-cycles}).
-
-Outside of the highest ranking 2-cycles, the typical 2-cycle signals a connection between distinct, yet very closely related concepts. 
-We also observe link patterns such as inventor to product (``Voere'' to ``VEC-91''), event to organizer (``Poetry Bus Tour'' to ``Weave Books''), and book to author (``Anatomy of Britain'' to ``Anthony Sampson'').
-
-Similarly, 3-cycles capture a synonymous or close relation among 3 articles: ``Tree of life (Biology),'' ``Tree of life (disambiguation),'' 
-and ``Tree of life''; ``Cinema of India,'' ``Indian Cinema,'' and ``Telugu Cinema''
-(see Fig.~\ref{fig:3-cycles}).
-Once we extend our cycle size beyond a length of 6 however, 
-``Philosophy'' along with the remaining list of high ranking articles by traversal visits dominate.
-The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day.
-Other lengthy cycles span $60-75$ articles including collections of articles on national histories such as ``Japanese Eras'' 
-or judicial bodies such as the ``Legislative Assembly of Ontario.''
-
-\subsection{Basins}
-
-We can group articles lying on the same path to identify {\it basins}: 
-a group of path-connected articles.
-Since cycles identify only groups of articles with a closed set of links, 
-we additionally measure and rank basins to capture groups of closely related
-articles branching outside of a cycle into the rest of the FLN.
-We rank basins by the total number of traversal visits for each article in the path. 
-Akin to river networks, these basins are areas of accumulation with a path 
-flowing outwards to the rest of the FLN.
-
-The highest ranking basins by the number of traversal visits are groups of articles
-around ``Philosophy.'' 
-The highest ranking paths include branches of philosophy flowing through 
-``Awareness,'' ``Existence,'' and ``Consciousness'' to ``Philosophy.'' Other paths
-include concepts around ``Mathematics,'' ``Scientific,'' ``Experiments,'' 
-``Biology,'' and ``Fact.''
-These paths link many specific articles to ``Philosophy,'' each funneled through a particular domain.
-
-Moving beyon basins around ``Philosophy,'' we find other basins around 
-foundational concepts such as ``Community,'' ``Landmass,'' ``Federal Government,'' 
-``Presentation,'' and ``Belief System.'' 
-The basins around each of these foundational notions are 
-various paths containing related articles. For example around 
-``Community'' we find basins
-flowing from ``United States'' to ``Federal Republic'' to ``Political Union'' to ``State'' culminating at ``Community''; we also find basins flowing from 
-``Public Policy'' to ``Executive (government)'' to ``Government'' to ``State'' and then 
-to ``Community''; we also find paths flowing through a similar chain beginning
-with ``Democracy,'' another beginning with ``Constitution,'' another at 
-``Dictatorship,'' and so on. The articles build from specific means of organizing
-a community (or society) and then build up to ``Community.'' 
-Other basins around landmass for example begin at specific geographical regions
-such as ``Eastern Europe'' building up to ``Continent'' and finally ``Landmass''.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
-  \caption{
-    \textbf{Distribution of Traversal Funnels.}
-We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed number of traversal funnels. 
-We find two regimes with the top regime (log(rank) < 4) 
-well-explained by a linear fit, yielding Pearson's 
-Correlation Coefficient of $-0.99$ and a 
-corresponding power law rank exponent $\alpha = -1.08$
-and a size exponent $\gamma = 0.074$. 
-The bottom horizontal regime corresponds to the more than $99\%$ of articles
-which hold zero traversal funnels.}
-  \label{fig:Funnels Distribution}
-\end{figure}
-
-\subsection{Traversal Funnels}
-
-To analyze the influence an article exerts in shaping the 
-structure of the FLN, we compute the number of traversal funnels for each.
-Articles directing more paths exert a greater influence over the structure
-of the FLN by increasing the accumulation of first links
-on a particular path. By measuring traversal funnels, we distinguish between an article that simply happened to fall within a cycle from an article funneling 
-many first links.
-
-Ranking articles by the number of traversal funnels we find 
-``Philosophy'' to be by far the highest-ranking article with 
-$7.37$ million paths
-(see Fig.~\ref{fig:Funnels}).
-Of any article, the number of traversal funnels Philosophy holds exceeds 
-all others by more than two orders of magnitude.
-The ``Philosophy'' cycle which contains ``Existence,'' ``Awareness,'' ``Reality,'' 
-and similar articles accumulates the overwhelming proportion of its 
-references through ``Philosophy'': $7.37$ million of the $7.4$ million references
-are funneled through ``Philosophy''.
-Second on the list of highest-ranking articles by traversal funnels is 
-``Presentation'' with only $30$ thousand paths. Similarly abstract 
-concepts also rank highly such as ``Tree of life'' (30 thousand), 
-``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).
-
-Many high-ranking articles are remarkably topical, culturally and politically important concepts.
-For example, ``Health Care,'' a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August, 2009 and February, 2010.
-Other high ranking articles include key historical events such as the ``Cold War'' or critical scarce resource with recent 
-media discussion such as ``Fossil Fuel.'' 
-The highest-ranking list also includes ``Hip Hop,'' ``Cancer,'' and ``Web Page.''
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/popularity_visits.pdf}
-  \caption{
-\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal visits.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal visits.
-While the most popular articles do not necessarily correspond to the articles
-with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-The greatest number of articles fall within roughly 1 million traversal visits and page views.
-Overall, the page views for the top 1000 articles articles by traversal visits appear to be log-normally distributed.
-}
-  \label{fig:Views and Visits}
-
-\end{figure}
-As a distribution, we find few articles influence the structure of the 
-FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
-more than $99\%$ with none---most articles are recipients of 
-the references flowing through the articles with at least one traversal funnel.
-When fit to a log-log linear model we find the $99\%$ of articles with zero
-traversal funnels form one regime (with log(rank) less than 4).
-The top regime, corresponding to the $17, 821$ articles with at least one 
-traversal funnel strongly fits a linear model with
-a Pearson's Correlation Coefficient of $-0.99$ and a 
-corresponding power law rank exponent $\alpha = -1.08$
-and a size exponent $\gamma = 0.074$ 
-(see Fig.~\ref{fig:Funnels Distribution}). Even within the few articles
-influencing the structure of the FLN, only a handful of these exert most of the 
-control.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/popularity_funnels.pdf}
-  \caption{
-\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal funnels.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal funnels.
-}
-  \label{fig:Views and Funnels}
-\end{figure}
-\subsection{Article Popularity}
-
-Wikipedia released an API to measure article popularity by page views
-starting November 2015. 
-Page views add another dimension to our
-findings by constrasting
-the number of users who access a particular article against
-the number of traversal visits and funnels.
-We measure popularity as the total number 
-of page views in the English edition of Wikipedia in the month of 
-October 2015---the earliest full month for which the data is available. 
-We find the highest ranking 1000 articles (by traversal visits) have an average of
-$70$ thousand page views in October with high variation: the standard deviation 
-is $1.1$ million page views. 
-The number of page views has a skewed distribution with $75\%$ of articles
-reaching fewer than $73$ thousand page views in October.
-The article for ``United States'' has the most page 
-views of the highest ranking $1000$ articles by traversal visits with 
-$1.4$ million views in October. The next most popular articles are 
-``India,'' ``World War II,'' and the ``United Kingdom'' each with roughly 1 million page views. 
-The ``Philosophy'' article, despite outranking every article by traversal visits,
-has only $240$ thousand page views in October.
-
-We also analyze article popularity for the highest ranking 1000 articles by 
-traversal funnels. In October, the average page views per article is 
-$22$ thousand with a standard deviation of $95$ thousand views. The distribution
-is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
-``Halloween'' is the most popular article with $2.8$ million views in October,
-likely a result of Halloween falling in the month of October. 
-Other popular articles include ``Scientology'' ($463$ thousand views), ``Clint Eastwood'' 
-($341$ thousand views),
-and the ``Cold War'' ($298$ thousand views) although each has significantly fewer views compared to the views for ``Halloween''.
-Other standout popular October articles include ``24-hour'' and ``12-hour'' clocks likely due
-to Daylight savings in October and ``Marriage'' possibly due to the drop in the number of
-weddings in the months following October
-\cite{weddings}.
-``Philosophy'' which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears nowhere in the top 20
-by traversal visits. ``Philosophy'' is a relatively popular article 
-among articles influencing the shape of the network, but less popular among 
-highly-ranked articles by accumulation.
-
-%========================Conclusion====================================
-
-\section{Concluding Remarks}
-
-The findings here should only be considered within the limitations of their context.
-We examined only the English edition of Wikipedia at a particular moment in time, 
-considering only the first link in in each article as a means to relate articles.
-Finally, Wikipedia, while the largest digital 
-collection of human knowledge, is rife with the biases of the many contributing editors
-\cite{bias}.
-Nevertheless, the findings do reveal
-generalizable relationships and point to foundational notions.
-
-Among our observations is the appearance of multiple scale-free distributions within the network. 
-Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
-articles are responsible for funneling most paths. When measured against the traversal funnels, 
-``Philosophy'' emerges as an exceptional article by orders of magnitude. 
-Nevertheless, many other foundational concepts emerged naturally within FLN. 
-Basins around ``Community'', ``State'', and ``Science'' reveal a foundational structure within the network. 
-More curious is the emergence of recently prominent political and economics topics such as ``Fossil Fuel'' and ``Health Care'' 
-within the highest ranking funnels. 
-Wikipedia seems to reflect not only timeless foundations, but also the topical (at least within English speaking society).
-
-Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the first link.
-These findings also form the basis for the creation of a taxonomy where 
-every idea, event, or object sits within a hierarchy of connected notions.
-The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
-Applications could range from an enhanced network of ideas to psychological insights into how humans form associations.
-Specifically, an ever-evolving reference of related hierarchical concepts could be 
-used to improve search engine algorithms or natural language processing.
-
-\acknowledgments
-MI is grateful to RJ for pointing out the Reddit post 
-\cite{reddit}
-describing how the majority of links lead to ``Philosophy,'' inspiring this research.
-
-The authors are also grateful for the suggestions and input provided by Randall Harp.
-
-PSD and CMD acknowledge support from NSF Big Data Grand $\#1447634$.
-
-The authors also acknowledge support from the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
-
-\newpage
-
 \section{Appendix}
@@
 An online appendix containing all of the code and data used can be accessed \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.
-
-\begin{thebibliography}{1}
-
-      \bibitem{editors} Iba, Takashi, et al.
-``Analyzing the creative editing behavior of Wikipedia editors: Through dynamic social network analysis.'' Procedia-Social and Behavioral Sciences 2.4 (2010): 6441-6456.
-
-      \bibitem{accuracy1} Holman Rector, Lucy. ``Comparison of Wikipedia and other encyclopedias for accuracy, breadth, and depth in historical articles.'' Reference services review 36.1 (2008): 7-22.
-
-      \bibitem{accuracy2} Giles, Jim. ``Internet encyclopaedias go head to head.'' Nature 438.7070 (2005): 900-901.
-
-       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. ``An analysis of topical coverage of Wikipedia.'' Journal of Computer Mediated Communication 13.2 (2008): 429-440.
-
-        \bibitem{bias_women} Hill, Benjamin Mako, and Aaron Shaw.
-``The Wikipedia gender gap revisited: characterizing survey response bias with propensity score estimation.'' PloS one 8.6 (2013): e65782.
-
-      \bibitem{missing_entries} Williams, Jake Ryland, Clark, Eric M., Bagrow, James P., Danforth, Christopher M.
-\& Dodds, Peter Sheridan (2015).
-Identifying missing dictionary entries with frequency-conserving context models.
-Phys.
-Rev.
-E, 92, 042808.
-
-      \bibitem{clustering} Banerjee, Somnath, Krishnan Ramanathan, and Ajay Gupta.
-``Clustering short texts using wikipedia.'' Proceedings of the 30th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2007.
-
-      \bibitem{semantic_relatedness} Gabrilovich, Evgeniy, and Shaul Markovitch. ``Computing Semantic Relatedness Using Wikipedia-based Explicit Semantic Analysis.'' IJCAI. Vol. 7. 2007.
-
-      \bibitem{disambiguating} Cucerzan, Silviu. ``Large-Scale Named Entity Disambiguation Based on Wikipedia Data.'' EMNLP-CoNLL. Vol. 7. 2007.
-
-      \bibitem{drugs} Clauson, Kevin A., et al. ``Scope, completeness, and accuracy of drug information in Wikipedia.'' Annals of Pharmacotherapy 42.12 (2008): 1814-1821.
-
-      \bibitem{stats} Wikipedia contributors. ``Wikipedia:Statistics.'' Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 9 Nov. 2015. Web. 13 Nov. 2015.
-
-      \bibitem{pagerank} Page, Lawrence, et al. ``The PageRank citation ranking: bringing order to the Web.'' (1999).
-
-      \bibitem{classifier} Chakrabarti, Soumen, Mukul Joshi, and Vivek Tawde.
-``Enhanced topic distillation using text, markup tags, and hyperlinks.'' Proceedings of the 24th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2001.
-
-      \bibitem{relevance} Kamps, Jaap, and Marijn Koolen. ``Is Wikipedia link structure different?.'' Proceedings of the second ACM international conference on Web search and data mining. ACM, 2009.
-
-      \bibitem{wiki_train} Wikipedia contributors. ``Train.'' Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 
-          \url{https://en.wikipedia.org/wiki/Train}
-
-    \bibitem{locke} Bolton, Martha Brandt.
-``The Taxonomy of Ideas in Locke\textquotesingle s {\it Essay},'' The Cambridge Companion to Locke's ``Essay Concerning Human Understanding.'' Ed.Lex Newman..
-1st ed.
-Cambridge: Cambridge University Press, 2007.
-67-100.
-Cambridge Companions Online.
-Web.
-13 November 2015.
-http://dx.doi.org/10.1017/CCOL0521834333.004.
-
-    \bibitem{aristotle} Studtmann, Paul, ``Aristotle's Categories,'' The Stanford Encyclopedia of Philosophy (Summer 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/sum2014/entries/aristotle-categories/}
-
-    \bibitem{descartes} Smith, Kurt, ``Descartes' Theory of Ideas,'' The Stanford Encyclopedia of Philosophy (Spring 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/spr2014/entries/descartes-ideas/}
-
-    \bibitem{hist_thesaurus} What is the Historical Thesaurus of the OED - Oxford English Dictionary (Oxford English Dictionary)
-    http://public.oed.com/historical-thesaurus-of-the-oed/what-is-the-historical-thesaurus-of-the-oed/
-
-    \bibitem{vacc}  Vermont Advanced Computing Core, \url{http://www.uvm.edu/vacc}.
-
-    \bibitem{wiki_views} Page Views for Wikipedia, Non-mobile site, Normalized (Page Views for Wikipedia, Non-mobile site, Normalized)
-    \url{https://stats.wikimedia.org/EN/TablesPageViewsMonthly.htm}
-
-    \bibitem{weddings} Wedding statistics in the United States (Wedding statistics in the United States)
-    \url{http://www.soundvision.com/article/wedding-statistics-in-the-united-states}
-
-    \bibitem{six_degrees} Elmacioglu, Ergin, and Dongwon Lee. ``On six degrees of separation in DBLP-DB and more.'' ACM SIGMOD Record 34.2 (2005): 33-40.
-
-    \bibitem{wordnet} Princeton University ``About WordNet.'' WordNet. Princeton University. 2010. 
-    \url{http://wordnet.princeton.edu}
-
-    \bibitem{newman} Newman, Mark EJ. ``The structure and function of complex networks.'' SIAM review 45.2 (2003): 167-256.
-
-    \bibitem{disease} Eubank, Stephen, et al. ``Modelling disease outbreaks in realistic urban social networks.'' Nature 429.6988 (2004): 180-184.
-
-    \bibitem{social_nets} Newman, Mark EJ, Duncan J.
-Watts, and Steven H.
-Strogatz.
-``Random graph models of social networks.'' Proceedings of the National Academy of Sciences 99.suppl 1 (2002): 2566-2572.
-
-    \bibitem{geo_basins} Horton, Robert E.
-``Erosional development of streams and their drainage basins; hydrophysical approach to quantitative morphology.'' Geological society of America bulletin 56.3 (1945): 275-370.
-    APA
-
-    \bibitem{dodds} Dodds, Peter Sheridan, and Daniel H. Rothman. ``Unified view of scaling laws for river networks.'' Physical Review E 59.5 (1999): 4865.
-
-    \bibitem{food_webs} Garlaschelli, Diego, Guido Caldarelli, and Luciano Pietronero. ``Universal scaling relations in food webs.'' Nature 423.6936 (2003): 165-168.
-
-    \bibitem{mat_blog} brain of mat kelcey (brain of mat kelcey). 
-    ``Do all first links lead to Philosophy?''
-    \url{http://matpalm.com/blog/2011/08/13/wikipedia-philosophy/}
-
-    \bibitem{Ilmari_first_links} User:Ilmari Karonen/First link (Wikipedia)
-    \url{https://en.wikipedia.org/wiki/User:Ilmari_Karonen/First_link}
-
-    \bibitem{xkcd} xkcd: Extended Mind (xkcd: Extended Mind)
-    \url{http://xkcd.com/903/}
-
-    \bibitem{reddit} Almost every Wikipedia page links to Philosophy if you keep clicking the first link in every article. 
-    \url{https://www.reddit.com/r/InternetIsBeautiful/comments/35asgt/almost_every_wikipedia_page_links_to_philosophy/}
-
-    \bibitem{bias} Wagner, Claudia, et al. ``It's a Man's Wikipedia? Assessing Gender Inequality in an Online Encyclopedia.'' arXiv preprint arXiv:1501.06307 (2015).
-
-    \bibitem{wiki_edits} Wikipedia:Statistics 
-    \url{https://en.wikipedia.org/wiki/Wikipedia:Statistics}
-
-    \bibitem{media_wiki_templates} MediaWiki: ``Help: Templates''
-    \url{https://www.mediawiki.org/wiki/Help:Templates}
-
-    \bibitem{links_that_speak} Ronen, Shahar, Bruno Gonçalves, Kevin Z.
-Hu, Alessandro Vespignani, Steven Pinker, and César A.
-Hidalgo.
-``'Links That Speak: The Global Language Network and Its Association with Global Fame.'' Proceedings of the National Academy of Sciences Proc Natl Acad Sci USA 111.52 (2014).
-
-\end{thebibliography}
-
-\end{document}

### 2016-04-22 writeup/wikipedia-network.tex
bibtex typos

@@
 Connecting every bit of knowledge: \\
-The structure of Wikipedia's Flrst Link Network
+The structure of Wikipedia's First Link Network
 }
@@
 articles on inventions, places, people, objects, and events are related and organized.  
-We traverse every path, measuring the accumulation of first links, path lengths, basins, cycles, 
+By traversing every path, we measure the accumulation of first links, path lengths, basins, cycles, 
 and the influence each article exerts in shaping the network.
@@
 Community, State, and Science. 
-Philosophy funnels more paths than any other article by two orders of magnitude. 
+Philosophy directs more paths than any other article by two orders of magnitude. 
 We also observe a gravitation towards topical articles such as 
@@
 %TODO:
-% Abstract 
-    % In traversing every path, we measure ?
-    % Our findings... braided network (what does this mean?)
 % Formatting
     % incorporate into template
-% Intro
-    % add month to monthly view stats
+% Abstract 
+    % Our findings... braided network (not sure I understand what this means? braided networks don't come up in the paper)
+        % is the idea to say "we developed an algorithm that's generalizable to a broader class of networks?"
 % Method
     % what to cite for drainage basin in geomorphology
-
-% continue on Page 3
-
+    % clarify basin discussion of method
+    % change name from traversal funnels to non-cycle traversal visits?
+% Results
+    % run MLE; what do i compare against?
 \section{Introduction}
@@
 
-\begin{thebibliography}{1}
-
-      \bibitem{editors} Iba, Takashi, et al.
-``Analyzing the creative editing behavior of Wikipedia editors: Through dynamic social network analysis.'' Procedia-Social and Behavioral Sciences 2.4 (2010): 6441-6456.
-
-      \bibitem{accuracy1} Holman Rector, Lucy. ``Comparison of Wikipedia and other encyclopedias for accuracy, breadth, and depth in historical articles.'' Reference services review 36.1 (2008): 7-22.
-
-      \bibitem{accuracy2} Giles, Jim. ``Internet encyclopaedias go head to head.'' Nature 438.7070 (2005): 900-901.
-
-       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. ``An analysis of topical coverage of Wikipedia.'' Journal of Computer Mediated Communication 13.2 (2008): 429-440.
-
-        \bibitem{bias_women} Hill, Benjamin Mako, and Aaron Shaw.
-``The Wikipedia gender gap revisited: characterizing survey response bias with propensity score estimation.'' PloS one 8.6 (2013): e65782.
-
-      \bibitem{missing_entries} Williams, Jake Ryland, Clark, Eric M., Bagrow, James P., Danforth, Christopher M.
-\& Dodds, Peter Sheridan (2015).
-Identifying missing dictionary entries with frequency-conserving context models.
-Phys.
-Rev.
-E, 92, 042808.
-
-      \bibitem{clustering} Banerjee, Somnath, Krishnan Ramanathan, and Ajay Gupta.
-``Clustering short texts using wikipedia.'' Proceedings of the 30th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2007.
-
-      \bibitem{semantic_relatedness} Gabrilovich, Evgeniy, and Shaul Markovitch. ``Computing Semantic Relatedness Using Wikipedia-based Explicit Semantic Analysis.'' IJCAI. Vol. 7. 2007.
-
-      \bibitem{disambiguating} Cucerzan, Silviu. ``Large-Scale Named Entity Disambiguation Based on Wikipedia Data.'' EMNLP-CoNLL. Vol. 7. 2007.
-
-      \bibitem{drugs} Clauson, Kevin A., et al. ``Scope, completeness, and accuracy of drug information in Wikipedia.'' Annals of Pharmacotherapy 42.12 (2008): 1814-1821.
-
-      \bibitem{stats} Wikipedia contributors. ``Wikipedia:Statistics.'' Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 9 Nov. 2015. Web. 13 Nov. 2015.
-
-      \bibitem{pagerank} Page, Lawrence, et al. ``The PageRank citation ranking: bringing order to the Web.'' (1999).
-
-      \bibitem{classifier} Chakrabarti, Soumen, Mukul Joshi, and Vivek Tawde.
-``Enhanced topic distillation using text, markup tags, and hyperlinks.'' Proceedings of the 24th annual international ACM SIGIR conference on Research and development in information retrieval.
-ACM, 2001.
-
-      \bibitem{relevance} Kamps, Jaap, and Marijn Koolen. ``Is Wikipedia link structure different?.'' Proceedings of the second ACM international conference on Web search and data mining. ACM, 2009.
-
-      \bibitem{wiki_train} Wikipedia contributors. ``Train.'' Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 
-          \url{https://en.wikipedia.org/wiki/Train}
-
-    \bibitem{locke} Bolton, Martha Brandt.
-``The Taxonomy of Ideas in Locke\textquotesingle s {\it Essay},'' The Cambridge Companion to Locke's ``Essay Concerning Human Understanding.'' Ed.Lex Newman..
-1st ed.
-Cambridge: Cambridge University Press, 2007.
-67-100.
-Cambridge Companions Online.
-Web.
-13 November 2015.
-http://dx.doi.org/10.1017/CCOL0521834333.004.
-
-    \bibitem{aristotle} Studtmann, Paul, ``Aristotle's Categories,'' The Stanford Encyclopedia of Philosophy (Summer 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/sum2014/entries/aristotle-categories/}
-
-    \bibitem{descartes} Smith, Kurt, ``Descartes' Theory of Ideas,'' The Stanford Encyclopedia of Philosophy (Spring 2014 Edition), Edward N.
-Zalta (ed.), \url{http://plato.stanford.edu/archives/spr2014/entries/descartes-ideas/}
-
-    \bibitem{hist_thesaurus} What is the Historical Thesaurus of the OED - Oxford English Dictionary (Oxford English Dictionary)
-    http://public.oed.com/historical-thesaurus-of-the-oed/what-is-the-historical-thesaurus-of-the-oed/
-
-    \bibitem{vacc}  Vermont Advanced Computing Core, \url{http://www.uvm.edu/vacc}.
-
-    \bibitem{wiki_views} Page Views for Wikipedia, Non-mobile site, Normalized (Page Views for Wikipedia, Non-mobile site, Normalized)
-    \url{https://stats.wikimedia.org/EN/TablesPageViewsMonthly.htm}
-
-    \bibitem{weddings} Wedding statistics in the United States (Wedding statistics in the United States)
-    \url{http://www.soundvision.com/article/wedding-statistics-in-the-united-states}
-
-    \bibitem{six_degrees} Elmacioglu, Ergin, and Dongwon Lee. ``On six degrees of separation in DBLP-DB and more.'' ACM SIGMOD Record 34.2 (2005): 33-40.
-
-    \bibitem{wordnet} Princeton University ``About WordNet.'' WordNet. Princeton University. 2010. 
-    \url{http://wordnet.princeton.edu}
-
-    \bibitem{newman} Newman, Mark EJ. ``The structure and function of complex networks.'' SIAM review 45.2 (2003): 167-256.
-
-    \bibitem{disease} Eubank, Stephen, et al. ``Modelling disease outbreaks in realistic urban social networks.'' Nature 429.6988 (2004): 180-184.
-
-    \bibitem{social_nets} Newman, Mark EJ, Duncan J.
-Watts, and Steven H.
-Strogatz.
-``Random graph models of social networks.'' Proceedings of the National Academy of Sciences 99.suppl 1 (2002): 2566-2572.
-
-    \bibitem{geo_basins} Horton, Robert E.
-``Erosional development of streams and their drainage basins; hydrophysical approach to quantitative morphology.'' Geological society of America bulletin 56.3 (1945): 275-370.
-    APA
-
-    \bibitem{dodds} Dodds, Peter Sheridan, and Daniel H. Rothman. ``Unified view of scaling laws for river networks.'' Physical Review E 59.5 (1999): 4865.
-
-    \bibitem{food_webs} Garlaschelli, Diego, Guido Caldarelli, and Luciano Pietronero. ``Universal scaling relations in food webs.'' Nature 423.6936 (2003): 165-168.
-
-    \bibitem{mat_blog} brain of mat kelcey (brain of mat kelcey). 
-    ``Do all first links lead to Philosophy?''
-    \url{http://matpalm.com/blog/2011/08/13/wikipedia-philosophy/}
-
-    \bibitem{Ilmari_first_links} User:Ilmari Karonen/First link (Wikipedia)
-    \url{https://en.wikipedia.org/wiki/User:Ilmari_Karonen/First_link}
-
-    \bibitem{xkcd} xkcd: Extended Mind (xkcd: Extended Mind)
-    \url{http://xkcd.com/903/}
-
-    \bibitem{reddit} Almost every Wikipedia page links to Philosophy if you keep clicking the first link in every article. 
-    \url{https://www.reddit.com/r/InternetIsBeautiful/comments/35asgt/almost_every_wikipedia_page_links_to_philosophy/}
-
-    \bibitem{bias} Wagner, Claudia, et al. ``It's a Man's Wikipedia? Assessing Gender Inequality in an Online Encyclopedia.'' arXiv preprint arXiv:1501.06307 (2015).
-
-    \bibitem{wiki_edits} Wikipedia:Statistics 
-    \url{https://en.wikipedia.org/wiki/Wikipedia:Statistics}
-
-    \bibitem{media_wiki_templates} MediaWiki: ``Help: Templates''
-    \url{https://www.mediawiki.org/wiki/Help:Templates}
-
-    \bibitem{links_that_speak} Ronen, Shahar, Bruno Gonçalves, Kevin Z.
-Hu, Alessandro Vespignani, Steven Pinker, and César A.
-Hidalgo.
-``'Links That Speak: The Global Language Network and Its Association with Global Fame.'' Proceedings of the National Academy of Sciences Proc Natl Acad Sci USA 111.52 (2014).
-
-\end{thebibliography}
+\bibliography{wikipedia-network}{}
+\bibliographystyle{plain}
 

### 2016-04-22 writeup/2016-04wikipedia-network/wikipedia-network.abs.tex
successfully complied separate files

--- 2016-04-22 writeup/wikipedia-network.tex
@@
-\documentclass[pre,twocolumn,twoside,superscriptaddress,floatfix, aps, 10pt]{revtex4-1}
-
-\usepackage{currfile}
-\lefthyphenmin=3
-\righthyphenmin=2
-
-\usepackage{graphicx,epsfig,verbatim,enumerate}
-\usepackage{amssymb,amsmath}
-\usepackage{ifthen}
-
-% Fonts and typesetting settings
-\usepackage[sc]{mathpazo}
-\usepackage[T1]{fontenc}
-\linespread{1.05} % Palatino needs more space between lines
-\usepackage{microtype}
-\usepackage{textcomp}
-
-\usepackage{longtable}
-
-\usepackage{mathtools}
-
-\newboolean{twocolswitch}
-
-\newcommand{\www}[1]{\url{#1}}
-\newcommand{\req}[1]{(\ref{#1})}
-\newcommand{\Req}[1]{Eq.~(\ref{#1})}
-
-% Lettrines
-\usepackage{lettrine}
-
-%useful shortcuts
-\def\R{\ensuremath{\mathbb{R}}} %\ensuremath adds math mode, if forgotten
-\def\Q{\ensuremath{\mathbb{Q}}}
-\def\N{\ensuremath{\mathbb{N}}}
-\def\Z{\ensuremath{\mathbb{Z}}}
-\def\C{\ensuremath{\mathbb{C}}}
-
-%shorcuts with arguments
-\newcommand{\abs}[1]{\left\vert#1\right\vert} %nice absolute values
-\newcommand{\bt}[1]{\textbf{#1}} %bold
-\newcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert} %vector norm
-\renewcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-
-%piecewise function
-
-%\begin{displaymath}
-%   f(x) = \left\{
-%     \begin{array}{lr}
-%       1 & : x \in \mathbb{Q}\\
-%       0 & : x \notin \mathbb{Q}
-%     \end{array}
-%   \right.
-%\end{displaymath}
-
-%environment
-\newcommand{\tab}{\phantom{ssss}}
-
-\usepackage{hyperref}
-\usepackage{color}
-\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}
-
-%colors
-\definecolor{javagreen}{rgb}{0.25,0.5,0.35} %dark green color
-\definecolor{lightblue}{rgb}{0.149,0.545,0.824} %solarized blue
-\definecolor{sred}{rgb}{0.863, 0.196, 0.184} %solarized red
-
-\newcommand{\blue}[1]{{\leavevmode\color{lightblue}{#1}}} %solarized blue 
-\newcommand{\green}[1]{{\leavevmode\color{javagreen}{#1}}} %command for green
-\newcommand{\red}[1]{{\leavevmode\color{sred}{#1}}} %solarized red
-\newcommand{\gray}[1]{{\leavevmode\color[gray]{0.5}{#1}}} %gray text
-\newcommand{\Prob}[1]{{\rm Pr}\left(#1\right)}
-
-\setboolean{twocolswitch}{true}
-
-\begin{document}
-
-\title{\protect
-Connecting every bit of knowledge: \\
-The structure of Wikipedia's First Link Network
-}
-
-\date{\today}
-
-\author{
-\firstname{Mark}
-\surname{Ibrahim}
-}
-\email{mark.s.ibrahim@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Christopher}
-\surname{M. Danforth}
-}
-\email{chris.danforth@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Peter}
-\surname{Sheridan Dodds}
-}
-
-\email{peter.dodds@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\begin{abstract}
-  \protect
 Apples, porcupines, and the most obscure Bob Dylan song---is every topic a few clicks from Philosophy? 
@@
 Wikipedia's ever growing store of knowledge.
-
-\end{abstract}
-
-\maketitle
-
-%TODO:
-% Formatting
-    % incorporate into template
-% Abstract 
-    % Our findings... braided network (not sure I understand what this means? braided networks don't come up in the paper)
-        % is the idea to say "we developed an algorithm that's generalizable to a broader class of networks?"
-% Method
-    % what to cite for drainage basin in geomorphology
-    % clarify basin discussion of method
-    % change name from traversal funnels to non-cycle traversal visits?
-% Results
-    % run MLE; what do i compare against?
-\section{Introduction}
-
-Wikipedia is a towering achievement of the Internet age. 
-At no point in history has a larger or more meticulously indexed collection of human knowledge 
-existed.
-Wikipedia contains 37 million articles in 283 languages, 
-with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
-\cite{clauson2008scope, stats}.
-Demonstrating its relevance to modern inquiry,
-Wikipedia is the sixth most visited
-site in the world, surpassing 18 billion page views and 10 million edits in January, 2013 alone.
-\cite{wiki_edits, wiki_views}.
-
-Wikipedia has naturally become the object of many studies. 
-Researchers have examined the cultural dynamics among editors
-\cite{iba2010analyzing, samoilenko2016linguistic},
-the accuracy of the content relative to traditional encyclopedias
-\cite{holman2008comparison, giles2005internet},
-the topics covered 
-\cite{halavais2008analysis},
-and bias against portions of the population
-\cite{hill2013wikipedia}.
-Wikipedia's content has also proven to be a powerful tool. 
-Researchers have used Wikipedia to identify missing dictionary entries
-\cite{williams2015identifying},
-cluster short text
-\cite{banerjee2007clustering},
-compute semantic relatedness
-\cite{gabrilovich2007computing},
-and disambiguate meaning \cite{cucerzan2007large}.
-
-While these many studies have dissected and fruitfully applied Wikipedia's content,
-here we examine the connections among the many articles.
-A hyperlink from one Wikipedia article to another
-naturally indicates a relationship between the two articles
-\cite{kamps2009wikipedia}.
-The notion that hyperlinks convey information about the content of a
-page has proved enormously successful in multiple domains from search engine algorithms 
-such as PageRank 
-\cite{page1999pagerank} 
-to topic classification
-\cite{chakrabarti2001integrating}.
-Here, we treat a hyperlink as a mechanism connecting two topics.
-
-The authors of a Wikipedia article choose where and whether to include a 
-reference to another Wikipedia article
-in the HTML markup.
-For example, as of November 2014, the authors of the 
-\href{https://en.wikipedia.org/wiki/Train}{``Train''}
-article
-had collectively chosen
-``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
-among others as relevant articles to reference in describing ``Train''
-\cite{wiki_train}
-.
-
-\begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
-  \caption{
-    \textbf{First Link Path For ``Train.''}
-    We follow the first link  to another Wikipedia article
-    in the main body of the article---the area
-    inside the green rectangle, which excludes 
-    side bar elements, the navigation bar and title; the first link is circled in red. 
-    In this example, the first link to another Wikipedia article is ``Rail Transport.'' 
-    We can again select the first link on the ``Rail Transport''
-    article, repeating the process to 
-    form a path of first links. After 11 links, we arrive at ``Philosophy.''
-  }
-  \label{fig:Train First Links}
-\end{figure*}
-
-By focusing our attention on the main body of an article---excluding
-elements in side bars and headings---we 
-attempt to systematically capture the core description of a topic.
-Within the body text, the first link marks the earliest moment in a topic's introduction where the 
-authors choose to directly reference another article. 
-While many links reference relevant details, 
-the first link is an association within a topic's initial 
-description. While 
-``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
-are links detailing particulars, the first link, ``rail transport,''
-is the topic the authors associate with ``Train'' in the introduction
-(Fig.~\ref{fig:Train First Links}).
-``Banana'' has a first link to ``fruit,'' ``Bob Dylan'' has a first link to 
-``Blowing in the Wind,'' and ``Physics'' has a first link to ``natural science.''
-Collectively, first links provide a pragmatic and interpretable 
-means to connect each article to another.
-
-By following the first hyperlink contained in all articles of the English edition of
-Wikipedia as of November, 2014, we form a directed network:
-{\it Wikipedia's First Link Network} (FLN).
-Inspired by the claim that the majority of first links lead to
-``Philosophy''---popularized by an 
-\href{https://xkcd.com/903/}{xkcd}
-comic and subsequently
-discussed in blog posts 
-\cite{xkcd, mat_blog, Ilmari_first_links, 
-bob_west}
----
-we holistically study how the FLN  yields a flow of connections.
-For methodological details, see Sec.~\ref{Appendix}.
-
-The FLN is a wealth of relations among inventions, places,
-figures, objects, and events across space and time.
-``Train'' for example, links to a parent node, ``Rail Transport,''
-while many child nodes such as ``Steel'' and ``Horsepower'' link to ``Train.''
-As shown in Fig.~\ref{fig:Train First Links}, the path starting at ``Train'' 
-contains ``Goods,'' ``Economics,'' ``Social Science,''
-leading ultimately to ``Philosophy''.
-Unlike previous taxonomies created by individuals
-\cite{Bolton2007, descartes, aristotle, hist_thesaurus},
-the relations in the FLN emerge without a centralized effort 
-as the aggregate of each article's authors choice of first link.
-
-Our goal is to study the structure of the FLN for insight into how the information on Wikipedia is organized and related.
-Informed by river network metrics,
-we consider the FLN as a flow, quantifying 
-the accumulation of first links around articles 
-and the influence an article exerts in shaping the FLN.
-Together with cycles, in-degree, depth, and the content of the articles, 
-we build our analysis of the relations among the ideas in Wikipedia.
-
-\section{Traversing the First Link Network}
-\label{Traversal Algorithm}
-
-An essential feature of a directed network's structure is the degree distribution 
-\cite{newman2003structure}. 
-The degree distribution has been used to study many phenomena from disease outbreak 
-\cite{eubank2004modelling} 
-to the dynamics of social networks 
-\cite{newman2002random}.
-The in-degree distribution in the FLN describes how many first links point to a 
-particular article. 
-Articles with zero in-degree have no references---they are outer leaves in the FLN 
-or are disconnected entirely.
-
-\begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/traversal_visit_algo_figure.pdf}  
-  \caption{
-    \textbf{Traversal Visit Algorithm on a sample network.}
-     The traversal visit vectors are an adjacency matrix for the paths through the network: 
-     the first column indicates the path formed starting with article A. 
-     The number of traversal visits for article A is then the number of paths containing A 
-     or the sum of the first row in our matrix:
-     $\sum_{i=1}^7 A_{\text{visit, i}} = 7$.
-  }
-  \label{fig:Traversal Visits}
-\end{figure*}
-
-Starting at each article, we construct a path through the FLN, 
-to map the flow of connections among articles.
-The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
-Previous studies have used flow to characterize the structure of river networks
-\cite{horton1945erosional, dodds1999unified} and describe the organization of 
-blood networks, food systems, and transportation networks
-\cite{garlaschelli2003universal}.
-
-The first metric we develop quantifies the accumulation of first links.
-The algorithm begins by selecting an article, then traversing the path formed
-by following the first links. Each time a first link references an article, we increment a count
-associated with the article. 
-We continue until a page is revisited or is invalid---defined here to mean pointing to a page outside of Wikipedia.
-We select a second article and repeat the process until we have 
-constructed a path for each article in the network. 
-We define the number of {\it traversal visits} of an article 
-to be the number of references flowing to the ideas in the article---equivalent
-to drainage basin area in geomorphology.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/traversal_funnel_algo_figure.pdf}
-  \caption{
-    \textbf{Traversal Funnel Algorithm on a simple network.}
-  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
-We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.
-For example
-  the number of traversal funnels for article $E$ is 
-  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
-  \label{fig:Traversal Funnels}
-
-\end{figure}
-
-We can characterize the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
-(Fig.~\ref{fig:Traversal Visits}), the path starting at article A is the 
-first column in the traversal visits matrix. 
-An entry of $1$ indicates the path contains a given article and 
-$0$ indicates the path does not.
-To compute the number of traversal visits for an article, we sum the corresponding row in our matrix.
-The traversal visits matrix for 
-our Wikipedia dataset consists of 121 million entries encoding each path out of the more than 
-googol possible paths ($\simeq4.7\times10^6\times2^{4.7\times10^6}$) through the FLN .
-
-By measuring the number of first links between
-two articles, we obtain an additional piece of information we call 
-{\it path length}, computed by summing 
-columns in the traversal visits matrix. 
-Path length describes how closely related topics are.
-Although ``Train'' is related to ``Economics'' for example, there are several articles
-bridging the connection:
-``Train'' is more specifically related to transportation, whose object
-is often goods. Goods are one of the fundamental objects of study in Economics.
-Described in links, this relationship is captured by a relatedness of $4$ 
-first links ultimately connecting ``Train'' to ``Economics.''
-
-One possible path through the FLN is a {\it cycle} or a group of articles 
-linking to one another inside a loop. In our sample network 
-(Fig.~\ref{fig:Traversal Visits}) 
-a 3-cycle exists among nodes A, B, and C. 
-We can readily identify
-the types of cycle structures within the FLN
-and rank each by 
-the number of references directed towards a cycle.
-
-We can also form and rank {\it basins} in the network by identifying groups of 
-path-connected articles, not necessarily forming a perfect cycle.
-A basin connects a group of articles and identifies the paths 
-to a particular article.
-
-While traversal visits measure accumulation, each article's first link also 
-influences the shape of the FLN. 
-At a point of accumulation, a single article's first link 
-can exert great influence over the shape of the FLN by directing many
-references on a particular path. To distinguish between an article 
-that simply happened to fall within a cycle from an article funneling 
-many first links, we develop a second metric called {\it traversal funnels}.
-To measure traversal funnels, we traverse the FLN in the same manner as we 
-did for traversal visits, but end a path once we enter a cycle.
-We are then able to distinguish between an article related to many other ideas
-only by virtue of its place in a cycle, from an article exerting influence over where the first links flow. 
-In our simple network 
-(see Fig.~\ref{fig:Traversal Funnels}), article C 
-directs the flow of links towards the 3-cycle, while articles A and B are 
-recipients of the flow---without exerting direct influence themselves.
-
-By considering the FLN not only as collection of directly linked pairs of articles, but
-as a flow, 
-With accumulated references, cycles, and basins, we can explore how the many articles in Wikipedia are organized and related.
-
-\section{Results}
-
-\subsection{Degree Distribution}
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
-  \caption{
-    \textbf{Highest Ranking Articles by in-degree.}
-    We rank each article by the number of direct first links to the article (in-degree).
-The highest-ranking articles tend to represent geographical and biological abstractions.
-A full online appendix of the results and data is available \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.}
-  \label{fig:indegree list}
-\end{figure}
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
-  \caption{
-    \textbf{FLN Degree Distribution.}
-We fit the in-degree distribution with a linear $\log_{10}-\log_{10}$ model. 
-The result is a 
-an excellent fit with a Pearson's Correlation Coefficient of $-0.98$, yielding a 
-power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$. 
-}
-  \label{fig:degree distribution}
-\end{figure}
-
-We rank all 11 million articles in the FLN by in-degree to find 
-the ``United States'' with 80,249 direct first links as the most referenced
-Wikipedia article 
-(Fig.~\ref{fig:indegree list}). 
-Other high-ranking articles
-include foundational abstract concepts such as ``village,'' ``species,''; 
-sports associations such as ``American Football,'' ``Association Football''; 
-and developed nations such as ``France,'' ``Japan,'' ``Russia,'' ``Australia,'' and 
-the ``Netherlands.'' These high ranking articles are useful abstractions: nations
-describe a collection of individuals with a common culture, language, or 
-geographical proximity; sports teams describe an ever changing collection of 
-players often associated with a cultural identity or a geographical 
-region. 
-Since abstractions such as nations and teams are inherently comprised
-of many parts, authors often reference the abstraction when describing a part.
-
-``Philosophy'' and other philosophical concepts
-are not among the highest-ranking articles by in-degree.
-``Philosophy'' has an in-degree of only 581, with direct first links from articles about Philosophers and areas of Philosophy: ``Existentialism and Humanism,'' ``Predeterminism,'' ''Synoptic Philosophy,'' ``Qualia,'' ``Dorothy Emmet,'' and ``Christopher W.
-Morris.''
-So while a great many FLN paths flow to 
-``Philosophy'' (see traversal visits discussion below), 
-the accumulation is not the 
-result of many articles directly referencing ``Philosophy.'' 
-Instead, first links flow towards ``Philosophy'' as the 
-ultimate anchor, by generalizing from specific to broad.
-
-The FLN's in-degree exhibits a decaying power law distribution where a few articles 
-receive most direct first references, while most articles receive few or none.
-The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5.
-Less than $1\%$ of articles have more than 100 direct first links and $75\%$ of articles
-have fewer than 9. 
-We fit the in-degree distribution against each article's rank with a log-log (base 10) linear regression model
-(see Fig.~\ref{fig:degree distribution}).
-We obtain a Pearson's Correlation Coefficient of $-0.98$
-with a corresponding  power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$.
-
-\subsection{Depth of the FLN}
-
-How many links does a connection among articles span? 
-We find the longest path length is 365,
-corresponding to the yearly calendar of Orthodox Liturgics.
-Each day's Liturgics links to the next day's. On the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
-(see discussion of cycles ~\ref{cycles}).
-We also find similarly lengthy paths following the evolution of a place or topic through time: 
-``1953 in Scotland'' or ``1560s Architecture,'' with articles sequentially proceeding by year, decade or era.
-In general, the longest paths connect temporally organized ideas.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/path_lengths_dist.png}
-  \caption{
-    \textbf{Path Length Distribution.}
-The network depth appears to have a bimodal distribution with a median of 29 first links in a path.
-The 365 cycle for ``Orthodox'' liturgics is the outlier to the right, with historical articles 
-describing nations such as ``Scotland'' and ``UK'' also lying outside the third quartile.
-More than $75\%$ of articles have path lengths between 
-$0$ and $50$ links.}
-  \label{fig:Path Length Distribution}
-\end{figure}
-
-Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
-  \caption{
-    \textbf{Distribution of Traversal Visits.}
-We fit the distribution of traversal visits against article rank with a log-log (base 10) linear regression model log-log model and uncover two regimes: 
-with a cutoff rank of $10^{5}$.
-The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
-Coefficient of $-0.99$ 
-with a corresponding power law rank exponent 
-$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
-The horizontal flattening around the highest
-ranking articles is a result of the cyclic structure (see discussion on cycles).
-The bottom regime has a Pearson's Correlation Coefficient of $-0.93$
-with a corresponding power law rank exponent 
-$\alpha = -0.579$ and a size exponent $\gamma = -0.727$.}
-  \label{fig:Distribution of Visits}
-
-\end{figure}
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/articles_ranked.png}
-  \caption{
-    \textbf{Highest ranking articles by number of traversal visits.}
-We compute the number of traversal visits for each article in the FLN (see 
-Sec.~\ref{Traversal Algorithm}). 
-In doing so, we can rank each article by the accumulation of first links.
-The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.}
-  \label{fig:highest visits}
-\end{figure}
-
-This roughly corresponds to the official number of articles on Wikipedia: 
-$~4.7$ million as of November 2014---approximately half of the 11 million 
-articles in the XML dump are redirects or disambiguations, not full articles.
-The most common path length is 29, with $75\%$ of path lengths ranging between 7--30 articles.
-As a distribution, more than $75\%$ of articles have a path length below 
-$50$ first links 
-while a few temporally organized paths exceed 50 links 
-(see Fig.~\ref{fig:Path Length Distribution}).
-
-\subsection{Traversal Visits}
-
-As a distribution, the number of traversal visits by article appears to follow a decaying power law. 
-The majority of articles have fewer than 30 traversal visits and
-first link references accumulate at a few articles.
-Specifically, $99.76\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
-Meanwhile, the highest ranking 30 articles have an extremely disproportionate number of traversal visits.
-
-\begin{figure*}[tph!]
-  \includegraphics[width=\textwidth]{graphics/2_and_3_cycles.pdf}
-  \caption{
-    \textbf{Highest ranking 2-cycles and 3-cycles.}
-We identify pairs of articles whose first links point to one another, forming
-a 2-cycle. We then rank each pair of articles by the total number of 
-traversal visits to gauge the most referenced groups of two articles linked
-to each other. We find 2-cycles often capture synonyms or articles representing nearly the 
-same concepts as opposed to distinct concepts. Similarly, we identify and 
-rank 3-cycles to find they appear to capture three closely related ideas or synonyms.}
-  \label{fig:cycles}
-\end{figure*}
-In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against rank
-to two regimes corresponding to power law rank exponents of 
-$\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
-On the aggregate, the distribution suggests
-a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none.
-The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.
-
-The highest ranking articles by traversal visits are broad topics spanning academic disciplines 
-or notions fundamental to society~(Fig. \ref{fig:highest visits}): ``Science,'' ``Mathematics,'' ``Geography,'' and ``Philosophy''
-as well as ``Community,'' ``State,'' ``Earth,'' ``Information,'' ``Power,'' and ``Communication.''
-The first links flow towards broader topics, accumulating at of these foundational notions.
-While ``Banana'' is a concrete fruit for example, 
-the first links flow from ``Fruit'' to ``Botany'' to ``Biology,'' and ultimately 
-culminate with ``Science'' and ``Philosophy.'' 
-We observe that first links direct the wealth of specific knowledge on Wikipedia to 
-a few foundational notions.
-
-\subsection{Network Cycles}
-\label{cycles}
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
-  \caption{
-    \textbf{Funnels.}
-We represent the highest-ranking articles by the number of traversal 
-funnels to gauge the influence each article exerts in shaping the 
-structure of the FLN. 
-We find ``Philosophy'' exerts an overwhelming proportion
-of the influence, with other abstract notions and topical concepts ranking
-next.}
-  \label{fig:Funnels}
-\end{figure}
-
-We first identify 2-cycles, meaning a pair of articles with first link pointing to one another.
-Of the 11 million articles, roughly 84,000 are members of 2-cycles. 
-The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
-``Health Care'' and ``Medical Case Management''; ``Broadcasting House'' and ``BBC''; and ``Secondary Education'' and ``Secondary School'' 
-(see Fig.~\ref{fig:cycles}).
-
-Outside of these prominent 2-cycles, the typical 2-cycle signals a connection between distinct, yet very closely related concepts. 
-For example, we observe link patterns such as inventor to product (``Voere'' to ``VEC-91''), event to organizer (``Poetry Bus Tour'' to ``Weave Books''), and book to author (``Anatomy of Britain'' to ``Anthony Sampson'').
-
-Similarly, 3-cycles capture a synonymous or close relation among 3 articles: ``Tree of life (Biology),'' ``Tree of life (disambiguation),'' 
-and ``Tree of life''; ``Cinema of India,'' ``Indian Cinema,'' and ``Telugu Cinema''
-(see Fig.~\ref{fig:cycles}).
-Once we extend our cycle size beyond a length of 6 however, 
-``Philosophy'' along with the remaining list of high ranking articles by traversal visits dominate.
-The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day.
-Other lengthy cycles span 60--75 articles including collections of articles on national histories such as ``Japanese Eras'' 
-or judicial bodies such as the ``Legislative Assembly of Ontario.''
-
-\subsection{Basins}
-
-The highest ranking basins (Sec.~\ref{Traversal Algorithm}) by the number of traversal visits are groups of articles around ``Philosophy.'' 
-The highest ranking paths include branches of philosophy flowing through 
-``Awareness,'' ``Existence,'' and ``Consciousness'' to ``Philosophy,'' 
-which form a cycle of seven articles. As of April 10, 2016 the basin flowing from 
-``Philosophy'' has changed to contain instead ``Ancient Greece,'' ``Greek,'' and 
-``Indo-European Languages'' (see network resilience Sec.~\ref{Traversal Funnels}).
-Other paths
-include concepts around ``Mathematics,'' ``Scientific,'' ``Experiments,'' 
-``Biology,'' and ``Fact.''
-These paths link many specific articles to ``Philosophy,'' each funneled through a particular domain.
-
-Moving beyond basins around ``Philosophy,'' we find other basins around 
-foundational concepts such as ``Community,'' ``Landmass,'' ``Federal Government,'' 
-``Presentation,'' and ``Belief System.'' 
-The basins around each of these foundational notions are 
-various paths containing related articles. For example around 
-``Community'' we find basins
-flowing from ``United States'' to ``Federal Republic'' to ``Political Union'' to ``State'' culminating at ``Community''; we also find basins flowing from 
-``Public Policy'' to ``Executive (government)'' to ``Government'' to ``State'' and then 
-to ``Community''; we also find paths flowing through a similar chain beginning
-with ``Democracy,'' another beginning with ``Constitution,'' another at 
-``Dictatorship,'' and so on. The articles build from specific means of organizing
-a community (or society) and then build up to ``Community.'' 
-Other basins around landmass for example begin at specific geographical regions
-such as ``Eastern Europe'' building up to ``Continent'' and finally ``Landmass''.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
-  \caption{
-    \textbf{Distribution of Traversal Funnels.}
-We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed number of traversal funnels. 
-We find two regimes with the top regime (log(rank) < 4) 
-well-explained by a linear fit, yielding Pearson's 
-Correlation Coefficient of $-0.99$ and a 
-corresponding power law rank exponent $\alpha = -1.08$
-and a size exponent $\gamma = 0.074$. 
-The bottom horizontal regime corresponds to the more than $99\%$ of articles
-which hold zero traversal funnels.}
-  \label{fig:Funnels Distribution}
-\end{figure}
-
-\subsection{Traversal Funnels}
-\label{Traversal Funnels}
-
-To analyze the influence an article exerts in shaping the 
-structure of the FLN, we compute the number of traversal funnels for each.
-Ranking articles by the number of traversal funnels we find 
-``Philosophy'' to be by far the highest ranking article with 
-$7.37$ million paths
-(see Fig.~\ref{fig:Funnels}).
-Of any article, the number of traversal funnels Philosophy holds exceeds 
-all others by more than two orders of magnitude.
-The ``Philosophy'' cycle which contains ``Existence,'' ``Awareness,'' ``Reality,'' 
-and similar articles accumulates the overwhelming proportion of its 
-references through ``Philosophy'': $7.37$ million of the $7.4$ million references
-are funneled through ``Philosophy''.
-Second on the list of highest-ranking articles by traversal funnels is 
-``Presentation'' with only $30$ thousand paths. Similarly abstract 
-concepts also rank highly such as ``Tree of life'' (30 thousand), 
-``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).
-
-Many high-ranking articles are remarkably topical, culturally and politically important concepts.
-For example, ``Health Care,'' a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August, 2009 and February, 2010.
-Other high ranking articles include key historical events such as the ``Cold War'' or critical scarce resource with recent 
-media discussion such as ``Fossil Fuel.'' 
-The highest-ranking list also includes ``Hip Hop,'' ``Cancer,'' and ``Web Page.''
-
-Traversal funnels are comparatively resilient to changes in the FLN over time.
-While a few links can alter a cycle, potentially changing the number of traversal visits for an article, traversal funnels do not depend on cycles.
-Even when the particular links between articles change, the overall influence measured by traversal funnels is relatively resilient.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/popularity_visits.pdf}
-  \caption{
-\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal visits.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal visits.
-While the most popular articles do not necessarily correspond to the articles
-with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-The greatest number of articles fall within roughly 1 million traversal visits and page views.
-Overall, the page views for the top 1000 articles articles by traversal visits appear to be log-normally distributed.
-}
-  \label{fig:Views and Visits}
-
-\end{figure}
-As a distribution, we find few articles influence the structure of the 
-FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
-more than $99\%$ with none---most articles are recipients of 
-the references flowing through the articles with at least one traversal funnel.
-When fit to a log-log linear model we find the $99\%$ of articles with zero
-traversal funnels form one regime (with $\log_{10}$(rank) less than 4).
-The top regime, corresponding to the $17, 821$ articles with at least one 
-traversal funnel strongly fits a linear model with
-a Pearson's Correlation Coefficient of $-0.99$ and a 
-corresponding power law rank exponent $\alpha = -1.08$
-and a size exponent $\gamma = 0.074$ 
-(see Fig.~\ref{fig:Funnels Distribution}). Even within the few articles
-influencing the structure of the FLN, only a handful of these exert most of the 
-control.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/popularity_funnels.pdf}
-  \caption{
-\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal funnels.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal funnels.
-}
-  \label{fig:Views and Funnels}
-\end{figure}
-\subsection{Article Popularity}
-
-Wikipedia released an API to measure article popularity by page views
-starting November 2015. 
-Page views add another dimension to our
-findings by constrasting
-the number of users who access a particular article against
-the number of traversal visits and funnels.
-We measure popularity as the total number 
-of page views in the English edition of Wikipedia in the month of 
-October 2015---the earliest full month for which the data is available. 
-We find the highest ranking 1000 articles (by traversal visits) have an average of
-$70$ thousand page views in October with high variation: the standard deviation 
-is $1.1$ million page views. 
-The number of page views has a skewed distribution with $75\%$ of articles
-reaching fewer than $73$ thousand page views in October.
-The article for ``United States'' has the most page 
-views of the highest ranking $1000$ articles by traversal visits with 
-$1.4$ million views in October. The next most popular articles are 
-``India,'' ``World War II,'' and the ``United Kingdom'' each with roughly 1 million page views. 
-The ``Philosophy'' article, despite outranking every article by traversal visits,
-has only $240$ thousand page views in October.
-
-We also analyze article popularity for the highest ranking 1000 articles by 
-traversal funnels. In October, the average page views per article is 
-$22$ thousand with a standard deviation of $95$ thousand views. The distribution
-is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
-``Halloween'' is the most popular article with $2.8$ million views in October,
-likely a result of Halloween falling in the month of October. 
-The second most popular article, ``Scientology'' ($463$ thousand views) appeared in popular news as actress Leah Remini announced she's leaving the Church of Scientology 
-\cite{scientology}.
-Other popular articles include ``Clint Eastwood'' ($341$ thousand views),
-and the ``Cold War'' ($298$ thousand views) although each has significantly fewer views compared to the views for ``Halloween''.
-Other standout popular October articles include ``24-hour'' and ``12-hour'' clocks likely due
-to Daylight savings in October and ``Marriage'' possibly due to the drop in the number of
-weddings in the months following October
-\cite{weddings}.
-``Philosophy'' which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears nowhere in the top 20
-by traversal visits with only $240$ thousand page views in October, 2015. ``Philosophy'' is a relatively popular article 
-among articles influencing the shape of the network, but less popular among 
-highly-ranked articles by accumulation.
-
-%========================Conclusion====================================
-
-\section{Concluding Remarks}
-
-The findings here should only be considered within the limitations of their context.
-We examined only the English edition of Wikipedia at a particular moment in time, 
-considering only the first link in in each article as a means to relate articles.
-Finally, Wikipedia, while the largest digital 
-collection of human knowledge, is rife with the biases of the many contributing editors
-\cite{wagner2015s}.
-Nevertheless, the findings do reveal
-generalizable relationships and point to foundational notions.
-
-Among our observations is the appearance of multiple scale-free distributions within the network. 
-Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
-articles are responsible for funneling most paths. When measured against the traversal funnels, 
-``Philosophy'' emerges as an exceptional article by orders of magnitude. 
-Nevertheless, many other foundational concepts emerged naturally within FLN. 
-Basins around ``Community'', ``State'', and ``Science'' reveal a foundational structure within the network. 
-More curious is the emergence of recently prominent political and economics topics such as ``Fossil Fuel'' and ``Health Care'' 
-within the highest ranking funnels. 
-Wikipedia seems to reflect not only timeless foundations, but also the topical (at least within English speaking society).
-
-Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the first link.
-These findings also form the basis for the creation of a taxonomy where 
-every idea, event, or object sits within a hierarchy of connected notions.
-The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
-Applications could range from an enhanced network of ideas to psychological insights into how humans form associations.
-Specifically, an ever-evolving reference of related hierarchical concepts could be 
-used to improve search engine algorithms or natural language processing.
-
-\acknowledgments
-MI is grateful to RJ for pointing out the Reddit post 
-\cite{reddit}
-describing how the majority of links lead to ``Philosophy,'' inspiring this research.
-
-The authors are also grateful for the suggestions and input provided by Randall Harp.
-
-PSD and CMD acknowledge support from NSF Big Data Grant $\#1447634$.
-
-The authors also acknowledge support from the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
-
-\newpage
-
-\section{Appendix}
-\label{Appendix}
-
-\subsection{Constructing The First Link Network}
-
-To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
-Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
-eliminating any statistical error due to sampling.
-We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
-The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
-and disambiguations.
-Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
-\cite{wiki_edits},
-our aim is to characterize the structure of the First Link Network.
-
-Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
-Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
-information.
-While a human could manually identify the first link, to map the entire First Link network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
-
-While some libraries exist for MediaWiki Markup,
-approaches using existing libraries led to several bugs 
-including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
-as well as compatibility with other libraries used to parse the XML.
-Consequently,  we developed an algorithm for parsing the first link in the XML version of each article.
-Our parsing algorithm aimed to: 
-1) accurately identify the first link among other page elements, and 
-2) efficiently do so---that is without needing for several passes through the data.
-To process an article in a single pass, we developed a hierarchical system of flags:
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/flags.pdf}  
-  \caption{
-    \textbf{Parsing Algorithm of Wikipedia's XML dump.}
-The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
-Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
-Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.}
-  \label{fig:parsing algorithm}
-\end{figure}
-
-The algorithm loops in three-character chunks to account for potentially nested elements, 
-shifting by one character steps through the article markup.
-If any markup triggers for a flag are detected, a flag is raised. 
-Once a flag is raised, we stop processing and proceed to the next character
-until the flag's closing markup.
-A first link is identified only if Flags 1, 2, and 3 are all off.
-In this case, the entire link is retrieved. 
-We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
-as well as common file extensions for 
-images, audio files, and the like 
-\cite{media_wiki_templates}.
-The first link of an article is then the earliest valid link with unraised flags.
-
-To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
-across 112 cores of the UVM supercomputer cluster
-\cite{vacc}.
-We then joined the results to form a hash table containing every Wikipedia article and its corresponding
-first link. The resulting network map is the basis of our analysis.
-An online appendix containing all of the code and data used can be accessed \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.
-
-\bibliography{wikipedia-network}{}
-\bibliographystyle{plain}
-
-\end{document}

### 2016-04-22 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
successfully complied separate files

@@
 \begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
+  \includegraphics[width=\textwidth]{../graphics/Train.pdf}  
   \caption{
@@
 \begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/traversal_visit_algo_figure.pdf}  
+  \includegraphics[width=\textwidth]{../graphics/traversal_visit_algo_figure.pdf}  
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/traversal_funnel_algo_figure.pdf}
+  \includegraphics[width=\columnwidth]{../graphics/traversal_funnel_algo_figure.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
+  \includegraphics[width=\columnwidth]{../graphics/articles_ndegree.png}
   \caption{
     \textbf{Highest Ranking Articles by in-degree.}
-    We rank each article by the number of direct first links to the article (in-degree).
-The highest-ranking articles tend to represent geographical and biological abstractions.
-A full online appendix of the results and data is available \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.}
+    We rank each article by the number of direct first links to the article (in-degree). The highest-ranking articles tend to represent geographical and biological abstractions. 
+}
   \label{fig:indegree list}
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
+  \includegraphics[width=\columnwidth]{../graphics/ndegree_loglog.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/path_lengths_dist.png}
+  \includegraphics[width=\columnwidth]{../graphics/path_lengths_dist.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
+  \includegraphics[width=\columnwidth]{../graphics/traversals_per_article.png} 
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/articles_ranked.png}
+  \includegraphics[width=\columnwidth]{../graphics/articles_ranked.png}
   \caption{
@@
 \begin{figure*}[tph!]
-  \includegraphics[width=\textwidth]{graphics/2_and_3_cycles.pdf}
+  \includegraphics[width=\textwidth]{../graphics/2_and_3_cycles.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
+  \includegraphics[width=\columnwidth]{../graphics/top_funnels.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
+  \includegraphics[width=\columnwidth]{../graphics/funnels_distribution.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/popularity_visits.pdf}
+  \includegraphics[width=\columnwidth]{../graphics/popularity_visits.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/popularity_funnels.pdf}
+  \includegraphics[width=\columnwidth]{../graphics/popularity_funnels.pdf}
   \caption{

### 2016-04-26 writeup/2016-04wikipedia-network/wikipedia-network-revtex4-combined.tex
MLE power law for indegree + compilation success

--- 2016-04-22 writeup/wikipedia-network.tex
@@
-\documentclass[pre,twocolumn,twoside,superscriptaddress,floatfix, aps, 10pt]{revtex4-1}
+\documentclass[pre,twocolumn,twoside,byrevtex,superscriptaddress,floatfix]{revtex4-1}
 
@@
 
+\usepackage{color}
+\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}
+
 \usepackage{graphicx,epsfig,verbatim,enumerate}
@@
 \usepackage{ifthen}
-
-% Fonts and typesetting settings
-\usepackage[sc]{mathpazo}
-\usepackage[T1]{fontenc}
-\linespread{1.05} % Palatino needs more space between lines
-\usepackage{microtype}
-\usepackage{textcomp}
 
@@
 
+\newcommand{\avg}[1]{\left\langle#1\right\rangle}
+\newcommand{\tavg}[1]{\langle#1\rangle}
+
+\newcommand{\LavgHa}[2]{\tavg{L_{#1,#2}}}
+\newcommand{\LavgHaN}[2]{\tavg{L_{#1,#2}}}
+\newcommand{\Lavg}{\tavg{L}}
+
+\newcommand{\sindex}[1]{}
+\newcommand{\nindex}[1]{}
+
+\newcommand{\etal}{\textit{et al.}}
 \newcommand{\www}[1]{\url{#1}}
@@
 
-% Lettrines
+% lettrines
 \usepackage{lettrine}
 
-%useful shortcuts
-\def\R{\ensuremath{\mathbb{R}}} %\ensuremath adds math mode, if forgotten
-\def\Q{\ensuremath{\mathbb{Q}}}
-\def\N{\ensuremath{\mathbb{N}}}
-\def\Z{\ensuremath{\mathbb{Z}}}
-\def\C{\ensuremath{\mathbb{C}}}
+\newcommand{\tbf}{\textbf}
+\newcommand{\tit}{\textit}
 
-%shorcuts with arguments
-\newcommand{\abs}[1]{\left\vert#1\right\vert} %nice absolute values
-\newcommand{\bt}[1]{\textbf{#1}} %bold
-\newcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert} %vector norm
-\renewcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
+\newcommand{\dee}[1]{\mbox{d}#1}
+\newcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
+\newcommand{\pdiffsq}[2]{\frac{\partial^2 #1}{{\partial #2}^2}}
+\newcommand{\diff}[2]{\frac{{\rm d}#1}{{\rm d}#2}}
+\newcommand{\diffsq}[2]{\frac{{\rm d}^{2}#1}{{\rm d} {#2}^2}}
+\newcommand{\tdiff}[2]{\mbox{d} #1/\mbox{d} #2}
+\newcommand{\tdiffsq}[2]{\mbox{d}^{2} #1/\mbox{d} {#2}^2}
+\newcommand{\tpdiff}[2]{\partial #1/\partial #2}
+\newcommand{\tpdiffsq}[2]{\partial^2 #1/\partial {#2}^2}
+\newcommand{\postdee}[1]{\,\mbox{d}#1}
 
-%piecewise function
-
-%\begin{displaymath}
-%   f(x) = \left\{
-%     \begin{array}{lr}
-%       1 & : x \in \mathbb{Q}\\
-%       0 & : x \notin \mathbb{Q}
-%     \end{array}
-%   \right.
-%\end{displaymath}
-
-%environment
-\newcommand{\tab}{\phantom{ssss}}
-
-\usepackage{hyperref}
-\usepackage{color}
-\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}
-
-%colors
-\definecolor{javagreen}{rgb}{0.25,0.5,0.35} %dark green color
-\definecolor{lightblue}{rgb}{0.149,0.545,0.824} %solarized blue
-\definecolor{sred}{rgb}{0.863, 0.196, 0.184} %solarized red
-
-\newcommand{\blue}[1]{{\leavevmode\color{lightblue}{#1}}} %solarized blue 
-\newcommand{\green}[1]{{\leavevmode\color{javagreen}{#1}}} %command for green
-\newcommand{\red}[1]{{\leavevmode\color{sred}{#1}}} %solarized red
-\newcommand{\gray}[1]{{\leavevmode\color[gray]{0.5}{#1}}} %gray text
 \newcommand{\Prob}[1]{{\rm Pr}\left(#1\right)}
 
+% \newcommand{\kstar}{d^{\ast}}
+% \newcommand{\kstari}{k_i^\ast}
+\newcommand{\kstar}{d^\ast}
+\newcommand{\kstari}{d_i^\ast}
+
+\newcommand{\dstar}{d^\ast}
+\newcommand{\dstari}{d_i^\ast}
+
+\newcommand{\phifix}{{\phi^{\ast}}}
+\newcommand{\phifixc}{{\phi_c^{\ast}}}
+\newcommand{\phifixb}{{\phi_b^{\ast}}}
+
+\newcommand{\phiiactive}{\phi_{i,t}}
+
+\newcommand{\phiiup}{{\phi_{i,\rm on}}}
+\newcommand{\phiidown}{{\phi_{i,\rm off}}}
+
+\newcommand{\phiup}{{\phi_{\rm on}}}
+\newcommand{\phidown}{{\phi_{\rm off}}}
+
+\newcommand{\Gfun}{G}
+
+\newcommand{\dstardist}{g}
+\newcommand{\dosedist}{f}
+\newcommand{\dosedistk}{f^{k\star}}
+
+% chaotic contagion
+\newcommand{\edgeinfprob}{\rho}
+\newcommand{\nodeinfprob}{\phi}
+
+\newcommand{\avgdegree}{k_{\rm avg}}
+
+\newcommand{\stateA}{S_{0}}
+\newcommand{\stateB}{S_{1}}
+
+\newcommand{\lcce}{H}
+
+\newcommand{\effectivecount}{f_{q,\textnormal{exp}}}
+\newcommand{\effectivecountq}[1]{f_{#1,\textnormal{exp}}}
+
+\newcommand{\simonrho}{\rho}
+\newcommand{\zipfexponent}{\alpha}
+
+\newcommand{\partitionprob}{q}
+\newcommand{\partitiontype}[1]{$\partitionprob$$=$$#1$}
+
+\newcommand{\onehalf}{\frac{1}{2}}
+\newcommand{\onequarter}{\frac{1}{4}}
+
+\newcommand{\clgamma}{\hat{\zipfexponent}}
+\newcommand{\xmin}{r_{\textnormal{max}}}
+\newcommand{\kstat}{D}
+\newcommand{\clp}{\mbox{$p$-value}}
+\newcommand{\sigamma}{1-\simonrho}
+
 \setboolean{twocolswitch}{true}
+
+\bibliographystyle{unsrtabbrv}
 
@@
 }
-
-\date{\today}
 
@@
 
+\date{\today}
+
 \begin{abstract}
@@
 
+\pacs{89.65.-s,89.75.Da,89.75.Fb,89.75.-k}
+
 \maketitle
 
-%TODO:
-% Formatting
-    % incorporate into template
-% Abstract 
-    % Our findings... braided network (not sure I understand what this means? braided networks don't come up in the paper)
-        % is the idea to say "we developed an algorithm that's generalizable to a broader class of networks?"
-% Method
-    % what to cite for drainage basin in geomorphology
-    % clarify basin discussion of method
-    % change name from traversal funnels to non-cycle traversal visits?
-% Results
-    % run MLE; what do i compare against?
 \section{Introduction}
@@
 
-\section{Appendix}
-\label{Appendix}
+\acknowledgments
+MI is grateful to RJ for pointing out the Reddit post 
+\cite{reddit}
+describing how the majority of links lead to ``Philosophy,'' inspiring this research.
 
-\subsection{Constructing The First Link Network}
+The authors are also grateful for the suggestions and input provided by Randall Harp.
 
-To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
-Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
-eliminating any statistical error due to sampling.
-We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
-The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
-and disambiguations.
-Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
-\cite{wiki_edits},
-our aim is to characterize the structure of the First Link Network.
+PSD and CMD acknowledge support from NSF Big Data Grant $\#1447634$.
 
-Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
-Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
-information.
-While a human could manually identify the first link, to map the entire First Link network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
+The authors also acknowledge support from the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
 
-While some libraries exist for MediaWiki Markup,
-approaches using existing libraries led to several bugs 
-including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
-as well as compatibility with other libraries used to parse the XML.
-Consequently,  we developed an algorithm for parsing the first link in the XML version of each article.
-Our parsing algorithm aimed to: 
-1) accurately identify the first link among other page elements, and 
-2) efficiently do so---that is without needing for several passes through the data.
-To process an article in a single pass, we developed a hierarchical system of flags:
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/flags.pdf}  
-  \caption{
-    \textbf{Parsing Algorithm of Wikipedia's XML dump.}
-The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
-Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
-Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.}
-  \label{fig:parsing algorithm}
-\end{figure}
+\begin{thebibliography}{10}
 
-The algorithm loops in three-character chunks to account for potentially nested elements, 
-shifting by one character steps through the article markup.
-If any markup triggers for a flag are detected, a flag is raised. 
-Once a flag is raised, we stop processing and proceed to the next character
-until the flag's closing markup.
-A first link is identified only if Flags 1, 2, and 3 are all off.
-In this case, the entire link is retrieved. 
-We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
-as well as common file extensions for 
-images, audio files, and the like 
-\cite{media_wiki_templates}.
-The first link of an article is then the earliest valid link with unraised flags.
+\bibitem{clauson2008scope}
+K.~A. Clauson, H.~H. Polen, M.~N.~K. Boulos, and J.~H. Dzenowagis.
+\newblock Scope, completeness, and accuracy of drug information in wikipedia.
+\newblock {\em Annals of Pharmacotherapy}, 42(12):1814--1821, 2008.
 
-To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
-across 112 cores of the UVM supercomputer cluster
-\cite{vacc}.
-We then joined the results to form a hash table containing every Wikipedia article and its corresponding
-first link. The resulting network map is the basis of our analysis.
-An online appendix containing all of the code and data used can be accessed \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.
+\bibitem{stats}
+Wikipedia.
+\newblock Wikipedia:statistics. from wikipedia, the free encyclopedia.
+\newblock [Online; accessed 22-April-2016].
 
-\bibliography{wikipedia-network}{}
-\bibliographystyle{plain}
+\bibitem{wiki_edits}
+E.~Zachte.
+\newblock Wikipedia statistics: edits per month, April 2016.
+\newblock [Online; accessed 22-April-2016].
+
+\bibitem{wiki_views}
+Wikipedia.
+\newblock Page views for wikipedia, non-mobile site, normalized.
+\newblock [Online; accessed 22-April-2016].
+
+\bibitem{iba2010analyzing}
+T.~Iba, K.~Nemoto, B.~Peters, and P.~A. Gloor.
+\newblock Analyzing the creative editing behavior of wikipedia editors: Through
+  dynamic social network analysis.
+\newblock {\em Procedia-Social and Behavioral Sciences}, 2(4):6441--6456, 2010.
+
+\bibitem{samoilenko2016linguistic}
+A.~Samoilenko, F.~Karimi, D.~Edler, J.~Kunegis, and M.~Strohmaier.
+\newblock Linguistic neighbourhoods: explaining cultural borders on wikipedia
+  through multilingual co-editing activity.
+\newblock {\em EPJ Data Science}, 5(1):1, 2016.
+
+\bibitem{holman2008comparison}
+L.~Holman~Rector.
+\newblock Comparison of wikipedia and other encyclopedias for accuracy,
+  breadth, and depth in historical articles.
+\newblock {\em Reference services review}, 36(1):7--22, 2008.
+
+\bibitem{giles2005internet}
+J.~Giles.
+\newblock Internet encyclopaedias go head to head.
+\newblock {\em Nature}, 438(7070):900--901, 2005.
+
+\bibitem{halavais2008analysis}
+A.~Halavais and D.~Lackaff.
+\newblock An analysis of topical coverage of wikipedia.
+\newblock {\em Journal of Computer-Mediated Communication}, 13(2):429--440,
+  2008.
+
+\bibitem{hill2013wikipedia}
+B.~M. Hill and A.~Shaw.
+\newblock The wikipedia gender gap revisited: Characterizing survey response
+  bias with propensity score estimation.
+\newblock {\em PloS one}, 8(6):e65782, 2013.
+
+\bibitem{williams2015identifying}
+J.~R. Williams, E.~M. Clark, J.~P. Bagrow, C.~M. Danforth, and P.~S. Dodds.
+\newblock Identifying missing dictionary entries with frequency-conserving
+  context models.
+\newblock {\em Physical Review E}, 92(4):042808, 2015.
+
+\bibitem{banerjee2007clustering}
+S.~Banerjee, K.~Ramanathan, and A.~Gupta.
+\newblock Clustering short texts using wikipedia.
+\newblock In {\em Proceedings of the 30th annual international ACM SIGIR
+  conference on Research and development in information retrieval}, pages
+  787--788. ACM, 2007.
+
+\bibitem{gabrilovich2007computing}
+E.~Gabrilovich and S.~Markovitch.
+\newblock Computing semantic relatedness using wikipedia-based explicit
+  semantic analysis.
+\newblock In {\em IJCAI}, volume~7, pages 1606--1611, 2007.
+
+\bibitem{cucerzan2007large}
+S.~Cucerzan.
+\newblock Large-scale named entity disambiguation based on wikipedia data.
+\newblock In {\em EMNLP-CoNLL}, volume~7, pages 708--716, 2007.
+
+\bibitem{kamps2009wikipedia}
+J.~Kamps and M.~Koolen.
+\newblock Is wikipedia link structure different?
+\newblock In {\em Proceedings of the second ACM international conference on Web
+  search and data mining}, pages 232--241. ACM, 2009.
+
+\bibitem{page1999pagerank}
+L.~Page, S.~Brin, R.~Motwani, and T.~Winograd.
+\newblock The pagerank citation ranking: bringing order to the web.
+\newblock {\em Stanford InfoLab}, 1999.
+
+\bibitem{chakrabarti2001integrating}
+S.~Chakrabarti.
+\newblock Integrating the document object model with hyperlinks for enhanced
+  topic distillation and information extraction.
+\newblock In {\em Proceedings of the 10th international conference on World
+  Wide Web}, pages 211--220. ACM, 2001.
+
+\bibitem{wiki_train}
+Wikipedia.
+\newblock Train --- wikipedia{,} the free encyclopedia, 2016.
+\newblock [Online; accessed 22-April-2016].
+
+\bibitem{xkcd}
+xkcd.
+\newblock xkcd: Extended mind.
+\newblock [Online; accessed 22-April-2016].
+
+\bibitem{mat_blog}
+M.~Kelcey.
+\newblock do all first links on wikipedia lead to philosophy?
+\newblock [Online; accessed 22-April-2016].
+
+\bibitem{Ilmari_first_links}
+Wikipedia.
+\newblock User:ilmari karonen/first link.
+\newblock [Online; accessed 22-April-2016].
+
+\bibitem{bob_west}
+B.~West.
+\newblock Wikipedia's fixed point.
+\newblock [Online; accessed 22-April-2016].
+
+\bibitem{Bolton2007}
+M.~B. Bolton.
+\newblock The taxonomy of ideas in locke's essay.
+\newblock In L.~Newman, editor, {\em The Cambridge Companion to Locke's}.
+  Cambridge University Press, 2007.
+
+\bibitem{descartes}
+K.~Smith.
+\newblock Descartes' theory of ideas.
+\newblock In E.~N. Zalta, editor, {\em The Stanford Encyclopedia of
+  Philosophy}. Spring 2014 edition, 2014.
+
+\bibitem{aristotle}
+P.~Studtmann.
+\newblock Aristotle's categories.
+\newblock In E.~N. Zalta, editor, {\em The Stanford Encyclopedia of
+  Philosophy}. Summer 2014 edition, 2014.
+
+\bibitem{hist_thesaurus}
+O.~E. Dictionary.
+\newblock What is the historical thesaurus of the oed - oxford english
+  dictionary.
+\newblock [Online; accessed 22-April-2016].
+
+\bibitem{newman2003structure}
+M.~E. Newman.
+\newblock The structure and function of complex networks.
+\newblock {\em SIAM review}, 45(2):167--256, 2003.
+
+\bibitem{eubank2004modelling}
+S.~Eubank, H.~Guclu, V.~A. Kumar, M.~V. Marathe, A.~Srinivasan, Z.~Toroczkai,
+  and N.~Wang.
+\newblock Modelling disease outbreaks in realistic urban social networks.
+\newblock {\em Nature}, 429(6988):180--184, 2004.
+
+\bibitem{newman2002random}
+M.~E. Newman, D.~J. Watts, and S.~H. Strogatz.
+\newblock Random graph models of social networks.
+\newblock {\em Proceedings of the National Academy of Sciences}, 99(suppl
+  1):2566--2572, 2002.
+
+\bibitem{horton1945erosional}
+R.~E. Horton.
+\newblock Erosional development of streams and their drainage basins;
+  hydrophysical approach to quantitative morphology.
+\newblock {\em Geological society of America bulletin}, 56:275--370, 1945.
+
+\bibitem{dodds1999unified}
+P.~S. Dodds and D.~H. Rothman.
+\newblock Unified view of scaling laws for river networks.
+\newblock {\em Physical Review E}, 59(5):4865, 1999.
+
+\bibitem{garlaschelli2003universal}
+D.~Garlaschelli, G.~Caldarelli, and L.~Pietronero.
+\newblock Universal scaling relations in food webs.
+\newblock {\em Nature}, 423(6936):165--168, 2003.
+
+\bibitem{scientology}
+.~USA Today.~Gannett.
+\newblock Leah remini on scientology: What we learned from 20/20.
+\newblock [Online; accessed 22-April-2016].
+
+\bibitem{weddings}
+SoundVision.com.
+\newblock Wedding statistics in the united states.
+\newblock [Online; accessed 22-April-2016].
+
+\bibitem{wagner2015s}
+C.~Wagner, D.~Garcia, M.~Jadidi, and M.~Strohmaier.
+\newblock It's a man's wikipedia? assessing gender inequality in an online
+  encyclopedia.
+\newblock {\em arXiv preprint arXiv:1501.06307}, 2015.
+
+\bibitem{reddit}
+Reddit.
+\newblock Almost every wikipedia page links to philosophy if you keep clicking
+  the first link in every article.
+\newblock [Online; accessed 22-April-2016].
+
+\end{thebibliography}
+
+\clearpage
 

### 2016-04-26 writeup/wikipedia-network.tex
MLE power law for indegree + compilation success

@@
     % clarify basin discussion of method
-    % change name from traversal funnels to non-cycle traversal visits?
+    % cite power law papers (clauset and package)
 % Results
@@
 \end{figure}
+
 \begin{figure}[tp!]
@@
     \textbf{FLN Degree Distribution.}
-We fit the in-degree distribution with a linear $\log_{10}-\log_{10}$ model. 
-The result is a 
-an excellent fit with a Pearson's Correlation Coefficient of $-0.98$, yielding a 
-power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$. 
+    We plot the indegree distribution against rank for all articles with an indegree of at least 1 on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
+    distribution using the Maximum Likelihood Estimator with xmin as the indegree minimizing Kolmogorov-Smirnov distance.
+We find xmin to be 5 links with a corresponding power law rank exponent $\alpha = 2.4$ and a size exponent $\gamma = 1.4$.
 }
@@
 have fewer than 9. 
-We fit the in-degree distribution against each article's rank with a log-log (base 10) linear regression model
+We plot the in-degree distribution against each article's rank with a log-log (base 10) and fit the data to a power law 
+distribution using the Maximum Likelihood Estimator with xmin as the indegree minimizing Kolmogorov-Smirnov distance.
 (see Fig.~\ref{fig:degree distribution}).
-We obtain a Pearson's Correlation Coefficient of $-0.98$
-with a corresponding  power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$.
+We find xmin to be 5 links with a corresponding power law rank exponent $\alpha = 2.4$ and a size exponent $\gamma = 1.4$. 
+Finally, we compare the fit against alternative distributions using a Kolmogorov-Smirnov Test.
+We find a significant log likelihood ratio favoring a  truncated power law ($R < 0, p=4.5\times10^{-13}$) given the limits on the maximum possible indegree.
+We find an inconclusive log likelihood ratio for a lognormal distribution and significant log likelihood ratios favoring a power law fit to an exponential ($R  > 0, p=0$) and to a stretched exponential ($R > 0, p = 3.0\times10^{-50}$).
 
@@
   \label{fig:Distribution of Visits}
-
 \end{figure}

### 2016-04-27 writeup/2016-04wikipedia-network/wikipedia-network-revtex4-combined.tex
updated readme with compilation instructions and recompiled universal format

@@
 \begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{../graphics/Train.pdf}  
+  \includegraphics[width=\textwidth]{.../graphics/Train.pdf}  
   \caption{
@@
 \begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{../graphics/traversal_visit_algo_figure.pdf}  
+  \includegraphics[width=\textwidth]{.../graphics/traversal_visit_algo_figure.pdf}  
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/traversal_funnel_algo_figure.pdf}
+  \includegraphics[width=\columnwidth]{.../graphics/traversal_funnel_algo_figure.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/articles_ndegree.png}
+  \includegraphics[width=\columnwidth]{.../graphics/articles_ndegree.png}
   \caption{
     \textbf{Highest Ranking Articles by in-degree.}
-    We rank each article by the number of direct first links to the article (in-degree). The highest-ranking articles tend to represent geographical and biological abstractions. 
-}
+    We rank each article by the number of direct first links to the article (in-degree).
+The highest-ranking articles tend to represent geographical and biological abstractions.
+A full online appendix of the results and data is available \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.}
   \label{fig:indegree list}
 \end{figure}
+
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/ndegree_loglog.png}
+  \includegraphics[width=\columnwidth]{.../graphics/ndegree_loglog.png}
   \caption{
     \textbf{FLN Degree Distribution.}
-We fit the in-degree distribution with a linear $\log_{10}-\log_{10}$ model. 
-The result is a 
-an excellent fit with a Pearson's Correlation Coefficient of $-0.98$, yielding a 
-power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$. 
+    We plot the indegree distribution against rank for all articles with an indegree of at least 1 on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
+    distribution using the Maximum Likelihood Estimator with xmin as the indegree minimizing Kolmogorov-Smirnov distance.
+We find xmin to be 5 links with a corresponding power law rank exponent $\alpha = 2.4$ and a size exponent $\gamma = 1.4$.
 }
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/path_lengths_dist.png}
+  \includegraphics[width=\columnwidth]{.../graphics/path_lengths_dist.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/traversals_per_article.png} 
+  \includegraphics[width=\columnwidth]{.../graphics/traversals_per_article.png} 
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/articles_ranked.png}
+  \includegraphics[width=\columnwidth]{.../graphics/articles_ranked.png}
   \caption{
@@
 \begin{figure*}[tph!]
-  \includegraphics[width=\textwidth]{../graphics/2_and_3_cycles.pdf}
+  \includegraphics[width=\textwidth]{.../graphics/2_and_3_cycles.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/top_funnels.png}
+  \includegraphics[width=\columnwidth]{.../graphics/top_funnels.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/funnels_distribution.png}
+  \includegraphics[width=\columnwidth]{.../graphics/funnels_distribution.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/popularity_visits.pdf}
+  \includegraphics[width=\columnwidth]{.../graphics/popularity_visits.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/popularity_funnels.pdf}
+  \includegraphics[width=\columnwidth]{.../graphics/popularity_funnels.pdf}
   \caption{

### 2016-04-27 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
updated readme with compilation instructions and recompiled universal format

@@
     \textbf{Highest Ranking Articles by in-degree.}
-    We rank each article by the number of direct first links to the article (in-degree). The highest-ranking articles tend to represent geographical and biological abstractions. 
-}
+    We rank each article by the number of direct first links to the article (in-degree).
+The highest-ranking articles tend to represent geographical and biological abstractions.
+A full online appendix of the results and data is available \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.}
   \label{fig:indegree list}
 \end{figure}
+
 \begin{figure}[tp!]

### 2016-04-27 writeup/2016-04wikipedia-network/wikipedia-network-revtex4-combined.tex
power law fit and descriptions complete

--- 2016-04-26 writeup/2016-04wikipedia-network/wikipedia-network-revtex4-combined.tex
@@
 \usepackage{ifthen}
+\usepackage{hyperref}
+
+% Fonts and typesetting settings
+\usepackage[sc]{mathpazo}
+\usepackage[T1]{fontenc}
+\linespread{1.05} % Palatino needs more space between lines
+\usepackage{microtype}
+\usepackage{textcomp}
 
@@
 have fewer than 9. 
-We fit the in-degree distribution against each article's rank with a log-log (base 10) linear regression model
+We plot the in-degree distribution against each article's rank with a log-log (base 10) and fit the data to a power law 
+distribution using the Maximum Likelihood Estimator with xmin as the indegree minimizing Kolmogorov-Smirnov distance.
 (see Fig.~\ref{fig:degree distribution}).
-We obtain a Pearson's Correlation Coefficient of $-0.98$
-with a corresponding  power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$.
+We find xmin to be 5 links with a corresponding power law rank exponent $\alpha = 2.4$ and a size exponent $\gamma = 1.4$. 
+Finally, we compare the fit against alternative distributions using a Kolmogorov-Smirnov Test.
+We find a significant log likelihood ratio favoring a  truncated power law ($R < 0, p=4.5\times10^{-13}$) given the maximum indegree is limited by the number of articles in the FLN.
+We find an inconclusive log likelihood ratio for a lognormal distribution and significant log likelihood ratios favoring a power law fit against an exponential ($R  > 0, p=0$) and against a stretched exponential ($R > 0, p = 3.0\times10^{-50}$).
 
@@
     \textbf{Distribution of Traversal Visits.}
-We fit the distribution of traversal visits against article rank with a log-log (base 10) linear regression model log-log model and uncover two regimes: 
-with a cutoff rank of $10^{5}$.
-The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
-Coefficient of $-0.99$ 
-with a corresponding power law rank exponent 
-$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
+    We plot the number of traversal visits for each article by rank on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
+    distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance.
+We find xmin to be 866 with a corresponding power law rank exponent $\alpha = 1.7$ and a size exponent $\gamma = 1.6$.
 The horizontal flattening around the highest
-ranking articles is a result of the cyclic structure (see discussion on cycles).
-The bottom regime has a Pearson's Correlation Coefficient of $-0.93$
-with a corresponding power law rank exponent 
-$\alpha = -0.579$ and a size exponent $\gamma = -0.727$.}
+ranking articles is a result of the cyclic structure (see discussion of Cycles ~\ref{cycles}).
+}
   \label{fig:Distribution of Visits}
-
 \end{figure}
@@
 \end{figure*}
-In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against rank
-to two regimes corresponding to power law rank exponents of 
-$\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
-On the aggregate, the distribution suggests
+In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against a power law distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance.
+We find xmin to be 866 with a corresponding power law rank exponent $\alpha = 1.7$ and a size exponent $\gamma = 1.6$.
+We also compared the fit against alternative distributions and found signficiant likelihood ratios favoring a power law fit to an exponential ($R>0, p=2.5\times10^{-60}$) and stretched exponential ($R > 0, p=.0001$).
+Similarly to indegree, we found an inconclusive comparison against a lognormal distribution and a signficiant loglikelihood ratio favoring a truncated power law given the max number of traversal visits is restricted by the number of articles in the FLN.
+On the aggregate, the distribution of traversal visits suggests
 a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none.
@@
     \textbf{Distribution of Traversal Funnels.}
-We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed number of traversal funnels. 
-We find two regimes with the top regime (log(rank) < 4) 
-well-explained by a linear fit, yielding Pearson's 
-Correlation Coefficient of $-0.99$ and a 
-corresponding power law rank exponent $\alpha = -1.08$
-and a size exponent $\gamma = 0.074$. 
-The bottom horizontal regime corresponds to the more than $99\%$ of articles
-which hold zero traversal funnels.}
+    We plot the number of traversal funnels for each article by rank on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
+    distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance.
+We find xmin to be 30 with a corresponding power law rank exponent $\alpha = 1.73$ and a size exponent $\gamma = 1.57$.
+    The outlier on the top left is Philosophy with more traversal funnels than the second highest ranking article by two orders of magnitude.
+}
   \label{fig:Funnels Distribution}
@@
 
-Traversal funnels are comparatively resilient to changes in the FLN over time.
-While a few links can alter a cycle, potentially changing the number of traversal visits for an article, traversal funnels do not depend on cycles.
-Even when the particular links between articles change, the overall influence measured by traversal funnels is relatively resilient.
+In contrast to traversal visits, traversal funnels are comparatively resilient to changes in the FLN over time. 
+While a few links can alter a cycle, potentially displacing an article that was the recipient of references, traversal funnels remain stable even when cycles change.
 
@@
 the references flowing through the articles with at least one traversal funnel.
-When fit to a log-log linear model we find the $99\%$ of articles with zero
-traversal funnels form one regime (with $\log_{10}$(rank) less than 4).
-The top regime, corresponding to the $17, 821$ articles with at least one 
-traversal funnel strongly fits a linear model with
-a Pearson's Correlation Coefficient of $-0.99$ and a 
-corresponding power law rank exponent $\alpha = -1.08$
-and a size exponent $\gamma = 0.074$ 
-(see Fig.~\ref{fig:Funnels Distribution}). Even within the few articles
+When to a power law 
+    distribution (using the Maximum Likelihood Estimator) with xmin as the number of traversal funnels minimizing Kolmogorov-Smirnov distance), we find xmin to be 30 with a corresponding power law rank exponent $\alpha = 1.73$ and a size exponent $\gamma = 1.57$
+(see Fig.~\ref{fig:Funnels Distribution}). 
+When compared against alternative distributions, we find inconclusive results for lognormal and truncated power law distributions and significant log likelihood ratios favoring a power law for exponential ($R>0, p=5.3\times10^{-6}$) and stretched exponential ($R>0, p = 1.8\times10^{-8}$) distributions.
+Even within the few articles
 influencing the structure of the FLN, only a handful of these exert most of the 
@@
 
-\newpage
-
-\acknowledgments
-MI is grateful to RJ for pointing out the Reddit post 
-\cite{reddit}
-describing how the majority of links lead to ``Philosophy,'' inspiring this research.
-
-The authors are also grateful for the suggestions and input provided by Randall Harp.
-
-PSD and CMD acknowledge support from NSF Big Data Grant $\#1447634$.
-
-The authors also acknowledge support from the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
-
 \begin{thebibliography}{10}

### 2016-04-27 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
power law fit and descriptions complete

@@
+%% :%s/{../graphics/{..graphics
+
 \section{Introduction}
@@
 Finally, we compare the fit against alternative distributions using a Kolmogorov-Smirnov Test.
-We find a significant log likelihood ratio favoring a  truncated power law ($R < 0, p=4.5\times10^{-13}$) given the limits on the maximum possible indegree.
-We find an inconclusive log likelihood ratio for a lognormal distribution and significant log likelihood ratios favoring a power law fit to an exponential ($R  > 0, p=0$) and to a stretched exponential ($R > 0, p = 3.0\times10^{-50}$).
+We find a significant log likelihood ratio favoring a  truncated power law ($R < 0, p=4.5\times10^{-13}$) given the maximum indegree is limited by the number of articles in the FLN.
+We find an inconclusive log likelihood ratio for a lognormal distribution and significant log likelihood ratios favoring a power law fit against an exponential ($R  > 0, p=0$) and against a stretched exponential ($R > 0, p = 3.0\times10^{-50}$).
 
@@
     \textbf{Distribution of Traversal Visits.}
-We fit the distribution of traversal visits against article rank with a log-log (base 10) linear regression model log-log model and uncover two regimes: 
-with a cutoff rank of $10^{5}$.
-The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
-Coefficient of $-0.99$ 
-with a corresponding power law rank exponent 
-$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
+    We plot the number of traversal visits for each article by rank on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
+    distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance.
+We find xmin to be 866 with a corresponding power law rank exponent $\alpha = 1.7$ and a size exponent $\gamma = 1.6$.
 The horizontal flattening around the highest
-ranking articles is a result of the cyclic structure (see discussion on cycles).
-The bottom regime has a Pearson's Correlation Coefficient of $-0.93$
-with a corresponding power law rank exponent 
-$\alpha = -0.579$ and a size exponent $\gamma = -0.727$.}
+ranking articles is a result of the cyclic structure (see discussion of Cycles ~\ref{cycles}).
+}
   \label{fig:Distribution of Visits}
@@
 used to improve search engine algorithms or natural language processing.
-
-\acknowledgments
-MI is grateful to RJ for pointing out the Reddit post 
-\cite{reddit}
-describing how the majority of links lead to ``Philosophy,'' inspiring this research.
-
-The authors are also grateful for the suggestions and input provided by Randall Harp.
-
-PSD and CMD acknowledge support from NSF Big Data Grant $\#1447634$.
-
-The authors also acknowledge support from the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
-
-\newpage

### 2016-04-27 writeup/2016-04wikipedia-network/wikipedia-network-revtex4-combined.tex
draft complete

@@
-\documentclass[pre,twocolumn,twoside,byrevtex,superscriptaddress,floatfix]{revtex4-1}
+\documentclass[pre,twocolumn,twoside,superscriptaddress,floatfix]{revtex4-1}
 
@@
 the number of references directed towards a cycle. 
-
-We can also form and rank {\it basins} in the network by identifying groups of 
-path-connected articles, not necessarily forming a perfect cycle.
-A basin connects a group of articles and identifies the paths 
-to a particular article.
+We can also identify and rank {\it groups of 
+path-connected articles}, not necessarily forming a perfect cycle.
 
@@
 as a flow, 
-With accumulated references, cycles, and basins, we can explore how the many articles in Wikipedia are organized and related.
+With accumulated references, cycles, and path-connected groups, we can explore how the many articles in Wikipedia are organized and related.
 
@@
     \textbf{FLN Degree Distribution.}
-    We plot the indegree distribution against rank for all articles with an indegree of at least 1 on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
-    distribution using the Maximum Likelihood Estimator with xmin as the indegree minimizing Kolmogorov-Smirnov distance.
+    We plot the in-degree distribution against rank for all articles with an in-degree of at least 1 on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
+    distribution using the Maximum Likelihood Estimator with xmin as the in-degree minimizing Kolmogorov-Smirnov distance.
 We find xmin to be 5 links with a corresponding power law rank exponent $\alpha = 2.4$ and a size exponent $\gamma = 1.4$.
@@
 We plot the in-degree distribution against each article's rank with a log-log (base 10) and fit the data to a power law 
-distribution using the Maximum Likelihood Estimator with xmin as the indegree minimizing Kolmogorov-Smirnov distance.
-(see Fig.~\ref{fig:degree distribution}).
+distribution using the Maximum Likelihood Estimator with xmin as the in-degree minimizing Kolmogorov-Smirnov distance
+(see Fig.~\ref{fig:degree distribution}) 
+\cite{clauset2009power, alstott2014powerlaw}
+.
 We find xmin to be 5 links with a corresponding power law rank exponent $\alpha = 2.4$ and a size exponent $\gamma = 1.4$. 
-Finally, we compare the fit against alternative distributions using a Kolmogorov-Smirnov Test.
-We find a significant log likelihood ratio favoring a  truncated power law ($R < 0, p=4.5\times10^{-13}$) given the maximum indegree is limited by the number of articles in the FLN.
-We find an inconclusive log likelihood ratio for a lognormal distribution and significant log likelihood ratios favoring a power law fit against an exponential ($R  > 0, p=0$) and against a stretched exponential ($R > 0, p = 3.0\times10^{-50}$).
+We also compare the fit against alternative distributions using the log likelikhood ratio of a Kolmogorov-Smirnov Test 
+(see Sec.~\ref{power_law_comp}).
 
@@
 
-Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
 \begin{figure}[tp!]
@@
 \end{figure}
-
+Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
 This roughly corresponds to the official number of articles on Wikipedia: 
@@
 We also compared the fit against alternative distributions and found signficiant likelihood ratios favoring a power law fit to an exponential ($R>0, p=2.5\times10^{-60}$) and stretched exponential ($R > 0, p=.0001$).
-Similarly to indegree, we found an inconclusive comparison against a lognormal distribution and a signficiant loglikelihood ratio favoring a truncated power law given the max number of traversal visits is restricted by the number of articles in the FLN.
+Similarly to in-degree, we found an inconclusive comparison against a lognormal distribution and a signficiant loglikelihood ratio favoring a truncated power law given the max number of traversal visits is restricted by the number of articles in the FLN.
 On the aggregate, the distribution of traversal visits suggests
@@
 
-\subsection{Basins}
-
-The highest ranking basins (Sec.~\ref{Traversal Algorithm}) by the number of traversal visits are groups of articles around ``Philosophy.'' 
+\subsection{Path-Connected Articles}
+
+The highest ranking groups of path-connected articles (Sec.~\ref{Traversal Algorithm}) by the number of traversal visits are articles around ``Philosophy.'' 
 The highest ranking paths include branches of philosophy flowing through 
 ``Awareness,'' ``Existence,'' and ``Consciousness'' to ``Philosophy,'' 
-which form a cycle of seven articles. As of April 10, 2016 the basin flowing from 
+which form a cycle of seven articles. As of April 10, 2016 the path-connected group around
 ``Philosophy'' has changed to contain instead ``Ancient Greece,'' ``Greek,'' and 
@@
 
-Moving beyond basins around ``Philosophy,'' we find other basins around 
+Moving beyond ``Philosophy,'' we find other groups around 
 foundational concepts such as ``Community,'' ``Landmass,'' ``Federal Government,'' 
 ``Presentation,'' and ``Belief System.'' 
-The basins around each of these foundational notions are 
-various paths containing related articles. For example around 
-``Community'' we find basins
-flowing from ``United States'' to ``Federal Republic'' to ``Political Union'' to ``State'' culminating at ``Community''; we also find basins flowing from 
+The groups around each of these foundational notions are 
+contain closely related articles. For example the group connecting
+``Community''
+flows from ``United States'' to ``Federal Republic'' to ``Political Union'' to ``State'' culminating at ``Community''; we also find groups flowing from 
 ``Public Policy'' to ``Executive (government)'' to ``Government'' to ``State'' and then 
@@
 a community (or society) and then build up to ``Community.'' 
-Other basins around landmass for example begin at specific geographical regions
+Other path-connected groups around landmass for example begin at specific geographical regions
 such as ``Eastern Europe'' building up to ``Continent'' and finally ``Landmass''.
@@
 Nevertheless, many other foundational concepts emerged naturally within FLN. 
-Basins around ``Community'', ``State'', and ``Science'' reveal a foundational structure within the network. 
+Path connected-groups around ``Community'', ``State'', and ``Science'' reveal a foundational structure within the network. 
 More curious is the emergence of recently prominent political and economics topics such as ``Fossil Fuel'' and ``Health Care'' 
@@
 Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the first link.
+In addition, the work provides a general algorithm for assessing the structure and measuring influence in a directed network (or any braided flow structure).
 These findings also form the basis for the creation of a taxonomy where 
@@
 
+\bibitem{clauset2009power}
+A.~Clauset, C.~R. Shalizi, and M.~E. Newman.
+\newblock Power-law distributions in empirical data.
+\newblock {\em SIAM review}, 51(4):661--703, 2009.
+
+\bibitem{alstott2014powerlaw}
+J.~Alstott, E.~Bullmore, and D.~Plenz.
+\newblock powerlaw: a python package for analysis of heavy-tailed
+  distributions.
+\newblock {\em PloS one}, 9(1):e85777, 2014.
+
 \bibitem{scientology}

### 2016-04-27 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
draft complete

@@
 
-\subsection{Basins}
+\subsection{Path-Connected Articles}
 
-The highest ranking basins (Sec.~\ref{Traversal Algorithm}) by the number of traversal visits are groups of articles around ``Philosophy.'' 
+The highest ranking groups of path-connected articles (Sec.~\ref{Traversal Algorithm}) by the number of traversal visits are articles around ``Philosophy.'' 
 The highest ranking paths include branches of philosophy flowing through 
 ``Awareness,'' ``Existence,'' and ``Consciousness'' to ``Philosophy,'' 
-which form a cycle of seven articles. As of April 10, 2016 the basin flowing from 
+which form a cycle of seven articles. As of April 10, 2016 the path-connected group around
 ``Philosophy'' has changed to contain instead ``Ancient Greece,'' ``Greek,'' and 

### 2016-04-27 writeup/2016-04wikipedia-network/wikipedia-network.methods.tex
draft complete

--- 2016-04-22 writeup/wikipedia-network.tex
@@
-\documentclass[pre,twocolumn,twoside,superscriptaddress,floatfix, aps, 10pt]{revtex4-1}
-
-\usepackage{currfile}
-\lefthyphenmin=3
-\righthyphenmin=2
-
-\usepackage{graphicx,epsfig,verbatim,enumerate}
-\usepackage{amssymb,amsmath}
-\usepackage{ifthen}
-
-% Fonts and typesetting settings
-\usepackage[sc]{mathpazo}
-\usepackage[T1]{fontenc}
-\linespread{1.05} % Palatino needs more space between lines
-\usepackage{microtype}
-\usepackage{textcomp}
-
-\usepackage{longtable}
-
-\usepackage{mathtools}
-
-\newboolean{twocolswitch}
-
-\newcommand{\www}[1]{\url{#1}}
-\newcommand{\req}[1]{(\ref{#1})}
-\newcommand{\Req}[1]{Eq.~(\ref{#1})}
-
-% Lettrines
-\usepackage{lettrine}
-
-%useful shortcuts
-\def\R{\ensuremath{\mathbb{R}}} %\ensuremath adds math mode, if forgotten
-\def\Q{\ensuremath{\mathbb{Q}}}
-\def\N{\ensuremath{\mathbb{N}}}
-\def\Z{\ensuremath{\mathbb{Z}}}
-\def\C{\ensuremath{\mathbb{C}}}
-
-%shorcuts with arguments
-\newcommand{\abs}[1]{\left\vert#1\right\vert} %nice absolute values
-\newcommand{\bt}[1]{\textbf{#1}} %bold
-\newcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert} %vector norm
-\renewcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
-
-%piecewise function
-
-%\begin{displaymath}
-%   f(x) = \left\{
-%     \begin{array}{lr}
-%       1 & : x \in \mathbb{Q}\\
-%       0 & : x \notin \mathbb{Q}
-%     \end{array}
-%   \right.
-%\end{displaymath}
-
-%environment
-\newcommand{\tab}{\phantom{ssss}}
-
-\usepackage{hyperref}
-\usepackage{color}
-\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}
-
-%colors
-\definecolor{javagreen}{rgb}{0.25,0.5,0.35} %dark green color
-\definecolor{lightblue}{rgb}{0.149,0.545,0.824} %solarized blue
-\definecolor{sred}{rgb}{0.863, 0.196, 0.184} %solarized red
-
-\newcommand{\blue}[1]{{\leavevmode\color{lightblue}{#1}}} %solarized blue 
-\newcommand{\green}[1]{{\leavevmode\color{javagreen}{#1}}} %command for green
-\newcommand{\red}[1]{{\leavevmode\color{sred}{#1}}} %solarized red
-\newcommand{\gray}[1]{{\leavevmode\color[gray]{0.5}{#1}}} %gray text
-\newcommand{\Prob}[1]{{\rm Pr}\left(#1\right)}
-
-\setboolean{twocolswitch}{true}
-
-\begin{document}
-
-\title{\protect
-Connecting every bit of knowledge: \\
-The structure of Wikipedia's First Link Network
-}
-
-\date{\today}
-
-\author{
-\firstname{Mark}
-\surname{Ibrahim}
-}
-\email{mark.s.ibrahim@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Christopher}
-\surname{M. Danforth}
-}
-\email{chris.danforth@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\author{
-\firstname{Peter}
-\surname{Sheridan Dodds}
-}
-
-\email{peter.dodds@uvm.edu}
-
-\affiliation{Department of Mathematics \& Statistics, 
-    Computational Story Lab,  \\
-    Vermont Complex Systems Center,
-    Vermont Advanced Computing Core,
-    The University of Vermont, Burlington, VT 05401.}
-
-\begin{abstract}
-  \protect
-Apples, porcupines, and the most obscure Bob Dylan song---is every topic a few clicks from Philosophy? 
-Within Wikipedia, the surprising answer is yes: nearly all 
-paths lead to Philosophy.
-Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
-More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
-By following the first link in each article, we algorithmically construct a directed network of 
-all 4.7 million articles: Wikipedia's First Link Network.
-Here, we study the English edition of Wikipedia's First Link Network for insight into how the many 
-articles on inventions, places, people, objects, and events are related and organized.  
-By traversing every path, we measure the accumulation of first links, path lengths, basins, cycles, 
-and the influence each article exerts in shaping the network.
-We find scale-free distributions describe path length, 
-accumulation, and influence. Far from dispersed, first links disproportionately accumulate 
-at a few articles---flowing from specific to general and culminating around fundamental notions such as
-Community, State, and Science. 
-Philosophy directs more paths than any other article by two orders of magnitude. 
-We also observe a gravitation towards topical articles such as 
-Health Care and Fossil Fuel. 
-These findings enrich our view of the connections and structure of
-Wikipedia's ever growing store of knowledge.
-
-\end{abstract}
-
-\maketitle
-
-%TODO:
-% Formatting
-    % incorporate into template
-% Abstract 
-    % Our findings... braided network (not sure I understand what this means? braided networks don't come up in the paper)
-        % is the idea to say "we developed an algorithm that's generalizable to a broader class of networks?"
-% Method
-    % what to cite for drainage basin in geomorphology
-    % clarify basin discussion of method
-    % change name from traversal funnels to non-cycle traversal visits?
-% Results
-    % run MLE; what do i compare against?
-\section{Introduction}
-
-Wikipedia is a towering achievement of the Internet age. 
-At no point in history has a larger or more meticulously indexed collection of human knowledge 
-existed.
-Wikipedia contains 37 million articles in 283 languages, 
-with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
-\cite{clauson2008scope, stats}.
-Demonstrating its relevance to modern inquiry,
-Wikipedia is the sixth most visited
-site in the world, surpassing 18 billion page views and 10 million edits in January, 2013 alone.
-\cite{wiki_edits, wiki_views}.
-
-Wikipedia has naturally become the object of many studies. 
-Researchers have examined the cultural dynamics among editors
-\cite{iba2010analyzing, samoilenko2016linguistic},
-the accuracy of the content relative to traditional encyclopedias
-\cite{holman2008comparison, giles2005internet},
-the topics covered 
-\cite{halavais2008analysis},
-and bias against portions of the population
-\cite{hill2013wikipedia}.
-Wikipedia's content has also proven to be a powerful tool. 
-Researchers have used Wikipedia to identify missing dictionary entries
-\cite{williams2015identifying},
-cluster short text
-\cite{banerjee2007clustering},
-compute semantic relatedness
-\cite{gabrilovich2007computing},
-and disambiguate meaning \cite{cucerzan2007large}.
-
-While these many studies have dissected and fruitfully applied Wikipedia's content,
-here we examine the connections among the many articles.
-A hyperlink from one Wikipedia article to another
-naturally indicates a relationship between the two articles
-\cite{kamps2009wikipedia}.
-The notion that hyperlinks convey information about the content of a
-page has proved enormously successful in multiple domains from search engine algorithms 
-such as PageRank 
-\cite{page1999pagerank} 
-to topic classification
-\cite{chakrabarti2001integrating}.
-Here, we treat a hyperlink as a mechanism connecting two topics.
-
-The authors of a Wikipedia article choose where and whether to include a 
-reference to another Wikipedia article
-in the HTML markup.
-For example, as of November 2014, the authors of the 
-\href{https://en.wikipedia.org/wiki/Train}{``Train''}
-article
-had collectively chosen
-``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
-among others as relevant articles to reference in describing ``Train''
-\cite{wiki_train}
-.
-
-\begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
-  \caption{
-    \textbf{First Link Path For ``Train.''}
-    We follow the first link  to another Wikipedia article
-    in the main body of the article---the area
-    inside the green rectangle, which excludes 
-    side bar elements, the navigation bar and title; the first link is circled in red. 
-    In this example, the first link to another Wikipedia article is ``Rail Transport.'' 
-    We can again select the first link on the ``Rail Transport''
-    article, repeating the process to 
-    form a path of first links. After 11 links, we arrive at ``Philosophy.''
-  }
-  \label{fig:Train First Links}
-\end{figure*}
-
-By focusing our attention on the main body of an article---excluding
-elements in side bars and headings---we 
-attempt to systematically capture the core description of a topic.
-Within the body text, the first link marks the earliest moment in a topic's introduction where the 
-authors choose to directly reference another article. 
-While many links reference relevant details, 
-the first link is an association within a topic's initial 
-description. While 
-``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
-are links detailing particulars, the first link, ``rail transport,''
-is the topic the authors associate with ``Train'' in the introduction
-(Fig.~\ref{fig:Train First Links}).
-``Banana'' has a first link to ``fruit,'' ``Bob Dylan'' has a first link to 
-``Blowing in the Wind,'' and ``Physics'' has a first link to ``natural science.''
-Collectively, first links provide a pragmatic and interpretable 
-means to connect each article to another.
-
-By following the first hyperlink contained in all articles of the English edition of
-Wikipedia as of November, 2014, we form a directed network:
-{\it Wikipedia's First Link Network} (FLN).
-Inspired by the claim that the majority of first links lead to
-``Philosophy''---popularized by an 
-\href{https://xkcd.com/903/}{xkcd}
-comic and subsequently
-discussed in blog posts 
-\cite{xkcd, mat_blog, Ilmari_first_links, 
-bob_west}
----
-we holistically study how the FLN  yields a flow of connections.
-For methodological details, see Sec.~\ref{Appendix}.
-
-The FLN is a wealth of relations among inventions, places,
-figures, objects, and events across space and time.
-``Train'' for example, links to a parent node, ``Rail Transport,''
-while many child nodes such as ``Steel'' and ``Horsepower'' link to ``Train.''
-As shown in Fig.~\ref{fig:Train First Links}, the path starting at ``Train'' 
-contains ``Goods,'' ``Economics,'' ``Social Science,''
-leading ultimately to ``Philosophy''.
-Unlike previous taxonomies created by individuals
-\cite{Bolton2007, descartes, aristotle, hist_thesaurus},
-the relations in the FLN emerge without a centralized effort 
-as the aggregate of each article's authors choice of first link.
-
-Our goal is to study the structure of the FLN for insight into how the information on Wikipedia is organized and related.
-Informed by river network metrics,
-we consider the FLN as a flow, quantifying 
-the accumulation of first links around articles 
-and the influence an article exerts in shaping the FLN.
-Together with cycles, in-degree, depth, and the content of the articles, 
-we build our analysis of the relations among the ideas in Wikipedia.
-
-\section{Traversing the First Link Network}
-\label{Traversal Algorithm}
-
-An essential feature of a directed network's structure is the degree distribution 
-\cite{newman2003structure}. 
-The degree distribution has been used to study many phenomena from disease outbreak 
-\cite{eubank2004modelling} 
-to the dynamics of social networks 
-\cite{newman2002random}.
-The in-degree distribution in the FLN describes how many first links point to a 
-particular article. 
-Articles with zero in-degree have no references---they are outer leaves in the FLN 
-or are disconnected entirely.
-
-\begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{graphics/traversal_visit_algo_figure.pdf}  
-  \caption{
-    \textbf{Traversal Visit Algorithm on a sample network.}
-     The traversal visit vectors are an adjacency matrix for the paths through the network: 
-     the first column indicates the path formed starting with article A. 
-     The number of traversal visits for article A is then the number of paths containing A 
-     or the sum of the first row in our matrix:
-     $\sum_{i=1}^7 A_{\text{visit, i}} = 7$.
-  }
-  \label{fig:Traversal Visits}
-\end{figure*}
-
-Starting at each article, we construct a path through the FLN, 
-to map the flow of connections among articles.
-The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
-Previous studies have used flow to characterize the structure of river networks
-\cite{horton1945erosional, dodds1999unified} and describe the organization of 
-blood networks, food systems, and transportation networks
-\cite{garlaschelli2003universal}.
-
-The first metric we develop quantifies the accumulation of first links.
-The algorithm begins by selecting an article, then traversing the path formed
-by following the first links. Each time a first link references an article, we increment a count
-associated with the article. 
-We continue until a page is revisited or is invalid---defined here to mean pointing to a page outside of Wikipedia.
-We select a second article and repeat the process until we have 
-constructed a path for each article in the network. 
-We define the number of {\it traversal visits} of an article 
-to be the number of references flowing to the ideas in the article---equivalent
-to drainage basin area in geomorphology.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/traversal_funnel_algo_figure.pdf}
-  \caption{
-    \textbf{Traversal Funnel Algorithm on a simple network.}
-  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
-We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.
-For example
-  the number of traversal funnels for article $E$ is 
-  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
-  \label{fig:Traversal Funnels}
-
-\end{figure}
-
-We can characterize the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
-(Fig.~\ref{fig:Traversal Visits}), the path starting at article A is the 
-first column in the traversal visits matrix. 
-An entry of $1$ indicates the path contains a given article and 
-$0$ indicates the path does not.
-To compute the number of traversal visits for an article, we sum the corresponding row in our matrix.
-The traversal visits matrix for 
-our Wikipedia dataset consists of 121 million entries encoding each path out of the more than 
-googol possible paths ($\simeq4.7\times10^6\times2^{4.7\times10^6}$) through the FLN .
-
-By measuring the number of first links between
-two articles, we obtain an additional piece of information we call 
-{\it path length}, computed by summing 
-columns in the traversal visits matrix. 
-Path length describes how closely related topics are.
-Although ``Train'' is related to ``Economics'' for example, there are several articles
-bridging the connection:
-``Train'' is more specifically related to transportation, whose object
-is often goods. Goods are one of the fundamental objects of study in Economics.
-Described in links, this relationship is captured by a relatedness of $4$ 
-first links ultimately connecting ``Train'' to ``Economics.''
-
-One possible path through the FLN is a {\it cycle} or a group of articles 
-linking to one another inside a loop. In our sample network 
-(Fig.~\ref{fig:Traversal Visits}) 
-a 3-cycle exists among nodes A, B, and C. 
-We can readily identify
-the types of cycle structures within the FLN
-and rank each by 
-the number of references directed towards a cycle.
-
-We can also form and rank {\it basins} in the network by identifying groups of 
-path-connected articles, not necessarily forming a perfect cycle.
-A basin connects a group of articles and identifies the paths 
-to a particular article.
-
-While traversal visits measure accumulation, each article's first link also 
-influences the shape of the FLN. 
-At a point of accumulation, a single article's first link 
-can exert great influence over the shape of the FLN by directing many
-references on a particular path. To distinguish between an article 
-that simply happened to fall within a cycle from an article funneling 
-many first links, we develop a second metric called {\it traversal funnels}.
-To measure traversal funnels, we traverse the FLN in the same manner as we 
-did for traversal visits, but end a path once we enter a cycle.
-We are then able to distinguish between an article related to many other ideas
-only by virtue of its place in a cycle, from an article exerting influence over where the first links flow. 
-In our simple network 
-(see Fig.~\ref{fig:Traversal Funnels}), article C 
-directs the flow of links towards the 3-cycle, while articles A and B are 
-recipients of the flow---without exerting direct influence themselves.
-
-By considering the FLN not only as collection of directly linked pairs of articles, but
-as a flow, 
-With accumulated references, cycles, and basins, we can explore how the many articles in Wikipedia are organized and related.
-
-\section{Results}
-
-\subsection{Degree Distribution}
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
-  \caption{
-    \textbf{Highest Ranking Articles by in-degree.}
-    We rank each article by the number of direct first links to the article (in-degree).
-The highest-ranking articles tend to represent geographical and biological abstractions.
-A full online appendix of the results and data is available \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.}
-  \label{fig:indegree list}
-\end{figure}
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
-  \caption{
-    \textbf{FLN Degree Distribution.}
-We fit the in-degree distribution with a linear $\log_{10}-\log_{10}$ model. 
-The result is a 
-an excellent fit with a Pearson's Correlation Coefficient of $-0.98$, yielding a 
-power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$. 
-}
-  \label{fig:degree distribution}
-\end{figure}
-
-We rank all 11 million articles in the FLN by in-degree to find 
-the ``United States'' with 80,249 direct first links as the most referenced
-Wikipedia article 
-(Fig.~\ref{fig:indegree list}). 
-Other high-ranking articles
-include foundational abstract concepts such as ``village,'' ``species,''; 
-sports associations such as ``American Football,'' ``Association Football''; 
-and developed nations such as ``France,'' ``Japan,'' ``Russia,'' ``Australia,'' and 
-the ``Netherlands.'' These high ranking articles are useful abstractions: nations
-describe a collection of individuals with a common culture, language, or 
-geographical proximity; sports teams describe an ever changing collection of 
-players often associated with a cultural identity or a geographical 
-region. 
-Since abstractions such as nations and teams are inherently comprised
-of many parts, authors often reference the abstraction when describing a part.
-
-``Philosophy'' and other philosophical concepts
-are not among the highest-ranking articles by in-degree.
-``Philosophy'' has an in-degree of only 581, with direct first links from articles about Philosophers and areas of Philosophy: ``Existentialism and Humanism,'' ``Predeterminism,'' ''Synoptic Philosophy,'' ``Qualia,'' ``Dorothy Emmet,'' and ``Christopher W.
-Morris.''
-So while a great many FLN paths flow to 
-``Philosophy'' (see traversal visits discussion below), 
-the accumulation is not the 
-result of many articles directly referencing ``Philosophy.'' 
-Instead, first links flow towards ``Philosophy'' as the 
-ultimate anchor, by generalizing from specific to broad.
-
-The FLN's in-degree exhibits a decaying power law distribution where a few articles 
-receive most direct first references, while most articles receive few or none.
-The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5.
-Less than $1\%$ of articles have more than 100 direct first links and $75\%$ of articles
-have fewer than 9. 
-We fit the in-degree distribution against each article's rank with a log-log (base 10) linear regression model
-(see Fig.~\ref{fig:degree distribution}).
-We obtain a Pearson's Correlation Coefficient of $-0.98$
-with a corresponding  power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$.
-
-\subsection{Depth of the FLN}
-
-How many links does a connection among articles span? 
-We find the longest path length is 365,
-corresponding to the yearly calendar of Orthodox Liturgics.
-Each day's Liturgics links to the next day's. On the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
-(see discussion of cycles ~\ref{cycles}).
-We also find similarly lengthy paths following the evolution of a place or topic through time: 
-``1953 in Scotland'' or ``1560s Architecture,'' with articles sequentially proceeding by year, decade or era.
-In general, the longest paths connect temporally organized ideas.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/path_lengths_dist.png}
-  \caption{
-    \textbf{Path Length Distribution.}
-The network depth appears to have a bimodal distribution with a median of 29 first links in a path.
-The 365 cycle for ``Orthodox'' liturgics is the outlier to the right, with historical articles 
-describing nations such as ``Scotland'' and ``UK'' also lying outside the third quartile.
-More than $75\%$ of articles have path lengths between 
-$0$ and $50$ links.}
-  \label{fig:Path Length Distribution}
-\end{figure}
-
-Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
-  \caption{
-    \textbf{Distribution of Traversal Visits.}
-We fit the distribution of traversal visits against article rank with a log-log (base 10) linear regression model log-log model and uncover two regimes: 
-with a cutoff rank of $10^{5}$.
-The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
-Coefficient of $-0.99$ 
-with a corresponding power law rank exponent 
-$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
-The horizontal flattening around the highest
-ranking articles is a result of the cyclic structure (see discussion on cycles).
-The bottom regime has a Pearson's Correlation Coefficient of $-0.93$
-with a corresponding power law rank exponent 
-$\alpha = -0.579$ and a size exponent $\gamma = -0.727$.}
-  \label{fig:Distribution of Visits}
-
-\end{figure}
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/articles_ranked.png}
-  \caption{
-    \textbf{Highest ranking articles by number of traversal visits.}
-We compute the number of traversal visits for each article in the FLN (see 
-Sec.~\ref{Traversal Algorithm}). 
-In doing so, we can rank each article by the accumulation of first links.
-The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.}
-  \label{fig:highest visits}
-\end{figure}
-
-This roughly corresponds to the official number of articles on Wikipedia: 
-$~4.7$ million as of November 2014---approximately half of the 11 million 
-articles in the XML dump are redirects or disambiguations, not full articles.
-The most common path length is 29, with $75\%$ of path lengths ranging between 7--30 articles.
-As a distribution, more than $75\%$ of articles have a path length below 
-$50$ first links 
-while a few temporally organized paths exceed 50 links 
-(see Fig.~\ref{fig:Path Length Distribution}).
-
-\subsection{Traversal Visits}
-
-As a distribution, the number of traversal visits by article appears to follow a decaying power law. 
-The majority of articles have fewer than 30 traversal visits and
-first link references accumulate at a few articles.
-Specifically, $99.76\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
-Meanwhile, the highest ranking 30 articles have an extremely disproportionate number of traversal visits.
-
-\begin{figure*}[tph!]
-  \includegraphics[width=\textwidth]{graphics/2_and_3_cycles.pdf}
-  \caption{
-    \textbf{Highest ranking 2-cycles and 3-cycles.}
-We identify pairs of articles whose first links point to one another, forming
-a 2-cycle. We then rank each pair of articles by the total number of 
-traversal visits to gauge the most referenced groups of two articles linked
-to each other. We find 2-cycles often capture synonyms or articles representing nearly the 
-same concepts as opposed to distinct concepts. Similarly, we identify and 
-rank 3-cycles to find they appear to capture three closely related ideas or synonyms.}
-  \label{fig:cycles}
-\end{figure*}
-In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against rank
-to two regimes corresponding to power law rank exponents of 
-$\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
-On the aggregate, the distribution suggests
-a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none.
-The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.
-
-The highest ranking articles by traversal visits are broad topics spanning academic disciplines 
-or notions fundamental to society~(Fig. \ref{fig:highest visits}): ``Science,'' ``Mathematics,'' ``Geography,'' and ``Philosophy''
-as well as ``Community,'' ``State,'' ``Earth,'' ``Information,'' ``Power,'' and ``Communication.''
-The first links flow towards broader topics, accumulating at of these foundational notions.
-While ``Banana'' is a concrete fruit for example, 
-the first links flow from ``Fruit'' to ``Botany'' to ``Biology,'' and ultimately 
-culminate with ``Science'' and ``Philosophy.'' 
-We observe that first links direct the wealth of specific knowledge on Wikipedia to 
-a few foundational notions.
-
-\subsection{Network Cycles}
-\label{cycles}
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
-  \caption{
-    \textbf{Funnels.}
-We represent the highest-ranking articles by the number of traversal 
-funnels to gauge the influence each article exerts in shaping the 
-structure of the FLN. 
-We find ``Philosophy'' exerts an overwhelming proportion
-of the influence, with other abstract notions and topical concepts ranking
-next.}
-  \label{fig:Funnels}
-\end{figure}
-
-We first identify 2-cycles, meaning a pair of articles with first link pointing to one another.
-Of the 11 million articles, roughly 84,000 are members of 2-cycles. 
-The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
-``Health Care'' and ``Medical Case Management''; ``Broadcasting House'' and ``BBC''; and ``Secondary Education'' and ``Secondary School'' 
-(see Fig.~\ref{fig:cycles}).
-
-Outside of these prominent 2-cycles, the typical 2-cycle signals a connection between distinct, yet very closely related concepts. 
-For example, we observe link patterns such as inventor to product (``Voere'' to ``VEC-91''), event to organizer (``Poetry Bus Tour'' to ``Weave Books''), and book to author (``Anatomy of Britain'' to ``Anthony Sampson'').
-
-Similarly, 3-cycles capture a synonymous or close relation among 3 articles: ``Tree of life (Biology),'' ``Tree of life (disambiguation),'' 
-and ``Tree of life''; ``Cinema of India,'' ``Indian Cinema,'' and ``Telugu Cinema''
-(see Fig.~\ref{fig:cycles}).
-Once we extend our cycle size beyond a length of 6 however, 
-``Philosophy'' along with the remaining list of high ranking articles by traversal visits dominate.
-The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day.
-Other lengthy cycles span 60--75 articles including collections of articles on national histories such as ``Japanese Eras'' 
-or judicial bodies such as the ``Legislative Assembly of Ontario.''
-
-\subsection{Basins}
-
-The highest ranking basins (Sec.~\ref{Traversal Algorithm}) by the number of traversal visits are groups of articles around ``Philosophy.'' 
-The highest ranking paths include branches of philosophy flowing through 
-``Awareness,'' ``Existence,'' and ``Consciousness'' to ``Philosophy,'' 
-which form a cycle of seven articles. As of April 10, 2016 the basin flowing from 
-``Philosophy'' has changed to contain instead ``Ancient Greece,'' ``Greek,'' and 
-``Indo-European Languages'' (see network resilience Sec.~\ref{Traversal Funnels}).
-Other paths
-include concepts around ``Mathematics,'' ``Scientific,'' ``Experiments,'' 
-``Biology,'' and ``Fact.''
-These paths link many specific articles to ``Philosophy,'' each funneled through a particular domain.
-
-Moving beyond basins around ``Philosophy,'' we find other basins around 
-foundational concepts such as ``Community,'' ``Landmass,'' ``Federal Government,'' 
-``Presentation,'' and ``Belief System.'' 
-The basins around each of these foundational notions are 
-various paths containing related articles. For example around 
-``Community'' we find basins
-flowing from ``United States'' to ``Federal Republic'' to ``Political Union'' to ``State'' culminating at ``Community''; we also find basins flowing from 
-``Public Policy'' to ``Executive (government)'' to ``Government'' to ``State'' and then 
-to ``Community''; we also find paths flowing through a similar chain beginning
-with ``Democracy,'' another beginning with ``Constitution,'' another at 
-``Dictatorship,'' and so on. The articles build from specific means of organizing
-a community (or society) and then build up to ``Community.'' 
-Other basins around landmass for example begin at specific geographical regions
-such as ``Eastern Europe'' building up to ``Continent'' and finally ``Landmass''.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
-  \caption{
-    \textbf{Distribution of Traversal Funnels.}
-We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed number of traversal funnels. 
-We find two regimes with the top regime (log(rank) < 4) 
-well-explained by a linear fit, yielding Pearson's 
-Correlation Coefficient of $-0.99$ and a 
-corresponding power law rank exponent $\alpha = -1.08$
-and a size exponent $\gamma = 0.074$. 
-The bottom horizontal regime corresponds to the more than $99\%$ of articles
-which hold zero traversal funnels.}
-  \label{fig:Funnels Distribution}
-\end{figure}
-
-\subsection{Traversal Funnels}
-\label{Traversal Funnels}
-
-To analyze the influence an article exerts in shaping the 
-structure of the FLN, we compute the number of traversal funnels for each.
-Ranking articles by the number of traversal funnels we find 
-``Philosophy'' to be by far the highest ranking article with 
-$7.37$ million paths
-(see Fig.~\ref{fig:Funnels}).
-Of any article, the number of traversal funnels Philosophy holds exceeds 
-all others by more than two orders of magnitude.
-The ``Philosophy'' cycle which contains ``Existence,'' ``Awareness,'' ``Reality,'' 
-and similar articles accumulates the overwhelming proportion of its 
-references through ``Philosophy'': $7.37$ million of the $7.4$ million references
-are funneled through ``Philosophy''.
-Second on the list of highest-ranking articles by traversal funnels is 
-``Presentation'' with only $30$ thousand paths. Similarly abstract 
-concepts also rank highly such as ``Tree of life'' (30 thousand), 
-``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).
-
-Many high-ranking articles are remarkably topical, culturally and politically important concepts.
-For example, ``Health Care,'' a recently high-contested legislative topic appears high on the list---Google trends indicates an uncharacteristic spike in search frequency between August, 2009 and February, 2010.
-Other high ranking articles include key historical events such as the ``Cold War'' or critical scarce resource with recent 
-media discussion such as ``Fossil Fuel.'' 
-The highest-ranking list also includes ``Hip Hop,'' ``Cancer,'' and ``Web Page.''
-
-Traversal funnels are comparatively resilient to changes in the FLN over time.
-While a few links can alter a cycle, potentially changing the number of traversal visits for an article, traversal funnels do not depend on cycles.
-Even when the particular links between articles change, the overall influence measured by traversal funnels is relatively resilient.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/popularity_visits.pdf}
-  \caption{
-\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal visits.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal visits.
-While the most popular articles do not necessarily correspond to the articles
-with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-The greatest number of articles fall within roughly 1 million traversal visits and page views.
-Overall, the page views for the top 1000 articles articles by traversal visits appear to be log-normally distributed.
-}
-  \label{fig:Views and Visits}
-
-\end{figure}
-As a distribution, we find few articles influence the structure of the 
-FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
-more than $99\%$ with none---most articles are recipients of 
-the references flowing through the articles with at least one traversal funnel.
-When fit to a log-log linear model we find the $99\%$ of articles with zero
-traversal funnels form one regime (with $\log_{10}$(rank) less than 4).
-The top regime, corresponding to the $17, 821$ articles with at least one 
-traversal funnel strongly fits a linear model with
-a Pearson's Correlation Coefficient of $-0.99$ and a 
-corresponding power law rank exponent $\alpha = -1.08$
-and a size exponent $\gamma = 0.074$ 
-(see Fig.~\ref{fig:Funnels Distribution}). Even within the few articles
-influencing the structure of the FLN, only a handful of these exert most of the 
-control.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/popularity_funnels.pdf}
-  \caption{
-\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal funnels.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal funnels.
-}
-  \label{fig:Views and Funnels}
-\end{figure}
-\subsection{Article Popularity}
-
-Wikipedia released an API to measure article popularity by page views
-starting November 2015. 
-Page views add another dimension to our
-findings by constrasting
-the number of users who access a particular article against
-the number of traversal visits and funnels.
-We measure popularity as the total number 
-of page views in the English edition of Wikipedia in the month of 
-October 2015---the earliest full month for which the data is available. 
-We find the highest ranking 1000 articles (by traversal visits) have an average of
-$70$ thousand page views in October with high variation: the standard deviation 
-is $1.1$ million page views. 
-The number of page views has a skewed distribution with $75\%$ of articles
-reaching fewer than $73$ thousand page views in October.
-The article for ``United States'' has the most page 
-views of the highest ranking $1000$ articles by traversal visits with 
-$1.4$ million views in October. The next most popular articles are 
-``India,'' ``World War II,'' and the ``United Kingdom'' each with roughly 1 million page views. 
-The ``Philosophy'' article, despite outranking every article by traversal visits,
-has only $240$ thousand page views in October.
-
-We also analyze article popularity for the highest ranking 1000 articles by 
-traversal funnels. In October, the average page views per article is 
-$22$ thousand with a standard deviation of $95$ thousand views. The distribution
-is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
-``Halloween'' is the most popular article with $2.8$ million views in October,
-likely a result of Halloween falling in the month of October. 
-The second most popular article, ``Scientology'' ($463$ thousand views) appeared in popular news as actress Leah Remini announced she's leaving the Church of Scientology 
-\cite{scientology}.
-Other popular articles include ``Clint Eastwood'' ($341$ thousand views),
-and the ``Cold War'' ($298$ thousand views) although each has significantly fewer views compared to the views for ``Halloween''.
-Other standout popular October articles include ``24-hour'' and ``12-hour'' clocks likely due
-to Daylight savings in October and ``Marriage'' possibly due to the drop in the number of
-weddings in the months following October
-\cite{weddings}.
-``Philosophy'' which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears nowhere in the top 20
-by traversal visits with only $240$ thousand page views in October, 2015. ``Philosophy'' is a relatively popular article 
-among articles influencing the shape of the network, but less popular among 
-highly-ranked articles by accumulation.
-
-%========================Conclusion====================================
-
-\section{Concluding Remarks}
-
-The findings here should only be considered within the limitations of their context.
-We examined only the English edition of Wikipedia at a particular moment in time, 
-considering only the first link in in each article as a means to relate articles.
-Finally, Wikipedia, while the largest digital 
-collection of human knowledge, is rife with the biases of the many contributing editors
-\cite{wagner2015s}.
-Nevertheless, the findings do reveal
-generalizable relationships and point to foundational notions.
-
-Among our observations is the appearance of multiple scale-free distributions within the network. 
-Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
-articles are responsible for funneling most paths. When measured against the traversal funnels, 
-``Philosophy'' emerges as an exceptional article by orders of magnitude. 
-Nevertheless, many other foundational concepts emerged naturally within FLN. 
-Basins around ``Community'', ``State'', and ``Science'' reveal a foundational structure within the network. 
-More curious is the emergence of recently prominent political and economics topics such as ``Fossil Fuel'' and ``Health Care'' 
-within the highest ranking funnels. 
-Wikipedia seems to reflect not only timeless foundations, but also the topical (at least within English speaking society).
-
-Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the first link.
-These findings also form the basis for the creation of a taxonomy where 
-every idea, event, or object sits within a hierarchy of connected notions.
-The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
-Applications could range from an enhanced network of ideas to psychological insights into how humans form associations.
-Specifically, an ever-evolving reference of related hierarchical concepts could be 
-used to improve search engine algorithms or natural language processing.
-
-\acknowledgments
-MI is grateful to RJ for pointing out the Reddit post 
-\cite{reddit}
-describing how the majority of links lead to ``Philosophy,'' inspiring this research.
-
-The authors are also grateful for the suggestions and input provided by Randall Harp.
-
-PSD and CMD acknowledge support from NSF Big Data Grant $\#1447634$.
-
-The authors also acknowledge support from the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.
-
-\newpage
-
 \section{Appendix}
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{graphics/flags.pdf}  
+  \includegraphics[width=\columnwidth]{../graphics/flags.pdf}  
   \caption{
@@
 
-\bibliography{wikipedia-network}{}
-\bibliographystyle{plain}
+\subsection{Power Law Alternative Distributions}
+\label{power_law_comp}
 
-\end{document}
+Using the powerlaw Python package developed by Alstott, we apply MLE to fit a power law exponent and compare the fit to alternative distributions
+\cite{clauset2009power, alstott2014powerlaw}.
+For in-degree, we find a significant log likelihood ratio favoring a  truncated power law ($R < 0, p=4.5\times10^{-13}$) given the maximum in-degree is limited by the number of articles in the FLN.
+We find an inconclusive log likelihood ratio for a lognormal distribution and significant log likelihood ratios favoring a power law fit against an exponential ($R  > 0, p=0$) and against a stretched exponential ($R > 0, p = 3.0\times10^{-50}$).

### 2016-04-27 writeup/wikipedia-network.tex
draft complete

@@
 articles on inventions, places, people, objects, and events are related and organized.  
-By traversing every path, we measure the accumulation of first links, path lengths, basins, cycles, 
+By traversing every path, we measure the accumulation of first links, path lengths, cycles, groups of path-connected articles 
 and the influence each article exerts in shaping the network.
@@
 
-%TODO:
-% Formatting
-    % incorporate into template
-% Abstract 
-    % Our findings... braided network (not sure I understand what this means? braided networks don't come up in the paper)
-        % is the idea to say "we developed an algorithm that's generalizable to a broader class of networks?"
-% Method
-    % what to cite for drainage basin in geomorphology
-    % clarify basin discussion of method
-    % cite power law papers (clauset and package)
-% Results
-    % run MLE; what do i compare against?
 \section{Introduction}
@@
 
+\subsection{Power Law Alternative Distributions}
+\label{power_law_comp}
+
+Using the powerlaw Python package developed by Alstott, we apply MLE to fit a power law exponent and compare the fit to alternative distributions
+\cite{clauset2009power, alstott2014powerlaw}.
+For in-degree, we find a significant log likelihood ratio favoring a  truncated power law ($R < 0, p=4.5\times10^{-13}$) given the maximum in-degree is limited by the number of articles in the FLN.
+We find an inconclusive log likelihood ratio for a lognormal distribution and significant log likelihood ratios favoring a power law fit against an exponential ($R  > 0, p=0$) and against a stretched exponential ($R > 0, p = 3.0\times10^{-50}$).
+
 \bibliography{wikipedia-network}{}

### 2016-04-30 writeup/2016-04wikipedia-network/manual_submission/wikipedia-network-revtex4-combined.tex
arxiv ready

--- 2016-04-27 writeup/2016-04wikipedia-network/wikipedia-network-revtex4-combined.tex
@@
 articles on inventions, places, people, objects, and events are related and organized.  
-By traversing every path, we measure the accumulation of first links, path lengths, basins, cycles, 
+By traversing every path, we measure the accumulation of first links, path lengths,
+groups of path-connected articles, cycles,
 and the influence each article exerts in shaping the network.
@@
 \begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{../graphics/Train.pdf}  
+  \includegraphics[width=\textwidth]{fig001_figures/Train.pdf}  
   \caption{
@@
 The FLN is a wealth of relations among inventions, places,
-figures, objects, and events across space and time.
+{figures, objects, and events across space and time.
 ``Train'' for example, links to a parent node, ``Rail Transport,''
@@
 \begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{../graphics/traversal_visit_algo_figure.pdf}  
+  \includegraphics[width=\textwidth]{fig002_figures/traversal_visit_algo_figure.pdf}  
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/traversal_funnel_algo_figure.pdf}
+  \includegraphics[width=\columnwidth]{fig003_figures/traversal_funnel_algo_figure.pdf}
   \caption{
@@
 By considering the FLN not only as collection of directly linked pairs of articles, but
-as a flow, 
-With accumulated references, cycles, and path-connected groups, we can explore how the many articles in Wikipedia are organized and related.
+as a flow, with accumulated references, cycles, and path-connected groups, we can explore how the many articles in Wikipedia are organized and related.
 
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/articles_ndegree.png}
+  \includegraphics[width=\columnwidth]{fig004_figures/articles_ndegree.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/ndegree_loglog.png}
+  \includegraphics[width=\columnwidth]{fig005_figures/ndegree_loglog.png}
   \caption{
@@
     distribution using the Maximum Likelihood Estimator with xmin as the in-degree minimizing Kolmogorov-Smirnov distance.
-We find xmin to be 5 links with a corresponding power law rank exponent $\alpha = 2.4$ and a size exponent $\gamma = 1.4$.
+We find xmin to be 5 links with a corresponding power law rank exponent $\alpha \simeq 2.44$ and a size exponent $\gamma \simeq 1.41$.
 }
@@
 .
-We find xmin to be 5 links with a corresponding power law rank exponent $\alpha = 2.4$ and a size exponent $\gamma = 1.4$. 
+A power law distribution obeys $Y = r^{-\alpha}$ where $r$ is the rank and $\alpha$ is Zipf's rank
+exponent
+\cite{zipf1949human}
+. We can also examine the size distribution by measuring the number of articles, size $S$, above
+a particular in-degree, which is distributed according to: $S^{-\gamma}$ ($\gamma$ is the size exponent). 
+The two exponents obey the relationship: $\alpha = \frac{1}{\gamma -1}$.
+For in-degree, we find xmin to be 5 links with a corresponding power law rank exponent $\alpha \simeq 2.44$ and a size exponent $\gamma \simeq 1.41$. 
 We also compare the fit against alternative distributions using the log likelikhood ratio of a Kolmogorov-Smirnov Test 
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/path_lengths_dist.png}
+  \includegraphics[width=\columnwidth]{fig006_figures/path_lengths_dist.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/traversals_per_article.png} 
+  \includegraphics[width=\columnwidth]{fig007_figures/traversals_per_article.png} 
   \caption{
@@
     distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance.
-We find xmin to be 866 with a corresponding power law rank exponent $\alpha = 1.7$ and a size exponent $\gamma = 1.6$.
+We find xmin to be 866 with a corresponding power law rank exponent $\alpha \simeq 1.72$ and a size exponent $\gamma \simeq 1.58$.
 The horizontal flattening around the highest
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/articles_ranked.png}
+  \includegraphics[width=\columnwidth]{fig008_figures/articles_ranked.png}
   \caption{
@@
 \begin{figure*}[tph!]
-  \includegraphics[width=\textwidth]{../graphics/2_and_3_cycles.pdf}
+  \includegraphics[width=\textwidth]{fig009_figures/2_and_3_cycles.pdf}
   \caption{
@@
 In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against a power law distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance.
-We find xmin to be 866 with a corresponding power law rank exponent $\alpha = 1.7$ and a size exponent $\gamma = 1.6$.
+We find xmin to be 866 with a corresponding power law rank exponent $\alpha \simeq 1.72$ and a size exponent $\gamma \simeq 1.58$.
 We also compared the fit against alternative distributions and found signficiant likelihood ratios favoring a power law fit to an exponential ($R>0, p=2.5\times10^{-60}$) and stretched exponential ($R > 0, p=.0001$).
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/top_funnels.png}
+  \includegraphics[width=\columnwidth]{fig010_figures/top_funnels.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/funnels_distribution.png}
+  \includegraphics[width=\columnwidth]{fig011_figures/funnels_distribution.png}
   \caption{
@@
     distribution using the Maximum Likelihood Estimator with xmin as the number of traversal visits minimizing Kolmogorov-Smirnov distance.
-We find xmin to be 30 with a corresponding power law rank exponent $\alpha = 1.73$ and a size exponent $\gamma = 1.57$.
+We find xmin to be 30 with a corresponding power law rank exponent $\alpha \simeq 1.73$ and a size exponent $\gamma \simeq 1.58$.
     The outlier on the top left is Philosophy with more traversal funnels than the second highest ranking article by two orders of magnitude.
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/popularity_visits.pdf}
+  \includegraphics[width=\columnwidth]{fig012_figures/popularity_visits.pdf}
   \caption{
@@
 When to a power law 
-    distribution (using the Maximum Likelihood Estimator) with xmin as the number of traversal funnels minimizing Kolmogorov-Smirnov distance), we find xmin to be 30 with a corresponding power law rank exponent $\alpha = 1.73$ and a size exponent $\gamma = 1.57$
+    distribution (using the Maximum Likelihood Estimator) with xmin as the number of traversal funnels minimizing Kolmogorov-Smirnov distance), we find xmin to be 30 with a corresponding power law rank exponent $\alpha \simeq 1.73$ and a size exponent $\gamma \simeq 1.58$
 (see Fig.~\ref{fig:Funnels Distribution}). 
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/popularity_funnels.pdf}
+  \includegraphics[width=\columnwidth]{fig013_figures/popularity_funnels.pdf}
   \caption{
@@
 
+\newpage
+
+\section{Appendix}
+\label{Appendix}
+
+\subsection{Constructing The First Link Network}
+
+To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
+Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
+eliminating any statistical error due to sampling.
+We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
+The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
+and disambiguations.
+Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
+\cite{wiki_edits},
+our aim is to characterize the structure of the First Link Network.
+
+Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
+Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
+information.
+While a human could manually identify the first link, to map the entire First Link network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
+
+While some libraries exist for MediaWiki Markup,
+approaches using existing libraries led to several bugs 
+including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
+as well as compatibility with other libraries used to parse the XML.
+Consequently,  we developed an algorithm for parsing the first link in the XML version of each article.
+Our parsing algorithm aimed to: 
+1) accurately identify the first link among other page elements, and 
+2) efficiently do so---that is without needing for several passes through the data.
+To process an article in a single pass, we developed a hierarchical system of flags:
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{fig014_figures/flags.pdf}  
+  \caption{
+    \textbf{Parsing Algorithm of Wikipedia's XML dump.}
+The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
+Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
+Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.}
+  \label{fig:parsing algorithm}
+\end{figure}
+
+The algorithm loops in three-character chunks to account for potentially nested elements, 
+shifting by one character steps through the article markup.
+If any markup triggers for a flag are detected, a flag is raised. 
+Once a flag is raised, we stop processing and proceed to the next character
+until the flag's closing markup.
+A first link is identified only if Flags 1, 2, and 3 are all off.
+In this case, the entire link is retrieved. 
+We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
+as well as common file extensions for 
+images, audio files, and the like 
+\cite{media_wiki_templates}.
+The first link of an article is then the earliest valid link with unraised flags.
+
+To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
+across 112 cores of the UVM supercomputer cluster
+\cite{vacc}.
+We then joined the results to form a hash table containing every Wikipedia article and its corresponding
+first link. The resulting network map is the basis of our analysis.
+An online appendix containing all of the code and data used can be accessed \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.
+
+\subsection{Power Law Alternative Distributions}
+\label{power_law_comp}
+
+Using the powerlaw Python package developed by Alstott, we apply MLE to fit a power law exponent and compare the fit to alternative distributions
+\cite{clauset2009power, alstott2014powerlaw}.
+For in-degree, we find a significant log likelihood ratio favoring a  truncated power law ($R < 0, p=4.5\times10^{-13}$) given the maximum in-degree is limited by the number of articles in the FLN.
+We find an inconclusive log likelihood ratio for a lognormal distribution and significant log likelihood ratios favoring a power law fit against an exponential ($R  > 0, p=0$) and against a stretched exponential ($R > 0, p = 3.0\times10^{-50}$).
+
 \acknowledgments
@@
 
+\bibitem{zipf1949human}
+G.~K. Zipf.
+\newblock $\{$Human Behaviour and the Principle of Least-Effort$\}$.
+\newblock 1949.
+
 \bibitem{scientology}
@@
 
+\bibitem{media_wiki_templates}
+Wikimedia.
+\newblock Help:templates.
+\newblock [Online; accessed 22-April-2016].
+
+\bibitem{vacc}
+UVM.
+\newblock University of vermont.
+\newblock [Online; accessed 22-April-2016].
+
 \bibitem{reddit}

### 2016-04-30 writeup/2016-04wikipedia-network/wikipedia-network-revtex4-combined.tex
arxiv ready

@@
 \begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{../graphics/Train.pdf}  
+  \includegraphics[width=\textwidth]{figures/Train.pdf}  
   \caption{
@@
 \begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{../graphics/traversal_visit_algo_figure.pdf}  
+  \includegraphics[width=\textwidth]{figures/traversal_visit_algo_figure.pdf}  
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/traversal_funnel_algo_figure.pdf}
+  \includegraphics[width=\columnwidth]{figures/traversal_funnel_algo_figure.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/articles_ndegree.png}
+  \includegraphics[width=\columnwidth]{figures/articles_ndegree.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/ndegree_loglog.png}
+  \includegraphics[width=\columnwidth]{figures/ndegree_loglog.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/path_lengths_dist.png}
+  \includegraphics[width=\columnwidth]{figures/path_lengths_dist.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/traversals_per_article.png} 
+  \includegraphics[width=\columnwidth]{figures/traversals_per_article.png} 
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/articles_ranked.png}
+  \includegraphics[width=\columnwidth]{figures/articles_ranked.png}
   \caption{
@@
 \begin{figure*}[tph!]
-  \includegraphics[width=\textwidth]{../graphics/2_and_3_cycles.pdf}
+  \includegraphics[width=\textwidth]{figures/2_and_3_cycles.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/top_funnels.png}
+  \includegraphics[width=\columnwidth]{figures/top_funnels.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/funnels_distribution.png}
+  \includegraphics[width=\columnwidth]{figures/funnels_distribution.png}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/popularity_visits.pdf}
+  \includegraphics[width=\columnwidth]{figures/popularity_visits.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{../graphics/popularity_funnels.pdf}
+  \includegraphics[width=\columnwidth]{figures/popularity_funnels.pdf}
   \caption{
@@
 
+\newpage
+
+\section{Appendix}
+\label{Appendix}
+
+\subsection{Constructing The First Link Network}
+
+To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
+Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
+eliminating any statistical error due to sampling.
+We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
+The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
+and disambiguations.
+Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
+\cite{wiki_edits},
+our aim is to characterize the structure of the First Link Network.
+
+Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
+Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
+information.
+While a human could manually identify the first link, to map the entire First Link network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
+
+While some libraries exist for MediaWiki Markup,
+approaches using existing libraries led to several bugs 
+including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
+as well as compatibility with other libraries used to parse the XML.
+Consequently,  we developed an algorithm for parsing the first link in the XML version of each article.
+Our parsing algorithm aimed to: 
+1) accurately identify the first link among other page elements, and 
+2) efficiently do so---that is without needing for several passes through the data.
+To process an article in a single pass, we developed a hierarchical system of flags:
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{figures/flags.pdf}  
+  \caption{
+    \textbf{Parsing Algorithm of Wikipedia's XML dump.}
+The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
+Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
+Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.}
+  \label{fig:parsing algorithm}
+\end{figure}
+
+The algorithm loops in three-character chunks to account for potentially nested elements, 
+shifting by one character steps through the article markup.
+If any markup triggers for a flag are detected, a flag is raised. 
+Once a flag is raised, we stop processing and proceed to the next character
+until the flag's closing markup.
+A first link is identified only if Flags 1, 2, and 3 are all off.
+In this case, the entire link is retrieved. 
+We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
+as well as common file extensions for 
+images, audio files, and the like 
+\cite{media_wiki_templates}.
+The first link of an article is then the earliest valid link with unraised flags.
+
+To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
+across 112 cores of the UVM supercomputer cluster
+\cite{vacc}.
+We then joined the results to form a hash table containing every Wikipedia article and its corresponding
+first link. The resulting network map is the basis of our analysis.
+An online appendix containing all of the code and data used can be accessed \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.
+
+\subsection{Power Law Alternative Distributions}
+\label{power_law_comp}
+
+Using the powerlaw Python package developed by Alstott, we apply MLE to fit a power law exponent and compare the fit to alternative distributions
+\cite{clauset2009power, alstott2014powerlaw}.
+For in-degree, we find a significant log likelihood ratio favoring a  truncated power law ($R < 0, p=4.5\times10^{-13}$) given the maximum in-degree is limited by the number of articles in the FLN.
+We find an inconclusive log likelihood ratio for a lognormal distribution and significant log likelihood ratios favoring a power law fit against an exponential ($R  > 0, p=0$) and against a stretched exponential ($R > 0, p = 3.0\times10^{-50}$).
+
 \acknowledgments

### 2016-04-30 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
arxiv ready

@@
-%% :%s/{../graphics/{..graphics
+%% :%s/{figures/{..graphics
 
@@
 used to improve search engine algorithms or natural language processing.
+
+\newpage
+
+\section{Appendix}
+\label{Appendix}
+
+\subsection{Constructing The First Link Network}
+
+To map Wikipedia's First Link Network, we use the freely-available XML dump of the English edition of Wikipedia. 
+Rather than rely on a sample of articles from which to generalize, we opted to process the entirety of Wikipedia, 
+eliminating any statistical error due to sampling.
+We analyze the snapshot provided on November 2014, representing the state of Wikipedia at the time.
+The November raw dump consists of 11 million articles: 4.7 million unique articles along with redirects
+and disambiguations.
+Knowing Wikipedia is an ever-evolving project with 10 edits every second and 750 new articles per day on average
+\cite{wiki_edits},
+our aim is to characterize the structure of the First Link Network.
+
+Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page.
+Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
+information.
+While a human could manually identify the first link, to map the entire First Link network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.
+
+While some libraries exist for MediaWiki Markup,
+approaches using existing libraries led to several bugs 
+including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
+as well as compatibility with other libraries used to parse the XML.
+Consequently,  we developed an algorithm for parsing the first link in the XML version of each article.
+Our parsing algorithm aimed to: 
+1) accurately identify the first link among other page elements, and 
+2) efficiently do so---that is without needing for several passes through the data.
+To process an article in a single pass, we developed a hierarchical system of flags:
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{figures/flags.pdf}  
+  \caption{
+    \textbf{Parsing Algorithm of Wikipedia's XML dump.}
+The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida.
+Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements.
+Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.}
+  \label{fig:parsing algorithm}
+\end{figure}
+
+The algorithm loops in three-character chunks to account for potentially nested elements, 
+shifting by one character steps through the article markup.
+If any markup triggers for a flag are detected, a flag is raised. 
+Once a flag is raised, we stop processing and proceed to the next character
+until the flag's closing markup.
+A first link is identified only if Flags 1, 2, and 3 are all off.
+In this case, the entire link is retrieved. 
+We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
+as well as common file extensions for 
+images, audio files, and the like 
+\cite{media_wiki_templates}.
+The first link of an article is then the earliest valid link with unraised flags.
+
+To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
+across 112 cores of the UVM supercomputer cluster
+\cite{vacc}.
+We then joined the results to form a hash table containing every Wikipedia article and its corresponding
+first link. The resulting network map is the basis of our analysis.
+An online appendix containing all of the code and data used can be accessed \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{here}.
+
+\subsection{Power Law Alternative Distributions}
+\label{power_law_comp}
+
+Using the powerlaw Python package developed by Alstott, we apply MLE to fit a power law exponent and compare the fit to alternative distributions
+\cite{clauset2009power, alstott2014powerlaw}.
+For in-degree, we find a significant log likelihood ratio favoring a  truncated power law ($R < 0, p=4.5\times10^{-13}$) given the maximum in-degree is limited by the number of articles in the FLN.
+We find an inconclusive log likelihood ratio for a lognormal distribution and significant log likelihood ratios favoring a power law fit against an exponential ($R  > 0, p=0$) and against a stretched exponential ($R > 0, p = 3.0\times10^{-50}$).

### 2016-05-01 writeup/2016-04wikipedia-network/package/wikipedia-network-revtex4-combined.tex
moved images to figures directory and renamed extensions to pdf- working version for arxiv

--- 2016-04-30 writeup/2016-04wikipedia-network/manual_submission/wikipedia-network-revtex4-combined.tex
@@
 \newcommand{\sigamma}{1-\simonrho}
+\graphicspath{{figures/}}
 
@@
 \begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{fig001_figures/Train.pdf}  
+  \includegraphics[width=\textwidth]{fig001_Train.pdf}  
   \caption{
@@
 \begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{fig002_figures/traversal_visit_algo_figure.pdf}  
+  \includegraphics[width=\textwidth]{fig002_traversal_visit_algo_figure.pdf}  
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig003_figures/traversal_funnel_algo_figure.pdf}
+  \includegraphics[width=\columnwidth]{fig003_traversal_funnel_algo_figure.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig004_figures/articles_ndegree.png}
+  \includegraphics[width=\columnwidth]{fig004_articles_ndegree.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig005_figures/ndegree_loglog.png}
+  \includegraphics[width=\columnwidth]{fig005_ndegree_loglog.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig006_figures/path_lengths_dist.png}
+  \includegraphics[width=\columnwidth]{fig006_path_lengths_dist.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig007_figures/traversals_per_article.png} 
+  \includegraphics[width=\columnwidth]{fig007_traversals_per_article.pdf} 
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig008_figures/articles_ranked.png}
+  \includegraphics[width=\columnwidth]{fig008_articles_ranked.pdf}
   \caption{
@@
 \begin{figure*}[tph!]
-  \includegraphics[width=\textwidth]{fig009_figures/2_and_3_cycles.pdf}
+  \includegraphics[width=\textwidth]{fig009_2_and_3_cycles.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig010_figures/top_funnels.png}
+  \includegraphics[width=\columnwidth]{fig010_top_funnels.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig011_figures/funnels_distribution.png}
+  \includegraphics[width=\columnwidth]{fig011_funnels_distribution.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig012_figures/popularity_visits.pdf}
+  \includegraphics[width=\columnwidth]{fig012_popularity_visits.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig013_figures/popularity_funnels.pdf}
+  \includegraphics[width=\columnwidth]{fig013_popularity_funnels.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig014_figures/flags.pdf}  
+  \includegraphics[width=\columnwidth]{fig014_flags.pdf}  
   \caption{

### 2016-05-01 writeup/2016-04wikipedia-network/wikipedia-network-revtex4-combined.tex
moved images to figures directory and renamed extensions to pdf- working version for arxiv

@@
 \begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{figures/Train.pdf}  
+  \includegraphics[width=\textwidth]{Train.pdf}  
   \caption{
@@
 \begin{figure*}[tp!]
-  \includegraphics[width=\textwidth]{figures/traversal_visit_algo_figure.pdf}  
+  \includegraphics[width=\textwidth]{traversal_visit_algo_figure.pdf}  
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{figures/traversal_funnel_algo_figure.pdf}
+  \includegraphics[width=\columnwidth]{traversal_funnel_algo_figure.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{figures/articles_ndegree.png}
+  \includegraphics[width=\columnwidth]{articles_ndegree.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{figures/ndegree_loglog.png}
+  \includegraphics[width=\columnwidth]{ndegree_loglog.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{figures/path_lengths_dist.png}
+  \includegraphics[width=\columnwidth]{path_lengths_dist.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{figures/traversals_per_article.png} 
+  \includegraphics[width=\columnwidth]{traversals_per_article.pdf} 
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{figures/articles_ranked.png}
+  \includegraphics[width=\columnwidth]{articles_ranked.pdf}
   \caption{
@@
 \begin{figure*}[tph!]
-  \includegraphics[width=\textwidth]{figures/2_and_3_cycles.pdf}
+  \includegraphics[width=\textwidth]{2_and_3_cycles.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{figures/top_funnels.png}
+  \includegraphics[width=\columnwidth]{top_funnels.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{figures/funnels_distribution.png}
+  \includegraphics[width=\columnwidth]{funnels_distribution.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{figures/popularity_visits.pdf}
+  \includegraphics[width=\columnwidth]{popularity_visits.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{figures/popularity_funnels.pdf}
+  \includegraphics[width=\columnwidth]{popularity_funnels.pdf}
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{figures/flags.pdf}  
+  \includegraphics[width=\columnwidth]{flags.pdf}  
   \caption{

### 2016-05-01 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
moved images to figures directory and renamed extensions to pdf- working version for arxiv

@@
-%% :%s/{figures/{..graphics
+%% :%s/{{..graphics
 

### 2016-05-02 writeup/2016-04wikipedia-network/correspondence/2016-05-02compsci-biographies.tex
compsci submission

\documentclass[11pt]{article}

\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{url}
\usepackage{hyperref}

\usepackage{marvosym}

\usepackage[svgnames]{xcolor}
%% \definecolor{lightgrey}{rgb}{0.7,0.7,0.7}
%% \definecolor{grey}{rgb}{0.5,0.5,0.5}

\setlength{\oddsidemargin}{20pt}
\setlength{\topmargin}{-50pt}
\setlength{\textwidth}{6in}
\setlength{\textheight}{670pt}

\setlength{\parindent}{0pt}
\setlength{\parskip}{10pt}

\pagestyle{empty}

\renewcommand{\baselinestretch}{1.1}

\newcommand{\etal}{\textit{et al.}}

\begin{document}

\sffamily

\bigskip
\bigskip

Mark Ibrahim is an applied mathematician who's interested in exploring how humans structure and interact with information. Before studying applied mathematics, Mark received a B.A. from Hamilton College where he also studied philosophy, economics, English, and French. Mark has experience in industry assessing market risk using portfolio trading data. He is currently building a project to search videos with speech and a second project serving a network search engine powered by Wikipedia.

Chris Danforth is the Flint Professor of Mathematical, Natural, and Technical Sciences at the University of Vermont. He co-directs the Computational Story Lab, a group of applied mathematicians at the undergraduate, masters, PhD, and postdoctoral level working on large-scale, system problems in many fields including sociology, nonlinear dynamics, networks, ecology, and physics. His research has been covered by the New York Times, Science Magazine, and the BBC among others. Descriptions of his projects are available at his website: http://uvm.edu/∼cdanfort

Peter Sheridan Dodds is a Professor at the University of Vermont (UVM) working on system-level problems in many fields, ranging from sociology to physics. He is Director of UVM's Complex Systems Center, co-Director of UVM's Computational Story Lab, a visiting faculty fellow at the Vermont Advanced Computing Core, and is appointed to the Department of Mathematics and Statistics. He maintains general research and teaching interests in complex systems and networks with a current focus on sociotechnical and psychological phenomena including collective emotional states, contagion, language, and stories. His methods encompass large-scale data collection and analysis, large-scale sociotechnical experiments, and the formulation, analysis, and simulation of theoretical models. Dodds's training is in theoretical physics, mathematics, and electrical engineering with extensive formal postdoctoral and research experience in the social sciences. Dodds's foundational funding was an  NSF CAREER grant awarded by the Social and Economic Sciences Directorate.

%% {\small
%% Professor\\
%% Director of the Vermont Complex Systems Center\\
%% Co-Director, Computational Story Lab\\
%% Vermont Advanced Computing Core\\
%% Department of Mathematics and Statistics\\
%% University of Vermont\\
%%

% $\bullet$ 
% Department of Mathematics and Statistics\\
% $\bullet$ 
% Vermont Advanced Computing Center\\
% $\bullet$ 
% Complex Systems Center\\
% University of Vermont\\
% pdodds@uvm.edu
% $\bullet$
% http://www.uvm.edu/{$\sim$}pdodds

%% \textbf{Suggested reviewers and their relevant expertise:}

\end{document}

### 2016-05-02 writeup/2016-04wikipedia-network/correspondence/2016-05-02compsci-highlights.tex
compsci submission

--- 2016-05-02 writeup/2016-04wikipedia-network/correspondence/2016-05-02compsci-biographies.tex
@@
 
-Mark Ibrahim is an applied mathematician who's interested in exploring how humans structure and interact with information.
-Before studying applied mathematics, Mark received a B.A.
-from Hamilton College where he also studied philosophy, economics, English, and French.
-Mark has experience in industry assessing market risk using portfolio trading data.
-He is currently building a project to search videos with speech and a second project serving a network search engine powered by Wikipedia.
+Our major contributions and findings are as follows:
 
-Chris Danforth is the Flint Professor of Mathematical, Natural, and Technical Sciences at the University of Vermont.
-He co-directs the Computational Story Lab, a group of applied mathematicians at the undergraduate, masters, PhD, and postdoctoral level working on large-scale, system problems in many fields including sociology, nonlinear dynamics, networks, ecology, and physics.
-His research has been covered by the New York Times, Science Magazine, and the BBC among others.
-Descriptions of his projects are available at his website: http://uvm.edu/∼cdanfort
+\textbf{1.}
+We connect all 4.7 million Wikipedia articles in a directed network via the first link.
+The network reflects how the many inventions, places, figures, objects, and events are related and organized.
 
-Peter Sheridan Dodds is a Professor at the University of Vermont (UVM) working on system-level problems in many fields, ranging from sociology to physics.
-He is Director of UVM's Complex Systems Center, co-Director of UVM's Computational Story Lab, a visiting faculty fellow at the Vermont Advanced Computing Core, and is appointed to the Department of Mathematics and Statistics.
-He maintains general research and teaching interests in complex systems and networks with a current focus on sociotechnical and psychological phenomena including collective emotional states, contagion, language, and stories.
-His methods encompass large-scale data collection and analysis, large-scale sociotechnical experiments, and the formulation, analysis, and simulation of theoretical models.
-Dodds's training is in theoretical physics, mathematics, and electrical engineering with extensive formal postdoctoral and research experience in the social sciences.
-Dodds's foundational funding was an  NSF CAREER grant awarded by the Social and Economic Sciences Directorate.
+\textbf{2.}
+We develop a novel algorithm to map the structure of a directed network as a flow. Our measures quantify: \\
+\mbox{}\ \textbf{a.}
+where references accumulate\\
+\mbox{}\ \textbf{b.}
+groups of path-connected nodes\\
+\mbox{}\ \textbf{c.}
+the influence each node exerts in shaping the network.
+
+\textbf{3.}
+We identify a gravitation of references on Wikipedia from  specific to general, which culminates disproportionately around only a few fundamental notions.
+
+Our findings form the basis for a new network-based topic search engine with applications in natural language processing and cognitive psychology.
 

### 2016-05-02 writeup/2016-04wikipedia-network/correspondence/2016-05-02compsci.tex
compsci submission

--- 2016-05-02 writeup/2016-04wikipedia-network/correspondence/2016-05-02compsci-highlights.tex
@@
 
+\raisebox{-4pt}{
+%% \includegraphics[width=0.4\textwidth]{uvm-towerlogoK.png}
+%% \includegraphics[width=0.4\textwidth]{uvm-towerlogo2C.jpg}
+\includegraphics[width=0.2\textwidth]{roboctopus.png}
+%% \includegraphics[height=68pt]{psd_1.jpg}
+}
+% \textcolor{DarkSlateBlue}{
+{ \small
+\begin{tabular}[b]{l}
+Prof.\ Peter Sheridan Dodds\\
+Director of the Vermont Complex Systems Center\\
+Co-Director, Computational Story Lab\\
+%% \Letter\
+%% Department of Mathematics and Statistics\\
+%% University of Vermont\\
+210 Colchester Avenue, Burlington, VT, 05401, USA\\
+%% \Lightning
+%% \Email\
+pdodds@uvm.edu
+$\bullet$
+http://www.uvm.edu/{$\sim$}pdodds
+% \Mobilefone
+% +1-917-533-8135 % (cell)
+% $\bullet$
+% \Telefon 
+% +1-802-656-3089 % (home)
+\end{tabular}
+}
+% \raisebox{-16pt}{
+% \includegraphics[height=68pt]{psd_1.jpg}
+% }
+
 \bigskip
 \bigskip
+
+To the Editors,
+
+We are very pleased to submit our manuscript for consideration at the Journal of Computational Science:\\
+\textbf{``Connecting every bit of knowledge:
+The structure of Wikipedia\textquotesingle s First Link Network''.}
 
@@
 Our findings form the basis for a new network-based topic search engine with applications in natural language processing and cognitive psychology.
+
+We look forward to hearing your decision.
+
+Yours sincerely and on behalf of the manuscript's authors,
+
+\includegraphics[width=0.2\textwidth]{signature}\\
+Mark Ibrahim
 

### 2016-09-06 code/centrality/btw_centrality.py
betweeness centrality computation

@@ -1,0 +1,42 @@
+"""
+computes the betweeness centrality on 
+the sample directed graph in the paper
+"""
+
+import networkx as nx
+
+
+edge_list = [("A", "B"),
+            ("B", "C"),
+            ("C", "A"),
+            ("D", "C"),
+            ("E", "C"),
+            ("F", "E"),
+            ("G", "E")]
+
+
+G = nx.DiGraph()
+G.add_edges_from(edge_list)
+
+"""
+In [1]: nx.betweenness_centrality(G)
+Out[1]:
+{'A': 0.16666666666666666,
+ 'B': 0.03333333333333333,
+ 'C': 0.3,
+ 'D': 0.0,
+ 'E': 0.2,
+ 'F': 0.0,
+ 'G': 0.0}
+
+In [3]: nx.eigenvector_centrality_numpy(G)
+Out[3]:
+{'A': 0.577350269189626,
+ 'B': 0.5773502691896255,
+ 'C': 0.577350269189626,
+ 'D': -3.652158313422749e-17,
+ 'E': 1.7124591974895072e-17,
+ 'F': -3.807937337545159e-17,
+ 'G': -8.262461530561738e-18}
+ """
+

### 2016-09-07 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
traversal funnels v centrality measures

@@
 
+\subsection{Traversal Visits}
+\label{Traversal Visits}
+
 The first metric we develop quantifies the accumulation of first links.
@@
 to drainage basin area in geomorphology.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{traversal_funnel_algo_figure.pdf}
-  \caption{
-    \textbf{Traversal Funnel Algorithm on a simple network.}
-  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
-We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.
-For example
-  the number of traversal funnels for article $E$ is 
-  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
-  \label{fig:Traversal Funnels}
-
-\end{figure}
 
@@
 
+\subsection{Traversal Funnels}
+\label{Traversal Funnels}
+
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{traversal_funnel_algo_figure.pdf}
+  \caption{
+    \textbf{Traversal Funnel Algorithm on a simple network.}
+  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
+We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.
+For example
+  the number of traversal funnels for article $E$ is 
+  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
+  \label{fig:Traversal Funnels}
+
+\end{figure}
+
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{centrality_measures.pdf}
+  \caption{
+    \textbf{Centrality Measures}
+    The table above compares standard centrality measures on the sample network against traversal funnels.
+    }
+  \label{fig:Centrality Measures}
+
+\end{figure}
+
 While traversal visits measure accumulation, each article's first link also 
@@
 only by virtue of its place in a cycle, from an article exerting influence over where the first links flow. 
-In our simple network 
+In the sample network 
 (see Fig.~\ref{fig:Traversal Funnels}), article C 
@@
 
-By considering the FLN not only as collection of directly linked pairs of articles, but
+To compute traversal funnels on a directed network, begin at each node: 1.
+Traverse the edge to the next node 2.
+Increment the next node's count 3.
+Repeat while the next node is unvisited.
+The resulting count is each node's number of traversal funnels: the number of paths the node directs up to the start of a cycle.
+
+As a measure of influence, traversal funnels differs from traditional centrality measures such as degree centrality, betweenness centrality, and eigenvector centrality. 
+Degree centrality characterizes how many articles have a first link to a particular topic, restricting influence to a single link. 
+Betweenness centrality and eigenvalue centrality capture flow within the network more than a single link away.
+Unlike traversal funnels however, betweenness centrality and eigenvalue centrality reward nodes that simply happen to be in a cycle.
+In our sample network (see Fig.~\ref{fig:Centrality Measures}), node A is relatively central using both betweenness and eigenvector centrality although node A is simply the recipient of paths directed by node C.
+Eigenvector centrality even assigns the same centrality to nodes A, B, and C.
+In contrast, traversal funnels defines a path without a predetermined end node, following the progression of links to its culmination at the start of a cycle.
+Furthermore, traversal funnels provides an interpretable metric: a node's traversal funnels is the number of paths the node directs to the start of a cycle.
+Consequently, traversal funnels allow us to isolate the paths directed by node C without any spill over effect on other nodes in its cycle.
+
+In addition, traversal funnels extends to directed networks with more than one outward edge. In such a network, there may be multiple paths associated with a single starting node. 
+The same interpretation for traversal funnels as the number of paths a node directs to a cycle still holds. 
+In contrast to betweenness centrality, traversal funnels would not favor the shortest path over any other. Traversal funnels equally weighs all paths up to a cycle.
+
+For our exploration of the FLN, by considering the FLN not only as collection of directly linked pairs of articles, but
 as a flow, with accumulated references, cycles, and path-connected groups, we can explore how the many articles in Wikipedia are organized and related.

### 2016-09-07 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
typo
citation for xefer added

@@
 \cite{xkcd, mat_blog, Ilmari_first_links, 
-bob_west}
+bob_west, xefer2011philosophy}
 ---
@@
     The table above compares standard centrality measures on the sample network against traversal funnels.
+While betweenness centrality and eigenvector centrality both reward node A for its place in the cycle, traversal funnels only rewards node C for directing paths into the cycle.
     }
@@
 Traverse the edge to the next node 2.
-Increment the next node's count 3.
+Increment the next node's count by one 3.
 Repeat while the next node is unvisited.
@@
 
-For our exploration of the FLN, by considering the FLN not only as collection of directly linked pairs of articles, but
-as a flow, with accumulated references, cycles, and path-connected groups, we can explore how the many articles in Wikipedia are organized and related.
+By considering the FLN not only as collection of directly linked pairs of articles, but
+as a flow, with accumulated references, cycles, and path-connected groups, we explore how the many articles in Wikipedia are organized and related.
 
@@
 
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{popularity_visits.pdf}
-  \caption{
-\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal visits.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal visits.
-While the most popular articles do not necessarily correspond to the articles
-with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-The greatest number of articles fall within roughly 1 million traversal visits and page views.
-Overall, the page views for the top 1000 articles articles by traversal visits appear to be log-normally distributed.
-}
-  \label{fig:Views and Visits}
-
-\end{figure}
 As a distribution, we find few articles influence the structure of the 
@@
 
+\subsection{Article Popularity}
+
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{popularity_visits.pdf}
+  \caption{
+\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal visits.}
+We use the total number of page views provided by Wikipedia for the month
+of October 2015 to compare each article's popularity against traversal visits.
+Overall, more page views do not correspond to higher traversal visits---Wikipedia users do not tend to visit articles with a higher accumulation in the FLN more so than they visit others. 
+However, the variation in popularity appears to decrease as the number of traversal visits increases.
+The greatest number of articles fall within roughly 1 million traversal visits and page views.
+Furthermore, page views for the top 1000 articles articles by traversal visits appear to be log-normally distributed.
+}
+  \label{fig:Views and Visits}
+
+\end{figure}
 \begin{figure}[tp!]
@@
 of October 2015 to compare each article's popularity to traversal funnels.
+Similar to traversal visits, more page views do not correspond to higher traversal funnels---Wikipedia users do not tend to visit articles with a higher influence in the FLN more so than they visit others.
 }
@@
 \end{figure}
-\subsection{Article Popularity}
 
@@
 highly-ranked articles by accumulation.
+
+Overall, the contrast between page views and traversal visits or funnels highlights a difference between popularity and influence in the FLN.
+Although traversal funnels and visits point to influential articles where many topics culminate, Wikipedia users do not visit these articles more than others.
+What is popular, is not necessarily foundational, and vice versa.
 

### 2016-09-08 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
revisions addressed

@@
 Unlike traversal funnels however, betweenness centrality and eigenvalue centrality reward nodes that simply happen to be in a cycle. 
-In our sample network (see Fig.~\ref{fig:Centrality Measures}), node A is relatively central using both betweenness and eigenvector centrality although node A is simply the recipient of paths directed by node C.
+In our sample network (see Fig.~\ref{fig:Centrality Measures}), node A is has betweenness centrality of 0.17 and eigenvalue centrality of 0.58 although node A is simply the recipient of paths directed by node C.
 Eigenvector centrality even assigns the same centrality to nodes A, B, and C.
@@
 By considering the FLN not only as collection of directly linked pairs of articles, but
-as a flow, with accumulated references, cycles, and path-connected groups, we explore how the many articles in Wikipedia are organized and related.
+as a flow using traversal funnels, with accumulated references, cycles, and path-connected groups, we explore how the many articles in Wikipedia are organized and related.
 

### 2016-09-12 writeup/2016-04wikipedia-network/correspondence/2016-08-22-compsci-reviewers/2016-09-12-response.body.tex
reponse to reviewers

\textbf{We thank the reviewers for their well-considered comments and
suggestions.}

\textbf{As we detail below, we have worked hard to address all concerns in full,
  and we believe our manuscript has been strengthened.}

\section*{Reviewer 1}

%% \reviewercomment{1}{}
%% \reply{}

\reviewercomment{1}{This article analyses the Wikipedia(English) image and introduces a directed network based on the ``First Link'' hyperlink through which one article first references the other. Following that the authors apply several network-analytical techniques to analyze the structure of the relationships. In the process, they also introduce concepts in the form of traversal visits and traversal funnels that are relatively novel when compared to traditional network-analytical methods. Following that they present several results, most importantly characterizing the scale-free nature of the degree distribution and the traversal visits. Further they utilize the traversal funnels metric as a measure of influence and characterize the articles on this scale. Further they rightly point out that one of the chief reasons behind some of the observations is the movement from ``Specifics'' to ``General'' nature of the articles thus explaining the influential nature of articles such as
``Philosophy''.}

\reviewercomment{1}{The article is well motivated especially given the recent attention related to topic of the central position of the ``Philosophy'' article on Wikipedia, as evidenced by posts on Reddit and Quora. The authors take a principled approach in analyzing the first link network. The distinction between traversal and visits and traversal funnels very nicely discriminates the influential articles from the ``accidentally'' influential ones. The results related to the scale-free nature of some of the metrics are noteworthy. The presentation is very clear and I did not find any major typos.  Citations to existing work are adequate.}

\reviewercomment{1}{
While there are several positives about the article, it did not however come across as very novel. The analyses are in general limited to path-based techniques and there are no comparisons with traditional network analytics (measures such as various notions of centrality). For example while the notion traversal funnels is taken to represent the influence of the articles, it does not come with any analyses or comparison with traditional notions of influence. Also it was not clear from the discussions on article popularity as to what exactly is the message conveyed by figures 12 and 13.

Specifically I would like the authors to address the following two aspects:

(a)     Clarify a bit more with a small network or through some comparisons, the notion of the influence on a directed graph as measured by traversal funnels \\
(b)     Add more clarity by discussing the results in figs. 12 and 13.}

\reply{In response to the reviewer's suggestions, we added a section comparing traversal funnels against traditional centrality measures such as degree centrality, betweenness centrality, and eigenvector centrality. We also included a table contrasting the centrality given by each method on a sample network to demonstrate the utility of traversal funnels as a measure of influence.

    In addition, we clarified the captions of figs. 12 and 13  and added a paragraph in our discussion of article popularity. We highlight the difference between popularity and the traversal results which aim to characterize foundational notions.
}

\section*{Reviewer 2}

\reviewercomment{1}{This paper examines the structure of the ``First Link'' network in Wikipedia.  They begin with the idea from an xkcd comic that claims that every path ends up at Philosophy, and observe that this is mostly true.  They analyze various other features of the graph, such as cycles, traversal funnels, etc.  The paper is largely descriptive.}

\reviewercomment{1}{Here are my major concerns:

1.  This paper is largely descriptive, and the originality seems low.  This idea has clearly been studied before, although admittedly not in scholarly publications.  Confirming these informal previous results does have value, but I would expect to see substantially more.  The notion of looking at how many paths use a node is not new (isn't this just betweenness centrality?), and the authors of the cited blog posts already looked at cycles.  I do like the notion of traversal funnels.

2.  Readers of this journal expect to see a new method/metric/etc. that can be generalized to other datasets.  In its current state, this paper might be a better fit in a computational social science journal.}

\reply{In response to the reviewer's suggestions, we elaborated on traversal funnels as a new generalizable method for measuring influence in a directed network. 
    We compare traversal funnels against traditional centrality measures such as degree centrality, betweenness centrality, and eigenvector centrality, highlighting the utility of traversal funnels as a new measure of influence.

    Specifically, traversal funnels isolates nodes directing paths into a cycle without spill over to other nodes in the cycle. Furthermore, traversal funnels provides an interpretable measure of influence in terms of paths. We include a table demonstrating these differences  on a sample network with a cycle. We also discuss how traversal funnels can be generalized to directed networks with more than a single outward edge.
}

\reviewercomment{1}{This paper has potential, and I like the ideas in it, but it doesn't contain enough novel content and is not generalizable.  It comes across as a sequence of observations about this particular network. These observations are certainly interesting, but are not enough for a journal paper.  Some suggestions:

1.  Can any of these concepts be generalized to other datasets? \\
}
\reply{As noted above, we include a new section discussing traversal funnels as a generalizable method for measuring influence in a directed network, including networks with more than a single outward edge.}

\reviewercomment{1}{
2.  What happens if you look past just the first link in a wikipedia article?  Say, the first k links?  How is the first link network different from the network generated by looking at all of the links?  Is this notion of considering only the single ``most important'' out-edge something that can be useful in other networks?  How does it affect, say, the community structure of the network?  If you begin with the complete link network, is it possible to somehow identify the first links (i.e., do they have some special structural characteristics)? \\
}
\reply{
    While examining the first k links would surely yield interesting results, we restrict our attention here to the first link in the main body text, noting its importance in anchoring a specific article to the next most general topic. Given our intent to characterize how topics in Wikipedia's First Link Network flow, the first link captures a more appropriate path compared to links appearing later in an article, which may link to detailed subtopics. 
}

\reviewercomment{1}{
3.  As is, the paper comes across as very descriptive with little new in the way of methods.   The idea of the traversal funnel seems like a good candidate for a new metric that is of general interest- can you look at how this concept works on other datasets?  Can you introduce an efficient algorithm (which doesn't traverse all paths) for finding these nodes?  Can you generalize to the case when nodes may have more than one out-edge?}

\reply{As noted above, we believe traversal funnels is a new generalizable measure of influence in a directed network. We include a comparison of traversal funnels against traditional centrality measures on a sample network and describe how traversal funnels can be generalized to networks with more than a single outward edge.}

### 2016-09-15 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
complete revisions and response to reviewers

@@
 Unlike traversal funnels however, betweenness centrality and eigenvalue centrality reward nodes that simply happen to be in a cycle. 
-In our sample network (see Fig.~\ref{fig:Centrality Measures}), node A is has betweenness centrality of 0.17 and eigenvalue centrality of 0.58 although node A is simply the recipient of paths directed by node C.
+We compute betweenness centrality and eigenvalue centrality on the sample network using \href{https://networkx.github.io/}{NetworkX}, a Python package for modelling complex networks (see the \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{online appendix} for details).
+In our sample network (see Fig.~\ref{fig:Centrality Measures}), node A has betweenness centrality of 0.17 and eigenvalue centrality of 0.58, although node A is simply the recipient of paths directed by node C.
 Eigenvector centrality even assigns the same centrality to nodes A, B, and C.
+
 In contrast, traversal funnels defines a path without a predetermined end node, following the progression of links to its culmination at the start of a cycle.
 Furthermore, traversal funnels provides an interpretable metric: a node's traversal funnels is the number of paths the node directs to the start of a cycle.
-Consequently, traversal funnels allow us to isolate the paths directed by node C without any spill over effect on other nodes in its cycle.
+Consequently, traversal funnels allow us to isolate the paths directed by node C, of which there are 4, without any spill over effect on other nodes in its cycle: A and B have zero traversal funnels.
 

### 2016-09-15 writeup/2016-04wikipedia-network/correspondence/2016-08-22-compsci-reviewers/2016-09-12-response.body.tex

@@
   and we believe our manuscript has been strengthened.}
+
+\textbf{Regarding the novelty of our work, while there have been a handful of explorations described on the internet (e.g.
+in blog posts), to our knowledge we are providing the first academic description of the first-link network.}
+
+\textbf{
+In our analysis of the network's structure, we also develop traveral funnels, a method for characterizing influence in a directed network.
+}
 

### 2016-09-15 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
changed figure placement

@@
 \end{figure*}
-
 Starting at each article, we construct a path through the FLN, 
@@
 
-\subsection{Traversal Funnels}
-\label{Traversal Funnels}
-
 \begin{figure}[tp!]
@@
 \end{figure}
+\subsection{Traversal Funnels}
+\label{Traversal Funnels}
 
@@
 \section{Results}
-
 \subsection{Degree Distribution}
@@
 \end{figure}
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{ndegree_loglog.pdf}
-  \caption{
-    \textbf{FLN Degree Distribution.}
-    We plot the in-degree distribution against rank for all articles with an in-degree of at least 1 on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
-    distribution using the Maximum Likelihood Estimator with xmin as the in-degree minimizing Kolmogorov-Smirnov distance.
-We find xmin to be 5 links with a corresponding power law rank exponent $\alpha \simeq 2.44$ and a size exponent $\gamma \simeq 1.41$.
-}
-  \label{fig:degree distribution}
-\end{figure}
-
 We rank all 11 million articles in the FLN by in-degree to find 
@@
 
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{ndegree_loglog.pdf}
+  \caption{
+    \textbf{FLN Degree Distribution.}
+    We plot the in-degree distribution against rank for all articles with an in-degree of at least 1 on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
+    distribution using the Maximum Likelihood Estimator with xmin as the in-degree minimizing Kolmogorov-Smirnov distance.
+We find xmin to be 5 links with a corresponding power law rank exponent $\alpha \simeq 2.44$ and a size exponent $\gamma \simeq 1.41$.
+}
+  \label{fig:degree distribution}
+\end{figure}
+
 \subsection{Depth of the FLN}
-
-How many links does a connection among articles span? 
-We find the longest path length is 365,
-corresponding to the yearly calendar of Orthodox Liturgics.
-Each day's Liturgics links to the next day's. On the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
-(see discussion of cycles ~\ref{cycles}).
-We also find similarly lengthy paths following the evolution of a place or topic through time: 
-``1953 in Scotland'' or ``1560s Architecture,'' with articles sequentially proceeding by year, decade or era.
-In general, the longest paths connect temporally organized ideas.
 
@@
 \end{figure}
+How many links does a connection among articles span? 
+We find the longest path length is 365,
+corresponding to the yearly calendar of Orthodox Liturgics.
+Each day's Liturgics links to the next day's. On the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
+(see discussion of cycles ~\ref{cycles}).
+We also find similarly lengthy paths following the evolution of a place or topic through time: 
+``1953 in Scotland'' or ``1560s Architecture,'' with articles sequentially proceeding by year, decade or era.
+In general, the longest paths connect temporally organized ideas.
+
+Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
+This roughly corresponds to the official number of articles on Wikipedia: 
+$~4.7$ million as of November 2014---approximately half of the 11 million 
+articles in the XML dump are redirects or disambiguations, not full articles.
+The most common path length is 29, with $75\%$ of path lengths ranging between 7--30 articles.
+As a distribution, more than $75\%$ of articles have a path length below 
+$50$ first links 
+while a few temporally organized paths exceed 50 links 
+(see Fig.~\ref{fig:Path Length Distribution}).
 
@@
 \end{figure}
-Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
-This roughly corresponds to the official number of articles on Wikipedia: 
-$~4.7$ million as of November 2014---approximately half of the 11 million 
-articles in the XML dump are redirects or disambiguations, not full articles.
-The most common path length is 29, with $75\%$ of path lengths ranging between 7--30 articles.
-As a distribution, more than $75\%$ of articles have a path length below 
-$50$ first links 
-while a few temporally organized paths exceed 50 links 
-(see Fig.~\ref{fig:Path Length Distribution}).
 
@@
 \label{cycles}
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{top_funnels.pdf}
-  \caption{
-    \textbf{Funnels.}
-We represent the highest-ranking articles by the number of traversal 
-funnels to gauge the influence each article exerts in shaping the 
-structure of the FLN. 
-We find ``Philosophy'' exerts an overwhelming proportion
-of the influence, with other abstract notions and topical concepts ranking
-next.}
-  \label{fig:Funnels}
-\end{figure}
 
@@
 \begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{top_funnels.pdf}
+  \caption{
+    \textbf{Funnels.}
+We represent the highest-ranking articles by the number of traversal 
+funnels to gauge the influence each article exerts in shaping the 
+structure of the FLN. 
+We find ``Philosophy'' exerts an overwhelming proportion
+of the influence, with other abstract notions and topical concepts ranking
+next.}
+  \label{fig:Funnels}
+\end{figure}
+
+\subsection{Traversal Funnels}
+\label{Traversal Funnels}
+
+\begin{figure}[tp!]
   \includegraphics[width=\columnwidth]{funnels_distribution.pdf}
@@
 \end{figure}
-
-\subsection{Traversal Funnels}
-\label{Traversal Funnels}
 

### 2016-11-21 writeup/2016-04wikipedia-network/correspondence/2016-11-06-compsci-reviewers/2016-11-06-response.body.tex
added pesudo-code and fixed centrality table

\textbf{We thank the reviewers for their well-considered comments and
suggestions.}

\textbf{As we detail below, we have worked hard to address all concerns in full,
  and we believe our manuscript has been strengthened.}

\section*{Reviewer 1}

%% \reviewercomment{1}{}
%% \reply{}

\reviewercomment{1}{
I am not fully satisfied with the updates that the authors have provided. After taking into account the comments by Reviewer 2, many of which are valid concerns, I would like to see another revision addressing the following two points.

(a) Present in an algorithm form,  the steps that the authors used to compute the "Traversal-Funnel" based influence measure instead of a descriptive discussion. Comment on the running time complexity of the same.

(b) Please double check the eigenvector centrality values. Normally the eigenvector corresponding to the top eigenvalue is selected and the entires are all non-negative. There are some negative entries present in the values shown.
}

\reply{In response to the reviewer's suggestions, we:

(a) Included pseudo-code for the traversal funnels algorithm as well as a description of the runtime and auxiliary memory usage in section B. Traversal Funnels. We have also made all code used for this work available in the online appendix.

(b) Re-rendered the Centrality Measures table (Fig. 4) to reflect the accurate, non-negative eigenvector centrality values. The previous submission contained incorrectly rendered scientific notation showing negative values. The true values included in our source code (see online appendix) are effectively zero:
 D: $-3.65*10^{-17}$, 
 F: $-3.81*10^{-17}$, 
 and 
G: $-8.26*10^{-18}$.
The Centrality Measures table (Fig 4) now correctly displays the true values.
}

\section*{Reviewer 2}

\reviewercomment{1}{
I am satisfied with the modifications made and the author response.  I would suggest adding something to the introduction pointing out that you are introducing the new notion of traversal funnels, and make it a bit more prominent that this is a novel measure you have created.
}

\reply{
In response to the reviewer's suggestion, we more prominently highlighted traversal funnels as new measure of influence in various sections including the abstract, introduction, and conclusion.
}

### 2016-11-21 writeup/2016-04wikipedia-network/wikipedia-network.abs.tex
added pesudo-code and fixed centrality table

--- 2016-04-22 writeup/2016-04wikipedia-network/wikipedia-network.abs.tex
@@
 articles on inventions, places, people, objects, and events are related and organized.
-By traversing every path, we measure the accumulation of first links, path lengths, basins, cycles, 
-and the influence each article exerts in shaping the network.
-We find scale-free distributions describe path length, 
+
+By traversing every path, we measure the accumulation of first links, path lengths,
+groups of path-connected articles, and cycles.
+We also develop a new method, traversal funnels, to measure the influence each article exerts in shaping the network. 
+Traversal funnels provides a new measure of influence for directed networks without spill-over into cycles, in contrast to traditional network centrality measures.
+Within Wikipedia's First Link Network, we find scale-free distributions describe path length, 
 accumulation, and influence. Far from dispersed, first links disproportionately accumulate 

### 2016-11-21 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
added pesudo-code and fixed centrality table

@@
 Informed by river network metrics,
-we consider the FLN as a flow, quantifying 
+we consider the FLN as a flow. We quantify
 the accumulation of first links around articles 
-and the influence an article exerts in shaping the FLN.
+and develop a new method, \textit{traversal funnels}, to measure the influence an article exerts in shaping the FLN.
 Together with cycles, in-degree, depth, and the content of the articles, 
@@
 \cite{garlaschelli2003universal}. 
+In the same vein, we use flow to measure accumulation and develop a new method for isolating influence in a directed network with cycles.
 
@@
 
-To compute traversal funnels on a directed network, begin at each node: 1.
-Traverse the edge to the next node 2.
-Increment the next node's count by one 3.
-Repeat while the next node is unvisited.
+The algorithm increments the traversal funnels count along each path, up to the start of a cycle.
 The resulting count is each node's number of traversal funnels: the number of paths the node directs up to the start of a cycle.
+Below is the pseudo-code.
+
+\begin{verbatim}
+Traversal Funnels Algorithm
+
+// 1. Determine Cycles 
+for node in Graph
+    next_node = node.link
+    visited = {node, next_node}
+    // advance to potential cycle
+    while next_node not in visited:
+        next_node = next_node.link
+        visited.add(next_node)
+    // mark cycle
+    if (node == next_node):
+        node.in_cycle = True
+
+// 2. Compute Traversal Funnels
+for node in Graph:
+    while node.in_cycle == False:
+        node.funnels += 1
+        node = node.link
+\end{verbatim}
+
+The algorithm's average runtime is proportional to the product of the number of nodes, $n$, and the average path length, $l$.
+If the average path length is fixed (independent of $n$), the runtime is $O(n)$ with $O(n + l)$ auxiliary memory for storing cycles and traversing paths.
+In the worst case where $l$ is proportional to $n$, the runtime is $O(n^2)$ with $O(n)$ auxiliary memory for storing cycles and traversing paths.
+In implementation, each iteration step in the algorithm can be parallelized since the algorithm is order agnostic, as long as each node is eventually selected.
 
@@
 In addition, the work provides a general algorithm for assessing the structure and measuring influence in a directed network (or any braided flow structure). 
-These findings also form the basis for the creation of a taxonomy where 
+Traversal funnels is a new measure of influence for directed networks without spill-over into cycles in contrast to traditional network centrality measures.
+When applied to Wikipedia, these findings also form the basis for the creation of a taxonomy where 
 every idea, event, or object sits within a hierarchy of connected notions.

### 2016-11-21 writeup/traversal_funnels_pseudo.py

@@ -1,69 +1,27 @@
 """
 Pseudo Code for Traversal Funnels
-
-Reviewer Comments
-Reviewer #1: I am not fully satisfied with the updates that the authors have provided. After taking into account the comments by Reviewer 2, many of which are valid concerns, I would like to see another revision addressing the following two points.
-
-(a) Present in an algorithm form,  the steps that the authors used to compute the "Traversal-Funnel" based influence measure instead of a descriptive discussion. Comment on the running time complexity of the same.
-(b) Please double check the eigenvector centrality values. Normally the eigenvector corresponding to the top eigenvalue is selected and the entires are all non-negative. There are some negative entries present in the values shown.
-
-Reviewer #2: I am satisfied with the modifications made and the author response.  I would suggest adding something to the introduction pointing out that you are introducing the new notion of traversal funnels, and make it a bit more prominent that this is a novel measure you have created.
 """
 
+Traversal Funnels Algorithm
 
-class Article():
-    funnels = 0
-    in_cycle = False
+// 1. Determine Cycles 
+for node in Graph
+# O(n)
+    next_node = node.link
+    visited = {node, next_node}
+    // advance to potential cycle
+    while next_node not in visited:
+        # O(path length)
+        next_node = next_node.link
+        visited.add(next_node)
+    // mark cycle
+    if (node == next_node):
+        node.in_cycle = True
 
-    title = ""
-    first_link = article
-
-    def discover_cycle(self):
-        """determines whether the current article is in a cycle"""
-        article = self.first_link
-        while article.title != self.title and article.first_link:
-            article = article.first_link
-        if article.title == self.title:
-            return True
-        return False
-
-# 1. Determine Cycles
-for article in Graph:
-    # O(n)
-    article.discover_cycle()
-    # worst case: O(n)
-
-# 2. Compute Traversal Funnels
-for article in Graph:
-    # O(n)
-    while not article.in_cycle:
-        # worst case: O(n)
-        article.funnels += 1
-        article = article.first_link
-
-# Complexity: O(2n^2) = O(n^2)
-#  and no auxilary memory
-    # For a graph with small cycles (fixed number of cycles, independent of n)
-        # O(n) (for a graph with a maximum number of cycles, say 5)
-
-# Paper Version
-    # Include Article class in appendix 
-    # Describe Graph as an iterable data structure containing articles
-# 1. Determine Cycles (detailed)        
-for article in Graph:
[…]

### 2016-11-21 writeup/2016-04wikipedia-network/correspondence/2016-11-06-compsci-reviewers/2016-11-06-response.body.tex
updated response to reviewers

@@
 
-(a) Included pseudo-code for the traversal funnels algorithm as well as a description of the runtime and auxiliary memory usage in section B.
-Traversal Funnels.
-We have also made all code used for this work available in the online appendix.
+(a) Included pseudo-code for the traversal funnels algorithm as well as a description of the runtime and auxiliary memory usage in section B (Traversal Funnels).
+We have also made all code used in this work available on the online appendix.
 

### 2016-12-07 writeup/2016-04wikipedia-network/package/wikipedia-network-revtex4-combined.tex
accepted arXiv version

@@
 articles on inventions, places, people, objects, and events are related and organized.
+
 By traversing every path, we measure the accumulation of first links, path lengths,
-groups of path-connected articles, cycles,
-and the influence each article exerts in shaping the network.
-We find scale-free distributions describe path length, 
+groups of path-connected articles, and cycles.
+We also develop a new method, traversal funnels, to measure the influence each article exerts in shaping the network. 
+Traversal funnels provides a new measure of influence for directed networks without spill-over into cycles, in contrast to traditional network centrality measures.
+Within Wikipedia's First Link Network, we find scale-free distributions describe path length, 
 accumulation, and influence. Far from dispersed, first links disproportionately accumulate 
@@
 \cite{garlaschelli2003universal}. 
+In the same vein, we use flow to measure accumulation and develop a new method for isolating influence in a directed network with cycles.
+
+\subsection{Traversal Visits}
+\label{Traversal Visits}
 
@@
 to drainage basin area in geomorphology.
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig003_traversal_funnel_algo_figure.pdf}
-  \caption{
-    \textbf{Traversal Funnel Algorithm on a simple network.}
-  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
-We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.
-For example
-  the number of traversal funnels for article $E$ is 
-  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
-  \label{fig:Traversal Funnels}
-
-\end{figure}
 
@@
 
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{fig003_traversal_funnel_algo_figure.pdf}
+  \caption{
+    \textbf{Traversal Funnel Algorithm on a simple network.}
+  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
+We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.
+For example
+  the number of traversal funnels for article $E$ is 
+  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
+  \label{fig:Traversal Funnels}
+
+\end{figure}
+\subsection{Traversal Funnels}
+\label{Traversal Funnels}
+
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{fig004_centrality_measures.pdf}
+  \caption{
+    \textbf{Centrality Measures}
+    The table above compares standard centrality measures on the sample network against traversal funnels.
+While betweenness centrality and eigenvector centrality both reward node A for its place in the cycle, traversal funnels only rewards node C for directing paths into the cycle.
+}
+  \label{fig:Centrality Measures}
+
+\end{figure}
+
 While traversal visits measure accumulation, each article's first link also 
@@
 
+The algorithm increments the traversal funnels count along each path, up to the start of a cycle.
+The resulting count is each node's number of traversal funnels: the number of paths the node directs up to the start of a cycle.
+Below is the pseudo-code.
+
+\begin{verbatim}
+Traversal Funnels Algorithm
+
+// 1. Determine Cycles 
+for node in Graph
+    next_node = node.link
+    visited = {node, next_node}
+    // advance to potential cycle
+    while next_node not in visited:
+        next_node = next_node.link
+        visited.add(next_node)
+    // mark cycle
+    if (node == next_node):
+        node.in_cycle = True
+
+// 2. Compute Traversal Funnels
+for node in Graph:
+    while node.in_cycle == False:
+        node.funnels += 1
+        node = node.link
+\end{verbatim}
+
+The algorithm's average runtime is proportional to the product of the number of nodes, $n$, and the average path length, $l$.
+If the average path length is fixed (independent of $n$), the runtime is $O(n)$ with $O(n + l)$ auxiliary memory for storing cycles and traversing paths.
+In the worst case where $l$ is proportional to $n$, the runtime is $O(n^2)$ with $O(n)$ auxiliary memory for storing cycles and traversing paths.
+In implementation, each iteration step in the algorithm can be parallelized since the algorithm is order agnostic, as long as each node is eventually selected.
+
+As a measure of influence, traversal funnels differs from traditional centrality measures such as degree centrality, betweenness centrality, and eigenvector centrality. 
+Degree centrality characterizes how many articles have a first link to a particular topic, restricting influence to a single link. 
+Betweenness centrality and eigenvalue centrality capture flow within the network more than a single link away. 
+Unlike traversal funnels however, betweenness centrality and eigenvalue centrality reward nodes that simply happen to be in a cycle. 
+We compute betweenness centrality and eigenvalue centrality on the sample network using \href{https://networkx.github.io/}{NetworkX}, a Python package for modelling complex networks (see the \href{http://compstorylab.org/share/papers/ibrahim2016a/index.html}{online appendix} for details).
+In our sample network (see Fig.~\ref{fig:Centrality Measures}), node A has betweenness centrality of 0.17 and eigenvalue centrality of 0.58, although node A is simply the recipient of paths directed by node C.
+Eigenvector centrality even assigns the same centrality to nodes A, B, and C.
+
+In contrast, traversal funnels defines a path without a predetermined end node, following the progression of links to its culmination at the start of a cycle.
+Furthermore, traversal funnels provides an interpretable metric: a node's traversal funnels is the number of paths the node directs to the start of a cycle.
+Consequently, traversal funnels allow us to isolate the paths directed by node C, of which there are 4, without any spill over effect on other nodes in its cycle: A and B have zero traversal funnels.
+
+In addition, traversal funnels extends to directed networks with more than one outward edge. In such a network, there may be multiple paths associated with a single starting node. 
+The same interpretation for traversal funnels as the number of paths a node directs to a cycle still holds. 
+In contrast to betweenness centrality, traversal funnels would not favor the shortest path over any other. Traversal funnels equally weighs all paths up to a cycle.
+
 By considering the FLN not only as collection of directly linked pairs of articles, but
-as a flow, with accumulated references, cycles, and path-connected groups, we can explore how the many articles in Wikipedia are organized and related.
+as a flow using traversal funnels, with accumulated references, cycles, and path-connected groups, we explore how the many articles in Wikipedia are organized and related.
 
 \section{Results}
-
 \subsection{Degree Distribution}
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig004_articles_ndegree.pdf}
+  \includegraphics[width=\columnwidth]{fig005_articles_ndegree.pdf}
   \caption{
@@
 \end{figure}
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig005_ndegree_loglog.pdf}
-  \caption{
-    \textbf{FLN Degree Distribution.}
-    We plot the in-degree distribution against rank for all articles with an in-degree of at least 1 on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
-    distribution using the Maximum Likelihood Estimator with xmin as the in-degree minimizing Kolmogorov-Smirnov distance.
-We find xmin to be 5 links with a corresponding power law rank exponent $\alpha \simeq 2.44$ and a size exponent $\gamma \simeq 1.41$.
-}
-  \label{fig:degree distribution}
-\end{figure}
-
 We rank all 11 million articles in the FLN by in-degree to find 
@@
 
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{fig006_ndegree_loglog.pdf}
+  \caption{
+    \textbf{FLN Degree Distribution.}
+    We plot the in-degree distribution against rank for all articles with an in-degree of at least 1 on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
+    distribution using the Maximum Likelihood Estimator with xmin as the in-degree minimizing Kolmogorov-Smirnov distance.
+We find xmin to be 5 links with a corresponding power law rank exponent $\alpha \simeq 2.44$ and a size exponent $\gamma \simeq 1.41$.
+}
+  \label{fig:degree distribution}
+\end{figure}
+
 \subsection{Depth of the FLN}
 
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{fig007_path_lengths_dist.pdf}
+  \caption{
+    \textbf{Path Length Distribution.}
+The network depth appears to have a bimodal distribution with a median of 29 first links in a path.
+The 365 cycle for ``Orthodox'' liturgics is the outlier to the right, with historical articles 
+describing nations such as ``Scotland'' and ``UK'' also lying outside the third quartile.
+More than $75\%$ of articles have path lengths between 
+$0$ and $50$ links.}
+  \label{fig:Path Length Distribution}
+\end{figure}
 How many links does a connection among articles span? 
@@
 
+Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
+This roughly corresponds to the official number of articles on Wikipedia: 
+$~4.7$ million as of November 2014---approximately half of the 11 million 
+articles in the XML dump are redirects or disambiguations, not full articles.
+The most common path length is 29, with $75\%$ of path lengths ranging between 7--30 articles.
+As a distribution, more than $75\%$ of articles have a path length below 
+$50$ first links 
+while a few temporally organized paths exceed 50 links 
+(see Fig.~\ref{fig:Path Length Distribution}).
+
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig006_path_lengths_dist.pdf}
-  \caption{
-    \textbf{Path Length Distribution.}
-The network depth appears to have a bimodal distribution with a median of 29 first links in a path.
-The 365 cycle for ``Orthodox'' liturgics is the outlier to the right, with historical articles 
-describing nations such as ``Scotland'' and ``UK'' also lying outside the third quartile.
-More than $75\%$ of articles have path lengths between 
-$0$ and $50$ links.}
-  \label{fig:Path Length Distribution}
-\end{figure}
-
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig007_traversals_per_article.pdf} 
+  \includegraphics[width=\columnwidth]{fig008_traversals_per_article.pdf} 
   \caption{
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig008_articles_ranked.pdf}
+  \includegraphics[width=\columnwidth]{fig009_articles_ranked.pdf}
   \caption{
@@
 \begin{figure*}[tph!]
-  \includegraphics[width=\textwidth]{fig009_2_and_3_cycles.pdf}
+  \includegraphics[width=\textwidth]{fig010_2_and_3_cycles.pdf}
   \caption{
@@
 \label{cycles}
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig010_top_funnels.pdf}
-  \caption{
-    \textbf{Funnels.}
-We represent the highest-ranking articles by the number of traversal 
-funnels to gauge the influence each article exerts in shaping the 
-structure of the FLN. 
-We find ``Philosophy'' exerts an overwhelming proportion
-of the influence, with other abstract notions and topical concepts ranking
-next.}
-  \label{fig:Funnels}
-\end{figure}
 
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig011_funnels_distribution.pdf}
+  \includegraphics[width=\columnwidth]{fig011_top_funnels.pdf}
+  \caption{
+    \textbf{Funnels.}
+We represent the highest-ranking articles by the number of traversal 
+funnels to gauge the influence each article exerts in shaping the 
+structure of the FLN. 
+We find ``Philosophy'' exerts an overwhelming proportion
+of the influence, with other abstract notions and topical concepts ranking
+next.}
+  \label{fig:Funnels}
+\end{figure}
+
+\subsection{Traversal Funnels}
+\label{Traversal Funnels}
+
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{fig012_funnels_distribution.pdf}
   \caption{
@@
 
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig012_popularity_visits.pdf}
-  \caption{
-\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal visits.}
-We use the total number of page views provided by Wikipedia for the month
-of October 2015 to compare each article's popularity to traversal visits.
-While the most popular articles do not necessarily correspond to the articles
-with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases.
-The greatest number of articles fall within roughly 1 million traversal visits and page views.
-Overall, the page views for the top 1000 articles articles by traversal visits appear to be log-normally distributed.
-}
-  \label{fig:Views and Visits}
-
-\end{figure}
 As a distribution, we find few articles influence the structure of the 
@@
 
+\subsection{Article Popularity}
+
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig013_popularity_funnels.pdf}
+  \includegraphics[width=\columnwidth]{fig013_popularity_visits.pdf}
+  \caption{
+\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal visits.}
+We use the total number of page views provided by Wikipedia for the month
+of October 2015 to compare each article's popularity against traversal visits.
+Overall, more page views do not correspond to higher traversal visits---Wikipedia users do not tend to visit articles with a higher accumulation in the FLN more so than they visit others. 
+However, the variation in popularity appears to decrease as the number of traversal visits increases.
+The greatest number of articles fall within roughly 1 million traversal visits and page views.
+Furthermore, page views for the top 1000 articles articles by traversal visits appear to be log-normally distributed.
+}
+  \label{fig:Views and Visits}
+
+\end{figure}
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{fig014_popularity_funnels.pdf}
   \caption{
@@
 
+Overall, the contrast between page views and traversal visits or funnels highlights a difference between popularity and influence in the FLN.
+Although traversal funnels and visits point to influential articles where many topics culminate, Wikipedia users do not visit these articles more than others.
+What is popular, is not necessarily foundational, and vice versa.
+
 %========================Conclusion====================================
@@
 \begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{fig014_flags.pdf}  
+  \includegraphics[width=\columnwidth]{fig015_flags.pdf}  
   \caption{
@@
 
+\bibitem{xefer2011philosophy}
+.~Xefer, May.
+\newblock All road lead to philosophy.
+\newblock [Online; accessed 22-April-2016].
+
 \bibitem{Bolton2007}

### 2016-12-07 writeup/2016-04wikipedia-network/wikipedia-network-revtex4-combined.tex
accepted arXiv version

@@
 
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{traversal_funnel_algo_figure.pdf}
+  \caption{
+    \textbf{Traversal Funnel Algorithm on a simple network.}
+  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path.
+We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row.
+For example
+  the number of traversal funnels for article $E$ is 
+  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
+  \label{fig:Traversal Funnels}
+
+\end{figure}
+\subsection{Traversal Funnels}
+\label{Traversal Funnels}
+
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{centrality_measures.pdf}
+  \caption{
+    \textbf{Centrality Measures}
+    The table above compares standard centrality measures on the sample network against traversal funnels.
+While betweenness centrality and eigenvector centrality both reward node A for its place in the cycle, traversal funnels only rewards node C for directing paths into the cycle.
+}
+  \label{fig:Centrality Measures}
+
+\end{figure}
+
 While traversal visits measure accumulation, each article's first link also 
@@
 
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{ndegree_loglog.pdf}
+  \caption{
+    \textbf{FLN Degree Distribution.}
+    We plot the in-degree distribution against rank for all articles with an in-degree of at least 1 on a $\log_{10}-\log_{10}$. We then fit the data to a power law 
+    distribution using the Maximum Likelihood Estimator with xmin as the in-degree minimizing Kolmogorov-Smirnov distance.
+We find xmin to be 5 links with a corresponding power law rank exponent $\alpha \simeq 2.44$ and a size exponent $\gamma \simeq 1.41$.
+}
+  \label{fig:degree distribution}
+\end{figure}
+
 \subsection{Depth of the FLN}
 
+\begin{figure}[tp!]
+  \includegraphics[width=\columnwidth]{path_lengths_dist.pdf}
+  \caption{
+    \textbf{Path Length Distribution.}
+The network depth appears to have a bimodal distribution with a median of 29 first links in a path.
+The 365 cycle for ``Orthodox'' liturgics is the outlier to the right, with historical articles 
+describing nations such as ``Scotland'' and ``UK'' also lying outside the third quartile.
+More than $75\%$ of articles have path lengths between 
+$0$ and $50$ links.}
+  \label{fig:Path Length Distribution}
+\end{figure}
 How many links does a connection among articles span? 
@@
 
-\begin{figure}[tp!]
-  \includegraphics[width=\columnwidth]{path_lengths_dist.pdf}
-  \caption{
-    \textbf{Path Length Distribution.}
-The network depth appears to have a bimodal distribution with a median of 29 first links in a path.
-The 365 cycle for ``Orthodox'' liturgics is the outlier to the right, with historical articles 
-describing nations such as ``Scotland'' and ``UK'' also lying outside the third quartile.
-More than $75\%$ of articles have path lengths between 
-$0$ and $50$ links.}
-  \label{fig:Path Length Distribution}
-\end{figure}
+Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
+This roughly corresponds to the official number of articles on Wikipedia: 
+$~4.7$ million as of November 2014---approximately half of the 11 million 
+articles in the XML dump are redirects or disambiguations, not full articles.
+The most common path length is 29, with $75\%$ of path lengths ranging between 7--30 articles.
+As a distribution, more than $75\%$ of articles have a path length below 
+$50$ first links 
+while a few temporally organized paths exceed 50 links 
+(see Fig.~\ref{fig:Path Length Distribution}).
 
